### 2025-02-09 notes.tex
Blank doc, Makefile.

\documentclass{acmart}

\begin{document}

\title{Approximation as Differentiation notes}
\maketitle

\end{document}

### 2025-02-09 notes.tex
Remove ACM ref, build twice to remove temp page.
Start on category of families.

@@
 \documentclass{acmart}
+
+\input{macros}
+
+\settopmatter{printacmref=false}
 
@@
 
+\section{Overview}
+
+Some notes on the key constructions:
+\begin{itemize}
+\item Categories of families, special case of Grothendieck construction
+\end{itemize}
+
+\section{Definitions}
+
+\subsection{Category of families}
+
+Special case of Grothendieck construction for a functor $F: C \to \Cat$, where we consider the special case
+where $F: C \to \Set$. (This is a ``special case'' inasmuch as there is an embedding from $\Set$ to $\Cat$
+sending each set to the corresponding discrete category.)
+
 \end{document}

### 2025-02-09 notes.tex
Start on category of families.
Clarify how a functor F: C \to Set is also a functor F: C \to Cat.

--- 2025-02-09 notes.tex
@@
 
+\section{Overview}
+
+Some notes on the key constructions:
+\begin{itemize}
+\item Categories of families, special case of Grothendieck construction
+\end{itemize}
+
+\section{Definitions}
+
+\subsection{Category of families}
+
+This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C \to
+\Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete categories.)
+
 \end{document}

### 2025-02-09 notes.tex
Category of families.

@@
 
-This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C \to
-\Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete categories.)
+For a functor $F: C \to Set$, we have the category where:
+\begin{itemize}
+\item objects are pairs $(A, x)$ where $x \in FA$
+\item morphisms from $(A, x)$ to $(B, y)$ are morphisms $f: A \to B$ in $C$ where $(Ff)(x) = y$
+\end{itemize}
+
+\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C
+\to \Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
+categories.
 

### 2025-02-09 notes.tex
Define Fam.

--- 2025-02-09 notes.tex
@@
 
+$\Set$ will often be $\Setoid$ in the Agda code but we'll ignore that here for now.
+
 \section{Definitions}
+
+\subsection{Indexed families of objects}
+
+For $I$ a set and $C$ a category, write $\Fam(I,C)$ for the set of $I$-indexed families of objects of $C$.
+
+Suppose $C$ a category and $I, J$ sets and $x$ an $I$-indexed family of objects of $C$ and $y$ a $J$-indexed
+family of objects of $C$. Define $y_f$ for any $f: I \to J$ to be the $I$-indexed family of objects of $C$.
 
@@
 
-This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C \to
-\Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete categories.)
+For a functor $F: C \to Set$, we have the category where:
+\begin{itemize}
+\item objects are pairs $(I, x)$ with $I$ an object of $C$ and $x \in FI$
+\item morphisms from $(I, x)$ to $(J, y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(x) = y$
+\end{itemize}
+
+\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C
+\to \Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
+categories.)
+
+In our code we give a slightly different construction. For a category $C$, define the category where:
+\begin{itemize}
+\item objects are pairs $(I, x)$ with $I$ a set and $x$ an $I$-indexed family of objects of $C$
+\item morphisms from $(I, x)$ to $(J, y)$ are functions $f: I \to J$ paired with morphisms
+\end{itemize}
 

### 2025-02-09 notes.tex
Fam(I,C) is a functor category where the naturality conditions are trivial because I is discrete.

@@
 
-$\Set$ will often be $\Setoid$ in the Agda code but we'll ignore that here for now.
+\noindent $\Set$ will often be $\Setoid$ in the Agda code but I'm glossing that here.
 
@@
 
-For $I$ a set and $C$ a category, write $\Fam(I,C)$ for the set of $I$-indexed families of objects of $C$.
+For any set $I$ and category $C$ write $\Fam(I,C)$ for the category where:
+\begin{itemize}
+\item the objects are the $I$-indexed families of objects of $C$;
+\item a morphism from $X$ to $Y$ is a family of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
+\end{itemize}
 
-Suppose $C$ a category and $I, J$ sets and $x$ an $I$-indexed family of objects of $C$ and $y$ a $J$-indexed
-family of objects of $C$. Define $y_f$ for any $f: I \to J$ to be the $I$-indexed family of objects of $C$.
+\noindent Equivalently, $\Fam(I,C)$ is the functor category $\Functors{I}{C}$.
+
+\noindent Suppose $x \in \Fam(I,C)$ and $y \in \Fam(J,C)$. Define $y_f$ for any $f: I \to J$ to be the
+$I$-indexed family of objects of $C$.
 

### 2025-02-09 notes.tex
Fam(I,C) is a functor category where the naturality conditions are trivial because I is discrete.
Reindexing.
Reindexing functor.

@@
 
-\noindent Suppose $x \in \Fam(I,C)$ and $y \in \Fam(J,C)$. Define $y_f$ for any $f: I \to J$ to be the
-$I$-indexed family of objects of $C$.
+\begin{definition}[Reindexing functor]
+For any $X \in \Fam(J,C)$ and $f: I \to J$ define $X[f] \in \Fam(I,C)$ where $X[f]_i = X_{fi}$.
+\end{definition}
+
+[$\Fam(-,C)$ is a fibred category over $\Set$?]
 
@@
 \begin{itemize}
-\item objects are pairs $(I, x)$ with $I$ a set and $x$ an $I$-indexed family of objects of $C$
-\item morphisms from $(I, x)$ to $(J, y)$ are functions $f: I \to J$ paired with morphisms
+\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$
+\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with morphisms $X \to Y[f]$ in
+$\Fam(I,C)$.
 \end{itemize}

### 2025-02-09 notes.tex
Ignore .pdf.

@@
 
-[$\Fam(-,C)$ is a fibred category over $\Set$?]
-
 \subsection{Category of families}
@@
 \begin{itemize}
-\item objects are pairs $(I, x)$ with $I$ an object of $C$ and $x \in FI$
-\item morphisms from $(I, x)$ to $(J, y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(x) = y$
+\item objects are pairs $(I, X)$ with $I$ an object of $C$ and $X \in FI$ an $I$-indexed family;
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(X) = Y$.
 \end{itemize}
@@
 \noindent This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C
-\to \Set$. (Where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
-categories.)
+\to \Set$ (where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
+categories).
 
-In our code we give a slightly different construction. For a category $C$, define the category where:
+In our code we give a slightly different construction, namely the fibred category $\Fam(C)$ over $\Set$ where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$
+\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
 \item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with morphisms $X \to Y[f]$ in

### 2025-02-10 notes.tex
nonacm.

@@
-\documentclass{acmart}
+\documentclass[nonacm]{acmart}
 
@@
 
-For a functor $F: C \to Set$, we have the category where:
+For a functor $F: C \to \Set$, we have the category where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ with $I$ an object of $C$ and $X \in FI$ an $I$-indexed family;
+\item objects are pairs $(I, X)$ of an object $I$ of $C$ and an $I$-indexed family $X \in FI$;
 \item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(X) = Y$.

### 2025-02-10 notes.tex
Functors -> FunCat.
LatGal.

@@
 \begin{itemize}
-\item Categories of families, special case of Grothendieck construction
+\item Category of families $\Fam$
+\item Category of bounded lattices and Galois connections
 \end{itemize}
@@
 
-\noindent Equivalently, $\Fam(I,C)$ is the functor category $\Functors{I}{C}$.
+\noindent Equivalently, $\Fam(I,C)$ is the functor category $\FunCat{I}{C}$.
 
@@
 
+\noindent \todo{Some properties of $\Fam(C)$.}
+
+\subsection{Category of bounded lattices and Galois connections}
+
+Define the category $\LatGal$ where:
+\begin{itemize}
+\item objects $L$ are bounded lattices;
+\item morphisms $(f, g)$ from $L$ to $L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
+join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq fx \iff gy \leq x$.
+\end{itemize}
+
+\noindent \note{The Agda defines an object of $\LatGal$ as a set equipped with meet- and join-semilattice
+structures but there is no requirement that they satisfy the absorption laws or induce the same partial
+order.}
+
 \end{document}

### 2025-02-10 notes.tex
The two semilattice structures share a partial order.
Ignoring lists for now.

@@
 \begin{itemize}
-\item objects $L$ are bounded lattices;
-\item morphisms $(f, g)$ from $L$ to $L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq fx \iff gy \leq x$.
+\item objects $L$ are the bounded lattices;
+\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
+join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq fx \iff gy \leq x$ for any $x \in L$ and any
+$y \in L'$.
 \end{itemize}
 
-\noindent \note{The Agda defines an object of $\LatGal$ as a set equipped with meet- and join-semilattice
-structures but there is no requirement that they satisfy the absorption laws or induce the same partial
-order.}
+\noindent $\LatGal$ has finite biproducts. \todo{Tie this to pre-additive structure.}
+
+\subsection{Semantic domain}
+
+We interpret our language into any category with:
+\begin{itemize}
+\item finite products (binary products and a terminal object)
+\end{itemize}
+\note{Ignoring lists for now.}
 

### 2025-02-10 notes.tex
Signatures.
acmsmall.

@@
-\documentclass[nonacm]{acmart}
+\documentclass[acmsmall,nonacm]{acmart}
+
+%\geometry{paperwidth=8.3in, paperheight=11.7in}
 
@@
 
-Some notes on the key constructions:
+Some of the key constructions:
 \begin{itemize}
 \item Category of families $\Fam$
-\item Category of bounded lattices and Galois connections
+\item Category of $\LatGal$ of bounded lattices and Galois connections
 \end{itemize}
 
-\noindent $\Set$ will often be $\Setoid$ in the Agda code but I'm glossing that here.
+\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but will gloss that here.
 
@@
 
-\noindent Equivalently, $\Fam(I,C)$ is the functor category $\FunCat{I}{C}$.
+\noindent Equivalently $\Fam(I,C)$ is the functor category $\FunCat{I}{C}$ where we interpret $I$ as a
+discrete category.
 
@@
 
+\section{Language}
+
+\subsection{Syntax}
+
+We parameterise our language over a signature $\Sigma = (S, \phi, R)$ consisting of:
+\begin{itemize}
+\item a set $S$ of sorts
+\item a set of operations $\phi$ with signature $\Sigma(\phi) \in S^* \times S$
+\item a set of relations $R$ with signature $\Sigma(R) \in S^*$
+\end{itemize}
+
 \subsection{Semantic domain}
@@
 \item finite products (binary products and a terminal object)
+\item Booleans
 \end{itemize}
-\note{Ignoring lists for now.}
+\note{Ignoring lists for now; Booleans can be derived from coproducts.}
 

### 2025-02-10 notes.tex
Brackets for function application.

@@
 \documentclass[acmsmall,nonacm]{acmart}
-
-%\geometry{paperwidth=8.3in, paperheight=11.7in}
 
@@
 \begin{definition}[Reindexing functor]
-For any $X \in \Fam(J,C)$ and $f: I \to J$ define $X[f] \in \Fam(I,C)$ where $X[f]_i = X_{fi}$.
+For any $X \in \Fam(J,C)$ and $f: I \to J$ define $X[f] \in \Fam(I,C)$ where $X[f]_i = X_{f(i)}$.
 \end{definition}
@@
 \item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq fx \iff gy \leq x$ for any $x \in L$ and any
+join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
 $y \in L'$.

### 2025-02-10 notes.tex
Force to A4.
Ok, think I see how Fam(C) is a Grothendieck construction.

@@
 
+\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
+\settopmatter{printacmref=false}
+
 \input{macros}
-
-\settopmatter{printacmref=false}
 
@@
 
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\FunCat{I}{C}$ where we interpret $I$ as a
-discrete category.
+\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
+category.
 
@@
 
-\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C
+\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$ for the special case where $F: C
 \to \Set$ (where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
@@
 
-In our code we give a slightly different construction, namely the fibred category $\Fam(C)$ over $\Set$ where:
+Define the fibred category $\Fam(C)$ over $\Set$ where:
 \begin{itemize}
@@
 \end{itemize}
+
+\noindent This is the Grothendieck construction for the functor $\Set^{\op} \to \Cat$ which sends any set $I$
+to the category $\Fam(I,C)$ and any function $f: I \to J$ to the reindexing functor $-[f]: \Fam(J,C) \to
+\Fam(I,C)$.
 
@@
 \end{itemize}
-\note{Ignoring lists for now; Booleans can be derived from coproducts.}
+\note{Ignoring lists for now. Booleans can be derived from coproducts.}
 

### 2025-02-10 notes.tex
Ok, think I see how Fam(C) is a Grothendieck construction.
Clarify reindexing as contravariant functor.
Split out into files.

--- 2025-02-10 notes.tex
@@
 
-\section{Definitions}
-
-\subsection{Indexed families of objects}
-
-For any set $I$ and category $C$ write $\Fam(I,C)$ for the category where:
-\begin{itemize}
-\item the objects are the $I$-indexed families of objects of $C$;
-\item a morphism from $X$ to $Y$ is a family of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
-\end{itemize}
-
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\FunCat{I}{C}$ where we interpret $I$ as a
-discrete category.
-
-\begin{definition}[Reindexing functor]
-For any $X \in \Fam(J,C)$ and $f: I \to J$ define $X[f] \in \Fam(I,C)$ where $X[f]_i = X_{f(i)}$.
-\end{definition}
-
-\subsection{Category of families}
-
-For a functor $F: C \to \Set$, we have the category where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $C$ and an $I$-indexed family $X \in FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(X) = Y$.
-\end{itemize}
-
-\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$, in the special case where $F: C
-\to \Set$ (where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
-categories).
-
-In our code we give a slightly different construction, namely the fibred category $\Fam(C)$ over $\Set$ where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with morphisms $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
-
-\noindent \todo{Some properties of $\Fam(C)$.}
-
-\subsection{Category of bounded lattices and Galois connections}
-
-Define the category $\LatGal$ where:
-\begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
-\end{itemize}
-
-\noindent $\LatGal$ has finite biproducts. \todo{Tie this to pre-additive structure.}
-
-\section{Language}
-
-\subsection{Syntax}
-
-We parameterise our language over a signature $\Sigma = (S, \phi, R)$ consisting of:
-\begin{itemize}
-\item a set $S$ of sorts
-\item a set of operations $\phi$ with signature $\Sigma(\phi) \in S^* \times S$
-\item a set of relations $R$ with signature $\Sigma(R) \in S^*$
-\end{itemize}
-
-\subsection{Semantic domain}
-
-We interpret our language into any category with:
-\begin{itemize}
-\item finite products (binary products and a terminal object)
-\item Booleans
-\end{itemize}
-\note{Ignoring lists for now; Booleans can be derived from coproducts.}
+\input{notes/definitions}
+\input{notes/language}
 

### 2025-02-10 notes/definitions.tex
Split out into files.

\section{Definitions}

\subsection{Indexed families of objects}

For any set $I$ and category $C$ write $\Fam(I,C)$ for the category where:
\begin{itemize}
\item the objects are the $I$-indexed families of objects of $C$;
\item a morphism from $X$ to $Y$ is a family of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
\end{itemize}

\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
category.

\begin{definition}[Reindexing]
For any $f: I \to J$ define the following contravariant \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C)
\to \Fam(I,C)$ that sends any $X \in \Fam(J,C)$ to the $I$-indexed family $X[f]$ which has $X[f]_i = X_{f(i)}$
for any $i \in I$.
\end{definition}

\subsection{Category of families}

For a functor $F: C \to \Set$, we have the category where:
\begin{itemize}
\item objects are pairs $(I, X)$ of an object $I$ of $C$ and an $I$-indexed family $X \in FI$;
\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(X) = Y$.
\end{itemize}

\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$ for the special case where $F: C
\to \Set$ (where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
categories).

Define the fibred category $\Fam(C)$ over $\Set$ where:
\begin{itemize}
\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with morphisms $X \to Y[f]$ in
$\Fam(I,C)$.
\end{itemize}

\noindent This is the Grothendieck construction for the functor $\Set^{\op} \to \Cat$ which sends any set $I$
to the category $\Fam(I,C)$ and any function $f: I \to J$ to the reindexing functor $-[f]: \Fam(J,C) \to
\Fam(I,C)$.

\todo{Some properties of $\Fam(C)$.}

\subsection{Category of bounded lattices and Galois connections}

Define the category $\LatGal$ where:
\begin{itemize}
\item objects $L$ are the bounded lattices;
\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
$y \in L'$.
\end{itemize}

\noindent $\LatGal$ has finite biproducts. \todo{Tie this to pre-additive structure.}

### 2025-02-10 notes/definitions.tex
Two different functors..

@@
 \noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category.
+category. $\Fam(-,C): \Set^{\op} \to \Cat$ is a functor which sends any set $I$ to the category $\Fam(I,C)$
+and any function $f: I \to J$ to the \emph{reindexing functor} $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
 
 \begin{definition}[Reindexing]
-For any $f: I \to J$ define the following contravariant \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C)
-\to \Fam(I,C)$ that sends any $X \in \Fam(J,C)$ to the $I$-indexed family $X[f]$ which has $X[f]_i = X_{f(i)}$
-for any $i \in I$.
+For any $f: I \to J$ the \emph{reindexing} functor sends any $J$-indexed family $X$ to the $I$-indexed family
+$X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$.
 \end{definition}

### 2025-02-10 notes/definitions.tex
Two different functors..

@@
 
-For a functor $F: C \to \Set$, we have the category where:
+For a functor $F: C \to \Cat$, we have the category where:
 \begin{itemize}
@@
 
-\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$ for the special case where $F: C
-\to \Set$ (where we read a $\Set$-valued functor as a $\Cat$-valued functor restricted to discrete
-categories).
+\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$.
 
@@
 
-\noindent This is the Grothendieck construction for the functor $\Set^{\op} \to \Cat$ which sends any set $I$
-to the category $\Fam(I,C)$ and any function $f: I \to J$ to the reindexing functor $-[f]: \Fam(J,C) \to
-\Fam(I,C)$.
+\noindent This is the Grothendieck construction for the functor $\Fam(-,C)$.
 

### 2025-02-10 notes/definitions.tex
Some clarity on Fam(C) and relationship to Grothendieck construction.

@@
 
-For a functor $F: C \to \Cat$, we have the category where:
+For a functor $F: D \to \Cat$, we have the Grothendieck construction for $F$ which is the category where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $C$ and an $I$-indexed family $X \in FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $C$ where $(Ff)(X) = Y$.
+\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an $I$-indexed family $X \in FI$;
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ where $(Ff)(X) = Y$.
 \end{itemize}
 
-\noindent This is the Grothendieck construction for a functor $F: C \to \Cat$.
-
-Define the fibred category $\Fam(C)$ over $\Set$ where:
+\noindent The fibred category $\Fam(C)$ over $\Set$ is the Grothendieck construction for the functor
+$\Fam(-,C)$:
 \begin{itemize}
@@
 \end{itemize}
-
-\noindent This is the Grothendieck construction for the functor $\Fam(-,C)$.
 

### 2025-02-10 notes/definitions.tex
Some progress.

@@
 \noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category. $\Fam(-,C): \Set^{\op} \to \Cat$ is a functor which sends any set $I$ to the category $\Fam(I,C)$
-and any function $f: I \to J$ to the \emph{reindexing functor} $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
+category. $\Fam(-,C): \Set^{\op} \to \Cat$ is a functor which sends any set $I$ to $\Fam(I,C)$ and any
+function $f: I \to J$ to the \emph{reindexing functor} $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
 
@@
 For any $f: I \to J$ the \emph{reindexing} functor sends any $J$-indexed family $X$ to the $I$-indexed family
-$X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$.
+$X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and similarly for morphisms.
 \end{definition}
@@
 
-For a functor $F: D \to \Cat$, we have the Grothendieck construction for $F$ which is the category where:
+For a functor $F: D \to \Cat$, the Grothendieck construction for $F$ is the category where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an $I$-indexed family $X \in FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ where $(Ff)(X) = Y$.
+\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object of $FI$;
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)(X) \to Y$.
 \end{itemize}
@@
 \noindent The fibred category $\Fam(C)$ over $\Set$ is the Grothendieck construction for the functor
-$\Fam(-,C)$:
+$\Fam(-,C): \Set^{\op} \to \Cat$:
 \begin{itemize}

### 2025-02-19 notes/definitions.tex
Tweak.

@@
 
-For any set $I$ and category $C$ write $\Fam(I,C)$ for the category where:
+For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
 \begin{itemize}
-\item the objects are the $I$-indexed families of objects of $C$;
-\item a morphism from $X$ to $Y$ is a family of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
+\item objects are the $I$-indexed families of objects of $C$;
+\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
 \end{itemize}
@@
 \noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category. $\Fam(-,C): \Set^{\op} \to \Cat$ is a functor which sends any set $I$ to $\Fam(I,C)$ and any
-function $f: I \to J$ to the \emph{reindexing functor} $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
+category. $\Fam(-,C): \Set^{\op} \to \Cat$ is the functor which sends any set $I$ to $\Fam(I,C)$ and any
+function $f: I \to J$ to the following \emph{reindexing functor}.
 
 \begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor sends any $J$-indexed family $X$ to the $I$-indexed family
-$X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and similarly for morphisms.
+For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
+$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and
+similarly for morphisms.
 \end{definition}
@@
 
-For a functor $F: D \to \Cat$, the Grothendieck construction for $F$ is the category where:
+The Grothendieck construction for a functor $F: D \to \Cat$ is the category where:
 \begin{itemize}
 \item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object of $FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)(X) \to Y$.
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)(X) \to Y$ in $FJ$.
 \end{itemize}
@@
 \item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with morphisms $X \to Y[f]$ in
+\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
 $\Fam(I,C)$.

### 2025-02-19 notes/definitions.tex
Some attempt to make sense of Grothendieck construction when F is contravariant.

@@
 \noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category. $\Fam(-,C): \Set^{\op} \to \Cat$ is the functor which sends any set $I$ to $\Fam(I,C)$ and any
+category. $\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any
 function $f: I \to J$ to the following \emph{reindexing functor}.
@@
 
-The Grothendieck construction for a functor $F: D \to \Cat$ is the category where:
+The \emph{Grothendieck construction} $\Grothendieck{D}F$ for a functor $F: D \to \Cat$ is the category where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object of $FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)(X) \to Y$ in $FJ$.
+\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $FI$;
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)X
+\to Y$ in $FJ$.
 \end{itemize}
 
-\noindent The fibred category $\Fam(C)$ over $\Set$ is the Grothendieck construction for the functor
-$\Fam(-,C): \Set^{\op} \to \Cat$:
+\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
+paired with a morphism $X \to (Ff)Y$ in $FI$. Writing $\Fam(C)$ for the Grothendieck construction
+$\Grothendieck{\Set^{\op}}\Fam(-,C)$ we then have that in $\Fam(C)$:
 \begin{itemize}

### 2025-02-19 notes/definitions.tex
Correct direction of f.

@@
 \begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
+\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
+\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: J \to I$ paired with a morphism $X \to Y[f]$ in
 $\Fam(I,C)$.

### 2025-02-21 notes/definitions.tex
Nope, original direction was right.
CMon.

--- 2025-02-19 notes/definitions.tex
@@
 \begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and indexed family $X$ in $\Fam(I,C)$;
+\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
 \item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
@@
 \noindent $\LatGal$ has finite biproducts. \todo{Tie this to pre-additive structure.}
+
+\subsection{Category of commutative monoids}
+
+$\CMon$ is the category whose objects are the commutative monoids $M$ and morphisms $M \to M'$ are the
+commutative monoid homomorphisms, with the trivial one-element monoid $\One$ as terminal object and binary
+products $M \tensor M'$ given by the direct product of monoids.
+
+\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
+
+A presheaf $F: \LatGal^\op \to \CMon$ assigns to every
+bounded lattice $L$ a commutative monoid $FL$ and to every Galois connection $f: L \to L'$ .
+$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-21 notes/definitions.tex
$\CMon$ is complete and cocomplete.

@@
 $\CMon$ is the category whose objects are the commutative monoids $M$ and morphisms $M \to M'$ are the
-commutative monoid homomorphisms, with the trivial one-element monoid $\One$ as terminal object and binary
-products $M \tensor M'$ given by the direct product of monoids.
+commutative monoid homomorphisms. $\CMon$ is symmetric closed monoidal, with the trivial one-element monoid
+$\One$ as terminal object and tensor product $M \tensor M'$ given by the direct product of $M$ and $M'$. In
+fact $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
 

### 2025-02-21 notes/definitions.tex
Biproducts.
Say what zero morphism and addition of morphisms is.
CMon-enriched category.

@@
 
+\subsection{$\CMon$-enriched category}
+
+A category is $\CMon$-enriched if $\Hom{A}{B}$ is closed under the \emph{zero} morphism $\zero_{A,B}$ and
+addition of morphisms $+_{A,B}$ satisfying the commutative monoid laws.
+
 \subsection{Category $\Func{\LatGal^\op}{\CMon}$}
@@
 $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.
+
+\subsection{Biproducts}
+
+Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
+object $A \oplus B$ of $C$ together with morphisms
+
+\begin{center}
+\begin{tikzcd}
+   A \arrow[r, "i_A", shift left] &
+   A \oplus B \arrow[l, "p_A", shift left] \arrow[r, "p_B"', shift right] &
+   B \arrow[l, "i_B"', shift right]
+\end{tikzcd}
+\end{center}
+
+\noindent satisfying
+
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   p_A \after i_A &= \id_A \\
+   p_B \after i_A &= \zero_{A,B}
+\end{salign*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   p_B \after i_B &= \id_B \\
+   p_A \after i_B &= \zero_{B,A}
+\end{salign*}
+\end{center}
+\end{minipage}
+
+\noindent and
+
+\begin{salign*}
+i_A \after p_A + i_B \after p_B &= \id_{A \oplus B}
+\end{salign*}

### 2025-02-21 notes/definitions.tex
CMon-enriched category.

@@
 
-A category is $\CMon$-enriched if $\Hom{A}{B}$ is closed under the \emph{zero} morphism $\zero_{A,B}$ and
-addition of morphisms $+_{A,B}$ satisfying the commutative monoid laws.
+A category is $\CMon$-enriched if every $\Hom{A}{B}$ is a commutative monoid with the \emph{zero} morphism
+$\zero_{A,B}$ as the unit and addition of morphisms $+_{A,B}$ as the commutative monoid operation.
 

### 2025-02-21 notes.tex
Two different functors..
Some progress.
Some attempt to make sense of Grothendieck construction when F is contravariant.
Setoid.
CMon.
Biproducts.
Some references.
Products (or coproducts) are enough to derive biproducts in a CMon-enriched category.

@@
 \documentclass[acmsmall,nonacm]{acmart}
+
+\usepackage{bbm}
+\usepackage{tikz-cd}
 
@@
 \begin{itemize}
-\item Category of families $\Fam$
+\item Category of families $\Fam(C)$ for any category $C$
 \item Category of $\LatGal$ of bounded lattices and Galois connections
@@
 
-\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but will gloss that here.
+Is one of the goals to define sufficient conditions on $C$ for $\Fam(C)$ to be Cartesian closed?
+
+\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but I'm glossing that detail here.
 
 \input{notes/definitions}
-\input{notes/language}
+% \input{notes/language}
+
+\section{Some references}
+
+\cite{karvonen2020}
+
+\bibliographystyle{plainnat}
+\bibliography{bib}
 

### 2025-02-21 notes/definitions.tex
Biproducts.
Products (or coproducts) are enough to derive biproducts in a CMon-enriched category.

--- 2025-02-21 notes/definitions.tex
@@
 
-A category is $\CMon$-enriched if $\Hom{A}{B}$ is closed under the \emph{zero} morphism $\zero_{A,B}$ and
-addition of morphisms $+_{A,B}$ satisfying the commutative monoid laws.
+A category is $\CMon$-enriched if every $\Hom{A}{B}$ is a commutative monoid with the \emph{zero} morphism
+$\zero_{A,B}$ and addition of morphisms $+_{A,B}$ as the commutative monoid operations, and composition
+is a commutative monoid homomorphism.
 
-\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
+\begin{proposition}
+Suppose $C$ a $\CMon$-enriched monoidal category.
+\begin{itemize}
+\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
+\item If $C$ has a terminal (or initial) object then it is a zero object.
+\end{itemize}
+\end{proposition}
 
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every
-bounded lattice $L$ a commutative monoid $FL$ and to every Galois connection $f: L \to L'$ .
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.
+If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
 
@@
 \end{salign*}
+
+Any category with biproducts has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A \oplus
+B, i_A, i_B)$.
+
+\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
+
+A presheaf $F: \LatGal^\op \to \CMon$ assigns to every
+bounded lattice $L$ a commutative monoid $FL$ and to every Galois connection $f: L \to L'$ .
+$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-21 notes/definitions.tex
Bilinearity.

@@
 commutative monoid homomorphisms. $\CMon$ is symmetric closed monoidal, with the trivial one-element monoid
-$\One$ as terminal object and tensor product $M \tensor M'$ given by the direct product of $M$ and $M'$. In
-fact $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
+$\One$ as terminal object and monoidal product $M \tensor M'$ given by the direct product of $M$ and $M'$.
+$\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
 
@@
 
-A category is $\CMon$-enriched if every $\Hom{A}{B}$ is a commutative monoid with the \emph{zero} morphism
-$\zero_{A,B}$ and addition of morphisms $+_{A,B}$ as the commutative monoid operations, and composition
-is a commutative monoid homomorphism.
+A category $C$ is \emph{$\CMon$-enriched} if every hom-object $\Hom{C}{A}{B}$ is a commutative monoid with a
+\emph{zero} morphism $\zero_{A,B}$ and addition of morphisms $+_{A,B}$ as the commutative monoid operations,
+and composition is a morphism $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
+i.e.~satisfies
+
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}
 
@@
 \begin{salign*}
-   p_A \after i_A &= \id_A \\
-   p_B \after i_A &= \zero_{A,B}
+   p_A \comp i_A &= \id_A \\
+   p_B \comp i_A &= \zero_{A,B}
 \end{salign*}
@@
 \begin{salign*}
-   p_B \after i_B &= \id_B \\
-   p_A \after i_B &= \zero_{B,A}
+   p_B \comp i_B &= \id_B \\
+   p_A \comp i_B &= \zero_{B,A}
 \end{salign*}
@@
 \begin{salign*}
-i_A \after p_A + i_B \after p_B &= \id_{A \oplus B}
+i_A \comp p_A + i_B \comp p_B &= \id_{A \oplus B}
 \end{salign*}

### 2025-02-21 notes/definitions.tex
Bilinearity.

@@
 \emph{zero} morphism $\zero_{A,B}$ and addition of morphisms $+_{A,B}$ as the commutative monoid operations,
-and composition is a morphism $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
+and composition is a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
 i.e.~satisfies
 
+\begin{salign*}
+f \comp \zero = f = \zero \comp f
+\end{salign*}
 \begin{salign*}

### 2025-02-21 notes/definitions.tex
Define commutative monoid.

@@
 
+\subsection{Commutative monoids}
+
+A \emph{commutative monoid} $A = (A, \varepsilon, +)$ is a set equipped with distinguished element
+$\varepsilon \in A$ called the \emph{unit} and associative binary operation $+$ satisfying $\varepsilon + x =
+x$ and $x + \varepsilon = x$ for any $x \in A$.
+
 \subsection{Category of commutative monoids}
 
-$\CMon$ is the category whose objects are the commutative monoids $M$ and morphisms $M \to M'$ are the
+$\CMon$ is the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are the
 commutative monoid homomorphisms. $\CMon$ is symmetric closed monoidal, with the trivial one-element monoid
-$\One$ as terminal object and monoidal product $M \tensor M'$ given by the direct product of $M$ and $M'$.
+$\One$ as terminal object and monoidal product $A \tensor B$ given by the direct product of $A$ and $B$.
 $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
+
+Although $\CMon$ is not Cartesian closed, every hom-object $\CMon(A,B)$ is a commutative monoid, with the
+constant homomorphism that sends every element of the carrier of $A$ to $1_A$ as the unit, and pointwise
+multiplication as the binary operation.
 

### 2025-02-21 notes/definitions.tex
CMon isn't Cartesian closed but there is a commutative monoid structure on the hom-objects.

@@
 $\varepsilon \in A$ called the \emph{unit} and associative binary operation $+$ satisfying $\varepsilon + x =
-x$ and $x + \varepsilon = x$ for any $x \in A$.
+x$ and $x + \varepsilon = x$ for any $x \in A$. A commutative monoid homomorphism from $A$ to $B$ is any
+function $f: A \to B$ which preserves $\varepsilon$ and $+$.
 
@@
 
-Although $\CMon$ is not Cartesian closed, every hom-object $\CMon(A,B)$ is a commutative monoid, with the
-constant homomorphism that sends every element of the carrier of $A$ to $1_A$ as the unit, and pointwise
-multiplication as the binary operation.
+Although $\CMon$ is not Cartesian closed, every hom-object $\CMon(A,B)$ is a commutative monoid, with unit
+given by the constant homomorphism that sends every element of $A$ to $\varepsilon_B$, and binary operation
+given by $(f + g)(a) = f(a) + f(b)$.
 

### 2025-02-21 notes/definitions.tex
Actually no, CMon is closed monoidal and therefore is CMon-enriched.

@@
 
-Although $\CMon$ is not Cartesian closed, every hom-object $\CMon(A,B)$ is a commutative monoid, with unit
-given by the constant homomorphism that sends every element of $A$ to $\varepsilon_B$, and binary operation
-given by $(f + g)(a) = f(a) + f(b)$.
-
 \subsection{$\CMon$-enriched category}
 
-A category $C$ is \emph{$\CMon$-enriched} if every hom-object $\Hom{C}{A}{B}$ is a commutative monoid with a
-\emph{zero} morphism $\zero_{A,B}$ and addition of morphisms $+_{A,B}$ as the commutative monoid operations,
-and composition is a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
+A category $C$ is \emph{$\CMon$-enriched} if:
+\begin{enumerate}
+\item every hom-object $\Hom{C}{A}{B}$ is a commutative monoid with a
+\emph{zero} morphism $\zero_{A,B}$ as unit and addition of morphisms $+_{A,B}$ as the monoid operation;
+\item composition is a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
 i.e.~satisfies
@@
 \end{salign*}
+\end{enumerate}
+
+$\CMon$ is closed monoidal, and so enriched over itself. Every hom-object $\CMon(A,B)$ is a commutative
+monoid, with:
+\begin{enumerate}
+\item unit given by the constant homomorphism that sends every element of $A$ to $\varepsilon_B$;
+\item binary operation given pointwise so that $(f + g)(a) = f(a) + f(b)$.
+\end{enumerate}
 

### 2025-02-22 notes/definitions.tex
Drop general notion of V-enriched category for arbitrary monoidal category for now.

@@
 
-\subsection{Commutative monoids}
+\subsection{$\CMon$-enriched categories}
 
-A \emph{commutative monoid} $A = (A, \varepsilon, +)$ is a set equipped with distinguished element
-$\varepsilon \in A$ called the \emph{unit} and associative binary operation $+$ satisfying $\varepsilon + x =
-x$ and $x + \varepsilon = x$ for any $x \in A$. A commutative monoid homomorphism from $A$ to $B$ is any
-function $f: A \to B$ which preserves $\varepsilon$ and $+$.
+\begin{definition}[Commutative monoid]
+A \emph{commutative monoid} $A = (A, \varepsilon, \bullet)$ is a set $A$ equipped with distinguished element
+$\varepsilon \in A$ called the \emph{unit} and associative binary operation $\bullet$ satisfying $\varepsilon
+\bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in A$.
+\end{definition}
 
-\subsection{Category of commutative monoids}
+A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ which preserves $\varepsilon$
+and $\bullet$.
 
-$\CMon$ is the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are the
-commutative monoid homomorphisms. $\CMon$ is symmetric closed monoidal, with the trivial one-element monoid
-$\One$ as terminal object and monoidal product $A \tensor B$ given by the direct product of $A$ and $B$.
-$\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
+\begin{definition}[Category of commutative monoids]
+Define $\CMon$ to be the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are
+the commutative monoid homomorphisms.
+\end{definition}
+
+$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object and
+monoidal product $A \tensor B$ given by the direct product of $A$ and $B$. $\CMon$ is complete and cocomplete,
+inheriting all limits and colimits from $\Set$.
 
@@
 
-A category $C$ is \emph{$\CMon$-enriched} if:
+We note that if a category $C$ is \emph{$\CMon$-enriched}, then:
 \begin{enumerate}
-\item every hom-object $\Hom{C}{A}{B}$ is a commutative monoid with a
-\emph{zero} morphism $\zero_{A,B}$ as unit and addition of morphisms $+_{A,B}$ as the monoid operation;
-\item composition is a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$,
-i.e.~satisfies
+\item every hom-object $\Hom{C}{A}{B}$ is a commutative monoid (writing $\zero_{A,B}$ for the unit and
+$+_{A,B}$ for the binary operation);
+\item composition is given by a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in
+$\CMon$, i.e.~satisfies
 
@@
 \begin{enumerate}
-\item unit given by the constant homomorphism that sends every element of $A$ to $\varepsilon_B$;
+\item unit given by the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
 \item binary operation given pointwise so that $(f + g)(a) = f(a) + f(b)$.
@@
 
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every
-bounded lattice $L$ a commutative monoid $FL$ and to every Galois connection $f: L \to L'$ .
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.
+A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $FL$ and to
+every Galois connection $f: L \to L'$ . $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of
+commutative monoids on $\LatGal$.

### 2025-02-22 notes/definitions.tex
Tweakery.

@@
 
-\todo{Some properties of $\Fam(C)$.}
-
 \subsection{Category of bounded lattices and Galois connections}
@@
 
-\noindent $\LatGal$ has finite biproducts. \todo{Tie this to pre-additive structure.}
+\noindent $\LatGal$ has finite biproducts.
 
@@
 $+_{A,B}$ for the binary operation);
-\item composition is given by a family of morphisms $\Hom{C}{B}{C} \tensor \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in
-$\CMon$, i.e.~satisfies
+\item composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
+\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
+$\Hom{C}{A}{B}$ separately:
 
@@
 
-$\CMon$ is closed monoidal, and so enriched over itself. Every hom-object $\CMon(A,B)$ is a commutative
-monoid, with:
+$\CMon$ is closed monoidal, and so enriched over itself; every hom-object $\CMon(A,B)$ is a commutative
+monoid with:
 \begin{enumerate}
@@
 
+\begin{proposition}
 If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
+\end{proposition}
 
@@
 
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $FL$ and to
-every Galois connection $f: L \to L'$ . $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of
-commutative monoids on $\LatGal$.
+A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
+every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
+$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-22 notes/definitions.tex
Move LatGal to after biproducts.
Clarifying self-enrichment of CMon.

@@
 
-\subsection{Indexed families of objects}
+\subsection{Category of families}
 
+\begin{definition}[$\Fam(I,C)$]
 For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
@@
 \end{itemize}
+\end{definition}
 
 \noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category. $\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any
-function $f: I \to J$ to the following \emph{reindexing functor}.
+category.
 
@@
 
-\subsection{Category of families}
+\begin{definition}[$\Fam(-,C)$ functor]
+$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
+$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
+\end{definition}
 
-The \emph{Grothendieck construction} $\Grothendieck{D}F$ for a functor $F: D \to \Cat$ is the category where:
+\begin{definition}[Grothendieck construction for $F$]
+Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
+category where:
 \begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $FI$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)X
-\to Y$ in $FJ$.
+\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
+\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
+\to Y$ in $F(J)$.
 \end{itemize}
+\end{definition}
 
 \noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to (Ff)Y$ in $FI$. Writing $\Fam(C)$ for the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$ we then have that in $\Fam(C)$:
+paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
+
+\begin{definition}[Category of families]
+For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
+$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
+\end{definition}
+
+In $\Fam(C)$:
 \begin{itemize}
@@
 \end{itemize}
-
-\subsection{Category of bounded lattices and Galois connections}
-
-Define the category $\LatGal$ where:
-\begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
-\end{itemize}
-
-\noindent $\LatGal$ has finite biproducts.
 
@@
 
-A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ which preserves $\varepsilon$
-and $\bullet$.
+A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
+$\bullet$.
 
@@
 
-We note that if a category $C$ is \emph{$\CMon$-enriched}, then:
+We note that if a category $C$ is \emph{$\CMon$-enriched}, we have that:
 \begin{enumerate}
-\item every hom-object $\Hom{C}{A}{B}$ is a commutative monoid (writing $\zero_{A,B}$ for the unit and
-$+_{A,B}$ for the binary operation);
-\item composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
+\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid; the unit $\zero_{A,B}$ is a zero morphism and
+the binary operation $+_{A,B}$ is addition of morphisms.
+\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
 \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
@@
 
-$\CMon$ is closed monoidal, and so enriched over itself; every hom-object $\CMon(A,B)$ is a commutative
-monoid with:
+$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
+monoid where:
 \begin{enumerate}
-\item unit given by the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item binary operation given pointwise so that $(f + g)(a) = f(a) + f(b)$.
+\item unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
+\item addition of morphisms $+$ is defined pointwise so that $+_{A}$ so that $(f + g)(a) = f(a) + g(a)$.
 \end{enumerate}
@@
 \begin{proposition}
-If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
+If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
 \end{proposition}
@@
 
+\subsection{Category of bounded lattices and Galois connections}
+
+Define the category $\LatGal$ where:
+\begin{itemize}
+\item objects $L$ are the bounded lattices;
+\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
+join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
+$y \in L'$.
+\end{itemize}
+
+\noindent $\LatGal$ has finite biproducts.
+
 \subsection{Category $\Func{\LatGal^\op}{\CMon}$}

### 2025-02-22 notes/definitions.tex
Clarifying self-enrichment of CMon.

@@
 \begin{enumerate}
-\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid; the unit $\zero_{A,B}$ is a zero morphism and
-the binary operation $+_{A,B}$ is addition of morphisms.
+\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$ (zero
+morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
 \item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
@@
 $\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
-monoid where:
+monoid of commutative monoid homomorphisms where:
 \begin{enumerate}
-\item unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item addition of morphisms $+$ is defined pointwise so that $+_{A}$ so that $(f + g)(a) = f(a) + g(a)$.
+\item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
+\item addition of morphisms $+$ is defined pointwise so that $(f + g)(a) = f(a) + g(a)$.
 \end{enumerate}

### 2025-02-24 notes/definitions.tex
Consolidate propositions about CMon-enriched categories and biproducts.

@@
 
-In $\Fam(C)$:
+In $\Fam(C)$ we then have that:
 \begin{itemize}
@@
 
-We note that if a category $C$ is \emph{$\CMon$-enriched}, we have that:
+We note that if a category $C$ is $\CMon$-enriched, then:
 \begin{enumerate}
@@
 \end{enumerate}
-
-\begin{proposition}
-Suppose $C$ a $\CMon$-enriched monoidal category.
-\begin{itemize}
-\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
-\end{itemize}
-\end{proposition}
 
@@
 
-Any category with biproducts has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A \oplus
-B, i_A, i_B)$.
+\begin{proposition}
+\item
+\begin{itemize}
+\item If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
+\oplus B, i_A, i_B)$.
+\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
+\item If $C$ has a terminal (or initial) object then it is a zero object.
+\end{itemize}
+\end{proposition}
 

### 2025-02-24 notes/definitions.tex
Remove mention of zero objects for now.

@@
 
+If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
+\oplus B, i_A, i_B)$.
+
 \begin{proposition}
@@
 \begin{itemize}
-\item If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
-\oplus B, i_A, i_B)$.
 \item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
+\item If $C$ has a terminal (or initial) object then it is both initial and terminal.
 \end{itemize}

### 2025-02-24 notes/definitions.tex
Define 'zero object'.

@@
 
-$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object and
-monoidal product $A \tensor B$ given by the direct product of $A$ and $B$. $\CMon$ is complete and cocomplete,
-inheriting all limits and colimits from $\Set$.
+$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object, monoidal
+product $A \tensor B$ given by the direct product of $A$ and $B$, and braiding given by the isomorphism $A
+\times B \iso B \times A$. $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
 
@@
 
-Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
-object $A \oplus B$ of $C$ together with morphisms
+\begin{definition}[Zero object]
+A \emph{zero} object is an object which is both terminal and initial.
+\end{definition}
+
+Suppose $C$ a $\CMon$-enriched category.
+
+\begin{definition}[Biproduct]
+For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \oplus B$ of $C$ together with
+morphisms
 
@@
 \end{salign*}
+\end{definition}
 
@@
 \begin{itemize}
-\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
-\item If $C$ has a terminal (or initial) object then it is both initial and terminal.
+\item If $C$ has products (or coproducts) then they are biproducts.
+\item If $C$ has a terminal (or initial) object then it is a zero object.
 \end{itemize}
 \end{proposition}
+
+\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.
 

### 2025-02-25 notes.tex
Drop general notion of V-enriched category for arbitrary monoidal category for now.
Separate files for each section. Move Fam towards the end.

@@
 
-Is one of the goals to define sufficient conditions on $C$ for $\Fam(C)$ to be Cartesian closed?
+% Is one of the goals to define sufficient conditions on $C$ for $\Fam(C)$ to be Cartesian closed?
 
-\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but I'm glossing that detail here.
+\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but we will gloss that detail for now.
 
-\input{notes/definitions}
+\section{Definitions}
+
+\input{notes/sec/cmon-enriched}
+\input{notes/sec/biproduct}
+\input{notes/sec/fam}
+\input{notes/sec/galois}
+
 % \input{notes/language}

### 2025-02-25 notes/sec/biproduct.tex
Separate files for each section. Move Fam towards the end.

--- 2025-02-24 notes/definitions.tex
@@
-\section{Definitions}
-
-\subsection{Category of families}
-
-\begin{definition}[$\Fam(I,C)$]
-For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
-\begin{itemize}
-\item objects are the $I$-indexed families of objects of $C$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
-\end{itemize}
-\end{definition}
-
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category.
-
-\begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and
-similarly for morphisms.
-\end{definition}
-
-\begin{definition}[$\Fam(-,C)$ functor]
-$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
-$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
-\end{definition}
-
-\begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
-category where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
-\to Y$ in $F(J)$.
-\end{itemize}
-\end{definition}
-
-\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
-
-\begin{definition}[Category of families]
-For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
-\end{definition}
-
-In $\Fam(C)$ we then have that:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
-
-\subsection{$\CMon$-enriched categories}
-
-\begin{definition}[Commutative monoid]
-A \emph{commutative monoid} $A = (A, \varepsilon, \bullet)$ is a set $A$ equipped with distinguished element
-$\varepsilon \in A$ called the \emph{unit} and associative binary operation $\bullet$ satisfying $\varepsilon
-\bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in A$.
-\end{definition}
-
-A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
-$\bullet$.
-
-\begin{definition}[Category of commutative monoids]
-Define $\CMon$ to be the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are
-the commutative monoid homomorphisms.
-\end{definition}
-
-$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object, monoidal
-product $A \tensor B$ given by the direct product of $A$ and $B$, and braiding given by the isomorphism $A
-\times B \iso B \times A$. $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
-
-\subsection{$\CMon$-enriched category}
-
-We note that if a category $C$ is $\CMon$-enriched, then:
-\begin{enumerate}
-\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$ (zero
-morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
-\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
-$\Hom{C}{A}{B}$ separately:
-
-\begin{salign*}
-f \comp \zero = f = \zero \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-\end{enumerate}
-
-$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
-monoid of commutative monoid homomorphisms where:
-\begin{enumerate}
-\item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item addition of morphisms $+$ is defined pointwise so that $(f + g)(a) = f(a) + g(a)$.
-\end{enumerate}
-
-\begin{proposition}
-If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
-\end{proposition}
-
 \subsection{Biproducts}
@@
 \noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.
-
-\subsection{Category of bounded lattices and Galois connections}
-
-Define the category $\LatGal$ where:
-\begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
-\end{itemize}
-
-\noindent $\LatGal$ has finite biproducts.
-
-\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
-
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
-every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-25 notes/sec/cmon-enriched.tex
Separate files for each section. Move Fam towards the end.

--- 2025-02-24 notes/definitions.tex
@@
-\section{Definitions}
-
-\subsection{Category of families}
-
-\begin{definition}[$\Fam(I,C)$]
-For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
-\begin{itemize}
-\item objects are the $I$-indexed families of objects of $C$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
-\end{itemize}
-\end{definition}
-
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category.
-
-\begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and
-similarly for morphisms.
-\end{definition}
-
-\begin{definition}[$\Fam(-,C)$ functor]
-$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
-$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
-\end{definition}
-
-\begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
-category where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
-\to Y$ in $F(J)$.
-\end{itemize}
-\end{definition}
-
-\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
-
-\begin{definition}[Category of families]
-For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
-\end{definition}
-
-In $\Fam(C)$ we then have that:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
-
 \subsection{$\CMon$-enriched categories}
@@
 
-\subsection{$\CMon$-enriched category}
+\subsubsection{$\CMon$-enriched categories}
 
@@
 \end{proposition}
-
-\subsection{Biproducts}
-
-\begin{definition}[Zero object]
-A \emph{zero} object is an object which is both terminal and initial.
-\end{definition}
-
-Suppose $C$ a $\CMon$-enriched category.
-
-\begin{definition}[Biproduct]
-For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \oplus B$ of $C$ together with
-morphisms
-
-\begin{center}
-\begin{tikzcd}
-   A \arrow[r, "i_A", shift left] &
-   A \oplus B \arrow[l, "p_A", shift left] \arrow[r, "p_B"', shift right] &
-   B \arrow[l, "i_B"', shift right]
-\end{tikzcd}
-\end{center}
-
-\noindent satisfying
-
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_A \comp i_A &= \id_A \\
-   p_B \comp i_A &= \zero_{A,B}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_B \comp i_B &= \id_B \\
-   p_A \comp i_B &= \zero_{B,A}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-i_A \comp p_A + i_B \comp p_B &= \id_{A \oplus B}
-\end{salign*}
-\end{definition}
-
-If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
-\oplus B, i_A, i_B)$.
-
-\begin{proposition}
-\item
-\begin{itemize}
-\item If $C$ has products (or coproducts) then they are biproducts.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
-\end{itemize}
-\end{proposition}
-
-\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.
-
-\subsection{Category of bounded lattices and Galois connections}
-
-Define the category $\LatGal$ where:
-\begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
-\end{itemize}
-
-\noindent $\LatGal$ has finite biproducts.
-
-\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
-
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
-every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-25 notes/sec/fam.tex
Separate files for each section. Move Fam towards the end.

--- 2025-02-24 notes/definitions.tex
@@
-\section{Definitions}
-
 \subsection{Category of families}
@@
 \end{itemize}
-
-\subsection{$\CMon$-enriched categories}
-
-\begin{definition}[Commutative monoid]
-A \emph{commutative monoid} $A = (A, \varepsilon, \bullet)$ is a set $A$ equipped with distinguished element
-$\varepsilon \in A$ called the \emph{unit} and associative binary operation $\bullet$ satisfying $\varepsilon
-\bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in A$.
-\end{definition}
-
-A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
-$\bullet$.
-
-\begin{definition}[Category of commutative monoids]
-Define $\CMon$ to be the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are
-the commutative monoid homomorphisms.
-\end{definition}
-
-$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object and
-monoidal product $A \tensor B$ given by the direct product of $A$ and $B$. $\CMon$ is complete and cocomplete,
-inheriting all limits and colimits from $\Set$.
-
-\subsection{$\CMon$-enriched category}
-
-We note that if a category $C$ is $\CMon$-enriched, then:
-\begin{enumerate}
-\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$ (zero
-morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
-\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
-$\Hom{C}{A}{B}$ separately:
-
-\begin{salign*}
-f \comp \zero = f = \zero \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-\end{enumerate}
-
-$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
-monoid of commutative monoid homomorphisms where:
-\begin{enumerate}
-\item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item addition of morphisms $+$ is defined pointwise so that $(f + g)(a) = f(a) + g(a)$.
-\end{enumerate}
-
-\begin{proposition}
-If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
-\end{proposition}
-
-\subsection{Biproducts}
-
-Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
-object $A \oplus B$ of $C$ together with morphisms
-
-\begin{center}
-\begin{tikzcd}
-   A \arrow[r, "i_A", shift left] &
-   A \oplus B \arrow[l, "p_A", shift left] \arrow[r, "p_B"', shift right] &
-   B \arrow[l, "i_B"', shift right]
-\end{tikzcd}
-\end{center}
-
-\noindent satisfying
-
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_A \comp i_A &= \id_A \\
-   p_B \comp i_A &= \zero_{A,B}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_B \comp i_B &= \id_B \\
-   p_A \comp i_B &= \zero_{B,A}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-i_A \comp p_A + i_B \comp p_B &= \id_{A \oplus B}
-\end{salign*}
-
-\begin{proposition}
-\item
-\begin{itemize}
-\item If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
-\oplus B, i_A, i_B)$.
-\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
-\end{itemize}
-\end{proposition}
-
-\subsection{Category of bounded lattices and Galois connections}
-
-Define the category $\LatGal$ where:
-\begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
-\end{itemize}
-
-\noindent $\LatGal$ has finite biproducts.
-
-\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
-
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
-every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-02-25 notes/sec/galois.tex
Separate files for each section. Move Fam towards the end.

--- 2025-02-24 notes/definitions.tex
@@
-\section{Definitions}
-
-\subsection{Category of families}
-
-\begin{definition}[$\Fam(I,C)$]
-For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
-\begin{itemize}
-\item objects are the $I$-indexed families of objects of $C$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
-\end{itemize}
-\end{definition}
-
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
-category.
-
-\begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and
-similarly for morphisms.
-\end{definition}
-
-\begin{definition}[$\Fam(-,C)$ functor]
-$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
-$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
-\end{definition}
-
-\begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
-category where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
-\to Y$ in $F(J)$.
-\end{itemize}
-\end{definition}
-
-\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
-
-\begin{definition}[Category of families]
-For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
-\end{definition}
-
-In $\Fam(C)$ we then have that:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
-
-\subsection{$\CMon$-enriched categories}
-
-\begin{definition}[Commutative monoid]
-A \emph{commutative monoid} $A = (A, \varepsilon, \bullet)$ is a set $A$ equipped with distinguished element
-$\varepsilon \in A$ called the \emph{unit} and associative binary operation $\bullet$ satisfying $\varepsilon
-\bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in A$.
-\end{definition}
-
-A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
-$\bullet$.
-
-\begin{definition}[Category of commutative monoids]
-Define $\CMon$ to be the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are
-the commutative monoid homomorphisms.
-\end{definition}
-
-$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object and
-monoidal product $A \tensor B$ given by the direct product of $A$ and $B$. $\CMon$ is complete and cocomplete,
-inheriting all limits and colimits from $\Set$.
-
-\subsection{$\CMon$-enriched category}
-
-We note that if a category $C$ is $\CMon$-enriched, then:
-\begin{enumerate}
-\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$ (zero
-morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
-\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
-$\Hom{C}{A}{B}$ separately:
-
-\begin{salign*}
-f \comp \zero = f = \zero \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-\end{enumerate}
-
-$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
-monoid of commutative monoid homomorphisms where:
-\begin{enumerate}
-\item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item addition of morphisms $+$ is defined pointwise so that $(f + g)(a) = f(a) + g(a)$.
-\end{enumerate}
-
-\begin{proposition}
-If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
-\end{proposition}
-
-\subsection{Biproducts}
-
-Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
-object $A \oplus B$ of $C$ together with morphisms
-
-\begin{center}
-\begin{tikzcd}
-   A \arrow[r, "i_A", shift left] &
-   A \oplus B \arrow[l, "p_A", shift left] \arrow[r, "p_B"', shift right] &
-   B \arrow[l, "i_B"', shift right]
-\end{tikzcd}
-\end{center}
-
-\noindent satisfying
-
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_A \comp i_A &= \id_A \\
-   p_B \comp i_A &= \zero_{A,B}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   p_B \comp i_B &= \id_B \\
-   p_A \comp i_B &= \zero_{B,A}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-i_A \comp p_A + i_B \comp p_B &= \id_{A \oplus B}
-\end{salign*}
-
-\begin{proposition}
-\item
-\begin{itemize}
-\item If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
-\oplus B, i_A, i_B)$.
-\item If $C$ has products (or coproducts), then they are biproducts and thus $C$ is semi-additive.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
-\end{itemize}
-\end{proposition}
-
 \subsection{Category of bounded lattices and Galois connections}

### 2025-02-25 notes/sec/fam.tex
Assume C locally small.

@@
 
-\begin{definition}[$\Fam(I,C)$]
+\begin{definition}[Category of $I$-indexed families of objects]
 For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
@@
 \end{itemize}
+
+\begin{proposition}
+Suppose $C$ a locally small category.
+\begin{enumerate}
+\item If $C$ has binary coproducts and set-indexed products, then $\Fam(C)$ is symmetric monoidal, with the
+coproduct as the monoidal product.
+\item If the coproducts are in fact biproducts, then $\Fam(C)$ is Cartesian closed.
+\end{enumerate}
+\end{proposition}
+
+\begin{itemize}
+\item The internal hom $[X, Y]$
+\end{itemize}

### 2025-02-25 notes/sec/grothendieck.tex
We need that Fam(-) preserves products.

--- 2025-02-25 notes/sec/fam.tex
@@
 \begin{proposition}
-Suppose $C$ a locally small category.
+If $C$ has products then $\Fam(C)$ has products.
+\end{proposition}
+
+\begin{proposition}
+Suppose $C$ locally small.
 \begin{enumerate}
@@
 
-\begin{itemize}
-\item The internal hom $[X, Y]$
-\end{itemize}
+\begin{proof}
+Suppose $X = (I, X)$ and $Y = (I, Y)$ are objects in $\Fam(C)$. Define $[X, Y]$ to be the object $(K,Z)$ in
+$\Fam(C)$ where $K$ is the set of morphisms from $\Fam(I,C)$ to $\Fam(J,C)$ and $Z$ is the $K$-indexed family
+which maps every $f: \Fam(I,C) \to \Fam(J,C)$ in $K$ to $\prod_{i \in I}\reindex{Y}{f}_i$. Then the following
+families of morphisms in $\Fam(C)$:
+\begin{enumerate}
+\item $\lambda_{X,Y,Z}: \Fam(C)(X \times Y, Z) \to \Fam(C)(X, [Y, Z])$
+\item $\eval_{X,Y}: [X, Y] \times X \to Y$
+\end{enumerate}
+\end{proof}

### 2025-02-25 notes/sec/grothendieck.tex
If 𝐶 has binary products then so does Fam(𝐶).
Constructing lambda.
Corrections.

--- 2025-02-25 notes/sec/fam.tex
@@
 \begin{proposition}
-Suppose $C$ a locally small category.
+If $C$ has binary products then so does $\Fam(C)$.
+\end{proposition}
+
+\begin{proposition}
+Suppose $C$ locally small.
 \begin{enumerate}
@@
 
-\begin{itemize}
-\item The internal hom $[X, Y]$
-\end{itemize}
+\begin{proof}
+Suppose $C$ locally small with binary biproducts given by $\biprod$. Suppose $X = (I, X)$ and $Y = (I, Y)$ are
+objects in $\Fam(C)$. Define $[X, Y]$ to be the object $(K,Z)$ in $\Fam(C)$ where $K$ is $\Fam(C)(X,Y)$ and
+$Z$ is the $K$-indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $K$ to $\prod_{i \in
+I}\reindex{Y}{f}_i$. Then define the following families of functions:
+
+\begin{enumerate}
+\item $\lambda_{X,Y,Z}$ which sends any morphism $(f: I \times J \to K, g: X \times Y \to \reindex{Z}{f}): X
+\times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X \to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ where
+$f'(i)(j) = f(i,j)$ and $g'(i) = $.
+\item $\eval_{X,Y}: [X, Y] \times X \to Y$
+\end{enumerate}
+\end{proof}

### 2025-02-25 notes/sec/biproduct.tex
specifying second component of f'(i).

@@
 \begin{definition}[Biproduct]
-For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \oplus B$ of $C$ together with
+For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \biprod B$ of $C$ together with
 morphisms
@@
 \begin{tikzcd}
-   A \arrow[r, "i_A", shift left] &
-   A \oplus B \arrow[l, "p_A", shift left] \arrow[r, "p_B"', shift right] &
-   B \arrow[l, "i_B"', shift right]
+   A \arrow[r, "\inj_A", shift left] &
+   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] &
+   B \arrow[l, "\inj_B"', shift right]
 \end{tikzcd}
@@
 \begin{salign*}
-   p_A \comp i_A &= \id_A \\
-   p_B \comp i_A &= \zero_{A,B}
+   \proj_A \comp \inj_A &= \id_A \\
+   \proj_B \comp \inj_A &= \zero_{A,B}
 \end{salign*}
@@
 \begin{salign*}
-   p_B \comp i_B &= \id_B \\
-   p_A \comp i_B &= \zero_{B,A}
+   \proj_B \comp \inj_B &= \id_B \\
+   \proj_A \comp \inj_B &= \zero_{B,A}
 \end{salign*}
@@
 \begin{salign*}
-i_A \comp p_A + i_B \comp p_B &= \id_{A \oplus B}
+\inj_A \comp \proj_A + \inj_B \comp \proj_B &= \id_{A \proj B}
 \end{salign*}
@@
 
-If $C$ has biproducts then it has products given by $(A \oplus B, p_A, p_B)$ and coproducts given by $(A
-\oplus B, i_A, i_B)$.
+If $C$ has biproducts then it has products given by $(A \biprod B, \proj_A, \proj_B)$ and coproducts given by $(A
+\biprod B, \inj_A, \inj_B)$.
 

### 2025-02-26 notes/sec/grothendieck.tex
specifying second component of f'(i).
Add assumption that C has set-indexed products.

--- 2025-02-25 notes/sec/fam.tex
@@
 
+\begin{proof}
+Suppose $C$ locally small with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
+given by $(\Pi,\pi)$. For any objects $X = (I, X)$ and $Y = (J, Y)$ in $\Fam(C)$, define $[X, Y]$ to be the
+object $(K,Z)$ in $\Fam(C)$ where $K$ is $\Fam(C)(X,Y)$ and $Z$ is the $K$-indexed family which maps every
+$(f: I \to J, g: X \to \reindex{Y}{f})$ in $K$ to $\prod_{i \in I}\reindex{Y}{f}_i$. Then define the following
+families of functions:
+
+\begin{enumerate}
+\item $\lambda_{X,Y,Z}$ which sends any morphism $(f: I \times J \to K, g: X \times Y \to \reindex{Z}{f}): X
+\times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X \to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ where:
 \begin{itemize}
-\item The internal hom $[X, Y]$
+\item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: \Fam(J,C))$
+\item $g'(i) = $.
 \end{itemize}
+\item $\eval_{X,Y}: [X, Y] \times X \to Y$
+\end{enumerate}
+\end{proof}

### 2025-02-26 notes/sec/grothendieck.tex
Restructure a bit.

@@
 Suppose $C$ locally small with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
-given by $(\Pi,\pi)$. For any objects $X = (I, X)$ and $Y = (J, Y)$ in $\Fam(C)$, define $[X, Y]$ to be the
-object $(K,Z)$ in $\Fam(C)$ where $K$ is $\Fam(C)(X,Y)$ and $Z$ is the $K$-indexed family which maps every
-$(f: I \to J, g: X \to \reindex{Y}{f})$ in $K$ to $\prod_{i \in I}\reindex{Y}{f}_i$. Then define the following
-families of functions:
+given by $(\prod,\pi)$. For any objects $X = (I, X), Y = (J, Y)$ and $Z = (K, Z)$ in $\Fam(C)$ define:
 
 \begin{enumerate}
-\item $\lambda_{X,Y,Z}$ which sends any morphism $(f: I \times J \to K, g: X \times Y \to \reindex{Z}{f}): X
-\times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X \to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ where:
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),W)$ in $\Fam(C)$ where $W$ is the indexed family which maps
+every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in I}\reindex{Y}{f}_i$.
+\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $(f: I \times J
+\to K, g: X \times Y \to \reindex{Z}{f}): X \times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X
+\to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ where:
 \begin{itemize}

### 2025-02-26 notes/sec/grothendieck.tex
Restructure a bit.
Give type of (−,𝑖) for clarify.
Onto g'.

@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),W)$ in $\Fam(C)$ where $W$ is the indexed family which maps
-every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in I}\reindex{Y}{f}_i$.
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where (overloading notation) $[X,Y]$ is
+the indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in
+I}\reindex{Y}{f}_i$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $(f: I \times J
@@
 \begin{itemize}
-\item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: \Fam(J,C))$
-\item $g'(i) = $.
+\item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: Y \to \reindex{Z}{f \comp (i,
+-)})$
+\item $g' = \{\}_{i \in I}$.
+\item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$
 \end{itemize}

### 2025-02-26 notes/sec/set-indexed-product.tex
Set-indexed products. Understand the role of the constant family now.

\subsection{Set-indexed products}

$C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
there exists an object $\prod_{i \in I}X_i$ in $C$ and family of morphisms $\eval_i$ and $\lambda$ such that
for any $I$-indexed family $Y$ and any family of morphisms $f_i: Y \to X_i$ in $C$,

\begin{center}
\begin{tikzcd}
   \prod_{i \in I}X_i \arrow[r, "\eval_i"] &
   X_i
   \\
   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda"]
\end{tikzcd}
\end{center}

Note that $f$ is a morphism from the $I$-indexed family that is constantly $Y$ to $X$ in $\Fam(I,C)$.

### 2025-02-26 notes/sec/set-indexed-product.tex
First pass over set-indexed products.

@@
 $C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
-there exists an object $\prod_{i \in I}X_i$ in $C$ and family of morphisms $\eval_i$ and $\lambda$ such that
-for any $I$-indexed family $Y$ and any family of morphisms $f_i: Y \to X_i$ in $C$,
+there exists an object $\prod_{i \in I}X_i$ in $C$ and family of morphisms $\eval_i$ such that for any object
+$Y$ of $C$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism
+$\lambda f$ making the following diagram commute:
 
@@
    \\
-   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda"]
+   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda f"]
 \end{tikzcd}
@@
 
-Note that $f$ is a morphism from the $I$-indexed family that is constantly $Y$ to $X$ in $\Fam(I,C)$.
+Note that $f$ is a morphism in $\Fam(I,C)$ from the constant family $\{Y\}_{i \in I}$ to $X$.

### 2025-02-26 notes/sec/set-indexed-product.tex
First pass over set-indexed products.

@@
 $C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
-there exists an object $\prod_{i \in I}X_i$ in $C$ and family of morphisms $\eval_i$ such that for any object
+there exists an object $\prod_{i \in I}X_i$ and family of morphisms $\eval_i$ in $C$ such that for any object
 $Y$ of $C$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism

### 2025-02-26 notes/sec/set-indexed-product.tex
First pass over set-indexed products.

@@
 there exists an object $\prod_{i \in I}X_i$ and family of morphisms $\eval_i$ in $C$ such that for any object
-$Y$ of $C$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism
-$\lambda f$ making the following diagram commute:
+$Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism $\lambda
+f$ making the following diagram commute:
 
@@
    \\
-   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda f"]
+   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda f"]
 \end{tikzcd}

### 2025-02-26 notes/sec/grothendieck.tex
Set-indexed products. Understand the role of the constant family now.
Some progress on different g's.

@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where (overloading notation) $[X,Y]$ is
-the indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where we also write $[X,Y]$ for the
+indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in
 I}\reindex{Y}{f}_i$.
@@
 \to K, g: X \times Y \to \reindex{Z}{f}): X \times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X
-\to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ where:
+\to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ with:
 \begin{itemize}
 \item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: Y \to \reindex{Z}{f \comp (i,
--)})$
-\item $g' = \{\}_{i \in I}$.
+-)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$;
+\item $g' = \{g_{i, j} \comp \inj_{X_i} \}_{i \in I}$ where $\inj_{X_i}: X_i \to X_i \biprod Y_j$;
 \item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$

### 2025-02-26 notes/sec/grothendieck.tex
Some progress on different g's.
Progress on using lambda of set-indexed products, not sure about the constant family though..
epsilon/lambda laws.
Tweaks.

@@
 Suppose $C$ locally small with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
-given by $(\prod,\pi)$. For any objects $X = (I, X), Y = (J, Y)$ and $Z = (K, Z)$ in $\Fam(C)$ define:
+given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X = (I, X), Y = (J, Y)$ and $Z = (K, Z)$ in
+$\Fam(C)$ define:
 
@@
 \item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: Y \to \reindex{Z}{f \comp (i,
--)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$;
-\item $g' = \{g_{i, j} \comp \inj_{X_i} \}_{i \in I}$ where $\inj_{X_i}: X_i \to X_i \biprod Y_j$;
-\item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$
+-)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
+\item $g' = \lambda_{\prod}(\{g_{i, j} \comp \inj_{X_i} \}_{i \in I})$ where $\inj_{X_i}: X_i \to X_i \biprod
+Y_j$
+\item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$.
 \end{itemize}
@@
 \end{enumerate}
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
+\begin{enumerate}
+\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f) \times \id_{Y}) = f$
+\item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$
+\end{enumerate}
+\todo{Prove this bit}
 \end{proof}

### 2025-02-27 notes/sec/grothendieck.tex
Rough pass over \epsilon.
Layout.

@@
 Y_j$
-\item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$.
 \end{itemize}
-\item $\eval_{X,Y}: [X, Y] \times X \to Y$
+using $(-, i)$ to denote the reindexing function $j \mapsto (i, j): J \to I \times J$.
+\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ where $\eval_{X,Y} = (((f, g),
+i) \mapsto f(i): \Fam(C)(X,Y) \times I \to J, \{[{\eval_{\prod}}_i, g_i] \mid ((f, g), i) \in \Fam(C)(X,Y)
+\times I\}: X \to \reindex{Y}{f})$ (using $[-,-]$ to denote coproduct of morphisms).
 \end{enumerate}

### 2025-02-27 notes/sec/grothendieck.tex
Layout.
Notationally quite intense..
Fractionally happier with notation.
Need to know that Fam(C) is locally small.

--- 2025-02-26 notes/sec/grothendieck.tex
@@
 \begin{proposition}
-If $C$ has binary products then so does $\Fam(C)$.
+\item
+\begin{enumerate}
+\item If $C$ is locally small then so is $\Fam(C)$.
+\item If $C$ has binary products then so does $\Fam(C)$.
+\end{enumerate}
 \end{proposition}
@@
 Y_j$
-\item $(-, i)$ is the reindexing function $j \mapsto (i, j): J \to I \times J$.
 \end{itemize}
-\item $\eval_{X,Y}: [X, Y] \times X \to Y$
+using $(-, i)$ to denote the reindexing function $j \mapsto (i, j): J \to I \times J$.
+\item The family of morphisms $\eval_{X,Y} = (f', g'): [X, Y] \times X \to Y$ in $\Fam(C)$ with
+\begin{itemize}
+\item $f': \Fam(C)(X,Y) \times I \to J = ((f, g), i) \mapsto f(i)$
+\item $g': [X, Y] \times X \to \reindex{Y}{f} = ((f, g), i) \mapsto [{\eval_{\prod}}_i, g_i]$
+\end{itemize}
+using $[-,-]$ to denote coproduct of morphisms.
 \end{enumerate}

### 2025-02-27 notes/sec/grothendieck.tex
Less general prop 2.12.

@@
 \begin{proposition}
-Suppose $C$ locally small.
-\begin{enumerate}
-\item If $C$ has binary coproducts and set-indexed products, then $\Fam(C)$ is symmetric monoidal, with the
-coproduct as the monoidal product.
-\item If the coproducts are in fact biproducts, then $\Fam(C)$ is Cartesian closed.
-\end{enumerate}
+Suppose $C$ locally small. If $C$ has binary biproducts and set-indexed products, then $\Fam(C)$ is Cartesian
+closed. \todo{Establish first that if $C$ has binary coproducts and set-indexed products, then $\Fam(C)$ is
+symmetric monoidal, with the coproduct as the monoidal product.}
 \end{proposition}
@@
 \begin{proof}
-Suppose $C$ locally small with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
+Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
 given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X = (I, X), Y = (J, Y)$ and $Z = (K, Z)$ in

### 2025-02-27 notes/sec/set-indexed-product.tex
Tweaks.
Simpler notation for \prod of a family.

--- 2025-02-26 notes/sec/set-indexed-product.tex
@@
 $C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
-there exists an object $\prod_{i \in I}X_i$ and family of morphisms $\eval_i$ in $C$ such that for any object
-$Y$ of $C$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism
-$\lambda f$ making the following diagram commute:
+there exists an object $\prod_{i \in I}X_i$ (also written $\prod X$) and family of morphisms $\eval_i$ in $C$
+such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a
+unique morphism $\lambda f$ making the following diagram commute:
 
@@
 \begin{tikzcd}
-   \prod_{i \in I}X_i \arrow[r, "\eval_i"] &
+   \prod X \arrow[r, "\eval_i"] &
    X_i
    \\
-   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda f"]
+   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda(f)"]
 \end{tikzcd}

### 2025-02-27 notes/sec/grothendieck.tex
Drop (i,-) notation for now.
Experiment with \mapsto instead of family notation.
Simpler notation for \prod of a family.

@@
 \item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where we also write $[X,Y]$ for the
-indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to $\prod_{i \in
-I}\reindex{Y}{f}_i$.
+indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to
+$\prod\reindex{Y}{f}$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $(f: I \times J
@@
 \begin{itemize}
-\item $f'(i) = (f \comp (i, -): J \to K, \{g_{i,j} \comp \inj_{Y_j}\}_{j \in J}: Y \to \reindex{Z}{f \comp (i,
--)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
-\item $g' = \lambda_{\prod}(\{g_{i, j} \comp \inj_{X_i} \}_{i \in I})$ where $\inj_{X_i}: X_i \to X_i \biprod
+\item $f'(i) = (j \mapsto f(i, j): J \to K, j \mapsto g_{i,j} \comp \inj_{Y_j}: Y \to \reindex{Z}{j \mapsto
+f(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
+\item $g' = \lambda_{\prod}(i \mapsto g_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i \to X_i \biprod
 Y_j$
 \end{itemize}
-using $(-, i)$ to denote the reindexing function $j \mapsto (i, j): J \to I \times J$.
 \item The family of morphisms $\eval_{X,Y} = (f', g'): [X, Y] \times X \to Y$ in $\Fam(C)$ with

### 2025-02-27 notes/sec/set-indexed-product.tex
Simpler notation for \prod of a family.

--- 2025-02-26 notes/sec/set-indexed-product.tex
@@
 $C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
-there exists an object $\prod_{i \in I}X_i$ and family of morphisms $\eval_i$ in $C$ such that for any object
-$Y$ of $C$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there exists a unique morphism
-$\lambda f$ making the following diagram commute:
+there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms $\eval_i$ in
+$C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
+exists a unique morphism $\lambda f$ making the following diagram commute:
 
@@
 \begin{tikzcd}
-   \prod_{i \in I}X_i \arrow[r, "\eval_i"] &
-   X_i
-   \\
-   Y \arrow[ru, "f_i"'] \arrow[u, "\lambda f"]
+   \prod_{I} X \arrow[r, "\eval_i"] & X_i \\
+   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda(f)"]
 \end{tikzcd}

### 2025-02-27 notes/sec/grothendieck.tex
Be explicit about type of projection.
Diagram easier to grok for eval/lambda.

@@
 closed. \todo{Establish first that if $C$ has binary coproducts and set-indexed products, then $\Fam(C)$ is
-symmetric monoidal, with the coproduct as the monoidal product.}
+symmetric monoidal, with the coproduct as monoidal product.}
 \end{proposition}
@@
 \item $f': \Fam(C)(X,Y) \times I \to J = ((f, g), i) \mapsto f(i)$
-\item $g': [X, Y] \times X \to \reindex{Y}{f} = ((f, g), i) \mapsto [{\eval_{\prod}}_i, g_i]$
+\item $g': [X, Y] \times X \to \reindex{Y}{f} = ((f, g), i) \mapsto [{\eval_{\prod}}_i, g_i]$ where
+${\eval_{\prod}}_i: \prod_I\reindex{Y}{f} \to Y_{f(i)}$
 \end{itemize}
@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$ the following diagrams commute:
 \begin{enumerate}
 \item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f) \times \id_{Y}) = f$
+
+\begin{center}
+\begin{tikzcd}
+   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
+   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
+\end{tikzcd}
+\end{center}
+
 \item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$

### 2025-03-03 notes/sec/set-indexed-product.tex
Clarify set-indexed product diagram commutes for every in I.

@@
 $C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
-exists a unique morphism $\lambda f$ making the following diagram commute:
+exists a unique morphism $\lambda f$ making the following diagram commute for every $i \in I$:
 

### 2025-03-03 notes/sec/grothendieck.tex
Clarify set-indexed product diagram commutes for every in I.
Tweak.

@@
 
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
+\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$, interpreting $I$ as a discrete
 category.
@@
 \item $g' = \lambda_{\prod}(i \mapsto g_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i \to X_i \biprod
-Y_j$
+Y_j$ \todo{whence $j$?}
 \end{itemize}
@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$ the following diagrams commute:
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have \todo{(prove these)}:
 \begin{enumerate}
-\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f) \times \id_{Y}) = f$
+\item The following diagram commutes:
 
@@
 \end{enumerate}
-\todo{Prove this bit}
 \end{proof}

### 2025-03-03 notes/sec/grothendieck.tex
Missing subscript.

@@
 For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$ and
+$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
 similarly for morphisms.
@@
 indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to
-$\prod\reindex{Y}{f}$.
+$\prod_{I}\reindex{Y}{f}$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $(f: I \times J

### 2025-03-03 notes/sec/grothendieck.tex
Tweaks.

@@
 \item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where we also write $[X,Y]$ for the
-indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f})$ in $\Fam(C)(X,Y)$ to
+indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f}): X \to Y$ in $\Fam(C)$ to
 $\prod_{I}\reindex{Y}{f}$.
@@
 f(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
-\item $g' = \lambda_{\prod}(i \mapsto g_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i \to X_i \biprod
-Y_j$ \todo{whence $j$?}
+\item $g'(i) = \lambda_{\prod_{J}}(j \mapsto g_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i \to X_i \biprod
+Y_j$ for every $j \in J$
 \end{itemize}

### 2025-03-03 notes/sec/grothendieck.tex
Tweaks.
Macros for [co]product of morphisms.
Probably happy with definitions..
Start experiment with accessors that project out components of objects/morphisms in Fam(C).

@@
 $\prod_{I}\reindex{Y}{f}$.
-\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $(f: I \times J
-\to K, g: X \times Y \to \reindex{Z}{f}): X \times Y \to Z$ to the morphism $(f': I \to \Fam(C)(Y,Z), g': X
-\to \reindex{[Y, Z]}{f'}): X \to [Y, Z]$ with:
+\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $f: X \times Y
+\to Z$ to the morphism $g: X \to [Y, Z]$ with:
 \begin{itemize}
-\item $f'(i) = (j \mapsto f(i, j): J \to K, j \mapsto g_{i,j} \comp \inj_{Y_j}: Y \to \reindex{Z}{j \mapsto
-f(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
-\item $g'(i) = \lambda_{\prod_{J}}(j \mapsto g_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i \to X_i \biprod
-Y_j$ for every $j \in J$
+\item $\idxf{g}(i) = (j \mapsto \idxf{f}(i, j): J \to K, j \mapsto \famf{f}_{i,j} \comp \inj_{Y_j}: Y \to
+\reindex{Z}{j \mapsto \idxf{f}(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
+\item $\famf{g}(i) = \lambda_{\prod_{J}}(j \mapsto \famf{f}_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i
+\to X_i \biprod Y_j$ for every $j \in J$
 \end{itemize}
-\item The family of morphisms $\eval_{X,Y} = (f', g'): [X, Y] \times X \to Y$ in $\Fam(C)$ with
+\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
 \begin{itemize}
-\item $f': \Fam(C)(X,Y) \times I \to J = ((f, g), i) \mapsto f(i)$
-\item $g': [X, Y] \times X \to \reindex{Y}{f} = ((f, g), i) \mapsto [{\eval_{\prod}}_i, g_i]$ where
-${\eval_{\prod}}_i: \prod_I\reindex{Y}{f} \to Y_{f(i)}$
+\item $\idxf{\eval}_{X,Y}((f, g), i) = f(i)$
+\item $\famf{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{I}\reindex{Y}{f} \biprod X_i \to Y_{f(i)}$
 \end{itemize}
-using $[-,-]$ to denote coproduct of morphisms.
+using $\coprodM{-}{-}$ to denote coproduct of morphisms.
 \end{enumerate}

### 2025-03-03 notes/sec/grothendieck.tex
New notation better.

@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[X, Y])$ in $\Fam(C)$ where we also write $[X,Y]$ for the
-indexed family which maps every $(f: I \to J, g: X \to \reindex{Y}{f}): X \to Y$ in $\Fam(C)$ to
-$\prod_{I}\reindex{Y}{f}$.
-\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) which sends any morphism $f: X \times Y
-\to Z$ to the morphism $g: X \to [Y, Z]$ with:
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idxf{X}, \idxf{Y}])$ in $\Fam(C)$ where $[\idxf{X},
+\idxf{Y}]$ denotes the indexed family which maps every $f: X \to Y$ in $\Fam(C)$ to
+$\prod_{I}\reindex{Y}{\idxf{f}}$.
+\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any morphism $f: X \times Y \to Z$
+to $g: X \to [Y, Z]$ with:
 \begin{itemize}

### 2025-03-03 notes/sec/grothendieck.tex
Think this is a bit better.

@@
 \begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
-\to Y$ in $F(J)$.
+\item objects $X$ are pairs of an object $\idx{X}$ of $D$ and an object $\faml{X}$ of $F(\idx{X})$;
+\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $D$ paired with morphisms
+$\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
 \end{itemize}
@@
 
-\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
+\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{J}$ in
+$D$ paired with morphisms $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
 
@@
 
-In $\Fam(C)$ we then have that:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
+\noindent In $\Fam(C)$ the objects $X$ are thus pairs of a set $\idx{X}$ and indexed family $\faml{X}$ in
+$\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with a
+morphism $\faml{f}: X \to \reindex{Y}{f}$ in $\Fam(\idx{X},C)$.
 
@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idxf{X}, \idxf{Y}])$ in $\Fam(C)$ where $[\idxf{X},
-\idxf{Y}]$ denotes the indexed family which maps every $f: X \to Y$ in $\Fam(C)$ to
-$\prod_{I}\reindex{Y}{\idxf{f}}$.
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X},
+\idx{Y}]$ denotes the indexed family which maps every $f: X \to Y$ in $\Fam(C)$ to
+$\prod_{I}\reindex{Y}{\idx{f}}$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any morphism $f: X \times Y \to Z$
@@
 \begin{itemize}
-\item $\idxf{g}(i) = (j \mapsto \idxf{f}(i, j): J \to K, j \mapsto \famf{f}_{i,j} \comp \inj_{Y_j}: Y \to
-\reindex{Z}{j \mapsto \idxf{f}(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
-\item $\famf{g}(i) = \lambda_{\prod_{J}}(j \mapsto \famf{f}_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i
+\item $\idx{g}(i) = (j \mapsto \idx{f}(i, j): J \to K, j \mapsto \faml{f}_{i,j} \comp \inj_{Y_j}: Y \to
+\reindex{Z}{j \mapsto \idx{f}(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
+\item $\faml{g}(i) = \lambda_{\prod_{J}}(j \mapsto \faml{f}_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i
 \to X_i \biprod Y_j$ for every $j \in J$
@@
 \begin{itemize}
-\item $\idxf{\eval}_{X,Y}((f, g), i) = f(i)$
-\item $\famf{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{I}\reindex{Y}{f} \biprod X_i \to Y_{f(i)}$
+\item $\idx{\eval}_{X,Y}((f, g), i) = f(i)$
+\item $\faml{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{I}\reindex{Y}{f} \biprod X_i \to Y_{f(i)}$
 \end{itemize}

### 2025-03-03 notes/sec/grothendieck.tex
Better bar.

@@
 \noindent In $\Fam(C)$ the objects $X$ are thus pairs of a set $\idx{X}$ and indexed family $\faml{X}$ in
-$\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with a
-morphism $\faml{f}: X \to \reindex{Y}{f}$ in $\Fam(\idx{X},C)$.
+$\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
+morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{f}$ in $\Fam(\idx{X},C)$.
 
@@
 Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
-given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X = (I, X), Y = (J, Y)$ and $Z = (K, Z)$ in
-$\Fam(C)$ define:
+given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X, Y, Z$ in $\Fam(C)$ define:
 
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X},
-\idx{Y}]$ denotes the indexed family which maps every $f: X \to Y$ in $\Fam(C)$ to
-$\prod_{I}\reindex{Y}{\idx{f}}$.
+\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X}, \idx{Y}]$
+denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to $\prod_{I}\reindex{Y}{\idx{f}}$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any morphism $f: X \times Y \to Z$

### 2025-03-03 notes/sec/grothendieck.tex
Now some further cleanup..

@@
 
-\noindent In $\Fam(C)$ the objects $X$ are thus pairs of a set $\idx{X}$ and indexed family $\faml{X}$ in
+\noindent In $\Fam(C)$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$ in
 $\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
@@
 \begin{itemize}
-\item $\idx{g}(i) = (j \mapsto \idx{f}(i, j): J \to K, j \mapsto \faml{f}_{i,j} \comp \inj_{Y_j}: Y \to
-\reindex{Z}{j \mapsto \idx{f}(i, j)})$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
+\item $\idx{g}(i) = g'$ where $\idx{g}'(j) = \idx{f}(i, j): J \to K$ and $\faml{g}'(j) = \faml{f}_{i,j} \comp
+\inj_{Y_j}: Y \to \reindex{Z}{j \mapsto \idx{f}(i, j)}$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
 \item $\faml{g}(i) = \lambda_{\prod_{J}}(j \mapsto \faml{f}_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i

### 2025-03-03 notes/sec/grothendieck.tex
Give type of each morphism in the family.
Now back to proof.
Now I'm paying for the notation a bit...

@@
 
-\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{J}$ in
+\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
 $D$ paired with morphisms $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
@@
 denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to $\prod_{I}\reindex{Y}{\idx{f}}$.
-\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any morphism $f: X \times Y \to Z$
-to $g: X \to [Y, Z]$ with:
+\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
+$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
 \begin{itemize}
-\item $\idx{g}(i) = g'$ where $\idx{g}'(j) = \idx{f}(i, j): J \to K$ and $\faml{g}'(j) = \faml{f}_{i,j} \comp
-\inj_{Y_j}: Y \to \reindex{Z}{j \mapsto \idx{f}(i, j)}$ where $\inj_{Y_j}: Y_j \to X_i \biprod Y_j$
-\item $\faml{g}(i) = \lambda_{\prod_{J}}(j \mapsto \faml{f}_{i, j} \comp \inj_{X_i})$ where $\inj_{X_i}: X_i
-\to X_i \biprod Y_j$ for every $j \in J$
+\item $\idx{\lambda}_{X,Y,Z}(f)(i) = g$ where $\idx{g}(j) = \idx{f}(i, j): \idx{Y} \to \idx{Z}$ and
+$\faml{g}(j) = \faml{f}_{i,j} \comp \inj_{\faml{Y}_j}: \faml{Y}_i \to \faml{Z}_{\idx{f}(i, j)}$ where
+$\inj_{\faml{Y}_j}: \faml{Y}_j \to \faml{X}_i \biprod \faml{Y}_j$
+\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\faml{J}}}(j \mapsto \faml{f}_{i, j} \comp
+\inj_{\faml{X}_i}: X_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i \biprod \faml{Y}_j$)
 \end{itemize}

### 2025-03-03 notes/sec/grothendieck.tex
Now I'm paying for the notation a bit...
Back to proof.#

@@
 $\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
-morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{f}$ in $\Fam(\idx{X},C)$.
+morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},C)$.
 
@@
 \item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X}, \idx{Y}]$
-denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to $\prod_{I}\reindex{Y}{\idx{f}}$.
+denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to
+$\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
@@
 \item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\faml{J}}}(j \mapsto \faml{f}_{i, j} \comp
-\inj_{\faml{X}_i}: X_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i \biprod \faml{Y}_j$)
+\inj_{\faml{X}_i}: \faml{X}_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i
+\biprod \faml{Y}_j$)
 \end{itemize}
@@
 \item $\idx{\eval}_{X,Y}((f, g), i) = f(i)$
-\item $\faml{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{I}\reindex{Y}{f} \biprod X_i \to Y_{f(i)}$
+\item $\faml{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{\idx{X}}\reindex{\faml{Y}}{f}
+\biprod \faml{X}_i \to \faml{Y}_{f(i)}$
 \end{itemize}

### 2025-03-03 notes/sec/grothendieck.tex
Need to be clearer about 'biproduct' of two morphisms..
Start on (family of) commutative triangles in C.
Fix up eval to new notation.

@@
 $\inj_{\faml{Y}_j}: \faml{Y}_j \to \faml{X}_i \biprod \faml{Y}_j$
-\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\faml{J}}}(j \mapsto \faml{f}_{i, j} \comp
+\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\idx{Y}}}(j \mapsto \faml{f}_{i, j} \comp
 \inj_{\faml{X}_i}: \faml{X}_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i
@@
 \begin{itemize}
-\item $\idx{\eval}_{X,Y}((f, g), i) = f(i)$
-\item $\faml{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{\idx{X}}\reindex{\faml{Y}}{f}
-\biprod \faml{X}_i \to \faml{Y}_{f(i)}$
+\item $\idx{\eval}_{X,Y}(f, i) = \idx{f}(i)$
+\item $\faml{\eval}_{X,Y}(f, i) = \coprodM{{\eval_{\prod}}_i}{\faml{f}_i}:
+\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_i \to \faml{Y}_{\idx{f}(i)}$
 \end{itemize}
@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have \todo{(prove these)}:
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have:
 \begin{enumerate}
-\item The following diagram commutes:
+\item To show that the diagram on the left in $\Fam(C)$ commutes, it suffices to show that the diagram on the
+right in $C$ commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:
 
@@
 \end{tikzcd}
+\hspace{1cm}
+\begin{tikzcd}
+   \prod_{\idx{Y}}\reindex{Z}{y \mapsto} \biprod \faml{Y}_y \arrow[dr, "\coprodM{\eval_{\prod_{\idx{Y}}}}{g_i}"] \\
+   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y\;\mapsto\;\faml{f}_{x, y}\;\comp\;
+\inj_{\faml{X}_x}) \biprod \id_{\faml{Y}_y}"] \arrow[r,
+   "\faml{f}_{x,y}"'] & \faml{Z}_{\idx{f}(x,y)}
+\end{tikzcd}
 \end{center}
 
-\item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$
+\item \todo{(exhibit $\lambda^{-1}$ and then show) $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$}
 \end{enumerate}

### 2025-03-03 notes/sec/grothendieck.tex
Fix up eval to new notation.
More consistent notation.

@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X}, \idx{Y}]$
-denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to
-$\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$.
+\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
+\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
@@
 \begin{tikzcd}
-   \prod_{\idx{Y}}\reindex{Z}{y \mapsto} \biprod \faml{Y}_y \arrow[dr, "\coprodM{\eval_{\prod_{\idx{Y}}}}{g_i}"] \\
+   \prod_{y \in \idx{Y}}\reindex{Z}{y \mapsto} \biprod \faml{Y}_y \arrow[dr, "\coprodM{\eval_{\prod_{\idx{Y}}}}{\faml{f}_i}"] \\
    \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y\;\mapsto\;\faml{f}_{x, y}\;\comp\;

### 2025-03-04 notes/sec/grothendieck.tex
More consistent notation.
Introduce Fam(I,C) first.

@@
 \subsection{Category of families}
-
-\begin{definition}[Category of $I$-indexed families of objects]
-For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
-\begin{itemize}
-\item objects are the $I$-indexed families of objects of $C$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
-\end{itemize}
-\end{definition}
-
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$, interpreting $I$ as a discrete
-category.
-
-\begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
-similarly for morphisms.
-\end{definition}
-
-\begin{definition}[$\Fam(-,C)$ functor]
-$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
-$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
-\end{definition}
 
@@
 \begin{tikzcd}
-   \prod_{y \in \idx{Y}}\reindex{Z}{y \mapsto} \biprod \faml{Y}_y \arrow[dr, "\coprodM{\eval_{\prod_{\idx{Y}}}}{\faml{f}_i}"] \\
-   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y\;\mapsto\;\faml{f}_{x, y}\;\comp\;
+   \prod_{\idx{Y}}\reindex{\faml{Z}}{y' \mapsto \idx{f}(x,y')} \biprod \faml{Y}_y \arrow[dr,
+   "\coprodM{\eval_{\prod_{\idx{Y}}}}{\faml{f}_x}"] \\
+   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y'\;\mapsto\;\faml{f}_{x, y'}\;\comp\;
 \inj_{\faml{X}_x}) \biprod \id_{\faml{Y}_y}"] \arrow[r,

### 2025-03-04 notes/sec/fam.tex
Spelling out the hom-object in Fam(C).
Hom-objects using the set-indexed product.
Defining eval, lambda.
Fam(I,C) inherits [co]limits from C.

--- 2025-02-25 notes/sec/fam.tex
@@
-\subsection{Category of families}
+\subsection{Category of $I$-indexed families}
 
-\begin{definition}[$\Fam(I,C)$]
+\begin{definition}[Category of $I$-indexed families of objects]
 For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
@@
 
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
+\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$, interpreting $I$ as a discrete
 category.
+
+\begin{definition}[Constant family]
+For any object $X$ of $C$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
+\end{definition}
 
@@
 
-\begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
-category where:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
-\to Y$ in $F(J)$.
-\end{itemize}
-\end{definition}
-
-\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
-paired with a morphism $X \to F(f)(Y)$ in $F(I)$.
-
-\begin{definition}[Category of families]
-For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
-\end{definition}
-
-In $\Fam(C)$ we then have that:
-\begin{itemize}
-\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
-\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
-$\Fam(I,C)$.
-\end{itemize}
+\begin{proposition}
+If $C$ has limits (resp.~colimits) then $\Fam(I,C)$ has limits (resp.~colimits) computed pointwise.
+\end{proposition}

### 2025-03-04 notes/sec/grothendieck.tex
Standardise x, y for indices instead of i, j.
Revert to point-free for reindexing functions.
And express biproduct injections as morphisms in Fam(I,C) rather than families of morphisms in C.
Dang, no that doesn't work.

@@
 \begin{itemize}
-\item $\idx{\lambda}_{X,Y,Z}(f)(i) = g$ where $\idx{g}(j) = \idx{f}(i, j): \idx{Y} \to \idx{Z}$ and
-$\faml{g}(j) = \faml{f}_{i,j} \comp \inj_{\faml{Y}_j}: \faml{Y}_i \to \faml{Z}_{\idx{f}(i, j)}$ where
-$\inj_{\faml{Y}_j}: \faml{Y}_j \to \faml{X}_i \biprod \faml{Y}_j$
-\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\idx{Y}}}(j \mapsto \faml{f}_{i, j} \comp
-\inj_{\faml{X}_i}: \faml{X}_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i
-\biprod \faml{Y}_j$)
+\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
+$\faml{g} = \faml{f} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
+\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\faml{f} \comp \inj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f}$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
+\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$)
 \end{itemize}
@@
 \begin{itemize}
-\item $\idx{\eval}_{X,Y}(f, i) = \idx{f}(i)$
-\item $\faml{\eval}_{X,Y}(f, i) = \coprodM{{\eval_{\prod}}_i}{\faml{f}_i}:
-\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_i \to \faml{Y}_{\idx{f}(i)}$
+\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
+\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
+\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
 \end{itemize}
@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have:
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have the following.
 \begin{enumerate}
-\item To show that the diagram on the left in $\Fam(C)$ commutes, it suffices to show that the diagram on the
-right in $C$ commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:
+\item The diagram on the left in $\Fam(C)$ commutes; it suffices to show that the diagram on the right in $C$
+commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:
 

### 2025-03-08 notes/sec/biproduct.tex
Just enough notation to state proposition that copairing distributes over parallel composition.
Fixing up right-hand commutative diagram.
More general statement of prod/coprod lemma.

@@
 \begin{salign*}
-\inj_A \comp \proj_A + \inj_B \comp \proj_B &= \id_{A \proj B}
+\inj_A \comp \proj_A + \inj_B \comp \proj_B &= \id_{A \biprod B}
 \end{salign*}
 \end{definition}
-
-If $C$ has biproducts then it has products given by $(A \biprod B, \proj_A, \proj_B)$ and coproducts given by $(A
-\biprod B, \inj_A, \inj_B)$.
 
@@
 \begin{itemize}
+\item If $C$ has biproducts then it has products given by $(A \biprod B, \proj_A, \proj_B)$ and coproducts
+given by $(A \biprod B, \inj_A, \inj_B)$.
 \item If $C$ has products (or coproducts) then they are biproducts.
@@
 \noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.
+
+\begin{proposition}
+\label{prop:biproduct:prod-coprod}
+Suppose $C$ a category with biproducts. Then the following diagram in $C$ commutes for any $f_1, g_1, f_2,
+g_2$:
+
+\begin{center}
+\begin{tikzcd}[row sep=3em]
+   & A' \\
+   A \arrow[ur, "f_1"] \arrow[r, "\inj_A", shift left] &
+   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] \arrow[u, pos=0.3, "\coprodM{f_1}{g_1}"'] &
+   B \arrow[l, "\inj_B"', shift right] \arrow[ul, "g_1"'] \\
+   & B' \arrow[ul, "f_2"] \arrow[u, pos=0.6, "\prodM{f_2}{g_2}"'] \arrow[ur, "g_2"']
+\end{tikzcd}
+\end{center}
+\end{proposition}

### 2025-03-08 notes/sec/grothendieck.tex
Ok can make it work if I can precompose family with (pairing with x).
Type index on eval is a bit painful, but keep it for consistency for now..
Closer.
Clarify f as morphism from constant family.
'precomposition with reindexing function' operation on families of morphism is better understood as the action of the functor induced by the reindexing function.
Slightly more forgiving indexing on eval.
Fix index.
Just enough notation to state proposition that copairing distributes over parallel composition.
Fixing up right-hand commutative diagram.
Corrections.
Correction. Finally getting there.
Ba

@@
 \item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
-$\faml{g} = \faml{f} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
+$\faml{g} = \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
 \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
-\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\faml{f} \comp \inj_{\faml{X}_x}:
-\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f}$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
-\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$)
+\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{\idx{f}})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
+\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
 \end{itemize}
@@
 \item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
-\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
+\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
 \end{itemize}
@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have the following.
+Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
 \begin{enumerate}
-\item The diagram on the left in $\Fam(C)$ commutes; it suffices to show that the diagram on the right in $C$
-commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:
+\item Suppose $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property
+of the set-indexed product; the commutativity of the diagram on the right is trivial.
 
@@
 \begin{tikzcd}
-   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
-   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
+   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \arrow[dr, "\eval_{\prod_{y}}"] \\
+   \faml{X}_x \arrow[u, "\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
+\inj_{\faml{X}_x})"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}"'] & \faml{Z}_{\idx{f}(x,y)}
+\end{tikzcd}
+\hspace{1cm}
+\begin{tikzcd}[column sep=5em]
+   \faml{Y}_y \arrow[dr,
+   "\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}"] \\
+   \faml{Y}_y \arrow[u, -, double, "\id_{\faml{Y}_y}"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}"'] & \faml{Z}_{\idx{f}(x,y)}
+\end{tikzcd}
+\end{center}
+
+The diagram on the left below then commutes by Proposition~\ref{prop:biproduct:prod-coprod}, with
+$\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}} =
+\faml{f}_{(x,y)}$ by the uniqueness of the coproduct of morphisms. Then the diagram on the right in $\Fam(C)$
+commutes by the definitions of $\lambda_{X,Y,Z}, \eval_{Y,Z}$ and $[Y,Z]$.
+
+\begin{center}
+\begin{tikzcd}[column sep=5.5em]
+   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \biprod \faml{Y}_y \arrow[dr,
+   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}}"] \\
+   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
+\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r, "\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] & \faml{Z}_{\idx{f}(x,y)}
 \end{tikzcd}
@@
 \begin{tikzcd}
-   \prod_{\idx{Y}}\reindex{\faml{Z}}{y' \mapsto \idx{f}(x,y')} \biprod \faml{Y}_y \arrow[dr,
-   "\coprodM{\eval_{\prod_{\idx{Y}}}}{\faml{f}_x}"] \\
-   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y'\;\mapsto\;\faml{f}_{x, y'}\;\comp\;
-\inj_{\faml{X}_x}) \biprod \id_{\faml{Y}_y}"] \arrow[r,
-   "\faml{f}_{x,y}"'] & \faml{Z}_{\idx{f}(x,y)}
+   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
+   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
 \end{tikzcd}

### 2025-03-09 notes/sec/biproduct.tex
More general statement of prod/coprod lemma.
Reorient prod/coprod diagram.

@@
 \label{prop:biproduct:prod-coprod}
-Suppose $C$ a category with biproducts. Then the following diagram in $C$ commutes for any $f_1, g_1, f_2,
-g_2$:
+Suppose $C$ a category with biproducts. Then the following commutes for any morphisms $f_1, g_1, f_2,
+g_2$ in $C$:
 
 \begin{center}
-\begin{tikzcd}[row sep=3em]
-   & A' \\
-   A \arrow[ur, "f_1"] \arrow[r, "\inj_A", shift left] &
-   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] \arrow[u, pos=0.3, "\coprodM{f_1}{g_1}"'] &
-   B \arrow[l, "\inj_B"', shift right] \arrow[ul, "g_1"'] \\
-   & B' \arrow[ul, "f_2"] \arrow[u, pos=0.6, "\prodM{f_2}{g_2}"'] \arrow[ur, "g_2"']
+\begin{tikzcd}[row sep=3em, column sep=3em]
+   & A \arrow[d, "\inj_A", shift left] \arrow[dr, "f_2"] \\
+   A' \arrow[r, "\prodM{f_1}{g_1}"] \arrow[dr, "g_1"'] \arrow[ur, "f_1"] & A \biprod B \arrow[u, "\proj_A", shift left] \arrow[d, "\proj_B"', shift right] \arrow[r, pos=0.35, "\coprodM{f_2}{g_2}"] & B' \\
+   & B \arrow[u, "\inj_B"', shift right] \arrow[ur, "g_2"']
 \end{tikzcd}

### 2025-03-10 notes/sec/grothendieck.tex
Correction to isomorphism property.
Indexing function for lambda inv.
Indexing component of lambda is an iso by Cartesian closure of Set.

@@
 \end{enumerate}
-Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
+
+Then:
+
 \begin{enumerate}
-\item Suppose $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property
-of the set-indexed product; the commutativity of the diagram on the right is trivial.
+\item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism by the ismorphism of $\idx{X} \times \idx{Y} \to \idx{Z}$ and
+$\idx{X} \to \idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the family of functions
+$\lambda^{-1}_{X,Y,Z}$ sending any $f: X \to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f): X \times Y \to Z$ with:
+\begin{itemize}
+\item $\idx{\lambda^{-1}}_{X,Y,Z}(f)(x,y) = \idx{g}(y)$ where $\idx{f}(x) = g$
+\item $\faml{\lambda^{-1}}_{X,Y,Z}(f)(x,y) =$
+\end{itemize}
+
+\todo{Show $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$ and $\lambda_{X,Y,Z}(\lambda^{-1}_{X,Y,Z}(f)) =
+f$}.
+
+\item Suppose $f: X \times Y \to Z$ in $\Fam(C)$ and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the
+left below commutes by the universal property of the set-indexed product; the commutativity of the diagram on
+the right is trivial.
 
@@
    \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
-\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r, "\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] & \faml{Z}_{\idx{f}(x,y)}
+\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r,
+"\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] &
+\faml{Z}_{\idx{f}(x,y)}
 \end{tikzcd}
@@
 \end{center}
-
-\item \todo{(exhibit $\lambda^{-1}$ and then show) $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$}
 \end{enumerate}

### 2025-03-10 notes/sec/grothendieck.tex
Indexing component of lambda is an iso by Cartesian closure of Set.
Inline g.

@@
 \begin{itemize}
-\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
-$\faml{g} = \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
-\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\item $\idx{\lambda}_{X,Y,Z}(f)(x) = (\idx{f} \comp (x, -), \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}):
+Y \to Z$ where $\inj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
 \item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
@@
 \begin{enumerate}
-\item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism by the ismorphism of $\idx{X} \times \idx{Y} \to \idx{Z}$ and
-$\idx{X} \to \idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the family of functions
-$\lambda^{-1}_{X,Y,Z}$ sending any $f: X \to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f): X \times Y \to Z$ with:
+\item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism because $\idx{X} \times \idx{Y} \to \idx{Z} \iso \idx{X} \to
+\idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the following family of functions sending any $f: X
+\to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f): X \times Y \to Z$ with:
 \begin{itemize}
-\item $\idx{\lambda^{-1}}_{X,Y,Z}(f)(x,y) = \idx{g}(y)$ where $\idx{f}(x) = g$
-\item $\faml{\lambda^{-1}}_{X,Y,Z}(f)(x,y) =$
+\item $\faml{\lambda^{-1}}_{X,Y,Z}(f)(x,y) = \faml{g}_y$ where $\idx{f}(x) = g$
 \end{itemize}

### 2025-03-12 notes/sec/grothendieck.tex
Cite nunes2023.
Simplify.
Some improvement to idx/fam notation but still heavyweight.
Notation faff.

@@
 
+The following is a special case of a construction due to \citet{nunes2023}:
+
 \begin{proposition}
 Suppose $C$ locally small. If $C$ has binary biproducts and set-indexed products, then $\Fam(C)$ is Cartesian
-closed. \todo{Establish first that if $C$ has binary coproducts and set-indexed products, then $\Fam(C)$ is
-symmetric monoidal, with the coproduct as monoidal product.}
+closed. \todo{Establish first that if $C$ has binary coproducts and set-indexed products,
+then $\Fam(C)$ is symmetric monoidal, with the coproduct as monoidal product.}
 \end{proposition}
@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
+\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_f =
 \prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
-\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
-$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
+\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $(f, g): X \times Y \to Z$ to
+$(f', g'): X \to [Y, Z]$ with:
 \begin{itemize}
-\item $\idx{\lambda}_{X,Y,Z}(f)(x) = (\idx{f} \comp (x, -), \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}):
+\item $f'(x) = (f \comp (x, -), \reindex{g}{(x,-)} \comp \inj_{\faml{Y}}):
 Y \to Z$ where $\inj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
-\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
-\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{\idx{f}})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
+\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g}{(x,-)} \comp \inj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
 \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
 \end{itemize}
-\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
+\item The family of morphisms $\eval_{X,Y} = (f',g'): [X, Y] \times X \to Y$ in $\Fam(C)$ with
 \begin{itemize}
-\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
-\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
-\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
+\item $f'((f, g), x) = f(x)$
+\item $g'_{(f,g), x} = \coprodM{{\eval_{\prod}}_x}{g_x}:
+\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_x \to \faml{Y}_{f(x)}$
 \end{itemize}
-using $\coprodM{-}{-}$ to denote coproduct of morphisms.
 \end{enumerate}
@@
 \item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism because $\idx{X} \times \idx{Y} \to \idx{Z} \iso \idx{X} \to
-\idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the following family of functions sending any $f: X
-\to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f): X \times Y \to Z$ with:
+\idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the following family of functions sending any
+$(f,g): X \to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f,g) = (f^\dagger, g^\dagger): X \times Y \to Z$ with:
 \begin{itemize}
-\item $\faml{\lambda^{-1}}_{X,Y,Z}(f)(x,y) = \faml{g}_y$ where $\idx{f}(x) = g$
+\item $f^\dagger(x,y) = f'(y)$ where $f(x) = (f',g')$
+\item $g^\dagger_{x,y} = g'_y \comp \proj_{\faml{Y}_y}$ where $f(x) = (f',g')$ and
+$\proj_{\faml{Y}_y}: \faml{X}_x \biprod \faml{Y}_y \to \faml{Y}_y$
 \end{itemize}

### 2025-03-13 notes/sec/grothendieck.tex
Think lambda-inv .famf needs to use both of the available .famf functions, combining them with copair.
Put new definition of lambda-inv in LaTeX.

@@
 \begin{enumerate}
-\item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism because $\idx{X} \times \idx{Y} \to \idx{Z} \iso \idx{X} \to
-\idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the following family of functions sending any
-$(f,g): X \to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f,g) = (f^\dagger, g^\dagger): X \times Y \to Z$ with:
+\item Define the family of functions $\lambda^{-1}_{X,Y,Z}$ sending any $(f,g): X \to [Y,Z]$ to $(f^\dagger,
+g^\dagger): X \times Y \to Z$ with:
 \begin{itemize}
-\item $f^\dagger(x,y) = f'(y)$ where $f(x) = (f',g')$
-\item $g^\dagger_{x,y} = g'_y \comp \proj_{\faml{Y}_y}$ where $f(x) = (f',g')$ and
-$\proj_{\faml{Y}_y}: \faml{X}_x \biprod \faml{Y}_y \to \faml{Y}_y$
+\item $f^\dagger(x,y) = f'(y)$
+\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{g'_y}$
+% $g'_y \comp \proj_{\faml{Y}_y}$ where $f(x) = (f',g')$ and $\proj_{\faml{Y}_y}: \faml{X}_x \biprod
+% \faml{Y}_y \to \faml{Y}_y$
 \end{itemize}
+\noindent where $f(x) = (f',g')$.
 

### 2025-03-14 notes.tex
We need that Fam(-) preserves products.
Start on set-indexed products.
Introduce Fam(I,C) first.
Just enough notation to state proposition that copairing distributes over parallel composition.
secrefs.

@@
 
-Some of the key constructions:
+Notions covered so far:
 \begin{itemize}
-\item Category of families $\Fam(C)$ for any category $C$
-\item Category of $\LatGal$ of bounded lattices and Galois connections
+\item $\CMon$-enriched categories (\secref{cmon-enriched})
+\item biproducts (\secref{biproduct})
+\item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
+\item set-indexed products (\secref{set-indexed-product})
+\item Grothendieck construction and category of families $\Fam(C)$ (\secref{grothendieck})
+\item category of $\LatGal$ of bounded lattices and Galois connections (\secref{galois})
 \end{itemize}
 
-% Is one of the goals to define sufficient conditions on $C$ for $\Fam(C)$ to be Cartesian closed?
+\noindent $\Set$ will usually be $\Setoid$ in the Agda implementation but we will gloss that detail for now.
 
-\noindent $\Set$ will often be $\Setoid$ in the Agda implementation but we will gloss that detail for now.
+\input{notes/sec/notation}
 
@@
 \input{notes/sec/fam}
+\input{notes/sec/set-indexed-product}
+\input{notes/sec/grothendieck}
 \input{notes/sec/galois}

### 2025-03-14 notes/sec/grothendieck.tex
Put new definition of lambda-inv in LaTeX.
Progress on lambda-inv/natural.
secrefs.

@@
 \subsection{Category of families}
+\label{sec:grothendieck}
 
@@
 \noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
-$D$ paired with morphisms $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
+$D$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
 
@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_f =
-\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
+\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_{f,g} =
+\prod_{\idx{X}}\reindex{\faml{Y}}{f}$ for every $(f,g): X \to Y$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $(f, g): X \times Y \to Z$ to
@@
 \begin{itemize}
-\item $f^\dagger(x,y) = f'(y)$
-\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{g'_y}$
-% $g'_y \comp \proj_{\faml{Y}_y}$ where $f(x) = (f',g')$ and $\proj_{\faml{Y}_y}: \faml{X}_x \biprod
-% \faml{Y}_y \to \faml{Y}_y$
+\item $f^\dagger(x,y) = f'(y)$ where $f(x) = (f',g')$
+\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{g'_y}$ where $f(x) = (f',g')$
 \end{itemize}
-\noindent where $f(x) = (f',g')$.
 
-\todo{Show $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$ and $\lambda_{X,Y,Z}(\lambda^{-1}_{X,Y,Z}(f)) =
-f$}.
+\todo{
+\begin{itemize}
+\item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f,g)) = (f,g)$ for any $(f,g): X \times Y \to Z$
+\item $\lambda_{X,Y,Z}(\lambda^{-1}_{X,Y,Z}(f,g)) = (f,g)$ for any $(f,g): X \to [Y,Z]$
+\end{itemize}
+}
 

### 2025-03-25 notes/sec/grothendieck.tex
Start on other triangle identity.
Start on indexing component.
Indexing part.
Onto family morphism part.
Family morphism part.
Progress on g'.

@@
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_{f,g} =
-\prod_{\idx{X}}\reindex{\faml{Y}}{f}$ for every $(f,g): X \to Y$.
+\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_f =
+\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
 \item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $(f, g): X \times Y \to Z$ to
@@
 \begin{itemize}
-\item $f'((f, g), x) = f(x)$
-\item $g'_{(f,g), x} = \coprodM{{\eval_{\prod}}_x}{g_x}:
-\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_x \to \faml{Y}_{f(x)}$
+\item $f'(f, x) = \idx{f}(x)$
+\item $g'_{f, x} = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
+\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
 \end{itemize}
@@
 Then:
-
 \begin{enumerate}
-\item Define the family of functions $\lambda^{-1}_{X,Y,Z}$ sending any $(f,g): X \to [Y,Z]$ to $(f^\dagger,
-g^\dagger): X \times Y \to Z$ with:
+\item Suppose $(f,g): X \to [Y,Z]$ and define $(f^\dagger, g^\dagger): X \times Y \to Z$ with:
 \begin{itemize}
-\item $f^\dagger(x,y) = f'(y)$ where $f(x) = (f',g')$
-\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{g'_y}$ where $f(x) = (f',g')$
+\item $f^\dagger(x,y) = \idx{f(x)}(y)$
+\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y}$
 \end{itemize}
 
-\todo{
+To show $\lambda_{X,Y,Z}(f^\dagger, g^\dagger) = (f,g)$, suppose $\lambda_{X,Y,Z}(f^\dagger, g^\dagger) = (f',
+g')$ with
 \begin{itemize}
-\item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f,g)) = (f,g)$ for any $(f,g): X \times Y \to Z$
-\item $\lambda_{X,Y,Z}(\lambda^{-1}_{X,Y,Z}(f,g)) = (f,g)$ for any $(f,g): X \to [Y,Z]$
+\item $f'(x) = (f^\dagger \comp (x, -), \reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}}): Y \to Z$ where
+$\inj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f^\dagger})$ where $\inj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
 \end{itemize}
-}
+\noindent To show $f' = f : \idx{X} \to \Fam(C)(Y,Z)$, suppose $x \in \idx{X}$. To show $f'(x) = f(x): Y \to
+Z$, suppose $y \in \idx{Y}$.
+\begin{enumerate}
+\item $(f^\dagger \comp (x, -))(y) = f^\dagger(x,y) = \idx{f(x)}(y)$.
+\item $(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}})_y = \reindex{g^\dagger}{(x,-)}_y \comp
+{\inj_{\faml{Y}}}_y = g^\dagger_{x,y} \comp {\inj_{\faml{Y}}}_y = \coprodM{{\eval_{\prod}}_y \comp
+g_x}{\faml{f}_y} \comp {\inj_{\faml{Y}}}_y = g_x$.
+\end{enumerate}
+
+To show $g' = g$, suppose $x \in \idx{X}$. To show $g'_x = g_x$:
+\begin{align*}
+\small
+g'_x
+&=
+\lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{X}_x}) \\
+&=
+\lambda_{\prod_{\idx{Y}}}(\{g^\dagger_{(x,y)} \comp (\inj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
+&=
+\lambda_{\prod_{\idx{Y}}}(\{\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y} \comp (\inj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
+&=
+\lambda_{\prod_{\idx{Y}}}(\{\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y}\}_{y \in \idx{Y}})
+\end{align*}
 

### 2025-03-26 notes/sec/grothendieck.tex
Done with g' case, just the indexed family part of f' to grok now.
Correction.
First version.
Clarify triangle identities.
C -> D.

@@
 \begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
+Suppose a functor $F: C \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{C}F$ for $F$ is the
 category where:
 \begin{itemize}
-\item objects $X$ are pairs of an object $\idx{X}$ of $D$ and an object $\faml{X}$ of $F(\idx{X})$;
-\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $D$ paired with morphisms
+\item objects $X$ are pairs of an object $\idx{X}$ of $C$ and an object $\faml{X}$ of $F(\idx{X})$;
+\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $C$ paired with morphisms
 $\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
@@
 \noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
-$D$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
+$C$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
 
@@
 \end{itemize}
-\item The family of morphisms $\eval_{X,Y} = (f',g'): [X, Y] \times X \to Y$ in $\Fam(C)$ with
+\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
 \begin{itemize}
-\item $f'(f, x) = \idx{f}(x)$
-\item $g'_{f, x} = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
+\item $\idx{\eval_{X,Y}}(f, x) = \idx{f}(x)$
+\item $\faml{\eval_{X,Y}}_{f, x} = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
 \prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
@@
 
-Then:
+Then the following triangle identities hold:
 \begin{enumerate}
-\item Suppose $(f,g): X \to [Y,Z]$ and define $(f^\dagger, g^\dagger): X \times Y \to Z$ with:
+\item $\lambda_{X,Y,Z}(\eval_{Y,Z} \comp (f \tensor \id_Y) = f$. Suppose $(f,g): X \to [Y,Z]$ and define
+$(f^\dagger, g^\dagger): X \times Y \to Z$ with:
 \begin{itemize}
@@
 To show $\lambda_{X,Y,Z}(f^\dagger, g^\dagger) = (f,g)$, suppose $\lambda_{X,Y,Z}(f^\dagger, g^\dagger) = (f',
-g')$ with
+g')$. Then:
 \begin{itemize}
@@
 \const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f^\dagger})$ where $\inj_{\faml{X}_x}:
-\const_{\idx{Y}}(\faml{X}_x) \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\const_{\idx{Y}}(\faml{X}_x) \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$.
 \end{itemize}
-\noindent To show $f' = f : \idx{X} \to \Fam(C)(Y,Z)$, suppose $x \in \idx{X}$. To show $f'(x) = f(x): Y \to
-Z$, suppose $y \in \idx{Y}$.
+\noindent To show $f' = f : \idx{X} \to \Fam(C)(Y,Z)$, suppose $x \in \idx{X}$ and show $f'(x) = f(x): Y \to
+Z$. Suppose $y \in \idx{Y}$.
 \begin{enumerate}
-\item $(f^\dagger \comp (x, -))(y) = f^\dagger(x,y) = \idx{f(x)}(y)$.
-\item $(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}})_y = \reindex{g^\dagger}{(x,-)}_y \comp
+\item $\idx{f'(x)}(y) = (f^\dagger \comp (x, -))(y) = f^\dagger(x,y) = \idx{f(x)}(y)$.
+\item $\faml{f'(x)}_y = (\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}})_y = \reindex{g^\dagger}{(x,-)}_y \comp
 {\inj_{\faml{Y}}}_y = g^\dagger_{x,y} \comp {\inj_{\faml{Y}}}_y = \coprodM{{\eval_{\prod}}_y \comp
-g_x}{\faml{f}_y} \comp {\inj_{\faml{Y}}}_y = g_x$.
+g_x}{\faml{f(x)}_y} \comp {\inj_{\faml{Y}}}_y = \faml{f(x)}_y$.
 \end{enumerate}
 
-To show $g' = g$, suppose $x \in \idx{X}$. To show $g'_x = g_x$:
+To show $g' = g$, suppose $x \in \idx{X}$.
 \begin{align*}
@@
 &=
-\lambda_{\prod_{\idx{Y}}}(\{\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y}\}_{y \in \idx{Y}})
+\lambda_{\prod_{\idx{Y}}}(\{{\eval_{\prod}}_y \comp g_x\}_{y \in \idx{Y}}) \\
+&=
+g_x
 \end{align*}
+\qed
 
-\item Suppose $f: X \times Y \to Z$ in $\Fam(C)$ and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the
-left below commutes by the universal property of the set-indexed product; the commutativity of the diagram on
-the right is trivial.
+\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f) \tensor \id_{Y}) = f$. Suppose $f: X \times Y \to Z$ in $\Fam(C)$
+and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property of
+the set-indexed product; the commutativity of the diagram on the right is trivial.
 

### 2025-03-26 notes.tex
C -> D.
Placeholder for FdVect.

@@
 
-\title{Approximation as Differentiation notes}
+\title{Approximation as Differentiation}
 \maketitle
@@
 
-Notions covered so far:
+Covered so far:
 \begin{itemize}
@@
 \item set-indexed products (\secref{set-indexed-product})
-\item Grothendieck construction and category of families $\Fam(C)$ (\secref{grothendieck})
-\item category of $\LatGal$ of bounded lattices and Galois connections (\secref{galois})
+\item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
+\item Examples of useful categories with biproducts:
+   \begin{itemize}
+   \item category $\LatGal$ of bounded lattices and Galois connections (\secref{categories-with-biproducts:latgal})
+   \item category $\FdVect_k$ of finite-dimensional vector spaces over a field $k$ (\secref{categories-with-biproducts:fdvect})
+   \end{itemize}
 \end{itemize}

### 2025-03-26 notes/sec/galois.tex
Assume C locally small.
secrefs.
Placeholder for FdVect.

@@
-\subsection{Category of bounded lattices and Galois connections}
+\subsection{Useful categories with biproducts}
+\label{sec:categories-with-biproducts}
+
+\subsubsection{Category of bounded lattices and Galois connections}
+\label{sec:categories-with-biproducts:latgal}
 
@@
 
-\subsection{Category $\Func{\LatGal^\op}{\CMon}$}
+\subsubsection{Category of finite-dimensional vector spaces}
+\label{sec:categories-with-biproducts:fdvect}
 
-A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
-every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
-$\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.
+% \subsubsection{$\CMon$-valued presheaves over $\LatGal$}
+%
+% A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
+% every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
+% $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Rename.

--- 2025-03-26 notes/sec/galois.tex
@@
 \begin{itemize}
-\item objects $L$ are the bounded lattices;
-\item morphisms $f \vdash g: L \to L'$ are meet-semilattice homomorphisms $f: L \to L'$ paired with
-join-semilattice homomorphisms $g: L' \to L$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in L$ and any
-$y \in L'$.
+\item objects $X, Y$ are the bounded lattices;
+\item morphisms $f \vdash g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
+join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
+$y \in Y$.
 \end{itemize}

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Enough setup to give the CMon-enrichment instance for LatGal.

@@
 
-Define the category $\LatGal$ where:
+\begin{definition}[Bounded lattice]
+A \emph{bounded lattice} $X = (X, \meet, \top, \join, \bot)$ is a partial order $X$ equipped with binary
+operations $\meet$ and $\join$ and distinguished elements $\top$ and $\bot$ such that $(X, \meet, \top)$ is a
+bounded meet-semilattice and $(X, \join, \bot)$ is a bounded join-semilattice.
+\end{definition}
+
+\noindent Since $(X, \meet, \top)$ is a commutative monoid, and thus a $\CMon$-enriched poset category, for
+any $X, Y$ there is a zero meet-semilattice homomorphism $0_{X,Y}: X \to \top_Y$ and addition of homomorphisms
+$X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join, \bot)$. Now
+define the category $\LatGal$ where:
 \begin{itemize}
@@
 
-\noindent $\LatGal$ has finite biproducts.
+\noindent $\LatGal$ is itself $\CMon$-enriched; for any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative
+monoid, with:
+\begin{itemize}
+\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \vdash 0_{Y,X}: X \to Y$
+\end{itemize}
+
+\noindent And moreover has finite biproducts:
 

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Enough setup to give the CMon-enrichment instance for LatGal.
To get to biproducts need to go via products.

@@
 any $X, Y$ there is a zero meet-semilattice homomorphism $0_{X,Y}: X \to \top_Y$ and addition of homomorphisms
-$X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join, \bot)$. Now
-define the category $\LatGal$ where:
+$f + g: X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join,
+\bot)$. Now define the category $\LatGal$ where:
 \begin{itemize}
@@
 
-\noindent $\LatGal$ is itself $\CMon$-enriched; for any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative
-monoid, with:
+\noindent and composition is component-wise.
+
+$\LatGal$ is itself $\CMon$-enriched. For any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative monoid,
+where:
 \begin{itemize}
 \item $0_{X,Y}$ is the Galois connection $0_{X,Y} \vdash 0_{Y,X}: X \to Y$
+\item $(f \vdash g) + (f' \vdash g') = (f + f') \vdash (g + g'): X \to Y$.
 \end{itemize}
 
-\noindent And moreover has finite biproducts:
+\noindent and biadditivity of composition arising component-wise. Moreover $\LatGal$ has binary products, with
+the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts:
 

### 2025-03-31 notes/sec/biproduct.tex
secrefs.
Or coproducts.

--- 2025-02-25 notes/sec/biproduct.tex
@@
 \subsection{Biproducts}
+\label{sec:biproduct}
 
@@
 \begin{tikzcd}
-   A \arrow[r, "\inj_A", shift left] &
-   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] &
-   B \arrow[l, "\inj_B"', shift right]
+   A \arrow[r, "\biinj_A", shift left] &
+   A \biprod B \arrow[l, "\biproj_A", shift left] \arrow[r, "\biproj_B"', shift right] &
+   B \arrow[l, "\biinj_B"', shift right]
 \end{tikzcd}
@@
 \begin{salign*}
-   \proj_A \comp \inj_A &= \id_A \\
-   \proj_B \comp \inj_A &= \zero_{A,B}
+   \biproj_A \comp \biinj_A &= \id_A \\
+   \biproj_B \comp \biinj_A &= \zero_{A,B}
 \end{salign*}
@@
 \begin{salign*}
-   \proj_B \comp \inj_B &= \id_B \\
-   \proj_A \comp \inj_B &= \zero_{B,A}
+   \biproj_B \comp \biinj_B &= \id_B \\
+   \biproj_A \comp \biinj_B &= \zero_{B,A}
 \end{salign*}
@@
 \begin{salign*}
-\inj_A \comp \proj_A + \inj_B \comp \proj_B &= \id_{A \proj B}
+\biinj_A \comp \biproj_A + \biinj_B \comp \biproj_B &= \id_{A \biprod B}
 \end{salign*}
 \end{definition}
-
-If $C$ has biproducts then it has products given by $(A \biprod B, \proj_A, \proj_B)$ and coproducts given by $(A
-\biprod B, \inj_A, \inj_B)$.
 
@@
 \begin{itemize}
+\item If $C$ has biproducts then it has products given by $(A \biprod B, \biproj_A, \biproj_B)$ and coproducts
+given by $(A \biprod B, \biinj_A, \biinj_B)$.
 \item If $C$ has products (or coproducts) then they are biproducts.
@@
 
-\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.
+\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts. Having products or coproducts
+is sufficient for semi-additivity:
+
+\begin{proposition}
+Suppose $C$ is $\CMon$-enriched.
+\begin{itemize}
+\item If $C$ has products given by $(A \times B, \proj_1, \proj_2)$ then:
+\begin{enumerate}
+\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
+\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1 + f_2}{g_1 + g_2}$
+\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
+biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
+\prodM{0_{A,B}}{\id_B}$.
+\end{enumerate}
+\item If $C$ has coproducts given by $(A + B, \inj_1, \inj_2)$ then
+\begin{enumerate}
+\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
+biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
+\coprodM{0_{A,B}}{\id_B}$.
+\end{enumerate}
+\end{itemize}
+\end{proposition}
+
+\begin{proposition}
+\label{prop:biproduct:prod-coprod}
+Suppose $C$ a category with biproducts. Then the following commutes for any morphisms $f_1, g_1, f_2,
+g_2$ in $C$:
+
+\begin{center}
+\begin{tikzcd}[row sep=3em, column sep=3em]
+   & A \arrow[d, "\biinj_A", shift left] \arrow[dr, "f_2"] \\
+   A' \arrow[r, "\prodM{f_1}{g_1}"] \arrow[dr, "g_1"'] \arrow[ur, "f_1"] & A \biprod B \arrow[u, "\biproj_A", shift left] \arrow[d, "\biproj_B"', shift right] \arrow[r, pos=0.35, "\coprodM{f_2}{g_2}"] & B' \\
+   & B \arrow[u, "\biinj_B"', shift right] \arrow[ur, "g_2"']
+\end{tikzcd}
+\end{center}
+\end{proposition}

### 2025-03-31 notes/sec/grothendieck.tex
Placeholder for FdVect.
Or coproducts.

@@
 Suppose $C$ locally small. If $C$ has binary biproducts and set-indexed products, then $\Fam(C)$ is Cartesian
-closed. \todo{Establish first that if $C$ has binary coproducts and set-indexed products,
-then $\Fam(C)$ is symmetric monoidal, with the coproduct as monoidal product.}
+closed.
 \end{proposition}
@@
 \begin{proof}
-Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\inj,\proj)$ and set-indexed products
+Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\biinj,\biproj)$ and set-indexed products
 given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X, Y, Z$ in $\Fam(C)$ define:
@@
 \begin{itemize}
-\item $f'(x) = (f \comp (x, -), \reindex{g}{(x,-)} \comp \inj_{\faml{Y}}):
-Y \to Z$ where $\inj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
-\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g}{(x,-)} \comp \inj_{\faml{X}_x}:
-\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
+\item $f'(x) = (f \comp (x, -), \reindex{g}{(x,-)} \comp \biinj_{\faml{Y}}):
+Y \to Z$ where $\biinj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g}{(x,-)} \comp \biinj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f})$ where $\biinj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
 \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
@@
 \begin{itemize}
-\item $f'(x) = (f^\dagger \comp (x, -), \reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}}): Y \to Z$ where
-$\inj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
-\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{X}_x}:
-\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f^\dagger})$ where $\inj_{\faml{X}_x}:
+\item $f'(x) = (f^\dagger \comp (x, -), \reindex{g^\dagger}{(x,-)} \comp \biinj_{\faml{Y}}): Y \to Z$ where
+$\biinj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
+\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \biinj_{\faml{X}_x}:
+\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f^\dagger})$ where $\biinj_{\faml{X}_x}:
 \const_{\idx{Y}}(\faml{X}_x) \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$.
@@
 \item $\idx{f'(x)}(y) = (f^\dagger \comp (x, -))(y) = f^\dagger(x,y) = \idx{f(x)}(y)$.
-\item $\faml{f'(x)}_y = (\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{Y}})_y = \reindex{g^\dagger}{(x,-)}_y \comp
-{\inj_{\faml{Y}}}_y = g^\dagger_{x,y} \comp {\inj_{\faml{Y}}}_y = \coprodM{{\eval_{\prod}}_y \comp
-g_x}{\faml{f(x)}_y} \comp {\inj_{\faml{Y}}}_y = \faml{f(x)}_y$.
+\item $\faml{f'(x)}_y = (\reindex{g^\dagger}{(x,-)} \comp \biinj_{\faml{Y}})_y = \reindex{g^\dagger}{(x,-)}_y \comp
+{\biinj_{\faml{Y}}}_y = g^\dagger_{x,y} \comp {\biinj_{\faml{Y}}}_y = \coprodM{{\eval_{\prod}}_y \comp
+g_x}{\faml{f(x)}_y} \comp {\biinj_{\faml{Y}}}_y = \faml{f(x)}_y$.
 \end{enumerate}
@@
 &=
-\lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \inj_{\faml{X}_x}) \\
+\lambda_{\prod_{\idx{Y}}}(\reindex{g^\dagger}{(x,-)} \comp \biinj_{\faml{X}_x}) \\
 &=
-\lambda_{\prod_{\idx{Y}}}(\{g^\dagger_{(x,y)} \comp (\inj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
+\lambda_{\prod_{\idx{Y}}}(\{g^\dagger_{(x,y)} \comp (\biinj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
 &=
-\lambda_{\prod_{\idx{Y}}}(\{\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y} \comp (\inj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
+\lambda_{\prod_{\idx{Y}}}(\{\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y} \comp (\biinj_{\faml{X}_x})_y\}_{y \in \idx{Y}}) \\
 &=
@@
    \faml{X}_x \arrow[u, "\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
-\inj_{\faml{X}_x})"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}"'] & \faml{Z}_{\idx{f}(x,y)}
+\biinj_{\faml{X}_x})"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\biinj_{\faml{X}_x}"'] & \faml{Z}_{\idx{f}(x,y)}
 \end{tikzcd}
@@
    \faml{Y}_y \arrow[dr,
-   "\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}"] \\
-   \faml{Y}_y \arrow[u, -, double, "\id_{\faml{Y}_y}"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}"'] & \faml{Z}_{\idx{f}(x,y)}
+   "\faml{f}_{(x,y)}\,\comp\,\biinj_{\faml{Y}_y}"] \\
+   \faml{Y}_y \arrow[u, -, double, "\id_{\faml{Y}_y}"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\biinj_{\faml{Y}_y}"'] & \faml{Z}_{\idx{f}(x,y)}
 \end{tikzcd}
@@
 The diagram on the left below then commutes by Proposition~\ref{prop:biproduct:prod-coprod}, with
-$\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}} =
+$\coprodM{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{Y}_y}} =
 \faml{f}_{(x,y)}$ by the uniqueness of the coproduct of morphisms. Then the diagram on the right in $\Fam(C)$
@@
    \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \biprod \faml{Y}_y \arrow[dr,
-   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}}"] \\
+   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\biinj_{\faml{Y}_y}}"] \\
    \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
-\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r,
-"\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] &
+\biinj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r,
+"\coprodM{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{Y}_y}}"'] &
 \faml{Z}_{\idx{f}(x,y)}

### 2025-03-31 notes/sec/cmon-enriched.tex
secrefs.
C -> D.
Rename.
Or coproducts.
copair-ε.

@@
 \subsection{$\CMon$-enriched categories}
+\label{sec:cmon-enriched}
 
@@
 $\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object, monoidal
-product $A \tensor B$ given by the direct product of $A$ and $B$, and braiding given by the isomorphism $A
+product $A \tensor B$ given by the Cartesian product of $A$ and $B$, and braiding given by the isomorphism $A
 \times B \iso B \times A$. $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
@@
 
-$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
-monoid of commutative monoid homomorphisms where:
+$\CMon$ is closed monoidal, and so enriched over itself. More specifically every hom-object $\CMon(A,B)$ is a
+commutative monoid of commutative monoid homomorphisms where:
+
 \begin{enumerate}

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Construct biproducts from coproducts on a cmon-category.

@@
 \item objects $X, Y$ are the bounded lattices;
-\item morphisms $f \vdash g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
+\item morphisms $f \dashv g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
 join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
@@
 \begin{itemize}
-\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \vdash 0_{Y,X}: X \to Y$
-\item $(f \vdash g) + (f' \vdash g') = (f + f') \vdash (g + g'): X \to Y$.
+\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \dashv 0_{Y,X}: X \to Y$
+\item $(f \dashv g) + (f' \dashv g') = (f + f') \dashv (g + g'): X \to Y$.
 \end{itemize}
 
-\noindent and biadditivity of composition arising component-wise. Moreover $\LatGal$ has binary products, with
-the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts:
+\noindent and biadditivity of composition arises component-wise. Moreover $\LatGal$ has binary products, with
+the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts
+(\propref{biproduct:from-product-or-coproduct}).
 

### 2025-03-31 notes/sec/biproduct.tex
copair-ε.
Construct biproducts from coproducts on a cmon-category.
Merge propositions 3.6 and 3.7.

@@
 
-\begin{proposition}
-\item
-\begin{itemize}
-\item If $C$ has biproducts then it has products given by $(A \biprod B, \biproj_A, \biproj_B)$ and coproducts
-given by $(A \biprod B, \biinj_A, \biinj_B)$.
-\item If $C$ has products (or coproducts) then they are biproducts.
-\item If $C$ has a terminal (or initial) object then it is a zero object.
-\end{itemize}
-\end{proposition}
-
 \noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts. Having products or coproducts
-is sufficient for semi-additivity:
+is sufficient for semi-additivity.
 
 \begin{proposition}
+\label{prop:biproduct:from-product-or-coproduct}
 Suppose $C$ is $\CMon$-enriched.
 \begin{itemize}
-\item If $C$ has products given by $(A \times B, \proj_1, \proj_2)$ then:
+\item If $C$ has biproducts given by $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
+\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
+\item If $C$ has binary products given by $(A \times B, \proj_1, \proj_2)$ then:
 \begin{enumerate}
@@
 \end{enumerate}
-\item If $C$ has coproducts given by $(A + B, \inj_1, \inj_2)$ then
+\item If $C$ has binary coproducts given by $(A + B, \inj_1, \inj_2)$ then
 \begin{enumerate}
+\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
+\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1 + f_2}{g_1 + g_2}$
 \item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
@@
 \end{enumerate}
+\item If $C$ has a terminal (or initial) object then it is a zero object.
 \end{itemize}

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Why LatGal has finite biproducts.

@@
 
-\noindent and biadditivity of composition arises component-wise. Moreover $\LatGal$ has binary products, with
-the product of Galois connections defined component-wise; it follows that $\LatGal$ has biproducts
+\noindent and biadditivity of composition is component-wise. Moreover $\LatGal$ has finite products, with the
+trivial 1-point lattice as terminal object, product lattice $X \times Y$ as binary product, and product of
+Galois connections defined component-wise; it follows that $\LatGal$ has finite biproducts
 (\propref{biproduct:from-product-or-coproduct}).

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Onto vector space over a field F.

@@
 
+\begin{definition}[Vector space over a field $F$]
+Suppose $F$ a field with addition of elements written $a + b$ and multiplication written $a \mult b$.
+Overloading $+$ and $\mult$, define a \emph{vector space $V = (V, +, \mult)$ over $F$} to be a set $V$
+equipped with vector addition $+: V^2 \to V$ () and scalar multiplication $\mult: F \times V \to V$ where
+$(V,+)$ is an Abelian group and the vector operations are compatible with the field operations in the
+following sense:
+\begin{itemize}
+\item
+\end{itemize}
+\end{definition}
+
 % \subsubsection{$\CMon$-valued presheaves over $\LatGal$}

### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Onto vector space over a field F.

--- 2025-03-31 notes/sec/some-categories-with-biproducts.tex
@@
 
+\begin{definition}[Vector space over a field $F$]
+Suppose $F$ a field with addition of elements written $a + b$ and multiplication written $a \mult b$.
+Overloading $+$ and $\mult$, define a \emph{vector space $V = (V, +, \mult)$ over $F$} to be a set $V$
+equipped with addition of vectors $+: V^2 \to V$ () and scalar multiplication $\mult: F \times V \to V$ where
+$(V,+)$ is an Abelian group and the vector operations are compatible with the field operations in that the
+following equations hold:
+\begin{itemize}
+\item $0 + a = a$
+\item $1 \mult a = a$
+\item $(a \mult b) \mult v = a \mult (b \mult v)$
+\item $(a + b) \mult v = (a \mult v) + (b \mult v)$
+\item $a \mult (u + v) = (a \mult u) + (b \mult v)$
+\end{itemize}
+\end{definition}
+
 % \subsubsection{$\CMon$-valued presheaves over $\LatGal$}

### 2025-04-01 notes.tex
Rename.
FdVect.

@@
 \item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
-\item Examples of useful categories with biproducts:
+\item Examples of useful categories with biproducts (\secref{categories-with-biproducts}):
    \begin{itemize}
    \item category $\LatGal$ of bounded lattices and Galois connections (\secref{categories-with-biproducts:latgal})
-   \item category $\FdVect_k$ of finite-dimensional vector spaces over a field $k$ (\secref{categories-with-biproducts:fdvect})
+   \item category $\FdVect_F$ of finite-dimensional vector spaces over a field $F$
+   (\secref{categories-with-biproducts:fdvect})
    \end{itemize}
@@
 \input{notes/sec/grothendieck}
-\input{notes/sec/galois}
+\input{notes/sec/some-categories-with-biproducts}
 

### 2025-04-01 notes/sec/some-categories-with-biproducts.tex
FdVect.

--- 2025-03-31 notes/sec/some-categories-with-biproducts.tex
@@
 
+\begin{definition}[Vector space over a field $F$]
+Suppose $F$ a field with addition of elements written $a + b$ and multiplication written $a \mult b$. Then a
+\emph{vector space $V = (V, +, \mult)$ over $F$} (overloading $+$ and $\mult$) is a set $V$ equipped with
+addition of vectors $+: V^2 \to V$ and scalar multiplication $\mult: F \times V \to V$ where $(V,+)$ is an
+abelian group (with identity element $0_V$ and additive inverse $-v$ for any $v \in V$) and the vector
+operations are compatible with the field operations in that the following equations hold:
+\begin{itemize}
+\item $1 \mult v = v$
+\item $(a \mult b) \mult v = a \mult (b \mult v)$
+\item $(a + b) \mult v = (a \mult v) + (b \mult v)$
+\item $a \mult (u + v) = (a \mult u) + (b \mult v)$
+\end{itemize}
+\end{definition}
+
+Up to isomorphism a finite-dimensional vector space over $F$ of dimension $n \in \Nat$ is $(F^n, +, \mult)$
+with $+$ and $\mult$ defined component-wise. Define $\FdVect_F$ to be the category where
+\begin{itemize}
+\item the objects are natural numbers $n$ paired with an $n$-dimensional vector space $F^n$
+\end{itemize}
+
 % \subsubsection{$\CMon$-valued presheaves over $\LatGal$}

### 2025-04-01 notes/sec/some-categories-with-biproducts.tex
biadditive -> bilinear seems more common.

@@
 
-\begin{definition}[Vector space over a field $F$]
+\begin{definition}[Vector space over a field]
 Suppose $F$ a field with addition of elements written $a + b$ and multiplication written $a \mult b$. Then a
@@
 Up to isomorphism a finite-dimensional vector space over $F$ of dimension $n \in \Nat$ is $(F^n, +, \mult)$
-with $+$ and $\mult$ defined component-wise. Define $\FdVect_F$ to be the category where
+with $+$ and $\mult$ defined component-wise. Define $\FinVect_F$ to be the category where
 \begin{itemize}
-\item the objects are natural numbers $n$ paired with an $n$-dimensional vector space $F^n$
+\item objects $V, W$ are natural numbers $n$ paired with an $n$-dimensional vector space $F^n$
+\item morphisms $f: V \to W$ are functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a \mult f(v)$.
 \end{itemize}
+
+$\FinVect_F$ is $\Ab$-enriched (and therefore $\CMon$-enriched). For any $V, W$ we have that $\FinVect_F(V,W)$
+is an abelian group, where:
+\begin{enumerate}
+\item $0_{V,W}(v) = 0_W$
+\item $(f + g)(v) = f(v) + g(v)$
+\item $(-f)(v) = -f(v)$
+\end{enumerate}
+\noindent with composition bilinear by (2) and the fact that addition of morphisms acts component-wise.
 

### 2025-04-01 notes.tex
biadditive -> bilinear seems more common.
Rename.

--- 2025-03-26 notes.tex
@@
 \item $\CMon$-enriched categories (\secref{cmon-enriched})
-\item biproducts (\secref{biproduct})
+\item biproducts and semi-additive categories (\secref{biproduct})
 \item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
@@
 \item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
-\item Examples of useful categories with biproducts:
+\item Examples of useful semi-additive categories (\secref{useful-semi-additive-categories}):
    \begin{itemize}
    \item category $\LatGal$ of bounded lattices and Galois connections (\secref{categories-with-biproducts:latgal})
-   \item category $\FdVect_k$ of finite-dimensional vector spaces over a field $k$ (\secref{categories-with-biproducts:fdvect})
+   \item category $\FinVect_F$ of finite-dimensional vector spaces over a field $F$
+   (\secref{categories-with-biproducts:fdvect})
    \end{itemize}
@@
 \input{notes/sec/grothendieck}
-\input{notes/sec/galois}
+\input{notes/sec/useful-semi-additive-categories}
 

### 2025-04-01 notes/sec/biproduct.tex
biadditive -> bilinear seems more common.
Rename.

@@
 
-Suppose $C$ a $\CMon$-enriched category.
+Suppose $C$ a $\CMon$-enriched category. \todo{I think we can also define ``biproduct'' using a universal
+property and zero morphisms, which might be nicer as we can then derive rather than assume $\CMon$-enrichment.}
 
@@
 \begin{itemize}
-\item If $C$ has biproducts given by $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
+\item If $C$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
 \biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
-\item If $C$ has binary products given by $(A \times B, \proj_1, \proj_2)$ then:
+\item If $C$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
 \begin{enumerate}
@@
 \end{enumerate}
-\item If $C$ has binary coproducts given by $(A + B, \inj_1, \inj_2)$ then
+\item If $C$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
 \begin{enumerate}

### 2025-04-01 notes/sec/cmon-enriched.tex
biadditive -> bilinear seems more common.
$\CMon569XZilmsenriched presheaf.

@@
 
-\subsubsection{$\CMon$-enriched categories}
+\subsubsection{$\CMon$-enriched category}
 
@@
 morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
+\item Composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
 \Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
@@
 \end{proposition}
+
+\subsubsection{$\CMon$-enriched presheaf}
+
+For any monoidal category $V$ and $V$-enriched categories $C$ and $D$, the action of a \emph{$V$-enriched}
+functor $F: C \to D$ on morphisms takes the form of a morphism $C(X,Y) \to D(F(X),F(Y))$ between hom-objects
+in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, we have that $F$
+satisfies:
+\begin{itemize}
+\item $F(0_{X,Y}) = 0_{FY,FX}$
+\item $F(f + g) = Ff + Fg$
+\end{itemize}

### 2025-04-01 notes/sec/cmon-enriched.tex
$\CMon569XZilmsenriched presheaf.
CMon-enriched functors also preserve bilinear composition.

@@
 functor $F: C \to D$ on morphisms takes the form of a morphism $C(X,Y) \to D(F(X),F(Y))$ between hom-objects
-in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, we have that $F$
-satisfies:
+in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$
+preserves the additive structure of morphisms:
 \begin{itemize}
-\item $F(0_{X,Y}) = 0_{FY,FX}$
-\item $F(f + g) = Ff + Fg$
+\item $F(0_{X,Y}) = 0_{F(Y),F(X)}$
+\item $F(f + g) = F(f) + F(g)$
 \end{itemize}
+and also preserves bilinear composition, i.e.:
+\begin{itemize}
+\item $F(f + g) \comp F(h) = F(h \comp f) + F(h \comp g)$
+\item $F(h) \comp F(f + g) = F(f \comp h) + F(g \comp h)$
+\end{itemize}
+in addition to the usual functorial properties.

### 2025-04-01 notes/sec/cmon-enriched.tex
CMon-enriched functors also preserve bilinear composition.
But the fact that F preserves bilinear composition is derivable rather than a requirement on F.
Defining your 'basic' Yoneda embedding.

@@
 For any monoidal category $V$ and $V$-enriched categories $C$ and $D$, the action of a \emph{$V$-enriched}
-functor $F: C \to D$ on morphisms takes the form of a morphism $C(X,Y) \to D(F(X),F(Y))$ between hom-objects
+functor $F: C \to D$ on morphisms takes the form of a morphism $C(X,Y) \to D(F(X),F(Y))$ between hom-monoids
 in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$
@@
 \end{itemize}
-and also preserves bilinear composition, i.e.:
+It then follows automatically that $F$ also preserves bilinear composition, i.e.:
 \begin{itemize}
+\item $F(f) \comp F(0) = F(f) = F(0) \comp F(f)$
 \item $F(f + g) \comp F(h) = F(h \comp f) + F(h \comp g)$
@@
 in addition to the usual functorial properties.
+
+\subsubsection{$\CMon$-enriched Yoneda embedding}
+
+Suppose $C$ a small category. The standard \emph{Yoneda embedding} for $C$ is the functor $\Yoneda: C \to
+\Func{C^\op}{\Set}$ sending:
+\begin{itemize}
+\item any object $X$ in $C$ to the contravariant hom-functor $C(-,X)$ which sends:
+\begin{itemize}
+\item any object $Y$ in $C$ to the hom-set $C(Y,X)$
+\item any morphism $f: Y \to Z$ in $C$ to the pre-composition function $(- \comp f): C(Z,X) \to C(Y,X)$
+\end{itemize}
+\item any morphism $f: X \to Y$ in $C$ to the natural transformation
+\end{itemize}

### 2025-04-01 notes/sec/cmon-enriched.tex
Defining your 'basic' Yoneda embedding.

@@
 
-$\CMon$ is closed monoidal, and so enriched over itself. More specifically every hom-object $\CMon(A,B)$ is a
-commutative monoid of commutative monoid homomorphisms where:
+$\CMon$ is closed monoidal, and so enriched over itself. More specifically every hom-object $\Hom{CMon}{A}{B}$
+is a commutative monoid of commutative monoid homomorphisms where:
 
@@
 For any monoidal category $V$ and $V$-enriched categories $C$ and $D$, the action of a \emph{$V$-enriched}
-functor $F: C \to D$ on morphisms takes the form of a morphism $C(X,Y) \to D(F(X),F(Y))$ between hom-monoids
-in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$
-preserves the additive structure of morphisms:
+functor $F: C \to D$ on morphisms takes the form of a morphism $\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between
+hom-monoids in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means
+that $F$ preserves the additive structure of morphisms:
 \begin{itemize}
@@
 \begin{itemize}
-\item any object $X$ in $C$ to the contravariant hom-functor $C(-,X)$ which sends:
+\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}$ which sends:
 \begin{itemize}
-\item any object $Y$ in $C$ to the hom-set $C(Y,X)$
-\item any morphism $f: Y \to Z$ in $C$ to the pre-composition function $(- \comp f): C(Z,X) \to C(Y,X)$
+\item any object $Y$ in $C$ to the hom-set $\Hom{C}{Y}{X}$;
+\item any morphism $f: Y \to Z$ in $C$ to the function $\Hom{C}{f}{X} = (- \comp f): \Hom{C}{Z}{X} \to
+\Hom{C}{Y}{X}$.
 \end{itemize}
-\item any morphism $f: X \to Y$ in $C$ to the natural transformation
+\item any morphism $f: X \to Y$ in $C$ to the natural transformation $\Yoneda(f)$ where $\Yoneda(f)_Z = (f
+\comp -): \Hom{C}{Z}{X} \to \Hom{C}{Z}{Y}$.
 \end{itemize}

### 2025-04-01 notes/sec/cmon-enriched.tex
Notation for V-enriched functor category.

@@
 functor $F: C \to D$ on morphisms takes the form of a morphism $\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between
-hom-monoids in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means
-that $F$ preserves the additive structure of morphisms:
+hom-objects in $V$. Write $\Func{C}{D}_V$ for the category of $V$-enriched functors from $C$ to $D$. In the
+specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$ preserves the
+additive structure of morphisms:
 \begin{itemize}
@@
 \end{itemize}
+
+If $C$ is $\CMon$-enriched then there is an enriched version of its Yoneda embedding, namely the
+$\CMon$-enriched functor $\Yoneda_{\CMon}: \Func{C^\op}{\CMon}$

### 2025-04-01 notes/sec/cmon-enriched.tex
Notation for V-enriched functor category.

@@
 If $C$ is $\CMon$-enriched then there is an enriched version of its Yoneda embedding, namely the
-$\CMon$-enriched functor $\Yoneda_{\CMon}: \Func{C^\op}{\CMon}$
+$\CMon$-enriched functor $\Yoneda_{\CMon}: \Func{C^\op}{\CMon}_{\CMon}$

### 2025-04-02 notes/sec/cmon-enriched.tex
Be more explicit about V-enriched functors.

--- 2025-04-01 notes/sec/cmon-enriched.tex
@@
 
-For any monoidal category $V$ and $V$-enriched categories $C$ and $D$, the action of a \emph{$V$-enriched}
-functor $F: C \to D$ on morphisms takes the form of a morphism $\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between
-hom-monoids in $V$. In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means
-that $F$ preserves the additive structure of morphisms:
+For any monoidal category $\mathscr{V}$ and $V$-enriched categories $C$ and $D$, the action of a
+\emph{$\mathscr{V}$-enriched} functor $F: C \to D$ on morphisms takes the form of a morphism $F_{X,Y}:
+\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ that is
+\emph{$\mathscr{V}$-natural}, i.e.~makes the following diagram commute:
+
+\begin{center}
+\begin{tikzcd}
+   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "- \comp -"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "F_{X,Z}"] \\
+   \Hom{D}{F(Y)}{F(Z)} \tensor \Hom{D}{F(X)}{F(Y)} \arrow[r, "- \comp -"'] & \Hom{D}{F(X)}{F(Z)}
+\end{tikzcd}
+\end{center}
+
+Write $\Func{C}{D}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from $C$ to $D$. In the
+specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$ preserves the
+additive structure of morphisms:
 \begin{itemize}
@@
 
-Suppose $C$ a small category. The standard \emph{Yoneda embedding} for $C$ is the functor $\Yoneda: C \to
+Suppose $C$ a small category. The usual \emph{Yoneda embedding} for $C$ is the functor $\Yoneda: C \to
 \Func{C^\op}{\Set}$ sending:
 \begin{itemize}
-\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}$ which sends:
+\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}: C \to \Set$ which
+sends:
    \begin{itemize}
@@
 \end{itemize}
+
+\noindent For a $\mathscr{V}$-enriched category $C$, the hom-functor $\Hom{C}{-}{X}: C \to \mathscr{V}$ sends
+objects $Y$ in $C$ to hom-objects $\Hom{C}{Y}{X}$ in $\mathscr{V}$ and morphisms $f: Y \to Z$ in $C$ to
+morphisms
+
+\noindent For $\CMon$-enriched $C$, there is an enriched version of the Yoneda embedding, namely the
+$\CMon$-enriched functor $\Yoneda_{\CMon}: C \to \Func{C^\op}{\CMon}_{\CMon}$. First we note that the
+hom-functor $\Hom{C}{-}{X}: C \to \CMon$ is $\CMon$-enriched because composition is bilinear and so $(- \comp
+f)$ preserves the additive structure of morphisms.

### 2025-04-02 notes/sec/cmon-enriched.tex
WIP.

@@
 
-\subsubsection{$\CMon$-enriched presheaf}
+\subsubsection{$\CMon$-enriched presheaves}
 
-For any monoidal category $\mathscr{V}$ and $V$-enriched categories $C$ and $D$, the action of a
-\emph{$\mathscr{V}$-enriched} functor $F: C \to D$ on morphisms takes the form of a morphism $F_{X,Y}:
-\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ that is
-\emph{$\mathscr{V}$-natural}, i.e.~makes the following diagram commute:
+\begin{definition}[$\mathscr{V}$-enriched functor]
+For any monoidal category $(\mathscr{V}, \tensor, I)$ and $V$-enriched categories $C$ and $D$, a
+\emph{$\mathscr{V}$-enriched} functor $F: C \to D$ on morphisms is a family of morphisms $F_{X,Y}:
+\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ that makes the following diagrams
+in $\mathscr{V}$ commute:
 
 \begin{center}
-\begin{tikzcd}
-   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "- \comp -"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "F_{X,Z}"] \\
-   \Hom{D}{F(Y)}{F(Z)} \tensor \Hom{D}{F(X)}{F(Y)} \arrow[r, "- \comp -"'] & \Hom{D}{F(X)}{F(Z)}
+\begin{tikzcd}[column sep=1.8cm]
+I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{C}{X}{X} \arrow[d, "F_{X,X}"] \\
+I \arrow[r, "\id_{F(X)}"'] & \Hom{D}{F(X)}{F(X)}
+\end{tikzcd}
+\hspace{5mm}
+\begin{tikzcd}[column sep=2.6cm]
+   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "F_{X,Z}"] \\
+   \Hom{D}{F(Y)}{F(Z)} \tensor \Hom{D}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] & \Hom{D}{F(X)}{F(Z)}
 \end{tikzcd}
 \end{center}
+\end{definition}
 
-Write $\Func{C}{D}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from $C$ to $D$. In the
-specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$ preserves the
+Write $\Func{C}{D}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from $C$ to $D$.
+
+In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$ preserves the
 additive structure of morphisms:
@@
 
-\noindent For a $\mathscr{V}$-enriched category $C$, the hom-functor $\Hom{C}{-}{X}: C \to \mathscr{V}$ sends
-objects $Y$ in $C$ to hom-objects $\Hom{C}{Y}{X}$ in $\mathscr{V}$ and morphisms $f: Y \to Z$ in $C$ to
-morphisms
+\noindent For a $\mathscr{V}$-enriched category $C$, the hom-functor $\Hom{C}{-}{X}: C \to \mathscr{V}$ is
+$\mathscr{V}$-enriched.
 

### 2025-04-02 notes/sec/cmon-enriched.tex
No need to mention that CMon is symmetric closed.

@@
 \begin{definition}[Category of commutative monoids]
-Define $\CMon$ to be the category whose objects are the commutative monoids $A$ and morphisms $A \to B$ are
-the commutative monoid homomorphisms.
+Define $\CMon$ to be the category whose objects $A$ are the commutative monoids $A$ and morphisms $f: A \to B$
+are the commutative monoid homomorphisms.
 \end{definition}
 
-$\CMon$ is symmetric closed monoidal, with the trivial one-element monoid $\One$ as terminal object, monoidal
-product $A \tensor B$ given by the Cartesian product of $A$ and $B$, and braiding given by the isomorphism $A
-\times B \iso B \times A$. $\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$.
+$\CMon$ is closed monoidal, with the trivial one-element monoid $\One$ as terminal object and monoidal product
+$A \tensor B$ given by the Cartesian product $A \times B$. $\CMon$ is complete and cocomplete, inheriting all
+limits and colimits from $\Set$.
 

### 2025-04-02 notes/sec/cmon-enriched.tex
Various tweaks.

@@
 
-We note that if a category $C$ is $\CMon$-enriched, then:
+Recall that if a category $C$ is $\CMon$-enriched, then:
 \begin{enumerate}
@@
 
-$\CMon$ is closed monoidal, and so enriched over itself. More specifically every hom-object $\Hom{CMon}{A}{B}$
-is a commutative monoid of commutative monoid homomorphisms where:
+Because $\CMon$ is closed monoidal, it is enriched over itself; every hom-object $\Hom{CMon}{A}{B}$ is a
+commutative monoid where:
 
@@
 \item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item addition of morphisms $+$ is defined pointwise so that $(f + g)(a) = f(a) + g(a)$.
+\item the binary operation $+$ is pointwise addition of homomorphisms $(f + g)(a) = f(a) + g(a)$.
 \end{enumerate}
@@
 \begin{proposition}
-If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
+If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
 \end{proposition}

### 2025-04-02 notes/sec/biproduct.tex
Enriched hom-functor.

@@
 
-Suppose $C$ a $\CMon$-enriched category. \todo{I think we can also define ``biproduct'' using a universal
-property and zero morphisms, which might be nicer as we can then derive rather than assume $\CMon$-enrichment.}
+Suppose $C$ a $\CMon$-enriched category. \todo{I think we can also define ``biproduct'' just using zero
+morphisms, and then derive rather than assume $\CMon$-enrichment.}
 

### 2025-04-02 notes/sec/cmon-enriched.tex
Enriched hom-functor.

@@
 
-Because $\CMon$ is closed monoidal, it is enriched over itself; every hom-object $\Hom{CMon}{A}{B}$ is a
+Because $\CMon$ is closed monoidal, it is enriched over itself; every hom-object $\Hom{\CMon}{A}{B}$ is a
 commutative monoid where:
@@
 \begin{definition}[$\mathscr{V}$-enriched functor]
-For any monoidal category $(\mathscr{V}, \tensor, I)$ and $V$-enriched categories $C$ and $D$, a
+For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $C$ and $D$, a
 \emph{$\mathscr{V}$-enriched} functor $F: C \to D$ on morphisms is a family of morphisms $F_{X,Y}:
@@
 
-In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, this means that $F$ preserves the
-additive structure of morphisms:
+In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, enrichment means that $F$
+preserves the additive structure of morphisms:
 \begin{itemize}
@@
 \end{itemize}
-in addition to the usual functorial properties.
+in addition to the usual enriched functor properties.
 
@@
 \begin{itemize}
-\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}: C \to \Set$ which
-sends:
+\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}: C^\op \to \Set$
+which sends:
    \begin{itemize}
@@
 
-\noindent For a $\mathscr{V}$-enriched category $C$, the hom-functor $\Hom{C}{-}{X}: C \to \mathscr{V}$ is
-$\mathscr{V}$-enriched.
+\noindent Now suppose $C$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal. Then the hom-functor
+$\Hom{C}{-}{X}: C^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched; the following diagram commutes by the
+associativity of composition:
 
-\noindent For $\CMon$-enriched $C$, there is an enriched version of the Yoneda embedding, namely the
-$\CMon$-enriched functor $\Yoneda_{\CMon}: C \to \Func{C^\op}{\CMon}_{\CMon}$. First we note that the
-hom-functor $\Hom{C}{-}{X}: C \to \CMon$ is $\CMon$-enriched because composition is bilinear and so $(- \comp
-f)$ preserves the additive structure of morphisms.
+\begin{center}
+\begin{tikzcd}[column sep=2.7cm]
+   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "\Hom{C}{-}{W}_{Y,Z}\,\tensor\,\Hom{C}{-}{W}_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "\Hom{C}{-}{W}_{X,Z}"] \\
+   \Hom{\mathscr{V}}{\Hom{C}{Z}{W}}{\Hom{C}{Y}{W}} \tensor \Hom{\mathscr{V}}{\Hom{C}{Y}{W}}{\Hom{C}{X}{W}} \arrow[r, "\comp_{\Hom{C}{X}{W},\Hom{C}{Y}{W},\Hom{C}{Z}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{Z}{W}}{\Hom{C}{X}{W}}
+\end{tikzcd}
+\end{center}
+
+\noindent \todo{Then there is an enriched version of the Yoneda embedding, namely a functor
+$\Yoneda_{\mathscr{V}}: C \to \Func{C^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to the above which is itself
+$\mathscr{V}$-enriched.}
+
+% First we note that the hom-functor $\Hom{C}{-}{X}: C \to \CMon$ is $\CMon$-enriched because composition is
+% bilinear and so $(- \comp f)$ preserves the additive structure of morphisms.

### 2025-04-02 notes/sec/cmon-enriched.tex
Just need to fix direction of morphisms in C^op.

@@
 
-\noindent Now suppose $C$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal. Then the hom-functor
-$\Hom{C}{-}{X}: C^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched; the following diagram commutes by the
-associativity of composition:
+\noindent Now suppose $C$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the hom-functor
+$\Hom{C}{-}{X}: C^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched. The following diagram commutes because
+precomposition with the identity is the identity on morphisms:
+
+\begin{center}
+\begin{tikzcd}[column sep=1.8cm]
+I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{C}{X}{X} \arrow[d, "\Hom{C}{-}{W}_{X,X}"] \\
+I \arrow[r, "\id_{\Hom{C}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{X}{W}}
+\end{tikzcd}
+\end{center}
+
+\noindent and the following by the associativity of composition:
 

### 2025-04-02 notes/sec/cmon-enriched.tex
Start on enriched Yoneda.

@@
 \begin{tikzcd}[column sep=2.7cm]
-   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "\Hom{C}{-}{W}_{Y,Z}\,\tensor\,\Hom{C}{-}{W}_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "\Hom{C}{-}{W}_{X,Z}"] \\
-   \Hom{\mathscr{V}}{\Hom{C}{Z}{W}}{\Hom{C}{Y}{W}} \tensor \Hom{\mathscr{V}}{\Hom{C}{Y}{W}}{\Hom{C}{X}{W}} \arrow[r, "\comp_{\Hom{C}{X}{W},\Hom{C}{Y}{W},\Hom{C}{Z}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{Z}{W}}{\Hom{C}{X}{W}}
+   \Hom{C}{Z}{Y} \tensor \Hom{C}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "\Hom{C}{-}{W}_{Y,Z}\,\tensor\,\Hom{C}{-}{W}_{X,Y}"'] & \Hom{C}{Z}{X} \arrow[d, "\Hom{C}{-}{W}_{X,Z}"] \\
+   \Hom{\mathscr{V}}{\Hom{C}{Y}{W}}{\Hom{C}{Z}{W}} \tensor \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{Y}{W}} \arrow[r, "\comp_{\Hom{C}{X}{W},\Hom{C}{Y}{W},\Hom{C}{Z}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{Z}{W}}
 \end{tikzcd}
@@
 \noindent \todo{Then there is an enriched version of the Yoneda embedding, namely a functor
-$\Yoneda_{\mathscr{V}}: C \to \Func{C^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to the above which is itself
+$\Yoneda_{\mathscr{V}}: C \to \Func{C^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to $\Yoneda$ which is itself
 $\mathscr{V}$-enriched.}

### 2025-04-02 notes/sec/biproduct.tex
Tweaks to biproducts.

@@
 
-Suppose $C$ a $\CMon$-enriched category. \todo{I think we can also define ``biproduct'' just using zero
-morphisms, and then derive rather than assume $\CMon$-enrichment.}
-
 \begin{definition}[Biproduct]
-For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \biprod B$ of $C$ together with
-morphisms
+Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
+object $A \biprod B$ of $C$ together with morphisms
 
@@
 
-\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts. Having products or coproducts
-is sufficient for semi-additivity.
+\noindent \todo{I think we can also define ``biproduct'' just using zero morphisms, and then derive rather
+than assume $\CMon$-enrichment.}
+
+We say a $\CMon$-enriched $C$ is \emph{semi-additive} if it also has finite biproducts; having products or
+coproducts is sufficient.
 

### 2025-04-10 appendix/notes/language.tex
Move notes to appendix.

\section{Language}

\subsection{Syntax}

We parameterise our language over a signature $\Sigma = (S, \phi, R)$ consisting of:
\begin{itemize}
\item a set $S$ of sorts
\item a set of operations $\phi$ with signature $\Sigma(\phi) \in S^* \times S$
\item a set of relations $R$ with signature $\Sigma(R) \in S^*$
\end{itemize}

\subsection{Semantic domain}

We interpret our language into any category with:
\begin{itemize}
\item finite products (binary products and a terminal object)
\item Booleans
\end{itemize}
\note{Ignoring lists for now. Booleans can be derived from coproducts.}

### 2025-04-10 appendix/notes/set-indexed-product.tex
Move notes to appendix.

\subsection{Set-indexed products}
\label{sec:set-indexed-product}

$C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms $\eval_i$ in
$C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i \in I$:

\begin{center}
\begin{tikzcd}
   \prod_{I} X \arrow[r, "\eval_i"] & X_i \\
   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda(f)"]
\end{tikzcd}
\end{center}

Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,C)$.

### 2025-04-10 main.tex
Update class file; rename main LaTeX file.
Authors.
Move notes to appendix.

--- 2025-03-26 notes.tex
@@
-\documentclass[acmsmall,nonacm]{acmart}
+\documentclass[acmsmall,screen]{acmart}
 
@@
 
-\title{Approximation as Differentiation}
+\title{Approximation as Differentiation notes}
+
+\author{Robert Atkey}
+\email{robert.atkey@strath.ac.uk}
+\orcid{0000-0002-4414-5047}
+\affiliation{%
+  \institution{University of Strathclyde}
+  \city{Glasgow}
+  \country{UK}}
+
+\author{Roly Perera}
+\email{roly.perera@cl.cam.ac.uk}
+\orcid{0000-0001-9249-9862}
+\affiliation{%
+  \institution{University of Cambridge}
+  \city{Cambridge}
+  \country{UK}
+}
+\additionalaffiliation{%
+   \institution{University of Bristol}
+   \city{Bristol}
+   \country{UK}
+}
+
+\begin{abstract}
+[Abstract]
+\end{abstract}
 \maketitle
 
-\section{Overview}
+\input{main/introduction}
 
-Covered so far:
-\begin{itemize}
-\item $\CMon$-enriched categories (\secref{cmon-enriched})
-\item biproducts (\secref{biproduct})
-\item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
-\item set-indexed products (\secref{set-indexed-product})
-\item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
-\item Examples of useful categories with biproducts:
-   \begin{itemize}
-   \item category $\LatGal$ of bounded lattices and Galois connections (\secref{categories-with-biproducts:latgal})
-   \item category $\FdVect_k$ of finite-dimensional vector spaces over a field $k$ (\secref{categories-with-biproducts:fdvect})
-   \end{itemize}
-\end{itemize}
-
-\noindent $\Set$ will usually be $\Setoid$ in the Agda implementation but we will gloss that detail for now.
-
-\input{notes/sec/notation}
-
-\section{Definitions}
-
-\input{notes/sec/cmon-enriched}
-\input{notes/sec/biproduct}
-\input{notes/sec/fam}
-\input{notes/sec/set-indexed-product}
-\input{notes/sec/grothendieck}
-\input{notes/sec/galois}
-
-% \input{notes/language}
-
-\section{Some references}
-
-\cite{karvonen2020}
+\pagebreak
+\appendix
+\input{appendix/notes}
 

### 2025-04-10 main/introduction.tex
Dummy introduction just to force a bibliography.

\section{Introduction}

\cite{karvonen2020} % just to generate a bibliography for now

### 2025-04-12 appendix/notes/cmon-enriched.tex
Going into a bit too much detail here..

--- 2025-04-02 notes/sec/cmon-enriched.tex
@@
 
-\begin{definition}[Category of commutative monoids]
+\subsubsection{Examples of commutative monoids}
+
+\begin{definition}[Bounded semilattice]
+A \emph{bounded semilattice} $A = (A, \varepsilon, \bullet)$ is a commutative monoid where $\bullet$ is
+\emph{idempotent}, i.e.~satisfies $x \bullet x = x$.
+\end{definition}
+
+\noindent The idempotence of $\bullet$ (together with commutativity and associativity) induces a partial order
+$\le_\bullet$ on $X$, with $x \le_{\bullet} y \iff x \bullet y = x$. With respect to this partial order, $x
+\bullet y$ is the greatest lower bound (meet) of $x$ and $y$ and $\varepsilon$ is the top element; with
+respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ is the bottom
+element. This thus provides the standard algebraic characterisation of the (dual) order-theoretic notions of
+bounded meet semilattices $(X, \top, \meet)$ and bounded join semilattices $(X, \bot, \join)$.
+
+\subsubsection{Category of commutative monoids}
+
+\begin{definition}[Category $\CMon$]
 Define $\CMon$ to be the category whose objects $A$ are the commutative monoids $A$ and morphisms $f: A \to B$

### 2025-04-12 appendix/notes/cmon-enriched.tex
Going into a bit too much detail here..

@@
 respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ is the bottom
-element. This thus provides the standard algebraic characterisation of the (dual) order-theoretic notions of
+element. This therefore provides an algebraic characterisation of the (dual) order-theoretic notions of
 bounded meet semilattices $(X, \top, \meet)$ and bounded join semilattices $(X, \bot, \join)$.

### 2025-04-12 appendix/notes/cmon-enriched.tex
Tweak 'bounded lattice'.

@@
 \begin{definition}[Bounded semilattice]
+\label{def:cmon-enriched:bounded-semilattice}
 A \emph{bounded semilattice} $A = (A, \varepsilon, \bullet)$ is a commutative monoid where $\bullet$ is
@@
 \bullet y$ is the greatest lower bound (meet) of $x$ and $y$ and $\varepsilon$ is the top element; with
-respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ is the bottom
-element. This therefore provides an algebraic characterisation of the (dual) order-theoretic notions of
-bounded meet semilattices $(X, \top, \meet)$ and bounded join semilattices $(X, \bot, \join)$.
+respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ and $\varepsilon$
+is the bottom element. This therefore provides an algebraic characterisation of the usual (dual)
+order-theoretic notions of bounded meet semilattices $(X, \top, \meet)$ and bounded join semilattices $(X,
+\bot, \join)$.
 

### 2025-04-15 appendix/notes.tex
Move notes to appendix.
mathcal for metavariables that range over categories.

--- 2025-04-01 notes.tex
@@
-\documentclass[acmsmall,nonacm]{acmart}
+\section{Notes}
 
-\usepackage{bbm}
-\usepackage{tikz-cd}
-
-\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
-\settopmatter{printacmref=false}
-
-\input{macros}
-
-\begin{document}
-
-\title{Approximation as Differentiation}
-\maketitle
-
-\section{Overview}
+\subsection{Overview}
 
@@
 \begin{itemize}
-\item $\CMon$-enriched categories (\secref{cmon-enriched})
+\item $\CMon$-enriched categories, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
 \item biproducts and semi-additive categories (\secref{biproduct})
-\item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
+\item category $\Fam(I,\cat{C})$ of $I$-indexed families of objects of $\cat{C}$ (\secref{fam})
 \item set-indexed products (\secref{set-indexed-product})
-\item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
+\item Grothendieck construction $\Grothendieck{\cat{C}}F$ and category of families $\Fam(\cat{C})$
+(\secref{grothendieck})
 \item Examples of useful semi-additive categories (\secref{useful-semi-additive-categories}):
@@
 
-\input{notes/sec/notation}
+\input{appendix/notes/notation}
+\input{appendix/notes/cmon-enriched}
+\input{appendix/notes/biproduct}
+\input{appendix/notes/fam}
+\input{appendix/notes/set-indexed-product}
+\input{appendix/notes/grothendieck}
+\input{appendix/notes/useful-semi-additive-categories}
 
-\section{Definitions}
-
-\input{notes/sec/cmon-enriched}
-\input{notes/sec/biproduct}
-\input{notes/sec/fam}
-\input{notes/sec/set-indexed-product}
-\input{notes/sec/grothendieck}
-\input{notes/sec/useful-semi-additive-categories}
-
-% \input{notes/language}
-
-\section{Some references}
-
-\cite{karvonen2020}
-
-\bibliographystyle{plainnat}
-\bibliography{bib}
-
-\end{document}
+% \input{appendix/notes/language}

### 2025-04-15 appendix/notes/cmon-enriched.tex
mathcal for metavariables that range over categories.

@@
 
-Recall that if a category $C$ is $\CMon$-enriched, then:
+Recall that if a category $\cat{C}$ is $\CMon$-enriched, then:
 \begin{enumerate}
-\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$ (zero
-morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
-\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
-$\Hom{C}{A}{B}$ separately:
+\item Every hom-object $\Hom{\cat{C}}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$
+(zero morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
+\item Composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{B}{C} \tensor
+\Hom{\cat{C}}{A}{B} \to \Hom{\cat{C}}{A}{C}$ in $\CMon$ that preserve the additive structure in
+$\Hom{\cat{C}}{B}{C}$ and $\Hom{\cat{C}}{A}{B}$ separately:
 
@@
 \begin{proposition}
-If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
+If $C$ is $\CMon$-enriched then any functor category $\Func{\cat{D}}{\cat{C}}$ is $\CMon$-enriched.
 \end{proposition}
@@
 \begin{definition}[$\mathscr{V}$-enriched functor]
-For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $C$ and $D$, a
-\emph{$\mathscr{V}$-enriched} functor $F: C \to D$ on morphisms is a family of morphisms $F_{X,Y}:
-\Hom{C}{X}{Y} \to \Hom{D}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ that makes the following diagrams
-in $\mathscr{V}$ commute:
+For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
+$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a family of
+morphisms $F_{X,Y}: \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$
+that makes the following diagrams in $\mathscr{V}$ commute:
 
@@
 \begin{tikzcd}[column sep=2.6cm]
-   \Hom{C}{Y}{Z} \tensor \Hom{C}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"'] & \Hom{C}{X}{Z} \arrow[d, "F_{X,Z}"] \\
-   \Hom{D}{F(Y)}{F(Z)} \tensor \Hom{D}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] & \Hom{D}{F(X)}{F(Z)}
+   \Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"']
+   & \Hom{\cat{C}}{X}{Z} \arrow[d, "F_{X,Z}"] \\
+   \Hom{\cat{D}}{F(Y)}{F(Z)} \tensor \Hom{\cat{D}}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] &
+   \Hom{\cat{D}}{F(X)}{F(Z)}
 \end{tikzcd}
@@
 
-Write $\Func{C}{D}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from $C$ to $D$.
+\begin{definition}[Category of enriched functors]
+Write $\Func{\cat{C}}{\cat{D}}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from
+$\cat{C}$ to $\cat{D}$.
+\end{definition}
 
-In the specific setting of a $\CMon$-enriched presheaf $F: C^\op \to \CMon$, enrichment means that $F$
+In the specific setting of a $\CMon$-enriched presheaf $F: \cat{C}^\op \to \CMon$, enrichment means that $F$
 preserves the additive structure of morphisms:
@@
 
-Suppose $C$ a small category. The usual \emph{Yoneda embedding} for $C$ is the functor $\Yoneda: C \to
-\Func{C^\op}{\Set}$ sending:
+Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
+\cat{C} \to \Func{\cat{C}^\op}{\Set}$ sending:
 \begin{itemize}
-\item any object $X$ in $C$ to the contravariant hom-functor $\Yoneda(X) = \Hom{C}{-}{X}: C^\op \to \Set$
+\item any object $X$ in $\cat{C}$ to the contravariant hom-functor $\Yoneda(X) = \Hom{\cat{C}}{-}{X}:
+\cat{C}^\op \to \Set$
 which sends:
    \begin{itemize}
-   \item any object $Y$ in $C$ to the hom-set $\Hom{C}{Y}{X}$;
-   \item any morphism $f: Y \to Z$ in $C$ to the function $\Hom{C}{f}{X} = (- \comp f): \Hom{C}{Z}{X} \to
-   \Hom{C}{Y}{X}$.
+   \item any object $Y$ in $\cat{C}$ to the hom-set $\Hom{\cat{C}}{Y}{X}$;
+   \item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
+   \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$.
    \end{itemize}
-\item any morphism $f: X \to Y$ in $C$ to the natural transformation $\Yoneda(f)$ where $\Yoneda(f)_Z = (f
-\comp -): \Hom{C}{Z}{X} \to \Hom{C}{Z}{Y}$.
+\item any morphism $f: X \to Y$ in $\cat{C}$ to the natural transformation $\Yoneda(f)$ where $\Yoneda(f)_Z =
+(f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
 \end{itemize}
 
-\noindent Now suppose $C$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the hom-functor
-$\Hom{C}{-}{X}: C^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched. The following diagram commutes because
-precomposition with the identity is the identity on morphisms:
+\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the
+hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched. The following
+diagram commutes because precomposition with the identity is the identity on morphisms:
 
@@
 \begin{tikzcd}[column sep=1.8cm]
-I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{C}{X}{X} \arrow[d, "\Hom{C}{-}{W}_{X,X}"] \\
-I \arrow[r, "\id_{\Hom{C}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{X}{W}}
+I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "\Hom{\cat{C}}{-}{W}_{X,X}"] \\
+I \arrow[r, "\id_{\Hom{\cat{C}}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{X}{W}}\
 \end{tikzcd}
@@
 \begin{tikzcd}[column sep=2.7cm]
-   \Hom{C}{Z}{Y} \tensor \Hom{C}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "\Hom{C}{-}{W}_{Y,Z}\,\tensor\,\Hom{C}{-}{W}_{X,Y}"'] & \Hom{C}{Z}{X} \arrow[d, "\Hom{C}{-}{W}_{X,Z}"] \\
-   \Hom{\mathscr{V}}{\Hom{C}{Y}{W}}{\Hom{C}{Z}{W}} \tensor \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{Y}{W}} \arrow[r, "\comp_{\Hom{C}{X}{W},\Hom{C}{Y}{W},\Hom{C}{Z}{W}}"'] & \Hom{\mathscr{V}}{\Hom{C}{X}{W}}{\Hom{C}{Z}{W}}
+   \Hom{\cat{C}}{Z}{Y} \tensor \Hom{\cat{C}}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d,
+   "\Hom{\cat{C}}{-}{W}_{Y,Z}\,\tensor\,\Hom{\cat{C}}{-}{W}_{X,Y}"'] & \Hom{\cat{C}}{Z}{X} \arrow[d,
+   "\Hom{\cat{C}}{-}{W}_{X,Z}"] \\
+   \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{W}}{\Hom{\cat{C}}{Z}{W}} \tensor
+   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Y}{W}} \arrow[r,
+   "\comp_{\Hom{\cat{C}}{X}{W},\Hom{\cat{C}}{Y}{W},\Hom{\cat{C}}{Z}{W}}"'] &
+   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Z}{W}}
 \end{tikzcd}
@@
 \noindent \todo{Then there is an enriched version of the Yoneda embedding, namely a functor
-$\Yoneda_{\mathscr{V}}: C \to \Func{C^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to $\Yoneda$ which is itself
-$\mathscr{V}$-enriched.}
+$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to $\Yoneda$
+which is itself $\mathscr{V}$-enriched.}
 

### 2025-04-15 appendix/notes/biproduct.tex
Assume universal quantification implicit.
Identities in LatGal.
mathcal for metavariables that range over categories.

--- 2025-04-02 notes/sec/biproduct.tex
@@
 \begin{definition}[Biproduct]
-Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
-object $A \biprod B$ of $C$ together with morphisms
+Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is
+an object $A \biprod B$ of $C$ together with morphisms
 
@@
 
-We say a $\CMon$-enriched $C$ is \emph{semi-additive} if it also has finite biproducts; having products or
-coproducts is sufficient.
+We say a $\CMon$-enriched $\cat{C}$ is \emph{semi-additive} if it also has finite biproducts; having products
+or coproducts is sufficient.
 
@@
 \label{prop:biproduct:from-product-or-coproduct}
-Suppose $C$ is $\CMon$-enriched.
+Suppose $\cat{C}$ is $\CMon$-enriched.
 \begin{itemize}
-\item If $C$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
+\item If $\cat{C}$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
 \biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
-\item If $C$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
+\item If $\cat{C}$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
 \begin{enumerate}
 \item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
-\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1 + f_2}{g_1 + g_2}$
+\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
 \item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
@@
 \end{enumerate}
-\item If $C$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
+\item If $\cat{C}$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
 \begin{enumerate}
 \item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
-\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1 + f_2}{g_1 + g_2}$
+\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
 \item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
@@
 \end{enumerate}
-\item If $C$ has a terminal (or initial) object then it is a zero object.
+\item If $\cat{C}$ has a terminal (or initial) object then it is a zero object.
 \end{itemize}
@@
 \label{prop:biproduct:prod-coprod}
-Suppose $C$ a category with biproducts. Then the following commutes for any morphisms $f_1, g_1, f_2,
-g_2$ in $C$:
+Suppose $\cat{C}$ a category with biproducts. Then the following diagram in $\cat{C}$ commutes:
 

### 2025-04-15 appendix/notes/fam.tex
Move notes to appendix.
mathcal for metavariables that range over categories.

--- 2025-03-04 notes/sec/fam.tex
@@
 \subsection{Category of $I$-indexed families}
+\label{sec:fam}
 
 \begin{definition}[Category of $I$-indexed families of objects]
-For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
+For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category where:
 \begin{itemize}
-\item objects are the $I$-indexed families of objects of $C$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
+\item objects are the $I$-indexed families of objects of $\cat{C}$;
+\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $\cat{C}$ for any $i \in I$.
 \end{itemize}
@@
 
-\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$, interpreting $I$ as a discrete
-category.
+\noindent Equivalently $\Fam(I,\cat{C})$ is the functor category $\Func{I}{\cat{C}}$, interpreting $I$ as a
+discrete category.
 
 \begin{definition}[Constant family]
-For any object $X$ of $C$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
+For any object $X$ of $\cat{C}$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
 \end{definition}
@@
 \begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
-$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
+For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$ sends
+any $J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
 similarly for morphisms.
@@
 
-\begin{definition}[$\Fam(-,C)$ functor]
-$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
-$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
+\begin{definition}[$\Fam(-,\cat{C})$ functor]
+$\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,\cat{C})$ and
+any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$.
 \end{definition}
@@
 \begin{proposition}
-If $C$ has limits (resp.~colimits) then $\Fam(I,C)$ has limits (resp.~colimits) computed pointwise.
+If $\cat{C}$ has limits (resp.~colimits) then $\Fam(I,\cat{C})$ has limits (resp.~colimits) computed
+pointwise.
 \end{proposition}

### 2025-04-15 appendix/notes/set-indexed-product.tex
mathcal for metavariables that range over categories.

@@
 
-$C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
-there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms $\eval_i$ in
-$C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
-exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i \in I$:
+$\cat{C}$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of
+$\cat{C}$, there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms
+$\eval_i$ in $\cat{C}$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$
+in $\cat{C}$, there exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i
+\in I$:
 
@@
 
-Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,C)$.
+Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,\cat{C})$.

### 2025-04-15 appendix/notes/grothendieck.tex
mathcal for metavariables that range over categories.

--- 2025-03-31 notes/sec/grothendieck.tex
@@
 \begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: C \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{C}F$ for $F$ is the
-category where:
+Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
+$F$ is the category where:
 \begin{itemize}
-\item objects $X$ are pairs of an object $\idx{X}$ of $C$ and an object $\faml{X}$ of $F(\idx{X})$;
+\item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
 \item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $C$ paired with morphisms
@@
 \begin{proof}
-Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\biinj,\biproj)$ and set-indexed products
-given by $(\prod,\eval_{\prod},\lambda_{\prod})$. For any objects $X, Y, Z$ in $\Fam(C)$ define:
+Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\biinj,\biproj)$ and set-indexed
+products given by $(\prod,\eval_{\prod},\lambda_{\prod})$. Write $(\times,\pi_1,\pi_2)$ for the product in
+$\Fam(C)$ inherited from the biproduct in $C$. For any objects $X, Y, Z$ in $\Fam(C)$ define:
 
@@
 
-Then the following triangle identities hold:
+Then the following triangle identities hold.
 \begin{enumerate}
-\item $\lambda_{X,Y,Z}(\eval_{Y,Z} \comp (f \tensor \id_Y) = f$. Suppose $(f,g): X \to [Y,Z]$ and define
-$(f^\dagger, g^\dagger): X \times Y \to Z$ with:
+\item $\lambda_{X,Y,Z}(\eval_{Y,Z} \comp (f, \id_Y)) = f$. Suppose $(f,g): X \to [Y,Z]$ and define
+$(f^\dagger, g^\dagger) = \eval_{Y,Z} \comp ((f,g), \id_Y): X \times Y \to Z$. We then have:
 \begin{itemize}
-\item $f^\dagger(x,y) = \idx{f(x)}(y)$
-\item $g^\dagger_{x,y} = \coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y}$
+\item $f^\dagger(x,y) = (\idx{\eval_{Y,Z}} \comp (f, \id_{\idx{Y}}))(x,y) = \idx{f(x)}(y)$
+\item $g^\dagger_{x,y} = (\faml{\eval_{Y,Z}} \comp (g, \id_{\faml{Y}}))_{x,y} = \faml{\eval_{Y,Z}}_{f(x), y}
+\comp (g, \id_{\faml{Y}})_{x,y} = \coprodM{{\eval_{\prod}}_y}{\faml{f(x)}_y} \comp (g_x, {\id_{\faml{Y}}}_y) =
+\coprodM{{\eval_{\prod}}_y \comp g_x}{\faml{f(x)}_y}$.
 \end{itemize}
@@
 
-\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f) \tensor \id_{Y}) = f$. Suppose $f: X \times Y \to Z$ in $\Fam(C)$
+\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f), \id_{Y}) = f$. Suppose $f: X \times Y \to Z$ in $\Fam(C)$
 and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property of
@@
 \begin{tikzcd}
-   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
-   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
+   {{[Y, Z]}} \times Y \arrow[dr, "\eval_{Y,Z}"] \\
+   X \times Y \arrow[u, "(\lambda_{X,Y,Z}(f){,}\,\id_Y)"] \arrow[r, "f"'] & Z
 \end{tikzcd}

### 2025-04-15 appendix/notes/grothendieck.tex
mathcal for metavariables that range over categories.

@@
 \item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
-\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $C$ paired with morphisms
+\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with morphisms
 $\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
@@
 \noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
-$C$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
+$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
 
 \begin{definition}[Category of families]
-For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
+For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
+$\Grothendieck{\Set^{\op}}\Fam(-,\cat{C})$.
 \end{definition}
 
-\noindent In $\Fam(C)$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$ in
-$\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
-morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},C)$.
+\noindent In $\Fam(\cat{C})$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$
+in $\Fam(\idx{X},\cat{C})$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
+morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},\cat{C})$.
 
@@
 \begin{enumerate}
-\item If $C$ is locally small then so is $\Fam(C)$.
-\item If $C$ has binary products then so does $\Fam(C)$.
+\item If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
+\item If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
 \end{enumerate}

### 2025-04-15 appendix/notes/grothendieck.tex
LatGal(X,Y) actually a semilattice.

@@
 \begin{proposition}
-Suppose $C$ locally small. If $C$ has binary biproducts and set-indexed products, then $\Fam(C)$ is Cartesian
-closed.
+Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and set-indexed products, then
+$\Fam(\cat{C})$ is Cartesian closed.
 \end{proposition}
@@
 \begin{proof}
-Suppose $C$ is locally small, with binary biproducts given by $(\biprod,\biinj,\biproj)$ and set-indexed
+Suppose $\cat{C}$ is locally small, with binary biproducts given by $(\biprod,\biinj,\biproj)$ and set-indexed
 products given by $(\prod,\eval_{\prod},\lambda_{\prod})$. Write $(\times,\pi_1,\pi_2)$ for the product in
-$\Fam(C)$ inherited from the biproduct in $C$. For any objects $X, Y, Z$ in $\Fam(C)$ define:
+$\Fam(\cat{C})$ inherited from the biproduct in $\cat{C}$. For any objects $X, Y, Z$ in $\Fam(\cat{C})$
+define:
 
 \begin{enumerate}
-\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(C)$ with $X' = \Fam(C)(X,Y)$ and $Y'_f =
+\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(\cat{C})$ with $X' = \Fam(\cat{C})(X,Y)$ and $Y'_f =
 \prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
@@
 \end{itemize}
-\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
+\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(\cat{C})$ with
 \begin{itemize}
@@
 \end{itemize}
-\noindent To show $f' = f : \idx{X} \to \Fam(C)(Y,Z)$, suppose $x \in \idx{X}$ and show $f'(x) = f(x): Y \to
-Z$. Suppose $y \in \idx{Y}$.
+\noindent To show $f' = f : \idx{X} \to \Fam(\cat{C})(Y,Z)$, suppose $x \in \idx{X}$ and show $f'(x) = f(x): Y
+\to Z$. Suppose $y \in \idx{Y}$.
 \begin{enumerate}
@@
 
-\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f), \id_{Y}) = f$. Suppose $f: X \times Y \to Z$ in $\Fam(C)$
+\item $\eval_{Y,Z} \comp (\lambda_{X,Y,Z}(f), \id_{Y}) = f$. Suppose $f: X \times Y \to Z$ in $\Fam(\cat{C})$
 and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property of
@@
 $\coprodM{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\biinj_{\faml{Y}_y}} =
-\faml{f}_{(x,y)}$ by the uniqueness of the coproduct of morphisms. Then the diagram on the right in $\Fam(C)$
-commutes by the definitions of $\lambda_{X,Y,Z}, \eval_{Y,Z}$ and $[Y,Z]$.
+\faml{f}_{(x,y)}$ by the uniqueness of the coproduct of morphisms. Then the diagram on the right in
+$\Fam(\cat{C})$ commutes by the definitions of $\lambda_{X,Y,Z}, \eval_{Y,Z}$ and $[Y,Z]$.
 

### 2025-04-15 appendix/notes/cmon-enriched.tex
Clarify enriched functor properties.

@@
 \begin{definition}[$\mathscr{V}$-enriched functor]
+\label{def:cmon-enriched:enriched-functor}
 For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
@@
 \begin{tikzcd}[column sep=1.8cm]
-I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{C}{X}{X} \arrow[d, "F_{X,X}"] \\
-I \arrow[r, "\id_{F(X)}"'] & \Hom{D}{F(X)}{F(X)}
+I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "F_{X,X}"] \\
+I \arrow[r, "\id_{F(X)}"'] & \Hom{\cat{D}}{F(X)}{F(X)}
 \end{tikzcd}
@@
 \end{definition}
+
+\noindent These generalise the usual functor laws to a setting where the hom-objects are not necessarily sets
+and thus do not support an elementwise formulation.
 
@@
 \end{itemize}
-in addition to the usual enriched functor properties.
+in addition to the usual enriched functor properties in \defref{cmon-enriched:enriched-functor}.
 
@@
 Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
-\cat{C} \to \Func{\cat{C}^\op}{\Set}$ sending:
+\cat{C} \to \Func{\cat{C}^\op}{\Set}$ defined as follows:
 \begin{itemize}
-\item any object $X$ in $\cat{C}$ to the contravariant hom-functor $\Yoneda(X) = \Hom{\cat{C}}{-}{X}:
-\cat{C}^\op \to \Set$
-which sends:
+\item on objects: $\Yoneda(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \Set$, the contravariant hom-functor
+sending:
    \begin{itemize}
@@
    \end{itemize}
-\item any morphism $f: X \to Y$ in $\cat{C}$ to the natural transformation $\Yoneda(f)$ where $\Yoneda(f)_Z =
-(f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
+\item on morphisms: for any $f: X \to Y$ in $\cat{C}$, $\Yoneda(f)$ is the natural transformation where
+$\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
 \end{itemize}
@@
 \noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the
-hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched. The following
-diagram commutes because precomposition with the identity is the identity on morphisms:
+hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched. The
+
+The following diagram commutes because precomposition with the identity is the identity on morphisms:
 

### 2025-04-15 appendix/notes/cmon-enriched.tex
Enriched Yoneda lemma.

@@
 
-$\CMon$ is closed monoidal, with the trivial one-element monoid $\One$ as terminal object and monoidal product
+$\CMon$ is monoidal closed, with the trivial one-element monoid $\One$ as terminal object and monoidal product
 $A \tensor B$ given by the Cartesian product $A \times B$. $\CMon$ is complete and cocomplete, inheriting all
@@
 
-Because $\CMon$ is closed monoidal, it is enriched over itself; every hom-object $\Hom{\CMon}{A}{B}$ is a
+Because $\CMon$ is monoidal closed, it is enriched over itself; every hom-object $\Hom{\CMon}{A}{B}$ is a
 commutative monoid where:
@@
 
-\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the
+\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ monoidal closed; then the
 hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched. The
@@
 
-\noindent \todo{Then there is an enriched version of the Yoneda embedding, namely a functor
-$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ analogous to $\Yoneda$
-which is itself $\mathscr{V}$-enriched.}
+\noindent We then have an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
+\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which sends:
+
+\begin{itemize}
+\item each object $X$ in $\cat{C}$ to the enriched hom-functor $\Yoneda_{\mathscr{V}}(X) =
+\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$
+\item each morphism $f: X \to Y$ in $\cat{C}$ to the natural transformation $\Yoneda_{\mathscr{V}}(f)$ where
+$\Yoneda_{\mathscr{V}}(f)_Z$ is again $(f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$
+\end{itemize}
+
+\noindent and an enriched Yoneda lemma which says that for any $\mathscr{V}$-enriched presheaf $F$ and any
+object $X \in \cat{C}$ the hom-object of natural transformations between the $\mathscr{V}$-enriched presheaf
+$\Hom{\cat{C}}{-}{X}$ and $F$, which also lives in $\mathscr{V}$, is isomorphic to $F(X)$:
+
+\begin{lemma}
+Suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ monoidal closed. For any $F: \cat{C}^\op \to
+\mathscr{V}$ and any object $X$ in $\cat{C}$:
+\[\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{F} \iso F(X)\]
+\end{lemma}
 

### 2025-04-16 appendix/notes/notation.tex
Standardise on X,Y,Z,W for objects.

\section{Notation}

In a category with coproducts or products we write:
\begin{itemize}
\item $\coprodM{f}{g}: X + Y \to Z$ for the coproduct of morphisms $f: X \to Z$ and $g: Y \to Z$
\item $\prodM{f}{g}: X \to Y \times Z$ for the product of morphisms $f: X \to Y$ and $g: X \to Z$
\item $(f,g): X \times Y \to Z \times W$ for $\prodM{f \comp \proj_1}{g \comp \proj_2}$ where $f: X \to Z$ and $g: Y \to W$.
\end{itemize}

### 2025-04-16 appendix/notes/cmon-enriched.tex
Standardise on X,Y,Z,W for objects.

@@
 \begin{definition}[Commutative monoid]
-A \emph{commutative monoid} $A = (A, \varepsilon, \bullet)$ is a set $A$ equipped with distinguished element
-$\varepsilon \in A$ called the \emph{unit} and associative binary operation $\bullet$ satisfying $\varepsilon
-\bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in A$.
+A \emph{commutative monoid} $X = (X, \bullet, \varepsilon)$ is a set $X$ equipped with distinguished element
+$\varepsilon \in X$ called the \emph{unit} and associative binary operation $\bullet: X^2 \to X$ satisfying
+$\varepsilon \bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in X$.
 \end{definition}
 
-A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
+A commutative monoid homomorphism from $X$ to $Y$ is any function $f: X \to Y$ preserving $\varepsilon$ and
 $\bullet$.
 
-\subsubsection{Examples of commutative monoids}
+\subsubsection{Category of commutative monoids}
+
+\begin{definition}[Category $\CMon$]
+Define $\CMon$ to be the category which has as objects $X$ all commutative monoids and as morphisms $f: X \to
+Y$ all commutative monoid homomorphisms.
+\end{definition}
+
+$\CMon$ is monoidal closed, with the trivial one-element monoid $\One$ as terminal object and monoidal product
+$A \tensor B$ given by the Cartesian product $A \times B$. $\CMon$ is complete and cocomplete, inheriting all
+limits and colimits from $\Set$.
+
+\subsubsection{Example of commutative monoid}
 
@@
 \label{def:cmon-enriched:bounded-semilattice}
-A \emph{bounded semilattice} $A = (A, \varepsilon, \bullet)$ is a commutative monoid where $\bullet$ is
+A \emph{bounded semilattice} $X = (X, \bullet, \varepsilon)$ is a commutative monoid where $\bullet$ is
 \emph{idempotent}, i.e.~satisfies $x \bullet x = x$.
@@
 is the bottom element. This therefore provides an algebraic characterisation of the usual (dual)
-order-theoretic notions of bounded meet semilattices $(X, \top, \meet)$ and bounded join semilattices $(X,
-\bot, \join)$.
+order-theoretic notions of bounded meet semilattices $(X, \meet, \top)$ and bounded join semilattices $(X,
+\join, \bot)$.
 
-\subsubsection{Category of commutative monoids}
-
-\begin{definition}[Category $\CMon$]
-Define $\CMon$ to be the category whose objects $A$ are the commutative monoids $A$ and morphisms $f: A \to B$
-are the commutative monoid homomorphisms.
+\begin{definition}[Category $\SemiLat$]
+Define $\SemiLat$ to be the category which has as objects all bounded semilattices and as morphisms all
+bounded semilattice homomorphisms.
 \end{definition}
-
-$\CMon$ is monoidal closed, with the trivial one-element monoid $\One$ as terminal object and monoidal product
-$A \tensor B$ given by the Cartesian product $A \times B$. $\CMon$ is complete and cocomplete, inheriting all
-limits and colimits from $\Set$.
 
@@
 \begin{enumerate}
-\item Every hom-object $\Hom{\cat{C}}{A}{B}$ is a commutative monoid of morphisms; we write $\zero_{A,B}$
-(zero morphism) for the unit, and $+_{A,B}$ (addition of morphisms) for the binary operation.
-\item Composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{B}{C} \tensor
-\Hom{\cat{C}}{A}{B} \to \Hom{\cat{C}}{A}{C}$ in $\CMon$ that preserve the additive structure in
-$\Hom{\cat{C}}{B}{C}$ and $\Hom{\cat{C}}{A}{B}$ separately:
+\item Every hom-object $\Hom{\cat{C}}{X}{Y}$ is a commutative monoid of morphisms; we write $\zero_{X,Y}$
+(zero morphism) for the unit, and $+_{X,Y}$ (addition of morphisms) for the binary operation (omitting the
+indices where implied by the context).
+\item Composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor
+\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in
+$\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
 
@@
 
-Because $\CMon$ is monoidal closed, it is enriched over itself; every hom-object $\Hom{\CMon}{A}{B}$ is a
-commutative monoid where:
+Because $\CMon$ is monoidal closed, it is enriched over itself; every hom-object $\Hom{\CMon}{X}{Y}$ is a
+commutative monoid with:
 
 \begin{enumerate}
-\item the unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
-\item the binary operation $+$ is pointwise addition of homomorphisms $(f + g)(a) = f(a) + g(a)$.
+\item unit $0_{X,Y} = \const(\varepsilon_Y)$, the constant homomorphism sending every element of $X$ to
+$\varepsilon_Y$;
+\item binary operation $+_{X,Y}$ given by pointwise addition of homomorphisms $(f + g)(x) = f(x) + g(x)$.
 \end{enumerate}

### 2025-04-16 appendix/notes/biproduct.tex
Standardise on X,Y,Z,W for objects.

@@
 \begin{definition}[Biproduct]
-Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is
-an object $A \biprod B$ of $C$ together with morphisms
+Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $X, Y$, the \emph{biproduct} of $X$ and $Y$ is
+an object $X \biprod Y$ of $C$ together with morphisms
 
@@
 \begin{tikzcd}
-   A \arrow[r, "\biinj_A", shift left] &
-   A \biprod B \arrow[l, "\biproj_A", shift left] \arrow[r, "\biproj_B"', shift right] &
-   B \arrow[l, "\biinj_B"', shift right]
+   X \arrow[r, "\biinj_X", shift left] &
+   X \biprod Y \arrow[l, "\biproj_X", shift left] \arrow[r, "\biproj_Y"', shift right] &
+   Y \arrow[l, "\biinj_Y"', shift right]
 \end{tikzcd}
@@
 \begin{salign*}
-   \biproj_A \comp \biinj_A &= \id_A \\
-   \biproj_B \comp \biinj_A &= \zero_{A,B}
+   \biproj_X \comp \biinj_X &= \id_X \\
+   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
 \end{salign*}
@@
 \begin{salign*}
-   \biproj_B \comp \biinj_B &= \id_B \\
-   \biproj_A \comp \biinj_B &= \zero_{B,A}
+   \biproj_Y \comp \biinj_Y &= \id_Y \\
+   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
 \end{salign*}
@@
 \begin{salign*}
-\biinj_A \comp \biproj_A + \biinj_B \comp \biproj_B &= \id_{A \biprod B}
+(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
 \end{salign*}
@@
 
-We say a $\CMon$-enriched $\cat{C}$ is \emph{semi-additive} if it also has finite biproducts; having products
-or coproducts is sufficient.
+We call a $\CMon$-enriched $\cat{C}$ \emph{semi-additive} if it also has finite biproducts; having products or
+coproducts is sufficient.
 
@@
 \begin{itemize}
-\item If $\cat{C}$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
-\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
-\item If $\cat{C}$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
+\item If $\cat{C}$ has biproducts $(X \biprod Y, \biinj_X, \biinj_Y, \biproj_X, \biproj_Y)$ then $(X
+\biprod Y, \biproj_X, \biproj_Y)$ is a product and $(X \biprod Y, \biinj_X, \biinj_Y)$ is a coproduct.
+\item If $\cat{C}$ has binary products $(X \times Y, \proj_1, \proj_2)$ then:
 \begin{enumerate}
-\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
+\item $\prodM{0_{X,Y}}{0_{X,Z}} = 0_{X,Y \times Z}$
 \item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
-\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
-biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
-\prodM{0_{A,B}}{\id_B}$.
+\item $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$ is a biproduct with $\biinj_X =
+\prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
 \end{enumerate}

### 2025-04-16 appendix/notes/cmon-enriched.tex
Standardise on X,Y,Z,W for objects.

@@
 \end{enumerate}
+
+Suppose $F, G: \cat{D} \to \cat{C}$ are functors where $\cat{C}$ is $\CMon$-enriched. Then the hom-object
+$\Hom{\Func{\cat{C}}{\cat{D}}}{F}{G}$ of natural transformations between $F$ and $G$ has a zero $0_{F,G}$ and
+addition $\eta + \mu$ given component-wise as $(0_{F,G})_X = 0_{F(X),G(X)}$ and $(\eta + \mu)_X = \eta_X +
+\mu_X$.
 

### 2025-04-16 appendix/notes/fam.tex
Tweaks.

@@
 \begin{definition}[Category of $I$-indexed families of objects]
-For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category where:
+For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category which has:
 \begin{itemize}
-\item objects are the $I$-indexed families of objects of $\cat{C}$;
-\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $\cat{C}$ for any $i \in I$.
+\item as objects all $I$-indexed families of objects of $\cat{C}$;
+\item as morphisms from $X$ to $Y$ all families of morphisms $f_i: X_i \to Y_i$ in $\cat{C}$ for any $i \in
+I$.
 \end{itemize}
@@
 
-\begin{definition}[$\Fam(-,\cat{C})$ functor]
-$\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,\cat{C})$ and
-any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$.
+\begin{definition}
+Then define $\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ to be the functor which sends any set $I$ to
+$\Fam(I,\cat{C})$ and any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to
+\Fam(I,\cat{C})$.
 \end{definition}

### 2025-04-16 appendix/notes/fam.tex
Minor tweaks.

@@
 \begin{definition}[Reindexing]
-For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$ sends
-any $J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
-similarly for morphisms.
+For any $f: I \to J$ define the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to
+\Fam(I,\cat{C})$ which:
+\begin{itemize}
+\item sends any $J$-indexed family $X$ to the $I$-indexed family $\reindex{X}{f}$ where $\reindex{X}{f}_i =
+X_{f(i)}$ for any $i \in I$;
+\item sends any morphism $g: X \to Y$ between $J$-indexed families to the morphism $\reindex{g}{f}:
+\reindex{X}{f} \to \reindex{X}{g}$ where $\reindex{g}{f}_i = g_{f(i)}$.
+\end{itemize}
 \end{definition}

### 2025-04-16 appendix/notes/grothendieck.tex
Clarify 'nudge' notation.
Minor correction.

@@
 Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
-$F$ is the category where:
+$F$ is the category which has:
 \begin{itemize}
-\item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
-\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with morphisms
-$\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
+\item as objects $X$ all pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
+\item as morphisms $f: X \to Y$ all pairs of a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ and a
+morphism $\faml{f}: F(\idx{f})(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
 \end{itemize}
@@
 \noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
-$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.
+$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(\idx{f})(\faml{Y})$ in $F(\idx{X})$.
 
@@
 \end{itemize}
+Here $(x,-): \idx{Y} \to \idx{X} \times \idx{Y}$ denotes the function $\prodM{\const(x)}{\id_{\idx{Y}}}$ which
+pairs any $y \in \idx{Y}$ with $x \in \idx{X}$.
 \item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(\cat{C})$ with

### 2025-04-16 appendix/notes/grothendieck.tex
Minor tweaks..

@@
 
-\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
-$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(\idx{f})(\faml{Y})$ in $F(\idx{X})$.
+\noindent When $F$ is contravariant, a morphism $f: X \to Y$ in $\Grothendieck{\cat{C}}{F}$ is a morphism
+$\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to
+F(\idx{f})(\faml{Y})$ in $F(\idx{X})$. \todo{Might want to unpack this a bit.}
 

### 2025-04-16 appendix/notes/cmon-enriched.tex
Enriched functors have a map on objects too.

@@
 For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
-$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a family of
-morphisms $F_{X,Y}: \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$
-that makes the following diagrams in $\mathscr{V}$ commute:
+$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a map sending
+every object $X$ of $\cat{C}$ to an object $F(X)$ of $\cat{D}$, plus a family of morphisms $F_{X,Y}:
+\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ making the
+following commute:
 
@@
 \noindent These generalise the usual functor laws to a setting where the hom-objects are not necessarily sets
-and thus do not support an elementwise formulation.
+and thus do not support an elementwise formulation, although here we are only interested in the situation
+where $\mathscr{V}$ is concrete.
 

### 2025-04-16 appendix/notes/cmon-enriched.tex
Enriched functors have a map on objects too.
Some progress on enriched Yoneda embedding.

@@
 \end{itemize}
-in addition to the usual enriched functor properties in \defref{cmon-enriched:enriched-functor}.
+in addition to the usual enriched functor properties from \defref{cmon-enriched:enriched-functor}.
 
@@
 \noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ monoidal closed; then the
-hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched. The
-
-The following diagram commutes because precomposition with the identity is the identity on morphisms:
+hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched, sending
+objects $Y$ to the hom-object $\Hom{\cat{C}}{Y}{X}$ in $\mathscr{V}$, and providing a family of morphisms
+$\Hom{\cat{C}}{-}{X}_{Y,Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{X}}{\Hom{\cat{C}}{Z}{X}}$ for any $X, Y$ in
+$\cat{C}$. The following diagram commutes because precomposition with the identity is the identity on
+morphisms:
 
@@
 
-\noindent and the following by the associativity of composition:
+\noindent and the following commutes by the associativity of composition:
 
@@
 
-\noindent We then have an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
-\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which sends:
+\noindent There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
+\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ sending objects $X$ to the enriched hom-functor
+$\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ and providing a family of
+morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to
+\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{\Hom{\cat{C}}{-}{Y}}$.
 
-\begin{itemize}
-\item each object $X$ in $\cat{C}$ to the enriched hom-functor $\Yoneda_{\mathscr{V}}(X) =
-\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$
-\item each morphism $f: X \to Y$ in $\cat{C}$ to the natural transformation $\Yoneda_{\mathscr{V}}(f)$ where
-$\Yoneda_{\mathscr{V}}(f)_Z$ is again $(f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$
-\end{itemize}
-
-\noindent and an enriched Yoneda lemma which says that for any $\mathscr{V}$-enriched presheaf $F$ and any
-object $X \in \cat{C}$ the hom-object of natural transformations between the $\mathscr{V}$-enriched presheaf
-$\Hom{\cat{C}}{-}{X}$ and $F$, which also lives in $\mathscr{V}$, is isomorphic to $F(X)$:
+The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in \cat{C}$
+the hom-object of natural transformations between the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$
+and $F$, which also lives in $\mathscr{V}$, is isomorphic to $F(X)$:
 

### 2025-04-16 appendix/notes/cmon-enriched.tex
More tweaks.

@@
 \noindent There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
-\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ sending objects $X$ to the enriched hom-functor
-$\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ and providing a family of
-morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to
-\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{\Hom{\cat{C}}{-}{Y}}$.
+\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$, which is itself $\mathscr{V}$-enriched, sending
+objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}:
+\cat{C}^\op \to \mathscr{V}$ and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations
+$\Yoneda_{\mathscr{V}}(f): \Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as
+post-composition with $f$.
 
-The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in \cat{C}$
-the hom-object of natural transformations between the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$
-and $F$, which also lives in $\mathscr{V}$, is isomorphic to $F(X)$:
+The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in
+\cat{C}$, the hom-object of $\mathscr{V}$-natural transformations (which also lives in $\mathscr{V}$) between
+the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ and $F$ is isomorphic to $F(X)$:
 

### 2025-04-18 appendix/notes/cmon-enriched.tex
Preliminaries.
Some cleanup/better org.

@@
 \label{sec:cmon-enriched}
-
-\begin{definition}[Commutative monoid]
-A \emph{commutative monoid} $X = (X, \bullet, \varepsilon)$ is a set $X$ equipped with distinguished element
-$\varepsilon \in X$ called the \emph{unit} and associative binary operation $\bullet: X^2 \to X$ satisfying
-$\varepsilon \bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in X$.
-\end{definition}
-
-A commutative monoid homomorphism from $X$ to $Y$ is any function $f: X \to Y$ preserving $\varepsilon$ and
-$\bullet$.
-
-\subsubsection{Category of commutative monoids}
-
-\begin{definition}[Category $\CMon$]
-Define $\CMon$ to be the category which has as objects $X$ all commutative monoids and as morphisms $f: X \to
-Y$ all commutative monoid homomorphisms.
-\end{definition}
-
-$\CMon$ is monoidal closed, with the trivial one-element monoid $\One$ as terminal object and monoidal product
-$A \tensor B$ given by the Cartesian product $A \times B$. $\CMon$ is complete and cocomplete, inheriting all
-limits and colimits from $\Set$.
-
-\subsubsection{Example of commutative monoid}
-
-\begin{definition}[Bounded semilattice]
-\label{def:cmon-enriched:bounded-semilattice}
-A \emph{bounded semilattice} $X = (X, \bullet, \varepsilon)$ is a commutative monoid where $\bullet$ is
-\emph{idempotent}, i.e.~satisfies $x \bullet x = x$.
-\end{definition}
-
-\noindent The idempotence of $\bullet$ (together with commutativity and associativity) induces a partial order
-$\le_\bullet$ on $X$, with $x \le_{\bullet} y \iff x \bullet y = x$. With respect to this partial order, $x
-\bullet y$ is the greatest lower bound (meet) of $x$ and $y$ and $\varepsilon$ is the top element; with
-respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ and $\varepsilon$
-is the bottom element. This therefore provides an algebraic characterisation of the usual (dual)
-order-theoretic notions of bounded meet semilattices $(X, \meet, \top)$ and bounded join semilattices $(X,
-\join, \bot)$.
-
-\begin{definition}[Category $\SemiLat$]
-Define $\SemiLat$ to be the category which has as objects all bounded semilattices and as morphisms all
-bounded semilattice homomorphisms.
-\end{definition}
-
-\subsubsection{$\CMon$-enriched category}
 
@@
 \subsubsection{$\CMon$-enriched presheaves}
-
-\begin{definition}[$\mathscr{V}$-enriched functor]
-\label{def:cmon-enriched:enriched-functor}
-For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
-$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a map sending
-every object $X$ of $\cat{C}$ to an object $F(X)$ of $\cat{D}$, plus a family of morphisms $F_{X,Y}:
-\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$ making the
-following commute:
-
-\begin{center}
-\begin{tikzcd}[column sep=1.8cm]
-I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "F_{X,X}"] \\
-I \arrow[r, "\id_{F(X)}"'] & \Hom{\cat{D}}{F(X)}{F(X)}
-\end{tikzcd}
-\hspace{5mm}
-\begin{tikzcd}[column sep=2.6cm]
-   \Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"']
-   & \Hom{\cat{C}}{X}{Z} \arrow[d, "F_{X,Z}"] \\
-   \Hom{\cat{D}}{F(Y)}{F(Z)} \tensor \Hom{\cat{D}}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] &
-   \Hom{\cat{D}}{F(X)}{F(Z)}
-\end{tikzcd}
-\end{center}
-\end{definition}
-
-\noindent These generalise the usual functor laws to a setting where the hom-objects are not necessarily sets
-and thus do not support an elementwise formulation, although here we are only interested in the situation
-where $\mathscr{V}$ is concrete.
-
-\begin{definition}[Category of enriched functors]
-Write $\Func{\cat{C}}{\cat{D}}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from
-$\cat{C}$ to $\cat{D}$.
-\end{definition}
 
@@
 \end{lemma}
-
-% First we note that the hom-functor $\Hom{C}{-}{X}: C \to \CMon$ is $\CMon$-enriched because composition is
-% bilinear and so $(- \comp f)$ preserves the additive structure of morphisms.

### 2025-04-18 appendix/notes/preliminaries.tex
Preliminaries.
Some cleanup/better org.

--- 2025-04-16 appendix/notes/notation.tex
@@
-\section{Notation}
+\subsection{Preliminaries}
 
-In a category with coproducts or products we write:
+\subsubsection{Notation}
+
+\paragraph{Coproduct and product of morphisms.} In a category with coproducts $X + Y$ or products $X \times
+Y$, we write:
 \begin{itemize}
@@
 \end{itemize}
+
+\paragraph{Constant function.} Write $\const(y): X \to Y$ for the function which sends any $x \in X$ to $y \in
+Y$.
+
+\paragraph{Functor category.} Write $\Func{\cat{C}}{\cat{D}}$ for the category of functors $\cat{C} \to
+\cat{D}$.
+
+\subsubsection{$\mathscr{V}$-categories (for concrete $\mathscr{V}$)}
+
+Let $\mathscr{V}$ range over monoidal categories $(\mathscr{V}, \tensor, I)$ that are \emph{concrete},
+i.e.~have a forgetful functor to $\Set$. In any $\mathscr{V}$-enriched category (or
+\emph{$\mathscr{V}$-category}) $\cat{C}$, the hom-objects $\Hom{\cat{C}}{X}{Y}$ have underlying sets and the
+functoriality and naturality conditions for $\mathscr{V}$-functors and $\mathscr{V}$-natural transformations
+can be formulated elementwise (as per the $\Set$-enriched case) rather than more generally in terms of
+morphisms from the tensor unit $I$. \todo{Be explicit about composition being a family of morphisms
+$\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\mathscr{V}$.}
+
+\begin{definition}[$\mathscr{V}$-functor]
+\label{def:cmon-enriched:enriched-functor}
+For any $\mathscr{V}$-enriched categories $\cat{C}$ and $\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F:
+\cat{C} \to \cat{D}$ on morphisms is a map sending every object $X$ of $\cat{C}$ to an object $F(X)$ of
+$\cat{D}$, plus a family of morphisms $F_{X,Y}: \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between
+hom-objects in $\mathscr{V}$, obeying the usual functor laws.
+\end{definition}
+
+\noindent The action of a $\mathscr{V}$-functor on morphisms thus necessarily preserves the
+$\mathscr{V}$-structure on the hom-objects, since it is is given by a morphism in $\mathscr{V}$. The notion of
+$\mathscr{V}$-natural transformation is defined similarly.
+
+\begin{definition}[$\mathscr{V}$-natural transformation]
+For any $\mathscr{V}$-enriched categories $\cat{C}$ and $\cat{D}$ and $\mathscr{V}$-functors $F, G: \cat{C}
+\to \cat{D}$, a {$\mathscr{V}$-natural transformation} $\eta: F \naturalto G$ is a family of morphisms
+$\eta_X: F(X) \to G(X)$ in $\cat{D}$ satisfying the usual naturality square for every $f: X \to Y$ in
+$\cat{C}$. \todo{This is sufficient because $\comp_{X,Y,Z}$ as well as $F$ and $G$ respects the
+$\mathscr{V}$-structure?}
+\end{definition}
+
+% making the following commute:
+%
+% \begin{center}
+% \begin{tikzcd}[column sep=1.8cm]
+% I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "F_{X,X}"] \\
+% I \arrow[r, "\id_{F(X)}"'] & \Hom{\cat{D}}{F(X)}{F(X)}
+% \end{tikzcd}
+% \hspace{5mm}
+% \begin{tikzcd}[column sep=2.6cm]
+%    \Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"']
+%    & \Hom{\cat{C}}{X}{Z} \arrow[d, "F_{X,Z}"] \\
+%    \Hom{\cat{D}}{F(Y)}{F(Z)} \tensor \Hom{\cat{D}}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] &
+%    \Hom{\cat{D}}{F(X)}{F(Z)}
+% \end{tikzcd}
+% \end{center}
+%
+% \noindent These generalise the usual functor laws to a setting where the hom-objects are not necessarily sets
+% and thus do not support an elementwise formulation, although here we are only interested in the situation
+% where $\mathscr{V}$ is concrete.
+
+\begin{definition}[Category of enriched functors]
+Write $\Func{\cat{C}}{\cat{D}}_{\mathscr{V}}$ for the category of $\mathscr{V}$-enriched functors from
+$\cat{C}$ to $\cat{D}$.
+\end{definition}

### 2025-04-18 appendix/notes/cmon-enriched.tex
Some cleanup/better org.
Should probably simplify bit about hom-functor being V-enriched (to state properties elementwise).

@@
 
-\noindent There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
-\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$, which is itself $\mathscr{V}$-enriched, sending
-objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}:
-\cat{C}^\op \to \mathscr{V}$ and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations
+\noindent \todo{Simplify because we can state the functor laws elementwise if $\mathscr{V}$ is concrete?}
+
+There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}: \cat{C} \to
+\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$, which is itself $\mathscr{V}$-enriched, sending objects $X$ to
+the $\mathscr{V}$-enriched hom-functor $\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to
+\mathscr{V}$ and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations
 $\Yoneda_{\mathscr{V}}(f): \Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as
-post-composition with $f$.
-
-The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in
-\cat{C}$, the hom-object of $\mathscr{V}$-natural transformations (which also lives in $\mathscr{V}$) between
-the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ and $F$ is isomorphic to $F(X)$:
+post-composition with $f$. The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and
+any object $X \in \cat{C}$, the hom-object of $\mathscr{V}$-natural transformations (which also lives in
+$\mathscr{V}$) between the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ and $F$ is isomorphic to
+$F(X)$:
 

### 2025-04-18 appendix/notes/cmon-enriched.tex
Cleanup of enriched functors and enriched Yoneda embedding (when enriching category is concrete).

--- 2025-04-18 appendix/notes/cmon-enriched.tex
@@
 
-\subsubsection{$\CMon$-enriched Yoneda embedding}
+\subsubsection{Enriched Yoneda embedding}
 
-Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
-\cat{C} \to \Func{\cat{C}^\op}{\Set}$ defined as follows:
+Suppose a monoidal category $\mathscr{V}$ and a $\mathscr{V}$-enriched category $\cat{C}$, again restricted to
+the case where $\mathscr{V}$ is concrete.
+
+\begin{definition}[Hom-functor]
+Define the contravariant \emph{hom-functor} $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ sending
 \begin{itemize}
-\item on objects: $\Yoneda(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \Set$, the contravariant hom-functor
-sending:
+\item any object $Y$ in $\cat{C}$ to the hom-object $\Hom{\cat{C}}{Y}{X}$ in $\mathscr{V}$;
+\item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
+\Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$ on the underlying hom-sets.
+\end{itemize}
+\end{definition}
+
+\begin{definition}[Yoneda embedding]
+The \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda: \cat{C} \to
+\Func{\cat{C}^\op}{\mathscr{V}}$ sending
 \begin{itemize}
-   \item any object $Y$ in $\cat{C}$ to the hom-set $\Hom{\cat{C}}{Y}{X}$;
-   \item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
-   \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$.
+\item any object $X$ to the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$;
+\item any morphism $f: X \to Y$ to the $\mathscr{V}$-natural transformation $\Hom{\cat{C}}{-}{X} \naturalto
+\Hom{\cat{C}}{-}{Y}$ where $\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
 \end{itemize}
-\item on morphisms: for any $f: X \to Y$ in $\cat{C}$, $\Yoneda(f)$ is the natural transformation where
-$\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
-\end{itemize}
+\end{definition}
 
-\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ monoidal closed; then the
-hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched, sending
-objects $Y$ to the hom-object $\Hom{\cat{C}}{Y}{X}$ in $\mathscr{V}$, and providing a family of morphisms
-$\Hom{\cat{C}}{-}{X}_{Y,Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{X}}{\Hom{\cat{C}}{Z}{X}}$ for any $X, Y$ in
-$\cat{C}$. The following diagram commutes because precomposition with the identity is the identity on
-morphisms:
+\noindent Now suppose $\mathscr{V}$ is monoidal closed. Then the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op
+\to \mathscr{V}$ is itself $\mathscr{V}$-enriched, providing a family of morphisms $\Hom{\cat{C}}{-}{X}_{Y,Z}:
+\Hom{\cat{C}}{Y}{Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Z}{X}}{\Hom{\cat{C}}{Y}{X}}$ in $\mathscr{V}$ for any
+$Y, Z$ in $\cat{C}$. Moreover there exists an enriched Yoneda embedding, namely a functor
+$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which is itself
+$\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
+and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations $\Yoneda_{\mathscr{V}}(f):
+\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as post-composition with $f$.
 
-\begin{center}
-\begin{tikzcd}[column sep=1.8cm]
-I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "\Hom{\cat{C}}{-}{W}_{X,X}"] \\
-I \arrow[r, "\id_{\Hom{\cat{C}}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{X}{W}}\
-\end{tikzcd}
-\end{center}
-
-\noindent and the following commutes by the associativity of composition:
-
-\begin{center}
-\begin{tikzcd}[column sep=2.7cm]
-   \Hom{\cat{C}}{Z}{Y} \tensor \Hom{\cat{C}}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d,
-   "\Hom{\cat{C}}{-}{W}_{Y,Z}\,\tensor\,\Hom{\cat{C}}{-}{W}_{X,Y}"'] & \Hom{\cat{C}}{Z}{X} \arrow[d,
-   "\Hom{\cat{C}}{-}{W}_{X,Z}"] \\
-   \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{W}}{\Hom{\cat{C}}{Z}{W}} \tensor
-   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Y}{W}} \arrow[r,
-   "\comp_{\Hom{\cat{C}}{X}{W},\Hom{\cat{C}}{Y}{W},\Hom{\cat{C}}{Z}{W}}"'] &
-   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Z}{W}}
-\end{tikzcd}
-\end{center}
-
-\noindent There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
-\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$, which is itself $\mathscr{V}$-enriched, sending
-objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}:
-\cat{C}^\op \to \mathscr{V}$ and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations
-$\Yoneda_{\mathscr{V}}(f): \Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as
-post-composition with $f$.
-
-The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in
-\cat{C}$, the hom-object of $\mathscr{V}$-natural transformations (which also lives in $\mathscr{V}$) between
-the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ and $F$ is isomorphic to $F(X)$:
+The \emph{enriched Yoneda lemma} states that, for any $\mathscr{V}$-enriched presheaf $F$ and any object $X
+\in \cat{C}$, there is an isomorphism between $F(X)$ and the hom-object (also living in $\mathscr{V}$) of
+$\mathscr{V}$-natural transformations from the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ to $F$:
 
@@
 \end{lemma}
+
+% The following diagram commutes because precomposition with the identity is the identity on morphisms:
+%
+% \begin{center}
+% \begin{tikzcd}[column sep=1.8cm]
+% I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "\Hom{\cat{C}}{-}{W}_{X,X}"] \\
+% I \arrow[r, "\id_{\Hom{\cat{C}}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{X}{W}}\
+% \end{tikzcd}
+% \end{center}
+%
+% \noindent and the following commutes by the associativity of composition:
+%
+% \begin{center}
+% \begin{tikzcd}[column sep=2.7cm]
+%    \Hom{\cat{C}}{Z}{Y} \tensor \Hom{\cat{C}}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d,
+%    "\Hom{\cat{C}}{-}{W}_{Y,Z}\,\tensor\,\Hom{\cat{C}}{-}{W}_{X,Y}"'] & \Hom{\cat{C}}{Z}{X} \arrow[d,
+%    "\Hom{\cat{C}}{-}{W}_{X,Z}"] \\
+%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{W}}{\Hom{\cat{C}}{Z}{W}} \tensor
+%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Y}{W}} \arrow[r,
+%    "\comp_{\Hom{\cat{C}}{X}{W},\Hom{\cat{C}}{Y}{W},\Hom{\cat{C}}{Z}{W}}"'] &
+%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Z}{W}}
+% \end{tikzcd}
+% \end{center}

### 2025-04-18 appendix/notes/cmon-enriched.tex
Cleanup of enriched functors and enriched Yoneda embedding (when enriching category is concrete).

@@
 $\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
-and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations $\Yoneda_{\mathscr{V}}(f):
-\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as post-composition with $f$.
+and providing a family of morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to $ to
+$\mathscr{V}$-natural transformations $\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each
+component as post-composition with $f$.
 

### 2025-04-23 main/introduction.tex
[ paper ] some words introducing the main ideas

@@
 
-\cite{karvonen2020} % just to generate a bibliography for now
+Provenance tracking aims to answer questions like ``where did these data come from'', ``why are these data in the output'' and ``how were these data computed''.
+Program slicing is a collection of techniques for provenance tracking that attempts to take a run of a program and and areas of interest in the output and turn them into the subset of the input and the program that were responsible for generating those specific outputs.
+
+Provenance tracking is important because in order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome.
+Existing approaches are often tied to particular programming languages or implementations.
+In this paper we develop a general categorical approach.
+
+\note{We will use biproducts, and use some results from \cite{karvonen2020}.} % just to generate a bibliography for now
+
+\subsection{Galois Program Slicing}
+
+Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications (FIXME: refs).
+Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{fluid} that allows interactive exploration of programmatically generated visualisations.
+
+At a high level, Galois Program Slicing assumes that, for each possible value that may be input or output by a program, there exists a lattice of {\em approximations} of that value.
+For a particular run of a program that takes input $x$ and produces output $y$, we also get a Galois connection between the lattice of approximations of $x$ and the lattice of approximations of $y$.
+The right half of the Galois connection is the ``forward direction'' taking approximations of the input to approximations of the output; the left half of the Galois connection is the ``backward direction'' that takes approximations of the output to the least (i.e., most approximate) approximation of the input that gives rise to this output approximation.
+
+\begin{example}
+  \todo{an example}
+\end{example}
+
+Previous work on Galois program slicing uses an operational semantics of programs that generates a trace of each execution which can be interpreted after the fact to compute the Galois connection described above.
+This becomes {\em program slicing} by including the source code of the program as part of the input, so that, in the backward direction, the least approximation of the input required for an output approximation includes the least part of the program required.
+
+\subsection{Galois Program Slicing and Automatic Differentiation}
+
+There is a close analogy between Galois Program Slicing and Automatic Differentiation for differentiable programs.
+We have already hinted at this in the description above, but let us now make it explicit.
+
+\begin{itemize}
+\item For Galois Program Slicing, we assume that every value has an associated lattice of {\em approximations}.
+For differentiable programs, every point has an associated vector space of {\em tangents}.
+\item For Galois Program Slicing, every program has an associated forward approximation map that takes approximations forward from the input to the output.
+This map {\em preserves meets}.
+For differentiable programs, every program has a forward derivative that takes tangents of the input to tangents of the output.
+The forward derivative map is {\em linear}, so it preserves addition of tangents and the zero tangent.
+\item For Galois Program Slicing, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input.
+This map {\em preserves joins}.
+For differentiable programs, every program has a reverse derivative that takes tangents of the output to tangents of the input.
+This map is again {\em linear}.
+\item For Galois Program Slicing, the forward and backward approximation maps are related by being a Galois connection.
+For differentiable programming, the forward and reverse derivatives are related by being each others' transpose.
+\end{itemize}
+
+Given this close connection between Galois program slicing and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model Galois program slicing.
+This will enable us to generalise and expand the scope of Galois program slicing to act as a foundation for data provenance in a wide range of computational settings.

### 2025-04-23 main/introduction.tex
[ paper ] an example to explain the basic idea

@@
 \begin{example}
-  \todo{an example}
+  The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then to compute the sum of the numbers:
+  \begin{displaymath}
+    \begin{array}{l}
+      \mathrm{query} :: \mathrm{Label} \to [(\mathrm{Label}, \mathrm{Int})] \to \mathrm{Int} \\
+      \mathrm{query}\,l\,\mathit{db} = \mathrm{sum}\,[ n \mid (l',n) \leftarrow \mathit{db}, l \equiv l' ]
+    \end{array}
+  \end{displaymath}
+  With $\mathit{db} = [(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$, we will have $\mathit{query}\,\mathsf{a}\,\mathit{db}$ and $\mathit{query}\,\mathsf{b}\,\mathit{db}$ both evaluating to $1$.
+
+  Now suppose that for a given run of the program, we are interested in which of the numerical parts of the input are used to compute the output for the query parameters $l = \mathsf{a}$ and $l = \mathsf{b}$.
+We can use Galois Program Slicing to do this.
+We arrange for the approximations of the input to form the following lattice, where the actual piece of data is at the top and information lost by approximation is represented by $\bot$s:
+  \begin{center}
+    \begin{tikzpicture}
+      \node (top) at (0,0) {$[(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$};
+      % row 2
+      \node [below of=top] (ioi) {$[(\mathsf{a}, 0), (\mathsf{b}, \bot), (\mathsf{a}, 1)]$};
+      \node [left of=ioi,xshift=-3cm] (oii) {$[(\mathsf{a}, \bot), (\mathsf{b}, 1), (\mathsf{a}, 1)]$};
+      \node [right of=ioi,xshift=3cm] (iio) {$[(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, \bot)]$};
+      % row 3
+      \node [below of=ioi] (oio) {$[(\mathsf{a}, \bot), (\mathsf{b}, 1), (\mathsf{a}, \bot)]$};
+      \node [left of=oio,xshift=-3cm] (ooi) {$[(\mathsf{a}, \bot), (\mathsf{b}, \bot), (\mathsf{a}, 1)]$};
+      \node [right of=oio,xshift=3cm] (ioo) {$[(\mathsf{a}, 0), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]$};
+      % row 4
+      \node [below of=oio] (bot) {$[(\mathsf{a}, \bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]$};
+
+      % links
+      \draw (top) -- (ioi);
+      \draw (top) -- (oii);
+      \draw (top) -- (iio);
+      \draw (ioi) -- (ooi);
+      \draw (ioi) -- (ioo);
+      \draw (oii) -- (ooi);
+      \draw (oii) -- (oio);
+      \draw (iio) -- (ioo);
+      \draw (iio) -- (oio);
+      \draw (ooi) -- (bot);
+      \draw (oio) -- (bot);
+      \draw (ioo) -- (bot);
+    \end{tikzpicture}
+  \end{center}
+  The output approximation lattice looks like this, where $1$ is the actual data point that was returned in both runs of the program, and $\bot$ indicates that we are approximating this piece of data away:
+  \begin{center}
+    \begin{tikzpicture}
+      \node (top) at (0,0) {$1$};
+      \node [below of=top] (bot) {$\bot$};
+      \draw (top) -- (bot);
+    \end{tikzpicture}
+  \end{center}
+  These are not the only choices of approximation lattices that we could have taken.
+For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not any of the other data or the overall structure.
+However, other choices are also useful.
+Indeed, one of the aims of this work is to clarify how to choose approximation structure approriate for different tasks by use of type information.
+We elaborate on this further in Section \todo{XYZ}.
+
+  Galois program slicing associates with each run of the program a Galois connection telling us how the inputs and outputs are related in that run.
+The backwards portion $\partial (\mathit{query}\,l)_r$ tells us, given an approximation of the output what the least approximation of the input is needed to generate that output.
+In the case of the two runs considered in this example, if we say we are not interested in the output by feeding in the least approximation $\bot$, then we find that we only need the least approximation of the input:
+  \begin{displaymath}
+    \partial (\mathit{query}\,l)_r(\bot) = [(\mathsf{a},\bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]
+  \end{displaymath}
+  for both $l = \mathsf{a}$ and $l = \mathsf{b}$.
+If instead we take the greatest approximation of the output (i.e., the output ``$1$'' itself), then the two query runs' backwards approximations return different results:
+  \begin{displaymath}
+    \begin{array}{l}
+      \partial (\mathit{query}\,\mathsf{a})_r(1) = [(\mathsf{a},0), (\mathsf{b},\bot), (\mathsf{a},1)] \\
+      \partial (\mathit{query}\,\mathsf{b})_r(1) = [(\mathsf{a},\bot), (\mathsf{b},1), (\mathsf{a},\bot)]
+    \end{array}
+  \end{displaymath}
+  Pieces of the input that were {\em not} used are replaced by $\bot$.
+As we expect, the run of the query with label $\mathsf{a}$ depends on the entries in the database labelled with $\mathsf{a}$, and likewise for the run with label $\mathsf{b}$.
+
+  In a simple query like this, it is easy to work out the dependency relationship between the input and output.
+However, the benefit of Galois Program Slicing is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output is.
+Moreover, by changing what we mean by ``approximation'' we can compute a range of different information about a program.
 \end{example}

### 2025-04-23 main/introduction.tex
[ paper ] explain the forward slicing too

@@
 
+  In this case, the forwards portion of the Galois connection tells us for each approximation of the input whether or not it is sufficient to compute the output.
+In this example, if we provide insufficient data to compute the output, then we will get an underapproximated output.
+For example, $\partial (\mathit{query}\, \mathsf{a})_f([(\mathsf{a},0),(\mathsf{b},\bot),(\mathsf{a},\bot)]) = \bot$ because we need all the values associated with the label $\mathsf{a}$ to compute their sum.
+
   In a simple query like this, it is easy to work out the dependency relationship between the input and output.

### 2025-04-23 main/introduction.tex
[ paper ] todo note
[ paper ] very rough outline of some content

@@
 
+\todo{Explain what the plan for the paper is}
+
+\todo{Perhaps more elaboration on the applications of program slicing}
+
 \subsection{Galois Program Slicing and Automatic Differentiation}
 
+\todo{explain why we want a more abstract view on galois program slicing}
 There is a close analogy between Galois Program Slicing and Automatic Differentiation for differentiable programs.
@@
 This will enable us to generalise and expand the scope of Galois program slicing to act as a foundation for data provenance in a wide range of computational settings.
+
+\subsection{Outline and Contributions}
+
+Our primary contribution in this article is to elucidate Galois Program Slicing by relating it to differentiable programming and automatic differentiation.
+In doing so, we aim to expand the applicability of program slicing, and to potentially transfer efficient automatic differentiation implementation techniques from automatic differentiation to program slicing and provenance tracking.
+
+\begin{enumerate}
+\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for Galois Program Slicing,
+\item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of Galois Program Slicing.
+\item With the benefit of our abstract setting, we can relate Galois Program Slicing to other parts of denotational semantics.
+In particular, we show that there is a close connection with Stable Domain Theory, a proposed solution to capturing sequentiality by recording more intensional information about programs' sensitivity to approximations.
+\item We also relate Galois Program Slicing to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
+\end{enumerate}

### 2025-04-23 main/formal-structures.tex
[ paper ] formal-structures bit

\section{Formal Structures for Generalised Automatic Differentiation}

\note{Probably a better way: (1) introduce the category of lattices and galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}

We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others \cite{nunes2023}. Here follows a rough overview of this approach and how we have applied it to Galois program slicing. \todo{Final version shouldn't be ``rough''!}

We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}. Not only does this mean that the constructions and proofs have been checked, the construction is executable and can be runn on small examples. We are also planning to use Agda's JavaScript backend to produce a demo.\todo{Do this?}

The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$. An object of the Grothendieck construction $\int T$ is a pair $(X \in \cat{C}, \partial X \in T(X)$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces. The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling. The maps in the indexed category are the maps of tangents or approximations, being derivatives and galois connections respectively.

In the case of differentiable programs and automatic differentiation, a basic model can be constructed by taking $\cat{C}$ to be $\Set$ and $T(X) = X \to \FinVect$: indexed collections of finite dimensional vector spaces, whose elements are interpreted as tangents at the given point, with linear maps. The Grothendieck category $\int T$ has objects that are pairs of a set $X$ and for each $x \in X$ a tangent vector space $\partial X(x)$. Morphisms are pairs of maps of points to points, accompanied by linear maps of tangents at those points. If we carefully choose an initial set of maps, we can read these maps as derivatives (note that nothing in the Grothendieck construction says that they must be derivatives!). Composition in the Grothendieck category is exactly the {\em chain rule} for composing derivatives, so we at least do know that if we start with derivatives on basic operations, then composing them will retain this property.

For Galois program slicing, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.

In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$. As is well-known\todo{reference}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$. In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the cartesian products from $\Set$. When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed. If the coproducts are actually {\em biproducts}, then $\Fam(\cat{X})$ is cartesian closed, and we can interpret higher-order programs.

\todo{Preceeding paragraph needs references to later in the document where these constructions are detailed in full.}

Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of them having products and being enriched in commutative monoids\todo{fwd ref}. However, neither of them have all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$). Neither of these are ideal. In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost. In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model. Moreover, in Agda, there are issues with predicativity when considering complete lattices.\todo{be more specific}

Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using Cmon-enriched presheaves.\todo{elaborate.}

### 2025-04-28 main/introduction.tex
Galois slicing citations.

@@
 
-\note{We will use biproducts, and use some results from \cite{karvonen2020}.} % just to generate a bibliography for now
+\note{We will use biproducts, and use some results from \cite{karvonen2020}.}
 
@@
 
-Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications (FIXME: refs).
-Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{fluid} that allows interactive exploration of programmatically generated visualisations.
+Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of
+program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
+Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid}
+that allows interactive exploration of programmatically generated visualisations.
 

### 2025-04-29 appendix/notes/auto-diff.tex
Tangent map.

\subsection{Automatic differentiation}

Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
${\pushf{f}}_x$ (tangent map or pushforward) of a function $f: \RR^m \to \RR^n$ at $x$ is a linear map
$\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$.

### 2025-04-29 appendix/notes/auto-diff.tex
Tangent map.

@@
 Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
-${\pushf{f}}_x$ (tangent map or pushforward) of a function $f: \RR^m \to \RR^n$ at $x$ is a linear map
-$\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
-$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$.
+${\pushf{f}}_x$ (tangent map or pushforward) of a differentiable function $f: \RR^m \to \RR^n$ at $x$ is a
+linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
+$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so ${\pushf{f}}_x: $\RR^m \linearto \RR^n$.

### 2025-04-29 appendix/notes/auto-diff.tex
Backward derivative.

--- 2025-04-29 appendix/notes/auto-diff.tex
@@
 Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
-${\pushf{f}}_x$ (tangent map or pushforward) of a function $f: \RR^m \to \RR^n$ at $x$ is a linear map
-$\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
-$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$.
+(tangent map or pushforward) ${\pushf{f}}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x$ is a
+linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
+$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually ${\pushf{f}}_x: \RR^m \linearto \RR^n$.
+The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n \linearto
+\RR^m$.

### 2025-04-29 appendix/notes/auto-diff.tex
Backward derivative.
Basic AD setup.

--- 2025-04-29 appendix/notes/auto-diff.tex
@@
 Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
-${\pushf{f}}_x$ (tangent map or pushforward) of a function $f: \RR^m \to \RR^n$ at $x$ is a linear map
-$\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean manifold,
-$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$.
+(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in
+\RR^m$ is a linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean
+manifold, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually $\pushf{f}_x: \RR^m \linearto
+\RR^n$. The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n
+\linearto \RR^m$.
+
+\subsubsection{Chain rule}
+
+Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$ we can define the following
+composite derivatives of $g \comp f: \RR^m \to \RR^k$:
+
+\begin{itemize}
+\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x$
+\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}$
+\end{itemize}
+
+\begin{definition}[Forward mode automatic differentiation functor]
+Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every smooth map $f:
+\RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m
+\linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at that
+point.
+\end{definition}
+
+For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, define $f_1 = \pi_1 \comp f$ and $f_2$ for
+$\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we
+can define composition on such maps as:
+\begin{align*}
+(g \comp f)_1(x) &= f_1(g_1(x)) \\
+(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
+\end{align*}
+\noindent and verify that $\tangents$ is indeed a functor. If $f_2(x)$ happens to be the derivative of $f_1$
+at $x$ for any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain
+rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
+
+\begin{definition}[Reverse mode automatic differentiation functor]
+Define the functor $\tangents^*$ which also sends every vector space $\RR^n$ to itself but which sends every
+smooth map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m \to
+\RR^n \times (\RR^n \to \RR^m)$.
+\end{definition}
+
+\noindent and again we can define a notion of composition for such maps that respects the (backward) chain
+rule.

### 2025-04-29 appendix/notes/language.tex
Move syntax fig to notes.
Make a start on syntax.

@@
-\section{Language}
+\section{Language Syntax and Interpretation}
 
@@
 
-We parameterise our language over a signature $\Sigma = (S, \phi, R)$ consisting of:
+We assume a language parameterised by a signature $\Sigma = (S, \Op, \Rel)$ consisting of:
 \begin{itemize}
-\item a set $S$ of sorts
-\item a set of operations $\phi$ with signature $\Sigma(\phi) \in S^* \times S$
-\item a set of relations $R$ with signature $\Sigma(R) \in S^*$
+\item a set $S$ of sorts;
+\item a set $\Op(s_1, \ldots, s_n; s)$ of operations with input sorts $s_1, \ldots, s_n$ and output sort $s$;
+\item a set $\Rel(s_1, \ldots, s_n)$ of $n$-ary relations over sorts $s_1, \ldots, s_n$
 \end{itemize}
 
-\subsection{Semantic domain}
+\noindent \todo{Are relations just for examples -- perhaps not needed in the formalism?}
 
-We interpret our language into any category with:
-\begin{itemize}
-\item finite products (binary products and a terminal object)
-\item Booleans
-\end{itemize}
-\note{Ignoring lists for now. Booleans can be derived from coproducts.}
+\input{fig/syntax}
+
+% \subsection{Semantic domain}
+%
+% We interpret our language into any category with:
+% \begin{itemize}
+% \item finite products (binary products and a terminal object)
+% \item Booleans
+% \end{itemize}
+% \note{Ignoring lists for now. Booleans can be derived from coproducts.}

### 2025-04-29 main/formal-structures.tex
lawvere63.

@@
 The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
-An object of the Grothendieck construction $\int T$ is a pair $(X \in \cat{C}, \partial X \in T(X)$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
+An object of the Grothendieck construction $\int T$ is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
 The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
@@
 In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$.
-As is well-known\todo{reference}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
+As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
 In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the cartesian products from $\Set$.

### 2025-05-04 appendix/notes/grothendieck.tex
Switch to contravariant formulation.

@@
 
+\subsubsection{Grothendieck construction}
+
+By an \emph{indexed category over $\cat{C}$} we shall mean a functor $\cat{C}^\op \to \Cat$.
+
 \begin{definition}[Grothendieck construction for $F$]
-Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
-$F$ is the category which has:
+Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction}
+for $T$ is the category $\Grothendieck{}T$ which has:
 \begin{itemize}
-\item as objects $X$ all pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
-\item as morphisms $f: X \to Y$ all pairs of a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ and a
-morphism $\faml{f}: F(\idx{f})(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
+\item as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $F(X)$;
+\item as morphisms $(X, \partial X) \to (Y, \partial Y)$ all pairs $(f, \partial f)$ of morphisms $f: X
+\to Y$ in $\cat{C}$ and morphisms $\partial f: F(X)(\partial X) \to \partial Y$ in $F(Y)$.
 \end{itemize}

### 2025-05-04 appendix/notes/fam.tex
Minor tweaks.
Switch to contravariant formulation.

@@
 \item sends any morphism $g: X \to Y$ between $J$-indexed families to the morphism $\reindex{g}{f}:
-\reindex{X}{f} \to \reindex{X}{g}$ where $\reindex{g}{f}_i = g_{f(i)}$.
+\reindex{X}{f} \to \reindex{Y}{f}$ where $\reindex{g}{f}_i = g_{f(i)}$.
 \end{itemize}
@@
 
+By \emph{indexed category over $\cat{C}$} we shall mean any functor $\cat{C}^\op \to \Cat$.
+
 \begin{definition}
-Then define $\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ to be the functor which sends any set $I$ to
+Then define $\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ to be the indexed category which sends any set $I$ to
 $\Fam(I,\cat{C})$ and any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to

### 2025-05-04 appendix/notes/grothendieck.tex
Switch to contravariant formulation.

@@
 \subsubsection{Grothendieck construction}
-
-By an \emph{indexed category over $\cat{C}$} we shall mean a functor $\cat{C}^\op \to \Cat$.
 
@@
 \item as morphisms $(X, \partial X) \to (Y, \partial Y)$ all pairs $(f, \partial f)$ of morphisms $f: X
-\to Y$ in $\cat{C}$ and morphisms $\partial f: F(X)(\partial X) \to \partial Y$ in $F(Y)$.
+\to Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to F(Y)(\partial Y)$ in $F(X)$.
 \end{itemize}
@@
 
-\noindent When $F$ is contravariant, a morphism $f: X \to Y$ in $\Grothendieck{\cat{C}}{F}$ is a morphism
-$\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to
-F(\idx{f})(\faml{Y})$ in $F(\idx{X})$. \todo{Might want to unpack this a bit.}
+\subsubsection{Category of families}
 
@@
 For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
-$\Grothendieck{\Set^{\op}}\Fam(-,\cat{C})$.
+$\Grothendieck{}\Fam(-,\cat{C})$.
 \end{definition}
 
-\noindent In $\Fam(\cat{C})$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$
-in $\Fam(\idx{X},\cat{C})$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
-morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},\cat{C})$.
+\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
+$\partial X$ in $\Fam(X,\cat{C})$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are
+functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in
+$\Fam(X,\cat{C})$.
 

### 2025-05-04 appendix/notes/grothendieck.tex
Homing in on chain rule in Grothendieck construction.

@@
 \begin{itemize}
-\item as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $F(X)$;
-\item as morphisms $(X, \partial X) \to (Y, \partial Y)$ all pairs $(f, \partial f)$ of morphisms $f: X
-\to Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to F(Y)(\partial Y)$ in $F(X)$.
+\item as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of
+$T(X)$;
+\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to
+Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
 \end{itemize}
 \end{definition}
+
+For any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y, \partial Y)
+\to (Z, \partial Z)$ in $\Grothendieck{}T$, we have:
+
+\begin{itemize}
+\item $\partial f: \partial X \to T(f)(\partial Y)$;
+\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$.
+\end{itemize}
+
+\noindent The composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is
+given by $(g \comp f, T(f)(\partial g) \comp \partial f)$.
 

### 2025-05-04 appendix/notes/auto-diff.tex
Move auto-diff to after definition of FinVect.

@@
 Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
-(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in
-\RR^m$ is a linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is a Euclidean
-manifold, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually $\pushf{f}_x: \RR^m \linearto
-\RR^n$. The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n
-\linearto \RR^m$.
+(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
+is a linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is Euclidean (flat),
+$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually $\pushf{f}_x: \RR^m \linearto \RR^n$. The
+\emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n \linearto \RR^m$.
 
@@
 rule.
+
+\subsubsection{Automatic differentiation via the Grothendieck construction}

### 2025-05-05 appendix/notes/auto-diff.tex
Spelling out auto-diff via Grothendieck.

@@
 \subsubsection{Automatic differentiation via the Grothendieck construction}
+
+Let $X, Y, Z$ range over sets and $U, V, W$ range over finite vector spaces. Recall that the category
+$\Fam(\FinVect) = \Grothendieck{}\Fam(-, \FinVect)$ (\defref{grothendieck:Grothendieck}) has:
+\begin{itemize}
+\item as objects, all pairs $(X, \{V_x\}_{x \in X})$ of a set $X$ and an object of $\Fam(X,\FinVect)$;
+\item as morphisms $(X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$, all pairs $(f, \partial f)$ of
+functions $f: X \to Y$ and families of morphisms $\partial f_x: \{V_x\}_{x \in X} \to \{W_{f(x)}\}_{x \in X}$
+in $\Fam(X,\FinVect)$.
+\end{itemize}
+
+For any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$ and $(g, \partial g): (Y,
+\{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have:
+\begin{itemize}
+\item $\partial f: \{V_x\}_{x \in X} \to \{W_{f(x)}\}_{x \in X}$;
+\item $\Fam(g, \FinVect)(\partial g): \Fam(g, \FinVect)(\{W_y\}_{y \in Y}) \to \Fam(g, \FinVect)(\{Z_{g(y)}\}_{y \in Y})$.
+\end{itemize}

### 2025-05-05 appendix/notes/fam.tex
Spelling out auto-diff via Grothendieck.

@@
 \begin{definition}[Category of $I$-indexed families of objects]
+\label{def:fam:Fam}
 For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category which has:

### 2025-05-05 appendix/notes/auto-diff.tex
Fix morphisms in Fam(FinVect).

@@
 \item as morphisms $(X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$, all pairs $(f, \partial f)$ of
-functions $f: X \to Y$ and families of morphisms $\partial f_x: \{V_x\}_{x \in X} \to \{W_{f(x)}\}_{x \in X}$
+functions $f: X \to Y$ and families of morphisms $\{\partial f_x: V_x \to W_{f(x)}\}_{x \in X}$
 in $\Fam(X,\FinVect)$.

### 2025-05-05 appendix/notes/auto-diff.tex
Fix morphisms in Fam(FinVect).
Move LatGal and FinVect to after section on biproducts/semi-additive categories.

@@
 
-For any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$ and $(g, \partial g): (Y,
-\{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have:
+\noindent Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor
+$\reindex{-}{f}: \Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ which sends any family of morphisms $\{g_y\}_{y \in
+Y}$ to $\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in
+Y})$ and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have:
 \begin{itemize}
-\item $\partial f: \{V_x\}_{x \in X} \to \{W_{f(x)}\}_{x \in X}$;
-\item $\Fam(g, \FinVect)(\partial g): \Fam(g, \FinVect)(\{W_y\}_{y \in Y}) \to \Fam(g, \FinVect)(\{Z_{g(y)}\}_{y \in Y})$.
+\item $\partial f_x: V_x \to W_{f(x)}$ for any $x \in X$;
+\item $\reindex{(\partial g_x)}{f}: \{W_{f(x)}\}_{x \in X} \to \{Z_{g(f(x))}\}_{x \in X}$.
 \end{itemize}

### 2025-05-05 appendix/notes/auto-diff.tex
Unpacking composition in Fam(FinVect).

--- 2025-05-05 appendix/notes/auto-diff.tex
@@
 
-For any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$ and $(g, \partial g): (Y,
-\{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have:
+\noindent Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor
+$\reindex{-}{f}: \Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{g_y\}_{y \in Y}$
+to $\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$
+and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
 \begin{itemize}
-\item $\partial f: \{V_x\}_{x \in X} \to \{W_{f(x)}\}_{x \in X}$;
-\item $\Fam(g, \FinVect)(\partial g): \Fam(g, \FinVect)(\{W_y\}_{y \in Y}) \to \Fam(g, \FinVect)(\{Z_{g(y)}\}_{y \in Y})$.
+\item $\partial f_x: V_x \to W_{f(x)}$;
+\item $\reindex{\partial g}{f}_x: W_{f(x)} \to Z_{g(f(x))}$.
 \end{itemize}

### 2025-05-05 appendix/notes/auto-diff.tex
Unpacking composition in Fam(FinVect).
Unpacked composition in Fam(FinVect).

@@
 
-\noindent Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor
-$\reindex{-}{f}: \Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{g_y\}_{y \in Y}$
-to $\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$
+Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor $\reindex{-}{f}:
+\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{g_y\}_{y \in Y}$ to
+$\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$
 and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
@@
 \item $\partial f_x: V_x \to W_{f(x)}$;
-\item $\reindex{\partial g}{f}_x: W_{f(x)} \to Z_{g(f(x))}$.
+\item $\reindex{\partial g}{f}_x: W_{f(x)} \to U_{g(f(x))}$.
 \end{itemize}
+
+\noindent The composition $(g, \partial g) \comp (f, \partial f): (X, \{V_x\}_{x \in X}) \to (Z, \{U_z\}_{z
+\in Z})$ is thus given by $(g \comp f, \reindex{\partial g}{f} \comp \partial f)$.

### 2025-05-05 main/formal-structures.tex
Case.

@@
 
-\note{Probably a better way: (1) introduce the category of lattices and galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
+\note{Probably a better way: (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
 
@@
 We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}.
-Not only does this mean that the constructions and proofs have been checked, the construction is executable and can be runn on small examples.
+Not only does this mean that the constructions and proofs have been checked, the construction is executable and can be run on small examples.
 We are also planning to use Agda's JavaScript backend to produce a demo.\todo{Do this?}
@@
 The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
-The maps in the indexed category are the maps of tangents or approximations, being derivatives and galois connections respectively.
+The maps in the indexed category are the maps of tangents or approximations, being derivatives and Galois connections respectively.
 

### 2025-05-05 main/introduction.tex
Switch to contravariant formulation.
Case.

@@
 
-Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of
-program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid}
-that allows interactive exploration of programmatically generated visualisations.
+Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
+Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
 
@@
 
-\todo{explain why we want a more abstract view on galois program slicing}
+\todo{explain why we want a more abstract view on Galois program slicing}
 There is a close analogy between Galois Program Slicing and Automatic Differentiation for differentiable programs.

### 2025-05-05 main/formal-structures.tex
Tweaks.

@@
 Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of them having products and being enriched in commutative monoids\todo{fwd ref}.
-However, neither of them have all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
+However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
 Neither of these are ideal.
@@
 
-Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using Cmon-enriched presheaves.\todo{elaborate.}
+Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using $\CMon$-enriched presheaves.\todo{elaborate.}

### 2025-05-05 appendix/notes/auto-diff.tex
Clarify that forward/backward derivatives are determined by f and given by the Jacobean matrix of f at x and its transpose.

@@
 (tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
-is a linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is Euclidean (flat),
-$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually $\pushf{f}_x: \RR^m \linearto \RR^n$. The
-\emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n \linearto \RR^m$.
+is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
+of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually
+$\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
+pullback) $\pullf{f}_x$ is the unique linear map $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of
+the Jacobean matrix of $f$ at $x$.
 

### 2025-05-05 appendix/notes/grothendieck.tex
Spelling out auto-diff via Grothendieck.
Clarify that forward/backward derivatives are determined by f and given by the Jacobean matrix of f at x and its transpose.

@@
 \begin{definition}[Grothendieck construction for $F$]
+\label{def:grothendieck:Grothendieck}
 Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction}
@@
 \end{definition}
+
+\noindent The $\partial X$ notation is justified by the fact that the Grothendieck construction formalises the
+idea of tangent spaces and derivatives.
 

### 2025-05-05 appendix/notes/auto-diff.tex
Correction.

@@
 is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
-of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually
+of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so in fact
 $\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
@@
 
-For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, define $f_1 = \pi_1 \comp f$ and $f_2$ for
+For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
 $\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we

### 2025-05-05 appendix/notes/auto-diff.tex
Clarify chain rule.

@@
 
-Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$ we can define the following
-composite derivatives of $g \comp f: \RR^m \to \RR^k$:
+Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
+we have:
 
 \begin{itemize}
-\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x$
-\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}$
+\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
+\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
 \end{itemize}
@@
 For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
-$\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we
-can define composition on such maps as:
+$\pi_2 \comp f$. We can define composition on such maps as:
 \begin{align*}
@@
 \end{align*}
-\noindent and verify that $\tangents$ is indeed a functor. If $f_2(x)$ happens to be the derivative of $f_1$
-at $x$ for any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain
-rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
+\noindent and verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
+\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
+any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
+$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
 
@@
 Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor $\reindex{-}{f}:
-\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{g_y\}_{y \in Y}$ to
-$\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$
-and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
+\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{\partial g_y\}_{y \in Y}$ to
+$\{\partial g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y
+\in Y})$ and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
 \begin{itemize}
 \item $\partial f_x: V_x \to W_{f(x)}$;
-\item $\reindex{\partial g}{f}_x: W_{f(x)} \to U_{g(f(x))}$.
+\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: W_{f(x)} \to U_{g(f(x))}$.
 \end{itemize}

### 2025-05-05 appendix/notes/auto-diff.tex
Category Diff.

@@
 
-For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
-$\pi_2 \comp f$. We can define composition on such maps as:
+Write $\Diff$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
+all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
+$\pi_2 \comp f$, composition in $\Diff$ is given by:
 \begin{align*}
@@
 \end{align*}
-\noindent and verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
+\noindent We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
 \linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for

### 2025-05-05 appendix/notes/auto-diff.tex
\int with a param notation for Grothendieck.

@@
 
-\subsubsection{Automatic differentiation via the Grothendieck construction}
+\subsubsection{Automatic differentiation via category of families}
 
 Let $X, Y, Z$ range over sets and $U, V, W$ range over finite vector spaces. Recall that the category
-$\Fam(\FinVect) = \Grothendieck{}\Fam(-, \FinVect)$ (\defref{grothendieck:Grothendieck}) has:
+$\Fam(\FinVect) = \Grothendieck{X}\Fam(X, \FinVect)$ (\defref{grothendieck:Grothendieck}) has:
 \begin{itemize}

### 2025-05-05 appendix/notes/auto-diff.tex
Diff^* category of cotangent maps.

@@
 
-Write $\Diff$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
+Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
 all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
-$\pi_2 \comp f$, composition in $\Diff$ is given by:
+$\pi_2 \comp f$, composition in $\Diff_*$ is given by:
 \begin{align*}
@@
 \end{align*}
-\noindent We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
+\noindent We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m
 \linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
@@
 
-\noindent and again we can define a notion of composition for such maps that respects the (backward) chain
-rule.
+\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition respects
+the (backward) chain rule.
 

### 2025-05-05 appendix/notes/auto-diff.tex
Ok think I get how to see Diff as subcategory of Fam(FinVect).

@@
 
-\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition respects
-the (backward) chain rule.
+\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition of the
+cotangent maps respects the (backward) chain rule.
 
@@
 \in Z})$ is thus given by $(g \comp f, \reindex{\partial g}{f} \comp \partial f)$.
+
+We can think of $\Diff$ as the subcategory of $\Fam(\FinVect)$ where the objects are just finite vector spaces
+$\RR^n$. There is no need to pair with an $\RR^n$-indexed family of finite vector spaces, since the tangent
+space for any $x \in \RR^n$ is simply $\RR^n$. Similarly the morphisms are fully specified by the pair of a
+smooth function $f: \RR^m \to \RR^n$ and constant family of linear maps $\{\delta f_x: \RR^m \linearto
+\RR^n\}_{x \in \RR^m}$.

### 2025-05-05 appendix/notes/auto-diff.tex
Diff is a full subcategory of Fam(FinVect).

--- 2025-05-05 appendix/notes/auto-diff.tex
@@
 
+\begin{definition}[$\Diff$]
+\label{def:auto-diff:Diff}
 Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
@@
 \end{align*}
-\noindent We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m
-\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
-any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
-$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
+\end{definition}
+
+We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
+\RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for any $x \in
+\RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp
+f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
 
@@
 \in Z})$ is thus given by $(g \comp f, \reindex{\partial g}{f} \comp \partial f)$.
+
+If we view any $\Diff$ object $\RR^n$ (\defref{auto-diff:Diff}) as implicitly paired with a constant family
+$\{\RR^n\}_{x \in \RR^n}$ and any $\Diff$ morphism $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the
+pair of a smooth function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta f_x: \RR^m
+\linearto \RR^n\}_{x \in \RR^m}$, then it is clear that $\Diff$ is equivalent to a full subcategory of
+$\Fam(\FinVect)$.

### 2025-05-05 appendix/notes/auto-diff.tex
smooth -> differentiable (don't need C^\infinity functions).

@@
 \begin{definition}[Forward mode automatic differentiation functor]
-Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every smooth map $f:
-\RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m
-\linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at that
-point.
+Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every differentiable map
+$f: \RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
+(\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
+that point.
 \end{definition}
+
+For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
+$\pi_2 \comp f$.
 
@@
 \label{def:auto-diff:Diff}
-Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
-all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
-$\pi_2 \comp f$, composition in $\Diff_*$ is given by:
+Define $\Diff_*$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
+are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, with $f_1$ differentiable. Composition in
+$\Diff_*$ is given by:
 \begin{align*}
@@
 Define the functor $\tangents^*$ which also sends every vector space $\RR^n$ to itself but which sends every
-smooth map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m \to
-\RR^n \times (\RR^n \to \RR^m)$.
+differentiable map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m
+\to \RR^n \times (\RR^n \to \RR^m)$.
 \end{definition}
@@
 
-If we view any $\Diff$ object $\RR^n$ (\defref{auto-diff:Diff}) as implicitly paired with a constant family
-$\{\RR^n\}_{x \in \RR^n}$ and any $\Diff$ morphism $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the
-pair of a smooth function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta f_x: \RR^m
-\linearto \RR^n\}_{x \in \RR^m}$, then it is clear that $\Diff$ is equivalent to a full subcategory of
-$\Fam(\FinVect)$.
+Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which pairs objects $\RR^n$ with the constant
+family $\{\RR^n\}_{x \in \RR^n}$ and sends morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
+the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta
+f_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. Then it is clear that $\Diff$ is equivalent to a full subcategory
+of $\Fam(\FinVect)$. For any morphisms $(f, \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^n,
+\{\RR^n\}_{y \in \RR^n})$ and $(g, \partial g): (\RR^n, \{\RR^n\}_{y \in \RR^n}) \to (\RR^k, \{\RR^k\}_{z \in
+\RR^k})$ in the image of that inclusion functor, we have for any $x \in \RR^m$:
+\begin{itemize}
+\item $\partial f_x: \RR^m \to \RR^n$;
+\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$
+\end{itemize}

### 2025-05-05 appendix/notes/auto-diff.tex
smooth -> differentiable (don't need C^\infinity functions).

@@
 Define $\Diff_*$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
-are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, with $f_1$ differentiable. Composition in
+are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ with $f_1$ differentiable. Composition in
 $\Diff_*$ is given by:

### 2025-05-05 appendix/notes/auto-diff.tex
Ok, recovered chain rule from composition in Fam(FinVect), somewhat pedantically.

@@
 We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
-\RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for any $x \in
-\RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp
-f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
+\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
+similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
+the derivative of $(g \comp f)_1$ at $x$.
 
@@
 
-Let $X, Y, Z$ range over sets and $U, V, W$ range over finite vector spaces. Recall that the category
-$\Fam(\FinVect) = \Grothendieck{X}\Fam(X, \FinVect)$ (\defref{grothendieck:Grothendieck}) has:
-\begin{itemize}
-\item as objects, all pairs $(X, \{V_x\}_{x \in X})$ of a set $X$ and an object of $\Fam(X,\FinVect)$;
-\item as morphisms $(X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$, all pairs $(f, \partial f)$ of
-functions $f: X \to Y$ and families of morphisms $\{\partial f_x: V_x \to W_{f(x)}\}_{x \in X}$
-in $\Fam(X,\FinVect)$.
-\end{itemize}
+Recall that the category $\Fam(\FinVect)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
+\FinVect)$. The functor $\Diff_* \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant
+families of tangent spaces $(\RR^n, \{\RR^n\}_{x \in \RR^n})$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m
+\linearto \RR^n)$ to pairs of differentiable functions $f_1: \RR^m \to \RR^n$ and constant families of linear
+maps $f_2: \{\delta f_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full
+subcategory of $\Fam(\FinVect)$.
 
-Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor $\reindex{-}{f}:
-\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{\partial g_y\}_{y \in Y}$ to
-$\{\partial g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y
-\in Y})$ and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
-\begin{itemize}
-\item $\partial f_x: V_x \to W_{f(x)}$;
-\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: W_{f(x)} \to U_{g(f(x))}$.
-\end{itemize}
-
-\noindent The composition $(g, \partial g) \comp (f, \partial f): (X, \{V_x\}_{x \in X}) \to (Z, \{U_z\}_{z
-\in Z})$ is thus given by $(g \comp f, \reindex{\partial g}{f} \comp \partial f)$.
-
-Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which pairs objects $\RR^n$ with the constant
-family $\{\RR^n\}_{x \in \RR^n}$ and sends morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
-the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta
-f_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. Then it is clear that $\Diff$ is equivalent to a full subcategory
-of $\Fam(\FinVect)$. For any morphisms $(f, \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^n,
-\{\RR^n\}_{y \in \RR^n})$ and $(g, \partial g): (\RR^n, \{\RR^n\}_{y \in \RR^n}) \to (\RR^k, \{\RR^k\}_{z \in
-\RR^k})$ in the image of that inclusion functor, we have for any $x \in \RR^m$:
+For any $(f, \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^n, \{\RR^n\}_{x \in \RR^n})$ and $(g,
+\partial g): (\RR^n, \{\RR^n\}_{x \in \RR^n}) \to (\RR^k, \{\RR^k\}_{x \in \RR^k})$ in the image of that
+functor, we have that for any $x \in \RR^m$:
 \begin{itemize}
@@
 \end{itemize}
+
+\noindent with the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f}
+\comp \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^k, \{\RR^k\}_{x \in \RR^k})$ corresponding
+exactly to the chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp
+{\partial f}_x$.

### 2025-05-05 main/formal-structures.tex
cartesian -> Cartesian.

@@
 As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
-In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the cartesian products from $\Set$.
+In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$.
 When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
-If the coproducts are actually {\em biproducts}, then $\Fam(\cat{X})$ is cartesian closed, and we can interpret higher-order programs.
+If the coproducts are actually {\em biproducts}, then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
 

### 2025-05-05 appendix/notes/auto-diff.tex
cartesian -> Cartesian.
Start on Galois slicing.

@@
 {\partial f}_x$.
+
+\subsubsection{Galois slicing via category of families}
+
+Now recall that the category $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
+\LatGal)$, with:
+\begin{itemize}
+\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for each $x \in X$, a bounded lattice
+$\partial X_x$;
+\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
+Y$ and Galois connections $\partial f: \partial X_x \to \partial Y_{f(x)}$.
+\end{itemize}

### 2025-05-06 appendix/notes/auto-diff.tex
naturally isomorphic -> canonically isomorphic.
Clean up backwards derivative.

@@
 
-Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
-(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
-is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
-of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so in fact
-$\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
-pullback) $\pullf{f}_x$ is the unique linear map $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of
-the Jacobean matrix of $f$ at $x$.
+For a differentiable manifold $M$, write $\tangents_x(M)$ for the tangent space at a point $x \in M$ and
+$\cotangents_x(M)$ for the cotangent space at $x$. Now consider flat manifolds of the form $\RR^n$.
+
+\begin{itemize}
+\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of a differentiable function
+$f: \RR^m \to \RR^n$ at $x \in \RR^m$ is the unique linear map $\tangents_x(\RR^m) \linearto
+\tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix of $f$ at $x$. Since $\RR^n$ is flat,
+$\tangents_x(\RR^n)$ is canonically isomorphic to $\RR^n$, so in fact $\pushf{f}_x$ is a linear map $\RR^m
+\linearto \RR^n$.
+\item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
+$\cotangents_{f(x)}(\RR^n) \linearto \cotangents_x(\RR^m)$ given by the transpose (adjoint) of the Jacobean
+matrix of $f$ at $x$. Since $\RR^n$ is flat, $\cotangents_x(\RR^n)$ is also canonically isomorphic to $\RR^n$,
+so in fact $\pullf{f}_x$ is a linear map $\RR^n \linearto \RR^m$.
+\end{itemize}
 
@@
 \begin{definition}[Forward mode automatic differentiation functor]
-Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every differentiable map
-$f: \RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
+Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
+$f: \RR^m \to \RR^n$ to the function $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
 (\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
@@
 \label{def:auto-diff:Diff}
-Define $\Diff_*$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
+Define $\Diff$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
 are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ with $f_1$ differentiable. Composition in
-$\Diff_*$ is given by:
+$\Diff$ is given by:
 \begin{align*}
@@
 
-We can verify that $\tangents_*$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
+We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
 \RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
@@
 \begin{definition}[Reverse mode automatic differentiation functor]
-Define the functor $\tangents^*$ which also sends every vector space $\RR^n$ to itself but which sends every
-differentiable map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m
-\to \RR^n \times (\RR^n \to \RR^m)$.
+Define the functor $\cotangents$ which also sends every vector space $\RR^n$ to itself but which sends every
+differentiable map $f: \RR^m \to \RR^n$ to the function $\cotangents(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m
+\to \RR^n \times (\RR^n \linearto \RR^m)$.
 \end{definition}
@@
 Recall that the category $\Fam(\FinVect)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
-\FinVect)$. The functor $\Diff_* \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant
+\FinVect)$. The functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant
 families of tangent spaces $(\RR^n, \{\RR^n\}_{x \in \RR^n})$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m

### 2025-05-06 appendix/notes/auto-diff.tex
Use const family notation.

@@
 
-For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
-$\pi_2 \comp f$.
+For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ define $f_1 = \pi_1 \comp f: \RR^m \to \RR^n$
+and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$.
 
@@
 Recall that the category $\Fam(\FinVect)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
-\FinVect)$. The functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant
-families of tangent spaces $(\RR^n, \{\RR^n\}_{x \in \RR^n})$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m
-\linearto \RR^n)$ to pairs of differentiable functions $f_1: \RR^m \to \RR^n$ and constant families of linear
-maps $f_2: \{\delta f_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full
-subcategory of $\Fam(\FinVect)$.
+\FinVect)$. The functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families
+of tangent spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto
+\RR^n)$ to pairs $(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps
+$f_2: \{{f_2}_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full subcategory
+of $\Fam(\FinVect)$.
 
-For any $(f, \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^n, \{\RR^n\}_{x \in \RR^n})$ and $(g,
-\partial g): (\RR^n, \{\RR^n\}_{x \in \RR^n}) \to (\RR^k, \{\RR^k\}_{x \in \RR^k})$ in the image of that
-functor, we have that for any $x \in \RR^m$:
+For any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g, \partial g): (\RR^n,
+\const(\RR^n)) \to (\RR^k, \const(\RR^k))$ in that subcategory, we have that for any $x \in \RR^m$:
 \begin{itemize}
@@
 
-\noindent with the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f}
-\comp \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^k, \{\RR^k\}_{x \in \RR^k})$ corresponding
-exactly to the chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp
-{\partial f}_x$.
+\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
+\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the chain rule, since
+$(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
 

### 2025-05-06 appendix/notes/auto-diff.tex
Use const family notation.
Experiment with diagram.
Diff* is (I think -- haven't checked) a full subcategory of Fam(FinVect^op).
Cleanup.

@@
 
-\begin{definition}[Forward mode automatic differentiation functor]
+\subsubsection{Forward mode automatic differentiation}
+
+\begin{definition}
 Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
@@
 
-\begin{definition}[Reverse mode automatic differentiation functor]
-Define the functor $\cotangents$ which also sends every vector space $\RR^n$ to itself but which sends every
-differentiable map $f: \RR^m \to \RR^n$ to the function $\cotangents(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m
-\to \RR^n \times (\RR^n \linearto \RR^m)$.
-\end{definition}
+\subsubsection{$\Diff$ as a subcategory of $\Fam(\FinVect)$}
 
-\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition of the
-cotangent maps respects the (backward) chain rule.
-
-\subsubsection{Automatic differentiation via category of families}
-
-Recall that the category $\Fam(\FinVect)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
-\FinVect)$. The functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families
-of tangent spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto
-\RR^n)$ to pairs $(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps
-$f_2: \{{f_2}_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full subcategory
-of $\Fam(\FinVect)$.
+Recall that $\Fam(\FinVect)$ denotes the Grothendieck construction $\Grothendieck{X}\Fam(X, \FinVect)$. The
+functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of tangent
+spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs
+$(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2: \{{f_2}_x:
+\RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full subcategory of
+$\Fam(\FinVect)$.
 
@@
 \item $\partial f_x: \RR^m \to \RR^n$;
-\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$
+\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
 \end{itemize}
@@
 \noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the chain rule, since
-$(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the forward chain rule,
+since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+
+\subsubsection{Reverse mode automatic differentiation}
+
+\begin{definition}
+Define the functor $\cotangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
+$f: \RR^m \to \RR^n$ to the function $\cotangents(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m \to \RR^n \times
+(\RR^n \linearto \RR^m)$.
+\end{definition}
+
+\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition of the
+cotangent maps respects the backward chain rule, with in this case $\Diff^*$ equivalent to a full subcategory
+of $\Fam(\FinVect^\op)$.
 
@@
 
-Now recall that the category $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X,
-\LatGal)$, with:
+Now recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$, with:
 \begin{itemize}

### 2025-05-06 appendix/notes/auto-diff.tex
Tweak.

@@
 
-\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition of the
-cotangent maps respects the backward chain rule, with in this case $\Diff^*$ equivalent to a full subcategory
-of $\Fam(\FinVect^\op)$.
+\noindent and again define a corresponding category $\Diff^*$ (equivalent to a full subcategory of
+$\Fam(\FinVect^\op)$) of such maps where composition of the cotangent maps respects the backward chain rule.
+
+\todo{The reason we can't combine $\Diff$ and $\Diff^*$ into a single category, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$ is..?}
 

### 2025-05-06 appendix/notes/fam.tex
Tweak.

@@
 \begin{definition}
-Then define $\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ to be the indexed category which sends any set $I$ to
+Define $\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ to be the indexed category which sends any set $I$ to
 $\Fam(I,\cat{C})$ and any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to

### 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-grothendieck.tex
Auto-diff separate.

--- 2025-05-06 appendix/notes/auto-diff.tex
@@
-\subsection{Automatic differentiation}
+\subsection{Galois slicing and automatic differentiation via Grothendieck construction}
 
-For a differentiable manifold $M$, write $\tangents_x(M)$ for the tangent space at a point $x \in M$ and
-$\cotangents_x(M)$ for the cotangent space at $x$. Now consider flat manifolds of the form $\RR^n$.
+\subsubsection{Galois slicing via category of families}
 
+Now recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$, with:
 \begin{itemize}
-\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of a differentiable function
-$f: \RR^m \to \RR^n$ at $x \in \RR^m$ is the unique linear map $\tangents_x(\RR^m) \linearto
-\tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix of $f$ at $x$. Since $\RR^n$ is flat,
-$\tangents_x(\RR^n)$ is canonically isomorphic to $\RR^n$, so in fact $\pushf{f}_x$ is a linear map $\RR^m
-\linearto \RR^n$.
-\item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
-$\cotangents_{f(x)}(\RR^n) \linearto \cotangents_x(\RR^m)$ given by the transpose (adjoint) of the Jacobean
-matrix of $f$ at $x$. Since $\RR^n$ is flat, $\cotangents_x(\RR^n)$ is also canonically isomorphic to $\RR^n$,
-so in fact $\pullf{f}_x$ is a linear map $\RR^n \linearto \RR^m$.
+\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for each $x \in X$, a bounded lattice
+$\partial X_x$;
+\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
+Y$ and Galois connections $\partial f: \partial X_x \to \partial Y_{f(x)}$.
 \end{itemize}
 
-\subsubsection{Chain rule}
-
-Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
-we have:
-
-\begin{itemize}
-\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
-\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
-\end{itemize}
-
-\subsubsection{Forward mode automatic differentiation}
-
-\begin{definition}
-Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
-$f: \RR^m \to \RR^n$ to the function $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
-(\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
-that point.
-\end{definition}
-
-For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ define $f_1 = \pi_1 \comp f: \RR^m \to \RR^n$
-and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$.
-
-\begin{definition}[$\Diff$]
-\label{def:auto-diff:Diff}
-Define $\Diff$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
-are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ with $f_1$ differentiable. Composition in
-$\Diff$ is given by:
-\begin{align*}
-(g \comp f)_1(x) &= f_1(g_1(x)) \\
-(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
-\end{align*}
-\end{definition}
-
-We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
-\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
-similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$.
-
-\subsubsection{$\Diff$ as a subcategory of $\Fam(\FinVect)$}
+\subsubsection{Automatic differentiation via categories of families}
 
@@
 since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
-
-\subsubsection{Reverse mode automatic differentiation}
-
-\begin{definition}
-Define the functor $\cotangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
-$f: \RR^m \to \RR^n$ to the function $\cotangents(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m \to \RR^n \times
-(\RR^n \linearto \RR^m)$.
-\end{definition}
-
-\noindent and again we can define a corresponding category $\Diff^*$ of such maps where composition of the
-cotangent maps respects the backward chain rule, with in this case $\Diff^*$ equivalent to a full subcategory
-of $\Fam(\FinVect^\op)$.
-
-\subsubsection{Galois slicing via category of families}
-
-Now recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$, with:
-\begin{itemize}
-\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for each $x \in X$, a bounded lattice
-$\partial X_x$;
-\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
-Y$ and Galois connections $\partial f: \partial X_x \to \partial Y_{f(x)}$.
-\end{itemize}

### 2025-05-06 appendix/notes/auto-diff.tex
Auto-diff separate.

@@
 
-\subsubsection{$\Diff$ as a subcategory of $\Fam(\FinVect)$}
-
-Recall that $\Fam(\FinVect)$ denotes the Grothendieck construction $\Grothendieck{X}\Fam(X, \FinVect)$. The
-functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of tangent
-spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs
-$(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2: \{{f_2}_x:
-\RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full subcategory of
-$\Fam(\FinVect)$.
-
-For any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g, \partial g): (\RR^n,
-\const(\RR^n)) \to (\RR^k, \const(\RR^k))$ in that subcategory, we have that for any $x \in \RR^m$:
-\begin{itemize}
-\item $\partial f_x: \RR^m \to \RR^n$;
-\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
-\end{itemize}
-
-\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the forward chain rule,
-since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
-
 \subsubsection{Reverse mode automatic differentiation}
@@
 \todo{The reason we can't combine $\Diff$ and $\Diff^*$ into a single category, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$ is..?}
-
-\subsubsection{Galois slicing via category of families}
-
-Now recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$, with:
-\begin{itemize}
-\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for each $x \in X$, a bounded lattice
-$\partial X_x$;
-\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
-Y$ and Galois connections $\partial f: \partial X_x \to \partial Y_{f(x)}$.
-\end{itemize}

### 2025-05-06 appendix/notes/auto-diff.tex
(strictly) indexed categories.

@@
 
-\todo{The reason we can't combine $\Diff$ and $\Diff^*$ into a single category, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$ is..?}
+\todo{Can't combine $\Diff$ and $\Diff^*$ into a single subcategory of $\Fam(\FinVect)$, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$, because $\Fam(\FinVect)$ is not Cartesian closed?}

### 2025-05-06 appendix/notes.tex
Preliminaries.
Some cleanup/better org.
Tangent map.
Move syntax fig to notes.
Move auto-diff to after definition of FinVect.
Move LatGal and FinVect to after section on biproducts/semi-additive categories.
Auto-diff separate.
Sync summary of notes.

@@
 \begin{itemize}
+\item automatic differentiation (\secref{auto-diff})
 \item $\CMon$-enriched categories, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
 \item biproducts and semi-additive categories (\secref{biproduct})
-\item category $\Fam(I,\cat{C})$ of $I$-indexed families of objects of $\cat{C}$ (\secref{fam})
-\item set-indexed products (\secref{set-indexed-product})
-\item Grothendieck construction $\Grothendieck{\cat{C}}F$ and category of families $\Fam(\cat{C})$
-(\secref{grothendieck})
-\item Examples of useful semi-additive categories (\secref{useful-semi-additive-categories}):
+\item useful semi-additive categories (\secref{useful-semi-additive-categories}):
    \begin{itemize}
@@
    \end{itemize}
+\item category $\Fam(I,\cat{C})$ of $I$-indexed families of objects of $\cat{C}$ (\secref{fam})
+\item set-indexed products (\secref{set-indexed-product})
+\item Grothendieck construction $\Grothendieck{\cat{C}}F$ and category of families $\Fam(\cat{C})$
+(\secref{grothendieck})
+\item useful categories of families for automatic approximation
+(\secref{galois-slicing-auto-diff-via-grothendieck}):
+   \begin{itemize}
+      \item Galois slicing via $\Fam(\LatGal)$ (\secref{galois-slicing-auto-diff-via-grothendieck:galois-slicing})
+      \item automatic differentiation via $\Fam(\FinVect)$ (\secref{galois-slicing-auto-diff-via-grothendieck:auto-diff})
+   \end{itemize}
 \end{itemize}
@@
 
-\input{appendix/notes/notation}
+\input{appendix/notes/preliminaries}
+\input{appendix/notes/auto-diff}
+\input{appendix/notes/cmon}
 \input{appendix/notes/cmon-enriched}
 \input{appendix/notes/biproduct}
+\input{appendix/notes/useful-semi-additive-categories}
 \input{appendix/notes/fam}
@@
 \input{appendix/notes/grothendieck}
-\input{appendix/notes/useful-semi-additive-categories}
-
-% \input{appendix/notes/language}
+\input{appendix/notes/auto-diff-galois-slicing-via-grothendieck}
+\input{appendix/notes/language}

### 2025-05-06 appendix/notes/auto-diff.tex
Sync summary of notes.

--- 2025-05-06 appendix/notes/auto-diff.tex
@@
 \subsection{Automatic differentiation}
+\label{sec:auto-diff}
 
@@
 
-\todo{The reason we can't combine $\Diff$ and $\Diff^*$ into a single category, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$ is..?}
+\todo{Can't combine $\Diff$ and $\Diff^*$ into a single subcategory of $\Fam(\FinVect)$, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$, because..?}

### 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-grothendieck.tex
Sync summary of notes.
Diff* is (I think -- haven't checked) a full subcategory of Fam(FinVect^op).

@@
 \subsection{Galois slicing and automatic differentiation via Grothendieck construction}
+\label{sec:galois-slicing-auto-diff-via-grothendieck}
 
-\subsubsection{Galois slicing via category of families}
+\subsubsection{Galois slicing via a category of families}
+\label{sec:galois-slicing-auto-diff-via-grothendieck:galois-slicing}
 
-Now recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$, with:
+Recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$ which has:
 \begin{itemize}
@@
 
+The action of $\Fam(-,\LatGal)$ on a morphism $f: X \to Y$. Composition of any $(f, \partial f):$
+
 \subsubsection{Automatic differentiation via categories of families}
+\label{sec:galois-slicing-auto-diff-via-grothendieck:auto-diff}
 
@@
 since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+
+Similarly, $\Diff^*$ is equivalent to a full subcategory of $\Fam(\FinVect^\op)$ with composition
+corresponding to the backward chain rule.

### 2025-05-06 appendix/notes.tex
Rename.

@@
 \item useful categories of families for automatic approximation
-(\secref{galois-slicing-auto-diff-via-grothendieck}):
+(\secref{galois-slicing-auto-diff-via-fam}):
    \begin{itemize}
-      \item Galois slicing via $\Fam(\LatGal)$ (\secref{galois-slicing-auto-diff-via-grothendieck:galois-slicing})
-      \item automatic differentiation via $\Fam(\FinVect)$ (\secref{galois-slicing-auto-diff-via-grothendieck:auto-diff})
+      \item Galois slicing via $\Fam(\LatGal)$ (\secref{galois-slicing-auto-diff-via-fam:galois-slicing})
+      \item automatic differentiation via $\Fam(\FinVect)$ (\secref{galois-slicing-auto-diff-via-fam:auto-diff})
    \end{itemize}
@@
 \input{appendix/notes/grothendieck}
-\input{appendix/notes/auto-diff-galois-slicing-via-grothendieck}
+\input{appendix/notes/auto-diff-galois-slicing-via-fam}
 \input{appendix/notes/language}

### 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-fam.tex
Rename.

--- 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-grothendieck.tex
@@
-\subsection{Galois slicing and automatic differentiation via Grothendieck construction}
-\label{sec:galois-slicing-auto-diff-via-grothendieck}
+\subsection{Galois slicing and automatic differentiation via categories of families}
+\label{sec:galois-slicing-auto-diff-via-fam}
 
-\subsubsection{Galois slicing via a category of families}
-\label{sec:galois-slicing-auto-diff-via-grothendieck:galois-slicing}
+\subsubsection{Galois slicing}
+\label{sec:galois-slicing-auto-diff-via-fam:galois-slicing}
 
@@
 
-\subsubsection{Automatic differentiation via categories of families}
-\label{sec:galois-slicing-auto-diff-via-grothendieck:auto-diff}
+\subsubsection{Automatic differentiation}
+\label{sec:galois-slicing-auto-diff-via-fam:auto-diff}
 
-Recall that $\Fam(\FinVect)$ denotes the Grothendieck construction $\Grothendieck{X}\Fam(X, \FinVect)$. The
-functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of tangent
-spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs
-$(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2: \{{f_2}_x:
-\RR^m \linearto \RR^n\}_{x \in \RR^m}$ exhibits $\Diff$ as equivalent to a full subcategory of
-$\Fam(\FinVect)$.
+Recall that $\Fam(\FinVect)$ denotes the Grothendieck construction $\Grothendieck{X}\Fam(X, \FinVect)$.
+Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of
+tangent spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
+pairs $(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2:
+\{{f_2}_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. This exhibits $\Diff$ as equivalent to a full subcategory
+of $\Fam(\FinVect)$.
 
-For any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g, \partial g): (\RR^n,
-\const(\RR^n)) \to (\RR^k, \const(\RR^k))$ in that subcategory, we have that for any $x \in \RR^m$:
+In that subcategory, for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
+\partial g): (\RR^n, \const(\RR^n)) \to (\RR^k, \const(\RR^k))$, we have that for any $x \in \RR^m$:
 \begin{itemize}

### 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-fam.tex
That's it for now, might switch to the view where we think of the objects in Fam(X,C) (X-indexed families of objects) as functors from X-as-discrete-category to C.

@@
 \label{sec:galois-slicing-auto-diff-via-fam}
+
+Recall that $\Fam(\cat{C})$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \cat{C})$.
 
@@
 
-Recall that $\Fam(\LatGal)$ is the Grothendieck construction $\Grothendieck{X}\Fam(X, \LatGal)$ which has:
+$\Fam(\LatGal)$ has:
 \begin{itemize}
@@
 
-The action of $\Fam(-,\LatGal)$ on a morphism $f: X \to Y$. Composition of any $(f, \partial f):$
+\todo{This is the basic setting for interpreting first-order programs.}
 
@@
 
-Recall that $\Fam(\FinVect)$ denotes the Grothendieck construction $\Grothendieck{X}\Fam(X, \FinVect)$.
 Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of

### 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-fam.tex
Clarify \partial f is family of Galois connections.

@@
 \begin{itemize}
-\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for each $x \in X$, a bounded lattice
+\item as objects $(X, \partial X)$ all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
 $\partial X_x$;
 \item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
-Y$ and Galois connections $\partial f: \partial X_x \to \partial Y_{f(x)}$.
+Y$ and families of Galois connections $\partial f_x: \partial X_x \to \partial Y_{f(x)}$ for every $x \in X$.
 \end{itemize}
+
+\noindent with the role of the pullback along $f$ in the composition $(g, \partial g) \comp (f, \partial f) =
+(g \comp f, \reindex{\partial g}{f} \comp \partial f)$ being to select the appropriate lattice of
+approximations and Galois connection at each point.
 

### 2025-05-06 main/formal-structures.tex

@@
 As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
-In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$.
+In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$).
 When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.

### 2025-05-07 main/formal-structures.tex
Addition of morphisms derived from biproduct structure.

@@
 \section{Formal Structures for Generalised Automatic Differentiation}
+
+\begin{definition}[Biproducts and Semi-additive categories]
+A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
+and inclusions) give an isomorphism
+\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
+\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
+initial and is called a \emph{zero} object.
+\end{definition}
+
+A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two
+morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
+forming a morphism $f + g: X \to Y$. Diagrammatically:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
+\end{tikzcd}
+\end{center}
+
+\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
+product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
+coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f}{g} = \coprodM{f}{g}$.
 

### 2025-05-07 main/formal-structures.tex
Addition of morphisms derived from biproduct structure.
Zero objects.
Biproducts and semi-additive categories.

@@
 A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and inclusions) give an isomorphism
+and injections) give an isomorphism
 \[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
@@
 product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
-coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f}{g} = \coprodM{f}{g}$.
+coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} =
+\coprodM{\inj_1 \comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
+composing the unique maps in and out of the zero object:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
+\end{tikzcd}
+\end{center}
+
+It is easy to verify that $+$ is associative and commutative and that $f + 0 = f = 0 + f$, and so every
+hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
+family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
+preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+
+\begin{salign*}
+f \comp \zero = f = \zero \comp f
+\end{salign*}
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}
 

### 2025-05-07 main/formal-structures.tex
Biproducts and semi-additive categories.

@@
 
-\begin{definition}[Biproducts and Semi-additive categories]
+\subsection{Biproducts and Semi-Additive Categories}
+
+\begin{definition}[Semi-additive category]
 A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
 and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
-\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object.
+\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
 \end{definition}
 
-A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two
-morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
+The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
+The unit $0$ is both terminal and initial and is called a \emph{zero} object. A semi-additive category
+$\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two morphisms $f, g: X \to Y$
+in $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X
+\to Y$. Diagrammatically:
 
+\vspace{-2mm}
 \begin{center}
@@
 
-\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
-product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
-coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} =
-\coprodM{\inj_1 \comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
+Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
+$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
+The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} = \coprodM{\inj_1
+\comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by composing the unique
+maps in and out of the zero object:
 

### 2025-05-07 main/formal-structures.tex
Biproducts and semi-additive categories.
Galois connections.

@@
 \section{Formal Structures for Generalised Automatic Differentiation}
+
+\subsection{Category of lattices and Galois connections}
+
+First we introduce the category of lattices and Galois connections, which will serve as our base category for
+interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
+posets, where each is the closest approximation to an inverse of the other: one from above, one from below.
+
+\begin{definition}[Galois connection]
+Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
+$f: X \to Y$ and $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
+\end{definition}
+
+\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
+adjunction between poset-as-categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
+referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
+compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition. Thus we can define the
+following category.
 

### 2025-05-07 main/formal-structures.tex
LatGal.

@@
 referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
-compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition. Thus we can define the
-following category.
+compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition, and thus form a category
+where the objects are posets and the morphisms are Galois connections.
+
+As indicated in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
+approximants of a point form a bounded lattice, with least element $\bot$ representing the approximation that
+discards all information and greatest element $\top$ representing the approximation that discards no
+information; so we define the following subcategory.
+
+\begin{definition}
+Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms from $X$ to
+$Y$, all Galois connections $f \adj g: X \to Y$.
+\end{definition}
+
+\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
+in $\LatGal$, $f$ and $g$ preserve the meet-semilattice structure of $X$ and join-semilattice structure of $Y$
+respectively.
 
@@
 
-It is easy to verify that $+$ is associative and commutative and that $f + 0 = f = 0 + f$, and so every
+It is easy to verify that $+$ is associative and commutative and that $f + 0 = f = 0 + f$, and thus that every
 hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
@@
 
-\note{Probably a better way: (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
+% Done:
+% (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category;
+
+\note{Probably a better way: (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
 

### 2025-05-07 main/formal-structures.tex
LatGal.

@@
 interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
-posets, where each is the closest approximation to an inverse of the other: one from above, one from below.
+posets, where each is the closest approximation (one from above, one from below) to an inverse of the other.
 
@@
 \noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
-adjunction between poset-as-categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
+adjunction between poset categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
 referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
@@
 
-As indicated in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
+As shown in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
 approximants of a point form a bounded lattice, with least element $\bot$ representing the approximation that
 discards all information and greatest element $\top$ representing the approximation that discards no
-information; so we define the following subcategory.
+information. So we define the following subcategory.
 
 \begin{definition}
-Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms from $X$ to
-$Y$, all Galois connections $f \adj g: X \to Y$.
+Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms all Galois
+connections $f \adj g: X \to Y$.
 \end{definition}

### 2025-05-08 main/formal-structures.tex
Clearer that composition is like multiplication with zero morphisms as the annihilator.

@@
 
-It is easy to verify that $+$ is associative and commutative and that $f + 0 = f = 0 + f$, and thus that every
+It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
 hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
@@
 \begin{salign*}
-f \comp \zero = f = \zero \comp f
+f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
 \end{salign*}

### 2025-05-08 main/formal-structures.tex
Separate doc for notes.

@@
 The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
-The unit $0$ is both terminal and initial and is called a \emph{zero} object. A semi-additive category
-$\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two morphisms $f, g: X \to Y$
-in $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X
-\to Y$. Diagrammatically:
+The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{Probably want to state the
+biproduct ``laws'' that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in
+$\CMon$, the category of commutative monoids: for any two morphisms $f, g: X \to Y$ in $\cat{C}$, the
+biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X \to Y$.
+Diagrammatically:
 

### 2025-05-08 notes.tex
Enriched hom-functor.
Separate doc for notes.

@@
-\documentclass[acmsmall,nonacm]{acmart}
+\documentclass[acmsmall,screen]{acmart}
 
+\usepackage{tabularx}
 \usepackage{bbm}
@@
 
-\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
+%\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
 \settopmatter{printacmref=false}
+\citestyle{acmauthoryear}
+\raggedbottom
 
@@
 \title{Approximation as Differentiation}
+
+\author{Robert Atkey}
+\email{robert.atkey@strath.ac.uk}
+\orcid{0000-0002-4414-5047}
+\affiliation{%
+  \institution{University of Strathclyde}
+  \city{Glasgow}
+  \country{UK}}
+
+\author{Roly Perera}
+\email{roly.perera@cl.cam.ac.uk}
+\orcid{0000-0001-9249-9862}
+\affiliation{%
+  \institution{University of Cambridge}
+  \city{Cambridge}
+  \country{UK}
+}
+\additionalaffiliation{%
+   \institution{University of Bristol}
+   \city{Bristol}
+   \country{UK}
+}
+
 \maketitle
 
-\section{Overview}
+\section{Notes}
+
+\subsection{Overview}
 
@@
 \begin{itemize}
-\item $\CMon$-enriched categories (\secref{cmon-enriched})
+\item automatic differentiation (\secref{auto-diff})
+\item $\CMon$-enriched categories, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
 \item biproducts and semi-additive categories (\secref{biproduct})
-\item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
-\item set-indexed products (\secref{set-indexed-product})
-\item Grothendieck construction $\Grothendieck{C}F$ and category of families $\Fam(C)$ (\secref{grothendieck})
-\item Examples of useful semi-additive categories (\secref{useful-semi-additive-categories}):
+\item useful semi-additive categories (\secref{useful-semi-additive-categories}):
    \begin{itemize}
@@
    \end{itemize}
+\item category $\Fam(I,\cat{C})$ of $I$-indexed families of objects of $\cat{C}$ (\secref{fam})
+\item set-indexed products (\secref{set-indexed-product})
+\item Grothendieck construction $\Grothendieck{\cat{C}}F$ and category of families $\Fam(\cat{C})$
+(\secref{grothendieck})
+\item useful categories of families for automatic approximation
+(\secref{galois-slicing-auto-diff-via-fam}):
+   \begin{itemize}
+      \item Galois slicing via $\Fam(\LatGal)$ (\secref{galois-slicing-auto-diff-via-fam:galois-slicing})
+      \item automatic differentiation via $\Fam(\FinVect)$ (\secref{galois-slicing-auto-diff-via-fam:auto-diff})
+   \end{itemize}
 \end{itemize}
@@
 
-\input{notes/sec/notation}
+\input{notes/preliminaries}
+\input{notes/auto-diff}
+\input{notes/cmon}
+\input{notes/cmon-enriched}
+\input{notes/biproduct}
+\input{notes/useful-semi-additive-categories}
+\input{notes/fam}
+\input{notes/set-indexed-product}
+\input{notes/grothendieck}
+\input{notes/auto-diff-galois-slicing-via-fam}
+\input{notes/language}
 
-\section{Definitions}
-
-\input{notes/sec/cmon-enriched}
-\input{notes/sec/biproduct}
-\input{notes/sec/fam}
-\input{notes/sec/set-indexed-product}
-\input{notes/sec/grothendieck}
-\input{notes/sec/useful-semi-additive-categories}
-
-% \input{notes/language}
-
-\section{Some references}
-
-\cite{karvonen2020}
-
-\bibliographystyle{plainnat}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{bib}

### 2025-05-08 notes/auto-diff-galois-slicing-via-fam.tex
Separate doc for notes.

--- 2025-05-06 appendix/notes/auto-diff-galois-slicing-via-fam.tex
@@
 $\partial X_x$;
-\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of functions $f: X \to
-Y$ and families of Galois connections $\partial f_x: \partial X_x \to \partial Y_{f(x)}$ for every $x \in X$.
+\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of a function $f: X \to
+Y$ and for every $x \in X$, a Galois connection $\partial f_x: \partial X_x \to \partial Y_{f(x)}$.
 \end{itemize}

### 2025-05-08 notes/auto-diff.tex
Separate doc for notes.

--- 2025-05-06 appendix/notes/auto-diff.tex
@@
 
-\noindent and again define a corresponding category $\Diff^*$ (equivalent to a full subcategory of
-$\Fam(\FinVect^\op)$) of such maps where composition of the cotangent maps respects the backward chain rule.
-
-\todo{Can't combine $\Diff$ and $\Diff^*$ into a single subcategory of $\Fam(\FinVect)$, with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$, because..?}
+\noindent and again define a corresponding category $\Diff^*$ of such maps where composition of the cotangent
+maps respects the backward chain rule. \todo{In practice can't combine $\Diff$ and $\Diff^*$ into a single
+category with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$
+because..?}

### 2025-05-08 notes/biproduct.tex
Separate doc for notes.

--- 2025-04-16 appendix/notes/biproduct.tex
@@
 \end{enumerate}
-\item If $\cat{C}$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
+\item If $\cat{C}$ has binary coproducts $(X + Y, \inj_1, \inj_2)$ then
 \begin{enumerate}
-\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
+\item $\coprodM{0_{X,Z}}{0_{Y,Z}} = 0_{X + Y,Z}$
 \item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
-\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
-biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
-\coprodM{0_{A,B}}{\id_B}$.
+\item $(X + Y, \inj_1, \inj_2, \biproj_X, \biproj_Y)$ is a biproduct with $\biproj_X: X + Y \to X =
+\coprodM{\id_X}{0_{Y,X}}$ and $\biproj_Y: X + Y \to Y = \coprodM{0_{X,Y}}{\id_Y}$.
 \end{enumerate}
@@
 \begin{tikzcd}[row sep=3em, column sep=3em]
-   & A \arrow[d, "\biinj_A", shift left] \arrow[dr, "f_2"] \\
-   A' \arrow[r, "\prodM{f_1}{g_1}"] \arrow[dr, "g_1"'] \arrow[ur, "f_1"] & A \biprod B \arrow[u, "\biproj_A", shift left] \arrow[d, "\biproj_B"', shift right] \arrow[r, pos=0.35, "\coprodM{f_2}{g_2}"] & B' \\
-   & B \arrow[u, "\biinj_B"', shift right] \arrow[ur, "g_2"']
+   & Y \arrow[d, "\biinj_Y", shift left] \arrow[dr, "f_2"] \\
+   X \arrow[r, "\prodM{f_1}{g_1}"] \arrow[dr, "g_1"'] \arrow[ur, "f_1"] & Y \biprod Z \arrow[u, "\biproj_Y", shift left] \arrow[d, "\biproj_Z"', shift right] \arrow[r, pos=0.35, "\coprodM{f_2}{g_2}"] & W \\
+   & B \arrow[u, "\biinj_Z"', shift right] \arrow[ur, "g_2"']
 \end{tikzcd}

### 2025-05-08 notes/language.tex
Separate doc for notes.

--- 2025-04-29 appendix/notes/language.tex
@@
 \item a set $S$ of sorts;
-\item a set $\Op(s_1, \ldots, s_n; s)$ of operations with input sorts $s_1, \ldots, s_n$ and output sort $s$;
-\item a set $\Rel(s_1, \ldots, s_n)$ of $n$-ary relations over sorts $s_1, \ldots, s_n$
+\item a set $\Op(s_1, \ldots, s_n; s)$ of $n$-ary operations with input sorts $s_1, \ldots, s_n$ and output
+sort $s$;
+\item a set $\Rel(s_1, \ldots, s_n)$ of $n$-ary relations over sorts $s_1, \ldots, s_n$.
 \end{itemize}

### 2025-05-08 main/formal-structures.tex
Progress on properties of LatGal.

--- 2025-05-08 main/formal-structures.tex
@@
 First we introduce the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
-posets, where each is the closest approximation (one from above, one from below) to an inverse of the other.
+interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
+X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
+of $g$, and $g$ the best approximation from above to an inverse of $f$.
 
@@
 Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
-$f: X \to Y$ and $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
+$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
 \end{definition}
 
-\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
-adjunction between poset categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
-referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
-compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition, and thus form a category
-where the objects are posets and the morphisms are Galois connections.
+\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
+Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
+$f$ and $g$ interpreted as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
+the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
+as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
+connections between them as morphisms.
 
-As shown in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
-approximants of a point form a bounded lattice, with least element $\bot$ representing the approximation that
-discards all information and greatest element $\top$ representing the approximation that discards no
-information. So we define the following subcategory.
+As sketched in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
+approximants of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$
+representing the approximation that discards all information about $x$, greatest element $\top$ representing
+the approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways
+to combine approximations. Thus rather than working in $\PosGal$, we define the following subcategory.
 
 \begin{definition}
-Define the category $\LatGal$ which has as objects $X, Y$ all bounded lattices, and as morphisms all Galois
-connections $f \adj g: X \to Y$.
+Define the category $\LatGal$ which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded lattices,
+and as morphisms all Galois connections $f \adj g: X \to Y$.
 \end{definition}
@@
 \noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $f$ and $g$ preserve the meet-semilattice structure of $X$ and join-semilattice structure of $Y$
-respectively.
+in $\LatGal$, $f$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
+$(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+
+\subsubsection{Additive structure on morphisms.}
+
+A bounded semilattice is an example of a \emph{commutative monoid} $(X, +, 0)$, namely a set $X$ equipped with
+an associative, commutative binary operation $+$ with unit $0$. One observation about commutative monoids is
+that pointwise lifting of $+$ to homomorphisms $f, g: X \to Y$, where $(f + g)(x) = f(x) + g(x): X \to Y$, is
+also ``additive'', in the sense of being associative and commutative and having a ``zero'' homomorphism
+$0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to $0_Y$, which acts as a unit for addition
+of homomorphisms. The homomorphisms from $X$ to $Y$ thus have a commutative monoid structure themselves.
+
+In particular $\LatGal$ has this structure..
+
+% Moreover since $(X, \top, \meet)$ is a commutative monoid, and thus a $\CMon$-enriched (poset)
+% category, for any $X, Y$ there is a zero meet-semilattice homomorphism $0^{\meet}_{X,Y} = \const(\top_Y): X
+% \to Y$ and also (idempotent) addition of homomorphisms $f +_{\meet} g: X \to Y$ given pointwise by $\meet$.
+% Dually, there is a zero join-semilattice homomorphism $0^{\join}_{X,Y} = \const(\bot_Y): X \to Y$ and
+% idempotent addition of homomorphisms $f +_{\join} g: X \to Y$ given pointwise by $\join$.
 
@@
 The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
-The unit $0$ is both terminal and initial and is called a \emph{zero} object. A semi-additive category
-$\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two morphisms $f, g: X \to Y$
-in $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X
-\to Y$. Diagrammatically:
+The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{State biproduct ``laws''
+that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of
+commutative monoids: for any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a
+way to ``add'' them together, forming a morphism $f + g: X \to Y$. Diagrammatically:
 

### 2025-05-08 main/formal-structures.tex
Progress on properties of LatGal.

@@
 
-As sketched in \secref{introduction:galois-slicing}, for Galois slicing we require a setting where the
-approximants of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$
-representing the approximation that discards all information about $x$, greatest element $\top$ representing
-the approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways
-to combine approximations. Thus rather than working in $\PosGal$, we define the following subcategory.
+As sketched in \secref{introduction:galois-slicing}, Galois slicing requires the approximants of a point $x$
+to form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
+approximation that discards all information about $x$, greatest element $\top$ representing the approximation
+that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
+approximations. Thus rather than working in $\PosGal$, we define the following subcategory.
 

### 2025-05-08 main/formal-structures.tex
Move semi-additive structure on morphisms para to beginning of section on semi-additive categories.

@@
 
-As sketched in \secref{introduction:galois-slicing}, Galois slicing requires the approximants of a point $x$
-to form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
+As sketched in \secref{introduction:galois-slicing}, Galois slicing requires that the approximants of a point
+$x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
 approximation that discards all information about $x$, greatest element $\top$ representing the approximation
 that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
-approximations. Thus rather than working in $\PosGal$, we define the following subcategory.
+approximations. Thus rather than working directly in $\PosGal$, we define the following subcategory.
 
@@
 
-\subsubsection{Additive structure on morphisms.}
+\subsection{Biproducts and Semi-Additive Categories}
 
-A bounded semilattice is an example of a \emph{commutative monoid} $(X, +, 0)$, namely a set $X$ equipped with
-an associative, commutative binary operation $+$ with unit $0$. One observation about commutative monoids is
-that pointwise lifting of $+$ to homomorphisms $f, g: X \to Y$, where $(f + g)(x) = f(x) + g(x): X \to Y$, is
-also ``additive'', in the sense of being associative and commutative and having a ``zero'' homomorphism
-$0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to $0_Y$, which acts as a unit for addition
-of homomorphisms. The homomorphisms from $X$ to $Y$ thus have a commutative monoid structure themselves.
-
-In particular $\LatGal$ has this structure..
-
-% Moreover since $(X, \top, \meet)$ is a commutative monoid, and thus a $\CMon$-enriched (poset)
-% category, for any $X, Y$ there is a zero meet-semilattice homomorphism $0^{\meet}_{X,Y} = \const(\top_Y): X
-% \to Y$ and also (idempotent) addition of homomorphisms $f +_{\meet} g: X \to Y$ given pointwise by $\meet$.
-% Dually, there is a zero join-semilattice homomorphism $0^{\join}_{X,Y} = \const(\bot_Y): X \to Y$ and
-% idempotent addition of homomorphisms $f +_{\join} g: X \to Y$ given pointwise by $\join$.
-
-\subsection{Biproducts and Semi-Additive Categories}
+An important observation about the semilattice homomorphisms that appear in adjoint pairs in $\LatGal$ is that
+pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) = f(x)
+\meet g(x): X \to Y$, gives a notion of ``addition'' for these maps which is associative and commutative, and
+which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to $0_Y$,
+acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from $X$ to
+$Y$, which also have a commutative monoid structure; Galois connections inherit this structure component-wise.
+We now say more generally how this so-called \emph{semi-additive} structure on morphisms arises and explain
+its significance to adapting CHAD to provide a semantic basis for Galois slicing as well as automatic
+differentiation.
 
@@
 
-\vspace{-2mm}
 \begin{center}
@@
 
-% Done:
-% (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category;
-
-\note{Probably a better way: (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
+\note{Probably a better way: (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
 

### 2025-05-08 main/formal-structures.tex
Move LatGal and semi-additive category to after current section 2 intro.

@@
 \section{Formal Structures for Generalised Automatic Differentiation}
-
-\subsection{Category of lattices and Galois connections}
-
-First we introduce the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
-X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
-of $g$, and $g$ the best approximation from above to an inverse of $f$.
-
-\begin{definition}[Galois connection]
-Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
-$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
-\end{definition}
-
-\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
-Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
-$f$ and $g$ interpreted as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
-the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
-as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
-connections between them as morphisms.
-
-As sketched in \secref{introduction:galois-slicing}, Galois slicing requires that the approximants of a point
-$x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
-approximation that discards all information about $x$, greatest element $\top$ representing the approximation
-that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
-approximations. Thus rather than working directly in $\PosGal$, we define the following subcategory.
-
-\begin{definition}
-Define the category $\LatGal$ which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded lattices,
-and as morphisms all Galois connections $f \adj g: X \to Y$.
-\end{definition}
-
-\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $f$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
-$(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
-
-\subsection{Biproducts and Semi-Additive Categories}
-
-An important observation about the semilattice homomorphisms that appear in adjoint pairs in $\LatGal$ is that
-pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) = f(x)
-\meet g(x): X \to Y$, gives a notion of ``addition'' for these maps which is associative and commutative, and
-which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to $0_Y$,
-acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from $X$ to
-$Y$, which also have a commutative monoid structure; Galois connections inherit this structure component-wise.
-We now say more generally how this so-called \emph{semi-additive} structure on morphisms arises and explain
-its significance to adapting CHAD to provide a semantic basis for Galois slicing as well as automatic
-differentiation.
-
-\begin{definition}[Semi-additive category]
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
-The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{State biproduct ``laws''
-that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of
-commutative monoids: for any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a
-way to ``add'' them together, forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} = \coprodM{\inj_1
-\comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by composing the unique
-maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
 
@@
 Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using $\CMon$-enriched presheaves.\todo{elaborate.}
+
+\subsection{Category of lattices and Galois connections}
+
+First we introduce the category of lattices and Galois connections, which will serve as our base category for
+interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
+X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
+of $g$, and $g$ the best approximation from above to an inverse of $f$.
+
+\begin{definition}[Galois connection]
+Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
+$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
+\end{definition}
+
+\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
+Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
+$f$ and $g$ interpreted as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
+the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
+as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
+connections between them as morphisms.
+
+As sketched in \secref{introduction:galois-slicing}, Galois slicing requires that the approximants of a point
+$x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
+approximation that discards all information about $x$, greatest element $\top$ representing the approximation
+that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
+approximations. Thus rather than working directly in $\PosGal$, we define the following subcategory.
+
+\begin{definition}
+Define the category $\LatGal$ which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded lattices,
+and as morphisms all Galois connections $f \adj g: X \to Y$.
+\end{definition}
+
+\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
+in $\LatGal$, $f$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
+$(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+
+\subsection{Biproducts and Semi-Additive Categories}
+
+An important observation about the semilattice homomorphisms that appear in adjoint pairs in $\LatGal$ is that
+the pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) =
+f(x) \meet g(x): X \to Y$, gives a notion of ``addition'' for those maps which is associative and commutative,
+and which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to
+$0_Y$, acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from
+$Y$ to $X$, which also have a commutative monoid structure; Galois connections then inherit this structure
+component-wise.
+
+We now say more generally when this so-called \emph{semi-additive} structure on morphisms arises and explain
+its role in generalising CHAD from automatic differentiation to providing a semantics for Galois slicing as
+well.
+
+\begin{definition}[Semi-additive category]
+A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
+and injections) give an isomorphism
+\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
+\end{definition}
+
+The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
+The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{State the biproduct
+``laws'' that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$, the
+category of commutative monoids: for any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure
+provides a way to ``add'' them together, forming a morphism $f + g: X \to Y$. Diagrammatically:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
+\end{tikzcd}
+\end{center}
+
+Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
+$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
+The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} = \coprodM{\inj_1
+\comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by composing the unique
+maps in and out of the zero object:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
+\end{tikzcd}
+\end{center}
+
+It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
+hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
+family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
+preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+
+\begin{salign*}
+f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
+\end{salign*}
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}

### 2025-05-08 main/formal-structures.tex
Intro to Fam section.

@@
 
+\subsection{A categorical framework for Galois slicing}
+
+We will consider a total language where types denote sets. The goal is to interpret programs as Galois
+connections relating approximations of their input to approximations of their output. Since the appropriate
+lattice of approximations depends on the value being approximated, the semantics of every type $\tau$ must
+have both a ``conventional'' component assigning $\tau$ to a set $X$, and an associated lattice of
+approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type $Y$ with a single free
+variable of type $X$ must be interpreted as a morphism that combining both a function $f: X \to Y$ and a
+family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$.
+
 \subsection{Category of lattices and Galois connections}
@@
 X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
-of $g$, and $g$ the best approximation from above to an inverse of $f$.
+of $g$, and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the
+bidirectional nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the
+least input slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the
+greatest output slice for a given slice of the input).
 

### 2025-05-08 main/formal-structures.tex
Intro to Fam section.

@@
 approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type $Y$ with a single free
-variable of type $X$ must be interpreted as a morphism that combining both a function $f: X \to Y$ and a
-family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$.
+variable of type $X$ must then be interpreted as a morphism which also has two components: a function $f: X
+\to Y$ giving the baseline (unapproximated) behaviour of the program, plus a family of Galois connections
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for program slicing.
 

### 2025-05-08 main/formal-structures.tex
Intro to Fam section.
Grothendieck.

--- 2025-05-08 main/formal-structures.tex
@@
 
+\subsection{A categorical framework for Galois slicing}
+
+For a total language where types denote sets, the goal is to interpret programs as Galois connections relating
+approximations of their input to approximations of their output. Since these lattice of approximations depends
+on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
+component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
+every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
+must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
+the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
+every $x \in X$ which can be used for slicing over that expression.
+
+This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
+category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
+= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
+the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
+between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
+gloms all such $\Set$-indexed families together, with reindexing morphisms to transport between different
+indexing sets.
+
 \subsection{Category of lattices and Galois connections}

### 2025-05-08 main/formal-structures.tex
Recap functor categories [X, C] where X a set.

@@
 
-\subsection{A categorical framework for Galois slicing}
-
-For a total language where types denote sets, the goal is to interpret programs as Galois connections relating
-approximations of their input to approximations of their output. Since these lattice of approximations depends
-on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
-component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
-every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
-must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
-the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
-every $x \in X$ which can be used for slicing over that expression.
-
-This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
-= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
-the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
-between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
-gloms all such $\Set$-indexed families together, with reindexing morphisms to transport between different
-indexing sets.
-
 \subsection{Category of lattices and Galois connections}
@@
 that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
-approximations. Thus rather than working directly in $\PosGal$, we define the following subcategory.
+approximations. Thus rather than working directly in $\PosGal$, we consider the following subcategory instead.
 
 \begin{definition}
-Define the category $\LatGal$ which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded lattices,
-and as morphisms all Galois connections $f \adj g: X \to Y$.
+Define $\LatGal$ to be the category which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded
+lattices, and as morphisms all Galois connections $f \adj g: X \to Y$.
 \end{definition}
@@
 $(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+
+\subsection{A categorical framework for Galois slicing}
+
+For a total language where types denote sets, our goal is to interpret programs as Galois connections relating
+approximations of their inputs to approximations of their outputs. Since these lattice of approximations
+depends on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
+component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
+every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
+must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
+the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
+every $x \in X$ which can be used for slicing over that expression.
+
+This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
+category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
+= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
+the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
+between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
+gloms all such $\Set$-indexed families of bounded lattices into the same category, with reindexing morphisms
+to transport between different indexing sets.
+
+\begin{definition}[Category of $X$-indexed families of objects]
+\label{def:fam:Fam}
+For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
+category $\Func{X}{C}$ has as objects all functors $X \to \cat{C}$, in other words all $X$-indexed families of
+objects of $\cat{C}$, and as morphisms from $X \to \cat{C}$ to $Y \to \cat{C}$, all natural transformations,
+i.e.~families of morphisms $f_x: X_x \to Y_x$ in $\cat{C}$ for every $x \in X$.
+\end{definition}
 

### 2025-05-08 main/formal-structures.tex
Recap functor categories [X, C] where X a set.

@@
 category $\Func{X}{C}$ has as objects all functors $X \to \cat{C}$, in other words all $X$-indexed families of
-objects of $\cat{C}$, and as morphisms from $X \to \cat{C}$ to $Y \to \cat{C}$, all natural transformations,
-i.e.~families of morphisms $f_x: X_x \to Y_x$ in $\cat{C}$ for every $x \in X$.
+objects of $\cat{C}$, and as morphisms from $X \to \cat{C}$ to $Y \to \cat{C}$, all families of morphisms
+$f_x: X_x \to Y_x$ in $\cat{C}$ for every $x \in X$, with naturality in $X$ immediate.
 \end{definition}

### 2025-05-08 main/formal-structures.tex
Recap functor categories [X, C] where X a set.

--- 2025-05-08 main/formal-structures.tex
@@
 
-\begin{definition}[Category of $X$-indexed families of objects]
+\begin{definition}[Functor category of $X$-indexed families]
 \label{def:fam:Fam}
 For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
-category $\Func{X}{C}$ has as objects all functors $X \to \cat{C}$, in other words all $X$-indexed families of
-objects of $\cat{C}$, and as morphisms from $X \to \cat{C}$ to $Y \to \cat{C}$, all natural transformations,
-i.e.~families of morphisms $f_x: X_x \to Y_x$ in $\cat{C}$ for every $x \in X$.
+category $\Func{X}{C}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of objects
+of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$
+for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
 \end{definition}

### 2025-05-08 main/formal-structures.tex
Recap functor categories [X, C] where X a set.

@@
 
+\begin{definition}[Reindexing]
+For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
+\Func(X,\cat{C})$ which:
+\begin{itemize}
+\item sends any $J$-indexed family $X$ to the $I$-indexed family $\reindex{X}{f}$ where $\reindex{X}{f}_i =
+X_{f(i)}$ for any $i \in I$;
+\item sends any morphism $g: X \to Y$ between $J$-indexed families to the morphism $\reindex{g}{f}:
+\reindex{X}{f} \to \reindex{Y}{f}$ where $\reindex{g}{f}_i = g_{f(i)}$.
+\end{itemize}
+\end{definition}
+
 \subsection{Biproducts and Semi-Additive Categories}

### 2025-05-08 main/formal-structures.tex
Reindexing.
Yes prefer this version of families because reindexing is nicer.

--- 2025-05-08 main/formal-structures.tex
@@
 gloms all such $\Set$-indexed families of bounded lattices into the same category, with reindexing morphisms
-to transport between different indexing sets.
+transporting between data indexed by different sets.
 
@@
 For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
-category $\Func{X}{C}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of objects
-of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$
-for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
+category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
+objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
+$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
+\end{definition}
+
+Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits from its codomain.
+
+\begin{proposition}
+If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
+pointwise.
+\end{proposition}
+
+\begin{definition}[Reindexing]
+For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
+\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
+$\reindex{F}{f} = F \comp f$, with $f$ interpreted as the unique functor between $X$ and $Y$ as discrete
+categories.
+\end{definition}
+
+\begin{definition}
+Let $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$ denote the indexed category which sends any set $X$ to the
+category $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the functor $\reindex{-}{f}: \Func{Y}{\cat{C}}
+\to \Func{X}{\cat{C}}$.
 \end{definition}

### 2025-05-08 main/formal-structures.tex
Grothendieck.
Galois slicing via Grothendieck.

@@
 \Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
-$\reindex{F}{f} = F \comp f$, with $f$ interpreted as the unique functor between $X$ and $Y$ as discrete
-categories.
+$\reindex{F}{f} = F \comp f$, regarding $f$ as a functor between $X$ and $Y$ as discrete categories.
 \end{definition}
@@
 \end{definition}
+
+\begin{definition}
+\label{def:grothendieck:Grothendieck}
+Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
+category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
+\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
+\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
+morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
+\end{definition}
+
+\noindent The $\partial X$ is intended to suggest the connection between the Grothendieck construction and the
+idea of tangent spaces and derivatives. To understand how the $\partial X$ components compose, consider any
+$(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial
+Z)$ in $\Grothendieck{}T$. We have:
+
+\begin{itemize}
+\item $\partial f: \partial X \to T(f)(\partial Y)$;
+\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$.
+\end{itemize}
+
+\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the functor $T(f)$. The composition $(g,
+\partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g \comp f,
+T(f)(\partial g) \comp \partial f)$.
+
+\subsubsection{Galois slicing via the Grothendieck construction}
+
+Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
+programs. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded
+lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of
+a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to
+\partial Y(f(x))$, as required. The role of the pullback (reindexing) along $f$ in the composition $(g,
+\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ is to select the
+appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
+function $f$.
 

### 2025-05-08 main/formal-structures.tex
Done for now.

@@
 
+\begin{definition}[Category of families]
+For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
+$\Grothendieck{X}\Func{X}{\cat{C}}$.
+\end{definition}
+
+\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
+$\partial X$ in $\Fam(X,\cat{C})$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are
+functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in
+$\Fam(X,\cat{C})$.
+
 \subsubsection{Galois slicing via the Grothendieck construction}
@@
 Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
-programs. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded
-lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of
-a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to
-\partial Y(f(x))$, as required. The role of the pullback (reindexing) along $f$ in the composition $(g,
-\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ is to select the
+programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
+$x \in X$, a bounded lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs
+$(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x):
+\partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition $(g,
+\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
 appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)

### 2025-05-08 main/formal-structures.tex

@@
 \noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $f$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
-$(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+in $\LatGal$, $g$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
+$(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
 
@@
 For a total language where types denote sets, our goal is to interpret programs as Galois connections relating
-approximations of their inputs to approximations of their outputs. Since these lattice of approximations
+approximations of their inputs to approximations of their outputs. Since these lattices of approximations
 depends on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
@@
 
-\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the functor $T(f)$. The composition $(g,
-\partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g \comp f,
-T(f)(\partial g) \comp \partial f)$.
+\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
+composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
+\comp f, T(f)(\partial g) \comp \partial f)$.
 

### 2025-05-08 notes/cmon-enriched.tex

--- 2025-04-18 appendix/notes/cmon-enriched.tex
@@
 $\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
-and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations $\Yoneda_{\mathscr{V}}(f):
-\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as post-composition with $f$.
+and providing a family of morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to
+\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{\Hom{\cat{C}}{-}{Y}}$ acting
+concretely at each component as post-composition with $f: X \to Y$.
 

### 2025-05-08 notes/cmon.tex

\subsection{Commutative monoids}

\begin{definition}[Commutative monoid]
A \emph{commutative monoid} $X = (X, \bullet, \varepsilon)$ is a set $X$ equipped with distinguished element
$\varepsilon \in X$ called the \emph{unit} and associative binary operation $\bullet: X^2 \to X$ satisfying
$\varepsilon \bullet x = x$ and $x \bullet \varepsilon = x$ for any $x \in X$.
\end{definition}

A commutative monoid homomorphism from $X$ to $Y$ is any function $f: X \to Y$ preserving $\varepsilon$ and
$\bullet$.

\subsubsection{Category of commutative monoids}

\begin{definition}[Category $\CMon$]
Define $\CMon$ to be the category which has as objects $X$ all commutative monoids and as morphisms $f: X \to
Y$ all commutative monoid homomorphisms.
\end{definition}

$\CMon$ is complete and cocomplete, inheriting all limits and colimits from $\Set$. $\CMon$ is also monoidal
closed. \todo{But the monoidal product is not the Cartesian product?}

\subsubsection{Example of commutative monoid}

\begin{definition}[Bounded semilattice]
\label{def:cmon-enriched:bounded-semilattice}
A \emph{bounded semilattice} $X = (X, \bullet, \varepsilon)$ is a commutative monoid where $\bullet$ is
\emph{idempotent}, i.e.~satisfies $x \bullet x = x$.
\end{definition}

\noindent The idempotence of $\bullet$ (together with commutativity and associativity) induces a partial order
$\le_\bullet$ on $X$, with $x \le_{\bullet} y \iff x \bullet y = x$. With respect to this partial order, $x
\bullet y$ is the greatest lower bound (meet) of $x$ and $y$ and $\varepsilon$ is the top element; with
respect to the opposite order, $x \bullet y$ is the least upper bound (join) of $x$ and $y$ and $\varepsilon$
is the bottom element. This therefore provides an algebraic characterisation of the usual (dual)
order-theoretic notions of bounded meet semilattices $(X, \meet, \top)$ and bounded join semilattices $(X,
\join, \bot)$.

\begin{definition}[Category $\SemiLat$]
Define $\SemiLat$ to be the category which has as objects all bounded semilattices and as morphisms all
bounded semilattice homomorphisms.
\end{definition}

### 2025-05-09 main/formal-structures.tex
Minor corrections.

@@
 
-\noindent The $\partial X$ is intended to suggest the connection between the Grothendieck construction and the
-idea of tangent spaces and derivatives. To understand how the $\partial X$ components compose, consider any
-$(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial
-Z)$ in $\Grothendieck{}T$. We have:
+\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
+construction and the idea of tangent spaces and derivatives. To understand how the $\partial X$ components
+compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
+\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
 
@@
 \noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
-$\partial X$ in $\Fam(X,\cat{C})$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are
-functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in
-$\Fam(X,\cat{C})$.
+$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
+$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in $\Fam(X,\cat{C})$.
 
@@
 programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
-$x \in X$, a bounded lattice $\partial X_x$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs
-$(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x):
-\partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition $(g,
-\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
+$x \in X$, a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all
+pairs $(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial
+f(x): \partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition
+$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
 appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)

### 2025-05-16 main/formal-structures.tex
Tweaks to 2.2 intro.

@@
 
-For a total language where types denote sets, our goal is to interpret programs as Galois connections relating
-approximations of their inputs to approximations of their outputs. Since these lattices of approximations
-depends on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
-component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
-every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
-must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
-the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
-every $x \in X$ which can be used for slicing over that expression.
+We will consider a pure, total language, where types can be interpreted as sets. Our goal is to interpret
+programs as Galois connections relating approximations of their inputs to approximations of their outputs.
+Since these lattices of approximations depends on the value being approximated, the semantics of every type
+$\tau$ must have both a ``conventional'' component, providing the set $X$ associated with $\tau$, plus an
+associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The semantics of an expression
+$e$ of type $Y$ with a single free variable of type $X$ must then also have two components: a function $f: X
+\to Y$ giving the baseline (unapproximated) meaning of the expression, plus a family of Galois connections
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for slicing over
+that expression.
 

### 2025-05-16 main/formal-structures.tex
No need to 'define' the [-,C] functor, just introduce informally before Fam(C).

@@
 \begin{definition}
-Let $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$ denote the indexed category which sends any set $X$ to the
-category $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the functor $\reindex{-}{f}: \Func{Y}{\cat{C}}
-\to \Func{X}{\cat{C}}$.
-\end{definition}
-
-\begin{definition}
 \label{def:grothendieck:Grothendieck}
@@
 \noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
-construction and the idea of tangent spaces and derivatives. To understand how the $\partial X$ components
+construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
 compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
@@
 \comp f, T(f)(\partial g) \comp \partial f)$.
+
+Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
+category which sends any set $X$ to the functor category $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to
+the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$.
 

### 2025-05-16 main/formal-structures.tex
Prefer direct composition notation for reindexing functor.

@@
 
+\subsubsection{Grothendieck construction for indexed categories}
+
 This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
@@
 the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
-between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
-gloms all such $\Set$-indexed families of bounded lattices into the same category, with reindexing morphisms
-transporting between data indexed by different sets.
+between them as morphisms. The Grothendieck construction $\Grothendieck{}T$
+(\defref{grothendieck:Grothendieck} below) is the category of all $\Set$-indexed families of bounded lattices,
+together with morphisms that account for both structure within each family and how families transform under
+reindexing.
 
-\begin{definition}[Functor category of $X$-indexed families]
+First we formalise $X$-indexed families and their categories.
+
+\begin{definition}[Category of $X$-indexed families]
 \label{def:fam:Fam}
@@
 
-\begin{definition}[Reindexing]
-For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
-\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
-$\reindex{F}{f} = F \comp f$, regarding $f$ as a functor between $X$ and $Y$ as discrete categories.
-\end{definition}
+Any function $f: X \to Y$ induces a \emph{reindexing} functor $- \comp f: \Func{Y}{\cat{C}} \to
+\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family $F \comp f$
+(regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any family of morphisms $\eta: F
+\to G$ to $\eta \comp f: F \comp f \to G \comp f$ with $(\eta \comp f)_x = \eta_{f(x)}$.
 
@@
 
-Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
-category which sends any set $X$ to the functor category $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to
-the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$.
+Of interest here is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
+category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
+$f: X \to Y$ to the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .
 
@@
 $\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
-$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in $\Fam(X,\cat{C})$.
+$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial Y \comp f$ in $\Func{X}{\cat{C}}$.
 

### 2025-05-16 main/formal-structures.tex
Prefer direct composition notation for reindexing functor.

@@
 f(x): \partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition
-$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
+$(g, \partial g) \comp (f, \partial f) = (g \comp f, \partial g \comp f \comp \partial f)$ selects the
 appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)

### 2025-05-16 main/formal-structures.tex
Prefer direct composition notation for reindexing functor.
Nope, need to keep reindexing notation distinct from composition otherwise things get confusing.

@@
 
-Any function $f: X \to Y$ induces a \emph{reindexing} functor $- \comp f: \Func{Y}{\cat{C}} \to
-\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family $F \comp f$
-(regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any family of morphisms $\eta: F
-\to G$ to $\eta \comp f: F \comp f \to G \comp f$ with $(\eta \comp f)_x = \eta_{f(x)}$.
+\begin{definition}[Reindexing]
+For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
+\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
+$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
+family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
+$\reindex{\eta}{f}(x) = \eta(f(x))$.
+\end{definition}
+
+Although reindexing just precomposes with $f$, writing it as $\reindex{-}{f}$ avoids notational confusion when
+combining composition of morphisms in $\Func{X}{C}$ with reindexing.
 

### 2025-05-16 main/formal-structures.tex
Nope, need to keep reindexing notation distinct from composition otherwise things get confusing.

@@
 
-Of interest here is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
+Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
 category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
-$f: X \to Y$ to the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .
+$f: X \to Y$ to the reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .
 

### 2025-05-16 main/formal-structures.tex
Going to introduce enrichment of LatGal earlier.

@@
 $(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+
+\todo{Introduce that $\LatGal$ is $\SemiLat$-enriched -- the hom-set $\LatGal(X)(Y)$ is has an (idempotent) addition of morphisms and a zero morphism.
+Moreover it has biproducts -- product of lattices $X \times Y$ is also a coproduct, although formal definition will come later.}
+
+% the pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) =
+% f(x) \meet g(x): X \to Y$, gives a notion of ``addition'' for those maps which is associative and commutative,
+% and which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to
+% $0_Y$, acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from
+% $Y$ to $X$, which also have a commutative monoid structure; Galois connections then inherit this structure
+% component-wise.
 
@@
 $\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
-$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial Y \comp f$ in $\Func{X}{\cat{C}}$.
+$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
+$\Func{X}{\cat{C}}$.
 
@@
 f(x): \partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition
-$(g, \partial g) \comp (f, \partial f) = (g \comp f, \partial g \comp f \comp \partial f)$ selects the
+$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
 appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
@@
 
-An important observation about the semilattice homomorphisms that appear in adjoint pairs in $\LatGal$ is that
-the pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) =
-f(x) \meet g(x): X \to Y$, gives a notion of ``addition'' for those maps which is associative and commutative,
-and which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to
-$0_Y$, acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from
-$Y$ to $X$, which also have a commutative monoid structure; Galois connections then inherit this structure
-component-wise.
-
-We now say more generally when this so-called \emph{semi-additive} structure on morphisms arises and explain
-its role in generalising CHAD from automatic differentiation to providing a semantics for Galois slicing as
-well.
+We noted that $\LatGal$ is $\CMon$-enriched (actually $\SemiLat$-enriched) and moreover that the product of
+lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact having
+biproducts is sufficient to ensure $\CMon$-enrichment; such a category is called \emph{semi-additive}.
 

### 2025-05-16 main/formal-structures.tex
LatGal has biproducts.

@@
 
-\todo{Introduce that $\LatGal$ is $\SemiLat$-enriched -- the hom-set $\LatGal(X)(Y)$ is has an (idempotent) addition of morphisms and a zero morphism.
-Moreover it has biproducts -- product of lattices $X \times Y$ is also a coproduct, although formal definition will come later.}
+One observation about the (bounded) meet-semilattice homomorphisms from $X$ to $Y$ is that they admit a
+meet-semilattice structure themselves: the pointwise lifting of $\meet$ to meet-preserving maps $f, g: X \to
+Y$, where $(f \meet g)(x) = f(x) \meet g(x): X \to Y$ also preserves meets, giving a notion of $\meet$ for
+morphisms, as does the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, and which acts a
+unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ also have a join-semilattice
+structure given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$
+to $\bot_Y$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
+semilattices and semilattice homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with
+unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
+connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$.
 
-% the pointwise lifting of $\meet$ to meet-semilattice homomorphisms $f, g: X \to Y$, where $(f \meet g)(x) =
-% f(x) \meet g(x): X \to Y$, gives a notion of ``addition'' for those maps which is associative and commutative,
-% and which has a ``zero'' homomorphism $0_{X,Y}: X \to Y$ given by the constant map sending any $x \in X$ to
-% $0_Y$, acting as a unit for addition of homomorphisms. The same can be said for the join-preserving maps from
-% $Y$ to $X$, which also have a commutative monoid structure; Galois connections then inherit this structure
-% component-wise.
+An further important property is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$,
+where $X \times Y$ denotes the product of lattices, have both and upper adjoints. This means $X \times Y$ is
+both a product and a coproduct in $\LatGal$, which we write as $X \biprod Y$, with projections $\biproj_X$ and
+$\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$.
+
+\vspace{-4mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   \biproj_X = \prodM{\id_X}{\bot_Y} \adj \proj_1: X \biprod Y \to X \\
+   \biproj_Y = \prodM{\bot_X}{\id_Y} \adj \proj_2: X \biprod Y \to Y
+\end{align*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   \biinj_X = \pi_1 \adj \prodM{\id_X}{\top_Y} : X \to X \biprod Y \\
+   \biinj_Y = \pi_2 \adj \prodM{\top_X}{\id_Y}: Y \to X \biprod Y
+\end{align*}
+\end{center}
+\end{minipage}
+\vspace{2mm}
+
+\noindent The trivial 1-point lattice is the unit for $\biprod$ and is both terminal and initial. Although the
+presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct structure
+will prove crucial for a establishing a Cartesian closed setting for interpreting higher-order programs.
 
@@
 
-We will consider a pure, total language, where types can be interpreted as sets. Our goal is to interpret
-programs as Galois connections relating approximations of their inputs to approximations of their outputs.
-Since these lattices of approximations depends on the value being approximated, the semantics of every type
-$\tau$ must have both a ``conventional'' component, providing the set $X$ associated with $\tau$, plus an
-associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The semantics of an expression
-$e$ of type $Y$ with a single free variable of type $X$ must then also have two components: a function $f: X
-\to Y$ giving the baseline (unapproximated) meaning of the expression, plus a family of Galois connections
-$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for slicing over
-that expression.
+We will consider a pure, total language, where types can be interpreted as sets. Our goal is to equip the
+interpretations of programs with Galois connections relating approximations of their inputs to approximations
+of their outputs. Since these lattices of approximations depends on the value being approximated, the
+semantics of every type $\tau$ must have both a ``conventional'' component, providing the set $X$ associated
+with $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The
+semantics of an expression $e$ of type $Y$ with a single free variable of type $X$ must then also have two
+components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of the expression, plus a
+family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can
+be used for slicing over that expression.
 
@@
 
-Although reindexing just precomposes with $f$, writing it as $\reindex{-}{f}$ avoids notational confusion when
+Although it just precomposes with $f$, writing reindexing as $\reindex{-}{f}$ avoids notational confusion when
 combining composition of morphisms in $\Func{X}{C}$ with reindexing.

### 2025-05-16 main/formal-structures.tex
LatGal has biproducts.

@@
 
-An further important property is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$,
-where $X \times Y$ denotes the product of lattices, have both and upper adjoints. This means $X \times Y$ is
-both a product and a coproduct in $\LatGal$, which we write as $X \biprod Y$, with projections $\biproj_X$ and
-$\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$.
+Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
+\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This means $X
+\times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
+projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$.
 
@@
 
-\noindent The trivial 1-point lattice is the unit for $\biprod$ and is both terminal and initial. Although the
-presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct structure
-will prove crucial for a establishing a Cartesian closed setting for interpreting higher-order programs.
+\noindent The trivial 1-point lattice, which is both terminal and initial, is the unit for $\biprod$. Although
+the presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct
+structure will prove crucial for a establishing a Cartesian closed setting for interpreting higher-order
+programs for Galois slicing.
 

### 2025-05-16 main/formal-structures.tex
LatGal has biproducts.

@@
 morphisms, as does the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, and which acts a
-unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ also have a join-semilattice
-structure given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$
-to $\bot_Y$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
+unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ have a join-semilattice structure
+given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to
+$\bot_X$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
 semilattices and semilattice homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with
@@
 \times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
-projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$.
+projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
 
@@
 the presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct
-structure will prove crucial for a establishing a Cartesian closed setting for interpreting higher-order
+structure will prove crucial for establishing a Cartesian closed setting for interpreting higher-order
 programs for Galois slicing.

### 2025-05-16 notes/preliminaries.tex
Notation for PSh(C) and PSc_V(C).

--- 2025-04-18 appendix/notes/preliminaries.tex
@@
 \begin{itemize}
-\item $\coprodM{f}{g}: X + Y \to Z$ for the coproduct of morphisms $f: X \to Z$ and $g: Y \to Z$
+\item $\coprodM{f}{g}: X \coprod Y \to Z$ for the coproduct of morphisms $f: X \to Z$ and $g: Y \to Z$
 \item $\prodM{f}{g}: X \to Y \times Z$ for the product of morphisms $f: X \to Y$ and $g: X \to Z$
@@
 \end{definition}
+
+\subsubsection{Presheaves and $\mathscr{V}$-enriched presheaves}
+
+Write:
+\begin{itemize}
+\item $\PSh(C)$ for the category of presheaves $\Func{C^\op}{\Set}$
+\item $\PSh_{\mathscr{V}}(C)$ for the category of $\mathscr{V}$-enriched presheaves
+$\Func{C^\op}{\mathscr{V}}$.
+\end{itemize}

### 2025-05-16 notes/cmon-enriched.tex
Roll out PSh_V notation.

--- 2025-04-18 appendix/notes/cmon-enriched.tex
@@
 $Y, Z$ in $\cat{C}$. Moreover there exists an enriched Yoneda embedding, namely a functor
-$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which is itself
-$\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
-and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations $\Yoneda_{\mathscr{V}}(f):
-\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as post-composition with $f$.
+$\Yoneda_{\mathscr{V}}: \cat{C} \to \PSh_{\mathscr{V}}(\cat{C})$ which is itself $\mathscr{V}$-enriched,
+sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$ and providing a family of
+morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to
+\Hom{\PSh_{\mathscr{V}}(\cat{C})}{\Hom{\cat{C}}{-}{X}}{\Hom{\cat{C}}{-}{Y}}$ acting concretely at each
+component as post-composition with $f: X \to Y$.
+
+The $\CMon$-enriched Yoneda embedding preserves biproducts:
+
+\begin{proposition}
+\end{proposition}
 
@@
 \mathscr{V}$ and any object $X$ in $\cat{C}$:
-\[\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{F} \iso F(X)\]
+\[\Hom{\PSh_{\mathscr{V}}{(\cat{C})}}{\Hom{\cat{C}}{-}{X}}{F} \iso F(X)\]
 \end{lemma}

### 2025-05-16 notes/cmon-enriched.tex
Show why CMon-enriched Yoneda embedding preserves biproducts.

@@
 
-The $\CMon$-enriched Yoneda embedding preserves biproducts:
+The $\CMon$-enriched Yoneda embedding preserves biproducts, where (as in any functor category) biproducts in
+$\PSh_{\CMon}(\cat{C})$ are computed pointwise.
 
 \begin{proposition}
+$\Yoneda_{\CMon}(X \biprod Y) \iso \Yoneda_{\CMon}(X) \biprod \Yoneda_{\CMon}(Y)$.
 \end{proposition}
+
+\begin{proof}
+Suppose $Z$ an object in $C$. Then $\Yoneda_{\CMon}(X \biprod Y)(Z) = \Hom{\cat{C}}{Z}{X \biprod Y}$. But for
+any $f: Z \to X \biprod Y$ in $\cat{C}$, there exist unique morphisms $f_1: Z \to X$ and $f_2: Z \to Y$ such
+that $f = \biinj_X \comp f_1 + \biinj_Y \comp f_2$ by the universal property of the biproduct. Thus we have
+$\Hom{\cat{C}}{Z}{X \biprod Y} \iso \Hom{\cat{C}}{Z}{X} \biprod \Hom{\cat{C}}{Z}{Y} = \Yoneda_{\CMon}(X)(Z)
+\biprod \Yoneda_{\CMon}(Y)(Z) = (\Yoneda_{\CMon}(X) \biprod \Yoneda_{\CMon}(Y))(Z)$.
+\end{proof}
 

### 2025-05-19 notes/cmon-enriched.tex
WIP.

@@
 
-The $\CMon$-enriched Yoneda embedding preserves biproducts, where (as in any functor category) biproducts in
-$\PSh_{\CMon}(\cat{C})$ are computed pointwise.
+The $\CMon$-enriched Yoneda embedding \todo{think this works for any monoidal closed $\mathscr{V} with
+biproducts?$} preserves biproducts, where (as in any functor category) biproducts in $\PSh_{\CMon}(\cat{C})$
+are computed pointwise, and noting that $\CMon$ itself has biproducts.
 

### 2025-05-19 notes/cmon-enriched.tex
CMon is itself semi-additive.

--- 2025-05-16 notes/cmon-enriched.tex
@@
 
-The $\CMon$-enriched Yoneda embedding preserves biproducts, where (as in any functor category) biproducts in
-$\PSh_{\CMon}(\cat{C})$ are computed pointwise.
+\todo{Generalise to any monoidal closed $\mathscr{V}$ with biproducts?} The $\CMon$-enriched Yoneda embedding
+preserves biproducts, where (as in any functor category) biproducts in $\PSh_{\CMon}(\cat{C})$ are computed
+pointwise, and noting that $\CMon$ itself has biproducts.
 

### 2025-05-19 notes/cmon.tex
CMon is itself semi-additive.

@@
 \begin{definition}[Category $\CMon$]
+\label{def:cmon:cmon}
 Define $\CMon$ to be the category which has as objects $X$ all commutative monoids and as morphisms $f: X \to

### 2025-05-19 notes/cmon-enriched.tex
Clean up CMon-enriched Yoneda preserves biproducts.

--- 2025-05-16 notes/cmon-enriched.tex
@@
 
-The $\CMon$-enriched Yoneda embedding preserves biproducts, where (as in any functor category) biproducts in
-$\PSh_{\CMon}(\cat{C})$ are computed pointwise.
+\todo{Generalise to any monoidal closed $\mathscr{V}$ with biproducts?} $\CMon$ is monoidal closed, and has
+biproducts (\secref{categories-with-biproducts:cmon}) which are inherited pointwise into
+$\PSh_{\CMon}(\cat{C})$ and preserved by the $\CMon$-enriched Yoneda embedding.
 

### 2025-05-19 main/formal-structures.tex

@@
 
-Although it just precomposes with $f$, writing reindexing as $\reindex{-}{f}$ avoids notational confusion when
-combining composition of morphisms in $\Func{X}{C}$ with reindexing.
+Although reindexing is just precomposition with $f$, writing it as $\reindex{-}{f}$ avoids notational
+confusion when combining composition of morphisms in $\Func{X}{C}$ with reindexing.
 
@@
 \begin{itemize}
-\item $\partial f: \partial X \to T(f)(\partial Y)$;
-\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$.
+\item $\partial f: \partial X \to T(f)(\partial Y)$, and
+\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$
 \end{itemize}

### 2025-05-20 main/formal-structures.tex
Fix def of GC.

@@
 Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
-$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
+$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
 \end{definition}

### 2025-05-20 main/formal-structures.tex
Fill in some fwd refs.

@@
 The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
-An object of the Grothendieck construction $\int T$ is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
+An object of the Grothendieck construction $\int T$~, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
 The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
@@
 
-In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$.
+In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$~(\secref{Fam}).
 As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
@@
 When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
-If the coproducts are actually {\em biproducts}, then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
-
-\todo{Preceeding paragraph needs references to later in the document where these constructions are detailed in full.}
+If the coproducts are actually {\em biproducts}~(\secref{biproducts}), then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
 
@@
 
-As sketched in \secref{introduction:galois-slicing}, Galois slicing requires that the approximants of a point
-$x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing the
-approximation that discards all information about $x$, greatest element $\top$ representing the approximation
-that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to combine
-approximations. Thus rather than working directly in $\PosGal$, we consider the following subcategory instead.
+As sketched in \secref{introduction:galois-slicing}, for Galois slicing we want a setting where the
+approximants of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$
+representing the approximation that discards all information about $x$, greatest element $\top$ representing
+the approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways
+to combine approximations. Thus rather than working directly in $\PosGal$, we consider the following
+subcategory instead.
 
@@
 \noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $g$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
-$(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.
+in $\LatGal$, $g$ is a (bounded) \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice
+structure $(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join,
+\bot)$.
 
-One observation about the (bounded) meet-semilattice homomorphisms from $X$ to $Y$ is that they admit a
-meet-semilattice structure themselves: the pointwise lifting of $\meet$ to meet-preserving maps $f, g: X \to
-Y$, where $(f \meet g)(x) = f(x) \meet g(x): X \to Y$ also preserves meets, giving a notion of $\meet$ for
-morphisms, as does the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, and which acts a
-unit for $\meet$. Dually the (bounded) join-preserving maps from $Y$ to $X$ have a join-semilattice structure
-given pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to
-$\bot_X$. Since these two constructions are adjoint, $\LatGal$ is enriched in the category $\SemiLat$ of
-semilattices and semilattice homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with
-unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
-connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$.
+The meet-semilattice homomorphisms from $X$ to $Y$ form a meet-semilattice themselves, with $\meet$ on
+homomorphisms defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and the constant map
+$\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$ providing the unit (and both preserving meets). Dually
+the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by $\join$ and the
+constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two constructions
+come in adjoint pairs, $\LatGal$ is enriched in the category $\SemiLat$ of semilattices and semilattice
+homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit $\top_{X,Y}$ given by the
+Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois connections $(f \adj g) \meet (f'
+\adj g') = (f \join f') \adj (g \meet g'): X \to Y$.
 
@@
 \subsubsection{Grothendieck construction for indexed categories}
+\label{sec:Grothendieck}
 
@@
 the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
-between them as morphisms. The Grothendieck construction $\Grothendieck{}T$
-(\defref{grothendieck:Grothendieck} below) is the category of all $\Set$-indexed families of bounded lattices,
-together with morphisms that account for both structure within each family and how families transform under
-reindexing.
+between them as morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) is
+the category of all $\Set$-indexed families of bounded lattices, together with morphisms that account for both
+structure within each family and how families transform under reindexing.
 
@@
 \begin{definition}[Category of $X$-indexed families]
-\label{def:fam:Fam}
 For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
@@
 \begin{definition}
-\label{def:grothendieck:Grothendieck}
+\label{def:Grothendieck}
 Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
@@
 
+\subsubsection{Category of families}
+\label{sec:Fam}
+
 Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
@@
 \begin{definition}[Category of families]
+\label{def:Fam}
 For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
@@
 \subsection{Biproducts and Semi-Additive Categories}
+\label{sec:biproducts}
 

### 2025-05-20 main/formal-structures.tex
Tweaks.

@@
 
-We noted that $\LatGal$ is $\CMon$-enriched (actually $\SemiLat$-enriched) and moreover that the product of
-lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact having
-biproducts is sufficient to ensure $\CMon$-enrichment; such a category is called \emph{semi-additive}.
+We noted that $\LatGal$ is enriched in $\CMon$, the category of commutative monoids (in fact in $\SemiLat$)
+and moreover that the product of lattices is also a coproduct, making $\LatGal$ into a category with
+\emph{biproducts}. In fact the presence of biproducts is sufficient for $\CMon$-enrichment; such a category is
+called \emph{semi-additive}.
 
@@
 
-The product/coproduct is called a \emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$.
+The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$.
 The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{State the biproduct
-``laws'' that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$, the
-category of commutative monoids: for any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure
-provides a way to ``add'' them together, forming a morphism $f + g: X \to Y$. Diagrammatically:
+``laws'' that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$: for
+any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them
+together, forming a morphism $f + g: X \to Y$. Diagrammatically:
 

### 2025-05-20 main/formal-structures.tex
Biproducts following from CMon enrichment plus products/coproducts.

@@
 
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$.
-The unit $0$ is both terminal and initial and is called a \emph{zero} object. \todo{State the biproduct
-``laws'' that follow from this formulation.} A semi-additive category $\cat{C}$ is enriched in $\CMon$: for
-any two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them
-together, forming a morphism $f + g: X \to Y$. Diagrammatically:
+The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
+projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
+initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
+two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
+forming a morphism $f + g: X \to Y$. Diagrammatically:
 
@@
 $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} = \coprodM{\inj_1
-\comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by composing the unique
-maps in and out of the zero object:
+The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
+\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
+composing the unique maps in and out of the zero object:
 
@@
 \end{salign*}
+
+\subsubsection{Biproduct laws}
+It is also easy to show that the following equations hold:
+
+\vspace{-3mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_X \comp \biinj_X &= \id_X \\
+   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
+\end{salign*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_Y \comp \biinj_Y &= \id_Y \\
+   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
+\end{salign*}
+\end{center}
+\end{minipage}
+
+\noindent and
+
+\begin{salign*}
+(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
+\end{salign*}
+
+We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
+\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
+providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
+biproducts.
+
+\begin{proposition}
+\label{prop:biproducts:from-product-or-coproduct}
+Suppose $\cat{C}$ is $\CMon$-enriched.
+\begin{enumerate}
+\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
+is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
+\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
+\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \inj_1, \inj_2, \biproj_X,
+\biproj_Y)$ is a biproduct, where $\biproj_X: X \coprod Y \to X = \coprodM{\id_X}{0_{Y,X}}$ and $\biproj_Y: X
+\coprod Y \to Y = \coprodM{0_{X,Y}}{\id_Y}$.
+\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
+\end{enumerate}
+\end{proposition}

### 2025-05-20 main/formal-structures.tex
Fwd ref to biproducts:from-product-or-coproduct.

@@
 
-Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of them having products and being enriched in commutative monoids\todo{fwd ref}.
+Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of their having products and being enriched in commutative monoids~(\propref{biproducts:from-product-or-coproduct} below).
 However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
@@
 \begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
+\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
 is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
 \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \inj_1, \inj_2, \biproj_X,
-\biproj_Y)$ is a biproduct, where $\biproj_X: X \coprod Y \to X = \coprodM{\id_X}{0_{Y,X}}$ and $\biproj_Y: X
-\coprod Y \to Y = \coprodM{0_{X,Y}}{\id_Y}$.
+\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
+\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
+$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
 \item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.

### 2025-05-21 main/formal-structures.tex
Move biproduct section after LatGal.

@@
 
+\subsection{Biproducts and Semi-Additive Categories}
+\label{sec:biproducts}
+
+We noted that $\LatGal$ is enriched in $\SemiLat$, the category of semilattices, and moreover that the product
+of lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact the presence
+of biproducts is sufficient for enrichment in $\CMon$, the category of commutative monoids (including
+semilattices); such a category is called \emph{semi-additive}.
+
+\begin{definition}[Semi-additive category]
+A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
+and injections) give an isomorphism
+\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
+\end{definition}
+
+The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
+projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
+initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
+two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
+forming a morphism $f + g: X \to Y$. Diagrammatically:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
+\end{tikzcd}
+\end{center}
+
+Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
+$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
+The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
+\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
+composing the unique maps in and out of the zero object:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
+\end{tikzcd}
+\end{center}
+
+It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
+hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
+family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
+preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+
+\begin{salign*}
+f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
+\end{salign*}
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}
+
+\subsubsection{Biproduct laws}
+It is also easy to show that the following equations hold:
+
+\vspace{-3mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_X \comp \biinj_X &= \id_X \\
+   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
+\end{salign*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_Y \comp \biinj_Y &= \id_Y \\
+   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
+\end{salign*}
+\end{center}
+\end{minipage}
+
+\noindent and
+
+\begin{salign*}
+(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
+\end{salign*}
+
+We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
+\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
+providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
+biproducts.
+
+\begin{proposition}
+\label{prop:biproducts:from-product-or-coproduct}
+Suppose $\cat{C}$ is $\CMon$-enriched.
+\begin{enumerate}
+\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
+is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
+\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
+\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
+\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
+$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
+\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
+\end{enumerate}
+\end{proposition}
+
 \subsection{A categorical framework for Galois slicing}
@@
 function $f$.
-
-\subsection{Biproducts and Semi-Additive Categories}
-\label{sec:biproducts}
-
-We noted that $\LatGal$ is enriched in $\CMon$, the category of commutative monoids (in fact in $\SemiLat$)
-and moreover that the product of lattices is also a coproduct, making $\LatGal$ into a category with
-\emph{biproducts}. In fact the presence of biproducts is sufficient for $\CMon$-enrichment; such a category is
-called \emph{semi-additive}.
-
-\begin{definition}[Semi-additive category]
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
-projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
-two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-\subsubsection{Biproduct laws}
-It is also easy to show that the following equations hold:
-
-\vspace{-3mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
-\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
-providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
-biproducts.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-Suppose $\cat{C}$ is $\CMon$-enriched.
-\begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
-is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
-\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
-\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
-$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
-\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
-\end{enumerate}
-\end{proposition}

### 2025-05-21 main/formal-structures.tex
Tweak.

@@
 
-\noindent The trivial 1-point lattice, which is both terminal and initial, is the unit for $\biprod$. Although
-the presence of biproducts is incompatible with $\LatGal$ itself being Cartesian closed, the biproduct
-structure will prove crucial for establishing a Cartesian closed setting for interpreting higher-order
-programs for Galois slicing.
+\noindent The trivial 1-point lattice, which is both terminal and initial and we will write as $0$, is the
+unit for $\biprod$. Such a structure $(\biprod, 0)$ is called a \emph{biproduct} (\secref{biproducts}).
+Although $\LatGal$ is not itself Cartesian closed (in particular the set of Galois connections $X \to Y$ does
+not in general form a bounded lattice), the biproduct structure will prove crucial for establishing a
+Cartesian closed setting for interpreting higher-order programs.
 
@@
 \begin{definition}[Semi-additive category]
+\label{def:biproducts:semi-additive}
 A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
@@
 \subsubsection{Biproduct laws}
-It is also easy to show that the following equations hold:
+It is also easy to show that in a semi-additive category, as defined in \defref{biproducts:semi-additive}, the
+following equations hold:
 
@@
 pairs $(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial
-f(x): \partial X(x) \to \partial Y(f(x))$, as required. The pullback (reindexing) along $f$ in the composition
-$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the
-appropriate lattice of approximations and Galois connection at each point, using the baseline (unapproximated)
-function $f$.
+f(x): \partial X(x) \to \partial Y(f(x))$, as we sketched informally at the beginning of this section. The
+pullback (reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
+\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
+connection at each point, using the baseline (unapproximated) function $f$.

### 2025-05-21 main/framework.tex
Bump 'categorical framework' to a new section.

--- 2025-05-21 main/formal-structures.tex
@@
-\section{Formal Structures for Generalised Automatic Differentiation}
-
-\note{Probably a better way: (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
-
-We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others \cite{nunes2023}.
-Here follows a rough overview of this approach and how we have applied it to Galois program slicing.
-\todo{Final version shouldn't be ``rough''!}
-
-We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}.
-Not only does this mean that the constructions and proofs have been checked, the construction is executable and can be run on small examples.
-We are also planning to use Agda's JavaScript backend to produce a demo.\todo{Do this?}
-
-The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
-An object of the Grothendieck construction $\int T$~, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
-The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
-The maps in the indexed category are the maps of tangents or approximations, being derivatives and Galois connections respectively.
-
-In the case of differentiable programs and automatic differentiation, a basic model can be constructed by taking $\cat{C}$ to be $\Set$ and $T(X) = X \to \FinVect$: indexed collections of finite dimensional vector spaces, whose elements are interpreted as tangents at the given point, with linear maps.
-The Grothendieck category $\int T$ has objects that are pairs of a set $X$ and for each $x \in X$ a tangent vector space $\partial X(x)$.
-Morphisms are pairs of maps of points to points, accompanied by linear maps of tangents at those points.
-If we carefully choose an initial set of maps, we can read these maps as derivatives (note that nothing in the Grothendieck construction says that they must be derivatives!).
-Composition in the Grothendieck category is exactly the {\em chain rule} for composing derivatives, so we at least do know that if we start with derivatives on basic operations, then composing them will retain this property.
-
-For Galois program slicing, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.
-
-In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$~(\secref{Fam}).
-As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
-In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$).
-When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
-If the coproducts are actually {\em biproducts}~(\secref{biproducts}), then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
-
-Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of their having products and being enriched in commutative monoids~(\propref{biproducts:from-product-or-coproduct} below).
-However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
-Neither of these are ideal.
-In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost.
-In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model.
-Moreover, in Agda, there are issues with predicativity when considering complete lattices.\todo{be more specific}
-
-Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using $\CMon$-enriched presheaves.\todo{elaborate.}
-
-\subsection{Category of lattices and Galois connections}
-
-First we introduce the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
-X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
-of $g$, and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the
-bidirectional nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the
-least input slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the
-greatest output slice for a given slice of the input).
-
-\begin{definition}[Galois connection]
-Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
-$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
-\end{definition}
-
-\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
-Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
-$f$ and $g$ interpreted as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
-the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
-as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
-connections between them as morphisms.
-
-As sketched in \secref{introduction:galois-slicing}, for Galois slicing we want a setting where the
-approximants of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$
-representing the approximation that discards all information about $x$, greatest element $\top$ representing
-the approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways
-to combine approximations. Thus rather than working directly in $\PosGal$, we consider the following
-subcategory instead.
-
-\begin{definition}
-Define $\LatGal$ to be the category which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded
-lattices, and as morphisms all Galois connections $f \adj g: X \to Y$.
-\end{definition}
-
-\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $g$ is a (bounded) \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice
-structure $(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join,
-\bot)$.
-
-The meet-semilattice homomorphisms from $X$ to $Y$ form a meet-semilattice themselves, with $\meet$ on
-homomorphisms defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and the constant map
-$\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$ providing the unit (and both preserving meets). Dually
-the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by $\join$ and the
-constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two constructions
-come in adjoint pairs, $\LatGal$ is enriched in the category $\SemiLat$ of semilattices and semilattice
-homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit $\top_{X,Y}$ given by the
-Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois connections $(f \adj g) \meet (f'
-\adj g') = (f \join f') \adj (g \meet g'): X \to Y$.
-
-Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
-\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This means $X
-\times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
-projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
-
-\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   \biproj_X = \prodM{\id_X}{\bot_Y} \adj \proj_1: X \biprod Y \to X \\
-   \biproj_Y = \prodM{\bot_X}{\id_Y} \adj \proj_2: X \biprod Y \to Y
-\end{align*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   \biinj_X = \pi_1 \adj \prodM{\id_X}{\top_Y} : X \to X \biprod Y \\
-   \biinj_Y = \pi_2 \adj \prodM{\top_X}{\id_Y}: Y \to X \biprod Y
-\end{align*}
-\end{center}
-\end{minipage}
-\vspace{2mm}
-
-\noindent The trivial 1-point lattice, which is both terminal and initial and we will write as $0$, is the
-unit for $\biprod$. Such a structure $(\biprod, 0)$ is called a \emph{biproduct} (\secref{biproducts}).
-Although $\LatGal$ is not itself Cartesian closed (in particular the set of Galois connections $X \to Y$ does
-not in general form a bounded lattice), the biproduct structure will prove crucial for establishing a
-Cartesian closed setting for interpreting higher-order programs.
-
-\subsection{Biproducts and Semi-Additive Categories}
-\label{sec:biproducts}
-
-We noted that $\LatGal$ is enriched in $\SemiLat$, the category of semilattices, and moreover that the product
-of lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact the presence
-of biproducts is sufficient for enrichment in $\CMon$, the category of commutative monoids (including
-semilattices); such a category is called \emph{semi-additive}.
-
-\begin{definition}[Semi-additive category]
-\label{def:biproducts:semi-additive}
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
-projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
-two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-\subsubsection{Biproduct laws}
-It is also easy to show that in a semi-additive category, as defined in \defref{biproducts:semi-additive}, the
-following equations hold:
-
-\vspace{-3mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
-\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
-providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
-biproducts.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-Suppose $\cat{C}$ is $\CMon$-enriched.
-\begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
-is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
-\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
-\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
-$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
-\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
-\end{enumerate}
-\end{proposition}
-
-\subsection{A categorical framework for Galois slicing}
+\section{A categorical framework for Galois slicing}
 
@@
 
-\subsubsection{Grothendieck construction for indexed categories}
+\subsection{Grothendieck construction for indexed categories}
 \label{sec:Grothendieck}
@@
 
-\subsubsection{Category of families}
+\subsection{Category of families}
 \label{sec:Fam}
@@
 
-\subsubsection{Galois slicing via the Grothendieck construction}
+\subsection{Galois slicing via the Grothendieck construction}
 

### 2025-05-21 main/framework.tex
Improvements to 3.1.

@@
 This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
-= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
-the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
-between them as morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) is
-the category of all $\Set$-indexed families of bounded lattices, together with morphisms that account for both
-structure within each family and how families transform under reindexing.
+category $\Set^\op \to \Cat$, namely the indexed category $\Func{-}{\LatGal}$ which assigns to every set $X$
+the functor category $\Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so
+$\Func{X}{\LatGal}$ is exactly the category of $X$-indexed families of bounded lattices, with $X$-indexed
+families of Galois connections between them as morphisms. The Grothendieck construction
+$\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$, sometimes called the \emph{total
+category} for $T$, is the category of all $\Set$-indexed families of objects of $T$, together with morphisms
+that account for both structure within each family and how families transform under reindexing.
 

### 2025-05-21 main/formal-structures.tex

@@
 \end{proposition}
-
-\subsection{A categorical framework for Galois slicing}
-
-We will consider a pure, total language, where types can be interpreted as sets. Our goal is to equip the
-interpretations of programs with Galois connections relating approximations of their inputs to approximations
-of their outputs. Since these lattices of approximations depends on the value being approximated, the
-semantics of every type $\tau$ must have both a ``conventional'' component, providing the set $X$ associated
-with $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The
-semantics of an expression $e$ of type $Y$ with a single free variable of type $X$ must then also have two
-components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of the expression, plus a
-family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can
-be used for slicing over that expression.
-
-\subsubsection{Grothendieck construction for indexed categories}
-\label{sec:Grothendieck}
-
-This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
-= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
-the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
-between them as morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) is
-the category of all $\Set$-indexed families of bounded lattices, together with morphisms that account for both
-structure within each family and how families transform under reindexing.
-
-First we formalise $X$-indexed families and their categories.
-
-\begin{definition}[Category of $X$-indexed families]
-For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
-category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
-objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
-$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
-\end{definition}
-
-Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits from its codomain.
-
-\begin{proposition}
-If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
-pointwise.
-\end{proposition}
-
-\begin{definition}[Reindexing]
-For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
-\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
-$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
-family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
-$\reindex{\eta}{f}(x) = \eta(f(x))$.
-\end{definition}
-
-Although reindexing is just precomposition with $f$, writing it as $\reindex{-}{f}$ avoids notational
-confusion when combining composition of morphisms in $\Func{X}{C}$ with reindexing.
-
-\begin{definition}
-\label{def:Grothendieck}
-Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
-category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
-\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
-\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
-morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
-\end{definition}
-
-\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
-construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
-compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
-\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
-
-\begin{itemize}
-\item $\partial f: \partial X \to T(f)(\partial Y)$, and
-\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$
-\end{itemize}
-
-\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
-composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
-\comp f, T(f)(\partial g) \comp \partial f)$.
-
-\subsubsection{Category of families}
-\label{sec:Fam}
-
-Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
-category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
-$f: X \to Y$ to the reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .
-
-\begin{definition}[Category of families]
-\label{def:Fam}
-For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
-$\Grothendieck{X}\Func{X}{\cat{C}}$.
-\end{definition}
-
-\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
-$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
-$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
-$\Func{X}{\cat{C}}$.
-
-\subsubsection{Galois slicing via the Grothendieck construction}
-
-Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
-programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
-$x \in X$, a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all
-pairs $(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial
-f(x): \partial X(x) \to \partial Y(f(x))$, as we sketched informally at the beginning of this section. The
-pullback (reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
-\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
-connection at each point, using the baseline (unapproximated) function $f$.

### 2025-05-21 notes/biproduct.tex

@@
 Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $X, Y$, the \emph{biproduct} of $X$ and $Y$ is
-an object $X \biprod Y$ of $C$ together with morphisms
+an object $X \biprod Y$ of $\cat{C}$ together with morphisms
 
@@
 \end{definition}
-
-\noindent \todo{I think we can also define ``biproduct'' just using zero morphisms, and then derive rather
-than assume $\CMon$-enrichment.}
 

### 2025-05-21 notes/stability.tex
Some notes on stable functions

\subsection{Stable Functions}
\label{sec:stable-functions}

\begin{definition}
  Let $f : X \to Y$ be a monotone function between posets $X$ and
  $Y$. The function $f$ is \emph{stable} if for all $x \in X$ and
  $y \leq f(x)$, there exists an $x_0 \leq x$ such that $y \leq f(x_0)$ which
  is minimal: for any other $x'_0 \leq x$ such that $y \leq f(x'_0)$ then
  $x_0 \leq x'_0$.
\end{definition}

The intuition behind stability is that for stable functions, we have a
way of determining what minimal approximation of the input is required
for any approximation of the output. The ordering on $X$ and $Y$
describes an ordering of information content between
elements. Monotonicity means that increases in information content on
the input are not lost on the output. Stability means that, if we use
$f$ to derive a result $f(x)$ from $x$, and then we identify some
approximation $y$ of $f(x)$, then there is a unique minimal
approximation of $x$ that gets us as close as possible to $y$.

We can rephrase stability in terms of Galois connections / adjunctions
between principal downsets.

\newcommand{\downset}[1]{\mathop{\downarrow}(#1)}

\begin{definition}
  For a poset $X$ and element $x \in X$, the \emph{principal
    downset of $x$} is $\downset{x} = \{ x' \in X \mid x' \leq x \}$.
\end{definition}

\begin{lemma}
  If $f : X \to Y$ is monotonic, then for all $x \in X$, the
  restriction of $f$ to $\downset{x}$ is a monotone function
  $f_x : \downset{x} \to \downset{f(x)}$.
\end{lemma}

\begin{proof}
  $f_x$ is well defined: for any $x' \leq x$,
  $f_x(x') = f(x') \leq f(x)$ by monotonicity, so
  $f_x(x') \in \downset{f(x)}$. The function $f_x$ is monotone because
  $f$ is.
\end{proof}

\begin{lemma}
  A monotone function $f : X \to Y$ is stable if and only if for all
  $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$
  has a left adjoint.
\end{lemma}

\begin{proof}
  If $f$ is stable, then define a left adjoint
  $g_x : \downset{f(x)} \to \downset{x}$ by setting $g_x(y)$ to be the
  minimal $x_0$ required by stability. (FIXME: this is monotone?). Let
  $x' \leq x$ and $y \leq f(x)$. Then if $g_x(y) \leq x'$, we have
  $y \leq f(g_x(y)) \leq f(x')$ by monotonicity of $f$ and the first
  part of stability. In the other direction, if we have
  $y \leq f(x')$, then by uniqueness we have $g_x(y) \leq x'$.

  TODO: left adjoint implies stability.
\end{proof}

\begin{enumerate}
\item Posets where the principal downsets are meet-semilattices. And
  posets where the principal downsets are lattices.
\item Stable functions on these.
\item Full and faithful functor into Fam(LatGal), copying the idea
  from automatic differentiation: we formally delineate points and
  their tangent bundles.
\item
  \begin{enumerate}
  \item Posets that are meet semilattices in their downsets, with
    monotone functions form a Tangent Category
  \item To get ``reverse derivatives'', we further assume stability.
  \end{enumerate}
\end{enumerate}

### 2025-05-22 notes/stability.tex
Stability <-> left adjoint on slices

@@
 \label{sec:stable-functions}
+
+Stable functions arise in domain theory as an attempt to capture the
+idea that any function describing a computational process must explore
+its input in a way that allows identification of minimal parts of the
+input that are needed for chosen approximations of the input. They
+were invented by Berry \cite{berry86} in an attempt to capture
+sequentiality, and were also instrumental in the discovery of Linear
+Logic via Coherence Spaces \cite{girard}. A textbook description of
+stable functions in the context of domain theory is given by Amadio
+and Curien \cite{amadio-curien}. Here, we are not concerned with
+completeness properties considered in domain theory. Stable functions
+can be defined between any two partially ordered sets:
 
@@
 \begin{definition}
-  For a poset $X$ and element $x \in X$, the \emph{principal
+  For a poset $X$ and $x \in X$, the \emph{principal
     downset of $x$} is $\downset{x} = \{ x' \in X \mid x' \leq x \}$.
@@
   $g_x : \downset{f(x)} \to \downset{x}$ by setting $g_x(y)$ to be the
-  minimal $x_0$ required by stability. (FIXME: this is monotone?). Let
-  $x' \leq x$ and $y \leq f(x)$. Then if $g_x(y) \leq x'$, we have
+  minimal $x_0$ required by stability. This is monotone: if
+  $y \leq y'$, then we know that $y \leq y' \leq f(g_x(y'))$ by
+  defintition of $g_x$, so $g_x(y) \leq g_x(y')$ by minimality of
+  $g_x(y)$. For the adjointness, let $x' \leq x$ and $y \leq
+  f(x)$. Then if $g_x(y) \leq x'$, we have
   $y \leq f(g_x(y)) \leq f(x')$ by monotonicity of $f$ and the first
@@
 
-  TODO: left adjoint implies stability.
+  If, for every $x$, $f_x$ has a left adjoint $g_x$, then for any
+  $x', y$ we have $y \leq f_x(x') \Leftrightarrow g_x(y) \leq x'$. So
+  $g_x(y)$ is the element that satisfies $y \leq f(g_x(y))$, and it is
+  minimal since if $y \leq f_x(x'_0)$ then $g_x(y) \leq x'_0$.
 \end{proof}

### 2025-05-23 notes/stability.tex
Hopefully illuminating examples of stable and non-stable functions

--- 2025-05-21 notes/stability.tex
@@
 \label{sec:stable-functions}
+
+Stable functions are an antecedent of Galois slicing that arose in
+domain theory as an attempt to capture the idea that any function
+describing a computational process must explore its input in a way
+that allows identification of minimal parts of the input that are
+needed for chosen approximations of the output.
+
+Stability was invented by Berry \cite{berry86} in an attempt to capture
+sequentiality, and were also instrumental in the discovery of Linear
+Logic via Coherence Spaces \cite{girard}. A textbook description of
+stable functions in the context of domain theory is given by Amadio
+and Curien \cite{amadio-curien} (Chapter 12). Here, we are not
+concerned with completeness properties considered in domain theory, so
+we start with only partially ordered sets. Stable functions can be
+defined as a certain subset of the set of monotone functions between
+two partially ordered sets:
 
@@
   $Y$. The function $f$ is \emph{stable} if for all $x \in X$ and
-  $y \leq f(x)$, there exists an $x_0 \leq x$ such that $y \leq f(x_0)$ which
-  is minimal: for any other $x'_0 \leq x$ such that $y \leq f(x'_0)$ then
-  $x_0 \leq x'_0$.
+  $y \leq f(x)$:
+  \begin{enumerate}
+  \item (\textsc{Existence}) there exists an $x_0 \leq x$ such that $y \leq f(x_0)$, and
+  \item (\textsc{Minimality}) for any $x'_0 \leq x$ such that $y \leq f(x'_0)$ then
+    $x_0 \leq x'_0$ .
+  \end{enumerate}
 \end{definition}
 
-The intuition behind stability is that for stable functions, we have a
-way of determining what minimal approximation of the input is required
-for any approximation of the output. The ordering on $X$ and $Y$
-describes an ordering of information content between
-elements. Monotonicity means that increases in information content on
-the input are not lost on the output. Stability means that, if we use
-$f$ to derive a result $f(x)$ from $x$, and then we identify some
-approximation $y$ of $f(x)$, then there is a unique minimal
-approximation of $x$ that gets us as close as possible to $y$.
+\begin{lemma}
+  Stable functions are closed under identities and composition.
+\end{lemma}
 
-We can rephrase stability in terms of Galois connections / adjunctions
-between principal downsets.
+\begin{example}[Stability reveals intensional information]
+  To see how stability works, consider the following examples of the
+  OR operation on the lifted booleans $\mathbb{B}_\bot$. Two functions
+  that are stable are the strict and left-short-circuiting
+  ORs\footnote{These clauses are shorthand for the graph of the
+    function. They are not to be understood as pattern matching
+    clauses in a language like Haskell.}:
+  \begin{displaymath}
+    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
+      \mathrm{strictOr}&\mathsf{tt}&\mathsf{tt}&=&\mathsf{tt} \\
+      \mathrm{strictOr}&\mathsf{tt}&\mathsf{ff}&=&\mathsf{tt} \\
+      \mathrm{strictOr}&\mathsf{ff}&\mathsf{tt}&=&\mathsf{tt} \\
+      \mathrm{strictOr}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
+      \mathrm{strictOr}&\bot&\_&=&\bot \\
+      \mathrm{strictOr}&\_&\bot&=&\bot \\
+    \end{array}
+    \qquad
+    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
+      \mathrm{shortCircuitOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
+      \mathsf{shortCircuitOR}&\mathsf{ff}&x&=&x \\
+      \mathsf{shortCircuitOR}&\bot&\_& =& \bot
+    \end{array}
+  \end{displaymath}
+  The function $\mathrm{strictOr}$ is stable. For example, for the
+  input-output pair $(\mathsf{tt},\mathsf{ff}), \mathsf{tt}$, the
+  minimal input that gives this output is exactly
+  $(\mathsf{tt}, \mathsf{ff})$. If we take the approximation
+  $\bot \leq \mathsf{tt}$ of the output, then the corresponding
+  minimal input is $(\bot, \bot)$.
+
+  The function $\mathrm{shortCircuitOR}$ is also stable. For the
+  input-output pair $(\mathsf{tt},\mathsf{ff}),\mathsf{tt}$, the
+  minimal input that gives this input is $(\mathsf{tt},\bot)$,
+  indicating that the presence of $\mathsf{ff}$ in the second argument
+  was not necessary to produce this output. As with
+  $\mathrm{strictOr}$, the minimal input required to produce the
+  output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.
+
+  The fact that these two functions's stability witnesses reveal
+  information about how they depend on their input is what we will
+  exploit in order to use the idea of stability for slicing.
+\end{example}
+
+\begin{example}[Functions that are not stable]
+  A function that is not stable is Plotkin's Parallel OR \cite{lcf77},
+  which short-circuits in both arguments, returning $\mathsf{tt}$ if
+  either argment is $\mathsf{tt}$, even if the other argument is not
+  defined:
+  \begin{displaymath}
+    \begin{array}{l@{(}l@{,~}l@{)~}c@{~}l}
+      \mathrm{parallelOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
+      \mathrm{parallelOR}&\_&\mathsf{tt}&=&\mathsf{tt} \\
+      \mathsf{parallelOR}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
+      \mathsf{parallelOR}&\bot&\bot&=&\bot
+    \end{array}
+  \end{displaymath}
+  For the input-output pair $(\mathsf{tt},\mathsf{tt}),\mathsf{tt}$,
+  there is no one minimal input that produces this output. We have
+  both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and
+  $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are
+  incomparable and their greatest lower bound $(\bot,\bot)$ gives the
+  output $\bot$.
+
+  Parallel OR is famous because it is not \emph{sequential}, meaning
+  intuitively that it cannot be implemented without running the two
+  arguments in parallel to see if one of them returns
+  $\mathsf{tt}$. That fact that it exists in the standard domain
+  theoretic semantics of PCF means that this semantics is not fully
+  abstract. Since Parallel OR is not stable, one might hope that
+  stability is enough to capture sequentiality, and hence potentially
+  give a fully abstract model of PCF. However, the following ternary
+  function $\mathbb{B}_\bot^3 \to \{\top,\bot\}$ is stable but admits
+  no sequential implementation that fixes an order that the arguments
+  are examined in:
+  \begin{displaymath}
+    \begin{array}{l@{(}l@{,~}l@{,~}l@{)~}c@{~}l}
+      \mathrm{gustave}&\mathsf{tt}&\mathsf{ff}&\_&=&\top \\
+      \mathrm{gustave}&\mathsf{ff}&\_&\mathsf{tt}&=&\top \\
+      \mathrm{gustave}&\_&\mathsf{tt}&\mathsf{ff}&=&\top \\
+      \mathrm{gustave}&\_&\_&\_&=&\bot
+    \end{array}
+  \end{displaymath}
+  Despite there being no one minimal input that achieves the output
+  $\top$, each of the minimal inputs that can achieve this output are
+  pairwise incomparable, so for each specific input that gets output
+  $\top$ there is a unique minimal input that achieves it (listed in
+  the first three lines of the definition).
+
+  In terms of program slicing, however, the $\mathrm{gustave}$
+  function does not present a problem. For any particular run of the
+  program, there is an unambiguous minimal input that achieves the
+  output, no matter that it was not achieved by a sequential
+  processing of the input.
+\end{example}
+
+The definition of stability has a nice operational meaning, but seems
+a little ad-hoc. We can rephrase stability in terms of Galois
+connections / adjunctions between principal downsets / slices of the
+partial orders.
 
@@
   $g_x : \downset{f(x)} \to \downset{x}$ by setting $g_x(y)$ to be the
-  minimal $x_0$ required by stability. (FIXME: this is monotone?). Let
-  $x' \leq x$ and $y \leq f(x)$. Then if $g_x(y) \leq x'$, we have
+  minimal $x_0$ required by stability. This is monotone: if
+  $y \leq y'$, then we know that $y \leq y' \leq f(g_x(y'))$ by
+  definition of $g_x$, so $g_x(y) \leq g_x(y')$ by minimality of
+  $g_x(y)$. For the adjointness, let $x' \leq x$ and $y \leq
+  f(x)$. Then if $g_x(y) \leq x'$, we have
   $y \leq f(g_x(y)) \leq f(x')$ by monotonicity of $f$ and the first

### 2025-05-24 notes/stability.tex

@@
   that are stable are the strict and left-short-circuiting
-  ORs\footnote{These clauses are shorthand for the graph of the
-    function. They are not to be understood as pattern matching
-    clauses in a language like Haskell.}:
+  ORs\footnote{The clauses in these examples are shorthand for the
+    graph of the function. They are not to be understood as pattern
+    matching clauses in a language like Haskell, where it is not
+    possible to match on $\bot$.}:
   \begin{displaymath}
@@
   $\bot \leq \mathsf{tt}$ of the output, then the corresponding
-  minimal input is $(\bot, \bot)$.
-
-  The function $\mathrm{shortCircuitOR}$ is also stable. For the
-  input-output pair $(\mathsf{tt},\mathsf{ff}),\mathsf{tt}$, the
-  minimal input that gives this input is $(\mathsf{tt},\bot)$,
-  indicating that the presence of $\mathsf{ff}$ in the second argument
-  was not necessary to produce this output. As with
-  $\mathrm{strictOr}$, the minimal input required to produce the
-  output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.
+  minimal input is $(\bot, \bot)$. The function
+  $\mathrm{shortCircuitOR}$ is also stable. For the input-output pair
+  $(\mathsf{tt},\mathsf{ff}),\mathsf{tt}$, the minimal input that
+  gives this input is $(\mathsf{tt},\bot)$, indicating that the
+  presence of $\mathsf{ff}$ in the second argument was not necessary
+  to produce this output. As with $\mathrm{strictOr}$, the minimal
+  input required to produce the output $\bot \leq \mathsf{tt}$ is
+  again $(\bot,\bot)$.
 
@@
   In terms of program slicing, however, the $\mathrm{gustave}$
-  function does not present a problem. For any particular run of the
-  program, there is an unambiguous minimal input that achieves the
-  output, no matter that it was not achieved by a sequential
-  processing of the input.
+  function does not present a problem. For any particular run (i.e.,
+  input-output pair) of the program, there is an unambiguous minimal
+  input that achieves the output, no matter that it was not achieved
+  by a sequential processing of the input.
 \end{example}
@@
 connections / adjunctions between principal downsets / slices of the
-partial orders.
+partial orders. The alternative presentation also has the advantage
+that it explicitly defines the stability witness as ``backwards''
+function from approximations of the output to approximations of the
+input.
 
@@
 
-\begin{enumerate}
-\item Posets where the principal downsets are meet-semilattices. And
-  posets where the principal downsets are lattices.
-\item Stable functions on these.
-\item Full and faithful functor into Fam(LatGal), copying the idea
-  from automatic differentiation: we formally delineate points and
-  their tangent bundles.
-\item
-  \begin{enumerate}
-  \item Posets that are meet semilattices in their downsets, with
-    monotone functions form a Tangent Category
-  \item To get ``reverse derivatives'', we further assume stability.
-  \end{enumerate}
-\end{enumerate}
+For what follows\footnote{Need a better way of motivating this.}, it
+will be useful for the principal downsets to be bounded lattices, so
+that we are guaranteed to have meets and joints of approximations of
+points. Note that the principal downsets always have a top element
+(the original element), but now we are also requiring a bottom
+element, as well as finite meets and joins.
+
+FIXME: is this a good name? Maybe just $\mathbf{Stable}$ or something
+like that?
+
+FIXME: if $x \leq x'$ then $\downset{x} \subseteq \downset{x'}$ and I
+assume that the inclusion preserves lattice structure?
+
+\begin{definition}
+  An \emph{approximation set} is a partial order such that every
+  principal downset $\downset{x}$ is a bounded lattice.
+
+  A morphism of approxmation sets is monotone function whose
+  restriction to each slice has a left adjoint.
+\end{definition}
+
+Note that, by standard properties of Galois connections/adjoints, the
+pairs $g_x \vdash f_x$ preserve the finite joins and meets respectively.
+
+\begin{proposition}
+  Approximation sets and their morphisms form a category.
+\end{proposition}
+
+FIXME: what properties does this category have? I think it has finite
+products, and probably finite coproducts?
+
+FIXME: there is a full and faithful functor from approximation sets to
+$\Fam(\LatGal)$.

### 2025-05-24 notes/stability.tex
Cite Berry PhD thesis.

@@
 
-Stability was invented by Berry \cite{berry86} in an attempt to capture
-sequentiality, and were also instrumental in the discovery of Linear
+Stability was invented by \citet{berry79} in an attempt to capture
+sequentiality, and was also instrumental in the discovery of Linear
 Logic via Coherence Spaces \cite{girard}. A textbook description of

### 2025-05-24 notes/stability.tex
Plotkin on LCF.

@@
 Logic via Coherence Spaces \cite{girard}. A textbook description of
-stable functions in the context of domain theory is given by Amadio
-and Curien \cite{amadio-curien} (Chapter 12). Here, we are not
+stable functions in the context of domain theory is given by
+\citet{amadio-curien} (Chapter 12). Here, we are not
 concerned with completeness properties considered in domain theory, so
@@
 \begin{example}[Functions that are not stable]
-  A function that is not stable is Plotkin's Parallel OR \cite{lcf77},
+  A function that is not stable is Plotkin's Parallel OR~\cite{lcf77},
   which short-circuits in both arguments, returning $\mathsf{tt}$ if
@@
 
-  A morphism of approxmation sets is monotone function whose
+  A morphism of approximation sets is monotone function whose
   restriction to each slice has a left adjoint.

### 2025-05-24 notes.tex
Sync notes summary.

@@
 
-\title{Approximation as Differentiation}
-
-\author{Robert Atkey}
-\email{robert.atkey@strath.ac.uk}
-\orcid{0000-0002-4414-5047}
-\affiliation{%
-  \institution{University of Strathclyde}
-  \city{Glasgow}
-  \country{UK}}
-
-\author{Roly Perera}
-\email{roly.perera@cl.cam.ac.uk}
-\orcid{0000-0001-9249-9862}
-\affiliation{%
-  \institution{University of Cambridge}
-  \city{Cambridge}
-  \country{UK}
-}
-\additionalaffiliation{%
-   \institution{University of Bristol}
-   \city{Bristol}
-   \country{UK}
-}
+\title{Approximation as Differentiation: Notes}
 
@@
 
-\section{Notes}
-
 \subsection{Overview}
 
-Covered so far:
+Covered here:
 \begin{itemize}
 \item automatic differentiation (\secref{auto-diff})
-\item $\CMon$-enriched categories, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
+\item stable functions (\secref{stable-functions})
+\item $\CMon$, the category of commutative monoids (\secref{cmon})
+\item $\CMon$-enrichment, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
 \item biproducts and semi-additive categories (\secref{biproduct})
@@
 \input{notes/auto-diff}
+\input{notes/stability}
 \input{notes/cmon}
@@
 \input{notes/auto-diff-galois-slicing-via-fam}
+
 \input{notes/language}

### 2025-05-24 main/framework.tex
Placeholder for Cartesian closure.
Correction to sec 3 intro.

@@
 of their outputs. Since these lattices of approximations depends on the value being approximated, the
-semantics of every type $\tau$ must have both a ``conventional'' component, providing the set $X$ associated
-with $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The
-semantics of an expression $e$ of type $Y$ with a single free variable of type $X$ must then also have two
-components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of the expression, plus a
-family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can
-be used for slicing over that expression.
+semantics of every type $\tau$ must have both a ``conventional'' component, providing the set $X$ which
+interprets $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The
+semantics of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with $\tau$ and
+$\sigma$ conventionally interpreted by $X$ and $Y$) must then also have two components: a function $f: X \to
+Y$ giving the conventional interpretation of $e$, and a family of Galois connections $\partial f(x): \partial
+X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for slicing over that expression.
 
@@
 connection at each point, using the baseline (unapproximated) function $f$.
+
+\subsection{Interpreting higher-order programs}
+
+\todo{conditions under which $\Fam(\cat{C})$ is Cartesian closed, why we need $\PSh_{\CMon}(\cat{C})$, etc}

### 2025-05-24 main/framework.tex
Tweak.

@@
 objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
-$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
+$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ has only identity morphisms.
 \end{definition}

### 2025-05-24 main/introduction.tex
Correction.
Minor corrections.

@@
 \subsection{Galois Program Slicing}
+\label{sec:introduction:galois-slicing}
 
@@
 However, other choices are also useful.
-Indeed, one of the aims of this work is to clarify how to choose approximation structure approriate for different tasks by use of type information.
+Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information.
 We elaborate on this further in Section \todo{XYZ}.
@@
   Galois program slicing associates with each run of the program a Galois connection telling us how the inputs and outputs are related in that run.
-The backwards portion $\partial (\mathit{query}\,l)_r$ tells us, given an approximation of the output what the least approximation of the input is needed to generate that output.
+The backwards portion $\partial (\mathit{query}\,l)_r$ tells us, given an approximation of the output, what the least approximation of the input is needed to generate that output.
 In the case of the two runs considered in this example, if we say we are not interested in the output by feeding in the least approximation $\bot$, then we find that we only need the least approximation of the input:
@@
   In a simple query like this, it is easy to work out the dependency relationship between the input and output.
-However, the benefit of Galois Program Slicing is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output is.
+However, the benefit of Galois Program Slicing is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
 Moreover, by changing what we mean by ``approximation'' we can compute a range of different information about a program.

### 2025-05-24 main/formal-structures.tex
Also cite vákár22.
Cancel \prod override.

@@
 
-We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others \cite{nunes2023}.
+We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others~\cite{vákár22,nunes2023}.
 Here follows a rough overview of this approach and how we have applied it to Galois program slicing.
@@
 and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables.
+\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
 \end{definition}

### 2025-05-24 main/introduction.tex
Tweaks questions in intro, placeholder for abstract.

@@
 
-Provenance tracking aims to answer questions like ``where did these data come from'', ``why are these data in the output'' and ``how were these data computed''.
+Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
 Program slicing is a collection of techniques for provenance tracking that attempts to take a run of a program and and areas of interest in the output and turn them into the subset of the input and the program that were responsible for generating those specific outputs.

### 2025-05-24 main/introduction.tex
Tweaks questions in intro, placeholder for abstract.

@@
 
-Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
+Perera and collaborators introduced the idea of {\em Galois Program Slicing} (hereafter, \GPS) as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
+\GPS forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
 
-At a high level, Galois Program Slicing assumes that, for each possible value that may be input or output by a program, there exists a lattice of {\em approximations} of that value.
+At a high level, \GPS assumes that, for each possible value that may be input or output by a program, there exists a lattice of {\em approximations} of that value.
 For a particular run of a program that takes input $x$ and produces output $y$, we also get a Galois connection between the lattice of approximations of $x$ and the lattice of approximations of $y$.

### 2025-05-24 main/introduction.tex
Minor reorg of intro.

--- 2025-05-24 main/introduction.tex
@@
 Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
-Program slicing is a collection of techniques for provenance tracking that attempts to take a run of a program and and areas of interest in the output and turn them into the subset of the input and the program that were responsible for generating those specific outputs.
+Provenance tracking has a wide range of applications, from debugging and program comprehension to reproducibility and transparency in scientific workflows.
+In order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome.
+\emph{Program slicing} is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.
 
-Provenance tracking is important because in order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome.
-Existing approaches are often tied to particular programming languages or implementations.
-In this paper we develop a general categorical approach.
-
-\note{We will use biproducts, and use some results from \cite{karvonen2020}.}
+Existing approaches to program slicing are often tied to particular programming languages or implementations. In this paper we develop a general categorical approach to program slicing and \todo{...}.
 
@@
 Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois program slicing forms the basis of Perera's data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
+Galois Program Slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
 
-At a high level, Galois Program Slicing assumes that, for each possible value that may be input or output by a program, there exists a lattice of {\em approximations} of that value.
+At a high level, \GPS assumes that, for each possible value that may be input or output by a program, there exists a lattice of {\em approximations} of that value.
 For a particular run of a program that takes input $x$ and produces output $y$, we also get a Galois connection between the lattice of approximations of $x$ and the lattice of approximations of $y$.

### 2025-05-24 main/formal-structures.tex
\GPS macro to standardise terminology.

@@
 We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others~\cite{vákár22,nunes2023}.
-Here follows a rough overview of this approach and how we have applied it to Galois program slicing.
-\todo{Final version shouldn't be ``rough''!}
+Here follows a rough overview of this approach and how we have applied it to \GPS.
 
@@
 
-For Galois program slicing, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.
+For \GPS, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.
 
@@
 First we introduce the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
+interpreting forward and backward \GPS. Galois connections are pairs of monotone functions $f: Y \to
 X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
@@
 
-As sketched in \secref{introduction:galois-slicing}, for Galois slicing we want a setting where the
-approximants of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$
-representing the approximation that discards all information about $x$, greatest element $\top$ representing
-the approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways
-to combine approximations. Thus rather than working directly in $\PosGal$, we consider the following
-subcategory instead.
+As sketched in \secref{introduction:galois-slicing}, for \GPS we would like a setting where the approximants
+of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing
+the approximation that discards all information about $x$, greatest element $\top$ representing the
+approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to
+combine approximations. Thus rather than working directly in $\PosGal$, we consider the following subcategory
+instead.
 

### 2025-05-24 main/framework.tex
Correction.
\GPS macro to standardise terminology.

@@
-\section{A categorical framework for Galois slicing}
+\section{A categorical framework for \GPS}
 
@@
 
-Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits from its codomain.
+Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.
 
@@
 
-\subsection{Galois slicing via the Grothendieck construction}
+\subsection{\GPS via the Grothendieck construction}
 
 Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
-programs for Galois slicing. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every
-$x \in X$, a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all
-pairs $(f, \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial
-f(x): \partial X(x) \to \partial Y(f(x))$, as we sketched informally at the beginning of this section. The
-pullback (reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
+programs for \GPS. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$,
+a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f,
+\partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial
+X(x) \to \partial Y(f(x))$, as we sketched informally at the beginning of this section. The pullback
+(reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
 \reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois

### 2025-05-24 main/introduction.tex
\GPS macro to standardise terminology.

@@
   Now suppose that for a given run of the program, we are interested in which of the numerical parts of the input are used to compute the output for the query parameters $l = \mathsf{a}$ and $l = \mathsf{b}$.
-We can use Galois Program Slicing to do this.
+We can use \GPS to do this.
 We arrange for the approximations of the input to form the following lattice, where the actual piece of data is at the top and information lost by approximation is represented by $\bot$s:
@@
 
-  Galois program slicing associates with each run of the program a Galois connection telling us how the inputs and outputs are related in that run.
+  \GPS associates with each run of the program a Galois connection telling us how the inputs and outputs are related in that run.
 The backwards portion $\partial (\mathit{query}\,l)_r$ tells us, given an approximation of the output, what the least approximation of the input is needed to generate that output.
@@
   In a simple query like this, it is easy to work out the dependency relationship between the input and output.
-However, the benefit of Galois Program Slicing is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
+However, the benefit of \GPS is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
 Moreover, by changing what we mean by ``approximation'' we can compute a range of different information about a program.
@@
 
-Previous work on Galois program slicing uses an operational semantics of programs that generates a trace of each execution which can be interpreted after the fact to compute the Galois connection described above.
+Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be interpreted after the fact to compute the Galois connection described above.
 This becomes {\em program slicing} by including the source code of the program as part of the input, so that, in the backward direction, the least approximation of the input required for an output approximation includes the least part of the program required.
@@
 
-\subsection{Galois Program Slicing and Automatic Differentiation}
+\subsection{Galois Slicing and Automatic Differentiation}
 
-\todo{explain why we want a more abstract view on Galois program slicing}
-There is a close analogy between Galois Program Slicing and Automatic Differentiation for differentiable programs.
-We have already hinted at this in the description above, but let us now make it explicit.
+\todo{explain why we want a more abstract view on \GPS}
+There is a close analogy between \GPS and automatic differentiation for differentiable programs. We have already hinted at this in the description above, but let us now make it explicit.
 
 \begin{itemize}
-\item For Galois Program Slicing, we assume that every value has an associated lattice of {\em approximations}.
-For differentiable programs, every point has an associated vector space of {\em tangents}.
-\item For Galois Program Slicing, every program has an associated forward approximation map that takes approximations forward from the input to the output.
+\item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
+\item For \GPS, every program has an associated forward approximation map that takes approximations forward from the input to the output.
 This map {\em preserves meets}.
@@
 The forward derivative map is {\em linear}, so it preserves addition of tangents and the zero tangent.
-\item For Galois Program Slicing, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input.
+\item For \GPS, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input.
 This map {\em preserves joins}.
@@
 This map is again {\em linear}.
-\item For Galois Program Slicing, the forward and backward approximation maps are related by being a Galois connection.
+\item For \GPS, the forward and backward approximation maps are related by being a Galois connection.
 For differentiable programming, the forward and reverse derivatives are related by being each others' transpose.
@@
 
-Given this close connection between Galois program slicing and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model Galois program slicing.
-This will enable us to generalise and expand the scope of Galois program slicing to act as a foundation for data provenance in a wide range of computational settings.
+Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.
+This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wide range of computational settings.
 
@@
 
-Our primary contribution in this article is to elucidate Galois Program Slicing by relating it to differentiable programming and automatic differentiation.
+Our primary contribution in this article is to elucidate \GPS by relating it to differentiable programming and automatic differentiation.
 In doing so, we aim to expand the applicability of program slicing, and to potentially transfer efficient automatic differentiation implementation techniques from automatic differentiation to program slicing and provenance tracking.
@@
 \begin{enumerate}
-\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for Galois Program Slicing,
-\item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of Galois Program Slicing.
-\item With the benefit of our abstract setting, we can relate Galois Program Slicing to other parts of denotational semantics.
+\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS,
+\item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.
+\item With the benefit of our abstract setting, we can relate \GPS to other parts of denotational semantics.
 In particular, we show that there is a close connection with Stable Domain Theory, a proposed solution to capturing sequentiality by recording more intensional information about programs' sensitivity to approximations.
-\item We also relate Galois Program Slicing to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
+\item We also relate \GPS to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
 \end{enumerate}

### 2025-05-24 notes/stability.tex
\GPS macro to standardise terminology.

@@
 
-Stable functions are an antecedent of Galois slicing that arose in
+Stable functions are an antecedent of \GPS that arose in
 domain theory as an attempt to capture the idea that any function

### 2025-05-24 main/introduction.tex
weiser81.

@@
 Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
-Provenance tracking has a wide range of applications, from debugging and program comprehension to reproducibility and transparency in scientific workflows.
+Provenance tracking has a wide range of applications, from debugging and program comprehension to improving reproducibility and transparency in scientific workflows.
 In order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome.
-\emph{Program slicing} is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.
+\emph{Program slicing}, first proposed by~\citet{weiser81}, is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.
 
@@
 
-Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois Program Slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
+Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
+Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
 

### 2025-05-24 main/introduction.tex
Move comment about prog slicing earlier.

@@
 The right half of the Galois connection is the ``forward direction'' taking approximations of the input to approximations of the output; the left half of the Galois connection is the ``backward direction'' that takes approximations of the output to the least (i.e., most approximate) approximation of the input that gives rise to this output approximation.
+This becomes {\em program slicing} by including the source code of the program as part of the input; then, in the backward direction, the least approximation of the input required for an output approximation includes the least part of the program required as well.
 
@@
 Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be interpreted after the fact to compute the Galois connection described above.
-This becomes {\em program slicing} by including the source code of the program as part of the input, so that, in the backward direction, the least approximation of the input required for an output approximation includes the least part of the program required.
-
-\todo{Explain what the plan for the paper is}
-
-\todo{Perhaps more elaboration on the applications of program slicing}
 

### 2025-05-24 main/introduction.tex
siskind08.

@@
 
-Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
+Data provenance aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
 Provenance tracking has a wide range of applications, from debugging and program comprehension to improving reproducibility and transparency in scientific workflows.
@@
 
-Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be interpreted after the fact to compute the Galois connection described above.
-
 \subsection{Galois Slicing and Automatic Differentiation}
 
-\todo{explain why we want a more abstract view on \GPS}
-There is a close analogy between \GPS and automatic differentiation for differentiable programs. We have already hinted at this in the description above, but let us now make it explicit.
+Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be rewound or replayed after the fact, to compute the Galois connection described above.
+It would be satisfying to have a denotational account of \GPS as well, for which traces would then offer a plausible implementation technique (perhaps one of several).
+As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08}, which points to a way to develop such an approach.
+We have already hinted at this in the description above, but let us now make it explicit.
 

### 2025-05-24 main/introduction.tex
siskind08.

@@
   In a simple query like this, it is easy to work out the dependency relationship between the input and output.
-However, the benefit of \GPS is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
+However, the benefit of \GPS, and other program analysis-based techniques, is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
 Moreover, by changing what we mean by ``approximation'' we can compute a range of different information about a program.

### 2025-05-24 main/formal-structures.tex
ACM requires title case for subsection headings.
Move comment about Agda impl to contribs.

@@
 
-\note{Probably a better way: (1) introduce the category of lattices and Galois connections, which is what we will use to interpret forward and backward approximation, and establish some properties of this category; (2) say that we want sets with each point having an associated lattice of approximations, and morphisms to match; (3) note that this is exactly the Grothendieck construction for a particular indexed category; (4) Say that CHAD is what we need here; (5) Recapitulate the bits of CHAD that we need, with our Agda formalisation; (6) Explain how we cope with higher-order structure, by using presheaves and the Yoneda embedding.}
-
-We take an approach based on the CHAD (Combinatory Homomorphic Automatic Differentiation) of Mattias Vákár and others~\cite{vákár22,nunes2023}.
-Here follows a rough overview of this approach and how we have applied it to \GPS.
-
-We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}.
-Not only does this mean that the constructions and proofs have been checked, the construction is executable and can be run on small examples.
-We are also planning to use Agda's JavaScript backend to produce a demo.\todo{Do this?}
+Our approach to a denotational account of \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
+Before we go into details, we present a high-level overview of this approach and how we have applied it to \GPS.
 
@@
 
-\subsection{Category of lattices and Galois connections}
+\subsection{Category of Lattices and Galois Connections}
 

### 2025-05-24 main/introduction.tex
Move comment about Agda impl to contribs.

--- 2025-05-24 main/introduction.tex
@@
   In a simple query like this, it is easy to work out the dependency relationship between the input and output.
-However, the benefit of \GPS is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
+However, the benefit of \GPS, and other language-based approaches, is that it is {\em automatic} for all programs, no matter how complex the relationship between input and output.
 Moreover, by changing what we mean by ``approximation'' we can compute a range of different information about a program.
@@
 \end{enumerate}
+
+We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}.
+Not only does this mean that the constructions and proofs have been checked, but also the construction is executable and can be run on small examples.
+We are also planning to use Agda's JavaScript backend to produce a demo.
+\todo{Do this?}

### 2025-05-24 main/formal-structures.tex
interpreted -> regarded.

@@
 Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
-$f$ and $g$ interpreted as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
+$f$ and $g$ regarded as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
 the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$

### 2025-05-24 main/framework.tex
ACM requires title case for subsection headings.
interpreted -> regarded.

@@
 
-\subsection{Grothendieck construction for indexed categories}
+\subsection{Grothendieck Construction for Indexed Categories}
 \label{sec:Grothendieck}
@@
 \begin{definition}[Category of $X$-indexed families]
-For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
+For any set $X$ (regarded as a discrete category) and any category $\cat{C}$, recall that the functor
 category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
@@
 
-\subsection{Category of families}
+\subsection{Category of Families}
 \label{sec:Fam}
@@
 
-\subsection{\GPS via the Grothendieck construction}
+\subsection{Galois Slicing via the Grothendieck Construction}
 
@@
 
-\subsection{Interpreting higher-order programs}
+\subsection{Interpreting Higher-Order Programs}
 

### 2025-05-24 notes/fam.tex
interpreted -> regarded.

--- 2025-05-06 appendix/notes/fam.tex
@@
 
-\noindent Equivalently $\Fam(I,\cat{C})$ is the functor category $\Func{I}{\cat{C}}$, interpreting $I$ as a
+\noindent Equivalently $\Fam(I,\cat{C})$ is the functor category $\Func{I}{\cat{C}}$, regarding $I$ as a
 discrete category.
@@
 
-By \emph{indexed category over $\cat{C}$} we shall mean any functor $\cat{C}^\op \to \Cat$.
+By \emph{(strictly) indexed category over $\cat{C}$} we shall mean any (strict) functor $\cat{C}^\op \to
+\Cat$.
 

### 2025-05-24 main/formal-structures.tex
Say something about CC on Fam(X).

@@
 In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$~(\secref{Fam}).
-As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$.
+These settings are adequate for interpreting first-order programs, but neither is Cartesian closed.
+To interpret higher-order programs in $\Fam(\cat{X})$ for any category $\cat{X}$, richer structure on $\cat{X}$ is required.
+As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ is the free coproduct completion of $\cat{X}$.
 In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$).

### 2025-05-24 main/formal-structures.tex
Say something about CC on Fam(X).
Start on todo's for new section.

@@
 These settings are adequate for interpreting first-order programs, but neither is Cartesian closed.
-To interpret higher-order programs in $\Fam(\cat{X})$ for any category $\cat{X}$, richer structure on $\cat{X}$ is required.
-As is well known~\cite{lawvere63}, $\Fam(\cat{X})$ is the free coproduct completion of $\cat{X}$.
+To interpret higher-order programs, the picture is more complicated.
+As is well known, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$~\cite{lawvere63}.
 In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$).
@@
 In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model.
-Moreover, in Agda, there are issues with predicativity when considering complete lattices.\todo{be more specific}
+Moreover, in Agda, there are issues with predicativity when considering complete lattices.
+\todo{be more specific}
 
-Vákár overcomes this by using different interpretations for forward and backward derivatives. we overcome this by using $\CMon$-enriched presheaves.\todo{elaborate.}
+Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
+We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{}, but first we flesh out the first-order setting.
 
@@
 Cartesian closed setting for interpreting higher-order programs.
-
-\subsection{Biproducts and Semi-Additive Categories}
-\label{sec:biproducts}
-
-We noted that $\LatGal$ is enriched in $\SemiLat$, the category of semilattices, and moreover that the product
-of lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact the presence
-of biproducts is sufficient for enrichment in $\CMon$, the category of commutative monoids (including
-semilattices); such a category is called \emph{semi-additive}.
-
-\begin{definition}[Semi-additive category]
-\label{def:biproducts:semi-additive}
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
-projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
-two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-\subsubsection{Biproduct laws}
-It is also easy to show that in a semi-additive category, as defined in \defref{biproducts:semi-additive}, the
-following equations hold:
-
-\vspace{-3mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\noindent and
-
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
-\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
-providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
-biproducts.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-Suppose $\cat{C}$ is $\CMon$-enriched.
-\begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
-is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
-\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
-\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
-$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
-\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
-\end{enumerate}
-\end{proposition}

### 2025-05-24 main/higher-order.tex
Start on todo's for new section.

\section{Interpreting Higher-Order Programs}

\todo{
\begin{enumerate}
\item conditions under which $\Fam(\cat{C})$ is Cartesian closed
\item $\PSh(\cat{C})$ has all small products (inherited from $\Set$)
\item why we need $\PSh_{\CMon}(\cat{C})$
\end{enumerate}
}

\subsection{Biproducts and Semi-Additive Categories}
\label{sec:biproducts}

We noted that $\LatGal$ is enriched in $\SemiLat$, the category of semilattices, and moreover that the product
of lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact the presence
of biproducts is sufficient for enrichment in $\CMon$, the category of commutative monoids (including
semilattices); such a category is called \emph{semi-additive}.

\begin{definition}[Semi-additive category]
\label{def:biproducts:semi-additive}
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and injections) give an isomorphism
\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
\end{definition}

The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
forming a morphism $f + g: X \to Y$. Diagrammatically:

\begin{center}
\begin{tikzcd}
   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
\end{tikzcd}
\end{center}

Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
composing the unique maps in and out of the zero object:

\begin{center}
\begin{tikzcd}
   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
\end{tikzcd}
\end{center}

It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:

\begin{salign*}
f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
\end{salign*}
\begin{salign*}
(f + g) \comp h &= (f \comp h) + (g \comp h) \\
h \comp (f + g) &= (h \comp f) + (h \comp g)
\end{salign*}

\subsubsection{Biproduct laws}
It is also easy to show that in a semi-additive category, as defined in \defref{biproducts:semi-additive}, the
following equations hold:

\vspace{-3mm}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_X \comp \biinj_X &= \id_X \\
   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
\end{salign*}
\end{center}
\end{minipage}%
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \biproj_Y \comp \biinj_Y &= \id_Y \\
   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
\end{salign*}
\end{center}
\end{minipage}

\noindent and

\begin{salign*}
(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
\end{salign*}

We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
biproducts.

\begin{proposition}
\label{prop:biproducts:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{enumerate}
\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
\end{enumerate}
\end{proposition}

### 2025-05-24 main/formal-structures.tex
Expand why LatGal isn't CC.

@@
 Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
-We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{}, but first we flesh out the first-order setting.
+We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{higher-order}, but first we flesh out the first-order setting.
 
@@
 
-The meet-semilattice homomorphisms from $X$ to $Y$ form a meet-semilattice themselves, with $\meet$ on
-homomorphisms defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and the constant map
-$\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$ providing the unit (and both preserving meets). Dually
-the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by $\join$ and the
-constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two constructions
-come in adjoint pairs, $\LatGal$ is enriched in the category $\SemiLat$ of semilattices and semilattice
-homomorphisms. The hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit $\top_{X,Y}$ given by the
-Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois connections $(f \adj g) \meet (f'
-\adj g') = (f \join f') \adj (g \meet g'): X \to Y$.
+The meet-semilattice homomorphisms from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$
+on homomorphisms is defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves
+meets; the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets,
+provides the unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given
+pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$.
+Since these two constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet
+semilattice with unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and
+meet of Galois connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$. Thus
+$\LatGal$ is enriched in $\SemiLat$, the category of semilattices and semilattice homomorphisms.
 
 Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
-\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This means $X
-\times Y$, which we shall hereafter write as $X \biprod Y$, acts as both a product and a coproduct, with
+\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This allows $X
+\times Y$, which we shall hereafter write as $X \biprod Y$, to act as both a product and a coproduct, with
 projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
@@
 \noindent The trivial 1-point lattice, which is both terminal and initial and we will write as $0$, is the
-unit for $\biprod$. Such a structure $(\biprod, 0)$ is called a \emph{biproduct} (\secref{biproducts}).
-Although $\LatGal$ is not itself Cartesian closed (in particular the set of Galois connections $X \to Y$ does
-not in general form a bounded lattice), the biproduct structure will prove crucial for establishing a
-Cartesian closed setting for interpreting higher-order programs.
+unit for $\biprod$. Such a structure $(\biprod, 0)$, where $\biprod$ acts as both a product and a coproduct,
+is called a \emph{biproduct} (\secref{biproducts}).
+
+$\LatGal$ is not Cartesian closed; the set of Galois connections $X \to Y$ does not in general form a bounded
+lattice, for a start. In fact if $\cat{C}$ has biproducts $(\biprod, 0)$ and is Cartesian closed, then every
+morphism in $\cat{C}$ is trivial. If $0$ is both initial and terminal, and therefore $0 \times X \iso X$, then
+$\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$ and so any morphism $f: X \to Y$ must be
+equal to the composite $X \to 0 \to Y$ of the unique morphisms in and out of $0$. The biproduct structure will
+prove crucial for establishing a Cartesian closed setting for interpreting higher-order programs.

### 2025-05-24 main/framework.tex
Start on todo's for new section.
Expand why LatGal isn't CC.

@@
-\section{A categorical framework for \GPS}
-
-We will consider a pure, total language, where types can be interpreted as sets. Our goal is to equip the
-interpretations of programs with Galois connections relating approximations of their inputs to approximations
-of their outputs. Since these lattices of approximations depends on the value being approximated, the
-semantics of every type $\tau$ must have both a ``conventional'' component, providing the set $X$ which
-interprets $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The
-semantics of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with $\tau$ and
-$\sigma$ conventionally interpreted by $X$ and $Y$) must then also have two components: a function $f: X \to
-Y$ giving the conventional interpretation of $e$, and a family of Galois connections $\partial f(x): \partial
-X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for slicing over that expression.
-
 \subsection{Grothendieck Construction for Indexed Categories}
 \label{sec:Grothendieck}
+
+What we want to do now is understand how to equip the interpretation of a (first-order) program with a Galois
+connections relating approximations of its inputs to approximations of its output. Since these lattices of
+approximations depend on the value being approximated, the semantics of every type $\tau$ must have both a
+``conventional'' component, providing the set $X$ which interprets $\tau$, plus an associated lattice of
+approximations $\partial X(x)$ for every point $x \in X$. The semantics of an expression $e$ of type $\sigma$
+with a single free variable of type $\tau$ (with $\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have
+two components: a function $f: X \to Y$ giving the conventional interpretation of $e$, and a family of Galois
+connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for
+slicing over that expression.
 
@@
 connection at each point, using the baseline (unapproximated) function $f$.
-
-\subsection{Interpreting Higher-Order Programs}
-
-\todo{conditions under which $\Fam(\cat{C})$ is Cartesian closed, why we need $\PSh_{\CMon}(\cat{C})$, etc}

### 2025-05-24 main/formal-structures.tex
End of 2.1/beginning of 2.2 currently a bit noddy..

@@
 Our approach to a denotational account of \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
-Before we go into details, we present a high-level overview of this approach and how we have applied it to \GPS.
+Before we go into details, we present a high-level overview of their approach and how we have applied it to \GPS.
 
 The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
-An object of the Grothendieck construction $\int T$~, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
+An object of the Grothendieck construction $\int T$, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
 The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
@@
 These settings are adequate for interpreting first-order programs, but neither is Cartesian closed.
-To interpret higher-order programs, the picture is more complicated.
+The picture for higher-order programs is more complicated.
 As is well known, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$~\cite{lawvere63}.
-In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$ (using the Cartesian products from $\Set$).
+In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$, using the Cartesian products from $\Set$.
 When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
@@
 Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
-We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{higher-order}, but first we flesh out the first-order setting.
+We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{higher-order} but first flesh out the first-order setting.
 
@@
 
-The meet-semilattice homomorphisms from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$
-on homomorphisms is defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves
-meets; the constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets,
-provides the unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given
-pointwise by $\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$.
-Since these two constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet
-semilattice with unit $\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and
-meet of Galois connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$. Thus
-$\LatGal$ is enriched in $\SemiLat$, the category of semilattices and semilattice homomorphisms.
+The meet-preserving maps from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$ on such
+maps is defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves meets; the
+constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets, provides the
+unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by
+$\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two
+constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit
+$\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
+connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$. Thus $\LatGal$ is
+enriched in $\SemiLat$, the category of semilattices and semilattice homomorphisms.
 
-Another notable property of $\LatGal$ is that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y
-\to Y$, where $X \times Y$ denotes the product of lattices, have both upper and lower adjoints. This allows $X
-\times Y$, which we shall hereafter write as $X \biprod Y$, to act as both a product and a coproduct, with
-projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
+$\LatGal$ is not Cartesian closed, and so we cannot use it directly to interpret higher-order programs. In
+fact Cartesian closure is incompatible with another notable property of $\LatGal$, which follows from the fact
+that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the
+product of lattices, have both upper and lower adjoints. This allows $X \times Y$, which we shall hereafter
+write as $X \biprod Y$, to act as both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$
+and injections $\biinj_X$ and $\biinj_Y$ given by:
 
@@
 
-\noindent The trivial 1-point lattice, which is both terminal and initial and we will write as $0$, is the
-unit for $\biprod$. Such a structure $(\biprod, 0)$, where $\biprod$ acts as both a product and a coproduct,
-is called a \emph{biproduct} (\secref{biproducts}).
-
-$\LatGal$ is not Cartesian closed; the set of Galois connections $X \to Y$ does not in general form a bounded
-lattice, for a start. In fact if $\cat{C}$ has biproducts $(\biprod, 0)$ and is Cartesian closed, then every
-morphism in $\cat{C}$ is trivial. If $0$ is both initial and terminal, and therefore $0 \times X \iso X$, then
-$\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$ and so any morphism $f: X \to Y$ must be
-equal to the composite $X \to 0 \to Y$ of the unique morphisms in and out of $0$. The biproduct structure will
-prove crucial for establishing a Cartesian closed setting for interpreting higher-order programs.
+\noindent The trivial 1-point lattice, which is both terminal and initial and we write as $0$, is the unit for
+$\biprod$. Such a structure $(\biprod, 0)$ where $\biprod$ acts as both a product and a coproduct is called a
+\emph{biproduct} (\secref{biproducts}). The presence of a zero object in a Cartesian closed category $\cat{C}$
+renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
+\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. Although the biproduct
+structure rules out Cartesian closure, it will prove crucial for establishing a Cartesian closed setting in
+\secref{higher-order}; first we develop a suitable categorical setting for first-order programs.

### 2025-05-24 main/formal-structures.tex
Slow prog..

@@
 
-$\LatGal$ is not Cartesian closed, and so we cannot use it directly to interpret higher-order programs. In
-fact Cartesian closure is incompatible with another notable property of $\LatGal$, which follows from the fact
-that the projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the
-product of lattices, have both upper and lower adjoints. This allows $X \times Y$, which we shall hereafter
-write as $X \biprod Y$, to act as both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$
-and injections $\biinj_X$ and $\biinj_Y$ given by:
+$\LatGal$ is not Cartesian closed, and so cannot serve directly as the basis for interpreting higher-order
+programs. In fact Cartesian closure is incompatible with another notable property of $\LatGal$. The
+projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the product of
+lattices, have both upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as
+$X \biprod Y$, is both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and
+injections $\biinj_X$ and $\biinj_Y$ given by:
 
@@
 renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. Although the biproduct
-structure rules out Cartesian closure, it will prove crucial for establishing a Cartesian closed setting in
-\secref{higher-order}; first we develop a suitable categorical setting for first-order programs.
+\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. However although the
+biproduct structure rules out Cartesian closure, it will prove crucial for establishing a Cartesian closed
+setting in \secref{higher-order}; but let us first see how to develop a suitable categorical setting for
+first-order programs.

### 2025-05-24 main/framework.tex
Slow prog..

@@
 
-What we want to do now is understand how to equip the interpretation of a (first-order) program with a Galois
-connections relating approximations of its inputs to approximations of its output. Since these lattices of
-approximations depend on the value being approximated, the semantics of every type $\tau$ must have both a
-``conventional'' component, providing the set $X$ which interprets $\tau$, plus an associated lattice of
-approximations $\partial X(x)$ for every point $x \in X$. The semantics of an expression $e$ of type $\sigma$
-with a single free variable of type $\tau$ (with $\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have
-two components: a function $f: X \to Y$ giving the conventional interpretation of $e$, and a family of Galois
+The goal now is to equip the interpretation of a (first-order) program with a Galois connection relating
+approximations of its inputs to approximations of its output. Since the specific lattices of approximations
+depend on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
+component, providing the set $X$ which interprets $\tau$, plus an associated lattice of approximations
+$\partial X(x)$ for every point $x \in X$. The semantics of an expression $e$ of type $\sigma$ with a single
+free variable of type $\tau$ (with $\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have two
+components: a function $f: X \to Y$ giving the conventional interpretation of $e$, and a family of Galois
 connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for
@@
 
-This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
-category $\Set^\op \to \Cat$, namely the indexed category $\Func{-}{\LatGal}$ which assigns to every set $X$
-the functor category $\Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so
-$\Func{X}{\LatGal}$ is exactly the category of $X$-indexed families of bounded lattices, with $X$-indexed
-families of Galois connections between them as morphisms. The Grothendieck construction
-$\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$, sometimes called the \emph{total
-category} for $T$, is the category of all $\Set$-indexed families of objects of $T$, together with morphisms
-that account for both structure within each family and how families transform under reindexing.
+This pattern of objects and morphisms is captured by the construction used in CHAD~\cite{vákár22,nunes2023} to
+describe automatic differentiation, namely the \emph{Grothendieck construction} for a particular indexed
+category $T: \Set^\op \to \Cat$. For CHAD $T(X)$ is the functor category $\Func{X}{\FinVect}$, regarded as the
+category of $X$-indexed families of finite vector spaces, with ; here $T(X) = \Func{X}{\LatGal}$, the category of
+$X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections between them as
+morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed
+category $T$, sometimes called the \emph{total category} for $T$, is the category of all $\Set$-indexed
+families of objects of $T$, together with morphisms that account for both structure within each family and how
+families transform under reindexing.
 

### 2025-05-24 main/framework.tex
Slow prog..

@@
 describe automatic differentiation, namely the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$. For CHAD $T(X)$ is the functor category $\Func{X}{\FinVect}$, regarded as the
-category of $X$-indexed families of finite vector spaces, with ; here $T(X) = \Func{X}{\LatGal}$, the category of
-$X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections between them as
-morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed
+category $T: \Set^\op \to \Cat$. For CHAD $T(X)$ will be the functor category $\Func{X}{\FinVect}$, regarded
+as the category of $X$-indexed families of finite vector spaces, with indexed linear maps; here $T(X) =
+\Func{X}{\LatGal}$, the category of $X$-indexed families of bounded lattices, with indexed Galois connections
+as morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed
 category $T$, sometimes called the \emph{total category} for $T$, is the category of all $\Set$-indexed

### 2025-05-24 main/category-of-families.tex
Rename file.

--- 2025-05-24 main/framework.tex
@@
 free variable of type $\tau$ (with $\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have two
-components: a function $f: X \to Y$ giving the conventional interpretation of $e$, and a family of Galois
+components: a function $f: X \to Y$ giving the conventional interpretation of $e$, plus a family of Galois
 connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for
@@
 describe automatic differentiation, namely the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$. For CHAD $T(X)$ will be the functor category $\Func{X}{\FinVect}$, regarded
-as the category of $X$-indexed families of finite vector spaces, with indexed linear maps; here $T(X) =
-\Func{X}{\LatGal}$, the category of $X$-indexed families of bounded lattices, with indexed Galois connections
-as morphisms. The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed
-category $T$, sometimes called the \emph{total category} for $T$, is the category of all $\Set$-indexed
-families of objects of $T$, together with morphisms that account for both structure within each family and how
-families transform under reindexing.
+category $T: \Set^\op \to \Cat$. For CHAD, $T(X)$ will be the functor category $\Func{X}{\FinVect}$, regarded
+as the category of $X$-indexed families of finite vector spaces, with indexed linear maps; for Galois slicing
+we will take $T(X) = \Func{X}{\LatGal}$, the category of $X$-indexed families of bounded lattices, with
+indexed Galois connections as morphisms. The Grothendieck construction
+$\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$, sometimes called the \emph{total
+category} for $T$, is the category of all $\Set$-indexed families of objects of $T$, together with morphisms
+that account for both structure within each family and how families transform under reindexing.
 
@@
 
-\subsection{Category of Families}
+\subsubsection{Category of Families}
 \label{sec:Fam}
@@
 \partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial
-X(x) \to \partial Y(f(x))$, as we sketched informally at the beginning of this section. The pullback
-(reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
-\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
-connection at each point, using the baseline (unapproximated) function $f$.
+X(x) \to \partial Y(f(x))$. This is precisely the setup we identified informally at the beginning of this
+section. The pullback (reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g
+\comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and
+Galois connection at each point, using the baseline (unapproximated) function $f$.
+
+\subsection{Automatic Differentiation via the Grothendieck Construction}
+
+Automatic differentiation for first-order programs can also be recovered in this setup, via the total category
+$\Fam(\FinVect)$.

### 2025-05-24 main/category-of-families.tex
Think I need to elide the Diff category I introduced in the notes..

@@
 
-\subsubsection{Category of Families}
+\subsection{Category of Families}
 \label{sec:Fam}
@@
 
-\subsection{Galois Slicing via the Grothendieck Construction}
+\subsection{Galois Slicing via a Category of Families}
 
-Now consider the total category $\Fam(\LatGal)$, which is our basic setting for interpreting first-order
-programs for \GPS. This has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$,
-a bounded lattice $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f,
-\partial f)$ of a function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial
-X(x) \to \partial Y(f(x))$. This is precisely the setup we identified informally at the beginning of this
-section. The pullback (reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g
-\comp f, \reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and
-Galois connection at each point, using the baseline (unapproximated) function $f$.
+The total category $\Fam(\LatGal)$ is then basic setting for interpreting first-order programs for \GPS. This
+has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
+$\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of a
+function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to \partial
+Y(f(x))$. This is precisely the setup we identified informally at the beginning of this section. The pullback
+(reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
+\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
+connection at each point, using the baseline (unapproximated) function $f$.
 
-\subsection{Automatic Differentiation via the Grothendieck Construction}
+\subsection{Automatic Differentiation via a Category of Families}
 
-Automatic differentiation for first-order programs can also be recovered in this setup, via the total category
-$\Fam(\FinVect)$.
+Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
+setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic idea here. For flat manifolds like
+$\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically isomorphic to $\RR^n$, so
+the forward and backward derivatives at $x$ of a differentiable function $f: \RR^m \to \RR^n$ have the
+following form.
+
+\begin{itemize}
+\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of $f$ at $x \in \RR^m$ is the
+unique linear map $\RR^m \linearto \RR^n$ given by the Jacobean matrix of $f$ at $x$.
+\item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
+$\RR^n \linearto \RR^m$ given by the transpose (adjoint) of the Jacobean matrix of $f$ at $x$.
+\end{itemize}
+
+\subsubsection{Chain rule}
+
+Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
+we have:
+
+\begin{itemize}
+\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
+\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
+\end{itemize}
+
+\subsubsection{Forward mode automatic differentiation}
+
+\begin{definition}
+Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
+$f: \RR^m \to \RR^n$ to the function $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
+(\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
+that point.
+\end{definition}
+
+For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ it is useful to define $f_1 = \pi_1 \comp f:
+\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$.
+
+\begin{definition}[$\Diff$]
+\label{def:auto-diff:Diff}
+\todo{Overkill -- try and give a construction that lives in $\Fam(\FinVect)$ directly} Define $\Diff$
+to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are all maps $f:
+\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ with $f_1$ differentiable. Composition in $\Diff$ is given by:
+\begin{align*}
+(g \comp f)_1(x) &= f_1(g_1(x)) \\
+(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
+\end{align*}
+\end{definition}
+
+We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
+\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
+similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
+the derivative of $(g \comp f)_1$ at $x$.
+
+\subsubsection{Embedding into $\Fam(\FinVect)$}
+\label{sec:galois-slicing-auto-diff-via-fam:auto-diff}
+
+Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of
+tangent spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
+pairs $(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2:
+\{{f_2}_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. This exhibits $\Diff$ as equivalent to a full subcategory
+of $\Fam(\FinVect)$.
+
+In that subcategory, for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
+\partial g): (\RR^n, \const(\RR^n)) \to (\RR^k, \const(\RR^k))$, we have that for any $x \in \RR^m$:
+\begin{itemize}
+\item $\partial f_x: \RR^m \to \RR^n$;
+\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
+\end{itemize}
+
+\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
+\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the forward chain rule,
+since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.

### 2025-05-24 main/category-of-families.tex
Think I need to elide the Diff category I introduced in the notes..

@@
 Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
-$f: \RR^m \to \RR^n$ to the function $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
+$f: \RR^m \to \RR^n$ to $(f, \pushf{f})$: \RR^m \to \RR^n \times
 (\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at

### 2025-05-24 main/category-of-families.tex
Some simplification of auto-diff.

--- 2025-05-24 main/category-of-families.tex
@@
 For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ it is useful to define $f_1 = \pi_1 \comp f:
-\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$.
-
-\begin{definition}[$\Diff$]
-\label{def:auto-diff:Diff}
-\todo{Overkill -- try and give a construction that lives in $\Fam(\FinVect)$ directly} Define $\Diff$
-to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are all maps $f:
-\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ with $f_1$ differentiable. Composition in $\Diff$ is given by:
+\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$. Then composition of such maps can be expressed as:
 \begin{align*}
@@
 \end{align*}
-\end{definition}
 
@@
 similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$.
+the derivative of $(g \comp f)_1$ at $x$. We can see these maps as living in $\Fam(\FinVect)$ if we regard
+every $\RR^n$ as paired with a constant family of tangent spaces also of the form $\RR^n$ and each map $f:
+\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs $(f_1, f_2)$ as the pair of a differentiable function
+$f_1: \RR^m \to \RR^n$ and a family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$.
 
-\subsubsection{Embedding into $\Fam(\FinVect)$}
-\label{sec:galois-slicing-auto-diff-via-fam:auto-diff}
-
-Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which send objects $\RR^n$ to constant families of
-tangent spaces $(\RR^n, \const(\RR^n))$ and morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
-pairs $(f_1, f_2)$ of differentiable functions $f_1: \RR^m \to \RR^n$ and families of linear maps $f_2:
-\{{f_2}_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. This exhibits $\Diff$ as equivalent to a full subcategory
-of $\Fam(\FinVect)$.
-
-In that subcategory, for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
+Then for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
 \partial g): (\RR^n, \const(\RR^n)) \to (\RR^k, \const(\RR^k))$, we have that for any $x \in \RR^m$:

### 2025-05-24 main/formal-structures.tex
Some simplification of auto-diff.

@@
 renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. However although the
-biproduct structure rules out Cartesian closure, it will prove crucial for establishing a Cartesian closed
-setting in \secref{higher-order}; but let us first see how to develop a suitable categorical setting for
-first-order programs.
+\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. However although biproducts
+rule out Cartesian closure, they will prove crucial for establishing a Cartesian closed setting in
+\secref{higher-order}; but let us first see how to develop a suitable categorical setting for first-order
+programs.

### 2025-05-24 main/higher-order.tex
Expand why LatGal isn't CC.
Some simplification of auto-diff.

@@
 \section{Interpreting Higher-Order Programs}
+\label{sec:higher-order}
 
@@
 \begin{enumerate}
-\item conditions under which $\Fam(\cat{C})$ is Cartesian closed
+\item if $\cat{C}$ has biproducts and all small products then $\Fam(\cat{C})$ is Cartesian closed
 \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$)
-\item why we need $\PSh_{\CMon}(\cat{C})$
+\item $\FinVect$ and $\LatGal$ have biproducts but not all small products
+\item but actually we need $\PSh_{\CMon}(\cat{C})$ rather than $\PSh(\cat{C})$ because $\CMon$-enriched Yoneda embedding preserves biproducts
 \end{enumerate}

### 2025-05-25 main/introduction.tex
elliott18.

@@
 It would be satisfying to have a denotational account of \GPS as well, for which traces would then offer a plausible implementation technique (perhaps one of several).
-As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08}, which points to a way to develop such an approach.
+As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18}, which points to a way to develop such an approach.
 We have already hinted at this in the description above, but let us now make it explicit.

### 2025-05-25 main/introduction.tex
spivak65.

@@
 
-Data provenance aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
-Provenance tracking has a wide range of applications, from debugging and program comprehension to improving reproducibility and transparency in scientific workflows.
-In order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome.
+To audit any computational process, we need robust and well-founded notions of \emph{provenance} to track how data are used.
+This allows us to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
+Provenance tracking has a wide range of applications, from debugging and program comprehension \todo{cite1} to improving reproducibility and transparency in scientific workflows \todo{cite2}.
 \emph{Program slicing}, first proposed by~\citet{weiser81}, is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.

### 2025-05-25 main/category-of-families.tex
spivak65.
Less noisy notation for constant family.

@@
 
-Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
-we have:
+Derivatives respect composition~\cite{spivak65}. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For
+any $x \in \RR^m$ we have:
 
@@
 the derivative of $(g \comp f)_1$ at $x$. We can see these maps as living in $\Fam(\FinVect)$ if we regard
-every $\RR^n$ as paired with a constant family of tangent spaces also of the form $\RR^n$ and each map $f:
-\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to pairs $(f_1, f_2)$ as the pair of a differentiable function
-$f_1: \RR^m \to \RR^n$ and a family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$.
+every finite vector space $\RR^n$ as paired with the constant family of tangent spaces which is always $\RR^n$
+(also written $\RR^n$, for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
+pairs $(f_1, f_2)$ as the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and a family of linear maps
+$f_2: \RR^m \to \RR^m \linearto \RR^n$.
 
-Then for any $(f, \partial f): (\RR^m, \const(\RR^m)) \to (\RR^n, \const(\RR^n))$ and $(g,
-\partial g): (\RR^n, \const(\RR^n)) \to (\RR^k, \const(\RR^k))$, we have that for any $x \in \RR^m$:
+Then for any $(f, \partial f): (\RR^m, \RR^m) \to (\RR^n, \RR^n)$ and $(g, \partial g): (\RR^n, \RR^n) \to
+(\RR^k, \RR^k)$, we have that for any $x \in \RR^m$:
 \begin{itemize}
@@
 \noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f): (\RR^m, \const(\RR^m)) \to (\RR^k, \const(\RR^k))$ corresponds exactly to the forward chain rule,
+\partial f): (\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule,
 since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.

### 2025-05-25 main/higher-order.tex
Less noisy notation for constant family.

@@
   \item if $\cat{C}$ has biproducts and all small products then $\Fam(\cat{C})$ is Cartesian closed
-\item $\PSh(\cat{C})$ has all small products (inherited from $\Set$)
-\item $\FinVect$ and $\LatGal$ have biproducts but not all small products
-\item but actually we need $\PSh_{\CMon}(\cat{C})$ rather than $\PSh(\cat{C})$ because $\CMon$-enriched Yoneda embedding preserves biproducts
+  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products (\todo{restrict $\LatGal$ to finite lattices?})
+  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
+  biproducts from $\cat{C}$
+  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
+  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
+  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how we will interpret base types?
+E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a suitable object 2}
   \end{enumerate}

### 2025-05-25 main/category-of-families.tex
Less noisy notation for constant family.
Cleanup.

@@
 
-\begin{definition}[Category of families]
+\begin{definition}[Categories of families]
 \label{def:Fam}
@@
 every finite vector space $\RR^n$ as paired with the constant family of tangent spaces which is always $\RR^n$
-(also written $\RR^n$, for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
-pairs $(f_1, f_2)$ as the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and a family of linear maps
-$f_2: \RR^m \to \RR^m \linearto \RR^n$.
+(also written $\RR^n$, for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as
+the pair $(f_1, f_2)$ of a differentiable function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2:
+\RR^m \to \RR^m \linearto \RR^n$.
 

### 2025-05-25 main/category-of-families.tex
Cleanup.

@@
 $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically isomorphic to $\RR^n$, so
-the forward and backward derivatives at $x$ of a differentiable function $f: \RR^m \to \RR^n$ have the
-following form.
+the forward and backward derivatives of a differentiable function $f: \RR^m \to \RR^n$ at a point $x$ have the
+following form:
 
@@
 
-\subsubsection{Chain rule}
+\subsubsection{Chain Rule}
 
@@
 
-\subsubsection{Forward mode automatic differentiation}
+\subsubsection{Forward-Mode Automatic Differentiation}
 
@@
 For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ it is useful to define $f_1 = \pi_1 \comp f:
-\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$. Then composition of such maps can be expressed as:
+\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$. Then composition of such maps
+can be expressed as:
 \begin{align*}

### 2025-05-25 main/category-of-families.tex
Identify a bit of reorg needed for Grothendieck section.

@@
 approximations of its inputs to approximations of its output. Since the specific lattices of approximations
-depend on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
-component, providing the set $X$ which interprets $\tau$, plus an associated lattice of approximations
-$\partial X(x)$ for every point $x \in X$. The semantics of an expression $e$ of type $\sigma$ with a single
-free variable of type $\tau$ (with $\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have two
-components: a function $f: X \to Y$ giving the conventional interpretation of $e$, plus a family of Galois
-connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$ which can be used for
-slicing over that expression.
+depend on the value being approximated, the semantics of every type $\tau$ must have both a conventional
+component as well as an approximating component, i.e.~a set $X$ which serves as the usual interpretation of
+$\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The semantics
+of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with $\tau$ and $\sigma$
+denoting $X$ and $Y$) must then also have two components: a function $f: X \to Y$ giving the conventional
+interpretation of $e$, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$
+for every $x \in X$ which can be used for slicing over that expression.
 
 This pattern of objects and morphisms is captured by the construction used in CHAD~\cite{vákár22,nunes2023} to
-describe automatic differentiation, namely the \emph{Grothendieck construction} for a particular indexed
-category $T: \Set^\op \to \Cat$. For CHAD, $T(X)$ will be the functor category $\Func{X}{\FinVect}$, regarded
-as the category of $X$-indexed families of finite vector spaces, with indexed linear maps; for Galois slicing
-we will take $T(X) = \Func{X}{\LatGal}$, the category of $X$-indexed families of bounded lattices, with
-indexed Galois connections as morphisms. The Grothendieck construction
-$\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$, sometimes called the \emph{total
-category} for $T$, is the category of all $\Set$-indexed families of objects of $T$, together with morphisms
-that account for both structure within each family and how families transform under reindexing.
+model automatic differentiation, namely the \emph{Grothendieck construction} for a particular set-indexed
+category $T: \Set^\op \to \Cat$. For automatic differentiation, $T(X)$ will be the functor category
+$\Func{X}{\FinVect}$, regarded as the category of $X$-indexed families of finite vector spaces, with indexed
+linear maps as morphisms. For Galois slicing, we will take $T(X) = \Func{X}{\LatGal}$, the category of
+$X$-indexed families of bounded lattices, with indexed Galois connections as morphisms.
 
-First we formalise $X$-indexed families and their categories.
+The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$,
+sometimes called the \emph{total category} for $T$, is the category of all set-indexed families of objects of
+$T$, together with morphisms that account for both structure within each family and how families transform
+under reindexing functions $f: X \to Y$. First we formalise $X$-indexed families and their categories.
 
@@
 \label{def:Grothendieck}
-Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
-category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
-\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
-\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
-morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
+\todo{Move earlier because doesn't assume $\cat{C} = \Set$.} Suppose an indexed category $T: \cat{C}^\op \to
+\Cat$. The \emph{Grothendieck construction} for $T$ is the category $\Grothendieck{X}T(X)$ (also written
+$\Grothendieck{}T$) which has as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an
+object $\partial X$ of $T(X)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial
+f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in
+$T(X)$.
 \end{definition}

### 2025-05-25 main/category-of-families.tex
First pass over forward-mode AD.

@@
 function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to \partial
-Y(f(x))$. This is precisely the setup we identified informally at the beginning of this section. The pullback
-(reindexing) along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
-\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
-connection at each point, using the baseline (unapproximated) function $f$.
+Y(f(x))$. This is precisely the setup we identified informally at the beginning of this section. The
+reindexing along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial
+g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois connection at each
+point, using the baseline (unapproximated) function $f$.
 
@@
 
-\begin{definition}
-Define the functor $\tangents$ which sends every vector space $\RR^n$ to itself and every differentiable map
-$f: \RR^m \to \RR^n$ to the function $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
-(\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
-that point.
-\end{definition}
-
-For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ it is useful to define $f_1 = \pi_1 \comp f:
-\RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$. Then composition of such maps
-can be expressed as:
+The basic idea of (forward-mode) automatic differentiation, as first described by \citet{linnainmaa76}, was to
+recognise that one could efficiently compute the pushforward map $\RR^m \to (\RR^m \linearto \RR^n)$ of a
+differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself by defining a single map
+$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Defining $f_1
+= \pi_1 \comp f: \RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$, composition of
+such maps can be expressed as:
 \begin{align*}
@@
 
-We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m \linearto
-\RR^n$ happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
+We can verify that $\tangents$ is functorial, regardless of whether the linear maps $\RR^m \linearto \RR^n$
+happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
 similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$. We can see these maps as living in $\Fam(\FinVect)$ if we regard
-every finite vector space $\RR^n$ as paired with the constant family of tangent spaces which is always $\RR^n$
-(also written $\RR^n$, for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as
-the pair $(f_1, f_2)$ of a differentiable function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2:
-\RR^m \to \RR^m \linearto \RR^n$.
+the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing composite forward
+derivatives automatically so long as the primitive operations of the system are appropriately equipped with
+them.
 
-Then for any $(f, \partial f): (\RR^m, \RR^m) \to (\RR^n, \RR^n)$ and $(g, \partial g): (\RR^n, \RR^n) \to
-(\RR^k, \RR^k)$, we have that for any $x \in \RR^m$:
-\begin{itemize}
-\item $\partial f_x: \RR^m \to \RR^n$;
-\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$.
-\end{itemize}
+We can see these maps as living in $\Fam(\FinVect)$ if we regard every finite vector space $\RR^n$ as paired
+with the family of tangent spaces which is constantly $\RR^n$ (also written $\RR^n$, for convenience), and
+each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$ of a differentiable
+function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$. Then the
+composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f):
+(\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial
+g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ computing
+the pushforward.
 
-\noindent The composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f): (\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule,
-since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+\subsubsection{Reverse-Mode Automatic Differentiation}
+
+\todo{brief summary}

### 2025-05-25 main/category-of-families.tex
First pass over forward-mode AD.

@@
 The basic idea of (forward-mode) automatic differentiation, as first described by \citet{linnainmaa76}, was to
-recognise that one could efficiently compute the pushforward map $\RR^m \to (\RR^m \linearto \RR^n)$ of a
-differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself by defining a single map
-$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Defining $f_1
-= \pi_1 \comp f: \RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$, composition of
-such maps can be expressed as:
+recognise that one could efficiently compute the pushforward map $\pushf{f}: \RR^m \to (\RR^m \linearto
+\RR^n)$ of a differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself by defining
+a single map $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$.
+Defining $f_1 = \pi_1 \comp f: \RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$,
+composition of such maps can be expressed as:
 \begin{align*}

### 2025-05-25 main/category-of-families.tex
Progress, bit clumsy still..

@@
 \subsection{Galois Slicing via a Category of Families}
+\label{sec:fam:galois-slicing}
 
@@
 similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing composite forward
-derivatives automatically so long as the primitive operations of the system are appropriately equipped with
-them.
+the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing the forward derivative
+of a composite function automatically, without computing each constituent $f(x)$ more than once, so long as
+the primitive operations of the system can provide derivatives. An efficient implementation of \GPS based on
+$\Fam(\LatGal)$ (\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to avoid
+recomputing $f(x)$ at each step during forward slicing.
 
-We can see these maps as living in $\Fam(\FinVect)$ if we regard every finite vector space $\RR^n$ as paired
-with the family of tangent spaces which is constantly $\RR^n$ (also written $\RR^n$, for convenience), and
-each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$ of a differentiable
-function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$. Then the
-composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f):
-(\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial
-g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ computing
-the pushforward.
+We can see these ``optimised'' maps as living in $\Fam(\FinVect)$ if we regard every finite vector space
+$\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$, for
+convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$. Then
+the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial
+f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x =
+{\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ selecting the appropriate pushforward
+by recomputing $f(x)$.
 

### 2025-05-25 main/category-of-families.tex
Progress, bit clumsy still..

@@
 the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing the forward derivative
-of a composite function automatically, without computing each constituent $f(x)$ more than once, so long as
-the primitive operations of the system can provide derivatives. An efficient implementation of \GPS based on
-$\Fam(\LatGal)$ (\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to avoid
-recomputing $f(x)$ at each step during forward slicing.
+of a composite function without have to recompute the function itself, so long as all atomic operations
+provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$ (\secref{fam:galois-slicing}
+above) would most likely apply a similar optimisation to avoid recomputing $f(x)$ at each step during forward
+slicing.
 

### 2025-05-25 main/category-of-families.tex
Prob ok for first pass on forward-mode AD.

--- 2025-05-25 main/category-of-families.tex
@@
 similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing composite forward
-derivatives automatically so long as the primitive operations of the system are appropriately equipped with
-them.
+the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing the forward derivative
+of a composite function, without have to recompute the function itself, as long as all atomic operations
+provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$ (\secref{fam:galois-slicing}
+above) would most likely apply a similar optimisation to make it possible to compute forwards slices without
+having to reexecute the program at the same time.
 
-We can see these maps as living in $\Fam(\FinVect)$ if we regard every finite vector space $\RR^n$ as paired
-with the family of tangent spaces which is constantly $\RR^n$ (also written $\RR^n$, for convenience), and
-each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$ of a differentiable
-function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$. Then the
-composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f):
-(\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial
-g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ computing
-the pushforward.
+We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
+vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
+for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
+Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
+\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
+f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ selecting the appropriate
+pushforward by recomputing $f(x)$.
 

### 2025-05-25 main/category-of-families.tex
Onto higher-order skeleton.

@@
 recognise that one could efficiently compute the pushforward map $\pushf{f}: \RR^m \to (\RR^m \linearto
-\RR^n)$ of a differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself by defining
+\RR^n)$ of a differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself, by defining
 a single map $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$.
@@
 similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing the forward derivative
-of a composite function, without have to recompute the function itself, as long as all atomic operations
-provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$ (\secref{fam:galois-slicing}
-above) would most likely apply a similar optimisation to make it possible to compute forwards slices without
-having to reexecute the program at the same time.
+the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides an algorithmic method for computing the
+forward derivative of a composite function, without have to recompute the function itself, as long as all
+atomic operations provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$
+(\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to make it possible to
+compute forwards slices without having to reexecute the program at the same time.
 

### 2025-05-25 main/higher-order.tex
Onto higher-order skeleton.
Set-wide products. If C has biproducts and set-wide products, then C is CC.

@@
 \label{sec:higher-order}
+
+Although $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide an appropriate notion of composition for automatic
+differentiation and Galois slicing respectively, neither is Cartesian closed. Both are therefore unsuitable
+for interpreting higher-order programs. However the following is a special case of a construction due to
+\citet{nunes2023}:
+
+\begin{proposition}
+Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
+$\Fam(\cat{C})$ is Cartesian closed.
+\end{proposition}
+
+$\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$ of objects
+of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family of
+projection morphisms $\{\eval_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
+and any family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product
+morphism $\lambda(f)$ making the following diagram commute for every $x \in X$:
+
+\begin{center}
+\begin{tikzcd}
+   \prod_{X} A \arrow[r, "\eval_x"] & A_x \\
+   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\lambda(f)"]
+\end{tikzcd}
+\end{center}
 
@@
 
-We can alternatively take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
+Alternatively, we can take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
 \biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore

### 2025-05-25 main/higher-order.tex
Set-wide products. If C has biproducts and set-wide products, then C is CC.

@@
 
-Although $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide an appropriate notion of composition for automatic
-differentiation and Galois slicing respectively, neither is Cartesian closed. Both are therefore unsuitable
-for interpreting higher-order programs. However the following is a special case of a construction due to
+$\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
+and Galois slicing respectively, but neither is Cartesian closed. Both are therefore unsuitable for
+interpreting higher-order programs. However the following is a special case of a construction due to
 \citet{nunes2023}:

### 2025-05-25 main/category-of-families.tex
Give the construction of the internal hom in Fam(C).

@@
 
-\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
-$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
-$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
+\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all sets $X$ paired with indexed families
+$\partial X: X \to \cat{C}$ and the morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are all
+functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
 $\Func{X}{\cat{C}}$.
+
+\todo{Maybe unpack why these are true}
+
+\begin{proposition}
+\label{prop:Grothendieck:fam-inherits}
+\item
+\begin{enumerate}
+\item If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
+\item If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
+\end{enumerate}
+\end{proposition}
 

### 2025-05-25 main/higher-order.tex
Give the construction of the internal hom in Fam(C).

@@
 $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed. Both are therefore unsuitable for
+and Galois slicing respectively, but neither is Cartesian closed. Neither is therefore suitable for
 interpreting higher-order programs. However the following is a special case of a construction due to
@@
 \end{center}
+
+If $\cat{C}$ has all small products and also biproducts, then the internal hom $\internalHom{(X, \partial
+X)}{(Y, \partial Y)}$ in $\Fam(\cat{C})$ is the object $(Z, \partial Z)$, where:
+\begin{itemize}
+\item the indexing set $Z$ is the hom-set $\Fam(\cat{C})(X,Y)$, which is indeed a set by
+\propref{Grothendieck:fam-inherits};
+\item the indexed family $\partial Z$ is the family of objects $\prod_{X}\reindex{\partial Y}{f}$ for every
+${(f, \partial f): (X, \partial X) \to (Y, \partial Y)}$.
+\end{itemize}
 

### 2025-05-25 main/higher-order.tex
Give the construction of the internal hom in Fam(C).

@@
 $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed. Neither is therefore suitable for
-interpreting higher-order programs. However the following is a special case of a construction due to
-\citet{nunes2023}:
+and Galois slicing respectively, but neither is Cartesian closed, and therefore suitable for interpreting
+higher-order programs. However the following is a special case of a construction due to \citet{nunes2023}:
 

### 2025-05-25 main/higher-order.tex
Progress on HO.

@@
 $\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed, and therefore suitable for interpreting
+and Galois slicing respectively, but neither is Cartesian closed and thereby a candidate for interpreting
 higher-order programs. However the following is a special case of a construction due to \citet{nunes2023}:

### 2025-05-26 main/higher-order.tex
Tweak proposition 3.3.

--- 2025-05-25 main/higher-order.tex
@@
 
-$\Fam(\FinVect)$ nor $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed. Neither is therefore suitable for
+$\Fam(\FinVect)$ and $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
+and Galois slicing respectively, but neither is Cartesian closed and thereby a potential candidate for
 interpreting higher-order programs. However the following is a special case of a construction due to
@@
 projection morphisms $\{\eval_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
-and any family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product
-morphism $\lambda(f)$ making the following diagram commute for every $x \in X$:
+and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
+$\lambda(f)$ making the following diagram commute for every $x \in X$:
 
@@
 \begin{itemize}
-\item the indexing set $Z$ is the hom-set $\Fam(\cat{C})(X,Y)$, which is indeed a set by
+\item the indexing set $Z$ is the $\Fam(\cat{C})((X, \partial X), (Y, \partial Y))$, which is indeed a set by
 \propref{Grothendieck:fam-inherits};
@@
 
+\todo{Explain role of biproducts}
+
 \todo{
 \begin{enumerate}
-  \item if $\cat{C}$ has biproducts and all small products then $\Fam(\cat{C})$ is Cartesian closed
   \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products (\todo{restrict $\LatGal$ to finite lattices?})
@@
 \subsubsection{Biproduct laws}
-It is also easy to show that in a semi-additive category, as defined in \defref{biproducts:semi-additive}, the
-following equations hold:
-
-\vspace{-3mm}
+It is also easy to show that in a semi-additive category the following equations hold:
+%\vspace{-4mm}
 \begin{minipage}[t]{0.45\textwidth}
@@
 
-\noindent and
-
 \begin{salign*}
@@
 
-Alternatively, we can take these to be the defining equations of a biproduct $(X \biprod Y, \biproj_X,
-\biproj_Y, \biinj_X, \biinj_Y)$ in any category $\cat{C}$ known to be $\CMon$-enriched, and therefore
-providing addition of morphisms; in that case any products or coproducts in $\cat{C}$ are necessarily
-biproducts.
+\vspace{3mm}
+\noindent It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
+in $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
+coproducts in $\cat{C}$ are necessarily biproducts.
 
@@
 \label{prop:biproducts:from-product-or-coproduct}
-Suppose $\cat{C}$ is $\CMon$-enriched.
+In a category $\cat{C}$ enriched in $\CMon$:
 \begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
-is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
-\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
-\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
-$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
-\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
+\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \inj_1, \inj_2)$ is a coproduct, where
+$\inj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\inj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
+\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \pi_1, \pi_2)$ is a product, where
+$\pi_1 = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and $\pi_2 = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to
+Y$.
+\item If $X$ is terminal (resp.~initial) then $X$ is initial (resp.~terminal).
 \end{enumerate}

### 2025-05-26 main/category-of-families.tex
Initial import of FinVect, but too lengthy..have specialised to the reals, but should also specialise to the finite case too.

@@
 
-\todo{Maybe unpack why these are true}
-
 \begin{proposition}
 \label{prop:Grothendieck:fam-inherits}
-\item
+\item \todo{Maybe unpack why these are true}
 \begin{enumerate}
@@
 Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
-setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic idea here. For flat manifolds like
-$\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically isomorphic to $\RR^n$, so
-the forward and backward derivatives of a differentiable function $f: \RR^m \to \RR^n$ at a point $x$ have the
-following form:
+setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic ideas here, starting with the
+category $\FinVect$ of finite vector spaces over $\RR$.
+
+\subsubsection{Category of finite-dimensional vector spaces}
+\label{sec:categories-with-biproducts:fdvect}
+
+\begin{definition}[Vector space over $\RR$]
+We define vector spaces over the reals, with addition of real numbers $a, b$ written $a + b$ and
+multiplication written $a \mult b$. Then a \emph{vector space $V = (V, +, \mult)$ over $\RR$} (overloading $+$
+and $\mult$) is a set $V$ equipped with addition of vectors $+: V^2 \to V$ and scalar multiplication $\mult:
+\RR \times V \to V$ where $(V,+)$ is an abelian group (with identity element $0_V$ and additive inverse $-v$
+for any $v \in V$) and the vector operations are compatible with the field operations in that the following
+equations hold:
+
+\vspace{-3mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   1 \mult v &= v \\
+   (a \mult b) \mult v &= a \mult (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
+   a \mult (u + v) &= (a \mult u) + (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}
+\end{definition}
+
+Up to isomorphism a finite-dimensional vector space over $\RR$ of dimension $n \in \Nat$ is the vector space
+$(\RR^n, +, \mult)$ where $+$ and $\mult$ are defined component-wise.
+
+\begin{definition}[Category $\FinVect$]
+Define $\FinVect$ to be the category which has:
+\begin{itemize}
+\item as objects $V, W$ all pairs of natural numbers $n$ and $n$-dimensional vector spaces $(\RR^n, +, \mult)$
+\item as morphisms $f: V \to W$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
+\mult f(v)$.
+\end{itemize}
+\end{definition}
+
+$\FinVect$ is enriched in $\Ab$, the category of abelian groups. For any $V, W$ we have that $\FinVect(V,W)$
+is an abelian group, where:
+\begin{enumerate}
+\item $0_{V,W}(v) = 0_W$
+\item $(f + g)(v) = f(v) + g(v)$
+\item $(-f)(v) = -f(v)$
+\end{enumerate}
+\noindent with composition bilinear by (2) and addition of morphisms computed component-wise. The direct sum
+$V \biprod W = (V \times W, +, \mult)$, with $+$ and $\mult$ defined component-wise, is a biproduct $(V
+\biprod W, \biinj_V, \biinj_W, \pi_1, \pi_2)$ with $\biinj_V = \prodM{\id_V}{0_{V,W}}$ and $\biinj_W =
+\prodM{0_{V,W}}{\id_W}$.
+
+\subsubsection{Derivatives of Differentiable Functions}
+
+For flat manifolds like $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically
+isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
+\RR^n$ at a point $x$ have the following form:
 

### 2025-05-26 main/category-of-families.tex
First pass over FinVect, still a bit subscript-heavy, but better.

--- 2025-05-25 main/category-of-families.tex
@@
 Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
-setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic idea here. For flat manifolds like
-$\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically isomorphic to $\RR^n$, so
-the forward and backward derivatives of a differentiable function $f: \RR^m \to \RR^n$ at a point $x$ have the
-following form:
+setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic ideas here, starting with the
+category $\FinVect$ of finite vector spaces over $\RR$, with addition of real numbers written $a + b$ and
+multiplication written $a \mult b$.
+
+\subsubsection{Category of finite-dimensional vector spaces}
+\label{sec:categories-with-biproducts:fdvect}
+
+\begin{definition}[Finite dimensional vector space over $\RR$]
+Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
+component-wise addition of vectors $+: \RR^n \times \RR^n \to \RR^n$ and scalar multiplication $\mult: \RR
+\times \RR^n \to \RR^n$ (overloading $+$ and $\mult$), where $(\RR^n,+)$ is an abelian group (with identity
+$0$ and additive inverse $-v$ again defined component-wise) and the vector operations are compatible with the
+field operations in that the following equations hold:
+
+\vspace{-4mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   1 \mult v &= v \\
+   (a \mult b) \mult v &= a \mult (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
+   a \mult (u + v) &= (a \mult u) + (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}
+\end{definition}
+
+\begin{definition}[Category $\FinVect$]
+Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces, and as
+morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
+\mult f(v)$.
+\end{definition}
+
+$\FinVect$ is enriched in $\Ab$, the category of abelian groups. For any objects $V = \RR^m$ and $W = \RR^n$
+the hom-set $\FinVect(V,W)$ is an abelian group, with identity $0_{V,W}$ given by the constant map $v \mapsto
+0_W$ and addition and additive inverse given pointwise:
+
+\vspace{-4mm}
+\begin{center}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{align*}
+(f + g)(v) &= f(v) + g(v)
+\end{align*}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{align*}
+(-f)(v) &= -f(v)
+\end{align*}
+\end{minipage}
+\end{center}
+
+\vspace{1mm}
+\noindent Composition is with bilinear. The direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian
+product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
+\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$.
+
+\subsubsection{Derivatives of Differentiable Functions}
+
+For flat manifolds like $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically
+isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
+\RR^n$ at a point $x$ have the following form:
 

### 2025-05-26 main/higher-order.tex
Avoid epsilon, lambda for set-indexed products.

@@
 of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family of
-projection morphisms $\{\eval_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
+projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
 and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
-$\lambda(f)$ making the following diagram commute for every $x \in X$:
+$\prodFam{f}$ making the following diagram commute for every $x \in X$:
 
@@
 \begin{tikzcd}
-   \prod_{X} A \arrow[r, "\eval_x"] & A_x \\
-   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\lambda(f)"]
+   \prod_{X} A \arrow[r, "\pi_x"] & A_x \\
+   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\prodFam{f}"]
 \end{tikzcd}

### 2025-05-26 main/category-of-families.tex
Avoid epsilon, lambda for set-indexed products.
First pass over currying in Fam(C).

@@
 \begin{proposition}
-\label{prop:Grothendieck:fam-inherits}
-\item \todo{Maybe unpack why these are true}
-\begin{enumerate}
-\item If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
-\item If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
-\end{enumerate}
+\label{prop:Grothendieck:fam-inherits-local-smallness}
+If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
+\end{proposition}
+
+\begin{proposition}
+\label{prop:Grothendieck:fam-inherits-products}
+If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
 \end{proposition}
@@
 \begin{center}
-\begin{minipage}[t]{0.45\textwidth}
+\begin{minipage}[t]{0.35\textwidth}
 \begin{align*}
@@
 \end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
+\begin{minipage}[t]{0.35\textwidth}
 \begin{align*}

### 2025-05-26 main/higher-order.tex
First pass over currying in Fam(C).

@@
 $\Fam(\FinVect)$ and $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed and thereby a potential candidate for
-interpreting higher-order programs. However the following is a special case of a construction due to
-\citet{nunes2023}:
+and Galois slicing respectively, but neither is Cartesian closed and thereby able to interpret higher-order
+programs directly. However the following is a special case of a construction due to \citet{nunes2023}:
 
@@
 
-$\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$ of objects
-of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family of
-projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
+A category $\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$
+of objects of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family
+of projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
 and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
@@
 
-If $\cat{C}$ has all small products and also biproducts, then the internal hom $\internalHom{(X, \partial
-X)}{(Y, \partial Y)}$ in $\Fam(\cat{C})$ is the object $(Z, \partial Z)$, where:
+Abusing notation somewhat, we shall write $X$ for an object $(X, \partial X)$ in $\Fam(C)$, and similarly for
+morphisms $f = (f, \partial f)$. If $\cat{C}$ has all small products and also biproducts, then the internal
+hom $\internalHom{X}{Y}$ in $\Fam(\cat{C})$ is the object $Z$ where:
 \begin{itemize}
-\item the indexing set $Z$ is the $\Fam(\cat{C})((X, \partial X), (Y, \partial Y))$, which is indeed a set by
-\propref{Grothendieck:fam-inherits};
-\item the indexed family $\partial Z$ is the family of objects $\prod_{X}\reindex{\partial Y}{f}$ for every
-${(f, \partial f): (X, \partial X) \to (Y, \partial Y)}$.
+\item the indexing set $Z$ is the $\Fam(\cat{C})(X, Y)$, which is indeed a set by
+\propref{Grothendieck:fam-inherits-local-smallness};
+\item the indexed family $\partial Z$ is the family of products $\prod_{X}\reindex{\partial Y}{f}$ for every
+${f: X \to Y}$.
 \end{itemize}
 
-\todo{Explain role of biproducts}
+The products in $\Fam(C)$ arise from the products in $\cat{C}$ (\propref{Grothendieck:fam-inherits-products}).
+The evaluation morphism $\eval_{X,Y}: (\internalHom{X}{Y}) \times X \to Y$ is a family of morphisms in
+$\cat{C}$ given by:
+\begin{align*}
+\eval_{X,Y}(f,x) &= f(x) \\
+\partial\eval_{X,Y}(f,x) &= \coprodM{\pi_x}{\partial f(x)}
+  : \textstyle\prod_{X}\reindex{\partial Y}{f} \biprod \partial X(x) \to \partial Y(f(x))
+\end{align*}
+\noindent using the coproduct of morphisms from the biproduct in $\cat{C}$.
+
+The currying isomorphism $\lambda_{X,Y,Z}: \Hom{\Fam(C)}{X \times Y}{Z} \to
+\Hom{\Fam(C)}{X}{\internalHom{Y}{Z}}$ natural in $X, Y, Z$ is given by:
+
+\vspace{-3mm}
+\begin{align*}
+\lambda_{X,Y,Z}(f)(x) &= g \quad\text{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
+\partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
+\end{align*}
+\noindent Here $(x,-): Y \to X \times Y$ is the ``reindexing'' function that pairs any $y$ with a fixed $x$,
+and for any $y$ $\biinj_{\partial X(x)}$ denotes the canonical injection $\partial X(x) \to \partial X(x)
+\biprod \partial Y(y)$.
 
@@
 \begin{enumerate}
-  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products (\todo{restrict $\LatGal$ to finite lattices?})
+  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products
   \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit

### 2025-05-26 main/category-of-families.tex
Note on reindexing functor for f, which is just the image of f in the indexed category [-,C].

@@
 
-The Grothendieck construction $\Grothendieck{}T$~(\defref{Grothendieck} below) for an indexed category $T$,
-sometimes called the \emph{total category} for $T$, is the category of all set-indexed families of objects of
-$T$, together with morphisms that account for both structure within each family and how families transform
-under reindexing functions $f: X \to Y$. First we formalise $X$-indexed families and their categories.
+The general form of the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
+\cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, assembles all the categories $T(X)$
+(for objects $X$ in $\cat{C}$) into a single category, together with morphisms that account for both internal
+structure and reindexing along morphisms $f: X \to Y$ in $\cat{C}$.
+
+\begin{definition}
+\label{def:Grothendieck}
+Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
+category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
+\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
+\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
+morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
+\end{definition}
+
+\noindent The $\partial X$ notation we adopt here is intended to suggest the connection between the
+Grothendieck construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$
+components compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g):
+(Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
+
+\begin{itemize}
+\item $\partial f: \partial X \to T(f)(\partial Y)$, and
+\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$
+\end{itemize}
+
+\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
+composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
+\comp f, T(f)(\partial g) \comp \partial f)$.
+
+\subsection{Category of Families}
+\label{sec:Fam}
+
+Now we formalise $X$-indexed families as functor categories $\Func{X}{\cat{C}}$ and consider the specific case
+of the Grothendieck construction for set-indexed categories $\Func{-}{\cat{C}}: \Set^\op \to \Cat$.
 
@@
 family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
-$\reindex{\eta}{f}(x) = \eta(f(x))$.
+$\reindex{\eta}{f}(x) = \eta(f(x))$. \todo{Should be clearer that $\reindex{-}{f}$ is actually
+$\Func{f}{\cat{C}}$.}
 \end{definition}
@@
 
-\begin{definition}
-\label{def:Grothendieck}
-\todo{Move earlier because doesn't assume $\cat{C} = \Set$.} Suppose an indexed category $T: \cat{C}^\op \to
-\Cat$. The \emph{Grothendieck construction} for $T$ is the category $\Grothendieck{X}T(X)$ (also written
-$\Grothendieck{}T$) which has as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an
-object $\partial X$ of $T(X)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial
-f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in
-$T(X)$.
-\end{definition}
-
-\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
-construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
-compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
-\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
-
-\begin{itemize}
-\item $\partial f: \partial X \to T(f)(\partial Y)$, and
-\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$
-\end{itemize}
-
-\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
-composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
-\comp f, T(f)(\partial g) \comp \partial f)$.
-
-\subsection{Category of Families}
-\label{sec:Fam}
-
-Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
-category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
-$f: X \to Y$ to the reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .
+The Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed category which sends
+any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the
+reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$, is called the \emph{families}
+construction for $\cat{C}$.
 

### 2025-05-26 main/category-of-families.tex
Minor tweaks.

@@
 \times \RR^n \to \RR^n$ (overloading $+$ and $\mult$), where $(\RR^n,+)$ is an abelian group (with identity
-$0$ and additive inverse $-v$ again defined component-wise) and the vector operations are compatible with the
-field operations in that the following equations hold:
+$0$ and additive inverse $-v$ again defined component-wise), and where the vector operations are compatible
+with the field operations in that the following equations hold:
 
@@
 morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
-\mult f(v)$.
+\mult f(v)$, with $f$ preserving the abelian group structure automatically.
 \end{definition}
@@
 product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
-\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$.
+\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$. \todo{Significance of biproducts}
 
@@
 
-Derivatives respect composition~\cite{spivak65}. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For
-any $x \in \RR^m$ we have:
+Derivatives respect composition~\cite{spivak65}. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. Then
+for any $x \in \RR^m$ we have:
 
@@
 Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
-f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ selecting the appropriate
-pushforward by recomputing $f(x)$.
+\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
+\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ recomputing $f(x)$ in
+order to select the appropriate pushforward.
 

### 2025-05-27 main/formal-structures.tex
Correction.
Tweaks.

@@
 
-Our approach to a denotational account of \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
+Our denotational approach to \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
 Before we go into details, we present a high-level overview of their approach and how we have applied it to \GPS.
@@
 However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
-Neither of these are ideal.
+Neither of these is ideal.
 In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost.
@@
 Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
-We overcome this by using $\CMon$-enriched presheaves; we turn to this in \secref{higher-order} but first flesh out the first-order setting.
+We overcome this by using $\CMon$-enriched presheaves, as described in \secref{higher-order}; first we present a suitable first-order setting.
 
@@
 
-First we introduce the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward \GPS. Galois connections are pairs of monotone functions $f: Y \to
-X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
-of $g$, and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the
-bidirectional nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the
-least input slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the
-greatest output slice for a given slice of the input).
+We start with the category of lattices and Galois connections, which will serve as our base category for
+interpreting forward and backward \GPS. Galois connections are pairs of monotone functions $f: Y \to X$ and
+$g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
+and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the bidirectional
+nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the least input
+slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the greatest
+output slice for a given slice of the input).
 
@@
 approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to
-combine approximations. Thus rather than working directly in $\PosGal$, we consider the following subcategory
+combine approximations. Thus rather than working directly in $\PosGal$, we will use the following subcategory
 instead.
@@
 The meet-preserving maps from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$ on such
-maps is defined pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves meets; the
+maps is given pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves meets; the
 constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets, provides the
 unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by
-$\join$ and the constant homomorphism $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two
+$\join$ and the constant map $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two
 constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit

### 2025-05-27 notes/stability.tex

--- 2025-05-24 notes/stability.tex
@@
 \label{sec:stable-functions}
+
+\newcommand{\downset}[1]{\mathop{\downarrow}(#1)}
+
+Central to the concept of Galois slicing is the idea that
+approximations of a value are a lattice, and that the functions
+transporting approximations forwards and backwards form a Galois
+connection between those lattices. We present a motivation for this
+theory starting from the idea of qualitative information ordering from
+Domain Theory. We argue that, with reference to differentiability,
+monotone functions between certain partially ordered sets are
+analogous to differentiable functions, and \emph{stable} functions
+between those sets are analogous to functions with a reverse
+derivative.
+
+The stable function approach works, but mixes values and their
+approximations into a single set. We use the families construction
+over the category of lattices and Galois connections to separate
+approximation from computation, and to provide a stepping stone to
+executable models of Galois slicing.
+
+\subsubsection{Domain Theory}
+
+Domain theory is a method for giving denotational semantics to
+programs that handle infinite data (i.e., functions,
+streams). \emph{Domains} are special partially ordered sets where the
+ordering denotes a relationship of qualitative information content: if
+$x \sqsubseteq y$, then $y$ may contain more information than $x$. For
+example, if $x$ and $y$ are functions, then $y$ may be defined at more
+places in its domain than $x$.
+
+Domains are typically closed under least upper bounds (\emph{lubs}) of
+directed sets, meaning that any collection of elements that is
+internally consistent has a ``completion'' that contains all the
+information covered by the set. These directed lubs are used to give
+semantics to recursive programs in terms of their finite
+approximations.
+
+Programs themselves are interpreted as monotone functions that
+preserve directed lubs. The latter property is called
+continuity. Monoticity captures the idea that if the input gets more
+defined, then the output can get more defined. Continuity intuitively
+expresses the idea that a function interpreting a program cannot
+``look at all the approximations'' when determining the output, which
+corresponds to the intuitive idea that a computable function cannot
+look at a non-finite amount of input to produce an output.
+
+For the purposes of Galois slicing, we are not interested in using
+approximations to model computation on infinite objects, but rather to
+use them for the related task of revealing how programs explore their
+inputs for provenance tracking. Therefore, we start by only
+considering certain partially ordered sets and monotone functions
+between them.
+
+\subsubsection{Forwards Slicing}
+
+We now set up an analogy between certain partial orders with monotone
+functions and manifolds and smooth functions.
+
+In a partially ordered set $X$, for any $x \in X$ the set of elements
+below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a
+partially ordered set. These approximations of $x$ we will think of as
+``derivatives'' of $x$, and the whole set $\downset{x}$ as the
+``tangent space''. Tangent spaces are vector spaces, since they are
+linear approximations to curves at that point. In the partially
+ordered setting, we take elements of $\downset{x}$ to be
+approximations of processes defined at $x$. As we can add tangents, we
+assume we can take meets of approximations in $\downset{x}$:
+
+\begin{definition}
+  A \emph{meet approximation space} is a partially ordered set $X$
+  where every for every $x \in X$, $\downset{x}$ is a meet
+  semilattice, with $x$ as the top element.
+\end{definition}
+
+Smooth functions have a ``pushforward'' derivative that takes tangents
+at a point $x$ to tangents at the point $f(x)$. In the partially
+ordered setting, we require that the restriction of a monotone
+function $f$ to each $\downset{x}$ preserves meets:
+
+\begin{definition}
+  A \emph{morphism of meet approximation spaces} $f : X \to Y$ is a
+  monotone function such that for all $x \in X$, the restriction
+  $f_x : \downset{x} \to \downset{f(x)}$ preserves meets.
+\end{definition}
+
+\begin{lemma}
+  Morphisms of meet approximation spaces are closed under identities
+  and composition, forming a category. FIXME: with products?
+  functions? sums?
+\end{lemma}
+
+\begin{example}
+  FIXME: various forms of OR:
+  \begin{itemize}
+  \item strict and left-short-circuiting
+  \item parallel or
+  \item gustave
+  \end{itemize}
+\end{example}
+
+\begin{example}
+  An example that preserves meets, but isn't stable (from Amadio and
+  Curien just before Lemma 12.2.3, originally due to Berry): Define
+  $f : D \to \{\bot, \top\}$, where:
+  \begin{displaymath}
+    D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots \sqsubseteq 1 \sqsubseteq 0
+  \end{displaymath}
+  as $f(\bot) = \bot$ and $f(n) = \top$. This is monotone, and
+  preserves meets in every $\downset{x}$. However, there is no
+  ``best'' (i.e., least) input that gives us any finite output. To
+  rectify this, we will further assume that our functions are stable.
+\end{example}
+
+\subsubsection{Backwards Slicing}
 
@@
 
-Stability was invented by \citet{berry79} in an attempt to capture
-sequentiality, and was also instrumental in the discovery of Linear
+Stability was invented by Berry \cite{berry86} in an attempt to capture
+sequentiality, and were also instrumental in the discovery of Linear
 Logic via Coherence Spaces \cite{girard}. A textbook description of
@@
 
-\newcommand{\downset}[1]{\mathop{\downarrow}(#1)}
-
 \begin{definition}

### 2025-05-27 main/category-of-families.tex

@@
 
-The general form of the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
+In its general form, the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
 \cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, assembles all the categories $T(X)$
@@
 
-\noindent The $\partial X$ notation we adopt here is intended to suggest the connection between the
-Grothendieck construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$
-components compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g):
-(Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
+\noindent The $\partial X$ notation used here is intended to suggest the connection between the Grothendieck
+construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
+compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
+\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
 
@@
 
-Although reindexing is just precomposition with $f$, writing it as $\reindex{-}{f}$ avoids notational
-confusion when combining composition of morphisms in $\Func{X}{C}$ with reindexing.
+\noindent Although reindexing is just precomposition with $f$, writing it as $\reindex{-}{f}$ avoids
+notational confusion when combining composition of morphisms in $\Func{X}{C}$ with reindexing.
 
 The Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed category which sends
-any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the
+any set $X$ to the category of indexed families $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the
 reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$, is called the \emph{families}
@@
 
-The total category $\Fam(\LatGal)$ is then basic setting for interpreting first-order programs for \GPS. This
-has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
+The total category $\Fam(\LatGal)$ is then a suitable setting for interpreting first-order programs for \GPS.
+It has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
 $\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of a

### 2025-05-27 main/formal-structures.tex

@@
 
-$\LatGal$ is not Cartesian closed, and so cannot serve directly as the basis for interpreting higher-order
-programs. In fact Cartesian closure is incompatible with another notable property of $\LatGal$. The
-projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the product of
-lattices, have both upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as
-$X \biprod Y$, is both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and
-injections $\biinj_X$ and $\biinj_Y$ given by:
+$\LatGal$ is not Cartesian closed, and so not directly suitable for interpreting higher-order programs. In
+fact Cartesian closure is incompatible with another important property of $\LatGal$. The projections $\pi_1: X
+\times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the product of lattices, have both
+upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is
+both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and
+$\biinj_Y$ given by:
 
@@
 renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and $\cat{C}$ has only a single object, up to isomorphism. However although biproducts
-rule out Cartesian closure, they will prove crucial for establishing a Cartesian closed setting in
-\secref{higher-order}; but let us first see how to develop a suitable categorical setting for first-order
-programs.
+\Rightarrow Y} \iso 1$, and so $\cat{C}$ has only a single object, up to isomorphism. \todo{Bridge better to
+next section; even if $\LatGal$ were Cartesian closed, directly interpreting into $\LatGal$ wouldn't allow
+each point to be associated with its own lattice of approximants.}

### 2025-05-27 main/introduction.tex

@@
 
-Previous work on \GPS uses an operational semantics of programs that generates a trace of each execution which can be rewound or replayed after the fact, to compute the Galois connection described above.
-It would be satisfying to have a denotational account of \GPS as well, for which traces would then offer a plausible implementation technique (perhaps one of several).
-As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18}, which points to a way to develop such an approach.
+Previous work on \GPS used a special operational semantics to generate a trace of each execution, and then separately used that trace to compute the Galois connection described above, by re-running forwards or backwards over the trace.
+It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately (but somewhat redundantly) defined ``backwards evaluation'' operation.
+As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vákár22}, which points to a way to develop such an approach.
 We have already hinted at this in the description above, but let us now make it explicit.
@@
 \begin{enumerate}
-\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS,
+\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS.
 \item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.

### 2025-05-30 main/approx-as-tangents.tex
[ tex ] rough attempt to develop the stablity<->differentiable analogy

\section{Approximations as Tangents}

We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a dentational setting. From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.

\subsection{Manifolds, Smooth Functions, and Automatic Differentiation}

For what follows we do not need to know in detail what a manifold is, or exactly how smooth functions are defined. We only state some definitions in order to fix terminology and to justify our claim of a connection to standard differentiablity. \bob{Need a good reference here.}

The general study of differentiable functions usually takes place on \emph{manifolds}, which are topological spaces that are ``locally'' like a Euclidean space $\RR^n$. The spaces $\RR^n$ are manifolds, but so are ``non-flat'' examples such as $n$-spheres and more exotic spaces. Every point $x$ in a manifold $M$ has an associated \emph{tangent vector space} $\tangents_x(M)$ and \emph{cotangent vector space} $\cotangents_x(M)$, the latter consisting of linear functions $\tangents_x(M) \linearto \RR$. The tangent and cotangent spaces are finite dimensional, so in the presence of a chosen basis they are canonically isomorphic. In the case when the manifold is $\RR^n$, then every tangent space is isomorphic to $\RR^n$ as well.

Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$. Each smooth function induces maps of the (co)tangent spaces:
\begin{itemize}
\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$. In the Eucdlidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
\item The \emph{backward derivative} (cotangent map, or pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$. In the Euclidean case, the backward derivative is represented by the transpose of the Jacobian of $f$ at $x$.
\end{itemize}

Computing the forward and backward derivatives of smooth functions $f$ has many applications of practical interest. For example, computation of the reverse derivative is of central interest in Machine Learning by Gradient Descent, the main technique used to train Deep Neural Networks. \bob{need a citation here}.

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output. \bob{cite something on autodiff}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms. \bob{cite: JAX, TensorFlow, Dex, Jesse Sigal, etc.}

\subsection{Stable Domain Theory as Differentiability}

Our thesis is that \GPS can profitably be seen as a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory. Smooth functions in this setting are Berry's \emph{stable functions} in domain theory. We now introduce these concepts and how they relate to \GPS.

\subsubsection{Domains as a Qualitative Theory of Approximation}

Domain theory is a method for defining the semantics of programs that
handle infinite data (e.g., functions or infinite
streams). \emph{Domains} are special partially ordered sets where the
ordering denotes a relationship of qualitative information content: if
$x \sqsubseteq y$, then $y$ may contain more information than $x$. For
example, if $x$ and $y$ are functions, then $y$ may be defined at more
places in its domain than $x$. Infinite objects are understood in
terms of their approximations in this sense, and domains are assumed
to be closed under least upper bounds (lubs) of directed sets, meaning
that any internally consistent collection of elements has a
``completion'' that contains all the information covered by the
set. Programs are interpreted as monotone functions that preserve
directed lubs. The latter property is called continuity. Monoticity
captures the idea that if the input gets more defined, then the output
can get more defined. Continuity intuitively expresses the idea that a
function interpreting a program cannot ``look at all the
approximations'' when determining the output, which corresponds to the
intuitive idea that a computable function cannot look at a non-finite
amount of input to produce an output.

For the purposes of \GPS, we are not interested in using
approximations to model computation on infinite objects, but rather to
use them for the related task of revealing how programs explore their
inputs for provenance tracking. Therefore, in what follows we dispense with completeness properties of the partially ordered sets we consider.

\subsubsection{Stable Functions as ``Smooth Functions''}

Continuous functions between domains capture the intuitive idea that
computable functions are restricted to finite observations of their
input for finite output. However, there are examples, such as
Plotkin's Parallel OR (\autoref{ex:parallel-or}, below) that are
continuous but do not treat their input in a way that is consistent
with a sequential implementation. Stability is an additional property
of monotone functions that was invented by Berry \cite{berry86} in an
attempt to capture sequentiality. This was not successful (see the
$\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will
see now how it is closely related to problem of computing forward and
backwards maps of approximations needed in \GPS. A textbook
description of stable functions in the context of domain theory is
given by \citet{amadio-curien} (Chapter 12). We will
refer to this where necessary.

In a partially ordered set $X$, for any $x \in X$ the set of elements
below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a
partially ordered set. These approximations of $x$ we will think of as
``derivatives'' of $x$, and the whole set $\downset{x}$ as the
``tangent space''. Tangent spaces are vector spaces, since they are
linear approximations to curves at that point. In the partially
ordered setting, we take elements of $\downset{x}$ to be
approximations of processes defined at $x$. As we can add tangents, we
assume we can take meets of approximations in $\downset{x}$:

\begin{definition}
  A \emph{bounded meet poset} is a partially ordered set $X$ where
  every for every $x \in X$, $\downset{x}$ is a meet semilattice, with
  $x$ as the top element.
\end{definition}

Smooth functions have a ``pushforward'' derivative that takes tangents
at a point $x$ to tangents at the point $f(x)$. In the partially
ordered setting, we require that the restriction of a monotone
function $f$ to each $\downset{x}$ preserves meets:

\begin{definition}
  A \emph{conditionally multiplicative} (cm) function $f : X \to Y$ is
  a monotone function such that for all $x \in X$, the restriction
  $f_x : \downset{x} \to \downset{f(x)}$ preserves meets.
\end{definition}

\begin{lemma}
  Morphisms of meet approximation spaces are closed under identities
  and composition, forming a category. FIXME: with products?
  functions? sums?
\end{lemma}

\begin{example}[Conditionally Multiplicative Functions]
  To see the effect of requiring preservation of bounded meets,
  consider several ways of defining disjunction on the lifted booleans
  $\mathbb{B}_\bot$. Two functions that are cm are the strict and
  left-short-circuiting ORs\footnote{The clauses in these examples are
    shorthand for the graph of the function. They are not to be
    understood as pattern matching clauses in a language like Haskell,
    where it is not possible to match on $\bot$.}:
  \begin{displaymath}
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{strictOr}&\mathsf{tt}&\mathsf{tt}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{tt}&\mathsf{ff}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{ff}&\mathsf{tt}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
      \mathrm{strictOr}&\bot&\_&=&\bot \\
      \mathrm{strictOr}&\_&\bot&=&\bot \\
    \end{array}
    \qquad
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{shortCircuitOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
      \mathsf{shortCircuitOR}&\mathsf{ff}&x&=&x \\
      \mathsf{shortCircuitOR}&\bot&\_& =& \bot
    \end{array}
  \end{displaymath}
  \bob{Demonstrate that they are conditionally multiplicative, by
    giving evidence of stability and stating that this implies cm}

  The function $\mathrm{strictOr}$ is stable. For example, for the
  input-output pair $(\mathsf{tt},\mathsf{ff}), \mathsf{tt}$, the
  minimal input that gives this output is exactly
  $(\mathsf{tt}, \mathsf{ff})$. If we take the approximation
  $\bot \leq \mathsf{tt}$ of the output, then the corresponding
  minimal input is $(\bot, \bot)$. The function
  $\mathrm{shortCircuitOR}$ is also stable. For the input-output pair
  $(\mathsf{tt},\mathsf{ff}),\mathsf{tt}$, the minimal input that
  gives this input is $(\mathsf{tt},\bot)$, indicating that the
  presence of $\mathsf{ff}$ in the second argument was not necessary
  to produce this output. As with $\mathrm{strictOr}$, the minimal
  input required to produce the output $\bot \leq \mathsf{tt}$ is
  again $(\bot,\bot)$.

  The fact that these two functions's stability witnesses reveal
  information about how they depend on their input is what we will
  exploit in order to use the idea of stability for slicing.
\end{example}

\begin{example}[A non-Conditionally Multiplicative Function]
  \label{ex:parallel-or}
  A function that is not cm is Plotkin's Parallel OR \cite{lcf77},
  which short-circuits in both arguments, returning $\mathsf{tt}$ if
  either argment is $\mathsf{tt}$, even if the other argument is not
  defined:
  \begin{displaymath}
    \begin{array}{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{parallelOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
      \mathrm{parallelOR}&\_&\mathsf{tt}&=&\mathsf{tt} \\
      \mathsf{parallelOR}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
      \mathsf{parallelOR}&\bot&\bot&=&\bot
    \end{array}
  \end{displaymath}
  For the input-output pair $(\mathsf{tt},\mathsf{tt}),\mathsf{tt}$,
  there is no one minimal input that produces this output. We have
  both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and
  $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are
  incomparable and their greatest lower bound $(\bot,\bot)$ gives the
  output $\bot$.

  Parallel OR is famous because it is not \emph{sequential}, meaning
  intuitively that it cannot be implemented without running the two
  arguments in parallel to see if one of them returns
  $\mathsf{tt}$. That fact that it exists in the standard domain
  theoretic semantics of PCF means that this semantics is complete for
  reasoning about observational equivalence in PCF. Since Parallel OR
  is not stable, one might hope that stability is enough to capture
  sequentiality, and hence potentially give a fully abstract model of
  PCF. However, the following ternary function
  $\mathbb{B}_\bot^3 \to \{\top,\bot\}$ is stable but admits no
  sequential implementation that fixes an order that the arguments are
  examined in:
  \begin{displaymath}
    \begin{array}{l@{(}l@{,~}l@{,~}l@{)~}c@{~}l}
      \mathrm{gustave}&\mathsf{tt}&\mathsf{ff}&\_&=&\top \\
      \mathrm{gustave}&\mathsf{ff}&\_&\mathsf{tt}&=&\top \\
      \mathrm{gustave}&\_&\mathsf{tt}&\mathsf{ff}&=&\top \\
      \mathrm{gustave}&\_&\_&\_&=&\bot
    \end{array}
  \end{displaymath}
  Despite there being no one minimal input that achieves the output
  $\top$, each of the minimal inputs that can achieve this output are
  pairwise incomparable, so for each specific input that gets output
  $\top$ there is a unique minimal input that achieves it (listed in
  the first three lines of the definition).

  In terms of \GPS the $\mathrm{gustave}$ function does not present a
  problem. For any particular run (i.e., input-output pair) of the
  program, there is an unambiguous minimal input that achieves the
  output, no matter that it was not achieved by a sequential
  processing of the input.
\end{example}

\begin{example}[Is Conditional Multiplicativity Enough?]
  \label{ex:non-stable-function}
  In the preceeding examples, we demonstrated conditional
  multiplicativity by giving a backwards mapping of approximations
  that forces preservation of meets. A natural question is whether or
  not we always have such a backwards map which we could use for
  backwards propagation of approximations. An example that is
  conditionally multiplicative, but does not admit such a backards map
  (from Amadio and Curien just before Lemma 12.2.3, originally due to
  Berry) is given by defining
  $\mathrm{unstable} : D \to \{\bot, \top\}$, where:
  \begin{displaymath}
    D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots \sqsubseteq 1 \sqsubseteq 0
  \end{displaymath}
  as $\mathrm{unstable}(\bot) = \bot$ and
  $\mathrm{unstable}(n) = \top$. This is monotone, and preserves meets
  in every $\downset{x}$. However, there is no ``best'' (i.e., least)
  input that gives us any finite output.
\end{example}

In light of the preceeding example, we can give an alternative
definition of a function that requires the existence of a reverse
mapping directly, without assuming that any meets exist:

\begin{definition}[Stable function]
  Let $f : X \to Y$ be a monotone function between posets $X$ and
  $Y$. The function $f$ is \emph{stable} if for all $x \in X$ and
  $y \leq f(x)$:
  \begin{enumerate}
  \item (\textsc{Existence}) there exists an $x_0 \leq x$ such that $y \leq f(x_0)$, and
  \item (\textsc{Minimality}) for any $x'_0 \leq x$ such that $y \leq f(x'_0)$ then
    $x_0 \leq x'_0$ .
  \end{enumerate}
\end{definition}

\begin{lemma}
  Stable functions are closed under identities and composition.
\end{lemma}

\begin{example}
  \bob{Elaborate on the examples}
  \begin{enumerate}
  \item $\mathrm{strictOr}$ and $\mathrm{shortCircuitOR}$ are stable, as demonstrated above.
  \item $\mathrm{parallelOR}$ is not stable.
  \item $\mathrm{gustave}$ is stable.
  \item $\mathrm{unstable}$ is not stable, but is conditionally multiplicative.
  \end{enumerate}
\end{example}

Stability has an alternative definition in terms of Galois connections, which will be more useful for what follows. This characterisation is due to Paul Taylor\bob{cite}.

\begin{lemma}
  A monotone function $f : X \to Y$ is stable if and only if for all
  $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$
  has a left Galois adjoint.
\end{lemma}

\begin{proof}
  If $f$ is stable, then define a left adjoint
  $f^*_x : \downset{f(x)} \to \downset{x}$ by setting $f^*_x(y)$ to be
  the minimal $x_0$ required by stability. This is monotone: if
  $y \leq y'$, then we know that $y \leq y' \leq f(f^*_x(y'))$ by
  definition of $f^*_x$, so $f^*_x(y) \leq f^*_x(y')$ by minimality of
  $f^*_x(y)$. For the adjointness, let $x' \leq x$ and $y \leq
  f(x)$. Then if $f^*_x(y) \leq x'$, we have
  $y \leq f(f^*_x(y)) \leq f(x')$ by monotonicity of $f$ and the first
  part of stability. In the other direction, if we have
  $y \leq f(x')$, then by uniqueness we have $f^*_x(y) \leq x'$.

  If, for every $x$, $f_x$ has a left adjoint $f^*_x$, then for any
  $x', y$ we have $y \leq f_x(x') \Leftrightarrow f^*_x(y) \leq
  x'$. So $f^*_x(y)$ is the element that satisfies
  $y \leq f(f^*_x(y))$, and it is minimal since if $y \leq f_x(x'_0)$
  then $f^*_x(y) \leq x'_0$.
\end{proof}

Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets (for forward approximation maps) and joins (for backwards approximation maps) exist:

\begin{definition}
  An \emph{L-poset} is a partially ordered set $X$ such that for every
  $x \in X$, the principal downset $\downset{x}$ is a bounded lattice
  (i.e., have all finite meets and joins).
\end{definition}

The following lemma is an instance of standard facts about Galois connections preserving meets and joins:

\begin{lemma}
  For L-posets $X$ and $Y$, a stable function $f : X \to Y$ preserves
  meets in its forward part $f_x$ and joins in its reverse part
  $f^*_x$.
\end{lemma}

The converse to this lemma (that functions that preserve meets in
their forward part have a left Galois adjoint) is not true, as was
demonstrated by the non-stable function in
\autoref{ex:non-stable-function}. In the case when the posets
$\downset{x}$ are \emph{complete}, and $f_x$ preserves infinitary
meets, then we are guaranteed a left Galois adjoint. In that example,
the infinite set $\{0, 1, 2, \dots, n, n-1, \dots\}$ not including
$\bot$ of approximations of $0$ does not have a greatest lower bound,
so the order is not complete.

We could have required that L-posets' principal downsets were complete
lattices, which would mean that we would only need to require that
forward approximation maps preserve all meets and we could compute the
backwards map automatically. However, our goal is a computable theory
for \GPS, analogous to automatic differentiation. Requiring
completeness to compute backwards maps of approximations would require
that complete joins are always computable, which will not be the case for approximations of functions.

\subsection{Summary}
\label{sec:diff-stab-summary}

We now summarise the analogy between differentiable functions between manifolds and stable functions between L-posets:

\begin{itemize}
\item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
\item For \GPS, every program has an associated forward approximation map that takes approximations forward from the input to the output. This map {\em preserves meets}. For differentiable programs, every program has a forward derivative that takes tangents of the input to tangents of the output. The forward derivative map is {\em linear}, so it preserves addition of tangents and the zero tangent.
\item For \GPS, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input. This map {\em preserves joins}. For differentiable programs, every program has a reverse derivative that takes tangents of the output to tangents of the input. This map is again {\em linear}.
\item For \GPS, the forward and backward approximation maps are related by being a Galois connection. For differentiable programming, the forward and reverse derivatives are related by being each others' transpose.
\end{itemize}

Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS. This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wide range of computational settings.

### 2025-05-31 main/higher-order.tex
Note on reindexing functor for f, which is just the image of f in the indexed category [-,C].
Start on typing rules.
WIP.
Some citations on categories for types.
Defining mu-polynomial functors.

@@
 
-\vspace{-3mm}
+\vspace{-4mm}
 \begin{align*}
-\lambda_{X,Y,Z}(f)(x) &= g \quad\text{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
+\lambda_{X,Y,Z}(f)(x) &= g \quad\textit{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
 \partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
@@
   \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
-  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how we will interpret base types?
-E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a suitable object 2}
+  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how
+  we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a
+  suitable object 2?}
   \end{enumerate}
@@
 \end{proposition}
+
+\subsection{Higher-order language}
+\todo{Introduce earlier?}
+
+We introduce a standard total functional languages with a (reasonably) expressive type
+system~\cite{crole94,pitts01,santocanale02}.
+
+\subsubsection{Syntax}
+
+\input{fig/mu-polynomial}
+
+\begin{definition}[$\mu$-polynomial]
+Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly$
+to be the smallest subcategory of $\Cat$ closed under the rules in \figref{mu-polynomial}.
+\end{definition}
+
+\input{fig/syntax}
+
+\figrefTwo{syntax}{typing} give the syntax and typing rules of the higher-order language.
+
+\subsubsection{Semantics}
+
+\input{fig/typing}

### 2025-06-01 main/higher-order.tex
Initial pass over mu-polynomial functors.
Explicitly define Pos/Neg.
First pass over semantics of well-kinded types.

@@
 
-\input{fig/mu-polynomial}
-
-\begin{definition}[$\mu$-polynomial]
-Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly$
-to be the smallest subcategory of $\Cat$ closed under the rules in \figref{mu-polynomial}.
-\end{definition}
-
+\input{fig/positivity}
 \input{fig/syntax}
+\input{fig/typing}
 
@@
 
-\input{fig/typing}
+\input{fig/mu-polynomial}
+\input{fig/set-semantics}
+
+\begin{definition}[$\mu$-polynomial]
+Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly_{\cat{C}}$ to be the
+smallest subcategory of $\Cat$ generated inductively by the rules in \figref{mu-polynomial}.
+\end{definition}
+
+\figref{set-semantics:types} gives the interpretation of a type $\tau$ with a single free type variable as an
+endofunctor on $\Set$.

### 2025-06-02 main/higher-order.tex
Moving from Set semantics to arbitrary C..

@@
 \input{fig/mu-polynomial}
-\input{fig/set-semantics}
+\input{fig/default-semantics}
 
@@
 
-\figref{set-semantics:types} gives the interpretation of a type $\tau$ with a single free type variable as an
-endofunctor on $\Set$.
+Suppose $\cat{C}$ a Cartesian closed category with finite coproducts, a strong monad $\mathcal{L}$ and
+$\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$. \figref{default-semantics:types} gives the
+interpretation of a type $\tau$ with a single free type variable as a $\mu$-polynomial endofunctor on
+$\cat{C}$.

### 2025-06-02 main/higher-order.tex
Notation for strength.
Progress on bind/return.
Var case.

@@
 
-Suppose $\cat{C}$ a Cartesian closed category with finite coproducts, a strong monad $\mathcal{L}$ and
-$\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$. \figref{default-semantics:types} gives the
-interpretation of a type $\tau$ with a single free type variable as a $\mu$-polynomial endofunctor on
-$\cat{C}$.
+Suppose $\cat{C}$ a Cartesian closed category with exponential $\internalHom{X}{Y}$, evaluation morphisms
+$\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
+\begin{enumerate}
+\item finite coproducts $(\coprod, \Zero)$;
+\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
+\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
+\mathcal{L}(X \times Y)$;
+\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$.
+\end{enumerate}
+\figref{default-semantics:types} gives the interpretation of a type $\tau$ with a single free type variable as
+a $\mu$-polynomial endofunctor on $\cat{C}$.

### 2025-06-03 main/higher-order.tex
First pass over type interpretation.

@@
 
-Suppose $\cat{C}$ a Cartesian closed category with exponential $\internalHom{X}{Y}$, evaluation morphisms
+Suppose $\cat{C}$ a Cartesian closed category with exponentials $\internalHom{X}{Y}$, evaluation morphisms
 $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
@@
 \end{enumerate}
-\figref{default-semantics:types} gives the interpretation of a type $\tau$ with a single free type variable as
-a $\mu$-polynomial endofunctor on $\cat{C}$.
+\figref{default-semantics:types} gives the interpretation of a type $\Delta \vdash \tau$ with $\length{\Delta}
+= n$ as a multifunctor $\cat{C}^n \to \cat{C}$ (as an object of $\cat{C}$ in the case $n = 0$).

### 2025-06-03 main/higher-order.tex
Notation for structure map and catamorphism for initial F-alg.
C bicartesian closed.

--- 2025-06-02 main/higher-order.tex
@@
 
-We introduce a standard total functional languages with a (reasonably) expressive type
-system~\cite{crole94,pitts01,santocanale02}.
+We introduce a standard total functional language with a (reasonably) expressive type
+system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types and a set $\PrimOp$
 
@@
 
-Suppose $\cat{C}$ a Cartesian closed category with exponential $\internalHom{X}{Y}$, evaluation morphisms
-$\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
+For any $\mu$-polynomial endofunctor $F: \cat{C} \to \cat{C}$ with initial algebra $\mu F$, write $\inMap_F$
+for the structure map $F\mu F \to \mu F$ and write $\cata_f$ for the unique morphism $\mu F \to X$ from the
+initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
+
+For a ``default'' (baseline) semantics, fix a bicartesian closed category $\cat{C}$ with finite products
+$(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
+morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
 \begin{enumerate}
-\item finite coproducts $(\coprod, \Zero)$;
 \item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
@@
 \end{enumerate}
-\figref{default-semantics:types} gives the interpretation of a type $\tau$ with a single free type variable as
-a $\mu$-polynomial endofunctor on $\cat{C}$.
+\figref{default-semantics:types} gives the default interpretation of types $\Delta \vdash \tau$ with
+$\length{\Delta} = n$ as multifunctors $\cat{C}^n \to \cat{C}$ (or as objects of $\cat{C}$ in the case $n =
+0$). \figref{default-semantics:terms} gives the default interpretation of terms $\Gamma \vdash t: \tau$ as
+morphisms $\sem{\Gamma} \to \sem{\tau}$.

### 2025-06-03 main/higher-order.tex
Syntax and typing for primitives.

@@
 We introduce a standard total functional language with a (reasonably) expressive type
-system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types and a set $\PrimOp$
+system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types $\rho$ and sets
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations.
 
@@
 $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
-morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
+morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the following additional structure:
 \begin{enumerate}
@@
 \mathcal{L}(X \times Y)$;
-\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$.
+\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$;
+\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$ of $\cat{C}$;
+\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\rho}:
+\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
 \end{enumerate}

### 2025-06-03 main/higher-order.tex
Higher-order language WIP; keep in notes for now.

--- 2025-06-03 main/higher-order.tex
@@
 \end{proposition}
-
-\subsection{Higher-order language}
-\todo{Introduce earlier?}
-
-We introduce a standard total functional languages with a (reasonably) expressive type
-system~\cite{crole94,pitts01,santocanale02}.
-
-\subsubsection{Syntax}
-
-\input{fig/positivity}
-\input{fig/syntax}
-\input{fig/typing}
-
-\figrefTwo{syntax}{typing} give the syntax and typing rules of the higher-order language.
-
-\subsubsection{Semantics}
-
-\input{fig/mu-polynomial}
-\input{fig/default-semantics}
-
-\begin{definition}[$\mu$-polynomial]
-Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly_{\cat{C}}$ to be the
-smallest subcategory of $\Cat$ generated inductively by the rules in \figref{mu-polynomial}.
-\end{definition}
-
-Suppose $\cat{C}$ a Cartesian closed category with exponentials $\internalHom{X}{Y}$, evaluation morphisms
-$\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, and the following additional structure:
-\begin{enumerate}
-\item finite coproducts $(\coprod, \Zero)$;
-\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
-\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
-\mathcal{L}(X \times Y)$;
-\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$.
-\end{enumerate}
-\figref{default-semantics:types} gives the interpretation of a type $\Delta \vdash \tau$ with $\length{\Delta}
-= n$ as a multifunctor $\cat{C}^n \to \cat{C}$ (as an object of $\cat{C}$ in the case $n = 0$).

### 2025-06-03 notes/language.tex
Separate doc for notes.
Higher-order language WIP; keep in notes for now.

--- 2025-04-10 appendix/notes/language.tex
@@
-\section{Language}
+\subsection{Higher-order language}
+\todo{Introduce earlier?}
 
-\subsection{Syntax}
+We introduce a standard total functional language with a (reasonably) expressive type
+system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types $\rho$ and sets
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations.
 
-We parameterise our language over a signature $\Sigma = (S, \phi, R)$ consisting of:
-\begin{itemize}
-\item a set $S$ of sorts
-\item a set of operations $\phi$ with signature $\Sigma(\phi) \in S^* \times S$
-\item a set of relations $R$ with signature $\Sigma(R) \in S^*$
-\end{itemize}
+\subsubsection{Syntax}
 
-\subsection{Semantic domain}
+\input{fig/positivity}
+\input{fig/syntax}
+\input{fig/typing}
 
-We interpret our language into any category with:
-\begin{itemize}
-\item finite products (binary products and a terminal object)
-\item Booleans
-\end{itemize}
-\note{Ignoring lists for now. Booleans can be derived from coproducts.}
+\figrefTwo{syntax}{typing} give the syntax and typing rules of the higher-order language.
+
+\subsubsection{Semantics}
+
+\input{fig/mu-polynomial}
+\input{fig/default-semantics}
+
+\begin{definition}[$\mu$-polynomial]
+Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly_{\cat{C}}$ to be the
+smallest subcategory of $\Cat$ generated inductively by the rules in \figref{mu-polynomial}.
+\end{definition}
+
+For any $\mu$-polynomial endofunctor $F: \cat{C} \to \cat{C}$ with initial algebra $\mu F$, write $\inMap_F$
+for the structure map $F\mu F \to \mu F$ and write $\cata_f$ for the unique morphism $\mu F \to X$ from the
+initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
+
+For a ``default'' (baseline) semantics, fix a bicartesian closed category $\cat{C}$ with finite products
+$(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
+morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the following additional structure:
+\begin{enumerate}
+\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
+\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
+\mathcal{L}(X \times Y)$;
+\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$;
+\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$ of $\cat{C}$;
+\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\rho}:
+\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
+\end{enumerate}
+\figref{default-semantics:types} gives the default interpretation of types $\Delta \vdash \tau$ with
+$\length{\Delta} = n$ as multifunctors $\cat{C}^n \to \cat{C}$ (or as objects of $\cat{C}$ in the case $n =
+0$). \figref{default-semantics:terms} gives the default interpretation of terms $\Gamma \vdash t: \tau$ as
+morphisms $\sem{\Gamma} \to \sem{\tau}$.

### 2025-06-03 notes/language.tex
Move to notes.

@@
 \subsection{Higher-order language}
-\todo{Introduce earlier?}
 
@@
 
-For a ``default'' (baseline) semantics, fix a bicartesian closed category $\cat{C}$ with finite products
-$(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
+For the sake of a ``default'' (baseline) semantics, fix a bicartesian closed category $\cat{C}$ with finite
+products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
 morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the following additional structure:

### 2025-06-03 notes/language.tex
Reference language sections.
Restrict to closed function types for now.

@@
 system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types $\rho$ and sets
-$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations.
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$.
 
 \subsubsection{Syntax}
+\label{sec:language:syntax}
 
-\input{fig/positivity}
+% \input{fig/positivity} -- follow nunes2023 instead
 \input{fig/syntax}
@@
 \subsubsection{Semantics}
+\label{sec:language:semantics}
 
@@
 \begin{definition}[$\mu$-polynomial]
-Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly_{\cat{C}}$ to be the
-smallest subcategory of $\Cat$ generated inductively by the rules in \figref{mu-polynomial}.
+Suppose $\cat{C}$ a category with finite coproducts $(\coprod, \Zero)$ and finite products $(\times, \One)$.
+Define $\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
+\figref{mu-polynomial}.
 \end{definition}
@@
 \end{enumerate}
+
 \figref{default-semantics:types} gives the default interpretation of types $\Delta \vdash \tau$ with

### 2025-06-03 main/related-work.tex

\section{Related work}

\subsection{Galois Program slicing}

\subsection{Automatic Differentiation}

\subsection{Stable Domain Theory}

\cite{berry79,berry82}

\subsection{Tangent Categories}

\cite{cockett14,cockett18}

\subsection{Lens Categories}

\cite{spivak19}

### 2025-06-03 notes/language.tex
Dropping in some todo's.
Tweaks.

@@
 
-For the sake of a ``default'' (baseline) semantics, fix a bicartesian closed category $\cat{C}$ with finite
-products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials $\internalHom{X}{Y}$, evaluation
-morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the following additional structure:
+To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
+category $\cat{C}$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
+$\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the
+following additional structure:
 \begin{enumerate}
@@
 \item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$ of $\cat{C}$;
-\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\rho}:
+\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
 \sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
@@
 \figref{default-semantics:types} gives the default interpretation of types $\Delta \vdash \tau$ with
-$\length{\Delta} = n$ as multifunctors $\cat{C}^n \to \cat{C}$ (or as objects of $\cat{C}$ in the case $n =
-0$). \figref{default-semantics:terms} gives the default interpretation of terms $\Gamma \vdash t: \tau$ as
+$\length{\Delta} = n$ as functors $\cat{C}^n \to \cat{C}$ (or as objects of $\cat{C}$ in the case $n = 0$).
+\figref{default-semantics:terms} gives the default interpretation of terms $\Gamma \vdash t: \tau$ as
 morphisms $\sem{\Gamma} \to \sem{\tau}$.
+
+\todo{Notation for partially applied functor used in \figref{default-semantics:types}}

### 2025-06-03 notes/language.tex
Tweaks.

@@
 
-\figref{default-semantics:types} gives the default interpretation of types $\Delta \vdash \tau$ with
-$\length{\Delta} = n$ as functors $\cat{C}^n \to \cat{C}$ (or as objects of $\cat{C}$ in the case $n = 0$).
-\figref{default-semantics:terms} gives the default interpretation of terms $\Gamma \vdash t: \tau$ as
-morphisms $\sem{\Gamma} \to \sem{\tau}$.
+\figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
+= n$ as functors $\cat{C}^n \to \cat{C}$, or as objects of $\cat{C}$ in the case $n = 0$.
+\figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
+$\sem{\Gamma} \to \sem{\tau}$.
 

### 2025-06-05 main/approx-as-tangents.tex
edit the "Approximation as Tangents" section

@@
 
-For what follows we do not need to know in detail what a manifold is, or exactly how smooth functions are defined.
-We only state some definitions in order to fix terminology and to justify our claim of a connection to standard differentiablity.
+% For what follows we do not need to know in detail what a manifold is, or exactly how smooth functions are defined.
+We only state some definitions in order to fix terminology and to justify our claim of a connection to manifolds and smooth functions.
 \bob{Need a good reference here.}
 
-The general study of differentiable functions usually takes place on \emph{manifolds}, which are topological spaces that are ``locally'' like a Euclidean space $\RR^n$.
-The spaces $\RR^n$ are manifolds, but so are ``non-flat'' examples such as $n$-spheres and more exotic spaces.
+The general study of differentiable functions takes place on \emph{manifolds}, which are topological spaces that ``locally'' look like a subset of the Euclidean space $\RR^n$.
+The spaces $\RR^n$ themselves are manifolds, but so are ``non-flat'' examples such as $n$-spheres and yet more exotic spaces.
 Every point $x$ in a manifold $M$ has an associated \emph{tangent vector space} $\tangents_x(M)$ and \emph{cotangent vector space} $\cotangents_x(M)$, the latter consisting of linear functions $\tangents_x(M) \linearto \RR$.
@@
 
-Our thesis is that \GPS can profitably be seen as a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory.
-Smooth functions in this setting are Berry's \emph{stable functions} in domain theory.
+Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory.
+Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry}.
 We now introduce these concepts and how they relate to \GPS.
@@
 
-Domain theory is a method for defining the semantics of programs that
-handle infinite data (e.g., functions or infinite
-streams). \emph{Domains} are special partially ordered sets where the
-ordering denotes a relationship of qualitative information content: if
-$x \sqsubseteq y$, then $y$ may contain more information than $x$. For
-example, if $x$ and $y$ are functions, then $y$ may be defined at more
-places in its domain than $x$. Infinite objects are understood in
-terms of their approximations in this sense, and domains are assumed
-to be closed under least upper bounds (lubs) of directed sets, meaning
-that any internally consistent collection of elements has a
-``completion'' that contains all the information covered by the
-set. Programs are interpreted as monotone functions that preserve
-directed lubs. The latter property is called continuity. Monoticity
-captures the idea that if the input gets more defined, then the output
-can get more defined. Continuity intuitively expresses the idea that a
-function interpreting a program cannot ``look at all the
-approximations'' when determining the output, which corresponds to the
-intuitive idea that a computable function cannot look at a non-finite
-amount of input to produce an output.
+Domain theory is a method for defining the semantics of programs that handle infinite data (e.g., functions or infinite streams).
+\emph{Domains} are certain partially ordered sets where the ordering denotes a relationship of qualitative information content: if $x \sqsubseteq y$, then $y$ may contain more information than $x$.
+For example, if $x$ and $y$ are functions, then $y$ may be defined for more values in its domain than $x$.
+Infinite objects are understood in terms of their approximations in this sense, and domains are assumed to be closed under least upper bounds (lubs) of directed sets, meaning that any internally consistent collection of elements has a ``completion'' that contains all the information covered by the set.
+Programs are interpreted as monotone functions that preserve directed lubs.
+Monotonicity captures the idea that if the input gets more defined, then the output can get more defined.
+Preservation of lubs, or \emph{continuity}, states that a function interpreting a program cannot act inconsistently on approximations and their completion, which corresponds to the intuitive idea that a function that is computable cannot look at a non-finite amount of input to produce a finite output.
 
-For the purposes of \GPS, we are not interested in using
-approximations to model computation on infinite objects, but rather to
-use them for the related task of revealing how programs explore their
-inputs for provenance tracking. Therefore, in what follows we dispense with completeness properties of the partially ordered sets we consider.
+For the purposes of \GPS, we are not interested in using approximations to model computation on infinite objects, but instead to use them for the related use of revealing how programs explore their inputs when producing parts of their output.
+Therefore, we ignore completeness properties of the partially ordered sets we consider.
 
@@
 
-Continuous functions between domains capture the intuitive idea that
-computable functions are restricted to finite observations of their
-input for finite output. However, there are examples, such as
-Plotkin's Parallel OR (\autoref{ex:parallel-or}, below) that are
-continuous but do not treat their input in a way that is consistent
-with a sequential implementation. Stability is an additional property
-of monotone functions that was invented by Berry \cite{berry86} in an
-attempt to capture sequentiality. This was not successful (see the
-$\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will
-see now how it is closely related to problem of computing forward and
-backwards maps of approximations needed in \GPS. A textbook
-description of stable functions in the context of domain theory is
-given by \citet{amadio-curien} (Chapter 12). We will
-refer to this where necessary.
+%FIXME: use cleverref
 
-In a partially ordered set $X$, for any $x \in X$ the set of elements
-below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a
-partially ordered set. These approximations of $x$ we will think of as
-``derivatives'' of $x$, and the whole set $\downset{x}$ as the
-``tangent space''. Tangent spaces are vector spaces, since they are
-linear approximations to curves at that point. In the partially
-ordered setting, we take elements of $\downset{x}$ to be
-approximations of processes defined at $x$. As we can add tangents, we
-assume we can take meets of approximations in $\downset{x}$:
+When giving a denotational semantics for sequential programming languages, such as PCF, continuous functions are too permissive.
+Plotkin's Parallel OR (\autoref{ex:parallel-or}, below) is continuous but does not treat its input in a way that is consistent with a sequential implementation \cite{plotkin77lcf}.
+Continuous functions can explore their input in a ``random access'' way, which is incompatible with the central idea in \GPS that we should be able to identify a \emph{minimal} part of the input that leads to a part of the output.
+
+Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality.
+This was not successful (see the $\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
+A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12).
+We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.
+
+In a partially ordered set $X$, for any $x \in X$ the set of elements below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a partially ordered set.
+These approximations of $x$ we will think of as ``derivatives'' of $x$, and the whole set $\downset{x}$ as the ``tangent space''.
+Tangent spaces are vector spaces, since they are linear approximations to curves at that point.
+In the partially ordered setting, we take elements of $\downset{x}$ to be approximations of processes defined at $x$.
+As we can add tangents, we assume we can take meets of approximations in $\downset{x}$:
 
 \begin{definition}
-  A \emph{bounded meet poset} is a partially ordered set $X$ where
-  every for every $x \in X$, $\downset{x}$ is a meet semilattice, with
-  $x$ as the top element.
+  A \emph{bounded meet poset} is a partially ordered set $X$ where for every $x \in X$, $\downset{x}$ is a meet semilattice, with $x$ as the top element.
 \end{definition}
 
-Smooth functions have a ``pushforward'' derivative that takes tangents
-at a point $x$ to tangents at the point $f(x)$. In the partially
-ordered setting, we require that the restriction of a monotone
-function $f$ to each $\downset{x}$ preserves meets:
+Smooth functions have a ``pushforward'' derivative that takes tangents at a point $x$ to tangents at the point $f(x)$ and is linear.
+In the partially ordered setting, we are going to take a function $f$'s pushforward derivative at $x$ to be its restriction to $\downset{x}$, taking approximations of the input to approximations of the output.
+Matching the linearity of the pushforward derivative smooth functions, we require that these restrictions preserve meets:
 
@@
     \end{array}
+    \qquad
+    \raisebox{-15ex}{
+      \begin{tikzpicture}
+        \node (top) at (0,0) {$(\mathsf{tt},\mathsf{ff})$};
+        \node [below of=top,xshift=-0.5cm] (oi) {$(\bot,\mathsf{ff})$};
+        \node [below of=top,xshift=0.5cm] (io) {$(\mathsf{tt},\bot)$};
+        \node [below of=oi,xshift=0.5cm] (oo) {$(\bot,\bot)$};
+        \draw (top) -- (oi);
+        \draw (top) -- (io);
+        \draw (oi) -- (oo);
+        \draw (io) -- (oo);
+      \end{tikzpicture}
+    }
   \end{displaymath}
-  \bob{Demonstrate that they are conditionally multiplicative, by
-    giving evidence of stability and stating that this implies cm}
-
-  The function $\mathrm{strictOr}$ is stable. For example, for the
-  input-output pair $(\mathsf{tt},\mathsf{ff}), \mathsf{tt}$, the
-  minimal input that gives this output is exactly
-  $(\mathsf{tt}, \mathsf{ff})$. If we take the approximation
-  $\bot \leq \mathsf{tt}$ of the output, then the corresponding
-  minimal input is $(\bot, \bot)$. The function
-  $\mathrm{shortCircuitOR}$ is also stable. For the input-output pair
-  $(\mathsf{tt},\mathsf{ff}),\mathsf{tt}$, the minimal input that
-  gives this input is $(\mathsf{tt},\bot)$, indicating that the
-  presence of $\mathsf{ff}$ in the second argument was not necessary
-  to produce this output. As with $\mathrm{strictOr}$, the minimal
-  input required to produce the output $\bot \leq \mathsf{tt}$ is
-  again $(\bot,\bot)$.
-
-  The fact that these two functions's stability witnesses reveal
-  information about how they depend on their input is what we will
-  exploit in order to use the idea of stability for slicing.
+  In the poset $\mathbb{B}_\bot^2$, a typical poset of approximations of a fully defined element is shown to the right.
+For $\mathrm{strictOr}$, any approximation that isn't the fully defined input is mapped to $\bot$, while $\mathrm{shortCircuitOr}$ maps the partially defined $(\mathsf{tt},\bot)$ to $\mathsf{tt}$.
+Thus, even though these functions operate identically on fully defined inputs, they differ in their \emph{derivatives} on partially defined input, exposing how .
+That they are cm can be checked by examining their restrictions' behaviour.
+If we take the approximations $(\bot,\mathsf{ff})$ and $(\mathsf{tt},\bot)$, then their meet is $(\bot,\bot)$; $\mathrm{strictOr}$ maps all three elements to $\bot$, so is cm here since $\bot \wedge \bot = \bot$; and $\mathrm{shortCircuitOR}$ has $(\bot,\mathsf{ff}) \mapsto \bot$ and $(\mathsf{tt},\bot) \mapsto \mathsf{tt}$, the meet of which is $\bot = \mathrm{shortCircuitOR}(\bot,\bot)$.
+Other combinations can be checked similarly.
 \end{example}
@@
   \label{ex:parallel-or}
-  A function that is not cm is Plotkin's Parallel OR \cite{lcf77},
-  which short-circuits in both arguments, returning $\mathsf{tt}$ if
-  either argment is $\mathsf{tt}$, even if the other argument is not
-  defined:
+  A function that is not cm is Plotkin's Parallel OR \cite{lcf77}, which short-circuits in both arguments.
+It returns $\mathsf{tt}$ if either argment is $\mathsf{tt}$ even if the other argument is not defined:
   \begin{displaymath}
@@
   \end{displaymath}
-  For the input-output pair $(\mathsf{tt},\mathsf{tt}),\mathsf{tt}$,
-  there is no one minimal input that produces this output. We have
-  both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and
-  $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are
-  incomparable and their greatest lower bound $(\bot,\bot)$ gives the
-  output $\bot$.
+  We have $\mathrm{parallelOR}(\mathsf{tt}, \bot) \wedge \mathrm{parallelOR}(\bot, \mathsf{tt}) = \mathsf{tt} \wedge \mathsf{tt} = \mathsf{tt}$ but $\mathrm{parallelOR}((\mathsf{tt},\bot) \wedge (\bot, \mathsf{tt})) = \mathrm{parallelOR}(\bot, \bot) = \bot$, so it is not cm.
 
-  Parallel OR is famous because it is not \emph{sequential}, meaning
-  intuitively that it cannot be implemented without running the two
-  arguments in parallel to see if one of them returns
-  $\mathsf{tt}$. That fact that it exists in the standard domain
-  theoretic semantics of PCF means that this semantics is complete for
-  reasoning about observational equivalence in PCF. Since Parallel OR
-  is not stable, one might hope that stability is enough to capture
-  sequentiality, and hence potentially give a fully abstract model of
-  PCF. However, the following ternary function
-  $\mathbb{B}_\bot^3 \to \{\top,\bot\}$ is stable but admits no
-  sequential implementation that fixes an order that the arguments are
-  examined in:
+  Parallel OR is famous because it is not \emph{sequential}, meaning intuitively that it cannot be implemented without running the two arguments in parallel to see if one of them returns $\mathsf{tt}$.
+The fact that it exists in the standard domain theoretic semantics of PCF means that this semantics is incomplete for reasoning about observational equivalence in PCF.
+Since Parallel OR is not cm, one might hope that cm-ness is enough to capture sequentiality, and hence potentially give a fully abstract model of PCF.
+However, the following ternary function $\mathbb{B}_\bot^3 \to \{\top,\bot\}$ is cm but admits no sequential implementation that fixes an order that the arguments are examined in:
   \begin{displaymath}
@@
   \end{displaymath}
-  Despite there being no one minimal input that achieves the output
-  $\top$, each of the minimal inputs that can achieve this output are
-  pairwise incomparable, so for each specific input that gets output
-  $\top$ there is a unique minimal input that achieves it (listed in
-  the first three lines of the definition).
+  Due to the way that the cases are defined, there is no way of constructing a pair of approximations for which preservation of their meet does not hold.
+In terms of derivatives, this makes sense in that we are only concerned about the intensional behaviour of a function at a point and its approximations.
+Parallel OR has two incompatible approximation behaviours at the point $(\mathsf{tt},\mathsf{tt})$.
+The $\mathrm{gustave}$ function does have consistent behaviour at each approximation for each point.
 
-  In terms of \GPS the $\mathrm{gustave}$ function does not present a
-  problem. For any particular run (i.e., input-output pair) of the
-  program, there is an unambiguous minimal input that achieves the
-  output, no matter that it was not achieved by a sequential
-  processing of the input.
+%  \bob{What have hyperstable functions?}
 \end{example}
+
+Conditional multiplicativity seems to be a reasonable analogue to functions with a well-defined notion of derivative.
+However, for \GPS we also require an analogue to the reverse derivative, where we map approximations backwards to give the least approximation of the input for a given approximation of the output.
+In the case of smooth functions, we are always guaranteed a reverse derivative but there is not always a best way to map approximations backwards for cm functions, as the following example shows.
 
@@
   \label{ex:non-stable-function}
-  In the preceeding examples, we demonstrated conditional
-  multiplicativity by giving a backwards mapping of approximations
-  that forces preservation of meets. A natural question is whether or
-  not we always have such a backwards map which we could use for
-  backwards propagation of approximations. An example that is
-  conditionally multiplicative, but does not admit such a backards map
-  (from Amadio and Curien just before Lemma 12.2.3, originally due to
-  Berry) is given by defining
-  $\mathrm{unstable} : D \to \{\bot, \top\}$, where:
+  % In the preceeding examples, we demonstrated conditional
+  % multiplicativity by giving a backwards mapping of approximations
+  % that forces preservation of meets. A natural question is whether or
+  % not we always have such a backwards map which we could use for
+  % backwards propagation of approximations.
+  An example that is conditionally multiplicative, but does not admit a backards map of approximations (from \citet{amadio-curien} just before Lemma 12.2.3, originally due to Berry) is given by defining $\mathrm{unstable} : D \to \{\bot, \top\}$, where:
   \begin{displaymath}
@@
   \end{displaymath}
-  as $\mathrm{unstable}(\bot) = \bot$ and
-  $\mathrm{unstable}(n) = \top$. This is monotone, and preserves meets
-  in every $\downset{x}$. However, there is no ``best'' (i.e., least)
-  input that gives us any finite output.
+  as $\mathrm{unstable}(\bot) = \bot$ and $\mathrm{unstable}(n) = \top$.
+This is monotone, and preserves meets in every $\downset{x}$.
+However, there is no ``best'' (i.e., least) input that gives us any finite output.
 \end{example}
@@
 definition of a function that requires the existence of a reverse
-mapping directly, without assuming that any meets exist:
+mapping directly, even without assuming that any meets exist:
 
@@
 \begin{example}
-  \bob{Elaborate on the examples}
   \begin{enumerate}
-  \item $\mathrm{strictOr}$ and $\mathrm{shortCircuitOR}$ are stable, as demonstrated above.
-  \item $\mathrm{parallelOR}$ is not stable.
-  \item $\mathrm{gustave}$ is stable.
-  \item $\mathrm{unstable}$ is not stable, but is conditionally multiplicative.
+  \item The function $\mathrm{strictOr}$ is stable.
+For example, for the input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input that gives this output is exactly $(\mathsf{tt}, \mathsf{ff})$.
+If we take the approximation $\bot \leq \mathsf{tt}$ of the output, then the corresponding minimal input is $(\bot, \bot)$.
+The function $\mathrm{shortCircuitOR}$ is also stable.
+For the input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input that gives this input is $(\mathsf{tt},\bot)$, indicating that the presence of $\mathsf{ff}$ in the second argument was not necessary to produce this output.
+As with $\mathrm{strictOr}$, the minimal input required to produce the output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.
+
+  \item The $\mathrm{parallelOR}$ function is not stable.
+For the input-output pair $(\mathsf{tt},\mathsf{tt}) \mapsto \mathsf{tt}$, there is no one minimal input that produces this output.
+We have both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are incomparable and their greatest lower bound $(\bot,\bot)$ gives the output $\bot$.
+
+  \item The $\mathrm{gustave}$ function is stable.
+Despite there being no one minimal input that achieves the output $\top$, each of the minimal inputs that can achieve this output are pairwise incomparable, so for each specific input that gets output $\top$ there is a unique minimal input that achieves it (listed in the first three lines of the definition).
+
+    In terms of \GPS the $\mathrm{gustave}$ function does not present a problem.
+For any particular run (i.e., input-output pair) of the program, there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
+
+  \item As discussed above, $\mathrm{unstable}$ is not stable, but is conditionally multiplicative.
   \end{enumerate}
 \end{example}
+
+The fact that these two functions's stability witnesses reveal information about how they depend on their input is what we will exploit in order to use the idea of stability for \GPS.
+It is interesting that sequentiality is not necessarily required in order to obtain backwards maps of approximations, as might be assumed by thinking about how to trace computations for provenance information.
 
@@
 \begin{definition}
-  An \emph{L-poset} is a partially ordered set $X$ such that for every
-  $x \in X$, the principal downset $\downset{x}$ is a bounded lattice
-  (i.e., have all finite meets and joins).
+  An \emph{L-poset} is a partially ordered set $X$ such that for every $x \in X$, the principal downset $\downset{x}$ is a bounded lattice (i.e., have all finite meets and joins).
 \end{definition}
@@
 
-The converse to this lemma (that functions that preserve meets in
-their forward part have a left Galois adjoint) is not true, as was
-demonstrated by the non-stable function in
-\autoref{ex:non-stable-function}. In the case when the posets
-$\downset{x}$ are \emph{complete}, and $f_x$ preserves infinitary
-meets, then we are guaranteed a left Galois adjoint. In that example,
-the infinite set $\{0, 1, 2, \dots, n, n-1, \dots\}$ not including
-$\bot$ of approximations of $0$ does not have a greatest lower bound,
-so the order is not complete.
+The converse to this lemma (that functions that preserve meets in their forward part have a left Galois adjoint) is not true, as was demonstrated by the non-stable function in \autoref{ex:non-stable-function}.
+In the case when the posets $\downset{x}$ are \emph{complete}, and $f_x$ preserves infinitary meets, then we are guaranteed a left Galois adjoint.
+In that example, the infinite set $\{0, 1, 2, \dots, n, n-1, \dots\}$ not including $\bot$ of approximations of $0$ does not have a greatest lower bound, so the order is not complete.
 
-We could have required that L-posets' principal downsets were complete
-lattices, which would mean that we would only need to require that
-forward approximation maps preserve all meets and we could compute the
-backwards map automatically. However, our goal is a computable theory
-for \GPS, analogous to automatic differentiation. Requiring
-completeness to compute backwards maps of approximations would require
-that complete joins are always computable, which will not be the case for approximations of functions.
+We could have required that L-posets' principal downsets were complete lattices, which would mean that we would only need to require that forward approximation maps preserve all meets and we could compute the backwards map automatically.
+However, our goal is a computable theory for \GPS, analogous to automatic differentiation.
+Requiring completeness to compute backwards maps of approximations would require that complete joins are always computable, which will not be the case for approximations of functions.
 
@@
 
-We now summarise the analogy between differentiable functions between manifolds and stable functions between L-posets:
+We now summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
 
@@
 
+\bob{Put a conjecture about Tangent Categories in somewhere}
+
 Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.

### 2025-06-06 notes/language.tex
C -> Sem for semantic category. Fix up Makefile for MacOS.

@@
 To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\cat{C}$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
+category $\Sem$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
 $\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the
@@
 \mathcal{L}(X \times Y)$;
-\item $\mu$-polynomial endofunctors given by $\muPoly_{\cat{C}}$;
-\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$ of $\cat{C}$;
+\item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
+\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$;
 \item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
@@
 \figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
-= n$ as functors $\cat{C}^n \to \cat{C}$, or as objects of $\cat{C}$ in the case $n = 0$.
+= n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
 \figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms

### 2025-06-07 main/approx-as-tangents.tex
Typo.

@@
 \item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$.
-In the Eucdlidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
+In the Euclidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
 \item The \emph{backward derivative} (cotangent map, or pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$.

### 2025-06-07 main/approx-as-tangents.tex
Unresolved citations.

@@
 Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory.
-Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry}.
+Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry79,berry82}.
 We now introduce these concepts and how they relate to \GPS.
@@
 
-Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality.
+Stability is a property that can be required of monotone functions that was invented by Berry~\cite{berry79} in an attempt to capture sequentiality.
 This was not successful (see the $\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.

### 2025-06-07 main/conclusion.tex
Note on grad-CAM.

\section{Conclusion and Future Work}

\todo{
\begin{itemize}
\item Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping
(Grad-CAM)~\cite{selvaraju20} use reverse-mode AD selectively on specific activations, calculating \emph{heat
maps} that highlight input regions contributing to a given classification decision. It might be possible to
compose these techniques with \GPS whilst still preserving useful round-tripping properties; if so this could
be a basis for end-to-end explainability in systems that combine neural networks with traditional symbolic
computation.
\end{itemize}
}

### 2025-06-07 main/higher-order.tex
Note on grad-CAM.
Add language section.

@@
-\section{Interpreting Higher-Order Programs}
+\section{CHAD at Higher-Order}
 \label{sec:higher-order}
 
-$\Fam(\FinVect)$ and $\Fam(\LatGal)$ provide appropriate notions of composition for automatic differentiation
-and Galois slicing respectively, but neither is Cartesian closed and thereby able to interpret higher-order
-programs directly. However the following is a special case of a construction due to \citet{nunes2023}:
+$\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
+composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed and
+thereby able to interpret higher-order programs directly. However the following is a special case of a
+construction due to \citet{nunes2023}:
 
@@
 \begin{itemize}
-\item the indexing set $Z$ is the $\Fam(\cat{C})(X, Y)$, which is indeed a set by
+\item the indexing set $Z$ is $\Fam(\cat{C})(X, Y)$, which is indeed a set by
 \propref{Grothendieck:fam-inherits-local-smallness};

### 2025-06-07 main/first-order.tex
Rename, merge files.

--- 2025-05-27 main/category-of-families.tex
@@
+\section{CHAD at First-Order}
+\label{sec:first-order}
+
+Our denotational approach to \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
+Before we go into details, we present a high-level overview of their approach and how we have applied it to \GPS.
+
+The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
+An object of the Grothendieck construction $\int T$, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
+The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
+The maps in the indexed category are the maps of tangents or approximations, being derivatives and Galois connections respectively.
+
+In the case of differentiable programs and automatic differentiation, a basic model can be constructed by taking $\cat{C}$ to be $\Set$ and $T(X) = X \to \FinVect$: indexed collections of finite dimensional vector spaces, whose elements are interpreted as tangents at the given point, with linear maps.
+The Grothendieck category $\int T$ has objects that are pairs of a set $X$ and for each $x \in X$ a tangent vector space $\partial X(x)$.
+Morphisms are pairs of maps of points to points, accompanied by linear maps of tangents at those points.
+If we carefully choose an initial set of maps, we can read these maps as derivatives (note that nothing in the Grothendieck construction says that they must be derivatives!).
+Composition in the Grothendieck category is exactly the {\em chain rule} for composing derivatives, so we at least do know that if we start with derivatives on basic operations, then composing them will retain this property.
+
+For \GPS, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.
+
+In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$~(\secref{Fam}).
+These settings are adequate for interpreting first-order programs, but neither is Cartesian closed.
+The picture for higher-order programs is more complicated.
+As is well known, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$~\cite{lawvere63}.
+In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$, using the Cartesian products from $\Set$.
+When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
+If the coproducts are actually {\em biproducts}~(\secref{biproducts}), then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
+
+Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of their having products and being enriched in commutative monoids~(\propref{biproducts:from-product-or-coproduct} below).
+However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
+Neither of these is ideal.
+In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost.
+In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model.
+Moreover, in Agda, there are issues with predicativity when considering complete lattices.
+\todo{be more specific}
+
+Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
+We overcome this by using $\CMon$-enriched presheaves, as described in \secref{higher-order}; first we present a suitable first-order setting.
+
+\subsection{Category of Lattices and Galois Connections}
+
+We start with the category of lattices and Galois connections, which will serve as our base category for
+interpreting forward and backward \GPS. Galois connections are pairs of monotone functions $f: Y \to X$ and
+$g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
+and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the bidirectional
+nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the least input
+slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the greatest
+output slice for a given slice of the input).
+
+\begin{definition}[Galois connection]
+Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
+$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
+\end{definition}
+
+\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
+Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
+$f$ and $g$ regarded as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
+the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
+as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
+connections between them as morphisms.
+
+As sketched in \secref{introduction:galois-slicing}, for \GPS we would like a setting where the approximants
+of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing
+the approximation that discards all information about $x$, greatest element $\top$ representing the
+approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to
+combine approximations. Thus rather than working directly in $\PosGal$, we will use the following subcategory
+instead.
+
+\begin{definition}
+Define $\LatGal$ to be the category which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded
+lattices, and as morphisms all Galois connections $f \adj g: X \to Y$.
+\end{definition}
+
+\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
+in $\LatGal$, $g$ is a (bounded) \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice
+structure $(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join,
+\bot)$.
+
+The meet-preserving maps from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$ on such
+maps is given pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves meets; the
+constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets, provides the
+unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by
+$\join$ and the constant map $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two
+constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit
+$\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
+connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$. Thus $\LatGal$ is
+enriched in $\SemiLat$, the category of semilattices and semilattice homomorphisms.
+
+$\LatGal$ is not Cartesian closed, and so not directly suitable for interpreting higher-order programs. In
+fact Cartesian closure is incompatible with another important property of $\LatGal$. The projections $\pi_1: X
+\times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the product of lattices, have both
+upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is
+both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and
+$\biinj_Y$ given by:
+
+\vspace{-4mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   \biproj_X = \prodM{\id_X}{\bot_Y} \adj \proj_1: X \biprod Y \to X \\
+   \biproj_Y = \prodM{\bot_X}{\id_Y} \adj \proj_2: X \biprod Y \to Y
+\end{align*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   \biinj_X = \pi_1 \adj \prodM{\id_X}{\top_Y} : X \to X \biprod Y \\
+   \biinj_Y = \pi_2 \adj \prodM{\top_X}{\id_Y}: Y \to X \biprod Y
+\end{align*}
+\end{center}
+\end{minipage}
+\vspace{2mm}
+
+\noindent The trivial 1-point lattice, which is both terminal and initial and we write as $0$, is the unit for
+$\biprod$. Such a structure $(\biprod, 0)$ where $\biprod$ acts as both a product and a coproduct is called a
+\emph{biproduct} (\secref{biproducts}). The presence of a zero object in a Cartesian closed category $\cat{C}$
+renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
+\Rightarrow Y} \iso 1$, and so $\cat{C}$ has only a single object, up to isomorphism. \todo{Bridge better to
+next section; even if $\LatGal$ were Cartesian closed, directly interpreting into $\LatGal$ wouldn't allow
+each point to be associated with its own lattice of approximants.}
+
 \subsection{Grothendieck Construction for Indexed Categories}

### 2025-06-07 main/first-order.tex
FinVect and LatGal as examples of semi-additive categories.

@@
 
-\subsection{Category of Lattices and Galois Connections}
+\subsection{Biproducts and Semi-Additive Categories}
+\label{sec:biproducts}
 
-We start with the category of lattices and Galois connections, which will serve as our base category for
-interpreting forward and backward \GPS. Galois connections are pairs of monotone functions $f: Y \to X$ and
+\todo{Introduce $\CMon$-category first.} We start with the idea of a biproduct and semi-additive categories,
+which are always enriched in $\CMon$, the category of commutative monoids.
+
+\begin{definition}[Semi-additive category]
+\label{def:biproducts:semi-additive}
+A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
+and injections) give an isomorphism
+\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
+\end{definition}
+
+The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
+projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
+initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
+two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
+forming a morphism $f + g: X \to Y$. Diagrammatically:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
+\end{tikzcd}
+\end{center}
+
+Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
+$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
+The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
+\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
+composing the unique maps in and out of the zero object:
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
+\end{tikzcd}
+\end{center}
+
+It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
+hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
+family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
+preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+
+\begin{salign*}
+f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
+\end{salign*}
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}
+
+\subsubsection{Biproduct laws}
+It is also easy to show that in a semi-additive category the following equations hold:
+%\vspace{-4mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_X \comp \biinj_X &= \id_X \\
+   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
+\end{salign*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_Y \comp \biinj_Y &= \id_Y \\
+   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
+\end{salign*}
+\end{center}
+\end{minipage}
+
+\begin{salign*}
+(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
+\end{salign*}
+
+\vspace{3mm}
+\noindent It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
+in $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
+coproducts in $\cat{C}$ are necessarily biproducts.
+
+\begin{proposition}
+\label{prop:biproducts:from-product-or-coproduct}
+In a category $\cat{C}$ enriched in $\CMon$:
+\begin{enumerate}
+\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \inj_1, \inj_2)$ is a coproduct, where
+$\inj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\inj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
+\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \pi_1, \pi_2)$ is a product, where
+$\pi_1 = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and $\pi_2 = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to
+Y$.
+\item If $X$ is terminal (resp.~initial) then $X$ is initial (resp.~terminal).
+\end{enumerate}
+\end{proposition}
+
+\subsubsection{Example: Category of Finite Vector Spaces}
+\label{sec:categories-with-biproducts:fdvect}
+
+The category $\FinVect$ of finite vector spaces over $\RR$ (with addition of real numbers written $a + b$ and
+multiplication written $a \mult b$) is semi-additive.
+
+\begin{definition}[Finite dimensional vector space over $\RR$]
+Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
+component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
+is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
+vector operations are compatible with the field operations in that the following equations hold \todo{probably
+more detail than we need to go into}:
+
+\vspace{-4mm}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   1 \mult v &= v \\
+   (a \mult b) \mult v &= a \mult (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{align*}
+   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
+   a \mult (u + v) &= (a \mult u) + (b \mult v)
+\end{align*}
+\end{center}
+\end{minipage}
+\end{definition}
+
+\vspace{1mm}
+\noindent Composition is with bilinear. The direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian
+product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
+\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$.
+
+\begin{definition}
+Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces, and as
+morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
+\mult f(v)$, with $f$ preserving the abelian group structure automatically.
+\end{definition}
+
+$\FinVect$ is enriched in $\Ab$, the category of abelian groups \todo{Introduce $\CMon$-enrichment first, then
+perhaps $\Ab$-enrichment parenthetically}. For any objects $V = \RR^m$ and $W = \RR^n$ the hom-set
+$\FinVect(V,W)$ is an abelian group, with identity $0_{V,W}$ given by the constant map $v \mapsto 0_W$ and
+addition and additive inverse given pointwise:
+
+\vspace{-4mm}
+\begin{center}
+\begin{minipage}[t]{0.35\textwidth}
+\begin{align*}
+(f + g)(v) &= f(v) + g(v)
+\end{align*}
+\end{minipage}%
+\begin{minipage}[t]{0.35\textwidth}
+\begin{align*}
+(-f)(v) &= -f(v)
+\end{align*}
+\end{minipage}
+\end{center}
+
+\subsubsection{Example: Category of Lattices and Galois Connections}
+
+\todo{Rework to account for new ordering.} Galois connections are pairs of monotone functions $f: Y \to X$ and
 $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
@@
 
+\todo{Again, introduce $\CMon$-enrichment first and relegate $\SemiLat$-enrichment to parenthetical remark.}
 The meet-preserving maps from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$ on such
@@
 renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and so $\cat{C}$ has only a single object, up to isomorphism. \todo{Bridge better to
-next section; even if $\LatGal$ were Cartesian closed, directly interpreting into $\LatGal$ wouldn't allow
-each point to be associated with its own lattice of approximants.}
+\Rightarrow Y} \iso 1$, and so $\cat{C}$ has only a single object, up to isomorphism.
 
@@
 Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
-setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic ideas here, starting with the
-category $\FinVect$ of finite vector spaces over $\RR$, with addition of real numbers written $a + b$ and
-multiplication written $a \mult b$.
-
-\subsubsection{Category of finite-dimensional vector spaces}
-\label{sec:categories-with-biproducts:fdvect}
-
-\begin{definition}[Finite dimensional vector space over $\RR$]
-Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
-component-wise addition of vectors $+: \RR^n \times \RR^n \to \RR^n$ and scalar multiplication $\mult: \RR
-\times \RR^n \to \RR^n$ (overloading $+$ and $\mult$), where $(\RR^n,+)$ is an abelian group (with identity
-$0$ and additive inverse $-v$ again defined component-wise), and where the vector operations are compatible
-with the field operations in that the following equations hold:
-
-\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   1 \mult v &= v \\
-   (a \mult b) \mult v &= a \mult (b \mult v)
-\end{align*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
-   a \mult (u + v) &= (a \mult u) + (b \mult v)
-\end{align*}
-\end{center}
-\end{minipage}
-\end{definition}
-
-\begin{definition}[Category $\FinVect$]
-Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces, and as
-morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
-\mult f(v)$, with $f$ preserving the abelian group structure automatically.
-\end{definition}
-
-$\FinVect$ is enriched in $\Ab$, the category of abelian groups. For any objects $V = \RR^m$ and $W = \RR^n$
-the hom-set $\FinVect(V,W)$ is an abelian group, with identity $0_{V,W}$ given by the constant map $v \mapsto
-0_W$ and addition and additive inverse given pointwise:
-
-\vspace{-4mm}
-\begin{center}
-\begin{minipage}[t]{0.35\textwidth}
-\begin{align*}
-(f + g)(v) &= f(v) + g(v)
-\end{align*}
-\end{minipage}%
-\begin{minipage}[t]{0.35\textwidth}
-\begin{align*}
-(-f)(v) &= -f(v)
-\end{align*}
-\end{minipage}
-\end{center}
-
-\vspace{1mm}
-\noindent Composition is with bilinear. The direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian
-product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
-\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$. \todo{Significance of biproducts}
+setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic ideas here.
 

### 2025-06-07 main/higher-order.tex
FinVect and LatGal as examples of semi-additive categories.

@@
 }
-
-\subsection{Biproducts and Semi-Additive Categories}
-\label{sec:biproducts}
-
-We noted that $\LatGal$ is enriched in $\SemiLat$, the category of semilattices, and moreover that the product
-of lattices is also a coproduct, making $\LatGal$ into a category with \emph{biproducts}. In fact the presence
-of biproducts is sufficient for enrichment in $\CMon$, the category of commutative monoids (including
-semilattices); such a category is called \emph{semi-additive}.
-
-\begin{definition}[Semi-additive category]
-\label{def:biproducts:semi-additive}
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
-projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
-two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-\subsubsection{Biproduct laws}
-It is also easy to show that in a semi-additive category the following equations hold:
-%\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-\vspace{3mm}
-\noindent It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
-in $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
-coproducts in $\cat{C}$ are necessarily biproducts.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-In a category $\cat{C}$ enriched in $\CMon$:
-\begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \inj_1, \inj_2)$ is a coproduct, where
-$\inj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\inj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \pi_1, \pi_2)$ is a product, where
-$\pi_1 = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and $\pi_2 = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to
-Y$.
-\item If $X$ is terminal (resp.~initial) then $X$ is initial (resp.~terminal).
-\end{enumerate}
-\end{proposition}

### 2025-06-07 main/first-order.tex
Pare back LatGal a bit.

@@
 
+A biproduct category $\cat{C}$ cannot be Cartesian closed without being trivial. If $0 \times X \iso X$ then
+$\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$, and so $\cat{C}$ is terminal, with only a
+single object up to isomorphism.
+
 \subsubsection{Example: Category of Finite Vector Spaces}
@@
 $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
-and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the bidirectional
-nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the least input
-slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the greatest
-output slice for a given slice of the input).
+and $g$ the best approximation from above to an inverse of $f$. The category $\LatGal$ of bounded lattices and
+Galois connections is semi-additive.
 
@@
 
-\noindent Galois connections thus generalise order isomorphisms. The $\adj$ notation is justified because a
-Galois connection $f \adj g: X \to Y$ can also be seen an adjunction between poset categories, with monotone
-$f$ and $g$ regarded as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
-the \emph{lower} (left) adjoint. Galois connections compose component-wise, with $\id_X \adj \id_X: X \to X$
-as the unit for composition, and thus form a category $\PosGal$ with all posets as objects and all Galois
-connections between them as morphisms.
-
-As sketched in \secref{introduction:galois-slicing}, for \GPS we would like a setting where the approximants
-of a point $x$ form a bounded lattice $(X, \meet, \join, \top, \bot)$, with least element $\bot$ representing
-the approximation that discards all information about $x$, greatest element $\top$ representing the
-approximation that retains all information about $x$, and $\meet$ and $\join$ providing two canonical ways to
-combine approximations. Thus rather than working directly in $\PosGal$, we will use the following subcategory
-instead.
+\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
+adjunction between poset categories, with monotone $f$ and $g$ regarded as functors; $f$ is usually referred
+to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections compose
+component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition.
 
@@
 
-$\LatGal$ is not Cartesian closed, and so not directly suitable for interpreting higher-order programs. In
-fact Cartesian closure is incompatible with another important property of $\LatGal$. The projections $\pi_1: X
-\times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the product of lattices, have both
-upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is
-both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and
-$\biinj_Y$ given by:
+The projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the
+product of lattices, have both upper and lower adjoints; this means that $X \times Y$, which we shall
+hereafter write as $X \biprod Y$, is both a product and a coproduct, with projections $\biproj_X$ and
+$\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
 
@@
 \noindent The trivial 1-point lattice, which is both terminal and initial and we write as $0$, is the unit for
-$\biprod$. Such a structure $(\biprod, 0)$ where $\biprod$ acts as both a product and a coproduct is called a
-\emph{biproduct} (\secref{biproducts}). The presence of a zero object in a Cartesian closed category $\cat{C}$
-renders $\cat{C}$ trivial; if $0 \times X \iso X$ then $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and so $\cat{C}$ has only a single object, up to isomorphism.
+$\biprod$.
 

### 2025-06-08 main/first-order.tex
How does the lattice of approximants in Fam(LatGal) actually get set up?

@@
 
-The goal now is to equip the interpretation of a (first-order) program with a Galois connection relating
-approximations of its inputs to approximations of its output. Since the specific lattices of approximations
-depend on the value being approximated, the semantics of every type $\tau$ must have both a conventional
-component as well as an approximating component, i.e.~a set $X$ which serves as the usual interpretation of
-$\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in X$. The semantics
-of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with $\tau$ and $\sigma$
-denoting $X$ and $Y$) must then also have two components: a function $f: X \to Y$ giving the conventional
-interpretation of $e$, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$
-for every $x \in X$ which can be used for slicing over that expression.
+\todo{Rework.} The goal now is to equip the interpretation of a (first-order) program with a Galois connection
+relating approximations of its inputs to approximations of its output. Since the specific lattices of
+approximations depend on the value being approximated, the semantics of every type $\tau$ must have both a
+conventional component as well as an approximating component, i.e.~a set $X$ which serves as the usual
+interpretation of $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in
+X$. The semantics of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with
+$\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have two components: a function $f: X \to Y$ giving
+the conventional interpretation of $e$, plus a family of Galois connections $\partial f(x): \partial X(x) \to
+\partial Y(f(x))$ for every $x \in X$ which can be used for slicing over that expression.
 
@@
 \label{prop:Grothendieck:fam-inherits-products}
-If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
+\item
+\begin{enumerate}
+If $\cat{C}$ has (symmetric) monoidal products then $\Fam(\cat{C})$ has (symmetric) monoidal products
+$(X, \partial X) \tensor (Y, \partial Y) = (X \times Y, \partial(X \times Y))$ indexed by the Cartesian
+product $X \times Y$, where $\partial(X \times Y)(x,y) = \partial X(x) \tensor \partial Y(y)$.
+\end{enumerate}
 \end{proposition}
 
-\subsection{Galois Slicing via a Category of Families}
+\subsection{First-Order Galois Slicing via CHAD}
 \label{sec:fam:galois-slicing}
@@
 
-\subsection{Automatic Differentiation via a Category of Families}
+\todo{Show how $\LatGal$ has a lifting monad?}
+
+\subsection{First-Order Automatic Differentiation via CHAD}
 

### 2025-06-08 main/first-order.tex
(Strong) monads preserved from C into Fam(C).

@@
 \label{prop:Grothendieck:fam-inherits-products}
-\item
-\begin{enumerate}
-If $\cat{C}$ has (symmetric) monoidal products then $\Fam(\cat{C})$ has (symmetric) monoidal products
-$(X, \partial X) \tensor (Y, \partial Y) = (X \times Y, \partial(X \times Y))$ indexed by the Cartesian
-product $X \times Y$, where $\partial(X \times Y)(x,y) = \partial X(x) \tensor \partial Y(y)$.
-\end{enumerate}
+If $\cat{C}$ is (symmetric) monoidal then $\Fam(\cat{C})$ is also (symmetric) monoidal, with tensor product
+$(X, \partial X) \tensor (Y, \partial Y) = (X \times Y,(x, y) \mapsto \partial X(x) \tensor \partial Y(y))$.
 \end{proposition}
+
+The action of the families construction, regarded as a covariant 2-functor $\Cat \to \Cat$, sends functors $F:
+\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that use $F$ to reassign the
+target category, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial
+f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation with $(F \comp
+\partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a
+(strong) monad on $\Fam(\cat{C})$, defined by lifting the original monad $F: \cat{C} \to \cat{C}$ pointwise
+over families.
 

### 2025-06-08 main/first-order.tex
Lifting.

@@
 \cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that use $F$ to reassign the
-target category, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial
-f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation with $(F \comp
-\partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a
-(strong) monad on $\Fam(\cat{C})$, defined by lifting the original monad $F: \cat{C} \to \cat{C}$ pointwise
-over families.
+target category by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and
+morphisms $(f, \partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural
+transformation with $(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on
+$\cat{C}$, then $\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by lifting the original monad $F:
+\cat{C} \to \cat{C}$ pointwise over families.
 
@@
 function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to \partial
-Y(f(x))$. This is precisely the setup we identified informally at the beginning of this section. The
-reindexing along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial
-g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois connection at each
-point, using the baseline (unapproximated) function $f$.
+Y(f(x))$. The reindexing along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
+\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
+connection at each point, using the baseline (unapproximated) function $f$.
+
+For a first-order language with a primitive type $\syn{nat}$ and primitive operations such as $+$, the
+``leaves'' of the approximation lattices would provided by a model of $\syn{nat}$ and its operations in
+$\Fam(\LatGal)$, for example assigning $\syn{nat}$ to the object $(\Nat,n \mapsto \Two)$, where the family is
+fully present ($\top$) or fully absent ($\bot$) from the approximation (as shown in the example in component
+is constantly $\Two = (\bot,\top)$, the two-point lattice indicating whether a given natural number
+\secref{introduction}). These ``leaf'' objects would then be combined by the interpretation of composite types
+like sums and products, providing suitable derived approximation structures for composite values.
+
+One might also like to allow the programmer be able to embed additional ``approximation points'' into
+composite structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual
+definition of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define a type
+of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along with an
+``approximating'' version of multiplication $*: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat} \to
+\syn{L}\;\syn{nat}$.
 
@@
 isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
-\RR^n$ at a point $x$ have the following form:
+\RR^n$ at a point $x$ have the following form \todo{Can probably delete now as covered in
+\secref{approx-as-tangents}}:
 

### 2025-06-08 main/first-order.tex
Approximating multiplication.

@@
 composite structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual
-definition of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define a type
-of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along with an
-``approximating'' version of multiplication $*: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat} \to
-\syn{L}\;\syn{nat}$.
+definition of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define within
+the language a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$,
+along with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times
+\syn{L}\;\syn{nat} \to \syn{L}\;\syn{nat}$ expressible as
+\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
+0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].
 

### 2025-06-08 main/first-order.tex
Approximating multiplication.

@@
 
-One might also like to allow the programmer be able to embed additional ``approximation points'' into
-composite structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual
-definition of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define within
-the language a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$,
-along with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times
-\syn{L}\;\syn{nat} \to \syn{L}\;\syn{nat}$ expressible as
+One might also like to allow the programmer to embed additional ``approximation points'' into composite
+structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual definition
+of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define within the
+language a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
+with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat}
+\to \syn{L}\;\syn{nat}$ expressible as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =

### 2025-06-09 main/approx-as-tangents.tex
Lifting.
minor edits to approx-as-tangents

--- 2025-06-05 main/approx-as-tangents.tex
@@
 
-\subsubsection{Stable Functions as ``Smooth Functions''}
-
-%FIXME: use cleverref
+\subsubsection{Bounded Meets and Conditional Multiplicativity}
+\label{sec:bounded-meets-and-cm}
 
 When giving a denotational semantics for sequential programming languages, such as PCF, continuous functions are too permissive.
-Plotkin's Parallel OR (\autoref{ex:parallel-or}, below) is continuous but does not treat its input in a way that is consistent with a sequential implementation \cite{plotkin77lcf}.
+Plotkin's Parallel OR (\exref{parallel-or}, below) is continuous but does not explore its input in a way consistent with a sequential implementation \cite{plotkin77lcf}.
 Continuous functions can explore their input in a ``random access'' way, which is incompatible with the central idea in \GPS that we should be able to identify a \emph{minimal} part of the input that leads to a part of the output.
@@
 Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality.
-This was not successful (see the $\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
+This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
 A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12).
@@
 The $\mathrm{gustave}$ function does have consistent behaviour at each approximation for each point.
-
-%  \bob{What have hyperstable functions?}
 \end{example}
@@
 
+\subsubsection{Stable functions and L-posets}
+
 In light of the preceeding example, we can give an alternative
@@
 \end{definition}
-
-\begin{lemma}
-  Stable functions are closed under identities and composition.
-\end{lemma}
 
@@
 
-Stability has an alternative definition in terms of Galois connections, which will be more useful for what follows. This characterisation is due to Paul Taylor\bob{cite}.
+Stability has an alternative definition in terms of Galois connections, which will be more useful for what follows.
+This characterisation is due to Paul Taylor\bob{cite}.
+We first formally define Galois connections, which we used informally in \exref{introduction-example}.
+
+\begin{definition}[Galois connection]
+  \label{def:galois-connection}
+  Suppose $X$ and $Y$ are posets.
+A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
+We refer to $f$ as the left adjoint and $g$ as the right adjoint.
+\end{definition}
 
 \begin{lemma}
-  A monotone function $f : X \to Y$ is stable if and only if for all
-  $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$
-  has a left Galois adjoint.
+  A monotone function $f : X \to Y$ is stable if and only if for all $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$ has a left Galois adjoint.
 \end{lemma}
@@
 
-The converse to this lemma (that functions that preserve meets in their forward part have a left Galois adjoint) is not true, as was demonstrated by the non-stable function in \autoref{ex:non-stable-function}.
+The converse to this lemma (that functions that preserve meets in their forward part have a left Galois adjoint) is not true, as was demonstrated by the non-stable function in \exref{non-stable-function}.
 In the case when the posets $\downset{x}$ are \emph{complete}, and $f_x$ preserves infinitary meets, then we are guaranteed a left Galois adjoint.
@@
 Requiring completeness to compute backwards maps of approximations would require that complete joins are always computable, which will not be the case for approximations of functions.
+
+\begin{proposition}
+  L-posets and stable functions form a category. \bob{and the main point is that composition of stability witnesses looks like a chain rule!}
+\end{proposition}
 

### 2025-06-09 main/introduction.tex
Lifting.
minor edits to approx-as-tangents

@@
 \begin{example}
+  \label{ex:introduction-example}
   The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then to compute the sum of the numbers:
@@
 
+\bob{Have stolen this text for \autoref{sec:diff-stab-summary}}
 \begin{itemize}

### 2025-06-09 main/approx-as-tangents.tex
Some blurb on CMon-enrichment.

--- 2025-06-07 main/approx-as-tangents.tex
@@
 \section{Approximations as Tangents}
+\label{sec:approx-as-tangents}
 
-We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a dentational setting.
+We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting.
 From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.

### 2025-06-09 main/first-order.tex
Some blurb on CMon-enrichment.

@@
 
-\subsection{Biproducts and Semi-Additive Categories}
+\subsection{$\CMon$-Categories and Biproducts}
 \label{sec:biproducts}
+
+For differentiable functions, we can combine linear approximations by adding tangent vectors. For stable
+functions, we can combine approximants by taking their meet. The additive structure in both cases form a
+commutative monoid with an addition + and zero element 0. If this additive structure is also present in the
+hom-sets $\cat{C}(X,Y)$ of a category, and composition preserves it, $\cat{C}$ is enriched in $\CMon$, the
+category of commutative monoids.
 
@@
 structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual definition
-of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define within the
-language a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
+of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define (within the
+language) a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
 with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat}

### 2025-06-09 main/first-order.tex
CMon.

@@
 functions, we can combine approximants by taking their meet. The additive structure in both cases form a
-commutative monoid with an addition + and zero element 0. If this additive structure is also present in the
-hom-sets $\cat{C}(X,Y)$ of a category, and composition preserves it, $\cat{C}$ is enriched in $\CMon$, the
+commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
+category of commutative monoids, if every hom-set $\cat{C}(X,Y)$ has a similar additive structure, providing
+an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and if composition is
+\emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to
+\Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and
+$\Hom{\cat{C}}{X}{Y}$ separately:
+
+\begin{salign*}
+f \comp \zero = f = \zero \comp f
+\end{salign*}
+\begin{salign*}
+(f + g) \comp h &= (f \comp h) + (g \comp h) \\
+h \comp (f + g) &= (h \comp f) + (h \comp g)
+\end{salign*}
+
+We start with the idea of a biproduct and semi-additive categories, which are always enriched in $\CMon$, the
 category of commutative monoids.
-
-\todo{Introduce $\CMon$-category first.} We start with the idea of a biproduct and semi-additive categories,
-which are always enriched in $\CMon$, the category of commutative monoids.
 

### 2025-06-09 main/first-order.tex
CMon.

@@
 functions, we can combine approximants by taking their meet. The additive structure in both cases form a
-commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
-category of commutative monoids, if every hom-set $\cat{C}(X,Y)$ has a similar additive structure, providing
-an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and if composition is
-\emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to
-\Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and
-$\Hom{\cat{C}}{X}{Y}$ separately:
+commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$ (or a
+$\CMon$-category), the category of commutative monoids, if every hom-set $\cat{C}(X,Y)$ has a similar additive
+structure, providing an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and if
+composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor
+\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in
+$\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
 

### 2025-06-09 main/first-order.tex
CMon.

--- 2025-06-09 main/first-order.tex
@@
 commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
-category of commutative monoids, if every hom-set $\cat{C}(X,Y)$ has a similar additive structure, providing
-an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and if composition is
+category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ has a similar additive structure, providing
+an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and composition is
 \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to

### 2025-06-09 main/first-order.tex
Clarify CMon monoidal.

--- 2025-06-09 main/first-order.tex
@@
 functions, we can combine approximants by taking their meet. The additive structure in both cases form a
-commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$ (or a
-$\CMon$-category), the category of commutative monoids, if every hom-set $\cat{C}(X,Y)$ has a similar additive
-structure, providing an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and if
+commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
+(monoidal) category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ has a similar additive
+structure, providing an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and
 composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor

### 2025-06-09 main/approx-as-tangents.tex
comment about cartesian closure in L-poset

--- 2025-06-09 main/approx-as-tangents.tex
@@
 
-Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$. Each smooth function induces maps of the (co)tangent spaces:
+Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$.
+Manifolds and smooth functions form a category $\Man$.
+Each smooth function induces maps of the (co)tangent spaces:
 \begin{itemize}
@@
     In terms of \GPS the $\mathrm{gustave}$ function does not present a problem.
-For any particular run (i.e., input-output pair) of the program, there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
+For any particular run (i.e., input$\mapsto$output pair), there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
 
@@
 
-Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets (for forward approximation maps) and joins (for backwards approximation maps) exist:
+Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets, preserved by forward approximation maps, and joins, preserved by backwards approximation maps, exist:
 
@@
 
-The following lemma is an instance of standard facts about Galois connections preserving meets and joins:
+This lemma is an instance of standard facts about Galois connections
+preserving meets and joins:
 
@@
 
+\bob{L-poset is has products and coproducts, but is not closed.
+To be closed would require each $\downset{x}$ to be a \emph{complete} lattice.
+However, this is tricky to implement as we need an infinitary operation to compute the reverse approximation map for function types.
+See \cite{amadio-curien} for the proof that bounded-complete lattices with stable functions form a cartesian closed category.}
+
 \subsection{Summary}
@@
 
-We now summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
+Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
 

### 2025-06-09 main/first-order.tex
Clarify CMon monoidal.

--- 2025-06-09 main/first-order.tex
@@
 commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
-category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ has a similar additive structure, providing
-an operation + to add morphisms together plus a zero morphism $0: X \to Y$, and composition is
-\emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to
-\Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and
-$\Hom{\cat{C}}{X}{Y}$ separately:
+(monoidal) category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ provides an operation + to add
+morphisms together plus a zero morphism $0: X \to Y$, and composition is \emph{bilinear}, i.e.~given by a
+family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
+preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
 
@@
 \end{salign*}
+
+\subsubsection{Biproducts}
+
+In any $\CMon$-category,
+
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_X \comp \biinj_X &= \id_X \\
+   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
+\end{salign*}
+\end{center}
+\end{minipage}%
+\begin{minipage}[t]{0.45\textwidth}
+\begin{center}
+\begin{salign*}
+   \biproj_Y \comp \biinj_Y &= \id_Y \\
+   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
+\end{salign*}
+\end{center}
+\end{minipage}
+
+\begin{salign*}
+(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
+\end{salign*}
+
+\hrule
 
@@
 \subsubsection{Biproduct laws}
-It is also easy to show that in a semi-additive category the following equations hold:
-%\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
 
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-\vspace{3mm}
-\noindent It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
+It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
 in $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or

### 2025-06-09 main/first-order.tex
Pared back version of biproducts.

@@
 
-For differentiable functions, we can combine linear approximations by adding tangent vectors. For stable
-functions, we can combine approximants by taking their meet. The additive structure in both cases form a
-commutative monoid with an addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
+For differentiable functions, we can combine linear approximations by adding tangent vectors; for stable
+functions, we can combine approximants by taking their meet. In both cases the additive structure is given by
+a commutative monoid with addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
 (monoidal) category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ provides an operation + to add
@@
 
-In any $\CMon$-category,
+In any $\CMon$-category, any products and coproducts coincide.
 
+\begin{proposition}
+\label{prop:biproducts:from-product-or-coproduct}
+In any $\CMon$-category:
+\begin{enumerate}
+\item If $X \times Y$ is a product then it is also coproduct, with $\inj_X = \prodM{\id_X}{0_{X,Y}}$ and
+$\inj_Y = \prodM{0_{X,Y}}{\id_Y}$.
+\item If $X \coprod Y$ is a coproduct then it is also a product, with $\pi_1 = \coprodM{\id_X}{0_{Y,X}}$ and
+$\pi_2 = \coprodM{0_{X,Y}}{\id_Y}$.
+\item $X$ is terminal iff $X$ is initial.
+\end{enumerate}
+\end{proposition}
+
+An object that is both terminal and initial is called a \emph{zero} object; an object that is both a product
+and a coproduct is called a \emph{biproduct}. More directly, a biproduct is as an object $X \biprod Y$
+together with morphisms
+
+\begin{center}
+\begin{tikzcd}
+   X \arrow[r, "\biinj_X", shift left] &
+   X \biprod Y \arrow[l, "\biproj_X", shift left] \arrow[r, "\biproj_Y"', shift right] &
+   Y \arrow[l, "\biinj_Y"', shift right]
+\end{tikzcd}
+\end{center}
+
+\noindent satisfying
+
+\vspace{-3mm}
 \begin{minipage}[t]{0.45\textwidth}
@@
 
-\hrule
+\noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
+with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial.
+Since $X \iso 0 \biprod X$, Cartesian closure implies $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow
+Y} \iso 1$, and all objects are isomorphic.
 
-We start with the idea of a biproduct and semi-additive categories, which are always enriched in $\CMon$, the
-category of commutative monoids.
-
-\begin{definition}[Semi-additive category]
-\label{def:biproducts:semi-additive}
-A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
-and injections) give an isomorphism
-\[\textstyle X \coprod Y \iso X \times Y\] that is natural in both variables.
-\end{definition}
-
-The product/coproduct is called a \emph{biproduct}, with the biproduct structure denoted by $(\biprod, 0)$ and
-projections $\biproj_X, \biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$. The unit $0$ is both terminal and
-initial and is called a \emph{zero} object. A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any
-two morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
-forming a morphism $f + g: X \to Y$. Diagrammatically:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
-\end{tikzcd}
-\end{center}
-
-Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-$\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-\coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-composing the unique maps in and out of the zero object:
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
-\end{tikzcd}
-\end{center}
-
-It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero_{X,Y} = 0_{X,Z} = \zero_{Y,Z} \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-\subsubsection{Biproduct laws}
-
-It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched
-in $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
-coproducts in $\cat{C}$ are necessarily biproducts.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-In a category $\cat{C}$ enriched in $\CMon$:
-\begin{enumerate}
-\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \inj_1, \inj_2)$ is a coproduct, where
-$\inj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\inj_Y = \prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
-\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \pi_1, \pi_2)$ is a product, where
-$\pi_1 = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and $\pi_2 = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to
-Y$.
-\item If $X$ is terminal (resp.~initial) then $X$ is initial (resp.~terminal).
-\end{enumerate}
-\end{proposition}
-
-A biproduct category $\cat{C}$ cannot be Cartesian closed without being trivial. If $0 \times X \iso X$ then
-$\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$, and so $\cat{C}$ is terminal, with only a
-single object up to isomorphism.
+% in any such category the canonical projections and injections give an isomorphism $\textstyle X \coprod Y
+% \iso X \times Y$ that is natural in both variables
+%
+% A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any two morphisms $f, g: X \to Y$ in
+% $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X \to
+% Y$. Diagrammatically:
+%
+% \begin{center} \begin{tikzcd} X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y
+% \arrow[r, "\codiag"] & Y \end{tikzcd} \end{center}
+%
+% Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
+% $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
+% The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
+% \coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
+% composing the unique maps in and out of the zero object:
+%
+% \begin{center} \begin{tikzcd} X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y \end{tikzcd} \end{center}
+%
+% It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
+% hom-object $\cat{C}(X,Y)$ is an object in $\CMon$.
+%
+% \subsubsection{Biproduct laws}
+%
+% It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched in
+% $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
+% coproducts in $\cat{C}$ are necessarily biproducts.
 

### 2025-06-09 main/first-order.tex
Much pared back FinVect.

@@
 
-The category $\FinVect$ of finite vector spaces over $\RR$ (with addition of real numbers written $a + b$ and
-multiplication written $a \mult b$) is semi-additive.
+The category $\FinVect$ of finite vector spaces over $\RR$ is semi-additive.
 
-\begin{definition}[Finite dimensional vector space over $\RR$]
-Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
-component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
-is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
-vector operations are compatible with the field operations in that the following equations hold \todo{probably
-more detail than we need to go into}:
-
-\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   1 \mult v &= v \\
-   (a \mult b) \mult v &= a \mult (b \mult v)
-\end{align*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
-   a \mult (u + v) &= (a \mult u) + (b \mult v)
-\end{align*}
-\end{center}
-\end{minipage}
+\begin{definition}
+Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces
+$(\RR^n, +, \mult)$, and as morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) +
+f(v)$ and $f(a \mult v) = a \mult f(v)$.
+% with $f$ preserving the zero and additive inverse structure automatically.
 \end{definition}
 
-\vspace{1mm}
-\noindent Composition is with bilinear. The direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian
-product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
-\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$.
+%  with addition of real numbers written $a + b$ and multiplication written $a \mult b$
 
-\begin{definition}
-Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces, and as
-morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
-\mult f(v)$, with $f$ preserving the abelian group structure automatically.
-\end{definition}
+% \begin{definition}[Finite dimensional vector space over $\RR$]
+% Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
+% component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
+% is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
+% vector operations are compatible with the field operations in that the following equations hold:
+%
+% \vspace{-4mm}
+% \begin{minipage}[t]{0.45\textwidth}
+% \begin{center}
+% \begin{align*}
+%    1 \mult v &= v \\
+%    (a \mult b) \mult v &= a \mult (b \mult v)
+% \end{align*}
+% \end{center}
+% \end{minipage}%
+% \begin{minipage}[t]{0.45\textwidth}
+% \begin{center}
+% \begin{align*}
+%    (a + b) \mult v &= (a \mult v) + (b \mult v) \\
+%    a \mult (u + v) &= (a \mult u) + (b \mult v)
+% \end{align*}
+% \end{center}
+% \end{minipage}
+% \end{definition}
 
-$\FinVect$ is enriched in $\Ab$, the category of abelian groups \todo{Introduce $\CMon$-enrichment first, then
-perhaps $\Ab$-enrichment parenthetically}. For any objects $V = \RR^m$ and $W = \RR^n$ the hom-set
-$\FinVect(V,W)$ is an abelian group, with identity $0_{V,W}$ given by the constant map $v \mapsto 0_W$ and
-addition and additive inverse given pointwise:
-
-\vspace{-4mm}
-\begin{center}
-\begin{minipage}[t]{0.35\textwidth}
-\begin{align*}
-(f + g)(v) &= f(v) + g(v)
-\end{align*}
-\end{minipage}%
-\begin{minipage}[t]{0.35\textwidth}
-\begin{align*}
-(-f)(v) &= -f(v)
-\end{align*}
-\end{minipage}
-\end{center}
+$\FinVect$ is enriched in $\CMon$ (in fact in $\Ab$, the category of abelian groups, but $\CMon$ will do for
+our purposes). For any objects $V = \RR^m$ and $W = \RR^n$, the hom-set $\FinVect(V,W)$ provides pointwise
+addition of morphisms $(f + g)(v) = f(v) + g(v)$; the constant map $v \mapsto 0$ is a zero morphism, and
+composition is bilinear. Moreover the direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian product
+is a biproduct, with $\biinj_{V} = \prodM{\id}{0}$ and $\biinj_{W} = \prodM{0}{\id}$.
 

### 2025-06-09 main/first-order.tex
Now paring back LatGal.

@@
 
-\todo{Rework to account for new ordering.} Galois connections are pairs of monotone functions $f: Y \to X$ and
-$g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
-and $g$ the best approximation from above to an inverse of $f$. The category $\LatGal$ of bounded lattices and
-Galois connections is semi-additive.
+The category $\LatGal$ of bounded lattices and Galois connections is semi-additive.
 
 \begin{definition}[Galois connection]
-Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
-$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
+A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$
+between posets satisfying $y \leq g(x) \iff f(y) \leq x$.
 \end{definition}
 
-\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
-adjunction between poset categories, with monotone $f$ and $g$ regarded as functors; $f$ is usually referred
-to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections compose
-component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition.
+\noindent A Galois connection $f \adj g: X \to Y$ is an adjunction between poset categories, regarding
+monotone $f$ and $g$ as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
+the \emph{lower} (left) adjoint. Galois connections are closed under component-wise composition and identities
+$\id_X \adj \id_X$.
 

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

@@
 
-\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
-in $\LatGal$, $g$ is a (bounded) \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice
-structure $(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join,
-\bot)$.
+\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
+Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
+the join-semilattice structure $(X, \join, \bot)$.
 
-\todo{Again, introduce $\CMon$-enrichment first and relegate $\SemiLat$-enrichment to parenthetical remark.}
-The meet-preserving maps from $X$ to $Y$ themselves form a meet-semilattice. The operation $\meet$ on such
-maps is given pointwise, so that $(f \meet g)(x) = f(x) \meet g(x): X \to Y$, and preserves meets; the
-constant map $\top_{X,Y}: X \to Y$ sending any $x \in X$ to $\top_Y$, which also preserves meets, provides the
-unit. Dually the join-preserving maps from $Y$ to $X$ have a join-semilattice structure given pointwise by
-$\join$ and the constant map $\bot_{Y,X}: Y \to X$ sending any $y \in Y$ to $\bot_X$. Since these two
-constructions come in adjoint pairs, the hom-set $\LatGal(X,Y)$ forms a bounded meet semilattice with unit
-$\top_{X,Y}$ given by the Galois connection $\bot_{Y,X} \adj \top_{X,Y}: X \to Y$ and meet of Galois
-connections $(f \adj g) \meet (f' \adj g') = (f \join f') \adj (g \meet g'): X \to Y$. Thus $\LatGal$ is
-enriched in $\SemiLat$, the category of semilattices and semilattice homomorphisms.
+$\LatGal$ is enriched in $\CMon$ (in fact in $\SemiLat$, the category of semilattices, but again $\CMon$ will
+do for our purposes). The meet-preserving maps from $X$ to $Y$ can be added pointwise using $\meet$, since $(f
+\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero;
+dually the join-preserving maps from $Y$ to $X$ can be added pointwise using $\join$, with the constant map
+$\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, the hom-set
+$\LatGal(X,Y)$ forms a commutative monoid with $0$ given by the Galois connection $\bot \adj \top$ and
+addition $(f \adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$.
 

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

@@
 $\LatGal(X,Y)$ forms a commutative monoid with $0$ given by the Galois connection $\bot \adj \top$ and
-addition $(f \adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$.
+addition $(f \adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$, with composition bilinear.
 

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

--- 2025-06-09 main/first-order.tex
@@
 Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
-the join-semilattice structure $(X, \join, \bot)$.
+the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps have their $\CMon$ structure
+structure (in fact $\SemiLat$ structure, but again $\CMon$ is sufficient): addition of maps is pointwise via
+$\meet$, since $(f \meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto
+\top_Y$ as a zero. Dually  the join-preserving maps can be added pointwise using $\join$, with the constant
+map $\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is
+enriched in $\CMon$. The hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f
+\adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is zero
+morphism, and composition is bilinear.
 
-$\LatGal$ is enriched in $\CMon$ (in fact in $\SemiLat$, the category of semilattices, but again $\CMon$ will
-do for our purposes). The meet-preserving maps from $X$ to $Y$ can be added pointwise using $\meet$, since $(f
-\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero;
-dually the join-preserving maps from $Y$ to $X$ can be added pointwise using $\join$, with the constant map
-$\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, the hom-set
-$\LatGal(X,Y)$ forms a commutative monoid with $0$ given by the Galois connection $\bot \adj \top$ and
-addition $(f \adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$.
-
-The projections $\pi_1: X \times Y \to X$ and $\pi_2: X \times Y \to Y$, where $X \times Y$ denotes the
-product of lattices, have both upper and lower adjoints; this means that $X \times Y$, which we shall
-hereafter write as $X \biprod Y$, is both a product and a coproduct, with projections $\biproj_X$ and
-$\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
+The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$, have both upper and lower adjoints;
+this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is both a product and a
+coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
 

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

@@
 
-The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$, have both upper and lower adjoints;
-this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$, is both a product and a
-coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$ given by:
+$\LatGal$ also has biproducts. The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$ have
+both upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$,
+is both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and
+$\biinj_Y$ given by:
 

### 2025-06-09 main/first-order.tex
LatGal as a CMon-category.

@@
 $\LatGal$ also has biproducts. The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$ have
-both upper and lower adjoints; this means that $X \times Y$, which we shall hereafter write as $X \biprod Y$,
-is both a product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and
-$\biinj_Y$ given by:
+both upper and lower adjoints; this means that $X \times Y$ (which we shall write as $X \biprod Y$), is both a
+product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$
+given by:
 
@@
 \begin{align*}
-   \biproj_X = \prodM{\id_X}{\bot_Y} \adj \proj_1: X \biprod Y \to X \\
-   \biproj_Y = \prodM{\bot_X}{\id_Y} \adj \proj_2: X \biprod Y \to Y
+   \biproj_X = \prodM{\id}{\bot} \adj \proj_1: X \biprod Y \to X \\
+   \biproj_Y = \prodM{\bot}{\id} \adj \proj_2: X \biprod Y \to Y
 \end{align*}
@@
 \begin{align*}
-   \biinj_X = \pi_1 \adj \prodM{\id_X}{\top_Y} : X \to X \biprod Y \\
-   \biinj_Y = \pi_2 \adj \prodM{\top_X}{\id_Y}: Y \to X \biprod Y
+   \biinj_X = \pi_1 \adj \prodM{\id}{\top} : X \to X \biprod Y \\
+   \biinj_Y = \pi_2 \adj \prodM{\top}{\id}: Y \to X \biprod Y
 \end{align*}
@@
 
-\noindent The trivial 1-point lattice, which is both terminal and initial and we write as $0$, is the unit for
-$\biprod$.
+\noindent with the trivial 1-point lattice as a zero object (terminal and initial).
 

### 2025-06-09 main/first-order.tex
Some cleanup of Grothendieck.

@@
 
-\todo{Rework.} The goal now is to equip the interpretation of a (first-order) program with a Galois connection
-relating approximations of its inputs to approximations of its output. Since the specific lattices of
-approximations depend on the value being approximated, the semantics of every type $\tau$ must have both a
-conventional component as well as an approximating component, i.e.~a set $X$ which serves as the usual
-interpretation of $\tau$, plus an associated lattice of approximations $\partial X(x)$ for every point $x \in
-X$. The semantics of an expression $e$ of type $\sigma$ with a single free variable of type $\tau$ (with
-$\tau$ and $\sigma$ denoting $X$ and $Y$) must then also have two components: a function $f: X \to Y$ giving
-the conventional interpretation of $e$, plus a family of Galois connections $\partial f(x): \partial X(x) \to
-\partial Y(f(x))$ for every $x \in X$ which can be used for slicing over that expression.
+We now consider how to adapt the CHAD framework~\cite{vákár22,nunes2023} to interpret programs as functions
+along with their forward and backward approximation maps. Since the appropriate lattice of approximation
+typically depends on the value being approximated, we will want to interpret a type $\tau$ as a pair $(X,
+\partial X)$ where $X$ provides the conventional interpretation as a set and $\partial X$ provides an
+appropriate lattice of approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type
+$\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
+conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, where $f$ provides the
+conventional interpretation of $e$ as a function $X \to Y$, and $\partial f$ is a family of Galois connections
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$, taking approximants to approximants.
 
-This pattern of objects and morphisms is captured by the construction used in CHAD~\cite{vákár22,nunes2023} to
-model automatic differentiation, namely the \emph{Grothendieck construction} for a particular set-indexed
-category $T: \Set^\op \to \Cat$. For automatic differentiation, $T(X)$ will be the functor category
-$\Func{X}{\FinVect}$, regarded as the category of $X$-indexed families of finite vector spaces, with indexed
-linear maps as morphisms. For Galois slicing, we will take $T(X) = \Func{X}{\LatGal}$, the category of
-$X$-indexed families of bounded lattices, with indexed Galois connections as morphisms.
+This pattern of objects and morphisms is captured by the construction used in CHAD to model automatic
+differentiation, namely the \emph{Grothendieck construction} for a particular set-indexed category $T:
+\Set^\op \to \Cat$. For automatic differentiation, $T(X)$ will be the functor category $\Func{X}{\FinVect}$,
+regarded as the category of $X$-indexed families of finite vector spaces, with families of linear maps as
+morphisms. For Galois slicing, we will take $T(X) = \Func{X}{\LatGal}$, the category of $X$-indexed families
+of bounded lattices, with families of Galois connections as morphisms.
 
 In its general form, the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
-\cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, assembles all the categories $T(X)$
-(for objects $X$ in $\cat{C}$) into a single category, together with morphisms that account for both internal
-structure and reindexing along morphisms $f: X \to Y$ in $\cat{C}$.
+\cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, incorporates all the categories
+$T(X)$ (for objects $X$ in $\cat{C}$) into a single category, together with morphisms that account for both
+internal structure and reindexing along morphisms $f: X \to Y$ in $\cat{C}$.
 
@@
 \label{def:Grothendieck}
-Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction} for $T$ is the
-category $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) which has as objects, all pairs $(X,
+Suppose $T: \cat{C}^\op \to \Cat$ an indexed category. The \emph{Grothendieck construction}
+$\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) is the category which has as objects, all pairs $(X,
 \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
@@
 \noindent The $\partial X$ notation used here is intended to suggest the connection between the Grothendieck
-construction and the idea of tangent spaces and derivatives. To understand how the $\partial f$ components
-compose, consider any $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ and $(g, \partial g): (Y,
-\partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$. We have:
+construction and the idea of tangent spaces and derivatives. For any $(f, \partial f): (X, \partial X) \to (Y,
+\partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$, the composition
+$(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is given by $(g \comp f,
+T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to ``reindex'' $\partial g$ along $f$,
+yielding a morphism $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial
+Z)$.
 
-\begin{itemize}
-\item $\partial f: \partial X \to T(f)(\partial Y)$, and
-\item $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial Z)$
-\end{itemize}
-
-\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
-composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
-\comp f, T(f)(\partial g) \comp \partial f)$.
+% which can then be composed with $\partial f: \partial X \to T(f)(\partial Y)$.
 

### 2025-06-09 main/first-order.tex
Minor clean-up.

@@
 $\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
-conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, where $f$ provides the
-conventional interpretation of $e$ as a function $X \to Y$, and $\partial f$ is a family of Galois connections
-$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$, taking approximants to approximants.
+conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, with $f$ providing the
+conventional interpretation as a function $X \to Y$, and $\partial f$ as a family of Galois connections
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$ taking approximants to approximants.
 
@@
 
-\subsection{Category of Families}
+\subsubsection{Category of Families}
 \label{sec:Fam}
 
-Now we formalise $X$-indexed families as functor categories $\Func{X}{\cat{C}}$ and consider the specific case
-of the Grothendieck construction for set-indexed categories $\Func{-}{\cat{C}}: \Set^\op \to \Cat$.
-
-\begin{definition}[Category of $X$-indexed families]
-For any set $X$ (regarded as a discrete category) and any category $\cat{C}$, recall that the functor
-category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
-objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
-$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ has only identity morphisms.
-\end{definition}
+For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$. In
+that case our indexed categories $T$ have the form $\Func{-}{\cat{C}}: \Set^\op \to \Cat$, associating to each
+set $X$ the functor category $\Func{X}{\cat{C}}$ which has as objects all functors $F: X \to \cat{C}$, namely
+$X$-indexed families of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x):
+F(x) \to G(x)$ in $\cat{C}$, with naturality trivial because $X$ has only identity morphisms.
 

### 2025-06-09 main/first-order.tex
Some minor improvements to families construction.

@@
 
-\noindent The $\partial X$ notation used here is intended to suggest the connection between the Grothendieck
+\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
 construction and the idea of tangent spaces and derivatives. For any $(f, \partial f): (X, \partial X) \to (Y,
@@
 
-For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$. In
-that case our indexed categories $T$ have the form $\Func{-}{\cat{C}}: \Set^\op \to \Cat$, associating to each
-set $X$ the functor category $\Func{X}{\cat{C}}$ which has as objects all functors $F: X \to \cat{C}$, namely
-$X$-indexed families of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x):
-F(x) \to G(x)$ in $\cat{C}$, with naturality trivial because $X$ has only identity morphisms.
+For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$ and
+each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
+of objects of $\cat{C}$ (regarded as functors $F: X \to \cat{C}$) , and as morphisms $F \to G$, all families
+of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only identity
+morphisms).
 
-Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.
-
-\begin{proposition}
-If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
-pointwise.
-\end{proposition}
+% Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.
+%
+% \begin{proposition}
+% If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
+% pointwise.
+% \end{proposition}
 
 \begin{definition}[Reindexing]
-For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
-\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
-$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
-family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
-$\reindex{\eta}{f}(x) = \eta(f(x))$. \todo{Should be clearer that $\reindex{-}{f}$ is actually
-$\Func{f}{\cat{C}}$.}
+For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f} = \Func{f}{\cat{C}}:
+\Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the
+$X$-indexed family $\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete
+categories) and any family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to
+\reindex{G}{f}$ with $\reindex{\eta}{f}(x) = \eta(f(x))$.
 \end{definition}

### 2025-06-09 main/approx-as-tangents.tex
Gentler notation of syntactic identifiers.

--- 2025-06-09 main/approx-as-tangents.tex
@@
 \begin{example}[Conditionally Multiplicative Functions]
+\label{ex:strict-short-circuit}
   To see the effect of requiring preservation of bounded meets,

### 2025-06-09 main/first-order.tex
Some minor improvements to families construction.
Gentler notation of syntactic identifiers.

@@
 
-The action of the families construction, regarded as a covariant 2-functor $\Cat \to \Cat$, sends functors $F:
-\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that use $F$ to reassign the
-target category by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and
-morphisms $(f, \partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural
-transformation with $(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on
-$\cat{C}$, then $\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by lifting the original monad $F:
-\cat{C} \to \cat{C}$ pointwise over families.
+The families construction, regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F:
+\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category
+by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f,
+\partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where
+$(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then
+$\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by lifting the original monad $F: \cat{C} \to
+\cat{C}$ pointwise over families.
 
@@
 ``leaves'' of the approximation lattices would provided by a model of $\syn{nat}$ and its operations in
-$\Fam(\LatGal)$, for example assigning $\syn{nat}$ to the object $(\Nat,n \mapsto \Two)$, where the family is
-fully present ($\top$) or fully absent ($\bot$) from the approximation (as shown in the example in component
-is constantly $\Two = (\bot,\top)$, the two-point lattice indicating whether a given natural number
-\secref{introduction}). These ``leaf'' objects would then be combined by the interpretation of composite types
-like sums and products, providing suitable derived approximation structures for composite values.
+$\Fam(\LatGal)$. For example, we might assign $\syn{nat}$ to the object $(\Nat,n \mapsto \{n, \bot\})$, where
+the family component is the two-point lattice indicating whether a given natural number is fully present ($n$)
+or fully absent ($\bot$) from the approximation, as shown in the example in \secref{introduction}, and we
+might then assign $+$ to the morphism $(\hat+,\partial(\hat+))$, with $\partial(\hat+)$ defined as
+\begin{displaymath}
+    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
+      \partial(\hat+)&n&m&=&n \mathbin{\hat+} m \\
+      \partial(\hat+)&\bot&\_&=&\bot \\
+      \partial(\hat+)&\_&\bot&=&\bot
+    \end{array}
+\end{displaymath}
+These elementary approximations would then be combined by the interpretation of composite types like sums and
+products, providing suitable derived approximation structures for composite values.
 
 One might also like to allow the programmer to embed additional ``approximation points'' into composite
-structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual definition
-of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define (within the
-language) a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
-with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat}
-\to \syn{L}\;\syn{nat}$ expressible as
+structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
+the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then define
+(within the language) a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat}
+\times \alpha)$, along with an approximating version of multiplication $*_{\text{appr}}:
+\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ given as as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
 0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].
+
+\noindent This would allow the programmer to emulate the short-circuiting behaviour of $\mathrm{strictOr}$ in
+\exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
+matrix convolutions, without baking this specific approximating behaviour into the interpretation of
+individual operators.
 

### 2025-06-09 main/first-order.tex
Onto AutoDiff.

@@
 \times \alpha)$, along with an approximating version of multiplication $*_{\text{appr}}:
-\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ given as as
+\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
@@
 
-\noindent This would allow the programmer to emulate the short-circuiting behaviour of $\mathrm{strictOr}$ in
-\exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
+\noindent This would allow the programmer to emulate the short-circuiting behaviour of $\mathrm{strictOr}$
+from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
 matrix convolutions, without baking this specific approximating behaviour into the interpretation of
@@
 Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
-setup, via the total category $\Fam(\FinVect)$. We recapitulate the basic ideas here.
+setup, via the total category $\Fam(\FinVect)$. First we recall the chain rule for the forward derivative
+$\pushf{f}_x$ and backward derivative $\pullf{f}_x$, which says that derivatives respect
+composition~\cite{spivak65}.
 
-\subsubsection{Derivatives of Differentiable Functions}
-
-For flat manifolds like $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically
-isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
-\RR^n$ at a point $x$ have the following form \todo{Can probably delete now as covered in
-\secref{approx-as-tangents}}:
-
-\begin{itemize}
-\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of $f$ at $x \in \RR^m$ is the
-unique linear map $\RR^m \linearto \RR^n$ given by the Jacobean matrix of $f$ at $x$.
-\item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
-$\RR^n \linearto \RR^m$ given by the transpose (adjoint) of the Jacobean matrix of $f$ at $x$.
-\end{itemize}
+% \subsubsection{Derivatives of Differentiable Functions}
+%
+% For flat manifolds like $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically
+% isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
+% \RR^n$ at a point $x$ have the following form \todo{Can probably delete now as covered in
+% \secref{approx-as-tangents}}:
+%
+% \begin{itemize}
+% \item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of $f$ at $x \in \RR^m$ is the
+% unique linear map $\RR^m \linearto \RR^n$ given by the Jacobean matrix of $f$ at $x$.
+% \item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
+% $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of the Jacobean matrix of $f$ at $x$.
+% \end{itemize}
 
@@
 
-Derivatives respect composition~\cite{spivak65}. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. Then
+Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. Then
 for any $x \in \RR^m$ we have:
@@
 forward derivative of a composite function, without have to recompute the function itself, as long as all
-atomic operations provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$
-(\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to make it possible to
-compute forwards slices without having to reexecute the program at the same time.
+atomic operations provide derivatives.
+
+% An efficient implementation of \GPS based on $\Fam(\LatGal)$
+% (\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to make it possible to
+% compute forwards slices without having to reexecute the program at the same time.
 
@@
 \partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
-\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ recomputing $f(x)$ in
-order to select the appropriate pushforward.
+\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$
+
+% with the reindexing along $f$ recomputing $f(x)$ in order to select the appropriate pushforward.
 

### 2025-06-09 main/first-order.tex
Minor tweaks.

@@
 
-Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
-setup, via the total category $\Fam(\FinVect)$. First we recall the chain rule for the forward derivative
-$\pushf{f}_x$ and backward derivative $\pullf{f}_x$, which says that derivatives respect
-composition~\cite{spivak65}.
+As per the CHAD work, automatic differentiation for first-order programs can also be recovered in this setup,
+using the total category $\Fam(\FinVect)$ instead of $\Fam(\LatGal)$. First we recall the chain rule for
+forward derivatives $\pushf{f}_x: \RR^m \linearto \RR^n$ and backward derivatives $\pullf{f}_x: \RR^n
+\linearto \RR^m$, which says that derivatives respect composition~\cite{spivak65}.
 
@@
 
-\subsubsection{Chain Rule}
-
-Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. Then
-for any $x \in \RR^m$ we have:
+\begin{proposition}[Chain Rule]
+Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$ differentiable. Then for any $x \in \RR^m$ we have:
 
@@
 \end{itemize}
+\end{proposition}
 
@@
 Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
-\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$
+\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
+f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
 

### 2025-06-09 main/first-order.tex
Tweaks to CMon categories.

@@
 
-For differentiable functions, we can combine linear approximations by adding tangent vectors; for stable
-functions, we can combine approximants by taking their meet. In both cases the additive structure is given by
-a commutative monoid with addition + and zero element 0. A category $\cat{C}$ is enriched in $\CMon$, the
-(monoidal) category of commutative monoids, when every hom-set $\cat{C}(X,Y)$ provides an operation + to add
-morphisms together plus a zero morphism $0: X \to Y$, and composition is \emph{bilinear}, i.e.~given by a
-family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
-preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+\subsubsection{$\CMon$-Categories}
+
+To say that a category $\cat{C}$ is enriched in $\CMon$, the (monoidal) category of commutative monoids, is to
+say that every hom-set $\cat{C}(X,Y)$ provides an operation + to add morphisms together plus a zero morphism
+$0: X \to Y$, and moreover that composition is \emph{bilinear}, i.e.~given by a family of morphisms
+$\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the
+additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
 
@@
 
-\subsubsection{Biproducts}
-
-In any $\CMon$-category, any products and coproducts coincide.
+In a $\CMon$-category, any products and coproducts coincide.
 
@@
 \begin{enumerate}
-\item If $X \times Y$ is a product then it is also coproduct, with $\inj_X = \prodM{\id_X}{0_{X,Y}}$ and
-$\inj_Y = \prodM{0_{X,Y}}{\id_Y}$.
-\item If $X \coprod Y$ is a coproduct then it is also a product, with $\pi_1 = \coprodM{\id_X}{0_{Y,X}}$ and
-$\pi_2 = \coprodM{0_{X,Y}}{\id_Y}$.
-\item $X$ is terminal iff $X$ is initial.
+\item If $X \times Y$ is a product then it is also coproduct, with $\inj_X = \prodM{\id}{0}$ and $\inj_Y =
+\prodM{0}{\id}$.
+\item If $X \coprod Y$ is a coproduct then it is also a product, with $\pi_1 = \coprodM{\id}{0}$ and $\pi_2 =
+\coprodM{0}{\id}$.
+\item $X$ is terminal iff it is initial.
 \end{enumerate}
@@
 An object that is both terminal and initial is called a \emph{zero} object; an object that is both a product
-and a coproduct is called a \emph{biproduct}. More directly, a biproduct is as an object $X \biprod Y$
-together with morphisms
+and a coproduct is called a \emph{biproduct}.
+
+\subsubsection{Biproducts}
+
+More directly, a biproduct is as an object $X \biprod Y$ together with morphisms
 

### 2025-06-09 main/first-order.tex
Tweaks.

@@
 \noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
-with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial.
-Since $X \iso 0 \biprod X$, Cartesian closure implies $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow
-Y} \iso 1$, and all objects are isomorphic.
+with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial:
+since $X \iso 0 \biprod X$, Cartesian closure would imply $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
+\Rightarrow Y} \iso 1$, and therefore that all objects are isomorphic.
 
@@
 Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
-the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps have their $\CMon$ structure
-structure (in fact $\SemiLat$ structure, but again $\CMon$ is sufficient): addition of maps is pointwise via
-$\meet$, since $(f \meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto
-\top_Y$ as a zero. Dually  the join-preserving maps can be added pointwise using $\join$, with the constant
-map $\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is
-enriched in $\CMon$. The hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f
-\adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is zero
-morphism, and composition is bilinear.
+the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps are enriched in $\CMon$ (in fact
+in $\SemiLat$, but again $\CMon$ is sufficient): addition of maps is computed pointwise via $\meet$, since $(f
+\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero.
+Dually the join-preserving maps can be added pointwise using $\join$, with the constant map $\bot = y \mapsto
+\bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is enriched in $\CMon$. The
+hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f \adj g) + (f' \adj g') =
+(f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is a zero morphism, and composition is
+bilinear.
 
@@
 each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
-of objects of $\cat{C}$ (regarded as functors $F: X \to \cat{C}$) , and as morphisms $F \to G$, all families
-of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only identity
-morphisms).
+of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
+$\cat{C}$ (with naturality trivial because $X$ has only identity morphisms).
 
@@
 
-\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all sets $X$ paired with indexed families
+\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all pairs of sets $X$ and indexed families
 $\partial X: X \to \cat{C}$ and the morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are all
-functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
+pairs of functions $f: X \to Y$ and morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
 $\Func{X}{\cat{C}}$.

### 2025-06-09 main/first-order.tex
Tweaks.

@@
 structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
-the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then define
-(within the language) a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat}
-\times \alpha)$, along with an approximating version of multiplication $*_{\text{appr}}:
-\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
+the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the
+language) define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times
+\alpha)$, along with an approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat}
+\times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
@@
 from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
-matrix convolutions, without baking this specific approximating behaviour into the interpretation of
+matrix convolutions, without baking this specific choice of approximation into the interpretation of
 individual operators.
@@
 Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
-f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
+\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
 

### 2025-06-09 main/language.tex

--- 2025-06-06 notes/language.tex
@@
 
-We introduce a standard total functional language with a (reasonably) expressive type
+We introduce a total higher-order functional language with a (reasonably) expressive type
 system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types $\rho$ and sets
@@
 
-\figrefTwo{syntax}{typing} give the syntax and typing rules of the higher-order language.
+\figrefTwo{syntax}{typing} give the syntax and typing rules.
 
@@
 \begin{definition}[$\mu$-polynomial]
-Suppose $\cat{C}$ a category with finite coproducts $(\coprod, \Zero)$ and finite products $(\times, \One)$.
-Define $\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
+Suppose $\cat{C}$ a category with finite coproducts $(\coprod, 0)$ and finite products $(\times, 1)$. Define
+$\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
 \figref{mu-polynomial}.
@@
 To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\Sem$ with finite products $(\times, \One)$, finite coproducts $(\coprod, \Zero)$, exponentials
+category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(\coprod, 0)$, exponentials
 $\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the

### 2025-06-17 main/higher-order.tex
Better to define 'small product' rather than 'category with all small products'.

@@
 $\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
-composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed and
-thereby able to interpret higher-order programs directly. However the following is a special case of a
-construction due to \citet{nunes2023}:
+composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed. To
+interpret higher-order programs directly, we use a special case of a construction due to \citet{nunes2023}
+(\propref{higher-order:fam-CC} below). First we introduce notation for the product over a set-indexed family.
 
-\begin{proposition}
-Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
-$\Fam(\cat{C})$ is Cartesian closed.
-\end{proposition}
-
-A category $\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$
-of objects of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family
-of projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
-and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
+\begin{definition}[Small product]
+Suppose $\cat{C}$ a category and $X$ a
+set. A \emph{product} of the family $A = \{A_x\}_{x \in X}$ is an object $\prod_{X}A$ plus a family of
+projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$ and
+family of morphisms $f = \{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
 $\prodFam{f}$ making the following diagram commute for every $x \in X$:
@@
 \end{center}
+\end{definition}
+
+\begin{proposition}
+\label{prop:higher-order:fam-CC}
+Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
+$\Fam(\cat{C})$ is Cartesian closed.
+\end{proposition}
 

### 2025-06-18 main/first-order.tex
Correction.
Tweak.

@@
 
-To say that a category $\cat{C}$ is enriched in $\CMon$, the (monoidal) category of commutative monoids, is to
-say that every hom-set $\cat{C}(X,Y)$ provides an operation + to add morphisms together plus a zero morphism
-$0: X \to Y$, and moreover that composition is \emph{bilinear}, i.e.~given by a family of morphisms
-$\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the
-additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
+A category $\cat{C}$ is enriched in $\CMon$, the (monoidal) category of commutative monoids, if every hom-set
+$\cat{C}(X,Y)$ provides an operation + to add morphisms together, as well as a zero morphism $0: X \to Y$, and
+composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor
+\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in
+$\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
 
@@
 \begin{enumerate}
-\item If $X \times Y$ is a product then it is also coproduct, with $\inj_X = \prodM{\id}{0}$ and $\inj_Y =
+\item If $X \times Y$ is a product then it is also coproduct, with $\inj_1 = \prodM{\id}{0}$ and $\inj_2 =
 \prodM{0}{\id}$.

### 2025-06-18 main/first-order.tex
Tweak.

@@
 $\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
-conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, with $f$ providing the
+conventional semantics) must then be interpreted by a pair $(f, \partial f)$, with $f$ providing the
 conventional interpretation as a function $X \to Y$, and $\partial f$ as a family of Galois connections

### 2025-06-18 main/first-order.tex
\downset.

@@
 For a first-order language with a primitive type $\syn{nat}$ and primitive operations such as $+$, the
-``leaves'' of the approximation lattices would provided by a model of $\syn{nat}$ and its operations in
-$\Fam(\LatGal)$. For example, we might assign $\syn{nat}$ to the object $(\Nat,n \mapsto \{n, \bot\})$, where
-the family component is the two-point lattice indicating whether a given natural number is fully present ($n$)
-or fully absent ($\bot$) from the approximation, as shown in the example in \secref{introduction}, and we
-might then assign $+$ to the morphism $(\hat+,\partial(\hat+))$, with $\partial(\hat+)$ defined as
+language would provide a model of $\syn{nat}$ and its operations in $\Fam(\LatGal)$. For example, we might
+interpret $\syn{nat}$ to the object $(\Nat,n \mapsto \downset{n})$, where the family component assigns to
+every natural number $n$ the two-point lattice $\downset{n} = \{n, \bot\}$ indicating whether $n$ is fully
+present ($n$) or fully absent ($\bot$) from an approximation of $n$, as shown in the example in
+\secref{introduction}. We might then interpret $+$ as the morphism $(+,\partial+)$, with $+$ here meaning the
+usual addition operator, and $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as
 \begin{displaymath}
     \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
-      \partial(\hat+)&n&m&=&n \mathbin{\hat+} m \\
-      \partial(\hat+)&\bot&\_&=&\bot \\
-      \partial(\hat+)&\_&\bot&=&\bot
+      \partial+_{n,m}&n&m&=&n + m \\
+      \partial+_{n,m}&\bot&\_&=&\bot \\
+      \partial+_{n,m}&\_&\bot&=&\bot
     \end{array}

### 2025-06-18 main/first-order.tex
\downset.

@@
 present ($n$) or fully absent ($\bot$) from an approximation of $n$, as shown in the example in
-\secref{introduction}. We might then interpret $+$ as the morphism $(+,\partial+)$, with $+$ here meaning the
-usual addition operator, and $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as
+\secref{introduction}. We could then interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition
+operator + pairs with an approximating version $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as
 \begin{displaymath}

### 2025-06-18 main/first-order.tex
New subsubsection for lifting.

--- 2025-06-18 main/first-order.tex
@@
 present ($n$) or fully absent ($\bot$) from an approximation of $n$, as shown in the example in
-\secref{introduction}. We might then interpret $+$ as the morphism $(+,\partial+)$, with $+$ here meaning the
-usual addition operator, and $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as
+\secref{introduction}. We could then interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition
+operator + paired with an approximating version $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as:
 \begin{displaymath}
@@
 \end{displaymath}
-These elementary approximations would then be combined by the interpretation of composite types like sums and
-products, providing suitable derived approximation structures for composite values.
+These elementary approximation structures would then be combined by the interpretation of composite types like
+sums and products, providing suitable derived structures for composite values.
+
+\subsubsection{Lifting}
 
@@
 structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
-the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the
-language) define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times
-\alpha)$, along with an approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat}
-\times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
-\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
+lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then (within the language)
+define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$,
+along with an approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
+\synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, expressible as
+\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
 0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].

### 2025-06-18 main/first-order.tex
Tweak.

@@
 
-The families construction, regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F:
-\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category
-by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f,
-\partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where
-$(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then
-$\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by lifting the original monad $F: \cat{C} \to
-\cat{C}$ pointwise over families.
-
 \subsection{First-Order Galois Slicing via CHAD}
@@
 define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$,
-along with an approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
+or perhaps a custom approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
 \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, expressible as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
-0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].
+0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\]
 
@@
 from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
-matrix convolutions, without baking this specific choice of approximation into the interpretation of
-individual operators.
+matrix convolutions, without baking specific choices of approximation into the interpretation of individual
+operators.
 
-\todo{Show how $\LatGal$ has a lifting monad?}
+To interpret the syntactic type constructor $\synVar{L}$, \todo{show how $\LatGal$ has a lifting monad.}
+
+The families construction, regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F:
+\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category
+by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f,
+\partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where
+$(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then
+$\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to
+\cat{C}$ componentwise to families. Thus the lifting monad $\mathcal{L}$ in $\LatGal$ is also a lifting
+monad in $\Fam(\LatGal)$.
 
@@
 the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides an algorithmic method for computing the
-forward derivative of a composite function, without have to recompute the function itself, as long as all
-atomic operations provide derivatives.
+forward derivative of a composite function, as long as all atomic operations provide derivatives, without have
+to recompute the function itself.
 

### 2025-06-18 main/first-order.tex
Tweak.

@@
 One might also like to allow the programmer to embed additional ``approximation points'' into composite
-structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
-lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then (within the language)
-define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$,
-or perhaps a custom approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
-\synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, expressible as
+structures, by providing them with a \emph{lifting} type constructor $\synVar{L}$. Rather than the usual
+definition of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then (within
+the language) define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat}
+\times \alpha)$, or perhaps a custom approximating version of multiplication $*_{\text{appr}}:
+\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, expressible as
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
@@
 
-To interpret the syntactic type constructor $\synVar{L}$, \todo{show how $\LatGal$ has a lifting monad.}
+To interpret the syntactic type constructor $\synVar{L}$, we note that $\LatGal$ has a lifting monad
+$\mathcal{L}$ which sends any bounded lattice $X = (X, \meet, \join, \top, \bot)$ to the lattice
+$\mathcal{L}(X) = X_\bot$.
 

### 2025-06-18 main/first-order.tex
Doesn't quite make sense what I've written..

@@
 
-To interpret the syntactic type constructor $\synVar{L}$, we note that $\LatGal$ has a lifting monad
-$\mathcal{L}$ which sends any bounded lattice $X = (X, \meet, \join, \top, \bot)$ to the lattice
-$\mathcal{L}(X) = X_\bot$.
+To interpret the syntactic type constructor $\synVar{L}$, consider the strong lifting monad $\mathcal{L}$
+which extends a poset $X$ with a distinguished bottom element $\bot$ and extends a monotone function $f$
+between posets with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ induces a strong monad $\mathcal{L}_\meet$
+in any meet-semilattice on $X$. For any join-semilattice on $X$, $\mathcal{L}$ induces a costrong comonad, as
+long as $X$ is bounded; the bottom element of $X$ is needed to implement the counit $\varepsilon_X:
+\mathcal{L}(X) \to X$. These two in turn form a strong monad in $\LatGal$ which sends any lattice on $X$ to a
+lattice on $\mathcal{L}(X)$
+
+we note that the category of join-semilattices has a
+strong monad $\mathcal{L}_{\join}$ which sends any join-semilattice $X = (X, \join, \bot)$ to the semilattice
+which extends $X$ with a new bottom element $\bot_{\mathcal{L}}$ while preserving the original ordering of $X$
+within the lifted structure. One can do the same for meet-semilattices, extending $X = (X, \meet, \top)$ with
+a new element $\bot^*$ which also acts as the bottom element. In both cases the action on
+morphisms $f$ is to map $\bot^*$ to $\bot^*$ and otherwise act as $f$:
 

### 2025-06-18 main/first-order.tex
Doesn't quite make sense what I've written..
Some progress, notationally a bit tricky...

--- 2025-06-18 main/first-order.tex
@@
 
-To interpret the syntactic type constructor $\synVar{L}$, we note that $\LatGal$ has a lifting monad
-$\mathcal{L}$ which sends any bounded lattice $X = (X, \meet, \join, \top, \bot)$ to the lattice
-$\mathcal{L}(X) = X_\bot$.
+To interpret the syntactic type constructor $\synVar{L}$, consider the functor $\mathcal{L}$ which extends any
+poset $X$ with a distinguished bottom element $\bot$ and any monotone function $f$ between posets with a
+mapping from $\bot$ to $\bot$. $\mathcal{L}$ extends to a strong monad $\mathcal{L}^\meet$ on
+meet-semilattices, and to a costrong comonad $\mathcal{L}^\join$ on bounded join-semilattices; the bottom
+element of $X$ is needed to implement the counit $\varepsilon_X: \mathcal{L}^\join(X) \to X$. These two in
+turn form a strong monad in $\LatGal$ with unit $\varepsilon^\join_X \dashv \eta^\meet_X$ and multiplication
+$\delta^\join_X \dashv \mu^\meet_X$.
 

### 2025-06-18 main/first-order.tex
Some progress, notationally a bit tricky...
Rough pass on why LatGal has a lifting monad.

--- 2025-06-18 main/first-order.tex
@@
 
-To interpret the syntactic type constructor $\synVar{L}$, we note that $\LatGal$ has a lifting monad
-$\mathcal{L}$ which sends any bounded lattice $X = (X, \meet, \join, \top, \bot)$ to the lattice
-$\mathcal{L}(X) = X_\bot$.
+To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
+which extends any poset $X$ with a distinguished bottom element $\bot$ and any monotone function $f$ between
+posets with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong monad on meet-semilattices. On
+(bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad $(\mathcal{L}, \varepsilon, \delta)$,
+with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \mathcal{L}(X) \to X$. These two
+combine to form a strong lifting monad in $\LatGal$ with unit $\varepsilon_X \dashv \eta_X$ and multiplication
+$\delta_X \dashv \mu_X$.
 
-The families construction, regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F:
-\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category
-by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f,
-\partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where
-$(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then
-$\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to
-\cat{C}$ componentwise to families. Thus the lifting monad $\mathcal{L}$ in $\LatGal$ is also a lifting
-monad in $\Fam(\LatGal)$.
+The lifting monad $\mathcal{L}$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction,
+regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to
+functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with
+$F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F
+\comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial
+f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong)
+monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to
+families.
 

### 2025-06-19 main/first-order.tex
Tweak.

@@
 We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
-vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
-for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
-Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
-\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
-\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
+vector space $\RR^n$ as paired with the family of tangent spaces which is constantly $\RR^n$, and each map $f:
+\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$. Then the composition $(g, \partial
+g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ again corresponds exactly to
+the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp
+{\partial f}_x$.
 

### 2025-06-19 main/first-order.tex
Some clarification on lifting.

@@
 present ($n$) or fully absent ($\bot$) from an approximation of $n$, as shown in the example in
-\secref{introduction}. We could then interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition
-operator + paired with an approximating version $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as:
+\secref{introduction}. A slight variant is to use the constant family which assigns to every $n$ the two-point
+lattice $\Two = \{\mathsf{tt},\mathsf{ff}\}$; using $\downset{n}$ is slightly more ergonomic as the
+approximants of a given value then contain, rather than merely determining, all the information about the
+value that hasn't been approximated away.
+
+We might then interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition operator + paired with an
+approximating version $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as:
 \begin{displaymath}
@@
 
-One might also like to allow the programmer to embed additional ``approximation points'' into composite
-structures, by providing them with a \emph{lifting} type constructor $\synVar{L}$. Rather than the usual
-definition of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then (within
-the language) define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat}
-\times \alpha)$, or perhaps a custom approximating version of multiplication $*_{\text{appr}}:
-\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, expressible as
+A programmer, or perhaps a language implementor, might also like to embed additional ``approximation points''
+into composite structures, to allow for example the list $1 : 2 : []$ to have the partial list $1 : \bot$ as
+an approximant, representing a list with at least one element. This can be accomplished by extending the
+syntax with a \emph{lifting} type constructor $\synVar{L}$. Then, as an alternative to the usual definition of
+lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define
+a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$, with
+monadic computations over approximated lists computing partial lists like $1 : \bot$. Similarly, one could
+define a custom approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
+\synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, as:
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
@@
 operators.
+
+With this lifting operator in the core language, one could then choose to fix certain choices of
+approximation, by providing a surface language without the lifting monad and interpreting it into the core
+language with lifting. The Galois slicing implementations discussed in \citet{perera12a} and
+\citet{ricciotti17}, for example, provide an approximation point at every composite type constructor; we can
+emulate this in our approach by defining a surface language where the data type of lists $\mu\alpha.1 +
+\synVar{nat} \times \alpha$ would desugar automatically into the approximating version
+$\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$.
 

### 2025-06-19 main/first-order.tex

@@
 language would provide a model of $\syn{nat}$ and its operations in $\Fam(\LatGal)$. For example, we might
-interpret $\syn{nat}$ to the object $(\Nat,n \mapsto \downset{n})$, where the family component assigns to
+interpret $\syn{nat}$ as the object $(\Nat,n \mapsto \downset{n})$, where the family component assigns to
 every natural number $n$ the two-point lattice $\downset{n} = \{n, \bot\}$ indicating whether $n$ is fully
@@
 \secref{introduction}. A slight variant is to use the constant family which assigns to every $n$ the two-point
-lattice $\Two = \{\mathsf{tt},\mathsf{ff}\}$; using $\downset{n}$ is slightly more ergonomic as the
-approximants of a given value then contain, rather than merely determining, all the information about the
-value that hasn't been approximated away.
+lattice $\Two = \{\mathsf{tt},\mathsf{ff}\}$; using $\downset{n}$ can be more useful in practice as the
+approximants of a given value then contain all the information about the value that hasn't been approximated
+away.
 
-We might then interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition operator + paired with an
+We might interpret $+$ as the morphism $(+,\partial+)$, i.e.~the usual addition operator + paired with an
 approximating version $\partial+_{n,m}: \downset{n} \to \downset{m}$ defined as:
@@
 
-A programmer, or perhaps a language implementor, might also like to embed additional ``approximation points''
-into composite structures, to allow for example the list $1 : 2 : []$ to have the partial list $1 : \bot$ as
-an approximant, representing a list with at least one element. This can be accomplished by extending the
-syntax with a \emph{lifting} type constructor $\synVar{L}$. Then, as an alternative to the usual definition of
-lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define
-a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$, with
+A programmer, or perhaps a language implementor, might want to embed additional ``approximation points'' into
+composite data types, to allow for richer approximation structures. For example, for a list like $1 : 2 : []$
+we might want to allow the partial lists $\bot$, $1 : \bot$ and $1 : 2 : \bot$ as approximants, in addition to
+the approximants obtained by replacing elements with $\bot$. This can be accomplished by extending the syntax
+with a \emph{lifting} type constructor $\synVar{L}$. Then, as an alternative to the usual definition of lists
+of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define a
+type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$, with
 monadic computations over approximated lists computing partial lists like $1 : \bot$. Similarly, one could
@@
 emulate this in our approach by defining a surface language where the data type of lists $\mu\alpha.1 +
-\synVar{nat} \times \alpha$ would desugar automatically into the approximating version
-$\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$.
+\synVar{nat} \times \alpha$ desugars automatically into the approximating version $\mu\alpha.\synVar{L}\,(1 +
+\synVar{L}\,\synVar{nat} \times \alpha)$.
 
 To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
-which extends any poset $X$ with a distinguished bottom element $\bot$ and any monotone function $f$ between
-posets with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong monad on meet-semilattices. On
-(bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad $(\mathcal{L}, \varepsilon, \delta)$,
-with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \mathcal{L}(X) \to X$. These two
-combine to form a strong lifting monad in $\LatGal$ with unit $\varepsilon_X \dashv \eta_X$ and multiplication
-$\delta_X \dashv \mu_X$.
+which acts on posets $X$ to extend them with a distinguished bottom element $\bot$ and which acts on monotone
+functions $f$ between posets to extend them with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a
+strong monad on meet-semilattices. On (bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad
+$(\mathcal{L}, \varepsilon, \delta)$, with the bottom element of $X$ needed to implement the counit
+$\varepsilon_X: \mathcal{L}(X) \to X$. These two combine to form a strong lifting monad in $\LatGal$ with unit
+$\varepsilon_X \dashv \eta_X$ and multiplication $\delta_X \dashv \mu_X$.
 

### 2025-06-19 main/introduction.tex

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 This allows us to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''.
-Provenance tracking has a wide range of applications, from debugging and program comprehension \todo{cite1} to improving reproducibility and transparency in scientific workflows \todo{cite2}.
+Provenance tracking has a wide range of applications, from debugging and program comprehension~\cite{buneman95,cheney07} to improving reproducibility and transparency in scientific workflows~\cite{kontogiannis08}.
 \emph{Program slicing}, first proposed by~\citet{weiser81}, is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.

### 2025-06-19 notes/stability.tex
[ tex ] rough attempt to develop the stablity<->differentiable analogy

@@
 \label{sec:stable-functions}
-
-\newcommand{\downset}[1]{\mathop{\downarrow}(#1)}
 
@@
 
-Stability was invented by Berry \cite{berry86} in an attempt to capture
+Stability was invented by Berry \cite{berry79} in an attempt to capture
 sequentiality, and were also instrumental in the discovery of Linear

### 2025-06-28 main/first-order.tex
Drop inductive types from syntax, in favour of lists.

@@
 To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
-which acts on posets $X$ to extend them with a distinguished bottom element $\bot$ and which acts on monotone
-functions $f$ between posets to extend them with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a
-strong monad on meet-semilattices. On (bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad
+which acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
+function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong
+monad on meet-semilattices. On (bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad
 $(\mathcal{L}, \varepsilon, \delta)$, with the bottom element of $X$ needed to implement the counit

### 2025-06-28 main/language.tex
Drop inductive types from syntax, in favour of lists.

@@
 
-We introduce a total higher-order functional language with a (reasonably) expressive type
-system~\cite{crole94,pitts01,santocanale02}, over a set $\PrimTy$ of primitive types $\rho$ and sets
-$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$.
+We introduce a total higher-order functional language, over a set $\PrimTy$ of primitive types $\rho$ and sets
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$, equipped with sums, products, functions
+and a data type of lists. We omit a treatment of general inductive or coinductive types.
 

### 2025-06-28 main/language.tex
Drop inductive types from syntax, in favour of lists.
Kinding not needed if we only have lists.
Separate figure for parameterised types.

@@
 \input{fig/syntax}
+% \input{fig/kinding}
 \input{fig/typing}
@@
 
-\input{fig/mu-polynomial}
-\input{fig/default-semantics}
+% \input{fig/mu-polynomial}
+\input{fig/semantics-parameterised-types}
+\input{fig/semantics}
 
-\begin{definition}[$\mu$-polynomial]
-Suppose $\cat{C}$ a category with finite coproducts $(\coprod, 0)$ and finite products $(\times, 1)$. Define
-$\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
-\figref{mu-polynomial}.
-\end{definition}
-
-For any $\mu$-polynomial endofunctor $F: \cat{C} \to \cat{C}$ with initial algebra $\mu F$, write $\inMap_F$
-for the structure map $F\mu F \to \mu F$ and write $\cata_f$ for the unique morphism $\mu F \to X$ from the
-initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
+% \begin{definition}[$\mu$-polynomial]
+% Suppose $\cat{C}$ a category with finite coproducts $(\coprod, 0)$ and finite products $(\times, 1)$. Define
+% $\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
+% \figref{mu-polynomial}.
+% \end{definition}
+%
+% For any $\mu$-polynomial endofunctor $F: \cat{C} \to \cat{C}$ with initial algebra $\mu F$, write $\inMap_F$
+% for the structure map $F\mu F \to \mu F$ and write $\cata_f$ for the unique morphism $\mu F \to X$ from the
+% initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
 
@@
 \mathcal{L}(X \times Y)$;
-\item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
+% \item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
 \item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$;
@@
 
-\figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
+\figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
 = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
-\figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
+\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
 $\sem{\Gamma} \to \sem{\tau}$.
 
-\todo{Notation for partially applied functor used in \figref{default-semantics:types}}
+% \todo{Notation for partially applied functor used in \figref{semantics:types}}

### 2025-06-28 notes/language.tex
Separate figure for parameterised types.

@@
 \input{fig/mu-polynomial}
-\input{fig/default-semantics}
+\input{fig/semantics}
 
@@
 
-\figref{default-semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
+\figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
 = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
-\figref{default-semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
+\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
 $\sem{\Gamma} \to \sem{\tau}$.
 
-\todo{Notation for partially applied functor used in \figref{default-semantics:types}}
+\todo{Notation for partially applied functor used in \figref{semantics:types}}

### 2025-06-28 main/language.tex
Put inductive types experiment in a subfolder.
Rehashing without inductive types.

@@
 % \input{fig/mu-polynomial}
-\input{fig/semantics-parameterised-types}
+% \input{fig/inductive-types/semantics}
 \input{fig/semantics}
@@
 To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(\coprod, 0)$, exponentials
-$\internalHom{X}{Y}$, evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$, plus the
-following additional structure:
+category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(\coprod, 0)$ and exponentials
+$\internalHom{X}{Y}$, with evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$,
+plus the following additional structure:
 \begin{enumerate}
@@
 % \item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
-\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}$;
+\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$;
 \item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
@@
 
-\figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
-= n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
-\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
-$\sem{\Gamma} \to \sem{\tau}$.
+% \figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
+% = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
+\figrefTwo{semantics:types}{semantics:contexts} give the interpretations of types $\tau$ and contexts $\Gamma$
+as objects of $\Sem$. \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as
+morphisms $\sem{\Gamma} \to \sem{\tau}$.
 

### 2025-06-28 main/first-order.tex
\Lift macro.

@@
 
-To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
-which acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
-function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong
-monad on meet-semilattices. On (bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad
-$(\mathcal{L}, \varepsilon, \delta)$, with the bottom element of $X$ needed to implement the counit
-$\varepsilon_X: \mathcal{L}(X) \to X$. These two combine to form a strong lifting monad in $\LatGal$ with unit
-$\varepsilon_X \dashv \eta_X$ and multiplication $\delta_X \dashv \mu_X$.
+To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\Lift, \eta, \mu)$ which
+acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
+function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\Lift$ is also a strong monad
+on meet-semilattices. On (bounded) join-semilattices, $\Lift$ induces a costrong comonad $(\Lift, \varepsilon,
+\delta)$, with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \Lift(X) \to X$. These
+two combine to form a strong lifting monad in $\LatGal$ with unit $\varepsilon_X \dashv \eta_X$ and
+multiplication $\delta_X \dashv \mu_X$.
 
-The lifting monad $\mathcal{L}$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction,
-regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to
-functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with
-$F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F
-\comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial
-f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong)
-monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to
-families.
+The lifting monad $\Lift$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction, regarded
+as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to functors
+$\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with $F$,
+sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F \comp
+\partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial f)(x) =
+F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong) monad on
+$\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to families.
 

### 2025-06-28 main/first-order.tex

@@
 the approximants obtained by replacing elements with $\bot$. This can be accomplished by extending the syntax
-with a \emph{lifting} type constructor $\synVar{L}$. Then, as an alternative to the usual definition of lists
+with a \emph{lifting} type constructor $\tyLift$. Then, as an alternative to the usual definition of lists
 of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define a
-type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times \alpha)$, with
+type of ``approximation lists'' $\mu\alpha.\tyLift\,(1 + \tyLift\,\synVar{nat} \times \alpha)$, with
 monadic computations over approximated lists computing partial lists like $1 : \bot$. Similarly, one could
-define a custom approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat} \times
-\synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$, as:
+define a custom approximating version of multiplication $*_{\text{appr}}: \tyLift\,\synVar{nat} \times
+\tyLift\,\synVar{nat} \to \tyLift\,\synVar{nat}$, as:
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
@@
 emulate this in our approach by defining a surface language where the data type of lists $\mu\alpha.1 +
-\synVar{nat} \times \alpha$ desugars automatically into the approximating version $\mu\alpha.\synVar{L}\,(1 +
-\synVar{L}\,\synVar{nat} \times \alpha)$.
+\synVar{nat} \times \alpha$ desugars automatically into the approximating version $\mu\alpha.\tyLift\,(1 +
+\tyLift\,\synVar{nat} \times \alpha)$.
 
-To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\Lift, \eta, \mu)$ which
-acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and which acts on a monotone
-function $f$ between posets to extend it with a mapping from $\bot$ to $\bot$. $\Lift$ is also a strong monad
-on meet-semilattices. On (bounded) join-semilattices, $\Lift$ induces a costrong comonad $(\Lift, \varepsilon,
+To interpret the syntactic type constructor $\tyLift$, consider the strong monad $(\Lift, \eta, \mu)$ which
+acts on a poset $X$ to extend it with a distinguished bottom element $\bot$ and on a monotone function $f$
+between posets to extend it with a mapping from $\bot$ to $\bot$. $\Lift$ is also a strong monad on
+meet-semilattices. On (bounded) join-semilattices, $\Lift$ induces a costrong comonad $(\Lift, \varepsilon,
 \delta)$, with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \Lift(X) \to X$. These

### 2025-06-28 main/language.tex
Rehashing without inductive types.
List requirements on Sem.
\Lift macro.
Might need to think about weakening..
Clarify 'fold' semantics.
Allow for context object in specification of fold..

@@
 % \input{fig/inductive-types/semantics}
-\input{fig/semantics}
-
+% \figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
+% = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
+% \todo{Notation for partially applied functor used in \figref{semantics:types}}
+%
 % \begin{definition}[$\mu$-polynomial]
@@
 
+\input{fig/semantics}
+
 To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
@@
 \begin{enumerate}
-\item a strong monad $\mathcal{L}$ (lifting) with unit $\eta_X: X \to \mathcal{L}(X)$, multiplication $\mu_X:
-\mathcal{L}^2(X) \to \mathcal{L}(X)$ and strength $\mathsf{st}_{X,Y}: X \times \mathcal{L}(Y) \to
-\mathcal{L}(X \times Y)$;
-% \item $\mu$-polynomial endofunctors given by $\muPoly_{\Sem}$;
+\item a strong monad $(\Lift, \eta, \mu)$ with strength $\mathsf{st}_{X,Y}: X \times \Lift(Y) \to \Lift(X
+\times Y)$;
 \item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$;
 \item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
-\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$.
+\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$
+\item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
+$\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to Y$ and
+$f_\cons: X \times Z \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times Y \times \List(X) \to Y$.
 \end{enumerate}
 
-% \figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
-% = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
 \figrefTwo{semantics:types}{semantics:contexts} give the interpretations of types $\tau$ and contexts $\Gamma$
@@
 morphisms $\sem{\Gamma} \to \sem{\tau}$.
-
-% \todo{Notation for partially applied functor used in \figref{semantics:types}}

### 2025-06-28 main/language.tex
Corrections.

@@
 $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to Y$ and
-$f_\cons: X \times Z \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times Y \times \List(X) \to Y$.
+$f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
 \end{enumerate}

### 2025-07-02 main/higher-order.tex
Remove old higher-order section.

--- 2025-06-07 main/higher-order.tex
@@
-\section{CHAD at Higher-Order}
+\section{Higher-Order Programs}
 \label{sec:higher-order}
-
-$\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
-composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed and
-thereby able to interpret higher-order programs directly. However the following is a special case of a
-construction due to \citet{nunes2023}:
-
-\begin{proposition}
-Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
-$\Fam(\cat{C})$ is Cartesian closed.
-\end{proposition}
-
-A category $\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$
-of objects of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family
-of projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
-and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
-$\prodFam{f}$ making the following diagram commute for every $x \in X$:
-
-\begin{center}
-\begin{tikzcd}
-   \prod_{X} A \arrow[r, "\pi_x"] & A_x \\
-   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\prodFam{f}"]
-\end{tikzcd}
-\end{center}
-
-Abusing notation somewhat, we shall write $X$ for an object $(X, \partial X)$ in $\Fam(C)$, and similarly for
-morphisms $f = (f, \partial f)$. If $\cat{C}$ has all small products and also biproducts, then the internal
-hom $\internalHom{X}{Y}$ in $\Fam(\cat{C})$ is the object $Z$ where:
-\begin{itemize}
-\item the indexing set $Z$ is $\Fam(\cat{C})(X, Y)$, which is indeed a set by
-\propref{Grothendieck:fam-inherits-local-smallness};
-\item the indexed family $\partial Z$ is the family of products $\prod_{X}\reindex{\partial Y}{f}$ for every
-${f: X \to Y}$.
-\end{itemize}
-
-The products in $\Fam(C)$ arise from the products in $\cat{C}$ (\propref{Grothendieck:fam-inherits-products}).
-The evaluation morphism $\eval_{X,Y}: (\internalHom{X}{Y}) \times X \to Y$ is a family of morphisms in
-$\cat{C}$ given by:
-\begin{align*}
-\eval_{X,Y}(f,x) &= f(x) \\
-\partial\eval_{X,Y}(f,x) &= \coprodM{\pi_x}{\partial f(x)}
-  : \textstyle\prod_{X}\reindex{\partial Y}{f} \biprod \partial X(x) \to \partial Y(f(x))
-\end{align*}
-\noindent using the coproduct of morphisms from the biproduct in $\cat{C}$.
-
-The currying isomorphism $\lambda_{X,Y,Z}: \Hom{\Fam(C)}{X \times Y}{Z} \to
-\Hom{\Fam(C)}{X}{\internalHom{Y}{Z}}$ natural in $X, Y, Z$ is given by:
-
-\vspace{-4mm}
-\begin{align*}
-\lambda_{X,Y,Z}(f)(x) &= g \quad\textit{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
-\partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
-\end{align*}
-\noindent Here $(x,-): Y \to X \times Y$ is the ``reindexing'' function that pairs any $y$ with a fixed $x$,
-and for any $y$ $\biinj_{\partial X(x)}$ denotes the canonical injection $\partial X(x) \to \partial X(x)
-\biprod \partial Y(y)$.
-
-\todo{
-\begin{enumerate}
-  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products
-  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
-  biproducts from $\cat{C}$
-  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
-  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
-  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how
-  we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a
-  suitable object 2?}
-  \end{enumerate}
-}

### 2025-07-02 old/higher-order.tex
Remove old higher-order section.

\section{CHAD at Higher-Order}
\label{sec:higher-order}

$\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed. To
interpret higher-order programs, we use a special case of a construction due to \citet{nunes2023}
(\propref{higher-order:fam-CC} below). First we recall the notion of a product over a set-indexed family.

\begin{definition}[Small product]
Suppose $\cat{C}$ a category and $X$ a set. A \emph{product} of the family $A = \{A_x\}_{x \in X}$ is an
object $\prod_{X}A$ plus a family of projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in
$\cat{C}$ such that for any object $B$ and family of morphisms $f = \{f_x: B \to A_x\}_{x \in X}$ in
$\cat{C}$, there exists a unique morphism $\prodFam{f}$ making the following diagram commute for every $x \in
X$:

\begin{center}
\begin{tikzcd}
   \prod_{X} A \arrow[r, "\pi_x"] & A_x \\
   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\prodFam{f}"]
\end{tikzcd}
\end{center}
\end{definition}

\begin{proposition}
\label{prop:higher-order:fam-CC}
Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
$\Fam(\cat{C})$ is Cartesian closed.
\end{proposition}

Abusing notation somewhat, we shall write $X$ for an object $(X, \partial X)$ in $\Fam(C)$, and similarly for
morphisms $f = (f, \partial f)$. If $\cat{C}$ has all small products and also biproducts, then the internal
hom $\internalHom{X}{Y}$ in $\Fam(\cat{C})$ is the object $Z$ where:
\begin{itemize}
\item the indexing set $Z$ is $\Fam(\cat{C})(X, Y)$, which is indeed a set by
\propref{Grothendieck:fam-inherits-local-smallness};
\item the indexed family $\partial Z$ is the family of products $\prod_{X}\reindex{\partial Y}{f}$ for every
${f: X \to Y}$.
\end{itemize}

The products in $\Fam(C)$ arise from the products in $\cat{C}$ (\propref{Grothendieck:fam-inherits-products}).
The evaluation morphism $\eval_{X,Y}: (\internalHom{X}{Y}) \times X \to Y$ is a family of morphisms in
$\cat{C}$ given by:
\begin{align*}
\eval_{X,Y}(f,x) &= f(x) \\
\partial\eval_{X,Y}(f,x) &= \coprodM{\pi_x}{\partial f(x)}
  : \textstyle\prod_{X}\reindex{\partial Y}{f} \biprod \partial X(x) \to \partial Y(f(x))
\end{align*}
\noindent using the coproduct of morphisms from the biproduct in $\cat{C}$.

The currying isomorphism $\lambda_{X,Y,Z}: \Hom{\Fam(C)}{X \times Y}{Z} \to
\Hom{\Fam(C)}{X}{\internalHom{Y}{Z}}$ natural in $X, Y, Z$ is given by:

\vspace{-4mm}
\begin{align*}
\lambda_{X,Y,Z}(f)(x) &= g \quad\textit{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
\partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
\end{align*}
\noindent Here $(x,-): Y \to X \times Y$ is the ``reindexing'' function that pairs any $y$ with a fixed $x$,
and for any $y$ $\biinj_{\partial X(x)}$ denotes the canonical injection $\partial X(x) \to \partial X(x)
\biprod \partial Y(y)$.

\todo{
\begin{enumerate}
  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products
  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
  biproducts from $\cat{C}$
  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how
  we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a
  suitable object 2?}
  \end{enumerate}
}

### 2025-07-02 main/related-work.tex
Move.

@@
 
-\subsection{Galois Program slicing}
+\subsection{Stable Domain Theory}
+
+Stable Domain Theory was originally proposed by \citet{berry79} as a refinement of domain theory aimed at
+capturing the intensional behaviour of sequential programs. Standard domain-theoretic models interpret
+programs as continuous functions (preserving directed joins) Berry observed that this continuity condition
+alone is too permissive to model sequentiality. His introduction of stability imposed additional constraints
+to reflect how functions preserve bounded meets of approximants, ensuring that the evaluation of a function
+respects a notion of computational order. Subsequent work by \citet{berry82} and Amadio and Curien [1998, Ch.
+12], elaborated these ideas, showing that stable functions admit local Galois adjoints on principal downsets,
+and offering connections to linear logic and sequential algorithms. Though stable functions do not fully
+characterise sequentiality (e.g., they admit Gustave-style counterexamples), they remain a foundational
+concept for refining program denotations by their sensitivity to approximation.
+
+Our use of stable domain theory diverges from the traditional aim of modelling infinite or partial data.
+Instead, we follow a growing line of work that reinterprets domains as qualitative approximation structures
+suitable for provenance and slicing applications. In particular, Paul Taylor’s characterisation of stable
+functions via local Galois connections provides the semantic underpinning for the reverse maps used in Galois
+slicing, allowing us to reconstruct slices from output approximants [Taylor, 1991]. This perspective has been
+fruitfully adopted in settings such as data provenance [Cheney et al. 2007], program differentiation [Elliott
+2018], and recently by Vákár and collaborators [Lucatelli Nunes and Vákár 2023] in their CHAD framework. Our
+work builds on these ideas by interpreting Galois slicing as a form of differentiable programming, using the
+machinery of stable functions and Galois adjoints to unify forward and backward provenance in a denotational
+setting.
 
@@
 
-\subsection{Stable Domain Theory}
+\subsection{Galois Program slicing}
 

### 2025-07-02 main/related-work.tex
Rough pass over Stable Domain Theory.

--- 2025-06-03 main/related-work.tex
@@
 \section{Related work}
+\label{sec:related-work}
 
-\subsection{Galois Program slicing}
+\subsection{Stable Domain Theory}
+
+Stable Domain Theory was originally proposed by \citet{berry79} as a refinement of domain theory aimed at
+capturing the intensional behaviour of sequential programs, and elaborated on subsequently by \citet{berry82}
+and \citet{amadio-curien}. Standard domain-theoretic models interpret programs as continuous functions,
+preserving directed joins; Berry observed that this continuity condition alone is too permissive to model
+sequentiality. Stability imposes additional constraints to reflect how functions preserve bounded meets of
+approximants, effectively requiring that the evaluation of a function respect a specific computational order.
+Though stable functions do not fully characterise sequentiality, because they admit $\mathrm{gustave}$-style
+counterexamples (\exref{parallel-or}), they remain an appropriate notion for studying the sensitivity of a
+program to partial data at a specific point.
+
+Our use of Stable Domain Theory diverges from the traditional aim of modelling infinite or partial data,
+however. Instead, we follow a line of work that uses partiality as a qualitative notion of approximation
+suitable for provenance and program slicing (discussed in more detail in \secref{related-work:galois-slicing}
+below). Paul Taylor’s characterisation of stable functions via local Galois connections on principle downsets
+provides the semantic underpinning for the reverse maps used in Galois slicing~\cite{taylor99}. Our work
+builds on these ideas by interpreting Galois slicing as a form of differentiable programming, using the
+machinery of stable functions and Galois connections to present Galois slicing in a denotational style.
 
@@
 
-\subsection{Stable Domain Theory}
+\subsection{\GPS}
+\label{sec:related-work:galois-slicing}
 

### 2025-07-02 main/related-work.tex
Initial pass over AD.

@@
 
+Automatic differentiation (AD), discussed in \secref{first-order:autodiff}, is the idea of computing
+derivatives of functions expressed as programs by systematically applying the chain rule. The observation that
+these derivative computations could be interleaved with the evaluation of the original program is due to
+\citet{linnainmaa76}, who showed how the forward derivative $\pushf{f}_x$ of $f$ at a point $x$ could be
+computed alongside $f(x)$ in a single pass, dramatically improving the efficiency of derivative evaluation
+over symbolic or numerical differentiation. This insight became the foundation of forward-mode AD, which
+underpins many optimisation and scientific computing tools, including JAX~\cite{jax2018github}.
+
+More recent approaches to automatic differentiation have emphasised its semantic foundations.
+\citet{elliott18} proposed a categorical model of AD that interprets programs not only as functions enriched
+with their derivatives, giving a compositional account of differentiation based on duality and linear maps.
+This line of work connects AD to denotational semantics, abstracting over operational concerns and enabling
+principled extensions to differentiable programming languages. Vákár and
+collaborators~\cite{vákár22,nunes2023} developed the CHAD framework, using Grothendieck constructions over
+indexed categories to capture both values and their tangents in a unified semantic structure. These semantic
+perspectives shed light on the categorical structure of AD and guide the design of systems that generalise
+beyond real analysis, including the application to data provenance and slicing explored in this paper.
+
 \subsection{\GPS}

### 2025-07-02 main/related-work.tex
Now onto Galois slicing.

@@
 collaborators~\cite{vákár22,nunes2023} developed the CHAD framework, using Grothendieck constructions over
-indexed categories to capture both values and their tangents in a unified semantic structure. These semantic
-perspectives shed light on the categorical structure of AD and guide the design of systems that generalise
-beyond real analysis, including the application to data provenance and slicing explored in this paper.
+indexed categories to capture both values and their tangents in a compositional semantic structure. These
+semantic perspectives shed light on the categorical structure of AD and guide the design of systems that
+generalise beyond real analysis, including the application to data provenance and slicing explored in this
+paper.
 
@@
 
-\cite{berry79,berry82}
+\cite{perera12a,perera13,perera16d,ricciotti17,perera22,bond25}
 

### 2025-07-02 main/introduction.tex
Rough stab at Galois slicing.

@@
 Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
+Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid}\cite{perera2025fluid} that allows interactive exploration of programmatically generated visualisations.
 

### 2025-07-02 main/related-work.tex
Rough stab at Galois slicing.

@@
 
-\cite{perera12a,perera13,perera16d,ricciotti17,perera22,bond25}
+Galois slicing was introduced in \cite{perera12a} as an operational approach to program slicing based on the
+idea of computing slices via Galois connections between lattices of input and output approximations. A
+connection to Berry's stable functions in relation to minimal slices for short-circuiting operations was
+alluded to in \citeauthor{perera13}'s PhD thesis but not elaborated upon. Subsequent work extended the
+approach to functional languages with assignment and exceptions~\cite{ricciotti17} and concurrent systems,
+applying Galois slicing to the $\pi$-calculus~\cite{perera16d}. In this latter work, the analysis shifted from
+functions to transition relations, considering individual transitions $P \longrightarrow Q$ between specific
+configurations $P$ and $Q$ as analogous to the ``edge'' between points $x$ and their images $f(x)$ in $f$, and
+building Galois connections between $\downset{P}$ and $\downset{Q}$. This required generalizing the semantics
+to capture behavioural provenance in the presence of nondeterminism. The main difference with the Galois
+slicing approach presented here is that the earlier work also computes \emph{program slices}, using
+approximation lattices that represent partially erased programs.
+
+More recent work explored the application of Galois slicing in interactive visualisations. \citet{perera22}
+presented an approach where slicing operates over Boolean algebras rather than plain lattices. In this setting
+every Galois connection $f \dashv g: A \to B$ has a conjugate $g^\circ \dashv f^\circ: B \to A$, where
+$f^\circ$ denotes the De Morgan dual $\neg \comp f \comp \neg$~\cite{jonsson51}. The provenance analysis can
+then be composed with its own conjugate to obtain a Galois connection which computes \emph{related outputs}
+(e.g., selecting a region of a chart and observing the regions of other charts which have overlapping data
+dependencies). \citet{bond25} revisited this approach using \emph{dynamic dependence graphs} as an
+implementation technique that decouples the derivation of dependency information from the analyses that make
+use of it, and observing that to compute the conjugate analysis one can just use the opposite graph.
 

### 2025-07-02 main/approx-as-tangents.tex
Rough pass over Stable Domain Theory.
JAX citation.

@@
 \bob{cite something on autodiff}.
-This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms.
-\bob{cite: JAX, TensorFlow, Dex, Jesse Sigal, etc.}
+This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github}.
+\bob{cite: TensorFlow, Dex, Jesse Sigal, etc.}
 
@@
 
-Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality.
+Stability is a property that can be required of monotone functions that was invented by \citet{berry79} in an attempt to capture sequentiality.
 This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.

### 2025-07-02 main/introduction.tex
JAX citation.

--- 2025-06-19 main/introduction.tex
@@
 Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}.
-Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
+Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid}~\cite{perera2025fluid} that allows interactive exploration of programmatically generated visualisations.
 

### 2025-07-02 main/first-order.tex
Initial pass over AD.
More on lifting, but wondering if the discussion would be better elsewhere..

@@
 the approximants obtained by replacing elements with $\bot$. This can be accomplished by extending the syntax
-with a \emph{lifting} type constructor $\tyLift$. Then, as an alternative to the usual definition of lists
-of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define a
-type of ``approximation lists'' $\mu\alpha.\tyLift\,(1 + \tyLift\,\synVar{nat} \times \alpha)$, with
-monadic computations over approximated lists computing partial lists like $1 : \bot$. Similarly, one could
-define a custom approximating version of multiplication $*_{\text{appr}}: \tyLift\,\synVar{nat} \times
-\tyLift\,\synVar{nat} \to \tyLift\,\synVar{nat}$, as:
+with a \emph{lifting} type constructor $\tyLift$. Then, as an alternative to the usual definition of lists of
+natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the language) define a type
+of ``approximation lists'' $\mu\alpha.\tyLift\,(1 + \tyLift\,\synVar{nat} \times \alpha)$, with monadic
+computations over approximated lists computing partial lists like $1 : \bot$. The lattice of approximants of
+$1 : 2 : []$ now includes not only the ``shape-preserving'' approximations but also the prefixes obtained by
+truncating successive tails with $\bot$:
+
+  \begin{center}
+    \begin{tikzpicture}
+      \node (top) at (0,0) {$1 : 2 : []$};
+      % row 2
+      \node [below of=top] (ioi) {$1 : \bot : []$};
+      \node [left of=ioi,xshift=-3cm] (oii) {$\bot : 2 : []$};
+      \node [right of=ioi,xshift=3cm] (iio) {$1 : 2 : \bot$};
+      % row 3
+      \node [below of=ioi] (oio) {$\bot : 2 : \bot$};
+      \node [left of=oio,xshift=-3cm] (ooi) {$\bot : \bot : []$};
+      \node [right of=oio,xshift=3cm] (ioo) {$1 : \bot : \bot$};
+      % row 4
+      \node [below of=oio] (bot) {$[(\mathsf{a}, \bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]$};
+
+      % links
+      \draw (top) -- (ioi);
+      \draw (top) -- (oii);
+      \draw (top) -- (iio);
+      \draw (ioi) -- (oio);
+      \draw (ioi) -- (ooi);
+      \draw (ioi) -- (ioo);
+    \end{tikzpicture}
+  \end{center}
+
+Similarly, one could define a custom approximating version of multiplication $*_{\text{appr}}:
+\tyLift\,\synVar{nat} \times \tyLift\,\synVar{nat} \to \tyLift\,\synVar{nat}$, as:
 \[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' \equiv
@@
 from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
-matrix convolutions, without baking specific choices of approximation into the interpretation of individual
-operators.
+matrix convolutions, without having to bake specific choices of approximation into the interpretation of
+individual operators.
 
 With this lifting operator in the core language, one could then choose to fix certain choices of
-approximation, by providing a surface language without the lifting monad and interpreting it into the core
+approximation, providing a surface language without the lifting monad and interpreting it into the core
 language with lifting. The Galois slicing implementations discussed in \citet{perera12a} and
@@
 \subsection{First-Order Automatic Differentiation via CHAD}
+\label{sec:first-order:autodiff}
 

### 2025-07-02 main/first-order.tex
Lattice of list slices for listing example.

@@
 computations over approximated lists computing partial lists like $1 : \bot$. The lattice of approximants of
-$1 : 2 : []$ now includes not only the ``shape-preserving'' approximations but also the prefixes obtained by
-truncating successive tails with $\bot$:
+$1 : 2 : []$ now includes not only all the ``shape-preserving'' approximations but also all prefixes obtained
+by truncating successive tails with $\bot$:
 
   \begin{center}
+    \tikzset{node distance=1cm}
     \begin{tikzpicture}
@@
       \node [below of=top] (ioi) {$1 : \bot : []$};
-      \node [left of=ioi,xshift=-3cm] (oii) {$\bot : 2 : []$};
-      \node [right of=ioi,xshift=3cm] (iio) {$1 : 2 : \bot$};
+      \node [left of=ioi,xshift=-1.5cm] (oii) {$\bot : 2 : []$};
+      \node [right of=ioi,xshift=1.5cm] (iio) {$1 : 2 : \bot$};
       % row 3
       \node [below of=ioi] (oio) {$\bot : 2 : \bot$};
-      \node [left of=oio,xshift=-3cm] (ooi) {$\bot : \bot : []$};
-      \node [right of=oio,xshift=3cm] (ioo) {$1 : \bot : \bot$};
+      \node [left of=oio,xshift=-1.5cm] (ooi) {$\bot : \bot : []$};
+      \node [right of=oio,xshift=1.5cm] (ioo) {$1 : \bot : \bot$};
       % row 4
-      \node [below of=oio] (bot) {$[(\mathsf{a}, \bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]$};
+      \node [below of=oio,xshift=-1cm] (ooo) {$\bot : \bot : \bot$};
+      \node [below of=oio,xshift=1cm] (io) {$1 : \bot$};
+      % row 5
+      \node [below of=ooo,xshift=1cm] (oo) {$\bot : \bot$};
+      % row 6
+      \node [below of=oo] (o) {$\bot$};
 
@@
       \draw (ioi) -- (ioo);
+      \draw (oii) -- (oio);
+      \draw (oii) -- (ooi);
+      \draw (iio) -- (oio);
+      \draw (iio) -- (ioo);
+      \draw (ooi) -- (ooo);
+      \draw (oio) -- (ooo);
+      \draw (ioo) -- (ooo);
+      \draw (ioo) -- (io);
+      \draw (ooo) -- (oo);
+      \draw (io) -- (oo);
+      \draw (oo) -- (o);
     \end{tikzpicture}

### 2025-07-02 main/approx-as-tangents.tex
Citations on gradient descent.

@@
 Computing the forward and backward derivatives of smooth functions $f$ has many applications of practical interest.
-For example, computation of the reverse derivative is of central interest in Machine Learning by Gradient Descent, the main technique used to train Deep Neural Networks.
-\bob{need a citation here}.
+For example, computation of the reverse derivative is of central interest in machine learning by gradient descent, the main technique used to train deep neural networks~\cite{rumelhart88,goodfellow16}.
 

### 2025-07-02 main/approx-as-tangents.tex
Cite Dex, Sigal, etc.

@@
 We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting.
-From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.
+From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps.
+\bob{and we did it in Agda, so we can run it}
 
@@
 Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$.
-However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output.
-\bob{cite something on autodiff}.
-This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github}.
-\bob{cite: TensorFlow, Dex, Jesse Sigal, etc.}
+However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output~\cite{linnainmaa76}.
+This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github,abadi16,elliott17,sigal24}.
 

### 2025-07-02 main/conclusion.tex
Some related work.

@@
 \section{Conclusion and Future Work}
+\label{sec:conclusion}
 
-\todo{
-\begin{itemize}
-\item Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping
-(Grad-CAM)~\cite{selvaraju20} use reverse-mode AD selectively on specific activations, calculating \emph{heat
-maps} that highlight input regions contributing to a given classification decision. It might be possible to
-compose these techniques with \GPS whilst still preserving useful round-tripping properties; if so this could
-be a basis for end-to-end explainability in systems that combine neural networks with traditional symbolic
-computation.
-\end{itemize}
-}
+\paragraph{Integrating XAI techniques with Galois slicing.} \todo{Sketch what this might look like.}
+Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping (Grad-CAM)~\cite{selvaraju20}
+use reverse-mode AD selectively on specific activations, calculating \emph{heat maps} that highlight input
+regions contributing to a given classification decision. It might be possible to compose these techniques with
+\GPS whilst still preserving useful round-tripping properties; if so this could be a basis for end-to-end
+explainability in systems that combine neural networks with traditional symbolic computation.

### 2025-07-02 main/related-work.tex
Some related work.

--- 2025-07-02 main/related-work.tex
@@
 builds on these ideas by interpreting Galois slicing as a form of differentiable programming, using the
-machinery of stable functions and Galois connections to present Galois slicing in a denotational style.
+machinery of CHAD to present Galois slicing in a denotational style.
 
@@
 More recent approaches to automatic differentiation have emphasised its semantic foundations.
-\citet{elliott18} proposed a categorical model of AD that interprets programs not only as functions enriched
-with their derivatives, giving a compositional account of differentiation based on duality and linear maps.
-This line of work connects AD to denotational semantics, abstracting over operational concerns and enabling
-principled extensions to differentiable programming languages. Vákár and
-collaborators~\cite{vákár22,nunes2023} developed the CHAD framework, using Grothendieck constructions over
-indexed categories to capture both values and their tangents in a unified semantic structure. These semantic
-perspectives shed light on the categorical structure of AD and guide the design of systems that generalise
-beyond real analysis, including the application to data provenance and slicing explored in this paper.
+\citet{elliott18} proposed a categorical model of AD that interprets programs as functions enriched with their
+derivatives, giving a compositional account of differentiation based on duality and linear maps. This line of
+work connects AD to denotational semantics, abstracting over operational concerns and enabling principled
+extensions to differentiable programming languages. Vákár and collaborators~\cite{vákár22,nunes2023} developed
+the CHAD framework upon which this paper was based, using Grothendieck constructions over indexed categories
+to capture both values and their tangents in a compositional semantic structure. These semantic perspectives
+shed light on the categorical structure of AD and guide the design of systems that generalise beyond real
+analysis, including the application to data provenance and slicing explored in this paper.
 
@@
 
-\cite{berry79,berry82}
+Galois slicing was introduced by \citet{perera12a} as an operational approach to program slicing for pure
+functional programs, based on Galois connections between lattices of input and output approximations. A
+connection to stable functions in relation to minimal slices for short-circuiting operations was alluded to
+in~\citet{perera13}. Subsequent work extended the approach to languages with assignment and
+exceptions~\cite{ricciotti17} and concurrent systems, applying Galois slicing to the
+$\pi$-calculus~\cite{perera16d}. For the $\pi$-calculus the analysis shifted from functions to transition
+relations, considering individual transitions $P \longrightarrow Q$ between configurations $P$ and $Q$ as
+analogous to the edge between $x$ and $f(x)$ in the graph of $f$, and building Galois connections between
+$\downset{P}$ and $\downset{Q}$. The main difference with the approach presented here is that the earlier work
+also computes \emph{program slices}, using approximation lattices that represent partially erased programs; we
+discuss this further in \secref{conclusion} below.
+
+More recent work explored Galois slicing for interactive visualisations. \citet{perera22} presented an
+approach where slicing operates over Boolean algebras rather than plain lattices. In this setting every Galois
+connection $f \dashv g: A \to B$ has a conjugate $g^\circ \dashv f^\circ: B \to A$, where $f^\circ$ denotes
+the De Morgan dual $\neg \comp f \comp \neg$~\cite{jonsson51}. The provenance analysis can then be composed
+with its own conjugate to obtain a Galois connection which computes \emph{related outputs} (e.g., selecting a
+region of a chart and observing the regions of other charts which share data dependencies). \citet{bond25}
+revisited this approach using \emph{dynamic dependence graphs} to decouple the derivation of dependency
+information from the analyses that make use of it, and observing that to compute the conjugate analysis one
+can just use the opposite graph.
 

### 2025-07-03 notes/stable-coproducts.tex
Start on stable coproducts.

\subsection{Stable coproducts}
\label{sec:stable-coproducts}

Important to the idea of an extensive category is how coproducts behave under composition with an arbitrary
morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under
pullback or change of base.

### 2025-07-03 main/approx-as-tangents.tex
Fix enumerated list inside example environment.

@@
 
-Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory.
+Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitative approximations of elements in the sense of domain theory.
 Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry79,berry82}.
@@
 \begin{example}
+  \leavevmode\\[-3ex]
   \begin{enumerate}

### 2025-07-03 main/related-work.tex
Fix enumerated list inside example environment.

@@
 underpins many optimisation and scientific computing tools, including JAX~\cite{jax2018github}.
+\citet{griewank89} showed how the Wengert list, the linear record of assignments used in forward-mode to
+compute derivatives efficiently, could be traversed in reverse to compute the pullback map. This two-pass
+approach is the foundation of reverse-mode AD, and closely resembles implementations of Galois slicing
+(\secref{related-work:galois-slicing} below) that record a trace during forward slicing for use in backward
+slicing.
 
-More recent approaches to automatic differentiation have emphasised its semantic foundations.
-\citet{elliott18} proposed a categorical model of AD that interprets programs as functions enriched with their
-derivatives, giving a compositional account of differentiation based on duality and linear maps. This line of
-work connects AD to denotational semantics, abstracting over operational concerns and enabling principled
-extensions to differentiable programming languages. Vákár and collaborators~\cite{vákár22,nunes2023} developed
-the CHAD framework upon which this paper was based, using Grothendieck constructions over indexed categories
-to capture both values and their tangents in a compositional semantic structure. These semantic perspectives
-shed light on the categorical structure of AD and guide the design of systems that generalise beyond real
-analysis, including the application to data provenance and slicing explored in this paper.
+More recent approaches to automatic differentiation have emphasised semantic foundations. \citet{elliott18}
+proposed a categorical model of AD that interprets programs as functions enriched with their derivatives,
+giving a compositional account of differentiation based on duality and linear maps. Vákár and
+collaborators~\cite{vákár22,nunes2023} developed the CHAD framework which inspired this paper, using
+Grothendieck constructions over indexed categories to capture both values and their tangents in a
+compositional semantic structure. These perspectives shed light on the categorical structure of AD and guide
+the design of systems that generalise AD beyond real analysis, including the application to data provenance
+and slicing explored in this paper.
 

### 2025-07-03 main/approx-as-tangents.tex
taylor99.

@@
     In terms of \GPS the $\mathrm{gustave}$ function does not present a problem.
-For any particular run (i.e., input$\mapsto$output pair), there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
+For any particular run (i.e., input $\mapsto$ output pair), there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
 
@@
 
-The fact that these two functions's stability witnesses reveal information about how they depend on their input is what we will exploit in order to use the idea of stability for \GPS.
+The fact that these two functions' stability witnesses reveal information about how they depend on their input is what we will exploit in order to use the idea of stability for \GPS.
 It is interesting that sequentiality is not necessarily required in order to obtain backwards maps of approximations, as might be assumed by thinking about how to trace computations for provenance information.
@@
 Stability has an alternative definition in terms of Galois connections, which will be more useful for what follows.
-This characterisation is due to Paul Taylor\bob{cite}.
+This characterisation is due to~\citet{taylor99}.
 We first formally define Galois connections, which we used informally in \exref{introduction-example}.

### 2025-07-03 main/first-order.tex
Fix enumerated list inside example environment.
Simplify AD section a bit, add something about reverse-mode in Fam(FinVect^op).

@@
 
-% which can then be composed with $\partial f: \partial X \to T(f)(\partial Y)$.
-
 \subsubsection{Category of Families}
@@
 
-As per the CHAD work, automatic differentiation for first-order programs can also be recovered in this setup,
-using the total category $\Fam(\FinVect)$ instead of $\Fam(\LatGal)$. First we recall the chain rule for
-forward derivatives $\pushf{f}_x: \RR^m \linearto \RR^n$ and backward derivatives $\pullf{f}_x: \RR^n
-\linearto \RR^m$, which says that derivatives respect composition~\cite{spivak65}.
+As per the CHAD work, automatic differentiation (AD) for first-order programs can also be recovered in this
+setup, using the total categories $\Fam(\FinVect)$ and $\Fam(\FinVect^\op)$ instead of $\Fam(\LatGal)$. First
+we recall the chain rule for forward derivatives $\pushf{f}_x: \RR^m \linearto \RR^n$ and backward derivatives
+$\pullf{f}_x: \RR^n \linearto \RR^m$, which says that derivatives respect composition~\cite{spivak65}.
 
@@
 
-\subsubsection{Forward-Mode Automatic Differentiation}
+\subsubsection{Automatic Differentiation}
 
-The basic idea of (forward-mode) automatic differentiation, as first described by \citet{linnainmaa76}, was to
-recognise that one could efficiently compute the pushforward map $\pushf{f}: \RR^m \to (\RR^m \linearto
-\RR^n)$ of a differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself, by defining
-a single map $\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$.
-Defining $f_1 = \pi_1 \comp f: \RR^m \to \RR^n$ and $f_2 = \pi_2 \comp f: \RR^m \to (\RR^m \linearto \RR^n)$,
-composition of such maps can be expressed as:
+The basic idea of (forward-mode) AD, as first described by \citet{linnainmaa76}, was to recognise that one
+could efficiently compute the pushforward map $\pushf{f}: \RR^m \to (\RR^m \linearto \RR^n)$ of a
+differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself, by defining a single map
+$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$
+for $\pi_1 \comp f$ and $f_2$ for $\pi_2 \comp f$, composition of such maps can be expressed as:
 \begin{align*}
@@
 
-We can verify that $\tangents$ is functorial, regardless of whether the linear maps $\RR^m \linearto \RR^n$
-happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
-similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
-the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides an algorithmic method for computing the
-forward derivative of a composite function, as long as all atomic operations provide derivatives, without have
-to recompute the function itself.
+We can verify that $\tangents$ is functorial, regardless of whether the linear maps happen to be derivatives.
+But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and similarly for $g$ then the
+derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp
+f)_1$ at $x$. Thus this setup provides an algorithmic method for computing the forward derivative of a
+composite function, as long as all atomic operations provide derivatives, without have to recompute the
+function itself.
 
@@
 vector space $\RR^n$ as paired with the family of tangent spaces which is constantly $\RR^n$, and each map $f:
-\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$. Then the composition $(g, \partial
-g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ again corresponds exactly to
-the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp
-{\partial f}_x$.
-
-% with the reindexing along $f$ recomputing $f(x)$ in order to select the appropriate pushforward.
-
-\subsubsection{Reverse-Mode Automatic Differentiation}
-
-\todo{brief summary}
+\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as a pair of maps $(f_1, f_2)$. Then the composition $(g,
+\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ again corresponds
+exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)}
+\comp {\partial f}_x$. Reverse-mode AD can be interpreted similarly in $\Fam(\FinVect^\op)$, reading the
+single map $g: \RR^m \to \RR^n \times (\RR^n \linearto \RR^m)$ as the pair of maps $(g_1, g_2)$.

### 2025-07-03 main/language.tex
Tweak.
Simplify AD section a bit, add something about reverse-mode in Fam(FinVect^op).

@@
 \item a strong monad $(\Lift, \eta, \mu)$ with strength $\mathsf{st}_{X,Y}: X \times \Lift(Y) \to \Lift(X
-\times Y)$;
-\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$;
+\times Y)$
+\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
 \item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
@@
 
-\figrefTwo{semantics:types}{semantics:contexts} give the interpretations of types $\tau$ and contexts $\Gamma$
-as objects of $\Sem$. \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as
-morphisms $\sem{\Gamma} \to \sem{\tau}$.
+\figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects of $\Sem$.
+\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms $\sem{\Gamma}
+\to \sem{\tau}$.

### 2025-07-03 main/first-order.tex
Simplify AD section a bit, add something about reverse-mode in Fam(FinVect^op).

@@
 differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself, by defining a single map
-$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$
-for $\pi_1 \comp f$ and $f_2$ for $\pi_2 \comp f$, composition of such maps can be expressed as:
+$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $g_1$
+for $\pi_1 \comp g$ and $g_2$ for $\pi_2 \comp g$, composition of such maps can be expressed as:
 \begin{align*}
-(g \comp f)_1(x) &= f_1(g_1(x)) \\
-(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
+(g \comp h)_1(x) &= g_1(h_1(x)) \\
+(g \comp h)_2(x) &= g_2(h_1(x)) \comp h_2(x)
 \end{align*}
@@
 We can verify that $\tangents$ is functorial, regardless of whether the linear maps happen to be derivatives.
-But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and similarly for $g$ then the
-derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp
-f)_1$ at $x$. Thus this setup provides an algorithmic method for computing the forward derivative of a
+But if $h_2(x)$ is the derivative of $h_1$ at $x$ for any $x \in \RR^m$ and similarly for $g$ then the
+derivatives compose according to the forward chain rule and $(g \comp h)_2(x)$ is the derivative of $(g \comp
+h)_1$ at $x$. Thus this setup provides an algorithmic method for computing the forward derivative of a
 composite function, as long as all atomic operations provide derivatives, without have to recompute the
@@
 We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
-vector space $\RR^n$ as paired with the family of tangent spaces which is constantly $\RR^n$, and each map $f:
-\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as a pair of maps $(f_1, f_2)$. Then the composition $(g,
-\partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f)$ again corresponds
-exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial f)_x = {\partial g}_{f(x)}
-\comp {\partial f}_x$. Reverse-mode AD can be interpreted similarly in $\Fam(\FinVect^\op)$, reading the
-single map $g: \RR^m \to \RR^n \times (\RR^n \linearto \RR^m)$ as the pair of maps $(g_1, g_2)$.
+vector space $\RR^n$ as paired with the family of tangent spaces which is constantly $\RR^n$, and each map $g:
+\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as a pair of maps $(g_1, g_2)$. Then the composition $(g,
+\partial g) \comp (h, \partial h) = (g \comp h, \reindex{\partial g}{h} \comp \partial h)$ again corresponds
+to the forward chain rule, since $(\reindex{\partial g}{h} \comp \partial h)_x = {\partial g}_{h(x)} \comp
+{\partial h}_x$. Analogously, reverse-mode maps $g: \RR^m \to \RR^n \times (\RR^n \linearto \RR^m)$ can be
+interpreted as living in $\Fam(\FinVect^\op)$ by reading them as pair of maps $(g_1, g_2)$ with composition
+being precisely the backwards chain rule.

### 2025-07-03 main/first-order.tex
Minor tweaks.

@@
 \noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
-with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial:
-since $X \iso 0 \biprod X$, Cartesian closure would imply $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
-\Rightarrow Y} \iso 1$, and therefore that all objects are isomorphic.
+with finite biproducts is called \emph{semi-additive}. A semi-additive category cannot be Cartesian closed
+without being trivial: since we have $X \iso 0 \biprod X$, Cartesian closure would imply $\Hom{\cat{C}}{X}{Y}
+\iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$, and therefore in such a category all objects are isomorphic.
 

### 2025-07-03 main/first-order.tex
Remove duplicate definition of Galois connection.

@@
 
-The category $\LatGal$ of bounded lattices and Galois connections is semi-additive.
-
-\begin{definition}[Galois connection]
-A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$
-between posets satisfying $y \leq g(x) \iff f(y) \leq x$.
-\end{definition}
-
-\noindent A Galois connection $f \adj g: X \to Y$ is an adjunction between poset categories, regarding
-monotone $f$ and $g$ as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
-the \emph{lower} (left) adjoint. Galois connections are closed under component-wise composition and identities
-$\id_X \adj \id_X$.
+Galois connections (\defref{galois-connection} above) are closed under component-wise composition and
+identities. The category $\LatGal$ of bounded lattices and Galois connections is semi-additive.
 

### 2025-07-03 main/first-order.tex
Fix confusing overloading of C in 3.2.

@@
 conventional interpretation as a function $X \to Y$, and $\partial f$ as a family of Galois connections
-$\partial f(x): \partial X(x) \to \partial Y(f(x))$ taking approximants to approximants.
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$ taking approximations of $x$ to approximations of $f(x)$.
 
@@
 In its general form, the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
-\cat{C}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, incorporates all the categories
-$T(X)$ (for objects $X$ in $\cat{C}$) into a single category, together with morphisms that account for both
-internal structure and reindexing along morphisms $f: X \to Y$ in $\cat{C}$.
+\cat{D}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, incorporates all the categories
+$T(X)$ (for objects $X$ in $\cat{D}$) into a single category, together with morphisms that account for both
+internal structure and reindexing along morphisms $f: X \to Y$ in $\cat{D}$.
 
@@
 \label{def:Grothendieck}
-Suppose $T: \cat{C}^\op \to \Cat$ an indexed category. The \emph{Grothendieck construction}
+Suppose $T: \cat{D}^\op \to \Cat$ an indexed category. The \emph{Grothendieck construction}
 $\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) is the category which has as objects, all pairs $(X,
-\partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
-\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{C}$ and
+\partial X)$ of an object $X$ of $\cat{D}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
+\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{D}$ and
 morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
@@
 
-For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$ and
-each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
-of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
-$\cat{C}$ (with naturality trivial because $X$ has only identity morphisms).
+For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{D} = \Set$ and
+the indexed category $T$ sends every set $X$ to the (functor) category $\Func{X}{\cat{C}}$ which has as
+objects all $X$-indexed families of objects of a fixed category $\cat{C}$, and as morphisms $F \to G$, all
+families of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only
+identity morphisms).
 

### 2025-07-03 main/first-order.tex
More consistent authors.

@@
 $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is given by $(g \comp f,
-T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to ``reindex'' $\partial g$ along $f$,
+T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to reindex $\partial g$ along $f$,
 yielding a morphism $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial
@@
 computations over approximated lists computing partial lists like $1 : \bot$. The lattice of approximants of
-$1 : 2 : []$ now includes not only all the ``shape-preserving'' approximations but also all prefixes obtained
-by truncating successive tails with $\bot$:
+$1 : 2 : []$ now includes not only all ``shape-preserving'' approximations but also all prefixes obtained by
+truncating successive tails with $\bot$:
 

### 2025-07-03 main/language.tex
'language' no longer needed in notes.

@@
 \input{fig/typing}
-
-\figrefTwo{syntax}{typing} give the syntax and typing rules.
 
@@
 % initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
-
-\input{fig/semantics}
-
-To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(\coprod, 0)$ and exponentials
-$\internalHom{X}{Y}$, with evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$,
-plus the following additional structure:
-\begin{enumerate}
-\item a strong monad $(\Lift, \eta, \mu)$ with strength $\mathsf{st}_{X,Y}: X \times \Lift(Y) \to \Lift(X
-\times Y)$
-\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
-\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
-\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$
-\item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
-$\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to Y$ and
-$f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
-\end{enumerate}
-
-\figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects of $\Sem$.
-\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms $\sem{\Gamma}
-\to \sem{\tau}$.

### 2025-07-03 notes/stable-coproducts.tex
First pass over stable coproducts.

@@
 Important to the idea of an extensive category is how coproducts behave under composition with an arbitrary
-morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under
-pullback or change of base.
+morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under pullback
+or change of base. Suppose $\cat{C}$ a category with coproducts $(\coprod, 0)$. The coproducts of $\cat{C}$
+are \emph{stable under pullback} if for every object $X \iso X_1 \coprod X_2$ and morphism $g: Y \to X$, there
+exists an object $Y \iso Y_1 \coprod Y_2$ and morphisms $h_1: Y_1 \to X_1$ and $h_2: Y_2 \to X_2$ such that
+the following diagram commutes for every $i \in \{1, 2\}$,:
+
+\begin{center}
+\begin{tikzcd}
+   & Y \arrow[rd, "g"] \\
+   Y_i \arrow[rd, "h_i"'] \arrow[ru, "\inMap_i"] && X \\
+   & X_i \arrow[ru, "\inMap_i"']
+\end{tikzcd}
+\end{center}

### 2025-07-03 notes/predicate-system.tex
Start on predicate systems.

\subsection{Predicate systems}
\label{sec:predicate-system}

### 2025-07-03 notes/predicate-system.tex
Start on predicate systems.
Defining predicate system.

@@
 \label{sec:predicate-system}
+
+A \emph{predicate system} over a category $\cat{C}$ with products is the following data:
+\begin{itemize}
+\item for every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$
+\item for every morphism $f: X \to Y$ of $\cat{C}$:
+   \begin{enumerate}
+   \item a function $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computing the reindexing or pullback of a predicate along $f$
+   \item a function $\directImage{-}{f}: \Pred(X) \to \Pred(Y)$ computing the pushforward or existential
+   image of a predicate along $f$
+   \end{enumerate}
+\end{itemize}
+
+The pullback $\reindex{-}{-}$ is functorial.
+
+% Reindexing [f] (pullback) and [f⟩ (pushforward or existential image) that satisfy an adjunction.
+%
+% Logical connectives (&&, ++, ==>, TT) and quantification (⋀).
+%
+% These operations satisfy standard laws such as monotonicity, closure under reindexing, etc.

### 2025-07-03 notes/predicate-system.tex
Defining predicate system.

@@
 \item for every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$
-\item for every morphism $f: X \to Y$ of $\cat{C}$:
+\item for every morphism $f: X \to Y$ of $\cat{C}$, two monotone functions:
    \begin{enumerate}
-   \item a function $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computing the reindexing or pullback of a predicate along $f$
-   \item a function $\directImage{-}{f}: \Pred(X) \to \Pred(Y)$ computing the pushforward or existential
+   \item $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computing the reindexing or pullback of a predicate along $f$
+   \item $\directImage{-}{f}: \Pred(X) \to \Pred(Y)$ computing the pushforward or existential
    image of a predicate along $f$

### 2025-07-03 notes/predicate-system.tex
Pred(X) is a Heyting algebra with universal quantification (?).

--- 2025-07-03 notes/predicate-system.tex
@@
 \label{sec:predicate-system}
+
+A \emph{predicate system} over a category $\cat{C}$ consists of the following data:
+\begin{enumerate}
+\item for every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$
+forming a Heyting algebra with universal quantification (i.e.~closed under the null connective $\top$ and
+binary connectives $\meet$, $\join$ and $\implies$ and unary connective $\forall$);
+\item for every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv \directImage{-}{f} : \Pred(X) \to \Pred(Y)$.
+\end{enumerate}
+\noindent The left adjoint $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computes the reindexing or pullback of a
+predicate along $f$; the right adjoint $\directImage{-}{f}: \Pred(X) \to \Pred(Y)$ computes the pushforward or
+existential image of a predicate along $f$, with the adjointness property meaning that
+\[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
+
+% These operations satisfy standard laws such as monotonicity, closure under reindexing, etc.

### 2025-07-04 notes/predicate-system.tex
Predicate system is indexed family of posets with logical operations.

@@
 A \emph{predicate system} over a category $\cat{C}$ consists of the following data:
-\begin{enumerate}
-\item for every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$
-forming a Heyting algebra with universal quantification (i.e.~closed under the null connective $\top$ and
-binary connectives $\meet$, $\join$ and $\implies$ and unary connective $\forall$);
-\item for every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv \directImage{-}{f} : \Pred(X) \to \Pred(Y)$.
-\end{enumerate}
-\noindent The left adjoint $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computes the reindexing or pullback of a
-predicate along $f$; the right adjoint $\directImage{-}{f}: \Pred(X) \to \Pred(Y)$ computes the pushforward or
-existential image of a predicate along $f$, with the adjointness property meaning that
+\begin{itemize}
+\item For every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
+\item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv
+\directImage{-}{f} : \Pred(X) \to \Pred(Y)$. The left adjoint $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computes
+the pullback (reindexing) of a predicate along $f$; the right adjoint $\directImage{-}{f}: \Pred(X) \to
+\Pred(Y)$ computes the pushforward (existential image) of a predicate along $f$, with the adjointness property
+meaning that
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
-
+\item For every fiber $\Pred(X)$ the following structure:
+   \begin{itemize}
+   \item $\Pred(X)$ is a Heyting algebra with universal quantification, i.e.~closed under top element $\top$,
+   meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary connective $\forall$, satisfying the usual
+   laws;
+   \item each operation is preserved by reindexing.
+   \end{itemize}
+\end{itemize}
 % These operations satisfy standard laws such as monotonicity, closure under reindexing, etc.

### 2025-07-04 notes/predicate-system.tex
Predicate system is indexed family of posets with logical operations.

@@
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
-\item For every fiber $\Pred(X)$ the following structure:
-   \begin{itemize}
-   \item $\Pred(X)$ is a Heyting algebra with universal quantification, i.e.~closed under top element $\top$,
-   meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary connective $\forall$, satisfying the usual
-   laws;
+\item For every $\Pred(X)$, the structure of a Heyting algebra with universal quantification, i.e.~$\Pred(X)$
+   is closed under top element $\top$, meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary
+   connective $\forall$, satisfying the usual laws;
    \item each operation is preserved by reindexing.

### 2025-07-04 notes/predicate-system.tex
Predicate system is indexed family of posets with logical operations.

@@
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
-\item For every $\Pred(X)$, the structure of a Heyting algebra with universal quantification, i.e.~$\Pred(X)$
-   is closed under top element $\top$, meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary
-   connective $\forall$, satisfying the usual laws;
-   \item each operation is preserved by reindexing.
-   \end{itemize}
+\item For every $\Pred(X)$, the structure of a Heyting algebra \todo{but no $\bot$?} with universal
+   quantification, i.e.~$\Pred(X)$ is closed under top element $\top$, meets $\meet$, joins $\join$, residual
+   $\Rightarrow$, and unary connective $\forall$, satisfying the usual laws, whereeach operation is preserved
+   by reindexing.
 \end{itemize}

### 2025-07-04 notes/predicate-system.tex
−[𝑓 ] is colax (and thus −⟨𝑓⟩ is lax) with respect to the Pred operations.

--- 2025-07-04 notes/predicate-system.tex
@@
 \item For every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
+\item For every $\Pred(X)$, the structure of a Heyting algebra \todo{but no $\bot$?} with universal
+   quantification, i.e.~$\Pred(X)$ is closed under top element $\top$, meets $\meet$, joins $\join$, residual
+   $\Rightarrow$, and unary connective $\forall$, satisfying the usual laws.
 \item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv
@@
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
-\item For every $\Pred(X)$, the structure of a Heyting algebra with universal quantification, i.e.~$\Pred(X)$
-   is closed under top element $\top$, meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary
-   connective $\forall$, satisfying the usual laws;
-   \item each operation is preserved by reindexing.
-   \end{itemize}
+\item $\reindex{-}{f}$ is colax (and thus $\directImage{-}{f}$ is lax) with respect to the $\Pred$ operations.
 \end{itemize}

### 2025-07-04 notes/predicate-system.tex
lax w.r.t. join. Universal quantifier is family of connectives.

--- 2025-07-04 notes/predicate-system.tex
@@
 \item For every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
+\item For every $\Pred(X)$, the structure of a Heyting algebra \todo{but no $\bot$?} with universal
+   quantification, i.e.~$\Pred(X)$ is closed under top element $\top$, meets $\meet$, joins $\join$, residual
+   $\residual$, and universal quantifiers $\forall_Y$ for all objects $Y$, satisfying the usual laws.
 \item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv
@@
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
-\item For every $\Pred(X)$, the structure of a Heyting algebra with universal quantification, i.e.~$\Pred(X)$
-   is closed under top element $\top$, meets $\meet$, joins $\join$, residual $\Rightarrow$, and unary
-   connective $\forall$, satisfying the usual laws;
-   \item each operation is preserved by reindexing.
+\item $\reindex{-}{f}$ is colax (and thus $\directImage{-}{f}$ is lax) with respect to $\top$, $\meet$,
+$\residual$ and $\forall_Y$
+\item $\reindex{-}{f}$ is lax (and thus $\directImage{-}{f}$ is colax) with respect to $\join$.
 \end{itemize}
-\end{itemize}
-% These operations satisfy standard laws such as monotonicity, closure under reindexing, etc.
+% These operations satisfy standard laws such as monotonicity, etc.

### 2025-07-04 notes/predicate-system.tex
Clarify C has finite products.
Some improvements.

@@
 
-A \emph{predicate system} over a category $\cat{C}$ consists of the following data:
+Suppose $\cat{C}$ at category with finite products. A \emph{predicate system} over $\cat{C}$ consists of the
+following data:
 \begin{itemize}
 \item For every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
-\item For every $\Pred(X)$, the structure of a Heyting algebra \todo{but no $\bot$?} with universal
-   quantification, i.e.~$\Pred(X)$ is closed under top element $\top$, meets $\meet$, joins $\join$, residual
-   $\residual$, and universal quantifiers $\forall_Y$ for all objects $Y$, satisfying the usual laws.
+\item For every $X$, the structure of a Heyting algebra (without $\bot$) on $\Pred(X)$, i.e.~ top element
+   $\top$, meets $\meet$, joins $\join$ and residual $\residual$ satisfying the usual laws.
 \item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv
@@
 \[\directImage{P}{f} \sqsubseteq Q \iff P \sqsubseteq \reindex{Q}{f} \]
+\item For every $X$ and $Y$, a right adjoint $\forall_Y: \Pred(X \times Y) \to \Pred(X)$ to the monotone map
+$\reindex{-}{\pi_1}$ called \emph{universal quantification}
 \item $\reindex{-}{f}$ is colax (and thus $\directImage{-}{f}$ is lax) with respect to $\top$, $\meet$,
-$\residual$ and $\forall_Y$
+$\residual$ and $\forall_Y$.
 \item $\reindex{-}{f}$ is lax (and thus $\directImage{-}{f}$ is colax) with respect to $\join$.

### 2025-07-04 notes/predicate-system.tex
That will do for now. Now onto presheaf version.

@@
    $\top$, meets $\meet$, joins $\join$ and residual $\residual$ satisfying the usual laws.
-\item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\reindex{-}{f} \dashv
-\directImage{-}{f} : \Pred(X) \to \Pred(Y)$. The left adjoint $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computes
-the pullback (reindexing) of a predicate along $f$; the right adjoint $\directImage{-}{f}: \Pred(X) \to
+\item For every morphism $f: X \to Y$ of $\cat{C}$, a Galois connection $\directImage{-}{f} \dashv
+\reindex{-}{f} : \Pred(Y) \to \Pred(X)$. The right adjoint $\reindex{-}{f}: \Pred(Y) \to \Pred(X)$ computes
+the pullback (reindexing) of a predicate along $f$; the left adjoint $\directImage{-}{f}: \Pred(X) \to
 \Pred(Y)$ computes the pushforward (existential image) of a predicate along $f$, with the adjointness property
@@
 \item For every $X$ and $Y$, a right adjoint $\forall_Y: \Pred(X \times Y) \to \Pred(X)$ to the monotone map
-$\reindex{-}{\pi_1}$ called \emph{universal quantification}
+$\reindex{-}{\pi_1}$ called \emph{universal quantification}. \todo{Is there a left adjoint $\exists_Y$?}
 \item $\reindex{-}{f}$ is colax (and thus $\directImage{-}{f}$ is lax) with respect to $\top$, $\meet$,
@@
 \end{itemize}
-% These operations satisfy standard laws such as monotonicity, etc.

### 2025-07-04 notes/predicate-system.tex
Fix directions of arrows.

@@
 
-Suppose $\cat{C}$ at category with finite products. A \emph{predicate system} over $\cat{C}$ consists of the
+\subsubsection{Predicate system over $\cat{C}$}
+
+Suppose $\cat{C}$ a category with finite products. A \emph{predicate system} over $\cat{C}$ consists of the
 following data:
@@
 \end{itemize}
+
+\subsubsection{Predicates on a presheaf}
+
+Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over $\Set$,
+$\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a predicate on $F$ to be a family of
+predicates $P_X \in \Pred(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
+\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$.

### 2025-07-04 notes/predicate-system.tex
Predicate system for presheaf category.

@@
 \begin{itemize}
-\item For every object $X$ of $\cat{C}$, a partial order $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
+\item For every object $X$ of $\cat{C}$, a poset $(\Pred(X), \sqsubseteq)$ of predicates on $X$.
 \item For every $X$, the structure of a Heyting algebra (without $\bot$) on $\Pred(X)$, i.e.~ top element
@@
 $\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a predicate on $F$ to be a family of
-predicates $P_X \in \Pred(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
-\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$.
+predicates $P_X \in \Pred_\mathcal{S}(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
+\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$. Define $\mathcal{S}^{\cat{C}}$ to be the predicate
+system which assigns to every presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$ ordered
+pointwise, and where $\top, \meet, \join, \residual$ and $\forall$ are all computed pointwise.

### 2025-07-04 notes/predicate-system.tex
Predicate system for presheaf category.

@@
 predicates $P_X \in \Pred_\mathcal{S}(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
-\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$. Define $\mathcal{S}^{\cat{C}}$ to be the predicate
-system which assigns to every presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$ ordered
-pointwise, and where $\top, \meet, \join, \residual$ and $\forall$ are all computed pointwise.
+\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$. Now define $\mathcal{S}^{\cat{C}}$ to be the
+predicate system which assigns to every presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$
+ordered pointwise, and where $\top, \meet, \join, \residual$ and $\forall$ are all computed pointwise.

### 2025-07-04 notes/predicate-system.tex
Ok need to fix the predicate system over Set.

--- 2025-07-04 notes/predicate-system.tex
@@
 Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over $\Set$,
-$\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a predicate on $F$ to be a family of
-predicates $P_X \in \Pred(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
+$\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a \emph{predicate on $F$} to be a family
+of predicates $P_X \in \Pred_\mathcal{S}(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
 \reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$.
+
+Now define the predicate system $\mathcal{S}^{\cat{C}}$ over $\Func{\cat{C}^\op}{\Set}$ which assigns to every
+presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$ ordered pointwise, and where $\top, \meet,
+\join, \residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation
+$\alpha: F \to G$, the pullback $\reindex{-}{\alpha}: \Pred(G) \to \Pred(F)$ ...

### 2025-07-04 notes/predicate-system.tex
Defining predicate system over Set.

@@
 
+\subsubsection{Predicate system over $\Set$}
+
+Suppose $X$ a set. Define a \emph{predicate on $X$} to be a family of (proof-irrelevant) propositions. For any
+function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and $\directImage{P}{f}(y)$ iff there exists
+$x = f(y)$ such that $P(x)$. Define $\mathcal{S}$ to be the predicate system over $\Set$ which assigns to any
+set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by implication, with $\top, \meet, \join$ and
+$\residual$ defined pointwise.
+
 \subsubsection{Predicates on a presheaf}
 
-Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over $\Set$,
-$\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a \emph{predicate on $F$} to be a family
-of predicates $P_X \in \Pred_\mathcal{S}(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y \sqsubseteq
-\reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$.
+Suppose $\cat{C}$ a category and $F: \cat{C}^\op \to \Set$ a presheaf. Define a \emph{predicate on $F$} to be
+a family of predicates $P_X \in \Pred_\mathcal{S}(F(X))$ for every object $X$ in $\cat{C}$ such that $P_Y
+\sqsubseteq \reindex{P_X}{F(f)}$ for every $f: X \to Y$ in $\cat{C}$.
 

### 2025-07-04 notes/predicate-system.tex
Predicate system over Set.
Add universal quantification.

@@
 function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and $\directImage{P}{f}(y)$ iff there exists
-$x = f(y)$ such that $P(x)$. Define $\mathcal{S}$ to be the predicate system over $\Set$ which assigns to any
-set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by implication, with $\top, \meet, \join$ and
-$\residual$ defined pointwise.
+$x$ such that $f(x) = y$ and $P(x)$. Define $\mathcal{S}$ to be the predicate system over $\Set$ which assigns
+to any set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by implication, with $\top, \meet, \join$ and
+$\residual$ defined pointwise and universal quantification defined as $\forall_Y(P)(x)$ iff $P(x,y)$ for all
+$y \in Y$.
 

### 2025-07-04 notes/predicate-system.tex
Reindex/direct image along a natural transformation.

@@
 
-Suppose $X$ a set. Define a \emph{predicate on $X$} to be a family of (proof-irrelevant) propositions. For any
-function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and $\directImage{P}{f}(y)$ iff there exists
-$x$ such that $f(x) = y$ and $P(x)$. Define $\mathcal{S}$ to be the predicate system over $\Set$ which assigns
-to any set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by implication, with $\top, \meet, \join$ and
-$\residual$ defined pointwise and universal quantification defined as $\forall_Y(P)(x)$ iff $P(x,y)$ for all
-$y \in Y$.
+Suppose $X$ a set. Define a \emph{predicate on $X$} to be a family of (proof-irrelevant) propositions for
+every $x \in X$. For any function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and
+$\directImage{P}{f}(y)$ iff there exists $x$ such that $f(x) = y$ and $P(x)$. Define $\mathcal{S}$ to be the
+predicate system over $\Set$ which assigns to any set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by
+implication, with $\top, \meet, \join$ and $\residual$ defined pointwise and universal quantification defined
+as $\forall_Y(P)(x)$ iff $P(x,y)$ for all $y \in Y$.
 
@@
 Now define the predicate system $\mathcal{S}^{\cat{C}}$ over $\Func{\cat{C}^\op}{\Set}$ which assigns to every
-presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$ ordered pointwise, and where $\top, \meet,
-\join, \residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation
-$\alpha: F \to G$, the pullback $\reindex{-}{\alpha}: \Pred(G) \to \Pred(F)$ ...
+presheaf $F$ the poset $\Pred(F)$ of predicates on $F$ ordered pointwise, and where $\top, \meet, \join,
+\residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation $\alpha: F
+\to G$, define $\reindex{P}{\alpha}_X = \reindex{P_X}{\alpha_X}$ and $\directImage{P}{\alpha}_X =
+\directImage{P_X}{\alpha_X}$.

### 2025-07-04 notes/predicate-system.tex
First pass over predicate system for the presheaf category for C.

@@
 Now define the predicate system $\mathcal{S}^{\cat{C}}$ over $\Func{\cat{C}^\op}{\Set}$ which assigns to every
-presheaf $F$ the poset $\Pred(F)$ of predicates on $F$ ordered pointwise, and where $\top, \meet, \join,
-\residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation $\alpha: F
-\to G$, define $\reindex{P}{\alpha}_X = \reindex{P_X}{\alpha_X}$ and $\directImage{P}{\alpha}_X =
-\directImage{P_X}{\alpha_X}$.
+presheaf $F$ the poset $\Pred(F)$ of predicates on $F$ ordered pointwise, where $\top, \meet, \join,
+\residual$ and $\forall$ in $\Pred(F)$ are all given pointwise, and to every natural transformation $\alpha: F
+\to G$, families of pullback and pushforward maps where $\reindex{P}{\alpha}_X = \reindex{P_X}{\alpha_X}$ and
+$\directImage{P}{\alpha}_X = \directImage{P_X}{\alpha_X}$.

### 2025-07-04 notes/predicate-system.tex
'Pred' not a category, so drop the bold..

@@
 $\directImage{P}{\alpha}_X = \directImage{P_X}{\alpha_X}$.
+
+\subsubsection{Closure operator over a predicate system}
+
+A \emph{closure operator} $C$ over a poset $X$ is a lax idempotent monad on $X$, with functoriality, unit and
+multiplication given by monotonicity, extensivity ($x \leq C(x)$) and idempotence ($C(C(x)) \leq C(x)$).
+Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over a category
+$\cat{C}$. A closure operator $C$ over $\mathcal{S}$ is a fibred lax idempotent monad $C$ on the fibration
+$\Pred: \cat{C}^\op \to \Poset$ such that for every object $X$ in $\cat{C}$ the fibre map $\Pred(X)$ is a
+closure operator, with a lax strength with respect to $\meet$.

### 2025-07-04 notes/predicate-system.tex
Ok, natural isomorphism between C \comp -[f] and -[f] \comp C is just the statement that C is a fibred functor.

--- 2025-07-04 notes/predicate-system.tex
@@
 $\directImage{P}{\alpha}_X = \directImage{P_X}{\alpha_X}$.
+
+\subsubsection{Closure operator over a predicate system}
+
+A \emph{closure operator} $C$ over a poset $(X, \sqleq)$ is a lax idempotent monad on $X$, with functoriality,
+unit and multiplication given by monotonicity, extensivity ($x \sqleq C(x)$) and idempotence ($C(C(x)) \sqleq
+C(x)$). Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over a
+category $\cat{C}$. A closure operator $C$ over $\mathcal{S}$ is a fibred lax idempotent monad $C$ on the
+fibration $\Pred: \cat{C}^\op \to \Poset$ such that:
+\begin{enumerate}
+\item for every object $X$ in $\cat{C}$ the fibre map $C_X: \Pred(X) \to \Pred(X)$ is a closure operator;
+\item $C$ is fibred, i.e. $C_X \comp \reindex{-}{f}$ is naturally isomorphic to $\reindex{-}{f} \comp C_Y$;
+\item $C$ is lax monoidal with respect to $\meet$;
+\item $C$ is lax strong, i.e.~$C(X \times Y) \sqleq X \times C(Y)$.
+\end{enumerate}

### 2025-07-04 notes/predicate-system.tex
Ok, natural isomorphism between C \comp -[f] and -[f] \comp C is just the statement that C is a fibred functor.
Assuming closure operator over a predicate system is a monad (but we don't have that data in the Agda).

--- 2025-07-04 notes/predicate-system.tex
@@
 $\directImage{P}{\alpha}_X = \directImage{P_X}{\alpha_X}$.
+
+\subsubsection{Closure operator over a predicate system}
+
+A \emph{closure operator} $C$ over a poset $(X, \sqleq)$ is a lax idempotent monad on $X$, with functoriality,
+unit and multiplication given by monotonicity, extensivity ($x \sqleq C(x)$) and idempotence ($C(C(x)) \sqleq
+C(x)$).
+
+Suppose $\mathcal{S} = (\Pred, \top, \meet, \join, \residual, \forall)$ a predicate system over a category
+$\cat{C}$. A closure operator $C$ over $\mathcal{S}$ is a fibred lax idempotent monad $C$ on the fibration
+$\Pred: \cat{C}^\op \to \Poset$ where for every object $X$ the fibre map $C_X: \Pred(X) \to \Pred(X)$ is a
+closure operator, $C$ is lax monoidal with respect to $\meet$ (i.e.~$C(P) \meet C(Q) \sqleq C(P \meet Q)$),
+and $C$ is lax strong, i.e.~$C(P) \meet Q \sqleq C(P \meet Q)$. $C$ being a fibred functor means it commutes
+with any reindexing functor $\reindex{-}{f}$, i.e.~$C_X \comp \reindex{-}{f} \iso \reindex{-}{f} \comp C_Y$.

### 2025-07-04 notes/stable-coproducts.tex
Start on predicate systems.
Now using stable coproducts to define the coproduct closure operator.

--- 2025-07-03 notes/stable-coproducts.tex
@@
 Important to the idea of an extensive category is how coproducts behave under composition with an arbitrary
-morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under
-pullback or change of base.
+morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under pullback
+or change of base. Suppose $\cat{C}$ a category with coproducts $(\coprod, 0)$. The coproducts of $\cat{C}$
+are \emph{stable under pullback} if for any objects $X \iso X_1 \coprod X_2$ and morphism $g: Y \to X$, there
+exist objects $Y \iso Y_1 \coprod Y_2$ and morphisms $h_1: Y_1 \to X_1$ and $h_2: Y_2 \to X_2$ such that the
+following diagram commutes for every $i \in \{1, 2\}$,:
+
+\begin{center}
+\begin{tikzcd}
+   & Y \arrow[rd, "g"] \\
+   Y_i \arrow[rd, "h_i"'] \arrow[ru, "\inMap_i"] && X \\
+   & X_i \arrow[ru, "\inMap_i"']
+\end{tikzcd}
+\end{center}

### 2025-07-05 main/first-order.tex
Tweak.

@@
 \begin{enumerate}
-\item If $X \times Y$ is a product then it is also coproduct, with $\inj_1 = \prodM{\id}{0}$ and $\inj_2 =
-\prodM{0}{\id}$.
-\item If $X \coprod Y$ is a coproduct then it is also a product, with $\pi_1 = \coprodM{\id}{0}$ and $\pi_2 =
-\coprodM{0}{\id}$.
+\item If $Z$ is a product of $X$ and $Y$ then it is also coproduct of $X$ and $Y$, with injections given by
+$\inj_1 = \prodM{\id}{0}$ and $\inj_2 = \prodM{0}{\id}$.
+\item If $Z$ is a coproduct of $X$ and $Y$, then it is also a product of $X$ and $Y$, with projections given by $\pi_1 =
+\coprodM{\id}{0}$ and $\pi_2 = \coprodM{0}{\id}$.
 \item $X$ is terminal iff it is initial.

### 2025-07-05 main/first-order.tex
Tweak.

@@
 \begin{enumerate}
-\item If $Z$ is a product of $X$ and $Y$ then it is also coproduct of $X$ and $Y$, with injections given by
-$\inj_1 = \prodM{\id}{0}$ and $\inj_2 = \prodM{0}{\id}$.
-\item If $Z$ is a coproduct of $X$ and $Y$, then it is also a product of $X$ and $Y$, with projections given by $\pi_1 =
+\item Any product of $X$ and $Y$ is also coproduct of $X$ and $Y$, with injections given by $\inj_1 =
+\prodM{\id}{0}$ and $\inj_2 = \prodM{0}{\id}$.
+\item Any coproduct of $X$ and $Y$  is also a product of $X$ and $Y$, with projections given by $\pi_1 =
 \coprodM{\id}{0}$ and $\pi_2 = \coprodM{0}{\id}$.

### 2025-07-07 main/approx-as-tangents.tex
Remove duplicate definition of Galois connection.
[ tex ] flesh out the approximations as tangents section

@@
 
-The general study of differentiable functions takes place on \emph{manifolds}, which are topological spaces that ``locally'' look like a subset of the Euclidean space $\RR^n$.
-The spaces $\RR^n$ themselves are manifolds, but so are ``non-flat'' examples such as $n$-spheres and yet more exotic spaces.
-Every point $x$ in a manifold $M$ has an associated \emph{tangent vector space} $\tangents_x(M)$ and \emph{cotangent vector space} $\cotangents_x(M)$, the latter consisting of linear functions $\tangents_x(M) \linearto \RR$.
-The tangent and cotangent spaces are finite dimensional, so in the presence of a chosen basis they are canonically isomorphic.
-In the case when the manifold is $\RR^n$, then every tangent space is isomorphic to $\RR^n$ as well.
+The general study of differentiable functions takes place on
+\emph{manifolds}, topological spaces that ``locally'' behave like an
+open subset of the Euclidean space $\RR^n$. The spaces $\RR^n$
+themselves are manifolds, but so are ``non-flat'' examples such as
+$n$-spheres and yet more exotic spaces. Every point $x$ in a manifold
+$M$ has an associated \emph{tangent vector space} $\tangents_x(M)$
+consisting of linear approximations of curves on the manifold passing
+through $x$. Each point also has a \emph{cotangent vector space}
+$\cotangents_x(M) = \tangents_x(M) \linearto \RR$. The tangent and
+cotangent spaces are finite dimensional, so in the presence of a
+chosen basis they are canonically isomorphic. In the case when the
+manifold is $\RR^n$, then every tangent space is isomorphic to $\RR^n$
+as well.
 
@@
 \begin{itemize}
-\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$.
+\item The \emph{forward derivative} (tangent map, pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$.
 In the Euclidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
-\item The \emph{backward derivative} (cotangent map, or pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$.
+\item The \emph{backward derivative} (cotangent map, pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$.
 In the Euclidean case, the backward derivative is represented by the transpose of the Jacobian of $f$ at $x$.
 \end{itemize}
+
+\bob{mention chain rule here? and tangent spaces?}
 
@@
 
-\subsection{Stable Domain Theory as Differentiability}
+\subsection{Stable Functions as Differentiable Functions}
 
-Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitative approximations of elements in the sense of domain theory.
+Our thesis is that \GPS is a generalised form of differentiable programming, where tangents are not linear approximations of curves but instead are qualitative information approximations of elements.
 Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry79,berry82}.
@@
 
-Domain theory is a method for defining the semantics of programs that handle infinite data (e.g., functions or infinite streams).
-\emph{Domains} are certain partially ordered sets where the ordering denotes a relationship of qualitative information content: if $x \sqsubseteq y$, then $y$ may contain more information than $x$.
-For example, if $x$ and $y$ are functions, then $y$ may be defined for more values in its domain than $x$.
-Infinite objects are understood in terms of their approximations in this sense, and domains are assumed to be closed under least upper bounds (lubs) of directed sets, meaning that any internally consistent collection of elements has a ``completion'' that contains all the information covered by the set.
-Programs are interpreted as monotone functions that preserve directed lubs.
-Monotonicity captures the idea that if the input gets more defined, then the output can get more defined.
-Preservation of lubs, or \emph{continuity}, states that a function interpreting a program cannot act inconsistently on approximations and their completion, which corresponds to the intuitive idea that a function that is computable cannot look at a non-finite amount of input to produce a finite output.
+Domain theory is a method for defining the semantics of programs that
+handle infinite data such as functions or infinite streams. Domains
+are certain partially ordered sets where the ordering denotes a
+relationship of qualitative information content: if $x \sqsubseteq y$,
+then $y$ may contain more information than $x$. For example, if $x$
+and $y$ are functions, then $y$ may be defined at more values than
+$x$. Infinite objects are understood in terms of their approximations
+in this sense, and domains are assumed to be closed under least upper
+bounds (lubs) of directed sets, meaning that any internally consistent
+collection of elements has a ``completion'' that contains all the
+information covered by the set. Programs are interpreted as monotone
+functions that preserve directed lubs. Monotonicity captures the idea
+that if the input gets more defined, then the output can get more
+defined. Preservation of lubs, or \emph{continuity}, states that a
+function interpreting a program cannot act inconsistently on
+approximations and their completion, which corresponds to the
+intuitive idea that a function that is computable cannot look at a
+non-finite amount of input to produce a finite output. \cite{abramsky-jung} is a comprehensive introduction to domain theory.
 
-For the purposes of \GPS, we are not interested in using approximations to model computation on infinite objects, but instead to use them for the related use of revealing how programs explore their inputs when producing parts of their output.
-Therefore, we ignore completeness properties of the partially ordered sets we consider.
+For the purposes of \GPS, we are not interested in using
+approximations to model computation on infinite objects, but instead
+for revealing how programs explore their inputs when producing parts
+of their output. Therefore, we ignore completeness properties of the
+partially ordered sets we consider\footnote{\bob{we might want to bring this back when considering general recursive programs.
+Are bidomains \cite{berry79} useful?
+they have two orderings.
+See also Laird's Bistable Biorders.}}.
 
@@
 
-When giving a denotational semantics for sequential programming languages, such as PCF, continuous functions are too permissive.
-Plotkin's Parallel OR (\exref{parallel-or}, below) is continuous but does not explore its input in a way consistent with a sequential implementation \cite{plotkin77lcf}.
-Continuous functions can explore their input in a ``random access'' way, which is incompatible with the central idea in \GPS that we should be able to identify a \emph{minimal} part of the input that leads to a part of the output.
+When giving a denotational semantics for sequential programming
+languages Scott-continuous functions are too permissive. Famously,
+\citet{plotkin77lcf}'s Parallel OR (\exref{parallel-or}, below) is
+continuous but does not explore its input in a way consistent with a
+sequential implementation. Continuous functions can explore their
+input in a non-deterministic way as long as the result is
+deterministic. This non-determinism results in functions that whose
+output cannot be assigned a unique minimal input that caused
+it. Therefore, continuous functions are incompatible with the central
+idea in \GPS that we should be able to identify a \emph{minimal} part
+of the input that leads to a part of the output.
 
-Stability is a property that can be required of monotone functions that was invented by \citet{berry79} in an attempt to capture sequentiality.
+\emph{Stability} is a property that can be required of monotone functions that was invented by \citet{berry79} in an attempt to capture sequentiality.
 This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
@@
 In a partially ordered set $X$, for any $x \in X$ the set of elements below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a partially ordered set.
-These approximations of $x$ we will think of as ``derivatives'' of $x$, and the whole set $\downset{x}$ as the ``tangent space''.
-Tangent spaces are vector spaces, since they are linear approximations to curves at that point.
-In the partially ordered setting, we take elements of $\downset{x}$ to be approximations of processes defined at $x$.
+These approximations of $x$ we will think of as ``tangents'' at $x$, and the whole set $\downset{x}$ as the ``tangent space''\footnote{\bob{Is it possible to derive the tangent spaces by thinking about continuous functions passing through $x$?}}.
+Tangent spaces are vector spaces, so in the partially ordered setting we take elements of $\downset{x}$ to be approximations of processes defined at $x$.
 As we can add tangents, we assume we can take meets of approximations in $\downset{x}$:
@@
 
-Smooth functions have a ``pushforward'' derivative that takes tangents at a point $x$ to tangents at the point $f(x)$ and is linear.
-In the partially ordered setting, we are going to take a function $f$'s pushforward derivative at $x$ to be its restriction to $\downset{x}$, taking approximations of the input to approximations of the output.
-Matching the linearity of the pushforward derivative smooth functions, we require that these restrictions preserve meets:
+For the approximation version of the forward derivative of $f$ at $x$, we take $f$'s restriction to $\downset{x}$, taking approximations of the input to approximations of the output.
+Matching the linearity of the forward derivative of smooth functions, we require that these restrictions preserve meets.
+This is exactly the definition of \emph{conditionally multiplicative} function from \citet{berry79}:
 
@@
 
-\begin{lemma}
-  Morphisms of meet approximation spaces are closed under identities
-  and composition, forming a category. FIXME: with products?
-  functions? sums?
-\end{lemma}
+\begin{theorem}
+  Bounded meet lattices and conditionally multiplicatve functions form
+  a category $\CM$. This category has products, coproducts and
+  exponentials.
+\end{theorem}
+
+\begin{proof}
+  \AGDA. See \citet{amadio-curien} for the case when the posets are
+  also cpos. The crucial technical step, identified by
+  \citet{berry79}, is that the ordering on conditionally
+  multiplicative functions is not the extensional ordering
+  ($f \sqsubseteq g$ iff $\forall x. f(x) \subseteq g(x)$) but instead
+  the stable ordering $f \sqsubseteq g$ iff \bob{FIXME}.
+\end{proof}
 
@@
   \label{ex:strict-short-circuit}
-  To see the effect of requiring preservation of bounded meets,
-  consider several ways of defining disjunction on the lifted booleans
+  To see the effect of conditional multiplicativity, consider several
+  ways of defining the OR on the lifted booleans
   $\mathbb{B}_\bot$. Two functions that are cm are the strict and
@@
 
-Conditional multiplicativity seems to be a reasonable analogue to functions with a well-defined notion of derivative.
+\begin{example}[Intervals and Maximal Elements]
+  \label{ex:intervals-and-maxima-elements}
+  The set of intervals
+  $\{ [l,u] \in \mathbb{R} \times \mathbb{R} \mid l \leq u \}$ ordered
+  by reverse inclusion forms a (Scott) domain
+  \cite{scott-outline}. The set of maximal elements is exactly
+  $\mathbb{R}$. This domain has been proposed as a model of
+  approximate real number computation \cite{escardo-real-pcf}. The
+  information approximation reading is intuitive: as intervals move up
+  the order they become tighter, containing more information about the
+  number they are approximating.
+
+  Given this reading, it makes sense to wonder if we can use interval
+  approximations as ``information tangents'' of real numbers, where
+  derivatives take approximating intervals to approximating
+  intervals. Since intervals form a Scott domain, they are bounded
+  complete and hence have bounded meets. However, the addition
+  function on intervals, $[l_1,u_1] + [l_2,u_2] = [l_1+l_2,u_1+u_2]$,
+  is not conditionally multiplicative, as can be easily checked.
+
+  A solution is to use the set of intervals with nominated points that
+  they are approximating: $\{[l,x,u] \mid l \leq x \leq u\}$. The
+  ordering now is that $[l_1,x_1,u_1] \sqsubseteq [l_2,x_2,u_2]$ iff
+  $x_1 = x_2$ and $l_1 \leq l_2$ and $u_2 \leq u_1$. Consequently, the
+  maximal elements are $[x,x,x]$, recovering $\mathbb{R}$ again, but
+  the approximations of each number all form independent sub-lattices.
+  Addition is defined as
+  $[l_1,x_1,u_1] + [l_2,x_2,u_2] = [(l_1+x_2) \sqcap
+  (l_2+x_1),x_1+x_2,(u_1+x_2)\sqcup(u_2+x_1)]$, which is conditionally
+  multiplicative. Note how this definition bears a resemblance to the
+  product rule for derivatives, with (in the lower end of the
+  interval) $\sqcap$ replacing $+$.
+
+  This example is important to us because it shows that (for total
+  programs) we are separately interested in the maximal elements and
+  their approximations, and that approximations of each maximal
+  element may have to be considered separately.
+
+  Relatedly, \citet{edalat-heckmann98} proposed a Scott domain of
+  formal balls on a metric space, where again the maximal elements are
+  the points of the original space. More generally,
+  \citet{continuous-lattice-book} (Section V-6) describes \emph{Domain
+    environments}, which are domains whose maximal elements are
+  exactly the points of a topological space. We are not aware of any
+  work linking domain environments in general to stable functions. It
+  would be interesting to see whether the situation for intervals,
+  where approximations must be relative to a nominated point, repeats
+  in general when considering conditionally multiplicative functions.
+\end{example}
+
+Given these examples, conditional multiplicativity seems to be a reasonable analogue to functions with a well-defined notion of derivative.
 However, for \GPS we also require an analogue to the reverse derivative, where we map approximations backwards to give the least approximation of the input for a given approximation of the output.
@@
   % backwards propagation of approximations.
-  An example that is conditionally multiplicative, but does not admit a backards map of approximations (from \citet{amadio-curien} just before Lemma 12.2.3, originally due to Berry) is given by defining $\mathrm{unstable} : D \to \{\bot, \top\}$, where:
-  \begin{displaymath}
-    D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots \sqsubseteq 1 \sqsubseteq 0
-  \end{displaymath}
-  as $\mathrm{unstable}(\bot) = \bot$ and $\mathrm{unstable}(n) = \top$.
-This is monotone, and preserves meets in every $\downset{x}$.
-However, there is no ``best'' (i.e., least) input that gives us any finite output.
+  An example that is conditionally multiplicative, but does not admit
+  a backards map of approximations (from \citet{amadio-curien} just
+  before Lemma 12.2.3, originally due to Berry) is given by defining
+  $\mathrm{unstable} : D \to \{\bot, \top\}$, where
+  $D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots
+  \sqsubseteq 1 \sqsubseteq 0$, as $\mathrm{unstable}(\bot) = \bot$
+  and $\mathrm{unstable}(n) = \top$. This is monotone, and preserves
+  meets in every $\downset{x}$. However, there is no ``best'' (i.e.,
+  least) input that gives us any finite output.
 \end{example}
@@
 
-In light of the preceeding example, we can give an alternative
-definition of a function that requires the existence of a reverse
-mapping directly, even without assuming that any meets exist:
+In light of the \exref{non-stable-function}, we consider
+\citet{berry79}'s definition of \emph{stable function} that requires
+the existence of a reverse mapping directly, even without assuming
+that any meets exist:
 
@@
   \begin{enumerate}
-  \item The function $\mathrm{strictOr}$ is stable.
-For example, for the input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input that gives this output is exactly $(\mathsf{tt}, \mathsf{ff})$.
-If we take the approximation $\bot \leq \mathsf{tt}$ of the output, then the corresponding minimal input is $(\bot, \bot)$.
-The function $\mathrm{shortCircuitOR}$ is also stable.
-For the input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input that gives this input is $(\mathsf{tt},\bot)$, indicating that the presence of $\mathsf{ff}$ in the second argument was not necessary to produce this output.
-As with $\mathrm{strictOr}$, the minimal input required to produce the output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.
+  \item The function $\mathrm{strictOr}$ is stable. For example, for
+    the input-output pair
+    $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input
+    that gives this output is exactly $(\mathsf{tt}, \mathsf{ff})$. If
+    we take the approximation $\bot \leq \mathsf{tt}$ of the output,
+    then the corresponding minimal input is $(\bot, \bot)$. The
+    function $\mathrm{shortCircuitOR}$ is also stable. For the
+    input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$,
+    the minimal input that gives this input is $(\mathsf{tt},\bot)$,
+    indicating that the presence of $\mathsf{ff}$ in the second
+    argument was not necessary to produce this output. As with
+    $\mathrm{strictOr}$, the minimal input required to produce the
+    output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.
 
-  \item The $\mathrm{parallelOR}$ function is not stable.
-For the input-output pair $(\mathsf{tt},\mathsf{tt}) \mapsto \mathsf{tt}$, there is no one minimal input that produces this output.
-We have both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are incomparable and their greatest lower bound $(\bot,\bot)$ gives the output $\bot$.
+  \item The $\mathrm{parallelOR}$ function is not stable. For the
+    input-output pair $(\mathsf{tt},\mathsf{tt}) \mapsto \mathsf{tt}$,
+    there is no one minimal input that produces this output. We have
+    both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and
+    $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are
+    incomparable and their greatest lower bound $(\bot,\bot)$ gives
+    the output $\bot$.
 
-  \item The $\mathrm{gustave}$ function is stable.
-Despite there being no one minimal input that achieves the output $\top$, each of the minimal inputs that can achieve this output are pairwise incomparable, so for each specific input that gets output $\top$ there is a unique minimal input that achieves it (listed in the first three lines of the definition).
+  \item The $\mathrm{gustave}$ function is stable. Despite there being
+    no one minimal input that achieves the output $\top$, each of the
+    minimal inputs that can achieve this output are pairwise
+    incomparable, so for each specific input that gets output $\top$
+    there is a unique minimal input that achieves it (listed in the
+    first three lines of the definition).
 
-    In terms of \GPS the $\mathrm{gustave}$ function does not present a problem.
-For any particular run (i.e., input $\mapsto$ output pair), there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.
+    In terms of \GPS the $\mathrm{gustave}$ function does not present
+    a problem. For any particular run (i.e., input $\mapsto$ output
+    pair), there is an unambiguous minimal input that achieves the
+    output, no matter that it was not achieved by a sequential
+    processing of the input.
 
-  \item As discussed above, $\mathrm{unstable}$ is not stable, but is conditionally multiplicative.
+  \item As discussed above, $\mathrm{unstable}$ is not stable, but is
+    conditionally multiplicative.
+
+  \item The addition function on intervals with nominated points in
+    \exref{intervals-and-maxima-elements} is stable. Given the input
+    $[l_1,x_1,u_1], [l_2,x_2,u_2]$ and an approximation
+    $[l,x_1+x_2,u]$ of the output, the minimal approximations of the
+    input are $[l-x_2,x_1,u-x_2], [l-x_1,x_2,u-x_1]$. We can read this
+    as saying if the output was the maximal element $x_1+x_2$ but we
+    only require the output to be in the range $[l,x_1+x_2,u]$, then
+    we can obtain intervals containing the input values that are
+    enough to obtain the desired output approximation \emph{assuming
+      that the other input is kept the same}. Note the analogy to
+    partial derivatives in multi-variable calculus, where the
+    derivative is computed in each variable independently.
   \end{enumerate}
@@
 
-The fact that these two functions' stability witnesses reveal information about how they depend on their input is what we will exploit in order to use the idea of stability for \GPS.
-It is interesting that sequentiality is not necessarily required in order to obtain backwards maps of approximations, as might be assumed by thinking about how to trace computations for provenance information.
+% The fact that these two functions' stability witnesses reveal
+% information about how they depend on their input is what we will
+% exploit in order to use the idea of stability for \GPS. It is
+% interesting that sequentiality is not necessarily required in order
+% to obtain backwards maps of approximations, as might be assumed by
+% thinking about how to trace computations for provenance information.
 
-Stability has an alternative definition in terms of Galois connections, which will be more useful for what follows.
-This characterisation is due to~\citet{taylor99}.
-We first formally define Galois connections, which we used informally in \exref{introduction-example}.
+Stability has an alternative definition in terms of Galois
+connections, which will be more useful for what follows. This
+characterisation is due to~\citet{taylor99}. We first define Galois
+connections, which we used informally in \exref{introduction-example}.
 
@@
 A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
-We refer to $f$ as the left adjoint and $g$ as the right adjoint.
+Since a Galois connection is also an adjunction, we refer to $f$ as the left adjoint and $g$ as the right adjoint.
 \end{definition}
@@
 \begin{lemma}
-  A monotone function $f : X \to Y$ is stable if and only if for all $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$ has a left Galois adjoint.
+  A monotone function $f : X \to Y$ is stable if and only if for all
+  $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$
+  has a left Galois adjoint.
 \end{lemma}
@@
 
-Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets, preserved by forward approximation maps, and joins, preserved by backwards approximation maps, exist:
+Even though stable functions can be defined on any partially ordered
+set, in light of the analogy with tangent spaces it makes sense to
+require that meets, preserved by forward approximation maps, and
+joins, preserved by backwards approximation maps, exist:
 
@@
 
-We could have required that L-posets' principal downsets were complete lattices, which would mean that we would only need to require that forward approximation maps preserve all meets and we could compute the backwards map automatically.
+\begin{theorem}
+  L-posets and stable functions form a category with products and
+  coproducts.
+\end{theorem}
+
+The category of L-posets and stable functions is not cartesian
+closed. To make it so, we would need to require that the principal
+downsets $\downset{x}$ are \emph{complete}
+lattices.
+\citet{amadio-curien} (Theorem 12.FIXME) details the proof.
+Intuitively, to generate the best approximation of an input value for a function, we need to take the infimum over all possible input values.
+
+Our goal is to model a higher-order language suitable for writing
+queries on databases as in \exref{introduction-example}, so why should
+we not just take complete L-posets and stable functions as our model?
+We have two reasons for moving to a slightly different model in \secref{model}:
+
+\begin{enumerate}
+\item Even without completeness, in Bounded meet posets and L-posets,
+  values and their approximations live in the same set. However, in
+  the \emph{total} query language we wish to model in
+  \secref{language}, we are not directly interested in the behaviour
+  of programs on approximations as we would be for partial programs
+  with general recursion. (Moreover, in the light of
+  \exref{intervals-and-maxima-elements} it is not clear whether
+  approximations for partiality and approximations for stability ought
+  to be the same thing. \bob{fwd reference to discussion.}) In
+  \exref{intervals-and-maxima-elements}, maximal elements could be
+  taken to be the ``proper values''. One idea is to restrict to
+  conditionally multiplicative or stable functions that preserve
+  maximal elements. However, this idea fails at higher order:
+  functions that take maximal elements to maximal elements are not
+  themselves maximal elements. We could devise a category of L-posets
+  with totality predicates (which would pick out maximal elements at
+  first-order) and totality preserving functions, but we prefer a more
+  direct method of separating values proper from their approximations
+  using the \emph{Category of Families} construction we detail in the
+  next section.
+\item A more local reason is that we wish to formalise our
+  construction in the proof assistant Agda \cite{agda} in order to get
+  an \emph{executable} model of the language in
+  \secref{language}. Agda's type theory is both predicative and
+  constructive. Predicativity means that it does not have complete
+  lattices in the classical sense: for a type $X$ we can only get
+  suprema and infinma of families in a lower universe level than
+  $X$. This is not necessarily a problem, as \citet{dejong21} shows
+  how to develop a large amount of domain theory in a predicative
+  setting. However, due to constructivity the lifting construction
+  (due to \citet{escardo-knapp}) requires a proof of definedness to
+  extract a value. Since we are modelling a total language, we prefer
+  to have a model that computes directly.
+\end{enumerate}
+
+\bob{For future work: what is the connection between FamCPO(LatGal)
+  and Laird's bistable biorders, or Berry bidomains?}
+
+% We could have required that L-posets' principal downsets were complete lattices, which would mean that we would only need to require that forward approximation maps preserve all meets and we could compute the backwards map automatically.
 However, our goal is a computable theory for \GPS, analogous to automatic differentiation.
@@
 
-\begin{proposition}
-  L-posets and stable functions form a category. \bob{and the main point is that composition of stability witnesses looks like a chain rule!}
-\end{proposition}
+% \begin{proposition}
+%   L-posets and stable functions form a category. \bob{and the main point is that composition of stability witnesses looks like a chain rule!}
+% \end{proposition}
 
-\bob{L-poset is has products and coproducts, but is not closed.
+% \bob{L-poset is has products and coproducts, but is not closed.
 To be closed would require each $\downset{x}$ to be a \emph{complete} lattice.
@@
 
-Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
+We have seen that \citet{berry79}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives.
+In the next section, we describe a model based on these ideas suitable for constructing executable models of total higher order languages.
 
-\begin{itemize}
-\item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
-\item For \GPS, every program has an associated forward approximation map that takes approximations forward from the input to the output.
+We end this section with a conjecture. Although we have argued above
+that there is an analogy between stable functions and smooth
+functions, we have not stated any mathematical theorems substantiating
+this. \emph{Tangent Categories} \cite{tangent-categories} are a
+categorical axiomatisation of the properties of manifolds and smooth
+functions in terms of the presence of tangent bundles $TX$ for every
+object $X$, forward derivatives and additivity of tangents. They
+generalise Cartesian Differential Categories \cite{cdcs}, which are an
+axiomatisation of Euclidean spaces and smooth functions.
+
+\begin{conjecture}
+  \label{con:tangent-stable-fns}
+  (1) Bounded meet posets and conditionally multiplicative functions
+  form a Tangent category where the tangent bundle
+  $TX = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
+  by meets. (2) L-posets and stable functions form a \emph{reverse
+    Tangent category} \cite{reverse-tangents}.
+\end{conjecture}
+
+% Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
+
+% \begin{itemize}
+% \item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
+% \item For \GPS, every program has an associated forward approximation map that takes approximations forward from the input to the output.
 This map {\em preserves meets}.
@@
 The forward derivative map is {\em linear}, so it preserves addition of tangents and the zero tangent.
-\item For \GPS, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input.
+% \item For \GPS, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input.
 This map {\em preserves joins}.
@@
 This map is again {\em linear}.
-\item For \GPS, the forward and backward approximation maps are related by being a Galois connection.
+% \item For \GPS, the forward and backward approximation maps are related by being a Galois connection.
 For differentiable programming, the forward and reverse derivatives are related by being each others' transpose.
-\end{itemize}
+% \end{itemize}
 
-\bob{Put a conjecture about Tangent Categories in somewhere}
+% \bob{Put a conjecture about Tangent Categories in somewhere}
 
-Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.
+% Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.
 This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wide range of computational settings.

### 2025-07-07 main/models.tex
[ tex ] start on the model construction

\section{Constructing Models of Total \GPS}
\label{sec:models-of-total-gps}

TODO:
\begin{enumerate}
\item Link to the discussion in the previous section
\item Category of Families construction over a category $\cat{C}$:
  \begin{enumerate}
  \item Has (all) coproducts
  \item Has products when $\cat{C}$ does, and is distributive and
    stable.
  \end{enumerate}
\item Categories of interest: $\LatGal$, $\MeetSLat$, $\JoinSLat$ and
  (for normal differentiation) $\FinVect$.
\item Some functors between these models and from L-poset. Some
  interesting objects in these categories.
\item For higher-order models: we use a theorem of Vákár and Lucatelli
  Nunes: if $\cat{C}$ has is cmon-enriched with biproducts and has all
  (small) products, then $\Fam(\cat{C})$ is cartesian closed.
\item We can construct a list object in $\Fam(\cat{C})$ when it has
  products.
\end{enumerate}

We have seen that the idea of stability from domain theory can be seen
as a bridging concept between smooth functions and their derivatives
on one hand, and \GPS on the other. For the next part of our
development, we want to transport a mathematical framework for
automatic differentation from normal differentation to \GPS. We take
the \emph{Combinatory Homomorphic Automatic Differentation} (CHAD)
framework of Vákár \etal{} and apply it to \GPS. In this section we
develop this framework for first-order programs only. In
\autoref{sec:chad-higher-order}, we extend the framework to deal with
the whole higher-order language we present in \autoref{sec:language}.

\subsection{CMon-enriched and Biproduct Categories}

As we noted above\bob{where?}, one commonality between tangent spaces
and approximations is the ability to add tangents and to take the meet
or join of two approximations. For our purposes, the suitable common
abstraction is given by categories enriched in commutative monoids.

\begin{definition}
  \bob{CMon-categories}
\end{definition}

\begin{definition}
  \bob{Biproducts}
\end{definition}

\begin{example}
  FinVect
\end{example}

\begin{example}
  $\LatGal$, the category of bounded lattices and Galois connections.
\end{example}

\subsection{Families}

The basic idea behind the CHAD framework is to use Cmon-categories
with biproducts as the representation of tangent objects. Tangent
objects are associated with points formally by constructing a category
where objects consist of a pair of a set (of points) and for each
point a corresponding tangent object. This is exactly the
\emph{families} construction that freely completes any category with
all coproducts:

\begin{definition}
  \label{def:category-of-families}
  For any category $\cat{C}$, the \emph{category of families over
    $\cat{C}$}, $\Fam(\cat{C})$ has:
  \begin{itemize}
  \item Objects $(A, X)$, where $A$ is a set and
    $X : A \to \Ob\cat{C}$ is an $A$-indexed familiy of objects of
    $\cat{C}$;
  \item Morphisms $(A, X) \to (B, Y)$ are pairs $(f, \partial f)$
    where $f : A \to B$ is a function and
    $\partial f : \Pi_{x : A}.~\cat{C}(X(x), Y(f\,x))$ is an
    $A$-indexed family of morphisms of $\cat{C}$.
  \end{itemize}
\end{definition}

The key observation is that the composition of morphisms in Fam(C)
matches the \emph{chain rule} for composition of
derivatives. Therefore, in Fam(FinVect), if we have two morphisms
$(f, \partial f)$ and $(g, \partial g)$ where the second components
are the actual derivatives of the first components, then their
composition's second component is the derivative of the
composition. More formally:\bob{is this in Vákár's work?}

\begin{proposition}
  There is a faithful functor $F : \Man \to \Fam(\FinVect)$ that sends
  a manifold $M$ to $(M, \lambda x. T_x(M))$, and each smooth function
  $f$ to $(f, f_*)$, the pair of $f$ and its forward derivative.
\end{proposition}

As similar result is given in \cite{cruttwell2022}, where Euclidean
spaces $\RR^n$ and smooth functions are embedded into a category of
lenses (the ``simply typed'' version of the $\Fam$ construction).

For \GPS, we have an analogous result:

\begin{lemma}
  Faithful functor from L-poset to LatGal.
\end{lemma}

\begin{lemma}
  Fam(C) has all coproducts.
\end{lemma}

\begin{lemma}
  Fam(C) has products when C does.
\end{lemma}

More generally, get monoidal products.

\subsection{\GPS in Fam(LatGal)}

\begin{enumerate}
\item The lifting monad
\item Show some examples
\end{enumerate}

### 2025-07-07 main/approx-as-tangents.tex
[ tex ] Models of Galois Slicing for a Total Language

@@
 \subsubsection{Stable functions and L-posets}
+\label{sec:stab}
 
@@
 
+As well as codifying exactly what we mean by differentiable structure
+on partially ordered sets, proving this conjecture would also tell us
+what higher derivatives mean in this context as well, something that
+we have not considered above. We will extend this conjecture to our
+refined model of partially ordered approximations in
+\secref{model-summary}.
+
 % Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:

### 2025-07-07 main/models.tex
[ tex ] Models of Galois Slicing for a Total Language

@@
-\section{Constructing Models of Total \GPS}
+\section{Constructing Models of \GPS for a Total Language}
 \label{sec:models-of-total-gps}
 
-TODO:
-\begin{enumerate}
-\item Link to the discussion in the previous section
-\item Category of Families construction over a category $\cat{C}$:
-  \begin{enumerate}
-  \item Has (all) coproducts
-  \item Has products when $\cat{C}$ does, and is distributive and
-    stable.
-  \end{enumerate}
-\item Categories of interest: $\LatGal$, $\MeetSLat$, $\JoinSLat$ and
-  (for normal differentiation) $\FinVect$.
-\item Some functors between these models and from L-poset. Some
-  interesting objects in these categories.
-\item For higher-order models: we use a theorem of Vákár and Lucatelli
-  Nunes: if $\cat{C}$ has is cmon-enriched with biproducts and has all
-  (small) products, then $\Fam(\cat{C})$ is cartesian closed.
-\item We can construct a list object in $\Fam(\cat{C})$ when it has
-  products.
-\end{enumerate}
+The previous section concluded that L-posets and stable functions give
+a model of \GPS analogous to manifolds and smooth functions. However,
+we noted a conceptual shortcoming of this model, for the purposes of
+modelling total computations, that proper values and their
+approximations live in the same category. In this section, we propose
+a model for total \GPS based on the \emph{Category of Families}
+construction. This construction, adn the more general Grothendieck
+construction, has been previously used by V{\'a}k{\'a}r and
+collaborators \bob{CITE} to model automatic differentiation for
+higher-order programs on the reals. We rely on some of their results,
+and discuss the commonalities as we go.
 
-We have seen that the idea of stability from domain theory can be seen
-as a bridging concept between smooth functions and their derivatives
-on one hand, and \GPS on the other. For the next part of our
-development, we want to transport a mathematical framework for
-automatic differentation from normal differentation to \GPS. We take
-the \emph{Combinatory Homomorphic Automatic Differentation} (CHAD)
-framework of Vákár \etal{} and apply it to \GPS. In this section we
-develop this framework for first-order programs only. In
-\autoref{sec:chad-higher-order}, we extend the framework to deal with
-the whole higher-order language we present in \autoref{sec:language}.
+\subsection{The Category of Families Construction}
 
-\subsection{CMon-enriched and Biproduct Categories}
-
-As we noted above\bob{where?}, one commonality between tangent spaces
-and approximations is the ability to add tangents and to take the meet
-or join of two approximations. For our purposes, the suitable common
-abstraction is given by categories enriched in commutative monoids.
+As we explained in \secref{stab}, L-posets can be seen as partially
+ordered sets where every principal downset $\downset{x}$ is a bounded
+lattice. That is, for an element $x$, $\downset{x}$ is its lattice of
+approximations/tangents. The shortcoming of this setup is that proper
+values and their approximations live in the same set. We fix this by
+changing our model to one where we have sets $X$ of values, and for
+each $x \in X$, a bounded lattice $\partial X(x)$ of approximations of
+$x$. This construction is an instance of the general \emph{Category of
+  Families} construction:
 
 \begin{definition}
-  \bob{CMon-categories}
+  Let $\cat{C}$ be a category. The \emph{Category of Families} over
+  $\cat{C}$, $\Fam(\cat{C})$, has as objects pairs $(X, \partial X)$,
+  where $X$ is a set and $\partial X : X \to \cat{C}$ is an
+  $X$-indexed family of objects in $\cat{C}$. A morphism
+  $f : (X, \partial X) \to (Y, \partial Y)$ consists of a pair of a
+  function $f : X \to Y$ and a familiy of morphisms of $\cat{C}$,
+  $\partial f : (x : X) \to \cat{C}(\partial X(x), \partial Y(f\,x))$.
 \end{definition}
 
-\begin{definition}
-  \bob{Biproducts}
-\end{definition}
-
-\begin{example}
-  FinVect
-\end{example}
-
-\begin{example}
-  $\LatGal$, the category of bounded lattices and Galois connections.
-\end{example}
-
-\subsection{Families}
-
-The basic idea behind the CHAD framework is to use Cmon-categories
-with biproducts as the representation of tangent objects. Tangent
-objects are associated with points formally by constructing a category
-where objects consist of a pair of a set (of points) and for each
-point a corresponding tangent object. This is exactly the
-\emph{families} construction that freely completes any category with
-all coproducts:
-
-\begin{definition}
-  \label{def:category-of-families}
-  For any category $\cat{C}$, the \emph{category of families over
-    $\cat{C}$}, $\Fam(\cat{C})$ has:
-  \begin{itemize}
-  \item Objects $(A, X)$, where $A$ is a set and
-    $X : A \to \Ob\cat{C}$ is an $A$-indexed familiy of objects of
-    $\cat{C}$;
-  \item Morphisms $(A, X) \to (B, Y)$ are pairs $(f, \partial f)$
-    where $f : A \to B$ is a function and
-    $\partial f : \Pi_{x : A}.~\cat{C}(X(x), Y(f\,x))$ is an
-    $A$-indexed family of morphisms of $\cat{C}$.
-  \end{itemize}
-\end{definition}
-
-The key observation is that the composition of morphisms in Fam(C)
-matches the \emph{chain rule} for composition of
-derivatives. Therefore, in Fam(FinVect), if we have two morphisms
-$(f, \partial f)$ and $(g, \partial g)$ where the second components
-are the actual derivatives of the first components, then their
-composition's second component is the derivative of the
-composition. More formally:\bob{is this in Vákár's work?}
+Composition $f \circ g $ of morphisms
+$f : (Y, \partial Y) \to (Z, \partial Z)$ and
+$g : (X, \partial X) \to (Y, \partial Y)$ in this category is given by
+normal function composition on the set components, and
+$\partial (f \circ g)(x) = \partial f(f, x) \circ \partial g(x)$,
+where the latter composition is in $\cat{C}$. The fact that
+composition of morphisms in $\Fam(\cat{C})$ resembles the chain rule
+from calculus is essential to its use in modelling automatic
+differentiation. Indeed, if we let $\FinVect$ be the category of
+finite dimensional real vector spaces, then:
 
 \begin{proposition}
-  There is a faithful functor $F : \Man \to \Fam(\FinVect)$ that sends
-  a manifold $M$ to $(M, \lambda x. T_x(M))$, and each smooth function
+  \label{prop:embed-manifolds}
+  There is a faithful functor $\Man \to \Fam(\FinVect)$ that sends a
+  manifold $M$ to $(M, \lambda x. T_x(M))$, and each smooth function
   $f$ to $(f, f_*)$, the pair of $f$ and its forward derivative.
@@
 
-As similar result is given in \cite{cruttwell2022}, where Euclidean
+As similar result is given by \citet{cruttwell2022}, where Euclidean
 spaces $\RR^n$ and smooth functions are embedded into a category of
-lenses (the ``simply typed'' version of the $\Fam$ construction).
+lenses (the ``simply typed'' version of the $\Fam$ construction). As
+in \cite{vakar2021}, the idea is to formally separate functions on
+points and their forward/reverse tangent maps for the purposes of
+implementation of automatic differentiation. In the case of smooth
+maps, this process throws away information on higher derivatives by
+turning smooth maps into pairs of plain functions and linear
+functions. We conjecture at the end of this section that the analogous
+construction in our partially ordered setting does not.
 
-For \GPS, we have an analogous result:
+For \GPS we replace the category $\FinVect$ with $\LatGal$:
 
-\begin{lemma}
-  Faithful functor from L-poset to LatGal.
-\end{lemma}
+\begin{definition}
+  The category $\LatGal$ has bounded lattices as objects and Galois
+  connections as morphisms, with the right adjunct going in the
+  ``forward'' direction.
+\end{definition}
 
-\begin{lemma}
-  Fam(C) has all coproducts.
-\end{lemma}
+We can get an analogous result to \propref{embed-manifolds} for L-posets:
 
-\begin{lemma}
-  Fam(C) has products when C does.
-\end{lemma}
+\begin{proposition}
+  There is a faithful functor $\Lposet \to \Fam(\LatGal)$ that maps an
+  L-poset $X$ to $(X, \lambda x. \downset{x})$ and stable functions
+  $f$ to $(f, \lambda x. (f_x, f^*_x))$.
+\end{proposition}
 
-More generally, get monoidal products.
+Despite its similarity, this proposition has a lesser status than
+\propref{embed-manifolds} because it is not clear that the category
+$\Lposet$ is in any way a canonical definition of approximable sets
+and functions with approximation derivatives. We could restrict
+$\Lposet$ to morphisms that preserve maximal elements and take
+L-posets $X$ to $(\mathrm{Max}(M), \lambda x. \downset{x})$. However,
+our current hypothesis is that $\Fam(\LatGal)$, where values and their
+approximations are separated by construction is the primary model of
+semantic \GPS. We now investigate the categorical properties of this
+category, with a view to modelling a higher-order total programming
+language.
 
-\subsection{\GPS in Fam(LatGal)}
+\subsection{Categorical Properties of $\Fam(\cat{C})$}
 
+\subsubsection{Coproducts and Products}
+
+The categories $\Fam(\cat{C})$ are the free coproduct completions of
+categories $\cat{C}$, so they are guaranteed to have all coproducts:
+
+\begin{proposition}
+  For any $\cat{C}$, $\Fam(\cat{C}$ has all coproducts, which can be
+  given on objects by:
+  \begin{displaymath}
+    \coprod_i (X_i, \partial X_i) = (\coprod_i X_i, \lambda (i, x_i).\, \partial X_i(x))
+  \end{displaymath}
+  Coproducts in $\Fam(\cat{C})$ are \emph{extensive}
+  \cite{carboni-lack-walters93} (Proposition 2.4).
+\end{proposition}
+
+For $\Fam(\cat{C})$ to have finite products, we need $\cat{C}$ to have
+finite products:
+
+\begin{proposition}
+  If $\cat{C}$ has finite products, then so does $\Fam(\cat{C})$. On
+  objects, binary products can be defined by:
+  \begin{displaymath}
+    (X, \partial X) \times (Y, \partial Y) = (X \times Y, \lambda (x, y). \partial X(x) \times \partial Y(y))
+  \end{displaymath}
+  Since $\Fam(\cat{C})$ is extensive, products and coproducts always
+  distribute.
+\end{proposition}
+
+Using the infinitary coproducts and finite products, we can construct
+a wide range of other useful semantic models of datatypes in
+$\Fam(\cat{C})$. For example lists can be constructed as the coproduct
+\begin{displaymath}
+  \mathrm{List}\,X = \coprod_{n \in \mathbb{N}} X^n
+\end{displaymath}
+where $X^0 = 1$ (the terminal object) and $X^{n+1} = X \times X^n$.
+
+Our category of interest, $\Fam(\LatGal)$ has coproducts and finite
+products, because $\LatGal$ has products. As we shall see below, these
+are also coproducts, which is essential to obtaining cartesian
+closure.
+
+\subsubsection{Cartesian Closure}
+
+For cartesian closure of categories of families, we rely on the
+following theorem of \citet{nunes2023}, specialised from their setting
+with the general Grothendieck construction to $\Fam(\cat{C})$. This
+relies on the definition of \emph{biproducts}, which we discuss below
+in \secref{biproducts}.
+
+\begin{theorem}[\cite{nunes2023}]
+  \label{thm:fam-closed}
+  \AGDA.  If $\cat{C}$ has biproducts and all products, then
+  $\Fam(\cat{C})$ is cartesian closed\footnote{More precisely, if
+    $\cat{C}$ has coproducts then we have a monoidal product on
+    $\Fam(\cat{C})$ which is closed by this construction. When these
+    coproducts are in fact biproducts, we get cartesian closure.}. On
+  objects, the internal hom can be given by:
+  \begin{displaymath}
+    (X, \partial X) \to (Y, \partial Y) = (\Pi_{x : X}. \Sigma_{y : Y}. \cat{C}(\partial X(x), \partial Y(y)), \lambda f. \Pi_{x : X}. \partial Y(\pi_1(f\, x)))
+  \end{displaymath}
+\end{theorem}
+
+Before we explore the hypotheses of this theorem, we give some
+intuition on how it relates to differentiation. The Set-component of
+$(X, \partial X) \to (Y, \partial Y)$ consists of exactly the
+morphisms of $\Fam(\cat{C})$, rephrased into a single object. When
+$\cat{C} = \FinVect$, these are functions with an associated linear
+map at every point, and when $\cat{C} = \LatGal$, these are functions
+with an associated Galois connection at every point. A tangent to a
+function is then defined to be a mapping from points in the domain to
+tangents in the codomain along the function.
+
+Unfortunately, neither of the categories $\LatGal$ or $\FinVect$
+satisfy the hypotheses of this theorem, because neither of them have
+infinite products. We will consider ways to rectify this below
+\bob{fwd ref}.
+
+There is another construction of internal homs on $\Fam(\cat{C})$
+arising from the use of fibrations for categorical logical relations
+proofs, due to \citet{hermida99} (Corollary 4.12). If we assume that
+$\cat{C}$ is itself cartesian closed and has all products, then we
+could construct an internal hom as:
+\begin{displaymath}
+  (X, \partial X) \to (Y, \partial Y) = (X \to Y, \lambda f. \Pi_{x : X}.\,\partial X(x) \to \partial Y(f\,x))
+\end{displaymath}
+However, for the purposes of modelling differentiable programs, this
+is fatally flawed in that neither of $\LatGal$ nor $\FinVect$ are
+cartesian closed, and there is no way of making them so without losing
+the property of having biproducts. We will use Hermida's construction
+in our definability proof in \secref{definability}, where we use a
+logical relations argument to show that every morphism definable in
+the higher order language is also first-order definable.
+
+\subsubsection{CMon-Categories and Biproducts}
+\label{sec:biproducts}
+
+Loosely stated, biproducts are objects that are both products and
+coproducts. They can be defined in any category, as shown by
+\citet{karvonen18}, but for our purposes it will be more convenient to
+use the shorter definition in categories enriched in commutative
+monoids:
+\begin{definition}
+  A category $\cat{C}$ is enriched in $\CMon$, the category of
+  commutative monoids, if every homset $\cat{C}(X,Y)$ is a commutative
+  monoid with $(+,0)$ and composition is bilinear:
+  \begin{salign*}
+    f \comp \zero = f = \zero \comp f
+  \end{salign*}
+  \begin{salign*}
+    (f + g) \comp h &= (f \comp h) + (g \comp h) \\
+    h \comp (f + g) &= (h \comp f) + (h \comp g)
+  \end{salign*}
+\end{definition}
+In any $\CMon$-category we can define what it means to be the
+biproduct of two objects:
+\begin{definition}
+  In a $\CMon$-category a biproduct is an object $X \biprod Y$
+  together with morphisms
+
+  \begin{center}
+    \begin{tikzcd}
+      X \arrow[r, "\biinj_X", shift left] &
+      X \biprod Y \arrow[l, "\biproj_X", shift left] \arrow[r, "\biproj_Y"', shift right] &
+      Y \arrow[l, "\biinj_Y"', shift right]
+    \end{tikzcd}
+  \end{center}
+
+  \noindent satisfying
+
+  \vspace{-3mm}
+  \begin{minipage}[t]{0.45\textwidth}
+    \begin{center}
+      \begin{salign*}
+        \biproj_X \comp \biinj_X &= \id_X \\
+        \biproj_Y \comp \biinj_X &= \zero_{X,Y}
+      \end{salign*}
+    \end{center}
+  \end{minipage}%
+  \begin{minipage}[t]{0.45\textwidth}
+    \begin{center}
+      \begin{salign*}
+        \biproj_Y \comp \biinj_Y &= \id_Y \\
+        \biproj_X \comp \biinj_Y &= \zero_{Y,X}
+      \end{salign*}
+    \end{center}
+  \end{minipage}
+
+  \begin{salign*}
+    (\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
+  \end{salign*}
+\end{definition}
+As the name suggests, biproducts in a category are both products and
+coproducts:
+
+\begin{proposition}
   \begin{enumerate}
-\item The lifting monad
-\item Show some examples
+  \item A $\CMon$-category that has biproducts $X \oplus Y$ for all
+    $X$ and $Y$ also has products and coproducts with
+    $X \times Y = X + Y = X \oplus Y$.
+  \item A $\CMon$-category with (co)products also has biproducts.
   \end{enumerate}
+\end{proposition}
+
+\begin{example}
+  The following categories are $\CMon$-enriched and have finite
+  products, and hence biproducts:
+  \begin{enumerate}
+  \item In $\FinVect$, morphisms are linear maps and so can be added
+    and have a zero map. Finite products are given by cartesian
+    products of the underlying sets, with the vector operations
+    defined pointwise.
+  \item In $\LatGal$, right adjoints are summed using meets and left
+    adjoints are summed using joins. The zero maps are given by the
+    constantly $\top$ and constrantly $\bot$ functions
+    respectively. Products are given by the cartesian product of the
+    underlying set.
+  \item The $\LatGal$ example splits into two separate categories
+    $\MeetSLat$ and $\JoinSLat$ of meet semi-lattices and join
+    semi-lattices respectively, with their respective homomorphisms as
+    the morphisms. As for $\LatGal$, these categories are
+    $\CMon$-enriched and have finite products, and so finite
+    biproducts.
+  \end{enumerate}
+\end{example}
+
+\subsubsection{Discrete Completeness}
+
+The second hypothesis of \thmref{fam-closed} is that the category
+$\cat{C}$ has all (i.e., infinite) products. This is required to
+gather together tangents for all of the points in the domain of the
+function. Unfortunately, neither $\FinVect$ nor $\LatGal$ is complete
+in this sense.
+
+In the case of $\FinVect$, the solution is to expand to the category
+of all vector spaces $\Vect$, where infinite direct products
+exist. Note that these infinite products are not biproducts because
+the vector space operations themselves are finitary. This is the
+solution that \cite{vakar2021} uses for the semantics of forward
+($\Fam(\Vect)$) and reverse ($\Fam(\Vect^\op)$) automatic
+differentiation for higher order programs. Since the forward and
+reverse derivatives of a smooth map intrinsically defined,
+\citet{vakar2021}'s correctness theorem can show that
+
+For $\LatGal$ we could expand to the category of complete lattices and
+Galois connections between them. From a classical mathematical point
+of view, this would give a model of \GPS that would likely be suitable
+for reasoning programs' behaviour and their forward and backward
+approximations. However, in terms of building an executable model
+inside the Agda proof assistant, and with an eye toward implementation
+strategies, we seek a more down-to-earth (i.e., predicative and
+constructive) solution. We will do this by separating the forward and
+backward parts of the Galois connections to act independently by
+moving to the product category $\MeetSLat \times
+\JoinSLat^\op$. Objects in this category consist of \emph{separate}
+meet- and join-semi-lattices and potentially unrelated forward
+meet-preserving and backward join-preserving maps. We first check that
+this category satisfies the hypotheses of \thmref{fam-closed}:
+
+\begin{proposition}
+  $\MeetSLat \times \JoinSLat^\op$ has biproducts and all products.
+\end{proposition}
+
+\begin{proof}
+  $\MeetSLat$ and $\JoinSLat$ are both $\CMon$-enriched and have
+  finite products, as noted above. The opposite of a category with
+  biproducts also has biproducts (by swapping the injections $i$ and
+  projections $p$), and products of categories with biproducts also
+  have biproducts pointwise. Hence $\MeetSLat \times \JoinSLat^\op$
+  has biproducts.
+
+  $\MeetSLat$ has all products, indeed all limits, because it is the
+  category of algebras for a Lawvere theory. Similarly, $\JoinSLat$
+  has all coproducts, indeed all colimits, for the same reason. Note
+  that these are very different constructions: elements of a product
+  of meet-semi-lattices consist of (possibly infinite) tuples of
+  elements, while elements of a coproduct of join-semilattices consist
+  of \emph{finite} formal joins of elements quotiented by the
+  join-semi-lattice equations. Since $\JoinSLat$ has all coproducts,
+  $\JoinSLat^\op$ has all products, and so
+  $\MeetSLat \times \JoinSLat^\op$ has all products, as required.
+\end{proof}
+
+\begin{corollary}
+  $\Fam(\MeetSLat \times \JoinSLat^\op)$ is cartesian closed and has
+  all coproducts.
+\end{corollary}
+
+This corollary means that, assuming a sensible intepretation of
+primitive types and operations, we can use
+$\Fam(\MeetSLat \times \JoinSLat^\op)$ to interpret the higher-order
+language we describe in the next section. We still regard the category
+$\Fam(\LatGal)$ as the reference model of approximable sets with
+forward and backward approximation maps; the category
+$\Fam(\MeetSLat \times \JoinSLat^\op)$ is a technical device to carry
+out the interpretation of higher-order programs. To get
+interpretations of first-order types and primitive operations, we can
+embed $\Fam(\LatGal)$ into $\Fam(\MeetSLat \times \JoinSLat^\op)$:
+
+\begin{proposition}
+  The functor
+  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ defined
+  on objects as
+  $F(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
+  X(x)))$. This functor is faithful and preserves coproducts and
+  finite products.
+\end{proposition}
+
+With this embedding functor, we can see that the interpretation of
+first-order types will be the same up to isormorphism in
+$\Fam(\LatGal)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$, as long as
+we interpret the base types as objects in $\Fam(\LatGal)$. At
+higher-order, however, the meet-semi-lattice and join-semi-lattice
+sides of the interpretation will diverge, and it is no longer clear
+that the interpretation of programs using higher-order functions
+internally will result in Galois connections. In \secref{definability}
+we will see that every program with first-order type (even if it uses
+higher-order functions internally) does in fact have an interpretation
+definable in $\Fam(\LatGal)$.
+
+% $\GLR(\Fam(\MeetSLat \times \JoinSLat^\op), F)$ fibred over
+% $\Fam(\MeetSLat \times \JoinSLat^\op)$ with a full and faithful functor $\Fam(\LatGal) \to \GLR(\Fam(\MeetSLat \times \JoinSLat^\op), F)$.
+
+\bob{In the conclusion, we have to admit to the possibility that
+  $\Fam(CLatGal)$ might be sensible. We could have also used
+  PShCMon(LatGal) and there is PShCMon(GraphLang) for
+  normalisation. We would have different problems with proving
+  correctness however.}
+
+\subsection{Semantic \GPS in $\Fam(\LatGal)$}
+\label{sec:semantic-gps}
+
+The above discussion has been somewhat abstract, so we now consider
+some examples in the category $\Fam(\LatGal)$ and how they relate to
+\GPS. We highlight how
+
+\subsubsection{Lifting Monad}
+
+\bob{Move the discussion in \texttt{first-order.tex} to here, but
+  construct the lifting monad in a more elementary way}
+
+\subsubsection{An Approximation Object and the Tagged Data Monad}
+
+\bob{flesh this out}
+
+The object $\mathbb{A} = (1, \lambda \_.\, \{\top,\bot\})$ represents
+a basic object for measuring approximations in terms of presence and
+absence. (If all we care about is presence or absence in our
+approximations, we could see this object as the analogue of the real
+numbers for normal differentiation.)
+
+This object is a commutative monoid in $\Fam(\LatGal)$, and so we can
+construct the writer monad $T(X) = \mathbb{A} \times X$ in
+$\Fam(\LatGal)$.
+
+On objects $X$ with trivial approximations, we have $L(X) \cong T(X)$,
+but for more structured objects we get usage information at every level.
+
+This is a bit like using the writer monad and Moggi's CBN translation
+as a basic way of doing provenance tracking, but here we can track
+backwards and forwards.
+
+\subsubsection{Approximating Numbers by Intervals}
+
+\bob{Describe how \exref{intervals-and-maxima-elements} fits into
+  $\Fam(\LatGal)$}
+
+\subsection{Summary}
+
+We have constructed the category
+$\Fam(\MeetSLat \times \JoinSLat^\op)$, which is a cartesian closed
+category with all coproducts, ienough to interpret the total
+higher-order language we define in the next section with primitive
+types and operations defined in $\Fam(\LatGal)$. However, we are not
+guaranteed by construction that at first-order type, the
+interpretations are in fact Galois connections. We will rectify this
+in \secref{definability} using a logical relations construction.
+
+As we did in \secref{approx-as-tangents}, we end the section with a
+conjecture relating our categories to Tangent categories.
+
+\begin{conjecture}
+  (1) The category $\Fam(\MeetSLat)$ is a Tangent category, with the
+  tangent bundle
+  $T(X, \partial X) = (\Sigma_{x : X}. \partial X(x), \lambda (x,
+  \delta x).\, \downset{\delta x})$. (2) The category $\Fam(\LatGal)$
+  is a reverse Tangent category with the analogous definition of
+  tangent bundle.
+\end{conjecture}
+
+Comparing this conjecture to the previous one, we can see that the
+difference between $\CM$, $\Lposet$ and $\Fam(\MeetSLat)$,
+$\Fam(\LatGal)$ is that the latter have a stronger separation of
+points from tangents, somewhat analogous to the situation with
+manifolds. Also note that, contrary to the $\Fam(\FinVect)$
+representation of manifolds and differentiable maps, (if this
+conjecture holds) we do not throw away information about higher
+derivatives. It is retained in the order structure of the tangent
+fibres.

### 2025-07-08 main/approx-as-tangents.tex
[ tex ] talk about chain rule more

@@
 
-\bob{mention chain rule here? and tangent spaces?}
+\begin{remark}[Chain Rule]
+  \label{rem:chain-rule}
+  A useful property of derivative maps is that they compose according
+  to the chain rule. Suppose that $f : M \to N$ and $g : N \to K$ are
+  smooth functions. Then for any $x \in M$, we have:
+  \begin{itemize}
+  \item $\pushf{(g \circ f)}_x = \pushf{g}_{f(x)} \circ \pushf{f}_x : \tangents_x(M) \linearto \tangents_{g(f(x)}(K)$
+  \item $\pullf{(g \circ f)}_x = \pullf{f}_x \circ \pullf{g}_{f(x)} : \cotangents_{f(x)}(K) \linearto \cotangents_x(M)$
+  \end{itemize}
+  The chain rule has the practical effect that we can compute
+  derivative maps of $f$ and $g$ independently and compose them,
+  instead of the potentially more difficult task of computing the
+  derivative maps of $g \circ f$. As we shall see below, stable maps
+  also obey a chain rule, and this forms the basis of the general
+  categorical approach to differentiability that we describe in
+  \secref{models-of-total-gps}.
+\end{remark}
 
@@
 In a partially ordered set $X$, for any $x \in X$ the set of elements below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a partially ordered set.
-These approximations of $x$ we will think of as ``tangents'' at $x$, and the whole set $\downset{x}$ as the ``tangent space''\footnote{\bob{Is it possible to derive the tangent spaces by thinking about continuous functions passing through $x$?}}.
+These approximations of $x$ we will think of as ``tangents'' at $x$, and the whole set $\downset{x}$ as the ``tangent space''.
 Tangent spaces are vector spaces, so in the partially ordered setting we take elements of $\downset{x}$ to be approximations of processes defined at $x$.
@@
 
+% \footnote{\bob{Is it possible to derive the tangent spaces by
+% thinking about continuous functions passing through $x$?}}
+
 \begin{definition}
-  A \emph{bounded meet poset} is a partially ordered set $X$ where for every $x \in X$, $\downset{x}$ is a meet semilattice, with $x$ as the top element.
+  A \emph{bounded meet poset} is a partially ordered set $X$ where for
+  every $x \in X$, $\downset{x}$ is a meet semilattice, with $x$ as
+  the top element.
 \end{definition}
@@
   Bounded meet lattices and conditionally multiplicatve functions form
-  a category $\CM$. This category has products, coproducts and
+  a category $\CM$. This category has products, coproducts, and
   exponentials.
@@
 \begin{proof}
-  \AGDA. See \citet{amadio-curien} for the case when the posets are
-  also cpos. The crucial technical step, identified by
-  \citet{berry79}, is that the ordering on conditionally
+  \AGDA. See \citet{amadio-curien} (FIXME: theorem reference) for the
+  case when the posets are also cpos. The crucial technical step,
+  identified by \citet{berry79}, is that the ordering on conditionally
   multiplicative functions is not the extensional ordering
-  ($f \sqsubseteq g$ iff $\forall x. f(x) \subseteq g(x)$) but instead
-  the stable ordering $f \sqsubseteq g$ iff \bob{FIXME}.
+  ($f \sqsubseteq_{\mathrm{ext}} g$ iff
+  $\forall x. f(x) \sqsubseteq g(x)$) but instead the stable ordering:
+  $f \sqsubseteq_{\mathrm{st}} g$ iff $f \sqsubseteq_{ext} g$ and
+  $\forall x, x'.\,x \sqsubseteq x' \Rightarrow f(x) = f(x') \land
+  g(x)$.
 \end{proof}
+
+\begin{remark}[Chain Rule]
+  \label{rem:cm-chain-rule}
+  It is almost a triviality at this point, but the crucial point is
+  that, for any cm functions $f : X \to Y$ and $g : Y \to Z$, the
+  restriction maps (``forward derivatives'') compose according to the
+  chain rule from \remref{chain-rule}:
+  \begin{displaymath}
+    (g \circ f)_x = g_{f(x)} \circ f_x
+  \end{displaymath}
+  We will see this phenomenon repeated in the definition of stable
+  functions below.
+\end{remark}
 
@@
 
-Given these examples, conditional multiplicativity seems to be a reasonable analogue to functions with a well-defined notion of derivative.
-However, for \GPS we also require an analogue to the reverse derivative, where we map approximations backwards to give the least approximation of the input for a given approximation of the output.
-In the case of smooth functions, we are always guaranteed a reverse derivative but there is not always a best way to map approximations backwards for cm functions, as the following example shows.
+Given these examples, conditional multiplicativity seems to be a
+reasonable analogue to functions with a well-defined notion of
+derivative. For \GPS we also require an analogue to the reverse
+derivative, where we map approximations backwards to give the least
+approximation of the input for a given approximation of the output. In
+the case of smooth functions, we are always guaranteed a reverse
+derivative. However, there is not always a best way to map
+approximations backwards for cm functions, as the following example
+shows.
 
@@
   \label{ex:non-stable-function}
-  % In the preceeding examples, we demonstrated conditional
-  % multiplicativity by giving a backwards mapping of approximations
-  % that forces preservation of meets. A natural question is whether or
-  % not we always have such a backwards map which we could use for
-  % backwards propagation of approximations.
   An example that is conditionally multiplicative, but does not admit
@@
   before Lemma 12.2.3, originally due to Berry) is given by defining
-  $\mathrm{unstable} : D \to \{\bot, \top\}$, where
+  $\mathrm{unstable} : D \to \{\bot \sqsubseteq \top\}$, where
   $D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots
@@
 
-In light of the \exref{non-stable-function}, we consider
-\citet{berry79}'s definition of \emph{stable function} that requires
+In light of the \exref{non-stable-function}, we turn to
+\citet{berry78}'s definition of \emph{stable function} that requires
 the existence of a reverse mapping directly, even without assuming
@@
 \begin{theorem}
-  L-posets and stable functions form a category with products and
-  coproducts.
+  L-posets and stable functions form a category $\Lposet$ with products
+  and coproducts.
 \end{theorem}
+
+\begin{remark}[Chain Rule]
+  As for the ``forward derivatives'' of conditionally multiplicative
+  functions, the forward and backwards parts of a stable function
+  compose according to the chain rule (c.f. \remref{chain-rule}):
+  \begin{itemize}
+  \item $(g \circ f)_x = g_{f(x)} \circ f_x : \downset{x} \linearto \downset{g(f(x))}$
+  \item $(g \circ f)^*_x = f*_x \circ g^*_{f(x)} : \downset{g(f(x))} \linearto \downset{x}$
+  \end{itemize}
+  We will use this property in \propref{embed-stable} to show that
+  $\Lposet$ embeds into our category of sets-with-approximation.
+\end{remark}
 
@@
 downsets $\downset{x}$ are \emph{complete}
-lattices.
-\citet{amadio-curien} (Theorem 12.FIXME) details the proof.
-Intuitively, to generate the best approximation of an input value for a function, we need to take the infimum over all possible input values.
+lattices. \citet{amadio-curien} (Theorem 12.FIXME) details the
+proof. Intuitively, to generate the best approximation of an input
+value for a function, we need to take the infimum over all possible
+input values.
 
@@
 queries on databases as in \exref{introduction-example}, so why should
-we not just take complete L-posets and stable functions as our model?
-We have two reasons for moving to a slightly different model in \secref{model}:
+we not just take complete L-posets and stable functions as our model
+of \GPS?  We have two reasons for moving to a different model in
+\secref{models-of-total-gps}:
 
@@
   next section.
-\item A more local reason is that we wish to formalise our
+\item A more practical reason is that we wish to formalise our
   construction in the proof assistant Agda \cite{agda} in order to get
@@
 
-We have seen that \citet{berry79}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives.
+We have seen that \citet{berry78}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives.
 In the next section, we describe a model based on these ideas suitable for constructing executable models of total higher order languages.
@@
 we have not considered above. We will extend this conjecture to our
-refined model of partially ordered approximations in
-\secref{model-summary}.
+refined model of lattice approximated sets in \secref{model-summary}.
 
@@
 % \end{itemize}
-
-% \bob{Put a conjecture about Tangent Categories in somewhere}
-
-% Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.
-This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wide range of computational settings.

### 2025-07-08 main/models.tex
[ tex ] talk about chain rule more
[ tex ] refine the discussion of Fam(C) a bit

@@
-\section{Constructing Models of \GPS for a Total Language}
+\section{Models of \GPS for a Total Language}
 \label{sec:models-of-total-gps}
@@
 a model for total \GPS based on the \emph{Category of Families}
-construction. This construction, adn the more general Grothendieck
+construction. This construction, and the more general Grothendieck
 construction, has been previously used by V{\'a}k{\'a}r and
 collaborators \bob{CITE} to model automatic differentiation for
-higher-order programs on the reals. We rely on some of their results,
+higher-order programs on the reals. We reuse some of their results,
 and discuss the commonalities as we go.
@@
 
-As we explained in \secref{stab}, L-posets can be seen as partially
-ordered sets where every principal downset $\downset{x}$ is a bounded
-lattice. That is, for an element $x$, $\downset{x}$ is its lattice of
-approximations/tangents. The shortcoming of this setup is that proper
-values and their approximations live in the same set. We fix this by
-changing our model to one where we have sets $X$ of values, and for
-each $x \in X$, a bounded lattice $\partial X(x)$ of approximations of
-$x$. This construction is an instance of the general \emph{Category of
-  Families} construction:
+L-posets are partially ordered sets where every principal downset
+$\downset{x}$ is a bounded lattice of approximations/tangents. As we
+explained in \secref{stab}, the shortcoming of this setup is that
+proper values and their approximations live in the same set. We fix
+this by changing our model to one where we have sets $X$ of values,
+and for each $x \in X$, a bounded lattice $\partial X(x)$ of
+approximations of $x$. This construction is an instance of the general
+\emph{Category of Families} construction:
 
@@
   function $f : X \to Y$ and a familiy of morphisms of $\cat{C}$,
-  $\partial f : (x : X) \to \cat{C}(\partial X(x), \partial Y(f\,x))$.
+  $\partial f : \Pi_{x : X}.\,\cat{C}(\partial X(x), \partial Y(f\,x))$.
 \end{definition}
 
-Composition $f \circ g $ of morphisms
+The reason for choosing the $\Fam$ construction is that composition in
+this category is an abstract version of the chain rule that we have
+seen in \remref{chain-rule}, \remref{chain-rule-cm}, and
+\remref{chain-rule-stable}. Composition $f \circ g $ of morphisms
 $f : (Y, \partial Y) \to (Z, \partial Z)$ and
@@
 $\partial (f \circ g)(x) = \partial f(f, x) \circ \partial g(x)$,
-where the latter composition is in $\cat{C}$. The fact that
-composition of morphisms in $\Fam(\cat{C})$ resembles the chain rule
-from calculus is essential to its use in modelling automatic
-differentiation. Indeed, if we let $\FinVect$ be the category of
-finite dimensional real vector spaces, then:
+where the latter composition is in $\cat{C}$.
+
+The fact that morphisms in $\Fam(C)$ compose according to a chain rule
+means that the categories we have seen so far embed into $\Fam(C)$ for
+appropriate $C$. If we let $\FinVect$ be the category of finite
+dimensional real vector spaces, then:
 
@@
 
-For \GPS we replace the category $\FinVect$ with $\LatGal$:
+For the categories $\CM$ and $\Lposet$, we pick the appropriate
+categories of partial orders and monotone maps:
 
 \begin{definition}
-  The category $\LatGal$ has bounded lattices as objects and Galois
-  connections as morphisms, with the right adjunct going in the
-  ``forward'' direction.
+  \begin{enumerate}
+  \item The category $\LatGal$ has bounded lattices as objects and
+    Galois connections as morphisms, with the right adjunct going in
+    the ``forward'' direction.
+  \item The category $\MeetSLat$ has meet-semi-lattices with top as
+    objects and monotone finite meet preserving functions as
+    morphisms.
+  \item The category $\JoinSLat$ has join-semi-lattices with bottom as
+    objects and monotone finite join preserving functions as
+    morphisms.
+  \end{enumerate}
 \end{definition}
@@
 \begin{proposition}
+  \label{prop:embed-stable}
   There is a faithful functor $\Lposet \to \Fam(\LatGal)$ that maps an
   L-poset $X$ to $(X, \lambda x. \downset{x})$ and stable functions
-  $f$ to $(f, \lambda x. (f_x, f^*_x))$.
+  $f$ to $(f, \lambda x. (f_x, f^*_x))$. Likewise, there is a faithful
+  functor $\CM \to \Fam(\MeetSLat)$.
 \end{proposition}
@@
 \propref{embed-manifolds} because it is not clear that the category
-$\Lposet$ is in any way a canonical definition of approximable sets
-and functions with approximation derivatives. We could restrict
-$\Lposet$ to morphisms that preserve maximal elements and take
-L-posets $X$ to $(\mathrm{Max}(M), \lambda x. \downset{x})$. However,
-our current hypothesis is that $\Fam(\LatGal)$, where values and their
-approximations are separated by construction is the primary model of
-semantic \GPS. We now investigate the categorical properties of this
-category, with a view to modelling a higher-order total programming
-language.
+$\Lposet$ (or $\CM$) is a canonical definition of approximable sets
+and functions with approximation derivatives, as we discussed at the
+end of \secref{stab}.
+% We could restrict $\Lposet$ to morphisms that preserve maximal
+% elements and take L-posets $X$ to
+% $(\mathrm{Max}(M), \lambda x. \downset{x})$.
+Our current hypothesis is that $\Fam(\LatGal)$, where values and their
+approximations are separated by construction, is the natural model of
+semantic \GPS in a total setting. We now investigate some categorical
+properties of this category, with a view to modelling a higher-order
+total programming language in \secref{language}.
 
@@
 \begin{proposition}
-  For any $\cat{C}$, $\Fam(\cat{C}$ has all coproducts, which can be
+  For any $\cat{C}$, $\Fam(\cat{C})$ has all coproducts, which can be
   given on objects by:
@@
   \end{displaymath}
-  Since $\Fam(\cat{C})$ is extensive, products and coproducts always
+  Since $\Fam(\cat{C})$ is extensive, products and coproducts
   distribute.
@@
 reverse derivatives of a smooth map intrinsically defined,
-\citet{vakar2021}'s correctness theorem can show that
+\citet{vakar2021}'s correctness theorem shows that, for programs with
+first-order type, the interpretation in $\Fam(\Vect)$ correctly yields
+the forward derivative of the defined function on the reals (and
+reverse derivative for $\Fam(\Vect^\op)$.
 
@@
 Galois connections between them. From a classical mathematical point
-of view, this would give a model of \GPS that would likely be suitable
-for reasoning programs' behaviour and their forward and backward
+of view, this would give a model of \GPS that would be suitable for
+reasoning about programs' behaviour and their forward and backward
 approximations. However, in terms of building an executable model
 inside the Agda proof assistant, and with an eye toward implementation
-strategies, we seek a more down-to-earth (i.e., predicative and
-constructive) solution. We will do this by separating the forward and
-backward parts of the Galois connections to act independently by
-moving to the product category $\MeetSLat \times
-\JoinSLat^\op$. Objects in this category consist of \emph{separate}
-meet- and join-semi-lattices and potentially unrelated forward
-meet-preserving and backward join-preserving maps. We first check that
-this category satisfies the hypotheses of \thmref{fam-closed}:
+strategies, we seek a finitary solution. (Note that the solution of
+moving to complete lattices is very different to moving to arbitrary
+dimension vector spaces: in the former we have infinitary operations,
+while the latter still has only finitary operations.)
+
+We will avoid the need for infinitary operations by separating the
+forward and backward parts of the Galois connections to act
+independently by moving to the product category
+$\MeetSLat \times \JoinSLat^\op$. Objects in this category consist of
+\emph{separate} meet- and join-semi-lattices and potentially unrelated
+forward meet-preserving and backward join-preserving maps. We first
+check that this category satisfies the hypotheses of
+\thmref{fam-closed}:
 

### 2025-07-08 main/models.tex
[ tex ] failure of cartesian closure on biproduct categories

@@
 
+\begin{remark}
+  \label{rem:hermida-exponentials}
   There is another construction of internal homs on $\Fam(\cat{C})$
@@
   is fatally flawed in that neither of $\LatGal$ nor $\FinVect$ are
-cartesian closed, and there is no way of making them so without losing
-the property of having biproducts. We will use Hermida's construction
-in our definability proof in \secref{definability}, where we use a
-logical relations argument to show that every morphism definable in
-the higher order language is also first-order definable.
+  cartesian closed, and there is no way of making them so without
+  losing the property of having biproducts. We will use Hermida's
+  construction in our definability proof in \secref{definability},
+  where we use a logical relations argument to show that every
+  morphism definable in the higher order language is also first-order
+  definable.
+\end{remark}
 
@@
   \end{salign*}
+  A zero object is an object that is both initial and terminal.
 \end{definition}
@@
 coproducts:
-
 \begin{proposition}
@@
     $X \times Y = X + Y = X \oplus Y$.
-  \item A $\CMon$-category with (co)products also has biproducts.
+  \item A $\CMon$-category with (co)products also has biproducts, and
+    any initial or terminal object is a zero object.
   \end{enumerate}
@@
 \end{example}
+
+The following proposition shows why the alternative construction of
+exponentials in $\Fam(C)$ described in \remref{hermida-exponentials}
+is not suitable:
+\begin{proposition}
+  A $\CMon$-category that is cartesian closed has trivial homsets:
+  $\cat{C}(X,Y) \cong 1$.
+\end{proposition}
+
+\begin{proof}
+  $\cat{C}(X, Y) \cong \cat{C}(0 + X,Y) \cong \cat{C}(0 \times X,Y)
+  \cong \cat{C}(0,X \to Y) \cong 1$.
+\end{proof}
 

### 2025-07-08 main/approx-as-tangents.tex
[ tex ] fiddling

@@
   approximations for partiality and approximations for stability ought
-  to be the same thing. \bob{fwd reference to discussion.}) In
-  \exref{intervals-and-maxima-elements}, maximal elements could be
-  taken to be the ``proper values''. One idea is to restrict to
-  conditionally multiplicative or stable functions that preserve
-  maximal elements. However, this idea fails at higher order:
-  functions that take maximal elements to maximal elements are not
-  themselves maximal elements. We could devise a category of L-posets
-  with totality predicates (which would pick out maximal elements at
-  first-order) and totality preserving functions, but we prefer a more
-  direct method of separating values proper from their approximations
-  using the \emph{Category of Families} construction we detail in the
-  next section.
+  to be the same thing. We discuss this further in
+  \secref{conclusion}.) In \exref{intervals-and-maxima-elements},
+  maximal elements could be taken to be the ``proper values''. One
+  idea is to restrict to conditionally multiplicative or stable
+  functions that preserve maximal elements. However, this idea fails
+  at higher order: functions that take maximal elements to maximal
+  elements are not themselves maximal elements. We could devise a
+  category of L-posets with totality predicates (which would pick out
+  maximal elements at first-order) and totality preserving functions,
+  but we prefer a more direct method of separating values proper from
+  their approximations using the \emph{Category of Families}
+  construction we detail in the next section.
 \item A more practical reason is that we wish to formalise our
@@
 \end{enumerate}
-
-\bob{For future work: what is the connection between FamCPO(LatGal)
-  and Laird's bistable biorders, or Berry bidomains?}
-
-% We could have required that L-posets' principal downsets were complete lattices, which would mean that we would only need to require that forward approximation maps preserve all meets and we could compute the backwards map automatically.
-However, our goal is a computable theory for \GPS, analogous to automatic differentiation.
-Requiring completeness to compute backwards maps of approximations would require that complete joins are always computable, which will not be the case for approximations of functions.
-
-% \begin{proposition}
-%   L-posets and stable functions form a category. \bob{and the main point is that composition of stability witnesses looks like a chain rule!}
-% \end{proposition}
-
-% \bob{L-poset is has products and coproducts, but is not closed.
-To be closed would require each $\downset{x}$ to be a \emph{complete} lattice.
-However, this is tricky to implement as we need an infinitary operation to compute the reverse approximation map for function types.
-See \cite{amadio-curien} for the proof that bounded-complete lattices with stable functions form a cartesian closed category.}
 

### 2025-07-08 main/conclusion.tex
[ tex ] fiddling

@@
 explainability in systems that combine neural networks with traditional symbolic computation.
+
+\paragraph{More Approximation Domains.} The intervals example is
+interesting, but is it useful, and are there more examples like this?
+One route might be to follow \cite{edalat-heckmann98} and explore
+whether metric spaces (which already provide a native notion of
+approximation) can be embedded in $\Fam(\LatGal)$.
+
+\paragraph{Alternative Higher-order Models}
+
+1. $\Fam(\PSh_{\CMon}(\LatGal))$
+
+2. Concrete sheaves over $\Fam(\LatGal)$ analogous to Diffeological spaces?
+
+\paragraph{Categorical Models of Differentation}
+
+\paragraph{Recursion and Partiality.}
+1. Despite using stable domain theory as our starting point, we have ignored
+
+2.
+Conjecture that the category with $(X : Cpo, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and galois connections is the right setting.
+Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains (Laird).
+
+\paragraph{Implementation}
+
+Would be better to have an implementation that worked by translation,
+instead of interpreting directly as Agda functions. See work by
+Smeding and Vákár on translations. To formalise this, we'd need
+to... See also $\Fam(\PSh_{\CMon}(Syn))$?

### 2025-07-08 main/models.tex
[ tex ] fiddling

@@
 satisfy the hypotheses of this theorem, because neither of them have
-infinite products. We will consider ways to rectify this below
-\bob{fwd ref}.
+infinite products. We will consider ways to rectify this below in
+\secref{fixing-completeness}.
 
@@
 \subsubsection{Discrete Completeness}
+\label{sec:fixing-completeness}
 
@@
 
-\bob{In the conclusion, we have to admit to the possibility that
-  $\Fam(CLatGal)$ might be sensible. We could have also used
-  PShCMon(LatGal) and there is PShCMon(GraphLang) for
-  normalisation. We would have different problems with proving
-  correctness however.}
+% \bob{In the conclusion, we have to admit to the possibility that
+%   $\Fam(CLatGal)$ might be sensible. We could have also used
+%   PShCMon(LatGal) and there is PShCMon(GraphLang) for
+%   normalisation. We would have different problems with proving
+%   correctness however.}
 
@@
 difference between $\CM$, $\Lposet$ and $\Fam(\MeetSLat)$,
-$\Fam(\LatGal)$ is that the latter have a stronger separation of
-points from tangents, somewhat analogous to the situation with
-manifolds. Also note that, contrary to the $\Fam(\FinVect)$
-representation of manifolds and differentiable maps, (if this
-conjecture holds) we do not throw away information about higher
-derivatives. It is retained in the order structure of the tangent
-fibres.
+$\Fam(\LatGal)$ is that the latter have a separation of points from
+tangents, somewhat analogous to the situation with manifolds. Also
+note that, contrary to the $\Fam(\FinVect)$ representation of
+manifolds and differentiable maps, (if this conjecture holds) we do
+not throw away information about higher derivatives. It is retained in
+the order structure of the tangent fibres.

### 2025-07-08 main/conclusion.tex
Purging section 4.

--- 2025-07-02 main/conclusion.tex
@@
 
-\paragraph{Integrating XAI techniques with Galois slicing.} \todo{Sketch what this might look like.}
+\paragraph{Integrating XAI techniques with Galois slicing.} \roly{Sketch what this might look like.}
 Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping (Grad-CAM)~\cite{selvaraju20}

### 2025-07-08 main/introduction.tex
Citations on gradient descent.
Cite Dex, Sigal, etc.
Purging section 4.

@@
 
-Existing approaches to program slicing are often tied to particular programming languages or implementations. In this paper we develop a general categorical approach to program slicing and \todo{...}.
+Existing approaches to program slicing are often tied to particular programming languages or implementations.
+In this paper we develop a general categorical approach to program slicing and \roly{complete}.
 
@@
 Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information.
-We elaborate on this further in Section \todo{XYZ}.
+We elaborate on this further in \secref{first-order}.
 
@@
 Not only does this mean that the constructions and proofs have been checked, but also the construction is executable and can be run on small examples.
-We are also planning to use Agda's JavaScript backend to produce a demo.
-\todo{Do this?}
+We are also planning to use Agda's JavaScript backend to produce a demo.% \todo{Do this?}

### 2025-07-08 main/models.tex
Purging section 4.

--- 2025-07-08 main/models.tex
@@
-\section{Models of \GPS for a Total Language}
+\section{Constructing Models of \GPS for a Total Language}
 \label{sec:models-of-total-gps}
@@
 a model for total \GPS based on the \emph{Category of Families}
-construction. This construction, and the more general Grothendieck
+construction. This construction, adn the more general Grothendieck
 construction, has been previously used by V{\'a}k{\'a}r and
 collaborators \bob{CITE} to model automatic differentiation for
-higher-order programs on the reals. We reuse some of their results,
+higher-order programs on the reals. We rely on some of their results,
 and discuss the commonalities as we go.
@@
 
-L-posets are partially ordered sets where every principal downset
-$\downset{x}$ is a bounded lattice of approximations/tangents. As we
-explained in \secref{stab}, the shortcoming of this setup is that
-proper values and their approximations live in the same set. We fix
-this by changing our model to one where we have sets $X$ of values,
-and for each $x \in X$, a bounded lattice $\partial X(x)$ of
-approximations of $x$. This construction is an instance of the general
-\emph{Category of Families} construction:
+As we explained in \secref{stab}, L-posets can be seen as partially
+ordered sets where every principal downset $\downset{x}$ is a bounded
+lattice. That is, for an element $x$, $\downset{x}$ is its lattice of
+approximations/tangents. The shortcoming of this setup is that proper
+values and their approximations live in the same set. We fix this by
+changing our model to one where we have sets $X$ of values, and for
+each $x \in X$, a bounded lattice $\partial X(x)$ of approximations of
+$x$. This construction is an instance of the general \emph{Category of
+  Families} construction:
 
@@
   function $f : X \to Y$ and a familiy of morphisms of $\cat{C}$,
-  $\partial f : \Pi_{x : X}.\,\cat{C}(\partial X(x), \partial Y(f\,x))$.
+  $\partial f : (x : X) \to \cat{C}(\partial X(x), \partial Y(f\,x))$.
 \end{definition}
 
-The reason for choosing the $\Fam$ construction is that composition in
-this category is an abstract version of the chain rule that we have
-seen in \remref{chain-rule}, \remref{chain-rule-cm}, and
-\remref{chain-rule-stable}. Composition $f \circ g $ of morphisms
+Composition $f \circ g $ of morphisms
 $f : (Y, \partial Y) \to (Z, \partial Z)$ and
@@
 $\partial (f \circ g)(x) = \partial f(f, x) \circ \partial g(x)$,
-where the latter composition is in $\cat{C}$.
-
-The fact that morphisms in $\Fam(C)$ compose according to a chain rule
-means that the categories we have seen so far embed into $\Fam(C)$ for
-appropriate $C$. If we let $\FinVect$ be the category of finite
-dimensional real vector spaces, then:
+where the latter composition is in $\cat{C}$. The fact that
+composition of morphisms in $\Fam(\cat{C})$ resembles the chain rule
+from calculus is essential to its use in modelling automatic
+differentiation. Indeed, if we let $\FinVect$ be the category of
+finite dimensional real vector spaces, then:
 
@@
 
-For the categories $\CM$ and $\Lposet$, we pick the appropriate
-categories of partial orders and monotone maps:
+For \GPS we replace the category $\FinVect$ with $\LatGal$:
 
 \begin{definition}
-  \begin{enumerate}
-  \item The category $\LatGal$ has bounded lattices as objects and
-    Galois connections as morphisms, with the right adjunct going in
-    the ``forward'' direction.
-  \item The category $\MeetSLat$ has meet-semi-lattices with top as
-    objects and monotone finite meet preserving functions as
-    morphisms.
-  \item The category $\JoinSLat$ has join-semi-lattices with bottom as
-    objects and monotone finite join preserving functions as
-    morphisms.
-  \end{enumerate}
+  The category $\LatGal$ has bounded lattices as objects and Galois
+  connections as morphisms, with the right adjunct going in the
+  ``forward'' direction.
 \end{definition}
@@
 \begin{proposition}
-  \label{prop:embed-stable}
   There is a faithful functor $\Lposet \to \Fam(\LatGal)$ that maps an
   L-poset $X$ to $(X, \lambda x. \downset{x})$ and stable functions
-  $f$ to $(f, \lambda x. (f_x, f^*_x))$. Likewise, there is a faithful
-  functor $\CM \to \Fam(\MeetSLat)$.
+  $f$ to $(f, \lambda x. (f_x, f^*_x))$.
 \end{proposition}
@@
 \propref{embed-manifolds} because it is not clear that the category
-$\Lposet$ (or $\CM$) is a canonical definition of approximable sets
-and functions with approximation derivatives, as we discussed at the
-end of \secref{stab}.
-% We could restrict $\Lposet$ to morphisms that preserve maximal
-% elements and take L-posets $X$ to
-% $(\mathrm{Max}(M), \lambda x. \downset{x})$.
-Our current hypothesis is that $\Fam(\LatGal)$, where values and their
-approximations are separated by construction, is the natural model of
-semantic \GPS in a total setting. We now investigate some categorical
-properties of this category, with a view to modelling a higher-order
-total programming language in \secref{language}.
+$\Lposet$ is in any way a canonical definition of approximable sets
+and functions with approximation derivatives. We could restrict
+$\Lposet$ to morphisms that preserve maximal elements and take
+L-posets $X$ to $(\mathrm{Max}(M), \lambda x. \downset{x})$. However,
+our current hypothesis is that $\Fam(\LatGal)$, where values and their
+approximations are separated by construction is the primary model of
+semantic \GPS. We now investigate the categorical properties of this
+category, with a view to modelling a higher-order total programming
+language.
 
@@
 \begin{proposition}
-  For any $\cat{C}$, $\Fam(\cat{C})$ has all coproducts, which can be
+  For any $\cat{C}$, $\Fam(\cat{C}$ has all coproducts, which can be
   given on objects by:
@@
   \end{displaymath}
-  Since $\Fam(\cat{C})$ is extensive, products and coproducts
+  Since $\Fam(\cat{C})$ is extensive, products and coproducts always
   distribute.
@@
 \begin{proposition}
+  \item
   \begin{enumerate}
@@
 reverse derivatives of a smooth map intrinsically defined,
-\citet{vakar2021}'s correctness theorem shows that, for programs with
-first-order type, the interpretation in $\Fam(\Vect)$ correctly yields
-the forward derivative of the defined function on the reals (and
-reverse derivative for $\Fam(\Vect^\op)$.
+\citet{vakar2021}'s correctness theorem can show that
 
@@
 Galois connections between them. From a classical mathematical point
-of view, this would give a model of \GPS that would be suitable for
-reasoning about programs' behaviour and their forward and backward
+of view, this would give a model of \GPS that would likely be suitable
+for reasoning programs' behaviour and their forward and backward
 approximations. However, in terms of building an executable model
 inside the Agda proof assistant, and with an eye toward implementation
-strategies, we seek a finitary solution. (Note that the solution of
-moving to complete lattices is very different to moving to arbitrary
-dimension vector spaces: in the former we have infinitary operations,
-while the latter still has only finitary operations.)
-
-We will avoid the need for infinitary operations by separating the
-forward and backward parts of the Galois connections to act
-independently by moving to the product category
-$\MeetSLat \times \JoinSLat^\op$. Objects in this category consist of
-\emph{separate} meet- and join-semi-lattices and potentially unrelated
-forward meet-preserving and backward join-preserving maps. We first
-check that this category satisfies the hypotheses of
-\thmref{fam-closed}:
+strategies, we seek a more down-to-earth (i.e., predicative and
+constructive) solution. We will do this by separating the forward and
+backward parts of the Galois connections to act independently by
+moving to the product category $\MeetSLat \times
+\JoinSLat^\op$. Objects in this category consist of \emph{separate}
+meet- and join-semi-lattices and potentially unrelated forward
+meet-preserving and backward join-preserving maps. We first check that
+this category satisfies the hypotheses of \thmref{fam-closed}:
 

### 2025-07-08 notes/auto-diff.tex
Purging section 4.

@@
 \noindent and again define a corresponding category $\Diff^*$ of such maps where composition of the cotangent
-maps respects the backward chain rule. \todo{In practice can't combine $\Diff$ and $\Diff^*$ into a single
+maps respects the backward chain rule. \roly{In practice can't combine $\Diff$ and $\Diff^*$ into a single
 category with morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n) \times (\RR^n \linearto \RR^m)$

### 2025-07-08 notes/cmon-enriched.tex
Purging section 4.

--- 2025-05-19 notes/cmon-enriched.tex
@@
 
-\todo{Generalise to any monoidal closed $\mathscr{V}$ with biproducts?} The $\CMon$-enriched Yoneda embedding
-preserves biproducts, where (as in any functor category) biproducts in $\PSh_{\CMon}(\cat{C})$ are computed
-pointwise, and noting that $\CMon$ itself has biproducts.
+\roly{Generalise to any monoidal closed $\mathscr{V}$ with biproducts?} $\CMon$ is monoidal closed, and has
+biproducts (\secref{categories-with-biproducts:cmon}) which are inherited pointwise into
+$\PSh_{\CMon}(\cat{C})$ and preserved by the $\CMon$-enriched Yoneda embedding.
 

### 2025-07-08 notes/preliminaries.tex
Purging section 4.

@@
 
-\paragraph{Coproduct and product of morphisms.} In a category with coproducts $X + Y$ or products $X \times
-Y$, we write:
+\paragraph{Coproduct and product of morphisms.} In a category with coproducts $X \coprod Y$ or products $X
+\times Y$, we write:
 \begin{itemize}
@@
 can be formulated elementwise (as per the $\Set$-enriched case) rather than more generally in terms of
-morphisms from the tensor unit $I$. \todo{Be explicit about composition being a family of morphisms
+morphisms from the tensor unit $I$. \roly{Be explicit about composition being a family of morphisms
 $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\mathscr{V}$.}
@@
 $\eta_X: F(X) \to G(X)$ in $\cat{D}$ satisfying the usual naturality square for every $f: X \to Y$ in
-$\cat{C}$. \todo{This is sufficient because $\comp_{X,Y,Z}$ as well as $F$ and $G$ respects the
+$\cat{C}$. \roly{This is sufficient because $\comp_{X,Y,Z}$ as well as $F$ and $G$ respects the
 $\mathscr{V}$-structure?}

### 2025-07-08 main/introduction.tex
Anonymise reference to GitHub repo.

@@
 
-We have \href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}.
+We have \anonymise{\href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}}{formalised this work in Agda}.
 Not only does this mean that the constructions and proofs have been checked, but also the construction is executable and can be run on small examples.
-We are also planning to use Agda's JavaScript backend to produce a demo.% \todo{Do this?}
+We are also planning to use Agda's JavaScript backend to produce a demo.

### 2025-07-08 main/first-order.tex
Pare back section 4 just to be on first-order Galois slicing in Fam(LatGal).

--- 2025-07-03 main/first-order.tex
@@
 
-Our denotational approach to \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
-Before we go into details, we present a high-level overview of their approach and how we have applied it to \GPS.
-
-The fundamental approach in CHAD is to consider Grothendieck constructions on indexed categories $T : \cat{C}^\op \to \Cat$.
-An object of the Grothendieck construction $\int T$, described in more detail in \secref{Grothendieck}, is a pair $(X \in \cat{C}, \partial X \in T(X))$, which we read as pairing a space $X$ of ``points'' from $\cat{C}$ with an associated bundle of tangent spaces.
-The maps from the ``base'' category $\cat{C}$ are used to interpret the programs we are modelling.
-The maps in the indexed category are the maps of tangents or approximations, being derivatives and Galois connections respectively.
-
-In the case of differentiable programs and automatic differentiation, a basic model can be constructed by taking $\cat{C}$ to be $\Set$ and $T(X) = X \to \FinVect$: indexed collections of finite dimensional vector spaces, whose elements are interpreted as tangents at the given point, with linear maps.
-The Grothendieck category $\int T$ has objects that are pairs of a set $X$ and for each $x \in X$ a tangent vector space $\partial X(x)$.
-Morphisms are pairs of maps of points to points, accompanied by linear maps of tangents at those points.
-If we carefully choose an initial set of maps, we can read these maps as derivatives (note that nothing in the Grothendieck construction says that they must be derivatives!).
-Composition in the Grothendieck category is exactly the {\em chain rule} for composing derivatives, so we at least do know that if we start with derivatives on basic operations, then composing them will retain this property.
-
-For \GPS, we again take $\cat{C}$ to be $\Set$ and now take $T(X) = X \to \LatGal$: indexed collections of lattices, whose elements are interpreted as approximations at the given point, with indexed Galois connections as the maps between them.
-
-In these special cases when $\cat{C} = \Set$, the Grothendieck construction is the {\em families} constructions $\Fam(\FinVect)$ and $\Fam(\LatGal)$~(\secref{Fam}).
-These settings are adequate for interpreting first-order programs, but neither is Cartesian closed.
-The picture for higher-order programs is more complicated.
-As is well known, $\Fam(\cat{X})$ for any category $\cat{X}$ is the free coproduct completion of $\cat{X}$~\cite{lawvere63}.
-In the case that $\cat{X}$ has a (symmetric) monoidal product, then we always get a symmetric monoidal product on $\Fam(\cat{X})$, using the Cartesian products from $\Set$.
-When $\cat{X}$ has all (small) products, and the monoidal product is actually a coproduct, then $\Fam(\cat{X})$ is symmetric monoidal closed.
-If the coproducts are actually {\em biproducts}~(\secref{biproducts}), then $\Fam(\cat{X})$ is Cartesian closed, and we can interpret higher-order programs.
-
-Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of their having products and being enriched in commutative monoids~(\propref{biproducts:from-product-or-coproduct} below).
-However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$).
-Neither of these is ideal.
-In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost.
-In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model.
-Moreover, in Agda, there are issues with predicativity when considering complete lattices.
-\todo{be more specific}
-
-Vákár and collaborators overcome this by using different interpretations for forward and backward derivatives.
-We overcome this by using $\CMon$-enriched presheaves, as described in \secref{higher-order}; first we present a suitable first-order setting.
-
-\subsection{$\CMon$-Categories and Biproducts}
-\label{sec:biproducts}
-
-\subsubsection{$\CMon$-Categories}
-
-A category $\cat{C}$ is enriched in $\CMon$, the (monoidal) category of commutative monoids, if every hom-set
-$\cat{C}(X,Y)$ provides an operation + to add morphisms together, as well as a zero morphism $0: X \to Y$, and
-composition is \emph{bilinear}, i.e.~given by a family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor
-\Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that preserve the additive structure in
-$\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:
-
-\begin{salign*}
-f \comp \zero = f = \zero \comp f
-\end{salign*}
-\begin{salign*}
-(f + g) \comp h &= (f \comp h) + (g \comp h) \\
-h \comp (f + g) &= (h \comp f) + (h \comp g)
-\end{salign*}
-
-In a $\CMon$-category, any products and coproducts coincide.
-
-\begin{proposition}
-\label{prop:biproducts:from-product-or-coproduct}
-In any $\CMon$-category:
-\begin{enumerate}
-\item If $X \times Y$ is a product then it is also coproduct, with $\inj_1 = \prodM{\id}{0}$ and $\inj_2 =
-\prodM{0}{\id}$.
-\item If $X \coprod Y$ is a coproduct then it is also a product, with $\pi_1 = \coprodM{\id}{0}$ and $\pi_2 =
-\coprodM{0}{\id}$.
-\item $X$ is terminal iff it is initial.
-\end{enumerate}
-\end{proposition}
-
-An object that is both terminal and initial is called a \emph{zero} object; an object that is both a product
-and a coproduct is called a \emph{biproduct}.
-
-\subsubsection{Biproducts}
-
-More directly, a biproduct is as an object $X \biprod Y$ together with morphisms
-
-\begin{center}
-\begin{tikzcd}
-   X \arrow[r, "\biinj_X", shift left] &
-   X \biprod Y \arrow[l, "\biproj_X", shift left] \arrow[r, "\biproj_Y"', shift right] &
-   Y \arrow[l, "\biinj_Y"', shift right]
-\end{tikzcd}
-\end{center}
-
-\noindent satisfying
-
-\vspace{-3mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_X \comp \biinj_X &= \id_X \\
-   \biproj_Y \comp \biinj_X &= \zero_{X,Y}
-\end{salign*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{salign*}
-   \biproj_Y \comp \biinj_Y &= \id_Y \\
-   \biproj_X \comp \biinj_Y &= \zero_{Y,X}
-\end{salign*}
-\end{center}
-\end{minipage}
-
-\begin{salign*}
-(\biinj_X \comp \biproj_X) + (\biinj_Y \comp \biproj_Y) &= \id_{X \biprod Y}
-\end{salign*}
-
-\noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
-with finite biproducts is called \emph{semi-additive}. A semi-additive category cannot be Cartesian closed
-without being trivial: since we have $X \iso 0 \biprod X$, Cartesian closure would imply $\Hom{\cat{C}}{X}{Y}
-\iso \Hom{\cat{C}}{0}{X \Rightarrow Y} \iso 1$, and therefore in such a category all objects are isomorphic.
-
-% in any such category the canonical projections and injections give an isomorphism $\textstyle X \coprod Y
-% \iso X \times Y$ that is natural in both variables
-%
-% A semi-additive category $\cat{C}$ is enriched in $\CMon$: for any two morphisms $f, g: X \to Y$ in
-% $\cat{C}$, the biproduct structure provides a way to ``add'' them together, forming a morphism $f + g: X \to
-% Y$. Diagrammatically:
-%
-% \begin{center} \begin{tikzcd} X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y
-% \arrow[r, "\codiag"] & Y \end{tikzcd} \end{center}
-%
-% Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the product and
-% $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the coproduct.
-% The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \biproj_X}{g \comp \biproj_Y} =
-% \coprodM{\biinj_X \comp f}{\biinj_Y \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
-% composing the unique maps in and out of the zero object:
-%
-% \begin{center} \begin{tikzcd} X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y \end{tikzcd} \end{center}
-%
-% It is easy to verify that $+$ is associative and commutative and that $f + 0_{X,Y} = f$, and thus that every
-% hom-object $\cat{C}(X,Y)$ is an object in $\CMon$.
-%
-% \subsubsection{Biproduct laws}
-%
-% It is also possible to use these laws to define biproducts, in the case where $\cat{C}$ is enriched in
-% $\CMon$ and so equipped with addition of morphisms and zero morphisms. In that case any products or
-% coproducts in $\cat{C}$ are necessarily biproducts.
-
-\subsubsection{Example: Category of Finite Vector Spaces}
-\label{sec:categories-with-biproducts:fdvect}
-
-The category $\FinVect$ of finite vector spaces over $\RR$ is semi-additive.
-
-\begin{definition}
-Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces
-$(\RR^n, +, \mult)$, and as morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) +
-f(v)$ and $f(a \mult v) = a \mult f(v)$.
-% with $f$ preserving the zero and additive inverse structure automatically.
-\end{definition}
-
-%  with addition of real numbers written $a + b$ and multiplication written $a \mult b$
-
-% \begin{definition}[Finite dimensional vector space over $\RR$]
-% Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
-% component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
-% is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
-% vector operations are compatible with the field operations in that the following equations hold:
-%
-% \vspace{-4mm}
-% \begin{minipage}[t]{0.45\textwidth}
-% \begin{center}
-% \begin{align*}
-%    1 \mult v &= v \\
-%    (a \mult b) \mult v &= a \mult (b \mult v)
-% \end{align*}
-% \end{center}
-% \end{minipage}%
-% \begin{minipage}[t]{0.45\textwidth}
-% \begin{center}
-% \begin{align*}
-%    (a + b) \mult v &= (a \mult v) + (b \mult v) \\
-%    a \mult (u + v) &= (a \mult u) + (b \mult v)
-% \end{align*}
-% \end{center}
-% \end{minipage}
-% \end{definition}
-
-$\FinVect$ is enriched in $\CMon$ (in fact in $\Ab$, the category of abelian groups, but $\CMon$ will do for
-our purposes). For any objects $V = \RR^m$ and $W = \RR^n$, the hom-set $\FinVect(V,W)$ provides pointwise
-addition of morphisms $(f + g)(v) = f(v) + g(v)$; the constant map $v \mapsto 0$ is a zero morphism, and
-composition is bilinear. Moreover the direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian product
-is a biproduct, with $\biinj_{V} = \prodM{\id}{0}$ and $\biinj_{W} = \prodM{0}{\id}$.
-
-\subsubsection{Example: Category of Lattices and Galois Connections}
-
-Galois connections (\defref{galois-connection} above) are closed under component-wise composition and
-identities. The category $\LatGal$ of bounded lattices and Galois connections is semi-additive.
-
-\begin{definition}
-Define $\LatGal$ to be the category which has as objects $X = (X, \meet, \join, \top, \bot)$ all bounded
-lattices, and as morphisms all Galois connections $f \adj g: X \to Y$.
-\end{definition}
-
-\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
-Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
-the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps are enriched in $\CMon$ (in fact
-in $\SemiLat$, but again $\CMon$ is sufficient): addition of maps is computed pointwise via $\meet$, since $(f
-\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero.
-Dually the join-preserving maps can be added pointwise using $\join$, with the constant map $\bot = y \mapsto
-\bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is enriched in $\CMon$. The
-hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f \adj g) + (f' \adj g') =
-(f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is a zero morphism, and composition is
-bilinear.
-
-$\LatGal$ also has biproducts. The projections $\pi_1$ and $\pi_2$ from the product lattice $X \times Y$ have
-both upper and lower adjoints; this means that $X \times Y$ (which we shall write as $X \biprod Y$), is both a
-product and a coproduct, with projections $\biproj_X$ and $\biproj_Y$ and injections $\biinj_X$ and $\biinj_Y$
-given by:
-
-\vspace{-4mm}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   \biproj_X = \prodM{\id}{\bot} \adj \proj_1: X \biprod Y \to X \\
-   \biproj_Y = \prodM{\bot}{\id} \adj \proj_2: X \biprod Y \to Y
-\end{align*}
-\end{center}
-\end{minipage}%
-\begin{minipage}[t]{0.45\textwidth}
-\begin{center}
-\begin{align*}
-   \biinj_X = \pi_1 \adj \prodM{\id}{\top} : X \to X \biprod Y \\
-   \biinj_Y = \pi_2 \adj \prodM{\top}{\id}: Y \to X \biprod Y
-\end{align*}
-\end{center}
-\end{minipage}
-\vspace{2mm}
-
-\noindent with the trivial 1-point lattice as a zero object (terminal and initial).
-
-\subsection{Grothendieck Construction for Indexed Categories}
-\label{sec:Grothendieck}
-
-We now consider how to adapt the CHAD framework~\cite{vákár22,nunes2023} to interpret programs as functions
-along with their forward and backward approximation maps. Since the appropriate lattice of approximation
-typically depends on the value being approximated, we will want to interpret a type $\tau$ as a pair $(X,
-\partial X)$ where $X$ provides the conventional interpretation as a set and $\partial X$ provides an
-appropriate lattice of approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type
-$\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
-conventional semantics) must then be interpreted by a pair $(f, \partial f)$, with $f$ providing the
-conventional interpretation as a function $X \to Y$, and $\partial f$ as a family of Galois connections
-$\partial f(x): \partial X(x) \to \partial Y(f(x))$ taking approximations of $x$ to approximations of $f(x)$.
-
-This pattern of objects and morphisms is captured by the construction used in CHAD to model automatic
-differentiation, namely the \emph{Grothendieck construction} for a particular set-indexed category $T:
-\Set^\op \to \Cat$. For automatic differentiation, $T(X)$ will be the functor category $\Func{X}{\FinVect}$,
-regarded as the category of $X$-indexed families of finite vector spaces, with families of linear maps as
-morphisms. For Galois slicing, we will take $T(X) = \Func{X}{\LatGal}$, the category of $X$-indexed families
-of bounded lattices, with families of Galois connections as morphisms.
-
-In its general form, the Grothendieck construction $\Grothendieck{}T$ for an arbitrary indexed category $T:
-\cat{D}^\op \to \Cat$, sometimes called the \emph{total category} for $T$, incorporates all the categories
-$T(X)$ (for objects $X$ in $\cat{D}$) into a single category, together with morphisms that account for both
-internal structure and reindexing along morphisms $f: X \to Y$ in $\cat{D}$.
-
-\begin{definition}
-\label{def:Grothendieck}
-Suppose $T: \cat{D}^\op \to \Cat$ an indexed category. The \emph{Grothendieck construction}
-$\Grothendieck{X}T(X)$ (also written $\Grothendieck{}T$) is the category which has as objects, all pairs $(X,
-\partial X)$ of an object $X$ of $\cat{D}$ and an object $\partial X$ of $T(X)$, and as morphisms $(X,
-\partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to Y$ in $\cat{D}$ and
-morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
-\end{definition}
-
-\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
-construction and the idea of tangent spaces and derivatives. For any $(f, \partial f): (X, \partial X) \to (Y,
-\partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$, the composition
-$(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is given by $(g \comp f,
-T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to reindex $\partial g$ along $f$,
-yielding a morphism $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial
-Z)$.
-
-\subsubsection{Category of Families}
-\label{sec:Fam}
-
-For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{D} = \Set$ and
-the indexed category $T$ sends every set $X$ to the (functor) category $\Func{X}{\cat{C}}$ which has as
-objects all $X$-indexed families of objects of a fixed category $\cat{C}$, and as morphisms $F \to G$, all
-families of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only
-identity morphisms).
-
-% Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.
-%
-% \begin{proposition}
-% If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
-% pointwise.
-% \end{proposition}
-
-\begin{definition}[Reindexing]
-For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f} = \Func{f}{\cat{C}}:
-\Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the
-$X$-indexed family $\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete
-categories) and any family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to
-\reindex{G}{f}$ with $\reindex{\eta}{f}(x) = \eta(f(x))$.
-\end{definition}
-
-\noindent Although reindexing is just precomposition with $f$, writing it as $\reindex{-}{f}$ avoids
-notational confusion when combining composition of morphisms in $\Func{X}{C}$ with reindexing.
-
-The Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed category which sends
-any set $X$ to the category of indexed families $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to the
-reindexing functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$, is called the \emph{families}
-construction for $\cat{C}$.
-
-\begin{definition}[Categories of families]
-\label{def:Fam}
-For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
-$\Grothendieck{X}\Func{X}{\cat{C}}$.
-\end{definition}
-
-\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all pairs of sets $X$ and indexed families
-$\partial X: X \to \cat{C}$ and the morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are all
-pairs of functions $f: X \to Y$ and morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
-$\Func{X}{\cat{C}}$.
-
-\begin{proposition}
-\label{prop:Grothendieck:fam-inherits-local-smallness}
-If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
-\end{proposition}
-
-\begin{proposition}
-\label{prop:Grothendieck:fam-inherits-products}
-If $\cat{C}$ is (symmetric) monoidal then $\Fam(\cat{C})$ is also (symmetric) monoidal, with tensor product
-$(X, \partial X) \tensor (Y, \partial Y) = (X \times Y,(x, y) \mapsto \partial X(x) \tensor \partial Y(y))$.
-\end{proposition}
+\todo{Salvage some of this for \secref{semantic-gps}}
 
@@
 
-A programmer, or perhaps a language implementor, might want to embed additional ``approximation points'' into
+A programmer (or perhaps language implementor) might want to embed additional ``approximation points'' into
 composite data types, to allow for richer approximation structures. For example, for a list like $1 : 2 : []$
@@
 $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to families.
-
-\subsection{First-Order Automatic Differentiation via CHAD}
-\label{sec:first-order:autodiff}
-
-As per the CHAD work, automatic differentiation (AD) for first-order programs can also be recovered in this
-setup, using the total categories $\Fam(\FinVect)$ and $\Fam(\FinVect^\op)$ instead of $\Fam(\LatGal)$. First
-we recall the chain rule for forward derivatives $\pushf{f}_x: \RR^m \linearto \RR^n$ and backward derivatives
-$\pullf{f}_x: \RR^n \linearto \RR^m$, which says that derivatives respect composition~\cite{spivak65}.
-
-% \subsubsection{Derivatives of Differentiable Functions}
-%
-% For flat manifolds like $\RR^n$, the tangent space and cotangent space at a point $x$ are both canonically
-% isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
-% \RR^n$ at a point $x$ have the following form \todo{Can probably delete now as covered in
-% \secref{approx-as-tangents}}:
-%
-% \begin{itemize}
-% \item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ of $f$ at $x \in \RR^m$ is the
-% unique linear map $\RR^m \linearto \RR^n$ given by the Jacobean matrix of $f$ at $x$.
-% \item The \emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$ is the unique linear map
-% $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of the Jacobean matrix of $f$ at $x$.
-% \end{itemize}
-
-\begin{proposition}[Chain Rule]
-Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$ differentiable. Then for any $x \in \RR^m$ we have:
-
-\begin{itemize}
-\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
-\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
-\end{itemize}
-\end{proposition}
-
-\subsubsection{Automatic Differentiation}
-
-The basic idea of (forward-mode) AD, as first described by \citet{linnainmaa76}, was to recognise that one
-could efficiently compute the pushforward map $\pushf{f}: \RR^m \to (\RR^m \linearto \RR^n)$ of a
-differentiable function $f: \RR^m \to \RR^n$ alongside the computation of $f$ itself, by defining a single map
-$\tangents(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $g_1$
-for $\pi_1 \comp g$ and $g_2$ for $\pi_2 \comp g$, composition of such maps can be expressed as:
-\begin{align*}
-(g \comp h)_1(x) &= g_1(h_1(x)) \\
-(g \comp h)_2(x) &= g_2(h_1(x)) \comp h_2(x)
-\end{align*}
-
-We can verify that $\tangents$ is functorial, regardless of whether the linear maps happen to be derivatives.
-But if $h_2(x)$ is the derivative of $h_1$ at $x$ for any $x \in \RR^m$ and similarly for $g$ then the
-derivatives compose according to the forward chain rule and $(g \comp h)_2(x)$ is the derivative of $(g \comp
-h)_1$ at $x$. Thus this setup provides an algorithmic method for computing the forward derivative of a
-composite function, as long as all atomic operations provide derivatives, without have to recompute the
-function itself.
-
-% An efficient implementation of \GPS based on $\Fam(\LatGal)$
-% (\secref{fam:galois-slicing} above) would most likely apply a similar optimisation to make it possible to
-% compute forwards slices without having to reexecute the program at the same time.
-
-We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
-vector space $\RR^n$ as paired with the family of tangent spaces which is constantly $\RR^n$, and each map $g:
-\RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as a pair of maps $(g_1, g_2)$. Then the composition $(g,
-\partial g) \comp (h, \partial h) = (g \comp h, \reindex{\partial g}{h} \comp \partial h)$ again corresponds
-to the forward chain rule, since $(\reindex{\partial g}{h} \comp \partial h)_x = {\partial g}_{h(x)} \comp
-{\partial h}_x$. Analogously, reverse-mode maps $g: \RR^m \to \RR^n \times (\RR^n \linearto \RR^m)$ can be
-interpreted as living in $\Fam(\FinVect^\op)$ by reading them as pair of maps $(g_1, g_2)$ with composition
-being precisely the backwards chain rule.

### 2025-07-08 main/approx-as-tangents.tex
abramsky-jung.
Missing \mathrm.
Citations.

@@
 From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps.
-\bob{and we did it in Agda, so we can run it}
 
@@
 intuitive idea that a function that is computable cannot look at a
-non-finite amount of input to produce a finite output. \cite{abramsky-jung} is a comprehensive introduction to domain theory.
+non-finite amount of input to produce a finite output. \citet{abramsky-jung} provide a comprehensive introduction to domain theory.
 
@@
   $\forall x. f(x) \sqsubseteq g(x)$) but instead the stable ordering:
-  $f \sqsubseteq_{\mathrm{st}} g$ iff $f \sqsubseteq_{ext} g$ and
+  $f \sqsubseteq_{\mathrm{st}} g$ iff $f \sqsubseteq_{\mathrm{ext}} g$ and
   $\forall x, x'.\,x \sqsubseteq x' \Rightarrow f(x) = f(x') \land
@@
   the points of the original space. More generally,
-  \citet{continuous-lattice-book} (Section V-6) describes \emph{Domain
+  \citet{continuous-lattice-book} (Section V-6) describe \emph{domain
     environments}, which are domains whose maximal elements are

### 2025-07-08 main/models.tex
Pare back section 4 just to be on first-order Galois slicing in Fam(LatGal).
Citations.

--- 2025-07-08 main/models.tex
@@
   $f : (X, \partial X) \to (Y, \partial Y)$ consists of a pair of a
-  function $f : X \to Y$ and a familiy of morphisms of $\cat{C}$,
+  function $f : X \to Y$ and a family of morphisms of $\cat{C}$,
   $\partial f : \Pi_{x : X}.\,\cat{C}(\partial X(x), \partial Y(f\,x))$.
@@
 lenses (the ``simply typed'' version of the $\Fam$ construction). As
-in \cite{vakar2021}, the idea is to formally separate functions on
+in \citet{vákár22}, the idea is to formally separate functions on
 points and their forward/reverse tangent maps for the purposes of
@@
 \begin{definition}
+  \item
   \begin{enumerate}
@@
 
-\subsection{Semantic \GPS in $\Fam(\LatGal)$}
+\subsection{Semantic Galois Slicing in $\Fam(\LatGal)$}
 \label{sec:semantic-gps}

### 2025-07-08 main/approx-as-tangents.tex
Improve some citations.

@@
 This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
-A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12).
+A textbook description of stable functions in the context of domain theory is given by \citet[Chapter 12]{amadio-curien}.
 We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.
@@
   the points of the original space. More generally,
-  \citet{continuous-lattice-book} (Section V-6) describe \emph{domain
+  \citet[Section V-6]{continuous-lattice-book} describe \emph{domain
     environments}, which are domains whose maximal elements are

### 2025-07-08 main/approx-as-tangents.tex
Citations.

@@
   An example that is conditionally multiplicative, but does not admit
-  a backards map of approximations (from \citet{amadio-curien} just
-  before Lemma 12.2.3, originally due to Berry) is given by defining
+  a backards map of approximations (from \citet[just
+  before Lemma 12.2.3]{amadio-curien}, originally due to Berry) is given by defining
   $\mathrm{unstable} : D \to \{\bot \sqsubseteq \top\}$, where
@@
 downsets $\downset{x}$ are \emph{complete}
-lattices. \citet{amadio-curien} (Theorem 12.FIXME) details the
+lattices. \citet[Theorem 12.FIXME]{amadio-curien} details the
 proof. Intuitively, to generate the best approximation of an input
@@
   their approximations using the \emph{Category of Families}
-  construction we detail in the next section.
+  construction we detail in \secref{models-of-total-gps:fam}.
 \item A more practical reason is that we wish to formalise our

### 2025-07-08 main/models.tex
Fix cite.
Citations.
Typo; more citations.

--- 2025-07-08 main/models.tex
@@
   arising from the use of fibrations for categorical logical relations
-  proofs, due to \citet{hermida99} (Corollary 4.12). If we assume that
+  proofs, due to \citet[Corollary 4.12]{hermida99}. If we assume that
   $\cat{C}$ is itself cartesian closed and has all products, then we
@@
 coproducts. They can be defined in any category, as shown by
-\citet{karvonen18}, but for our purposes it will be more convenient to
+\citet{karvonen20}, but for our purposes it will be more convenient to
 use the shorter definition in categories enriched in commutative
@@
     adjoints are summed using joins. The zero maps are given by the
-    constantly $\top$ and constrantly $\bot$ functions
+    constantly $\top$ and constantly $\bot$ functions
     respectively. Products are given by the cartesian product of the

### 2025-07-08 main/models.tex
semi-lattice -> semilattice.

@@
     the ``forward'' direction.
-  \item The category $\MeetSLat$ has meet-semi-lattices with top as
+  \item The category $\MeetSLat$ has meet-semilattices with top as
     objects and monotone finite meet preserving functions as
     morphisms.
-  \item The category $\JoinSLat$ has join-semi-lattices with bottom as
+  \item The category $\JoinSLat$ has join-semilattices with bottom as
     objects and monotone finite join preserving functions as
@@
   \item The $\LatGal$ example splits into two separate categories
-    $\MeetSLat$ and $\JoinSLat$ of meet semi-lattices and join
-    semi-lattices respectively, with their respective homomorphisms as
+    $\MeetSLat$ and $\JoinSLat$ of meet semilattices and join
+    semilattices respectively, with their respective homomorphisms as
     the morphisms. As for $\LatGal$, these categories are
@@
 $\MeetSLat \times \JoinSLat^\op$. Objects in this category consist of
-\emph{separate} meet- and join-semi-lattices and potentially unrelated
+\emph{separate} meet- and join-semilattices and potentially unrelated
 forward meet-preserving and backward join-preserving maps. We first
@@
   that these are very different constructions: elements of a product
-  of meet-semi-lattices consist of (possibly infinite) tuples of
+  of meet-semilattices consist of (possibly infinite) tuples of
   elements, while elements of a coproduct of join-semilattices consist
   of \emph{finite} formal joins of elements quotiented by the
-  join-semi-lattice equations. Since $\JoinSLat$ has all coproducts,
+  join-semilattice equations. Since $\JoinSLat$ has all coproducts,
   $\JoinSLat^\op$ has all products, and so
@@
 we interpret the base types as objects in $\Fam(\LatGal)$. At
-higher-order, however, the meet-semi-lattice and join-semi-lattice
+higher-order, however, the meet-semilattice and join-semilattice
 sides of the interpretation will diverge, and it is no longer clear

### 2025-07-08 main/models.tex

@@
 \subsection{The Category of Families Construction}
+\label{sec:models-of-total-gps:fam}
 
@@
   Coproducts in $\Fam(\cat{C})$ are \emph{extensive}
-  \cite{carboni-lack-walters93} (Proposition 2.4).
+  \cite[Proposition 2.4]{carboni-lack-walters93}.
 \end{proposition}

### 2025-07-08 main/approx-as-tangents.tex
berry78 -> berry79.

@@
 In light of the \exref{non-stable-function}, we turn to
-\citet{berry78}'s definition of \emph{stable function} that requires
+\citet{berry79}'s definition of \emph{stable function} that requires
 the existence of a reverse mapping directly, even without assuming
@@
 
-We have seen that \citet{berry78}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives.
+We have seen that \citet{berry79}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives.
 In the next section, we describe a model based on these ideas suitable for constructing executable models of total higher order languages.

### 2025-07-08 main/models.tex
Minor corrections.

@@
     semilattices respectively, with their respective homomorphisms as
-    the morphisms. As for $\LatGal$, these categories are
+    the morphisms. As with $\LatGal$, these categories are
     $\CMon$-enriched and have finite products, and so finite
@@
   The functor
-  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ defined
+  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
   on objects as
@@
 With this embedding functor, we can see that the interpretation of
-first-order types will be the same up to isormorphism in
+first-order types will be the same up to isomorphism in
 $\Fam(\LatGal)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$, as long as

### 2025-07-08 main/approx-as-tangents.tex
Tangent categores, Cartesian differential categories.

@@
 functions, we have not stated any mathematical theorems substantiating
-this. \emph{Tangent Categories} \cite{tangent-categories} are a
+this. \emph{Tangent Categories} \cite{rosický84,cockett14,cockett18} are a
 categorical axiomatisation of the properties of manifolds and smooth

### 2025-07-08 main/related-work.tex
'language' no longer needed in notes.
Tweaks.
Purging section 4.
Bit more blurb on tangent categories.

@@
-\section{Related work}
+\section{Related Work}
 \label{sec:related-work}
@@
 
-\cite{cockett14,cockett18}
+\emph{Tangent categories}, due originally to \citet{rosický84} and developed by \citet{cockett14,cockett18},
+provide an abstract categorical framework for reasoning about differentiation, inspired by the structure of
+the tangent bundle in differential geometry. In a tangent category, each object $X$ is equipped with a tangent
+bundle $T(X)$, and each morphism $f: X \to Y$ has a corresponding differential map $T(f): TX \to TY$
+satisfying axioms analogous to the chain rule and linearity of differentiation. Tangent categories generalise
+Cartesian differential categories \cite{cdcs}, which model differentiation over Cartesian closed categories
+using a syntactic derivative operator. The motivation behind tangent categories is to \roly{Bob to say
+something here..}
 

### 2025-07-08 main/approx-as-tangents.tex
Typos.

--- 2025-07-08 main/approx-as-tangents.tex
@@
 functions, we have not stated any mathematical theorems substantiating
-this. \emph{Tangent Categories} \cite{tangent-categories} are a
+this. \emph{Tangent Categories} \cite{cockett14,cockett18} are a
 categorical axiomatisation of the properties of manifolds and smooth

### 2025-07-08 main/models.tex
Typos.

@@
 differentiation for higher order programs. Since the forward and
-reverse derivatives of a smooth map intrinsically defined,
+reverse derivatives of a smooth map are intrinsically defined,
 \citet{vakar2021}'s correctness theorem shows that, for programs with
@@
 the forward derivative of the defined function on the reals (and
-reverse derivative for $\Fam(\Vect^\op)$.
+reverse derivative for $\Fam(\Vect^\op)$).
 

### 2025-07-08 main/introduction.tex
vákár -> vakar (easier).

@@
 It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately (but somewhat redundantly) defined ``backwards evaluation'' operation.
-As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vákár22}, which points to a way to develop such an approach.
+As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vakar22}, which points to a way to develop such an approach.
 We have already hinted at this in the description above, but let us now make it explicit.

### 2025-07-08 main/models.tex
vákár -> vakar (easier).

@@
 lenses (the ``simply typed'' version of the $\Fam$ construction). As
-in \citet{vákár22}, the idea is to formally separate functions on
+in \citet{vakar22}, the idea is to formally separate functions on
 points and their forward/reverse tangent maps for the purposes of

### 2025-07-08 main/related-work.tex
vákár -> vakar (easier).

@@
 giving a compositional account of differentiation based on duality and linear maps. Vákár and
-collaborators~\cite{vákár22,nunes2023} developed the CHAD framework which inspired this paper, using
+collaborators~\cite{vakar22,nunes2023} developed the CHAD framework which inspired this paper, using
 Grothendieck constructions over indexed categories to capture both values and their tangents in a

### 2025-07-08 old/main/first-order.tex
Purging section 4.
vákár -> vakar (easier).

--- 2025-07-05 main/first-order.tex
@@
 
-Our denotational approach to \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vákár22,nunes2023}.
+Our denotational approach to \GPS is based on the CHAD (Combinatory Homomorphic Automatic Differentiation) work of Mattias Vákár and others~\cite{vakar22,nunes2023}.
 Before we go into details, we present a high-level overview of their approach and how we have applied it to \GPS.
@@
 
-We now consider how to adapt the CHAD framework~\cite{vákár22,nunes2023} to interpret programs as functions
+We now consider how to adapt the CHAD framework~\cite{vakar22,nunes2023} to interpret programs as functions
 along with their forward and backward approximation maps. Since the appropriate lattice of approximation
@@
 % isomorphic to $\RR^n$, so the forward and backward derivatives of a differentiable function $f: \RR^m \to
-% \RR^n$ at a point $x$ have the following form \todo{Can probably delete now as covered in
+% \RR^n$ at a point $x$ have the following form \roly{Can probably delete now as covered in
 % \secref{approx-as-tangents}}:

### 2025-07-08 main/models.tex
Fix Vakar citation.

@@
 the vector space operations themselves are finitary. This is the
-solution that \cite{vakar2021} uses for the semantics of forward
+solution that \cite{vakar22} uses for the semantics of forward
 ($\Fam(\Vect)$) and reverse ($\Fam(\Vect^\op)$) automatic
@@
 reverse derivatives of a smooth map are intrinsically defined,
-\citet{vakar2021}'s correctness theorem shows that, for programs with
+\citet{vakar22}'s correctness theorem shows that, for programs with
 first-order type, the interpretation in $\Fam(\Vect)$ correctly yields

### 2025-07-08 main/approx-as-tangents.tex
Fix chain-rule refs.

@@
 \begin{remark}[Chain Rule]
-  \label{rem:cm-chain-rule}
+  \label{rem:chain-rule-cm}
   It is almost a triviality at this point, but the crucial point is
@@
   An example that is conditionally multiplicative, but does not admit
-  a backards map of approximations (from \citet[just
+  a backwards map of approximations (from \citet[just
   before Lemma 12.2.3]{amadio-curien}, originally due to Berry) is given by defining
@@
 \begin{remark}[Chain Rule]
+  \label{rem:chain-rule-stable}
   As for the ``forward derivatives'' of conditionally multiplicative

### 2025-07-08 main/related-work.tex
Fix autodiff ref.

@@
 
-Automatic differentiation (AD), discussed in \secref{first-order:autodiff}, is the idea of computing
+Automatic differentiation (AD), discussed in \secref{approx-as-tangents:autodiff}, is the idea of computing
 derivatives of functions expressed as programs by systematically applying the chain rule. The observation that

### 2025-07-08 main/models.tex
Standardise on 'cartesian'.

@@
 $\Fam(\MeetSLat \times \JoinSLat^\op)$, which is a cartesian closed
-category with all coproducts, ienough to interpret the total
+category with all coproducts, enough to interpret the total
 higher-order language we define in the next section with primitive
@@
 tangents, somewhat analogous to the situation with manifolds. Also
-note that, contrary to the $\Fam(\FinVect)$ representation of
-manifolds and differentiable maps, (if this conjecture holds) we do
+note that, if this conjecture holds, then contrary to the $\Fam(\FinVect)$ representation of
+manifolds and differentiable maps, we do
 not throw away information about higher derivatives. It is retained in

### 2025-07-08 main/related-work.tex
Standardise on 'cartesian'.

@@
 satisfying axioms analogous to the chain rule and linearity of differentiation. Tangent categories generalise
-Cartesian differential categories \cite{cdcs}, which model differentiation over Cartesian closed categories
+Cartesian differential categories \cite{cdcs}, which model differentiation over cartesian closed categories
 using a syntactic derivative operator. The motivation behind tangent categories is to \roly{Bob to say

### 2025-07-09 build.sh
[build script] Start.

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+set -xe
+
+AGDA_FLAGS="--ignore-interfaces"
+
+ENTRYPOINTS=(
+  src/example.agda
+)
+
+for file in "${ENTRYPOINTS[@]}"; do
+  agda $AGDA_FLAGS "$file"
+done

### 2025-07-09 build.sh
'everything.agda' to gather entrypoints.  is out of date.

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-set -xe
-
-AGDA_FLAGS="--ignore-interfaces"
-
-ENTRYPOINTS=(
-  src/example.agda
-)
-
-for file in "${ENTRYPOINTS[@]}"; do
-  agda $AGDA_FLAGS "$file"
-done

### 2025-07-09 main/definability.tex
[ tex ] definability section

\section{Correctness of the Higher-order Interpretation}
\label{sec:definability}

Introduction: what correctness properties we want; why we can't
compute the left adjoints from the forward maps (no completeness).

\begin{theorem}[\citet{fiore-simpson99}]
  \label{thm:glr-definability}
  Let $\cat{C}$ be an extensive bicartesian category, $\cat{D}$ be a
  bicartesian closed category, and $F : \cat{C} \to \cat{D}$ a functor
  preserving finite products and coproducts. Then there is a category
  $\GLR(F)$ and functors $p : \GLR(F) \to \cat{D}$ and
  $\hat{F} : \cat{C} \to \GLR(F)$, such that:
  \begin{enumerate}
  \item $\GLR(\cat{D},F)$ is bicartesian closed;
  \item $F = p \circ \hat{F} : \cat{C} \to \cat{D}$;
  \item The functor $p$ strictly preserves the bicartesian closed structure; and
  \item The functor $\hat{F}$ is full and preserves the bicartesian structure.
  \end{enumerate}
\end{theorem}

\begin{theorem}
  \label{thm:language-definability}
  For all $\Gamma \vdash M : \tau$, where $\Gamma$ and $\tau$ are
  first-order, there exists a morphism
  $g \in \Fam(\LatGal)(\sem{\Gamma}_{\mathit{fo}},
  \sem{\tau}_{\mathit{fo}})$ such that
  $\sem{\Gamma \vdash M : \tau} = F(g)$.
\end{theorem}

\begin{remark}
  We have formalised \thmref{glr-definability} and
  \thmref{language-definability} in Agda, except for the fact that
  $\Fam(C)$ is extensive\footnote{For reviewers: we intend to fix this
    before any final version.}. Our formalisation uses E-categories
  (\cite{palmgren-wilander}) so we cannot state
  \thmref{glr-definability} as given because the concept of strict
  preservation does not make sense. Therefore, the formalised version
  of \thmref{thm:language-definability} is ``up to isomorphism''.

  We believe that this result also extends to the list type, since it
  is constructed from an infinte coproduct, but have not yet checked
  all the details.
\end{remark}

\begin{proof}
  Instantiate \thmref{glr-definability} with
  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$. By
  \propref{XXX} we know that $F$ preserves finite products and
  coproducts. $\Fam(\LatGal)$ has extensive coproducts because it is
  the free coproduct completion. Thus we get the category $\GLR(F)$ in
  which we can interpret the calculus with the property that this
\end{proof}

FIXME: another property we might want: The interpretations in
$\Fam(\MeetSLat)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$ agree at
first order.

### 2025-07-09 main/language.tex
language subsection -> section.
Cleanup.
Tweaks.
Reinstate missing semantics.
[ tex ] fiddling
Just use + for binary coprod

--- 2025-07-03 main/language.tex
@@
-\subsection{Higher-order language}
+\section{Higher-Order Language}
+\label{sec:language}
 
-We introduce a total higher-order functional language, over a set $\PrimTy$ of primitive types $\rho$ and sets
-$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$, equipped with sums, products, functions
-and a data type of lists. We omit a treatment of general inductive or coinductive types.
+To model Galois slicing semantically for higher-order programs, we define a simple total functional
+programming language, extending the simply-typed lambda calculus.
 
@@
 
-% \input{fig/positivity} -- follow nunes2023 instead
+The language includes base types $\rho$ drawn from a set $\PrimTy$, along with standard type formers for sums,
+products, functions and lists, and an additional lifting type constructor $\tyLift$ to allow approximation
+points to be added explicitly, as discussed in \secref{models-of-total-gps}. Terms include variables, the
+usual introduction and elimination forms, a monadic return and bind for lifted terms, and primitive operations
+$\phi$ of arity $n$ drawn from a family of sets $\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$.
+
+The language is intentionally minimal: it excludes general recursion, and general inductive or coinductive
+types, which will consider in future work (\secref{conclusion}). Typing judgments for terms are standard and
+shown in \figref{typing}, with the usual rules for products, sums, functions, lists and lifting.
+
 \input{fig/syntax}
-% \input{fig/kinding}
 \input{fig/typing}
-
-\figrefTwo{syntax}{typing} give the syntax and typing rules.
 
@@
 
-% \input{fig/mu-polynomial}
-% \input{fig/inductive-types/semantics}
-% \figref{semantics:types} gives the interpretation of types $\Delta \vdash \tau$ with $\length{\Delta}
-% = n$ as functors $\Sem^n \to \Sem$, or as objects of $\Sem$ in the case $n = 0$.
-% \todo{Notation for partially applied functor used in \figref{semantics:types}}
-%
-% \begin{definition}[$\mu$-polynomial]
-% Suppose $\cat{C}$ a category with finite coproducts $(\coprod, 0)$ and finite products $(\times, 1)$. Define
-% $\muPoly_{\cat{C}}$ to be the smallest subcategory of $\Cat$ generated inductively by the rules in
-% \figref{mu-polynomial}.
-% \end{definition}
-%
-% For any $\mu$-polynomial endofunctor $F: \cat{C} \to \cat{C}$ with initial algebra $\mu F$, write $\inMap_F$
-% for the structure map $F\mu F \to \mu F$ and write $\cata_f$ for the unique morphism $\mu F \to X$ from the
-% initial $F$-algebra to any $F$-algebra $(X, f: FX \to X)$.
-
 \input{fig/semantics}
@@
 To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(\coprod, 0)$ and exponentials
+category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(+, 0)$ and exponentials
 $\internalHom{X}{Y}$, with evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$,

### 2025-07-09 build-suppl.sh
Script to build supplementary materials.

@@ -1,0 +1,3 @@
+#!/bin/bash
+set -xe
+zip -r suppl.zip agda -x "agda/_build/*"

### 2025-07-09 main/related-work.tex
WIP.
fam -> indexed-family.

@@
 connection to stable functions in relation to minimal slices for short-circuiting operations was alluded to
-in~\citet{perera13}. Subsequent work extended the approach to languages with assignment and
+in~\citet{perera13}, but not explored. Subsequent work extended the approach to languages with assignment and
 exceptions~\cite{ricciotti17} and concurrent systems, applying Galois slicing to the
@@
 the tangent bundle in differential geometry. In a tangent category, each object $X$ is equipped with a tangent
-bundle $T(X)$, and each morphism $f: X \to Y$ has a corresponding differential map $T(f): TX \to TY$
+bundle $T(X)$, and each morphism $f: X \to Y$ has a corresponding differential map $T(f): T(X) \to T(Y)$
 satisfying axioms analogous to the chain rule and linearity of differentiation. Tangent categories generalise

### 2025-07-09 main/approx-as-tangents.tex
Fix autodiff ref.
Add bounded-meet to 'everything' and reference paper theorem.
Correction to chain rule.

@@
 \subsection{Manifolds, Smooth Functions, and Automatic Differentiation}
+\label{sec:approx-as-tangents:autodiff}
 
@@
   \begin{itemize}
-  \item $\pushf{(g \circ f)}_x = \pushf{g}_{f(x)} \circ \pushf{f}_x : \tangents_x(M) \linearto \tangents_{g(f(x)}(K)$
-  \item $\pullf{(g \circ f)}_x = \pullf{f}_x \circ \pullf{g}_{f(x)} : \cotangents_{f(x)}(K) \linearto \cotangents_x(M)$
+  \item $\pushf{(g \circ f)}_x = \pushf{g}_{f(x)} \circ \pushf{f}_x : \tangents_x(M) \linearto \tangents_{g(f(x))}(K)$
+  \item $\pullf{(g \circ f)}_x = \pullf{f}_x \circ \pullf{g}_{f(x)} : \cotangents_{g(f(x))}(K) \linearto \cotangents_x(M)$
   \end{itemize}
@@
 \begin{theorem}
-  Bounded meet lattices and conditionally multiplicatve functions form
+  Bounded meet lattices and conditionally multiplicative functions form
   a category $\CM$. This category has products, coproducts, and
@@
   \end{displaymath}
-  In the poset $\mathbb{B}_\bot^2$, a typical poset of approximations of a fully defined element is shown to the right.
-For $\mathrm{strictOr}$, any approximation that isn't the fully defined input is mapped to $\bot$, while $\mathrm{shortCircuitOr}$ maps the partially defined $(\mathsf{tt},\bot)$ to $\mathsf{tt}$.
-Thus, even though these functions operate identically on fully defined inputs, they differ in their \emph{derivatives} on partially defined input, exposing how .
-That they are cm can be checked by examining their restrictions' behaviour.
-If we take the approximations $(\bot,\mathsf{ff})$ and $(\mathsf{tt},\bot)$, then their meet is $(\bot,\bot)$; $\mathrm{strictOr}$ maps all three elements to $\bot$, so is cm here since $\bot \wedge \bot = \bot$; and $\mathrm{shortCircuitOR}$ has $(\bot,\mathsf{ff}) \mapsto \bot$ and $(\mathsf{tt},\bot) \mapsto \mathsf{tt}$, the meet of which is $\bot = \mathrm{shortCircuitOR}(\bot,\bot)$.
-Other combinations can be checked similarly.
+  In the poset $\mathbb{B}_\bot^2$, a typical poset of approximations of a fully defined element is shown to
+  the right. For $\mathrm{strictOr}$, any approximation that isn't the fully defined input is mapped to
+  $\bot$, while $\mathrm{shortCircuitOr}$ maps the partially defined $(\mathsf{tt},\bot)$ to $\mathsf{tt}$.
+  Thus, even though these functions operate identically on fully defined inputs, they differ in their
+  \emph{derivatives} on partially defined input, exposing how they explore their arguments differently. That
+  they are cm can be checked by examining their restrictions' behaviour. If we take the approximations
+  $(\bot,\mathsf{ff})$ and $(\mathsf{tt},\bot)$, then their meet is $(\bot,\bot)$; $\mathrm{strictOr}$ maps
+  all three elements to $\bot$, so is cm here since $\bot \wedge \bot = \bot$; and $\mathrm{shortCircuitOR}$
+  has $(\bot,\mathsf{ff}) \mapsto \bot$ and $(\mathsf{tt},\bot) \mapsto \mathsf{tt}$, the meet of which is
+  $\bot = \mathrm{shortCircuitOR}(\bot,\bot)$. Other combinations can be checked similarly.
 \end{example}

### 2025-07-09 main/introduction.tex
Fix Vakar citation.
Correction to chain rule.

@@
 Existing approaches to program slicing are often tied to particular programming languages or implementations.
-In this paper we develop a general categorical approach to program slicing and \roly{complete}.
+In this paper we develop a general categorical approach to program slicing and \todo{complete}.
 
@@
   \label{ex:introduction-example}
-  The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then to compute the sum of the numbers:
+  The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then computing the sum of the numbers:
   \begin{displaymath}
@@
   \end{center}
-  The output approximation lattice looks like this, where $1$ is the actual data point that was returned in both runs of the program, and $\bot$ indicates that we are approximating this piece of data away:
+  In both runs of the program, the output approximation lattice looks like this, where $1$ is the actual data point that was returned, and $\bot$ indicates that we are approximating this piece of data away:
   \begin{center}
@@
   \end{center}
-  These are not the only choices of approximation lattices that we could have taken.
+  These are not the only choices of approximation lattices that we could have made.
 For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not any of the other data or the overall structure.
@@
 Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information.
-We elaborate on this further in \secref{first-order}.
+We elaborate on this further in \secref{models-of-total-gps}.
 
@@
   In this case, the forwards portion of the Galois connection tells us for each approximation of the input whether or not it is sufficient to compute the output.
-In this example, if we provide insufficient data to compute the output, then we will get an underapproximated output.
+If we provide insufficient data to compute the output, then we will get an underapproximated output.
 For example, $\partial (\mathit{query}\, \mathsf{a})_f([(\mathsf{a},0),(\mathsf{b},\bot),(\mathsf{a},\bot)]) = \bot$ because we need all the values associated with the label $\mathsf{a}$ to compute their sum.

### 2025-07-09 main/approx-as-tangents.tex
Theorem number for Amadio and Curien ref.

@@
 
-For the purposes of \GPS, we are not interested in using
-approximations to model computation on infinite objects, but instead
+For the purposes of \GPS, we are interested in using
+approximations not to model computation on infinite objects, but instead
 for revealing how programs explore their inputs when producing parts
@@
 When giving a denotational semantics for sequential programming
-languages Scott-continuous functions are too permissive. Famously,
+languages, Scott-continuous functions are too permissive. Famously,
 \citet{plotkin77lcf}'s Parallel OR (\exref{parallel-or}, below) is
@@
 deterministic. This non-determinism results in functions that whose
-output cannot be assigned a unique minimal input that caused
+output cannot be assigned a unique minimal input that account for
 it. Therefore, continuous functions are incompatible with the central
@@
 
-\emph{Stability} is a property that can be required of monotone functions that was invented by \citet{berry79} in an attempt to capture sequentiality.
-This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS.
+\emph{Stability} is a property that can be required of monotone functions, and was invented by \citet{berry79} in an attempt to capture sequentiality.
+This was unsuccessful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to the problem of computing forward and backwards maps of approximations needed in \GPS.
 A textbook description of stable functions in the context of domain theory is given by \citet[Chapter 12]{amadio-curien}.
@@
 \begin{proof}
-  \AGDA. See \citet{amadio-curien} (FIXME: theorem reference) for the
+  \AGDA. See \citet[Theorem 12.2.8]{amadio-curien} for the
   case when the posets are also cpos. The crucial technical step,

### 2025-07-09 main/approx-as-tangents.tex
Theorem number for Amadio and Curien ref.
Minor tweaks.

@@
 deterministic. This non-determinism results in functions that whose
-output cannot be assigned a unique minimal input that account for
+output cannot be assigned a unique minimal input that accounts for
 it. Therefore, continuous functions are incompatible with the central
@@
 
-\emph{Stability} is a property that can be required of monotone functions, and was invented by \citet{berry79} in an attempt to capture sequentiality.
+\emph{Stability} is a property that can be required of monotone functions, that was invented by \citet{berry79} in an attempt to capture sequentiality.
 This was unsuccessful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to the problem of computing forward and backwards maps of approximations needed in \GPS.

### 2025-07-10 main/conclusion.tex
Moving unused Agda files to .
Just use + for binary coprod
laird07, fix ref.

@@
 2.
-Conjecture that the category with $(X : Cpo, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and galois connections is the right setting.
-Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains (Laird).
+Conjecture that the category with $(X : \namedcat{Cpo}, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and galois connections is the right setting.
+Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains (\cite{laird07}).
 
@@
 Smeding and Vákár on translations. To formalise this, we'd need
-to... See also $\Fam(\PSh_{\CMon}(Syn))$?
+to... See also $\Fam(\PSh_{\CMon}(\namedcat{Syn}))$?

### 2025-07-10 main/definability.tex
laird07, fix ref.
Citations.

@@
   preservation does not make sense. Therefore, the formalised version
-  of \thmref{thm:language-definability} is ``up to isomorphism''.
+  of \thmref{language-definability} is ``up to isomorphism''.
 
   We believe that this result also extends to the list type, since it
-  is constructed from an infinte coproduct, but have not yet checked
+  is constructed from an infinite coproduct, but have not yet checked
   all the details.

### 2025-07-10 main/models.tex
Citations.

@@
   \item The category $\MeetSLat$ has meet-semilattices with top as
-    objects and monotone finite meet preserving functions as
+    objects and finite meet preserving functions as
     morphisms.
   \item The category $\JoinSLat$ has join-semilattices with bottom as
-    objects and monotone finite join preserving functions as
+    objects and finite join preserving functions as
     morphisms.

### 2025-07-10 main/first-order.tex
Pare back 4.1.

@@
 \label{sec:fam:galois-slicing}
-
-The total category $\Fam(\LatGal)$ is then a suitable setting for interpreting first-order programs for \GPS.
-It has as objects $(X, \partial X)$, all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
-$\partial X(x)$, and as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of a
-function $f: X \to Y$ and for every $x \in X$, a Galois connection $\partial f(x): \partial X(x) \to \partial
-Y(f(x))$. The reindexing along $f$ in the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f,
-\reindex{\partial g}{f} \comp \partial f)$ selects the appropriate lattice of approximations and Galois
-connection at each point, using the baseline (unapproximated) function $f$.
 

### 2025-07-10 main/models.tex
Suggestion to give F a more meaningful name.

@@
 \begin{proposition}
+  \label{prop:ho-embedding}
   The functor

### 2025-07-10 main/definability.tex
section numbering.
Suggestion to give F a more meaningful name.

@@
-\section{Correctness of the Higher-order Interpretation}
+\section{Correctness of the Higher-Order Interpretation}
 \label{sec:definability}
@@
 
+\roly{Clarify $F$ here refers to \propref{ho-embedding} (esp.~as \thmref{glr-definability} quantifies over
+arbitrary $F$), perhaps by giving it a more meaningful name}
+
 \begin{remark}
@@
   $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$. By
-  \propref{XXX} we know that $F$ preserves finite products and
+  \propref{ho-embedding} we know that $F$ preserves finite products and
   coproducts. $\Fam(\LatGal)$ has extensive coproducts because it is

### 2025-07-10 main/language.tex
section numbering.
How Fam(MeetSLat×JoinSLatop) implements Sem.

@@
 
-\subsubsection{Syntax}
+\subsection{Syntax}
 \label{sec:language:syntax}
@@
 
-\subsubsection{Semantics}
+\subsection{Semantics}
 \label{sec:language:semantics}
@@
 \times Y)$
-\item for each primitive type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
+\item for each base type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
 \item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
@@
 \to \sem{\tau}$.
+
+\subsubsection{Interpretation for higher-order \GPS}
+
+$\Fam(\MeetSLat \times \JoinSLat^\op)$ meets the requirements of $\Sem$: it is bicartesian closed
+(\corref{mslat-jslat-bcc}), can model $\List(X)$ as an infinite coproduct
+(\secref{models-of-total-gps:coproducts-and-products}), \roly{$\Lift$ monad, base types and operations.}

### 2025-07-10 main/models.tex
Standardise notation of List and L (lifting) functors.
How Fam(MeetSLat×JoinSLatop) implements Sem.

@@
 \subsubsection{Coproducts and Products}
+\label{sec:models-of-total-gps:coproducts-and-products}
 
@@
 \begin{displaymath}
-  \mathrm{List}\,X = \coprod_{n \in \mathbb{N}} X^n
+  \List(X) = \coprod_{n \in \mathbb{N}} X^n
 \end{displaymath}
@@
 \begin{corollary}
+  \label{cor:mslat-jslat-bcc}
   $\Fam(\MeetSLat \times \JoinSLat^\op)$ is cartesian closed and has
@@
 
-On objects $X$ with trivial approximations, we have $L(X) \cong T(X)$,
+On objects $X$ with trivial approximations, we have $\Lift(X) \cong T(X)$,
 but for more structured objects we get usage information at every level.

### 2025-07-10 main/language.tex
Explain how interpretation of primitive types arrives in \Fam(\MeetSLat \times \JoinSLat^\op).

@@
 
-$\Fam(\MeetSLat \times \JoinSLat^\op)$ meets the requirements of $\Sem$: it is bicartesian closed
-(\corref{mslat-jslat-bcc}), can model $\List(X)$ as an infinite coproduct
-(\secref{models-of-total-gps:coproducts-and-products}), \roly{$\Lift$ monad, base types and operations.}
+Given an interpretation of base types and operations in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times
+\JoinSLat^\op)$ can implement $\Sem$. It is bicartesian closed (\corref{mslat-jslat-bcc}) and can model
+$\List(X)$ as an infinite coproduct (\secref{models-of-total-gps:coproducts-and-products}). An interpretation
+of base types and operations can be obtained from their interpretation in $\Fam(\LatGal)$ via the embedding
+...
+\roly{$\Lift$ monad.}

### 2025-07-10 main/language.tex
WIP.
Formalise signature, model of signature.

@@
 \item for each base type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
-\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}:
-\sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$
+\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism
+$\sem{\phi}_{\Op}: \sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$
 \item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
@@
 
+\paragraph{Base types and operations}
+Define a \emph{signature} to be a set $\PrimTy$ of primitive types $\rho$ and a family of sets
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$. A \emph{model} of a signature in a
+category $\cat{C}$ with finite products and a terminal object assigns to each base type $\rho \in \PrimTy$ an
+object $\sem{\rho}_{\PrimTy}$ in $\cat{C}$ and to each primitive operation $\phi \in
+\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots \times
+\sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$.
+
 Given an interpretation of base types and operations in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times

### 2025-07-10 main/language.tex
Introduce signatures in syntax.

--- 2025-07-10 main/language.tex
@@
 
-The language includes base types $\rho$ drawn from a set $\PrimTy$, along with standard type formers for sums,
-products, functions and lists, and an additional lifting type constructor $\tyLift$ to allow approximation
-points to be added explicitly, as discussed in \secref{models-of-total-gps}. Terms include variables, the
-usual introduction and elimination forms, a monadic return and bind for lifted terms, and primitive operations
-$\phi$ of arity $n$ drawn from a family of sets $\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$.
+The language is parameterised by a signature $\Sigma = (\PrimTy, \PrimOp)$ consisting of a set $\PrimTy$ of
+base types $\rho$ and a family of sets $\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$ of
+arity $n$ over those base types. The syntax is defined in \figref{syntax}. Types includes base types $\rho$
+drawn from $\PrimTy$, along with standard type formers for sums, products, functions and lists, plus a type
+constructor $\tyLift$ to allow approximation points to be added explicitly, as discussed in
+\secref{models-of-total-gps}. Terms include variables, the usual introduction and elimination forms, a monadic
+return and bind for lifted terms, and primitive operations $\phi$.
 
 The language is intentionally minimal: it excludes general recursion, and general inductive or coinductive
-types, which will consider in future work (\secref{conclusion}). Typing judgments for terms are standard and
-shown in \figref{typing}, with the usual rules for products, sums, functions, lists and lifting.
+types, which we will consider in future work (\secref{conclusion}). Typing judgments for terms are standard
+and shown in \figref{typing}, with the usual rules for products, sums, functions, lists and lifting.
 
@@
 
+\paragraph{Base types and operations}
+A \emph{model} of a signature in a category $\cat{C}$ with finite products and a terminal object assigns to
+each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimTy}$ in $\cat{C}$ and to each primitive
+operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy}
+\times \ldots \times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$.
+
 Given an interpretation of base types and operations in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times

### 2025-07-10 main/language.tex
Do I want to say what a model of List looks like too?

@@
 To model Galois slicing semantically for higher-order programs, we define a simple total functional
-programming language, extending the simply-typed lambda calculus.
+programming language over a signature of base types and primitive operations, extending the simply-typed
+lambda calculus.
 
@@
 
-To give the semantics for the language defined in \figrefTwo{syntax}{typing}, we fix a bicartesian closed
-category $\Sem$ with finite products $(\times, 1)$, finite coproducts $(+, 0)$ and exponentials
-$\internalHom{X}{Y}$, with evaluation morphisms $\eval_{X,Y}$ and currying isomorphisms $\lambda_{X,Y,Z}$,
-plus the following additional structure:
+Models of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object; such a
+model $M$ assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each
+primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}:
+\sem{\rho_1}_{M} \times \ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$.
+
+To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ plus a model $M$ of
+$\Sigma$ in $\Sem$, with the following additional structure:
 \begin{enumerate}
@@
 \times Y)$
-\item for each base type $\rho \in \PrimTy$, an object $\sem{\rho}_{\PrimTy}$
-\item for each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism
-$\sem{\phi}_{\Op}: \sem{\rho_1} \times \ldots \times \sem{\rho_n} \to \sem{\rho}$
 \item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
@@
 \paragraph{Base types and operations}
-A \emph{model} of a signature in a category $\cat{C}$ with finite products and a terminal object assigns to
-each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimTy}$ in $\cat{C}$ and to each primitive
-operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy}
-\times \ldots \times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$.
-
 Given an interpretation of base types and operations in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times

### 2025-07-10 main/models.tex
[ tex ] Some discussion of using Fam(LatGal)

@@
 The fact that morphisms in $\Fam(C)$ compose according to a chain rule
-means that the categories we have seen so far embed into $\Fam(C)$ for
-appropriate $C$. If we let $\FinVect$ be the category of finite
-dimensional real vector spaces, then:
+means that the categories we considered in \secref{approx-as-tangents}
+embed into $\Fam(C)$ for appropriate $C$. If we let $\FinVect$ be the
+category of finite dimensional real vector spaces and linear maps,
+then:
 
@@
 
-\begin{definition}
-  \item
 \begin{enumerate}
@@
   Galois connections as morphisms, with the right adjunct going in
-    the ``forward'' direction.
+  the ``forward'' direction. The category $\Fam(\LatGal)$ is our
+  preferred model for total functions with Galois sliicing. We
+  explore some specific examples in this category in
+  \secref{semantic-gps}.
 \item The category $\MeetSLat$ has meet-semilattices with top as
-    objects and finite meet preserving functions as
-    morphisms.
+  objects and monotone finite meet preserving functions as
+  morphisms. The category $\Fam(\MeetSLat)$ provides a model of
+  total functions with ``forward derivatives'' only.
 \item The category $\JoinSLat$ has join-semilattices with bottom as
-    objects and finite join preserving functions as
-    morphisms.
+  objects and monotone finite join preserving functions as
+  morphisms. The category $\Fam(\JoinSLat^\op)$ provides a model of
+  total functions with ``backwards derivatives'' only.
 \end{enumerate}
-\end{definition}
 
-We can get an analogous result to \propref{embed-manifolds} for L-posets:
+We can get an analogous result to \propref{embed-manifolds} for
+L-posets and stable maps:
 
@@
 The categories $\Fam(\cat{C})$ are the free coproduct completions of
-categories $\cat{C}$, so they are guaranteed to have all coproducts:
+categories $\cat{C}$, so they have all coproducts:
 
@@
 a wide range of other useful semantic models of datatypes in
-$\Fam(\cat{C})$. For example lists can be constructed as the coproduct
+$\Fam(\cat{C})$. For example, lists can be constructed as a coproduct
 \begin{displaymath}
@@
 Our category of interest, $\Fam(\LatGal)$ has coproducts and finite
-products, because $\LatGal$ has products. As we shall see below, these
-are also coproducts, which is essential to obtaining cartesian
-closure.
+products, because $\LatGal$ has products. Similarly for
+$\Fam(\MeetSLat)$. As we shall see below, the products in $\LatGal$
+(and $\MeetSLat$ and $\JoinSLat)$ are also coproducts, which is
+essential to obtaining cartesian closure.
 
@@
 
-For cartesian closure of categories of families, we rely on the
-following theorem of \citet{nunes2023}, specialised from their setting
-with the general Grothendieck construction to $\Fam(\cat{C})$. This
-relies on the definition of \emph{biproducts}, which we discuss below
-in \secref{biproducts}.
+For cartesian closure of the categories $\Fam(C)$ that we are
+interested in, we rely on the following theorem of \citet{nunes2023},
+specialised from their setting with the general Grothendieck
+construction to $\Fam(\cat{C})$. This relies on the definition of
+\emph{biproducts}, which we discuss below in \secref{biproducts}.
 
@@
   \label{thm:fam-closed}
-  \AGDA.  If $\cat{C}$ has biproducts and all products, then
-  $\Fam(\cat{C})$ is cartesian closed\footnote{More precisely, if
-    $\cat{C}$ has coproducts then we have a monoidal product on
-    $\Fam(\cat{C})$ which is closed by this construction. When these
-    coproducts are in fact biproducts, we get cartesian closure.}. On
-  objects, the internal hom can be given by:
+  \AGDA.  If $\cat{C}$ has biproducts (\defref{biproducts}) and all
+  products, then $\Fam(\cat{C})$ is cartesian closed\footnote{More
+    precisely, if $\cat{C}$ has coproducts then we have a monoidal
+    product on $\Fam(\cat{C})$ which is closed by this
+    construction. When these coproducts are in fact biproducts, we get
+    cartesian closure.}. On objects, the internal hom can be given by:
   \begin{displaymath}
@@
 
-Before we explore the hypotheses of this theorem, we give some
-intuition on how it relates to differentiation. The Set-component of
-$(X, \partial X) \to (Y, \partial Y)$ consists of exactly the
-morphisms of $\Fam(\cat{C})$, rephrased into a single object. When
-$\cat{C} = \FinVect$, these are functions with an associated linear
-map at every point, and when $\cat{C} = \LatGal$, these are functions
-with an associated Galois connection at every point. A tangent to a
-function is then defined to be a mapping from points in the domain to
-tangents in the codomain along the function.
+The $\Set$-component of $(X, \partial X) \to (Y, \partial Y)$ consists
+of exactly the morphisms of $\Fam(\cat{C})$, rephrased into a single
+object. When $\cat{C} = \FinVect$, these are functions with an
+associated linear map at every point, and when $\cat{C} = \LatGal$,
+these are functions with an associated Galois connection at every
+point. A tangent to a function is then defined to be a mapping from
+points in the domain to tangents in the codomain along the function.
 
-Unfortunately, neither of the categories $\LatGal$ or $\FinVect$
-satisfy the hypotheses of this theorem, because neither of them have
-infinite products. We will consider ways to rectify this below in
+The category $\MeetSLat$ satisfies the hypoetheses of
+\thmref{fam-closed}, so:
+\begin{corollary}
+  $\Fam(\MeetSLat)$ is cartesian closed and has all coproducts.
+\end{corollary}
+Unfortunately, neither of the $\LatGal$ or $\FinVect$ satisfy the
+hypotheses of this theorem, because neither of them have infinite
+products. We will consider ways to rectify this below in
 \secref{fixing-completeness}.
@@
   There is another construction of internal homs on $\Fam(\cat{C})$
-  arising from the use of fibrations for categorical logical relations
-  proofs, due to \citet[Corollary 4.12]{hermida99}. If we assume that
-  $\cat{C}$ is itself cartesian closed and has all products, then we
-  could construct an internal hom as:
+  arising from the use of fibrations for categorical logical
+  relations, due to \citet[Corollary 4.12]{hermida99}. If we assume
+  that $\cat{C}$ is itself cartesian closed and has all products, then
+  we could construct an internal hom as:
   \begin{displaymath}
@@
   cartesian closed, and there is no way of making them so without
-  losing the property of having biproducts. We will use Hermida's
-  construction in our definability proof in \secref{definability},
-  where we use a logical relations argument to show that every
-  morphism definable in the higher order language is also first-order
-  definable.
+  losing the property of being able to conjunct or add tangents, as we
+  shall see below. We will implicitly use Hermida's construction in
+  our definability proof in \secref{definability}, where we use a
+  logical relations argument to show that every morphism definable in
+  the higher order language is also first-order definable.
 \end{remark}
@@
 Loosely stated, biproducts are objects that are both products and
-coproducts. They can be defined in any category, as shown by
+coproducts. The concept can be defined in any category, as shown by
 \citet{karvonen20}, but for our purposes it will be more convenient to
@@
   monoid with $(+,0)$ and composition is bilinear:
-  \begin{salign*}
-    f \comp \zero = f = \zero \comp f
-  \end{salign*}
-  \begin{salign*}
-    (f + g) \comp h &= (f \comp h) + (g \comp h) \\
-    h \comp (f + g) &= (h \comp f) + (h \comp g)
-  \end{salign*}
+  \begin{displaymath}
+    f \comp \zero = 0 = \zero \comp f
+  \end{displaymath}
+  \begin{displaymath}
+    (f + g) \comp h = (f \comp h) + (g \comp h) \qquad
+    h \comp (f + g) = (h \comp f) + (h \comp g)
+  \end{displaymath}
 \end{definition}
@@
 \begin{definition}
+  \label{def:biproducts}
   In a $\CMon$-category a biproduct is an object $X \biprod Y$
@@
 \begin{example}
-  The following categories are $\CMon$-enriched and have finite
-  products, and hence biproducts:
+  The following are $\CMon$-enriched and have finite products, and
+  hence biproducts:
   \begin{enumerate}
@@
     respectively. Products are given by the cartesian product of the
-    underlying set.
-  \item The $\LatGal$ example splits into two separate categories
-    $\MeetSLat$ and $\JoinSLat$ of meet semilattices and join
-    semilattices respectively, with their respective homomorphisms as
-    the morphisms. As with $\LatGal$, these categories are
-    $\CMon$-enriched and have finite products, and so finite
-    biproducts.
+    underlying set and the one-element lattice for the
+    terminal/initial/zero object.
+  \item $\MeetSLat$ and $\JoinSLat$ are both $\CMon$-enriched and have
+    finite products for similar reasons to $\LatGal$.
   \end{enumerate}
@@
 
-The following proposition shows why the alternative construction of
-exponentials in $\Fam(C)$ described in \remref{hermida-exponentials}
-is not suitable:
-\begin{proposition}
-  A $\CMon$-category that is cartesian closed has trivial homsets:
-  $\cat{C}(X,Y) \cong 1$.
-\end{proposition}
-
-\begin{proof}
-  $\cat{C}(X, Y) \cong \cat{C}(0 + X,Y) \cong \cat{C}(0 \times X,Y)
-  \cong \cat{C}(0,X \to Y) \cong 1$.
-\end{proof}
+\begin{remark}
+  Categories with zero objects cannot be cartesian closed without
+  being trivial in the sense of having exactly one morphism between
+  every pair of objects because
+  $\cat{C}(X, Y) \cong \cat{C}(1 \times X,Y) \cong \cat{C}(0 \times
+  X,Y) \cong \cat{C}(0,X \to Y) \cong 1$. Consequently, we cannot
+  apply the alternative construction of exponentials described in
+  \remref{hermida-exponentials}.
+\end{remark}
 
@@
   The functor
-  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
+  $\HoEmbed : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
   on objects as
-  $F(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
+  $\HoEmbed(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
   X(x)))$. This functor is faithful and preserves coproducts and
@@
 
-With this embedding functor, we can see that the interpretation of
+With this embedding functor, we see in
+\propref{first-order-type-interp} that the interpretation of
 first-order types will be the same up to isomorphism in
@@
 we interpret the base types as objects in $\Fam(\LatGal)$. At
-higher-order, however, the meet-semilattice and join-semilattice
-sides of the interpretation will diverge, and it is no longer clear
-that the interpretation of programs using higher-order functions
-internally will result in Galois connections. In \secref{definability}
-we will see that every program with first-order type (even if it uses
+higher-order, however, the meet-semilattice and join-semilattice sides
+of the interpretation will diverge, and it is no longer clear that the
+interpretation of programs using higher-order functions internally
+will result in Galois connections. In \secref{definability} we will
+see that every program with first-order type (even if it uses
 higher-order functions internally) does in fact have an interpretation
 definable in $\Fam(\LatGal)$.
-
-% $\GLR(\Fam(\MeetSLat \times \JoinSLat^\op), F)$ fibred over
-% $\Fam(\MeetSLat \times \JoinSLat^\op)$ with a full and faithful functor $\Fam(\LatGal) \to \GLR(\Fam(\MeetSLat \times \JoinSLat^\op), F)$.
 
@@
 
-The above discussion has been somewhat abstract, so we now consider
-some examples in the category $\Fam(\LatGal)$ and how they relate to
-\GPS. We highlight how
+Our thesis is that $\Fam(\LatGal)$ is a suitable setting for
+interpreting first-order programs for \GPS. The above discussion has
+been somewhat abstract, so we now consider some examples in the
+category $\Fam(\LatGal)$ and how they relate to \GPS.
+
+\todo{Fix a notation for the forwards and backwards maps}
+
+Spelt out in full, it has as objects $(X, \partial X)$, all pairs of a
+set $X$ and and for every $x \in X$, a bounded lattice
+$\partial X(x)$, and as morphisms
+$(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of
+a function $f: X \to Y$ and for every $x \in X$, a Galois connection
+$\partial f(x): \partial X(x) \to \partial Y(f(x))$. The reindexing
+along $f$ in the composition
+$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial
+  g}{f} \comp \partial f)$ selects the appropriate lattice of
+approximations and Galois connection at each point, using the baseline
+(unapproximated) function $f$.
+
+\subsubsection{Unapproximated Functions}
+
+$\LatGal$ has a terminal (also zero) object $\mathbb{1}$, so there is
+a functor $\Disc : \Set \to \Fam(\LatGal)$ that maps a set $X$ to
+$(X, \lambda x. \mathbb{1})$ and functions $f$ to morphisms
+$(f, \lambda\_.\,\mathrm{id}_{\mathbb{1}})$. This functor preserves
+products and coproducts. Therefore, we can take any sets and functions
+of interest for modelling primitive types and operations of a
+programming language and embed, albeit without any interesting
+approximation information.
 
@@
 
-\bob{Move the discussion in \texttt{first-order.tex} to here, but
-  construct the lifting monad in a more elementary way}
+The operation of adding a new bottom element to a bounded lattice
+forms part of a monad $L$ on $\LatGal$. This monad extends to a
+(strong) monad $\Lift$ on $\Fam(\LatGal)$ with
+$\Lift(X, \partial X) = (X, L \circ \partial X)$. The monad $\Lift$
+does not affect the points of the original object, but adds a new
+minimum approximation. This monad allows us to construct morphisms in
+$\Fam(\LatGal)$ that track information about how values are used
+similar to the examples of cm and stable functions we saw in
+\secref{approx-as-tangents}.
+
+\newcommand{\Bool}{\mathrm{Bool}}
+
+Let $\Bool = \Disc(\{\mathsf{tt},\mathsf{ff}\})$ be the
+(unapproximated) embedding of the booleans and
+$\mathrm{or} : \Bool \times \Bool \to \Bool$ be the (unapproximated)
+boolean or function. The, using a Moggi-style let notation
+\cite{notions-of-computation} for morphisms constructed using the
+Monad structure of $\Lift$, we can reproduce the functions
+$\mathrm{strictOr}$ and $\mathrm{shortCircuitOr}$ functions from
+\exref{strict-short-circuit} (we also assume an if-then-else operation
+on booleans, definable from the fact that $\Fam(\LatGal)$ has
+coproducts and $\Disc$ preserves them). Both of these expressions
+define morphisms $\Lift(\Bool) \times \Lift(\Bool) \to \Lift(\Bool)$
+in $\Fam(\LatGal)$:
+\begin{displaymath}
+  \begin{array}{lcl}
+    \mathrm{strictOr}(x,y)&=&\mathrm{let}\,b_1 \Leftarrow x\,\mathrm{in}\,\mathrm{let}\,b_2 \Leftarrow y\,\mathrm{in}\,\eta(\mathrm{or}(b_1,b_2)) \\
+    \mathrm{shortCircuitOr}(x,y)&=&\mathrm{let}\,b_1 \Leftarrow x\,\mathrm{in}\,\mathrm{if}\,b_1\,\mathrm{then}\,\eta(\mathsf{tt})\,\mathrm{else}\,y
+  \end{array}
+\end{displaymath}
+Examining the morphisms so defined in $\Fam(\LatGal)$, we can see
+that, in the $\Set$ component, they are both exactly the normal
+boolean-or operation. However, they have different approximation
+behaviour, reflecting the different ways that they examine their
+inputs. Let us write $\top,\bot$ for the elements of the approximation
+lattice at each point of $\Lift(\Bool)$, then applying the reverse
+derivative at the point $(\mathsf{tt},\mathsf{tt})$ to the tangent
+$\top$ reveals which of the inputs contributed to the output for each
+function:
+\begin{displaymath}
+  \begin{array}{lcl}
+    (\partial \mathrm{strictOr})_r(\mathsf{tt},\mathsf{tt})(\top) &=& (\top, \top)  \\
+    (\partial \mathrm{shortCircuitOr})_r(\mathsf{tt},\mathsf{tt})(\top) &=& (\top, \bot)
+  \end{array}
+\end{displaymath}
+In comparison to the categories $\CM$ and $\Lposet$ from
+\secref{approx-as-tangents}, we have retained the usage information in
+the forward and reverse tangents, but we also accurately model
+totality of the functions. That is, the constantly $\bot$ function is
+also present in both $\CM$ and $\Lposet$, but is not expressible in
+$\Fam(\LatGal)$.
+
+An analogue of the $\mathrm{parallelOr}$ function from
+\exref{parallel-or} is not definable in $\Fam(\LatGal)$. We would have
+to have
+$(\partial \mathrm{parallelOr})_f(\mathsf{tt},\mathsf{tt})(\top,\bot)
+= (\partial \mathrm{parallelOr})_f(\mathsf{tt},\mathsf{tt})(\bot,\top)
+= \top$ to reflect the desired property that either of the inputs
+being $\mathsf{tt}$ is enough to determine the output. We also must
+have
+$(\partial \mathrm{parallelOr})_f(\mathsf{tt},\mathsf{tt})(\bot,\bot)
+= \bot$, to reflect the fact that we will get no information in the
+output if we required that neither of the inputs is examined. However,
+this means that
+$(\partial \mathrm{parallelOr})_f(\mathsf{tt},\mathsf{tt})$ will not
+presere meets because $(\top,\bot) \sqcap (\bot,\top) = (\bot,\bot)$
+but $\top \neq \bot$.
+
+An analogue of the $\mathrm{gustave}$ function from
+\exref{parallel-or} is definable in $\Fam(\LatGal)$, but not using the
+lifting monad structure as we could for $\mathrm{strictOr}$ and
+$\mathrm{shortCircuitOr}$.
+
+\bob{We don't get a theorem like ``if we didn't input $i$ to obtain
+  the output, then we could replace it with any other value.'' How
+  would we extend $\Fam(\LatGal)$ to do this?}
 
 \subsubsection{An Approximation Object and the Tagged Data Monad}
-
-\bob{flesh this out}
 

### 2025-07-10 main/language.tex
Some progress..

@@
 
-Models of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object; such a
-model $M$ assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each
-primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}:
-\sem{\rho_1}_{M} \times \ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$.
+A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object, and
+assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each primitive
+operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
+\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in $\cat{C}$ and
+consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
+\List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
+Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
 
-To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ plus a model $M$ of
-$\Sigma$ in $\Sem$, with the following additional structure:
-\begin{enumerate}
-\item a strong monad $(\Lift, \eta, \mu)$ with strength $\mathsf{st}_{X,Y}: X \times \Lift(Y) \to \Lift(X
-\times Y)$
-\item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
-$\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to Y$ and
-$f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
-\end{enumerate}
+To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ with a strong monad
+$(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
+in $\Sem$.
 

### 2025-07-10 main/language.tex
WIP.

@@
 operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
-\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in $\cat{C}$ and
-consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
+\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in such a category,
+and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
 \List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
-Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
+Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$,
+satisfying the usual laws for lists.
 
@@
 $(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
-in $\Sem$.
-
-\figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects of $\Sem$.
-\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms $\sem{\Gamma}
-\to \sem{\tau}$.
+in $\Sem$. \figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects
+of $\Sem$. \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
+$\sem{\Gamma} \to \sem{\tau}$.
 

### 2025-07-10 main/models.tex
[ tex ] discuss a possible refinement of the model

--- 2025-07-10 main/models.tex
@@
 
-\bob{We don't get a theorem like ``if we didn't input $i$ to obtain
-  the output, then we could replace it with any other value.'' How
-  would we extend $\Fam(\LatGal)$ to do this?}
+\begin{remark}
+  \label{rem:further-structure}
+  These examples highlight a potential criticism of $\Fam(\LatGal)$ as
+  a category for modelling \GPS. For $\mathrm{shortCircuitOr}$, we had
+  $(\partial \mathrm{shortCircuitOr})_r(\mathsf{tt},\mathsf{tt})(\top)
+  = (\top, \bot)$, indicating that the second argument was not needed
+  for computing the output. However, there is no way, in
+  $\Fam(\LatGal)$, of turning this into a rigorous statement that the
+  $\Set$-component of this morphism does not actually depend on its
+  second argument. We conjecture that this can be rectified by
+  requiring some kind of additional structure on each object
+  $(X, \partial X)$ consisting of a map
+  $\Pi_{x : X}.\partial X(x) \to \mathcal{P}(X)$, where $\mathcal{P}$
+  is the powerset, which identifies for each $x$ which elements are
+  indistinguishable from $x$ at this level of approximation. One would
+  also presumably have to require additional conditions for this to
+  act like an equivalence relation, and to respect the lattice
+  structure\footnote{This additional structure is reminiscent of the
+    additional structure on \emph{directed containers} defined by
+    \citet{ahman-uustalu}. They require a map
+    $\Pi_{x : X}.\partial X(x) \to X$ picking out a specific $X$
+    ``jumped to'' by some change $\delta x : \partial X(x)$ at $x$. In
+    our proposed setup, we follow the approximation methodology of
+    \GPS by having a \emph{set} of things that could be taken instead
+    of the original $x$.}.
+\end{remark}
 

### 2025-07-10 main/language.tex
Bit of progress.

@@
 
-\paragraph{Base types and operations}
-Given an interpretation of base types and operations in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times
-\JoinSLat^\op)$ can implement $\Sem$. It is bicartesian closed (\corref{mslat-jslat-bcc}) and can model
-$\List(X)$ as an infinite coproduct (\secref{models-of-total-gps:coproducts-and-products}). An interpretation
-of base types and operations can be obtained from their interpretation in $\Fam(\LatGal)$ via the embedding
-...
-\roly{$\Lift$ monad.}
+Given a model $M$ of $\Sigma$ in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times \JoinSLat^\op)$ can
+implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\List(X)$ as an
+infinite coproduct (\secref{models-of-total-gps:coproducts-and-products}). A model $H(M)$ of $\Sigma$ can be
+obtained via the embedding $\HoEmbed: \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ given in
+\propref{ho-embedding}, because .... Finally $\LatGal$ has a lifting monad which is preserved by the 2-functor
+$\Fam: \Cat \to \Cat$ (\secref{first-order:lifting}). The correctness of this interpretation is considered in
+the next section.

### 2025-07-10 main/language.tex
Get a bit clear on how we get the model of Sigma.

@@
 obtained via the embedding $\HoEmbed: \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ given in
-\propref{ho-embedding}, because .... Finally $\LatGal$ has a lifting monad which is preserved by the 2-functor
-$\Fam: \Cat \to \Cat$ (\secref{first-order:lifting}). The correctness of this interpretation is considered in
-the next section.
+\propref{ho-embedding}, because the functor $\Fam: \Cat \to \Cat$ preserves finite products and so does the
+embedding. Finally $\LatGal$ has a lifting monad which is also preserved by $\Fam$
+(\secref{first-order:lifting}). The correctness of this interpretation is considered in the next section.

### 2025-07-10 main/language.tex
Rough pass.

--- 2025-07-10 main/language.tex
@@
 
-A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object, and
-assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each primitive
-operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
-\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in such a category,
-and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
-\List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
-Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$,
-satisfying the usual laws for lists.
+A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products, and assigns to each base
+type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimOp}$ in $\cat{C}$, and to each primitive operation $\phi
+\in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots
+\times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$. A model of $\tyList$ can also be given in such a
+category, and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms
+$\nil: 1 \to \List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms
+$f_\nil: Z \to Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times
+\List(X) \to Y$, with $\nil$, $\cons$ and $\fold$ satisfying the usual laws for lists.
 
-To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ with a strong monad
-$(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
-in $\Sem$. \figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects
-of $\Sem$. \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
-$\sem{\Gamma} \to \sem{\tau}$.
+The semantics for the full language is then given in terms of an arbitrary bicartesian closed category $\Sem$,
+models of $\Sigma$ and $\tyList$ in $\Sem$, and a strong monad $(\Lift, \eta, \mu)$ with strength
+$\mathsf{st}$ for interpreting $\tyLift$. \figref{semantics:types} gives the interpretations of types $\tau$
+and contexts $\Gamma$ as objects; \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t:
+\tau$ as morphisms $\sem{\Gamma} \to \sem{\tau}$.
 
@@
 Given a model $M$ of $\Sigma$ in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times \JoinSLat^\op)$ can
-implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\List(X)$ as an
-infinite coproduct (\secref{models-of-total-gps:coproducts-and-products}). A model $H(M)$ of $\Sigma$ can be
-obtained via the embedding $\HoEmbed: \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ given in
-\propref{ho-embedding}, because .... Finally $\LatGal$ has a lifting monad which is preserved by the 2-functor
-$\Fam: \Cat \to \Cat$ (\secref{first-order:lifting}). The correctness of this interpretation is considered in
-the next section.
+implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\tyList$ using infinite
+coproducts (\secref{models-of-total-gps:coproducts-and-products}). The embedding $\HoEmbed: \Fam(\LatGal) \to
+\Fam(\MeetSLat \times \JoinSLat^\op)$ from \propref{ho-embedding} preserves finite products, providing a model
+$H(M)$ of $\Sigma$. Finally the lifting monad in $\Fam(\LatGal)$ (\secref{first-order:lifting}) is also
+preserved by $H$. The correctness of this interpretation is considered in the next section.

### 2025-07-10 main/language.tex
Improvements, not sure H(M) technically makes sense though.

@@
 To model Galois slicing semantically for higher-order programs, we define a simple total functional
-programming language over a signature of base types and primitive operations, extending the simply-typed
-lambda calculus.
+programming language, extending the simply-typed lambda calculus. The language is parameterised by a signature
+$\Sigma = (\PrimTy, \PrimOp)$ consisting of a set $\PrimTy$ of base types $\rho$ and a family of sets
+$\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$ of arity $n$ over those base types.
 
@@
 
-The language is parameterised by a signature $\Sigma = (\PrimTy, \PrimOp)$ consisting of a set $\PrimTy$ of
-base types $\rho$ and a family of sets $\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$ of
-arity $n$ over those base types. The syntax is defined in \figref{syntax}. Types includes base types $\rho$
-drawn from $\PrimTy$, along with standard type formers for sums, products, functions and lists, plus a type
-constructor $\tyLift$ to allow approximation points to be added explicitly, as discussed in
-\secref{models-of-total-gps}. Terms include variables, the usual introduction and elimination forms, a monadic
-return and bind for lifted terms, and primitive operations $\phi$.
+The syntax is defined in \figref{syntax}. Types includes base types $\rho$ drawn from $\PrimTy$, along with
+standard type formers for sums, products, functions and lists, plus a type constructor $\tyLift$ to allow
+approximation points to be added explicitly, as discussed in \secref{models-of-total-gps}. Terms include
+variables, the usual introduction and elimination forms, a monadic return and bind for lifted terms, and
+primitive operations $\phi$.
 
@@
 
-A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products, and assigns to each base
-type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimOp}$ in $\cat{C}$, and to each primitive operation $\phi
-\in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots
+A model of $\Sigma$ can be given in any category $\cat{C}$ with finite products, and assigns to each base type
+$\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimOp}$ in $\cat{C}$, and to each primitive operation $\phi \in
+\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots
 \times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$. A model of $\tyList$ can also be given in such a
@@
 
-The semantics for the full language is then given in terms of an arbitrary bicartesian closed category $\Sem$,
-models of $\Sigma$ and $\tyList$ in $\Sem$, and a strong monad $(\Lift, \eta, \mu)$ with strength
+The semantics for the full language is then given in terms of an arbitrary bicartesian closed category $\Sem$
+providing models of $\Sigma$ and $\tyList$, as well as a strong monad $(\Lift, \eta, \mu)$ with strength
 $\mathsf{st}$ for interpreting $\tyLift$. \figref{semantics:types} gives the interpretations of types $\tau$

### 2025-07-10 main/models.tex
[ tex ] more discussion of semantics GPS in Fam(LatGal)

@@
 
-\todo{Fix a notation for the forwards and backwards maps}
-
-Spelt out in full, it has as objects $(X, \partial X)$, all pairs of a
-set $X$ and and for every $x \in X$, a bounded lattice
-$\partial X(x)$, and as morphisms
-$(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of
-a function $f: X \to Y$ and for every $x \in X$, a Galois connection
-$\partial f(x): \partial X(x) \to \partial Y(f(x))$. The reindexing
-along $f$ in the composition
-$(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial
-  g}{f} \comp \partial f)$ selects the appropriate lattice of
-approximations and Galois connection at each point, using the baseline
-(unapproximated) function $f$.
+Spelt out in full, $\Fam(\LatGal)$ has as objects $(X, \partial X)$,
+all pairs of a set $X$ and and for every $x \in X$, a bounded lattice
+$\partial X(x)$. Morphisms $(X, \partial X) \to (Y, \partial Y)$, are
+triples $(f, \partial f_f, \partial f_r)$ of functions $f : X \to Y$
+and families of monotone maps
+$\partial f_f : \Pi_{x : X}.\partial X(x) \multimap \partial Y(f\,x)$
+(``forward derivative'') and
+$\partial f_r : \Pi_{x : X}. \partial Y(f\,x) \multimap \partial X(x)$
+(``reverse derivative''), such that for all $x$,
+$\partial f_r(x) \dashv \partial f_f(x)$.
 
@@
   also presumably have to require additional conditions for this to
-  act like an equivalence relation, and to respect the lattice
-  structure\footnote{This additional structure is reminiscent of the
+  respect the lattice structure and be preserved by
+  morphisms\footnote{This additional structure is reminiscent of the
     additional structure on \emph{directed containers} defined by
@@
     ``jumped to'' by some change $\delta x : \partial X(x)$ at $x$. In
-    our proposed setup, we follow the approximation methodology of
-    \GPS by having a \emph{set} of things that could be taken instead
-    of the original $x$.}.
+    our proposed setup, we follow the approximation theme of \GPS by
+    having a \emph{set} of things that could be taken instead of the
+    original $x$. Objects of $\Fam(\LatGal)$ are ``directed'' in this
+    sense because the approximation map can pick out the element of
+    the original poset that was approximating $x$.}.
 \end{remark}
 
-\subsubsection{An Approximation Object and the Tagged Data Monad}
+The $\Lift$ monad provides a controllable way of adding
+presence/absence approximation points to composite data, and its monad
+structure makes explicit in the program structure exactly how such
+approximations are propagated through computations. The fact that
+there are different choices of this kind of approximation tracking
+provides freedom to the language implementor to decide what
+information is worth tracking. The Galois slicing implementations
+discussed in \citet{perera12a} and \citet{ricciotti17}, for example,
+bake-in an approximation point at every composite type constructor. We
+will see in \secref{cbn-translation} that this choice can be
+systematised in our setting by considering a monadic CBN translation
+to uniformly add $\Lift$ approximation points to composite data types.
 
-The object $\mathbb{A} = (1, \lambda \_.\, \{\top,\bot\})$ represents
-a basic object for measuring approximations in terms of presence and
-absence. (If all we care about is presence or absence in our
-approximations, we could see this object as the analogue of the real
-numbers for normal differentiation.)
+\subsubsection{An Approximation Object and the Tagging Monad}
 
-This object is a commutative monoid in $\Fam(\LatGal)$, and so we can
-construct the writer monad $T(X) = \mathbb{A} \times X$ in
-$\Fam(\LatGal)$.
+The $\Lift$ monad provides a way of tagging first-order data with
+presence and absence information. The object $\Lift(1)$, the lifting
+of the terminal object in $\Fam(\LatGal)$, yields an object that
+consists purely of presence/absence information:
+$\mathbb{A} = (1, \lambda\_.\,\{\top,\bot\})$. This object is the
+carrier of a commutative monoid in $\Fam(\LatGal)$, where the forward
+maps take the meet (both the inputs are required for the output to be
+present) and the backwards maps duplicate.% \bob{Is it also a comonoid?}
 
-On objects $X$ with trivial approximations, we have $\Lift(X) \cong T(X)$,
-but for more structured objects we get usage information at every level.
+Since $\mathbb{A}$ is a monoid, we can define the writer monad
+$T(X) = \mathbb{A} \times X$ in $\Fam(\LatGal)$ which ``tags'' $X$
+values with approximation information. This is similar to the $\Lift$
+monad in that it adds approximation information to an object. On
+discrete objects it agrees with the lifting:
+$T(\Disc(X)) \cong L(\Disc(X))$. However, on composite data, the two
+monads give different approximation lattices. Let $A$ and $B$ be sets.
+Then $T(T(\Disc(A)) \times T(T(\Disc(B))))$ has approximation lattices
+at $(a,b)$ that are always isomorphic to $\{\top,\bot\}^3$. The
+corresponding $\Lift(\Lift(\Disc(A)) \times \Lift(\Disc(B)))$ object's
+approximation lattice at $(a,b)$ is always isomorphic to
+$(\{\top,\bot\}^2)_\bot$.
+In terms of usage tracking, the object using the $\Lift$ monad is more
+appealing. The approximation lattice resulting from the use of the $T$
+monad contains apparently nonsensical elements corresponding to
+``using'' one or other components of the product \emph{without using
+  the product itself}. (This arises from the fact that we can project
+the $X$ out of $\mathbb{A} \times X$ without touching the
+$\mathbb{A}$.)
 
-This is a bit like using the writer monad and Moggi's CBN translation
-as a basic way of doing provenance tracking, but here we can track
-backwards and forwards.
+This example shows that we have to be careful about how we choose the
+interpretation of approximable sets in $\Fam(\LatGal)$, and again
+highlights the point we made in \remref{further-structure} that
+perhaps $\Fam(\LatGal)$ does not have quite enough structure to
+determine ``sensible'' approximation information. On the other hand,
+the use of the $T$ monad does have the advantage that the
+approximation lattices built from discrete sets, products, and
+coproducts, are always Boolean lattices, meaning that we can take
+complements of usage information. The ability to take boolean
+complements of approximations has been used by \citet{perera22} to
+compute \emph{related} outputs, as we discuss in
+\secref{related-work:galois-slicing}.
 
@@
 
-\bob{Describe how \exref{intervals-and-maxima-elements} fits into
-  $\Fam(\LatGal)$}
+So far, the approximation lattices we have looked at in
+$\Fam(\LatGal)$ have only consisted of those constructed from finite
+products and lifting, and only track binary usage/non-usage
+information. \exref{intervals-and-maxima-elements} shows how we can go
+beyond this to get more ``quantitative'' approximation
+information. Let the object of reals with interval approximations in
+$\Fam(\LatGal)$ be
+$(\mathbb{R}, \lambda x.\,\{[l,u] \mid l \leq x \leq u\} \cup
+\{\bot\})$ where the lattices of intervals are reverse ordered by
+inclusion with $\bot$ at the bottom. Then, following the examples in
+\exref{intervals-and-maxima-elements} and \exref{stable-functions}, we
+can define addition, negation, and scaling by $r \geq 0$:
+\begin{displaymath}
+  \begin{array}{lcl}
+    \mathrm{add}&=&(\begin{array}[t]{@{}l}
+      \lambda (x_1,x_2).\,x_1+x_2, \\
+      \lambda (x_1,x_2)\,([l_1,u_1],[l_2,u_2]).\,[(l_1+x_2) \sqcap (l_2+x_1),(u_1+x_2) \sqcup (u_2+x_1)] \\
+      \lambda (x_1,x_2)\,[l,u].\,([l-x_2,u-x_2],[l-x_1,u-x_1]))
+    \end{array}\\
+    \mathrm{neg}&=&(\lambda x.\,-x, \lambda x\,[l,u].\,[-u,-l], \lambda x\,[l,u].\,[-u,-l]) \\
+    \mathrm{scale}(r)&=&(\lambda x.\,rx, \lambda x\,[l,u].\,[rl,ru], \lambda x\,[l,u].\,\mathrm{if}\,r=0\,\mathrm{then}\,\bot\,\mathrm{else}\,[\frac{l}{r},\frac{u}{r}])
+  \end{array}
+\end{displaymath}
+Scaling by negative numbers is also possible with swapping of lower
+and upper bounds, as is multiplication.
 

### 2025-07-10 main/models.tex
Tweak.

--- 2025-07-10 main/models.tex
@@
-\section{Models of \GPS for a Total Language}
+\section{Models of Galois Slicing for a Total Language} % \GPS doesn't use title caps
 \label{sec:models-of-total-gps}
@@
 construction, has been previously used by V{\'a}k{\'a}r and
-collaborators \bob{CITE} to model automatic differentiation for
+collaborators \cite{vakar22} to model automatic differentiation for
 higher-order programs on the reals. We reuse some of their results,

### 2025-07-10 main/conclusion.tex
Some blurb on source transformation approach.

@@
 
-Would be better to have an implementation that worked by translation,
-instead of interpreting directly as Agda functions. See work by
-Smeding and Vákár on translations. To formalise this, we'd need
-to... See also $\Fam(\PSh_{\CMon}(\namedcat{Syn}))$?
+% To formalise this, we'd need to... See also $\Fam(\PSh_{\CMon}(\namedcat{Syn}))$?
+
+An interesting implementation technique, as an alternative the denotational approach presented here or the
+trace-based approaches used in earlier \GPS implementations, would be to develop a source-to-source
+transformation, in direct analogy with the CHAD approach to automatic differentiation
+\cite{vakar22,nunes2023}. In their approach, forward and reverse-mode AD are implemented as compositional
+transformations on source code, guided by a universal property: they arise as the unique structure-preserving
+functors from the source language to a suitably structured target language formalised as a Grothendieck
+construction. Adapting this to Galois slicing would allow slicing to ``compiled in'', avoiding the need for a
+custom interpreter and presenting opportunities for optimisation.

### 2025-07-10 main/conclusion.tex
Expand XAI bit.

@@
 
-\paragraph{Integrating XAI techniques with Galois slicing.} \roly{Sketch what this might look like.}
-Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping (Grad-CAM)~\cite{selvaraju20}
-use reverse-mode AD selectively on specific activations, calculating \emph{heat maps} that highlight input
-regions contributing to a given classification decision. It might be possible to compose these techniques with
-\GPS whilst still preserving useful round-tripping properties; if so this could be a basis for end-to-end
-explainability in systems that combine neural networks with traditional symbolic computation.
+\paragraph{Quantitative Galois slicng and XAI.}  Explainable AI (XAI) techniques like Gradient-weighted Class
+Activation Mapping (Grad-CAM)~\cite{selvaraju20} use reverse-mode AD selectively to calculate \emph{heat maps}
+(or \emph{saliency maps}) that highlight input regions contributing to a given classification or other
+outcome. We would like to investigate quantitative approximation structures where $\top$ represents the
+original input image and lower elements represent ``slices'' of the image where individual pixels have been
+ablated to some degree (partly removed or blurred). This might allow for composing some of these techniques
+with Galois slicing, for use in hybrid systems such as physical simulations with ML-based parameterisations.
 

### 2025-07-10 main/conclusion.tex
Tweaks.

@@
 
-\paragraph{Quantitative Galois slicng and XAI.}  Explainable AI (XAI) techniques like Gradient-weighted Class
+\paragraph{Quantitative slicing and XAI}  Explainable AI (XAI) techniques like Gradient-weighted Class
 Activation Mapping (Grad-CAM)~\cite{selvaraju20} use reverse-mode AD selectively to calculate \emph{heat maps}
@@
 
-\paragraph{Alternative Higher-order Models}
+\paragraph{Alternative Higher-Order Models}
 
@@
 
-\paragraph{Implementation}
+make\paragraph{General Inductive and Coinductive Types}
+
+\paragraph{Source-To-Source Translation Techniques}
 
@@
 
-An interesting implementation technique, as an alternative the denotational approach presented here or the
-trace-based approaches used in earlier \GPS implementations, would be to develop a source-to-source
-transformation, in direct analogy with the CHAD approach to automatic differentiation
-\cite{vakar22,nunes2023}. In their approach, forward and reverse-mode AD are implemented as compositional
-transformations on source code, guided by a universal property: they arise as the unique structure-preserving
-functors from the source language to a suitably structured target language formalised as a Grothendieck
-construction. Adapting this to Galois slicing would allow slicing to ``compiled in'', avoiding the need for a
-custom interpreter and presenting opportunities for optimisation.
+An interesting alternative to the denotational approach presented here, and to the trace-based approaches used
+in earlier \GPS implementations, would be to develop a source-to-source transformation, in direct analogy with
+the CHAD approach to automatic differentiation \cite{vakar22,nunes2023}. In their approach, forward and
+reverse-mode AD are implemented as compositional transformations on source code, guided by a universal
+property: they arise as the unique structure-preserving functors from the source language to a suitably
+structured target language formalised as a Grothendieck construction. Adapting this to Galois slicing would
+allow slicing to ``compiled in'', avoiding the need for a custom interpreter and potentially exposing
+opportunities for optimisation.

### 2025-07-10 main/conclusion.tex
Some blurb on inductive/coinductive types.

@@
 2.
-Conjecture that the category with $(X : \namedcat{Cpo}, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and galois connections is the right setting.
-Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains (\cite{laird07}).
+Conjecture that the category with $(X : \namedcat{Cpo}, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and Galois connections is the right setting.
+Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains \cite{laird07}.
 
-make\paragraph{General Inductive and Coinductive Types}
+\paragraph{General Inductive and Coinductive Types}
+
+For expedience, lists are the only recursive data type in our source language, so important future work is
+supporting general inductive and coinductive types. \citet{nunes2023} support automatic differentiation for
+datatypes defined as the least or greatest fixed points of $\mu\nu$-polynomial functors; we could potentially
+adopt a similar approach. Full inductive types would allow us to embed an interpreter for untyped lambda
+calculus; combined with the CBN monadic translation described in \secref{todo} which uniformly inserts
+approximation points, we should be able to obtain the program slicing behaviour of earlier \GPS work ``for
+free''. Coinductive types (e.g.~streams) present additional challenges, especially for defining
+join-preserving backward maps, but also open the door to slicing (finite prefixes of) infinite data sources,
+with some likely relationship to the problem of dealing with partial or non-terminating computations.
 

### 2025-07-10 main/conclusion.tex
Some blurb on inductive/coinductive types.

@@
 
-For expedience, lists are the only recursive data type in our source language, so important future work is
+Lists are the only recursive data type we provided in our source language, so important future work is
 supporting general inductive and coinductive types. \citet{nunes2023} support automatic differentiation for

### 2025-07-10 main/language.tex
Improvements, not sure H(M) technically makes sense though.
Pass over section 5.
[ tex ] write up the semantics and some examples
[ tex ] fix bib

@@
 The syntax is defined in \figref{syntax}. Types includes base types $\rho$ drawn from $\PrimTy$, along with
-standard type formers for sums, products, functions and lists, plus a type constructor $\tyLift$ to allow
-approximation points to be added explicitly, as discussed in \secref{models-of-total-gps}. Terms include
-variables, the usual introduction and elimination forms, a monadic return and bind for lifted terms, and
+standard type formers for sums, products, functions and lists% , plus a type constructor $\tyLift$ to allow
+% approximation points to be added explicitly, as discussed in \secref{models-of-total-gps}
+. Terms include
+variables, the usual introduction and elimination forms% , a monadic return and bind for lifted terms
+, and
 primitive operations $\phi$.
@@
 types, which we will consider in future work (\secref{conclusion}). Typing judgments for terms are standard
-and shown in \figref{typing}, with the usual rules for products, sums, functions, lists and lifting.
+and shown in \figref{typing}, with the usual rules for products, sums, functions, and lists.
 
@@
 
-A model of $\Sigma$ can be given in any category $\cat{C}$ with finite products, and assigns to each base type
-$\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimOp}$ in $\cat{C}$, and to each primitive operation $\phi \in
-\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots
-\times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$. A model of $\tyList$ can also be given in such a
-category, and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms
-$\nil: 1 \to \List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms
-$f_\nil: Z \to Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times
-\List(X) \to Y$, with $\nil$, $\cons$ and $\fold$ satisfying the usual laws for lists.
+An interpretation of a signature $\Sigma = (\PrimTy, \PrimOp)$ can be given in any category $\cat{C}$ with
+finite products, and assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimTy}$ in
+$\cat{C}$, and to each primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism
+$\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots \times \sem{\rho_n}_{\PrimTy} \to
+\sem{\rho}_{\PrimTy}$.
 
-The semantics for the full language is then given in terms of an arbitrary bicartesian closed category $\Sem$
-providing models of $\Sigma$ and $\tyList$, as well as a strong monad $(\Lift, \eta, \mu)$ with strength
-$\mathsf{st}$ for interpreting $\tyLift$. \figref{semantics:types} gives the interpretations of types $\tau$
-and contexts $\Gamma$ as objects; \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t:
-\tau$ as morphisms $\sem{\Gamma} \to \sem{\tau}$.
+Assuming that $\cat{C}$ is bicartesian closed and has a list object (\propref{lists}), then we can extend an
+interpretation of a signature $\Sigma$ to an interpretation of the whole language over
+$\Sigma$. \figref{semantics:types} and \figref{semantics:contexts} define the interpretation of types and
+contexts as objects of $\cat{C}$ respectively. Terms are interpreted as morphisms between the interpretations
+of the context and type, as defined in \figref{semantics:terms}. We have used the notations $\pi_i$ for
+projections, $\prodM{f}{g}$ for pairing, $\coprodM{f}{g}$ for (parameterised) copairing, $!_x$ for morphisms
+to the terminal object, and $\lambda$ and $\eval$ for currying and evaluation for exponentials.
 
-\subsubsection{Interpretation for higher-order \GPS}
+For the first-order definability result in \secref{definability}, we will need another interpretation
+$\sem{-}_{\mathit{fo}}$ of the first-order types (those constructed from primitive types, sums, unit and
+products) in any bicartesian category. Such interpretations are preserved by finite coproduct and coproduct
+preserving functors:
 
-Given a model $M$ of $\Sigma$ in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times \JoinSLat^\op)$ can
-implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\tyList$ using infinite
-coproducts (\secref{models-of-total-gps:coproducts-and-products}). The embedding $\HoEmbed: \Fam(\LatGal) \to
-\Fam(\MeetSLat \times \JoinSLat^\op)$ from \propref{ho-embedding} preserves finite products, providing a model
-$H(M)$ of $\Sigma$. Finally the lifting monad in $\Fam(\LatGal)$ (\secref{first-order:lifting}) is also
-preserved by $H$. The correctness of this interpretation is considered in the next section.
+\begin{lemma}\label{lem:first-order-agreement-types}
+  If $\cat{C}$ and $\cat{D}$ are bicartesian and bicartesian closed categories with interpretations of the
+  signature $\Sigma$, $F : \cat{C} \to \cat{D}$ is a bicartesian functor, and
+  $F(\sem{\rho}_{\PrimTy}) \cong \sem{\rho}_{\PrimTy}$ for all $\rho$, then for all first-order types $\tau$,
+  $F(\sem{\tau}_{\mathit{fo}}) \cong \sem{\tau}$, and similar for contexts only containing first-order types.
+\end{lemma}
+
+\subsubsection{Interpretation for Higher-Order \GPS}
+
+Given the above, we can now interpret the language in any of the bicartesian closed categories with list
+objects we constructed in \secref{models-of-total-gps}. Specifically, we assume that we have an interpretation
+of our chosen signature in $\Fam(\LatGal)$. Signatures are first-order, so it does not matter that
+$\Fam(\LatGal)$ is not closed. Any such interpretation can be transported to
+$\Fam(\MeetSLat \times \JoinSLat^\op)$ along the functor $H$ from \propref{ho-embedding} because it preserves
+finite products. We can then interpret the whole language in $\Fam(\MeetSLat \times \JoinSLat^\op)$.
+
+Interpreting a whole program $\Gamma \vdash t : \tau$ yield morphisms in
+$\Fam(\MeetSLat \times \JoinSLat^\op)$ which, as in \secref{semantic-gps}, are triples
+$(f, \partial f_f, \partial f_r)$ of the underlying function and the forward and backward approximation
+maps. However, unlike in $\Fam(\LatGal)$ are not guaranteed that the forward and backward maps even operate on
+the same lattices, let alone form a Galois connection. \lemref{first-order-agreement-types} guarantees that
+the lattices agree, but the fact that the pair form a Galois connection is less trivial. We will prove this
+property in \secref{definability}.
+
+% Given a model of $\Sigma$ in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times \JoinSLat^\op)$ can
+% implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\tyList$ using infinite
+% coproducts (\secref{models-of-total-gps:coproducts-and-products}). The embedding $\HoEmbed: \Fam(\LatGal) \to
+% \Fam(\MeetSLat \times \JoinSLat^\op)$ from \propref{ho-embedding} preserves finite products, and so can
+% transport each $\sem{\rho}_{\PrimTy}$ and $\sem{\phi}_{\Op}$ into the higher-order setting to provide a model
+% of $\Sigma$. Finally the lifting monad in $\Fam(\LatGal)$ (\secref{first-order:lifting}) is also preserved by
+% $H$. \roly{don't think we actually show this}
+
+% The correctness of this interpretation is considered in the next section.
+
+\subsection{Examples}
+
+Let the signature
+$\Sigma_{\mathsf{num}} = (\{\mathsf{num}\}, \{\mathrm{zero} : \tyUnit \to \mathsf{num}, \mathrm{add} :
+\mathsf{num} \times \mathsf{num} \to \mathsf{num}\})$. This signature suffices to write the simple query
+function in \exref{introduction-example}, where we interpret the $\mathrm{Label}$ type as the sum
+$\tyUnit \tySum \tyUnit$ and the labels $\mathsf{a}$ and $\mathsf{b}$ as $\tmInl{\tmUnit}$ and
+$\tmInr{\tmUnit}$. We consider several interpretations of $\Sigma_{\mathsf{num}}$ in $\Fam(\LatGal)$ and their
+behaviour on the query from \exref{introduction-example}.
+\begin{enumerate}
+\item If we take $\sem{\mathsf{num}}_{\PrimTy} = \Disc(\mathbb{R})$, with $\sem{\mathrm{zero}}_{\PrimOp}$ and
+  $\sem{\mathrm{add}}_{\PrimOp}$ the embeddings of the usual zero and addition functions via $\Disc$, then the
+  resulting interpretation contains no approximation information. The approximation lattices for all types are
+  always the trivial one-point lattice, and the forward and reverse approximation maps are necessarily the
+  identity.
+\item To get interesting approximation information, we take
+  $\sem{\mathsf{num}}_{\PrimTy} = \Lift(\Disc(\mathbb{R}))$, using the lifting monad from
+  \secref{models-of-total-gps:lifting-monad}, with $\sem{\mathrm{zero}}_{\PrimOp}$ and $\sem{\mathrm{add}}_{\PrimOp}$ defined from
+  the unlifted interpretations above and the monad structure. This interpretation recovers the behaviour given
+  in \exref{introduction-example}: running the reverse approximation map at the input $\mathsf{a}$ with the
+  top approximation maps indeed reveals that only the numbers in the rows tagged with that label in the input
+  are used, and likewise for $\mathsf{b}$.
+\item More quantitative approximation information can be obtained by using the interval approximation
+  interpretation from \exref{intervals-and-maxima-elements} and \secref{interval-approx}. We let
+  $\sem{\mathsf{num}}_{\PrimTy} = \mathbb{R}_{\mathit{intv}}$ and interpret addition using the morphism given
+  in \secref{interval-approx}.
+
+  Recall that $\mathrm{query}\,\mathsf{a}\,\mathit{db} = 1$, so in the reverse direction we can pick an interval containing $1$ to discover the largest (i.e.
+least in the order) intervals that will give rise to this output as \emph{independent} changes to the input.
+For example, if we pick $[-0.9,1]$ as the interval, then:
+  \begin{displaymath}
+    (\partial\mathrm{query}\,\mathsf{a}\,\mathsf{db})_r([0.9,1.1]) = [(\cdot,[-0.1,0.1]),(\cdot,\bot),(\cdot,[0.9,1.1])]
+  \end{displaymath}
+  Indicating that, to achieve an output within $[0.9,1.1]$ for this query, either the first $\mathsf{a}$ row
+  could be in $[-0.1,0.1]$ or the second one could be in $[0.9,1.1]$.
+\end{enumerate}
+
+\subsection{Systematic Insertion of Approximation via Moggi's CBN translation}
+\label{sec:cbn-translation}
+
+We can now carry out the systematic insertion of approximation points that we foreshadowed in
+\secref{models-of-total-gps:lifting-monad}, using \citet[\S 3.1]{notions-of-computation}'s monadic CBN translation. We use the $T$
+monad from \secref{tagging-monad} because it can be defined in terms of the language constructed we already
+have. This requires that we have a signature $\Sigma$ that includes a primitive type to be interpreted as the
+approximation object $\mathbb{A}$ and primitive operations to be interpreted as the monoid operations on this
+object.
+
+TODO: describe the syntactic CBN translation; show an example on the query example.

### 2025-07-10 main/models.tex
Moggi.
[ tex ] write up the semantics and some examples
[ tex ] fix bib

@@
 \item The category $\LatGal$ has bounded lattices as objects and
-  Galois connections as morphisms, with the right adjunct going in
-  the ``forward'' direction. The category $\Fam(\LatGal)$ is our
-  preferred model for total functions with Galois sliicing. We
-  explore some specific examples in this category in
-  \secref{semantic-gps}.
+  Galois connections as morphisms, with the right adjoint going in the
+  ``forward'' direction. The category $\Fam(\LatGal)$ is our preferred
+  model for total functions with Galois sliicing. We explore some
+  specific examples in this category in \secref{semantic-gps}.
 \item The category $\MeetSLat$ has meet-semilattices with top as
@@
 % $(\mathrm{Max}(M), \lambda x. \downset{x})$.
-Our current hypothesis is that $\Fam(\LatGal)$, where values and their
-approximations are separated by construction, is the natural model of
-semantic \GPS in a total setting. We now investigate some categorical
+Our working hypothesis is that $\Fam(\LatGal)$, where values and their
+approximations are separated by construction, is a natural model of
+semantic \GPS in a total setting, though we note some shortcomings in
+\remref{further-structure}. We now investigate some categorical
 properties of this category, with a view to modelling a higher-order
@@
 \subsubsection{Unapproximated Functions}
+\label{sec:unapproximated-functions}
 
@@
 \subsubsection{Lifting Monad}
+\label{sec:models-of-total-gps:lifting}
 
@@
 \subsubsection{An Approximation Object and the Tagging Monad}
+\label{sec:tagging-monad}
 
@@
 \subsubsection{Approximating Numbers by Intervals}
+\label{sec:interval-approx}
 
@@
 $\Fam(\LatGal)$ be
-$(\mathbb{R}, \lambda x.\,\{[l,u] \mid l \leq x \leq u\} \cup
-\{\bot\})$ where the lattices of intervals are reverse ordered by
-inclusion with $\bot$ at the bottom. Then, following the examples in
-\exref{intervals-and-maxima-elements} and \exref{stable-functions}, we
-can define addition, negation, and scaling by $r \geq 0$:
+$\mathbb{R}_{\mathit{intv}} = (\mathbb{R}, \lambda x.\,\{[l,u] \mid l
+\leq x \leq u\} \cup \{\bot\})$ where the lattices of intervals are
+reverse ordered by inclusion with $\bot$ at the bottom. Then,
+following the examples in \exref{intervals-and-maxima-elements} and
+\exref{stable-functions}, we can define addition, negation, and
+scaling by $r \geq 0$:
 \begin{displaymath}
@@
 \end{displaymath}
-Scaling by negative numbers is also possible with swapping of lower
-and upper bounds, as is multiplication.
+(we only define the forward and backward maps on intervals, their
+behaviour on $\bot$ is determined.)  Scaling by negative numbers is
+also possible with swapping of lower and upper bounds, as is
+multiplication.
 

### 2025-07-10 main/approx-as-tangents.tex
laird07, fix ref.
[ tex ] more discussion of semantics GPS in Fam(LatGal)
[ tex ] Fix up some references and citations

@@
 of their output. Therefore, we ignore completeness properties of the
-partially ordered sets we consider\footnote{\bob{we might want to bring this back when considering general recursive programs.
-Are bidomains \cite{berry79} useful?
-they have two orderings.
-See also Laird's Bistable Biorders.}}.
+partially ordered sets we consider.
 
@@
 input in a non-deterministic way as long as the result is
-deterministic. This non-determinism results in functions that whose
+deterministic. This non-determinism results in functions whose
 output cannot be assigned a unique minimal input that accounts for
@@
 
-\begin{example}
+\begin{example}\label{ex:stable-functions}
   \leavevmode\\[-3ex]
@@
 we have not considered above. We will extend this conjecture to our
-refined model of lattice approximated sets in \secref{model-summary}.
+refined model of lattice approximated sets in \secref{models:summary}.
 

### 2025-07-10 main/conclusion.tex
[ tex ] Fix up some references and citations

@@
 
+Are bidomains \cite{berry79} useful? they have two orderings. See also
+Laird's Bistable Biorders \cite{laird07}.
+
 \paragraph{General Inductive and Coinductive Types}
@@
 adopt a similar approach. Full inductive types would allow us to embed an interpreter for untyped lambda
-calculus; combined with the CBN monadic translation described in \secref{todo} which uniformly inserts
+calculus; combined with the CBN monadic translation described in \secref{cbn-translation} which uniformly inserts
 approximation points, we should be able to obtain the program slicing behaviour of earlier \GPS work ``for

### 2025-07-10 main/language.tex
[ tex ] Fix up some references and citations

@@
 
-Assuming that $\cat{C}$ is bicartesian closed and has a list object (\propref{lists}), then we can extend an
+Assuming that $\cat{C}$ is bicartesian closed and has a list object (\eqnref{lists}), then we can extend an
 interpretation of a signature $\Sigma$ to an interpretation of the whole language over
@@
   $\sem{\mathsf{num}}_{\PrimTy} = \Lift(\Disc(\mathbb{R}))$, using the lifting monad from
-  \secref{models-of-total-gps:lifting-monad}, with $\sem{\mathrm{zero}}_{\PrimOp}$ and $\sem{\mathrm{add}}_{\PrimOp}$ defined from
-  the unlifted interpretations above and the monad structure. This interpretation recovers the behaviour given
-  in \exref{introduction-example}: running the reverse approximation map at the input $\mathsf{a}$ with the
-  top approximation maps indeed reveals that only the numbers in the rows tagged with that label in the input
-  are used, and likewise for $\mathsf{b}$.
+  \secref{models-of-total-gps:lifting}, with $\sem{\mathrm{zero}}_{\PrimOp}$ and
+  $\sem{\mathrm{add}}_{\PrimOp}$ defined from the unlifted interpretations above and the monad structure. This
+  interpretation recovers the behaviour given in \exref{introduction-example}: running the reverse
+  approximation map at the input $\mathsf{a}$ with the top approximation maps indeed reveals that only the
+  numbers in the rows tagged with that label in the input are used, and likewise for $\mathsf{b}$.
 \item More quantitative approximation information can be obtained by using the interval approximation
@@
 
-  Recall that $\mathrm{query}\,\mathsf{a}\,\mathit{db} = 1$, so in the reverse direction we can pick an interval containing $1$ to discover the largest (i.e.
-least in the order) intervals that will give rise to this output as \emph{independent} changes to the input.
-For example, if we pick $[-0.9,1]$ as the interval, then:
+  Recall that $\mathrm{query}\,\mathsf{a}\,\mathit{db} = 1$, so in the reverse direction we can pick an
+  interval containing $1$ to discover the largest (i.e. least in the order) intervals that will give rise to
+  this output as \emph{independent} changes to the input. For example, if we pick
+  $[\frac{9}{10},\frac{11}{10}]$ as the interval, then:
   \begin{displaymath}
-    (\partial\mathrm{query}\,\mathsf{a}\,\mathsf{db})_r([0.9,1.1]) = [(\cdot,[-0.1,0.1]),(\cdot,\bot),(\cdot,[0.9,1.1])]
+    (\partial\mathrm{query}\,\mathsf{a}\,\mathsf{db})_r([\frac{9}{10},\frac{11}{10}]) = [(\cdot,[-\frac{1}{10},\frac{1}{10}]),(\cdot,\bot),(\cdot,[\frac{9}{10},\frac{11}{10}])]
   \end{displaymath}
-  Indicating that, to achieve an output within $[0.9,1.1]$ for this query, either the first $\mathsf{a}$ row
-  could be in $[-0.1,0.1]$ or the second one could be in $[0.9,1.1]$.
+  Indicating that, to achieve an output within $[\frac{9}{10},\frac{11}{10}]$ for this query, either the first
+  $\mathsf{a}$ row could be in $[-\frac{1}{10},\frac{1}{10}]$ or the second one could be in
+  $[\frac{9}{10},\frac{11}{10}]$.
 \end{enumerate}
@@
 We can now carry out the systematic insertion of approximation points that we foreshadowed in
-\secref{models-of-total-gps:lifting-monad}, using \citet[\S 3.1]{notions-of-computation}'s monadic CBN translation. We use the $T$
-monad from \secref{tagging-monad} because it can be defined in terms of the language constructed we already
-have. This requires that we have a signature $\Sigma$ that includes a primitive type to be interpreted as the
-approximation object $\mathbb{A}$ and primitive operations to be interpreted as the monoid operations on this
-object.
+\secref{models-of-total-gps:lifting}, using \citet[\S 3.1]{notions-of-computation}'s monadic CBN
+translation. We use the $T$ monad from \secref{tagging-monad} because it can be defined in terms of the
+language constructed we already have. This requires that we have a signature $\Sigma$ that includes a
+primitive type to be interpreted as the approximation object $\mathbb{A}$ and primitive operations to be
+interpreted as the monoid operations on this object.
 

### 2025-07-10 main/models.tex
[ tex ] Fix up some references and citations

@@
 $\Fam(\cat{C})$. For example, lists can be constructed as a coproduct
-\begin{displaymath}
+\begin{equation}
+  \label{eqn:lists}
   \List(X) = \coprod_{n \in \mathbb{N}} X^n
-\end{displaymath}
+\end{equation}
 where $X^0 = 1$ (the terminal object) and $X^{n+1} = X \times X^n$.
@@
 
-With this embedding functor, we see in
-\propref{first-order-type-interp} that the interpretation of
+With this embedding functor, we will see in
+\lemref{first-order-agreement-types} that the interpretation of
 first-order types will be the same up to isomorphism in
@@
     additional structure on \emph{directed containers} defined by
-    \citet{ahman-uustalu}. They require a map
+    \citet{ahman-chapman-uustalu2012}. They require a map
     $\Pi_{x : X}.\partial X(x) \to X$ picking out a specific $X$
@@
 \subsection{Summary}
+\label{sec:models:summary}
 

### 2025-07-10 main/conclusion.tex
[ tex ] some future work

--- 2025-07-10 main/conclusion.tex
@@
 \label{sec:conclusion}
+
+We have presented a semantic version of \GPS, inspired by connections to differentiable programming and
+automatic differentation, and shown that it can be used to interpret an expressive higher-order language
+suitable for writing simple queries and data manipulation. Our model elucidates some of the decisions
+implicitly taken in previous works on \GPS (\secref{cbn-translation}), and reveals new applications such as
+approximation by intervals \secref{interval-approx}. Our categorical approach admits a modular construction of
+our model, and the use of general theorems, such as Fiore and Simpson's \thmref{glr-definability}, to prove
+properties of the interpretation.
 
@@
 
-\paragraph{More Approximation Domains.} The intervals example is
-interesting, but is it useful, and are there more examples like this?
-One route might be to follow \cite{edalat-heckmann98} and explore
-whether metric spaces (which already provide a native notion of
-approximation) can be embedded in $\Fam(\LatGal)$.
+\paragraph{Refinement of the model} As we discussed in \remref{further-structure}, there are possible ways
+that the model we have proposed here could be refined to both remove nonsensical elements of the model, and to
+augment the model with enough power to prove additional properties such as ``functions are insenstive to
+unused inputs''. We are also planning to explore more examples of approximation along the lines of the
+intervals example in \secref{interval-approx}. One route might be to follow \citet{edalat-heckmann98}'s
+embedding of metric spaces in Scott domains and explore whether metric spaces (which already provide a native
+notion of approximation) can be embedded in $\Fam(\LatGal)$.
 
-\paragraph{Alternative Higher-Order Models}
-
-1. $\Fam(\PSh_{\CMon}(\LatGal))$
-
-2. Concrete sheaves over $\Fam(\LatGal)$ analogous to Diffeological spaces?
-
-\paragraph{Categorical Models of Differentation}
-
-\paragraph{Recursion and Partiality.}
-1. Despite using stable domain theory as our starting point, we have ignored
-
-2.
-Conjecture that the category with $(X : \namedcat{Cpo}, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and Galois connections is the right setting.
-Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains \cite{laird07}.
+\paragraph{Categorical Models of Differentation} We have already identified \conref{tangent-stable-fns} and
+\conref{tangent-fam} as potential connections to categorical models of differentation. It would also be
+interesting to discover any possible links to Differential Linear Logic
+\cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
+give an account of reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
+of \citet{vakar22}.
 
@@
 datatypes defined as the least or greatest fixed points of $\mu\nu$-polynomial functors; we could potentially
-adopt a similar approach. Full inductive types would allow us to embed an interpreter for untyped lambda
-calculus; combined with the CBN monadic translation described in \secref{todo} which uniformly inserts
+adopt a similar approach. Full inductive types would allow us to embed an interpreter for a small language;
+combined with the CBN monadic translation described in \secref{cbn-translation} which uniformly inserts
 approximation points, we should be able to obtain the program slicing behaviour of earlier \GPS work ``for
@@
 with some likely relationship to the problem of dealing with partial or non-terminating computations.
+
+\paragraph{Recursion and Partiality} This work has only examined the case for total programs. Even though we
+took stable domain theory as our starting point, we did not make any use of directed completeness or similar
+properties of domains. We expect that an account of recursion in an extension of the framework discussed so
+far would likely involve families of bounded lattices indexed by DCPOs, where the order of the DCPO would be
+reflected in embedding-projection relationships between the lattices. There are also other refinements of
+stable domain theory such as \citet{berry79}'s bidomains and \citet{laird07} bistable biorders that separate
+the extensional and stable orders on the same set, in a way that might be similar to
+\exref{intervals-and-maxima-elements}.
 

### 2025-07-10 main/models.tex
[ tex ] some future work

@@
 
-We have constructed the category
-$\Fam(\MeetSLat \times \JoinSLat^\op)$, which is a cartesian closed
-category with all coproducts, enough to interpret the total
-higher-order language we define in the next section with primitive
-types and operations defined in $\Fam(\LatGal)$. However, we are not
-guaranteed by construction that at first-order type, the
-interpretations are in fact Galois connections. We will rectify this
-in \secref{definability} using a logical relations construction.
+We have seen that the category $\Fam(\MeetSLat \times \JoinSLat^\op)$,
+which is a cartesian closed category with all coproducts, is enough to
+interpret the total higher-order language we define in the next
+section with primitive types and operations defined in
+$\Fam(\LatGal)$. However, we are not guaranteed by construction that
+at first-order type, the interpretations are in fact Galois
+connections. We will rectify this in \secref{definability} using a
+logical relations construction.
 
@@
 \begin{conjecture}
+  \label{con:tangent-fam}
   (1) The category $\Fam(\MeetSLat)$ is a Tangent category, with the
@@
 tangents, somewhat analogous to the situation with manifolds. Also
-note that, if this conjecture holds, then contrary to the $\Fam(\FinVect)$ representation of
-manifolds and differentiable maps, we do
-not throw away information about higher derivatives. It is retained in
-the order structure of the tangent fibres.
+note that, if this conjecture holds, then contrary to the
+$\Fam(\FinVect)$ representation of manifolds and differentiable maps,
+we do not throw away information about higher derivatives. It is
+retained in the order structure of the tangent fibres.

### 2025-07-10 main/related-work.tex
[ tex ] some future work

@@
 Cartesian differential categories \cite{cdcs}, which model differentiation over cartesian closed categories
-using a syntactic derivative operator. The motivation behind tangent categories is to \roly{Bob to say
-something here..}
+using a syntactic derivative operator. Reverse Tangent categories \cite{reverse-tangents} further axiomatise
+the existence of reverse derivatives. In \conref{tangent-stable-fns} and \conref{tangent-fam}, we have
+conjectured that the categories we have identified in this paper as models of \GPS are Tangent categories. As
+well as providing some substantiation to the claim that \GPS is a form of differentiation, we should be able
+to apply general results about Tangent categories to our setting. % BOB: such as?
 
-\subsection{Lens Categories}
+% \subsection{Differential Linear Logic}
 
-\cite{spivak19}
+% Differential Linear Logic \cite{dill} is an extension of Linear Logic with a derivative operator that acts
+% like
+
+% \subsection{Lens Categories}
+
+% \cite{spivak19}

### 2025-07-11 main/definability.tex
How Fam(MeetSLat×JoinSLatop) implements Sem.
WIP.
[ tex ] Finish correctness section

--- 2025-07-10 main/definability.tex
@@
-\section{Correctness of the Higher-order Interpretation}
+\section{Correctness of the Higher-Order Interpretation}
 \label{sec:definability}
 
-Introduction: what correctness properties we want; why we can't
-compute the left adjoints from the forward maps (no completeness).
+As we noted at the end of \secref{language:gps-interpretation}, our
+interpretation of the higher-order language is in the category
+$\Fam(\MeetSLat \times \JoinSLat^\op)$, so it is not a priori evident
+that we get a Galois connection from the interpretation of a program
+with first-order type (that may use higher-order functions
+internally). \citet{vakar22} and \citet{nunes2023} construct custom
+instances of categorical sconing arguments to prove correctness of
+their higher-order interpretation with respect to normal
+differentiation. Instead of doing this, we make use of a general
+\emph{syntax free} theorem due to \citet{fiore-simpson99}. The proof
+of this depends on the construction of a Grothendieck Logical Relation
+over the extensive topology on the category $\cat{C}$, but the
+statement of the theorem does not rely on this. We have formalised
+this proof in Agda (see \texttt{conservativity.agda} in the
+supplementary material\footnote{Our Agda development is complete
+  except for a proof that $\Fam(C)$ has extensive coproducts. We plan
+  to complete this part of the proof before any final
+  version. Moreover, this result does not yet apply to infinitary
+  coproducts, though we believe it is a relatively minor extension to
+  the proof to do so.}).
 
@@
 
+\begin{remark}
+  Compared to the exact result stated at the end of
+  \citet{fiore-simpson99}'s paper, we have made two modifications,
+  justified by our Agda proof. First, we generalise to the case where
+  $\cat{C}$ is not cartesian closed, and the functor $F$ does not
+  preserve exponentials. Examination of the proof reveals that if this
+  is the case, then $\hat{F}$ also preserves exponentials, but it is
+  not needed for the result stated. Second, Fiore and Simpson restrict
+  to the case when $\cat{C}$ is small to be able to construct
+  Grothendieck sheaves on this category. We use Agda's universe
+  hierarchy to simply construct ``large'' sheaves at the the
+  appropriate universe level.
+\end{remark}
+
 \begin{theorem}
@@
   \sem{\tau}_{\mathit{fo}})$ such that
-  $\sem{\Gamma \vdash M : \tau} = F(g)$.
+  $H(g) = (\cong) \circ \sem{\Gamma \vdash M : \tau} \circ (\cong)$,
+  where the isomorphisms arise from
+  \lemref{first-order-agreement-types}.
 \end{theorem}
-
-\begin{remark}
-  We have formalised \thmref{glr-definability} and
-  \thmref{language-definability} in Agda, except for the fact that
-  $\Fam(C)$ is extensive\footnote{For reviewers: we intend to fix this
-    before any final version.}. Our formalisation uses E-categories
-  (\cite{palmgren-wilander}) so we cannot state
-  \thmref{glr-definability} as given because the concept of strict
-  preservation does not make sense. Therefore, the formalised version
-  of \thmref{language-definability} is ``up to isomorphism''.
-
-  We believe that this result also extends to the list type, since it
-  is constructed from an infinite coproduct, but have not yet checked
-  all the details.
-\end{remark}
 
@@
   Instantiate \thmref{glr-definability} with
-  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$. By
-  \propref{XXX} we know that $F$ preserves finite products and
-  coproducts. $\Fam(\LatGal)$ has extensive coproducts because it is
-  the free coproduct completion. Thus we get the category $\GLR(F)$ in
-  which we can interpret the calculus with the property that this
+  $\HoEmbed : \Fam(\LatGal) \to \Fam(\MeetSLat \times
+  \JoinSLat^\op)$. By \propref{ho-embedding} we know that $F$
+  preserves finite products and coproducts. The fullness of
+  $\hat{\HoEmbed}$ means that for any morphism
+  $f : \hat{\HoEmbed}(\sem{\Gamma}_{\mathit{fo}}) \to
+  \hat{H}(\sem{\tau}_{\mathit{fo}})$ in $\GLR(\HoEmbed)$ there exists
+  a $g : \sem{\Gamma}_{\mathit{fo}} \to \sem{\tau}_{\mathit{fo}}$ in
+  $\Fam(\LatGal)$ such that $H(g) = f$. Since $\GLR(\HoEmbed)$ has
+  enough structure, we can interpret the term $M$ in it to get a
+  morphism
+  $\sem{\Gamma \vdash M : \tau}_{\GLR(\HoEmbed)} : \sem{\Gamma} \to
+  \sem{\tau}$ in $\GLR(\HoEmbed)$. Applying
+  \lemref{first-order-agreement-types} and the fact that the
+  strictness of $p$ means that
+  $p(\sem{\Gamma \vdash M : \tau}_{\GLR(\HoEmbed)}) = \sem{\Gamma
+    \vdash M : \tau}_{\Fam(\MeetSLat \times \JoinSLat^\op)}$ yields
+  the result.
 \end{proof}
 
-FIXME: another property we might want: The interpretations in
-$\Fam(\MeetSLat)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$ agree at
-first order.
+\begin{remark}
+  Note that even if we modified our base interpretation of semantic
+  \GPS along the lines of \remref{further-structure} to give a refined
+  version $\cat{G}$ of $\Fam(\LatGal)$, as long as there is a finite
+  product and coproduct preserving functor
+  $\cat{G} \to \Fam(\MeetSLat \times \JoinSLat^\op)$, then the
+  analogous result to \thmref{language-definability} still holds.
+\end{remark}
+
+We can also use \thmref{glr-definability} to show that the
+interpretation of the language in the category $\Set$ agrees with the
+higher-order interpretation in $\Fam(\MeetSLat \times \JoinSLat^\op)$
+on the underlying function at first order. This shows that the
+higher-order interpretation does what we expect in the underlying
+interpretation of terms, and that the approximation information does
+not interfere.
+
+\begin{theorem}
+  \label{thm:underlying-interp-equal} For all
+  $\Gamma \vdash M : \tau$, where $\Gamma$ and $\tau$ are first-order,
+  the interpretation $\sem{\Gamma \vdash M : \tau}_{\Set}$ in $\Set$
+  and the interpretation
+  $\sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
+    \JoinSLat^\op)}$ agree on their first components.
+\end{theorem}
+
+\begin{proof}
+  Instantiate \thmref{glr-definability} with the functor
+  $\langle \mathrm{Id} , \pi_1 \rangle : \Fam(\MeetSLat \times
+  \JoinSLat^\op) \to \Fam(\MeetSLat \times \JoinSLat^\op) \times \Set$
+  that is the identity in the first component and projects out the
+  underlying function in the second. We get, for each
+  $\Gamma \vdash M : \tau$, a $g$ such that
+  $g = \sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
+    \JoinSLat^\op)}$ and
+  $\pi_1(g) = \sem{\Gamma \vdash M : \tau}_\Set$. Therefore,
+  $\pi_1(\sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
+    \JoinSLat^\op)}) = \sem{\Gamma \vdash M : \tau}_\Set$.
+\end{proof}

### 2025-07-11 main/language.tex
[ tex ] Finish correctness section

@@
 \subsubsection{Interpretation for Higher-Order \GPS}
+\label{sec:language:gps-interpretation}
 
@@
 $\tmInr{\tmUnit}$. We consider several interpretations of $\Sigma_{\mathsf{num}}$ in $\Fam(\LatGal)$ and their
-behaviour on the query from \exref{introduction-example}.
+behaviour on the selection-and-sum query from \exref{introduction-example}.
 \begin{enumerate}
@@
 
+We have tested these examples on our fully formalised Agda implementation. See the file \texttt{example.agda}
+in the supplementary material for details.
+
 \subsection{Systematic Insertion of Approximation via Moggi's CBN translation}

### 2025-07-11 main.tex
Appendix after references.
Drop 'notes' from title.
[ paper ] some words introducing the main ideas
Start on syntax.
Bump 'categorical framework' to a new section.
Tweaks questions in intro, placeholder for abstract.
Minor reorg of intro.
Start on todo's for new section.
Rename file.
Related work section.
Start on typing rules.
[ tex ] rough attempt to develop the stablity<->differentiable analogy
Defining mu-polynomial functors.
Notation for strength.
Note on grad-CAM.
Add language section.
[ tex ] start on the model construction
Purging section 4.
Anonymise reference to GitHub repo.
Fix Vaka

@@
-\documentclass[acmsmall,screen]{acmart}
+\documentclass[acmsmall,screen,anonymous]{acmart}
 
+\usepackage{tabularx}
 \usepackage{bbm}
+\usepackage{bbold}
+\usepackage{mathpartir}
+\usepackage{subcaption}
 \usepackage{tikz-cd}
+\usepackage{xspace}
 
-\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
+%\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
 \settopmatter{printacmref=false}
+\citestyle{acmauthoryear}
+\raggedbottom
 
@@
 
-\title{Approximation as Differentiation notes}
+\title{Galois Slicing as Differentiation}
 
@@
 \begin{abstract}
-[Abstract]
+  \GPS is a technique for program slicing for provenance, developed by
+  Perera and collaborators. \GPS aims to explain program executions by
+  demonstrating how to track approximations of the input and output
+  forwards and backwards along a particular execution. In this paper,
+  we explore a analogy between \GPS and differentiable programming,
+  seeing the implementation of \GPS as a kind of automatic
+  differentiation. Using the CHAD approach to automatic differentation
+  due to V{\'a}k{\'a} and collaborators, we reformulate \GPS via a
+  categorical semantics. In doing so, we are able to explore
+  extensions of the \GPS idea to quantitative interval analysis, and
+  to clarify the implicit choices made in existing \GPS
+  instantiations.
 \end{abstract}
@@
 \input{main/introduction}
+\input{main/approx-as-tangents}
+\input{main/models}
+\input{main/language}
+\input{main/definability}
+\input{main/related-work}
+\input{main/conclusion}
 
-\pagebreak
-\appendix
-\input{appendix/notes}
-
-\bibliographystyle{plainnat}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{bib}
 
+% \pagebreak
+% \appendix
+% \input{appendix/notes}
+
 \end{document}

### 2025-07-11 main/language.tex
Tweaks.

@@
 $(f, \partial f_f, \partial f_r)$ of the underlying function and the forward and backward approximation
-maps. However, unlike in $\Fam(\LatGal)$ are not guaranteed that the forward and backward maps even operate on
+maps. However, unlike in $\Fam(\LatGal)$, it is not guaranteed that the forward and backward maps even operate on
 the same lattices, let alone form a Galois connection. \lemref{first-order-agreement-types} guarantees that

### 2025-07-11 main/introduction.tex
[ tex ] write up the semantics and some examples
[ tex ] fix bib
Fix borked line break.

@@
   \begin{displaymath}
-    \partial (\mathit{query}\,l)_r(\bot) = [(\mathsf{a},\bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]
+    \partial (\mathit{query}\,l\,\mathit{db})_r(\bot) = [(\mathsf{a},\bot), (\mathsf{b}, \bot), (\mathsf{a}, \bot)]
   \end{displaymath}
@@
     \begin{array}{l}
-      \partial (\mathit{query}\,\mathsf{a})_r(1) = [(\mathsf{a},0), (\mathsf{b},\bot), (\mathsf{a},1)] \\
-      \partial (\mathit{query}\,\mathsf{b})_r(1) = [(\mathsf{a},\bot), (\mathsf{b},1), (\mathsf{a},\bot)]
+      \partial (\mathit{query}\,\mathsf{a}\,\mathit{db})_r(1) = [(\mathsf{a},0), (\mathsf{b},\bot), (\mathsf{a},1)] \\
+      \partial (\mathit{query}\,\mathsf{b}\,\mathit{db})_r(1) = [(\mathsf{a},\bot), (\mathsf{b},1), (\mathsf{a},\bot)]
     \end{array}
@@
 
-  In this case, the forwards portion of the Galois connection tells us for each approximation of the input whether or not it is sufficient to compute the output.
+  In this case, the forwards portion of the Galois connection tells us, for each approximation of the input, whether or not it is sufficient to compute the output.
 If we provide insufficient data to compute the output, then we will get an underapproximated output.
-For example, $\partial (\mathit{query}\, \mathsf{a})_f([(\mathsf{a},0),(\mathsf{b},\bot),(\mathsf{a},\bot)]) = \bot$ because we need all the values associated with the label $\mathsf{a}$ to compute their sum.
+Here for example we will find that $\partial (\mathit{query}\, \mathsf{a})_f([(\mathsf{a},0),(\mathsf{b},\bot),(\mathsf{a},\bot)]) = \bot$ because we need all the values associated with the label $\mathsf{a}$ to compute their sum.
 
@@
 
-\bob{Have stolen this text for \autoref{sec:diff-stab-summary}}
 \begin{itemize}
@@
 
+\todo{Rewrite this to match the actual contents}
+
 Our primary contribution in this article is to elucidate \GPS by relating it to differentiable programming and automatic differentiation.

### 2025-07-11 main/approx-as-tangents.tex
12.2.8 -> 12.5.9 of Amadio and Curien.
Fix Amadio and Curien theorem refs.
[ tex ] little tweaks: pretty sure it is Thm 12.5.10

@@
 \begin{proof}
-  \AGDA. See \citet[Theorem 12.2.8]{amadio-curien} for the
+  \AGDA. See \citet[Theorem 12.1.9]{amadio-curien} for the
   case when the posets are also cpos. The crucial technical step,
@@
 closed. To make it so, we would need to require that the principal
-downsets $\downset{x}$ are \emph{complete}
-lattices. \citet[Theorem 12.FIXME]{amadio-curien} details the
-proof. Intuitively, to generate the best approximation of an input
-value for a function, we need to take the infimum over all possible
-input values.
+downsets $\downset{x}$ are \emph{complete} lattices. \citet[Theorem
+12.5.10]{amadio-curien} details the proof. Intuitively, to generate
+the best approximation of an input value for a function, we need to
+take the infimum over all possible input values.
 
@@
 \begin{enumerate}
-\item Even without completeness, in Bounded meet posets and L-posets,
+\item Even without completeness, in bounded meet posets and L-posets
   values and their approximations live in the same set. However, in
@@
   their approximations using the \emph{Category of Families}
-  construction we detail in \secref{models-of-total-gps:fam}.
+  construction as we explain in \secref{models-of-total-gps:fam}.
 \item A more practical reason is that we wish to formalise our
@@
   suprema and infinma of families in a lower universe level than
-  $X$. This is not necessarily a problem, as \citet{dejong21} shows
-  how to develop a large amount of domain theory in a predicative
+  $X$. This is not necessarily a problem, as \citet{dejong21} show how
+  to develop a large amount of domain theory in a predicative
   setting. However, due to constructivity the lifting construction

### 2025-07-11 main/models.tex
Tweaks.
Typo.
Obsolete.

@@
   ``forward'' direction. The category $\Fam(\LatGal)$ is our preferred
-  model for total functions with Galois sliicing. We explore some
+  model for total functions with Galois slicing. We explore some
   specific examples in this category in \secref{semantic-gps}.
@@
 
+  \vspace{-1mm}
   \noindent satisfying
 
-  \vspace{-3mm}
+  \vspace{-5mm}
   \begin{minipage}[t]{0.45\textwidth}
@@
 does not affect the points of the original object, but adds a new
-minimum approximation. This monad allows us to construct morphisms in
-$\Fam(\LatGal)$ that track information about how values are used
-similar to the examples of cm and stable functions we saw in
-\secref{approx-as-tangents}.
+minimum approximation.
 
@@
 $\mathrm{or} : \Bool \times \Bool \to \Bool$ be the (unapproximated)
-boolean or function. The, using a Moggi-style let notation
+boolean OR function. Using a Moggi-style let notation
 \cite{notions-of-computation} for morphisms constructed using the
@@
 lattice at each point of $\Lift(\Bool)$, then applying the reverse
-derivative at the point $(\mathsf{tt},\mathsf{tt})$ to the tangent
-$\top$ reveals which of the inputs contributed to the output for each
+derivative at $(\mathsf{tt},\mathsf{tt})$ to the tangent $\top$
+reveals which of the inputs contributed to the output for each
 function:
@@
     our proposed setup, we follow the approximation theme of \GPS by
-    having a \emph{set} of things that could be taken instead of the
-    original $x$. Objects of $\Fam(\LatGal)$ are ``directed'' in this
-    sense because the approximation map can pick out the element of
-    the original poset that was approximating $x$.}.
+    having a \emph{set} of things that could be used to replace the
+    original $x$. We observe that objects of $\Fam(\LatGal)$ arising
+    from $\Lposet$ are ``directed'' in the
+    \citet{ahman-chapman-uustalu2012} sense because the map can pick
+    out the element of the original poset that was approximating
+    $x$.}.
 \end{remark}
@@
 coproducts, are always Boolean lattices, meaning that we can take
-complements of usage information. The ability to take boolean
-complements of approximations has been used by \citet{perera22} to
-compute \emph{related} outputs, as we discuss in
+complements of usage information. The ability to take complements of
+approximations has been used by \citet{perera22} to compute
+\emph{related} outputs, as we discuss in
 \secref{related-work:galois-slicing}.
@@
     \end{array}\\
-    \mathrm{neg}&=&(\lambda x.\,-x, \lambda x\,[l,u].\,[-u,-l], \lambda x\,[l,u].\,[-u,-l]) \\
+    \mathrm{neg}&=&(\lambda x.\,{-}x, \lambda x\,[l,u].\,[-u,-l], \lambda x\,[l,u].\,[-u,-l]) \\
     \mathrm{scale}(r)&=&(\lambda x.\,rx, \lambda x\,[l,u].\,[rl,ru], \lambda x\,[l,u].\,\mathrm{if}\,r=0\,\mathrm{then}\,\bot\,\mathrm{else}\,[\frac{l}{r},\frac{u}{r}])
@@
 behaviour on $\bot$ is determined.)  Scaling by negative numbers is
-also possible with swapping of lower and upper bounds, as is
-multiplication.
+also possible with swapping of bounds, as is multiplication.
 
@@
 
-Comparing this conjecture to the previous one, we can see that the
-difference between $\CM$, $\Lposet$ and $\Fam(\MeetSLat)$,
+Comparing this conjecture to \conref{tangent-stable-fns}, we can see
+that the difference between $\CM$, $\Lposet$ and $\Fam(\MeetSLat)$,
 $\Fam(\LatGal)$ is that the latter have a separation of points from
-tangents, somewhat analogous to the situation with manifolds. Also
-note that, if this conjecture holds, then contrary to the
-$\Fam(\FinVect)$ representation of manifolds and differentiable maps,
-we do not throw away information about higher derivatives. It is
-retained in the order structure of the tangent fibres.
+tangents, somewhat analogous to the situation with manifolds. If this
+conjecture holds, then contrary to the $\Fam(\FinVect)$ representation
+of manifolds and differentiable maps, we do not throw away information
+about higher derivatives. It is retained in the order structure of the
+tangent fibres.

### 2025-07-11 main/language.tex
Typo.
[ tex ] expand on the examples; emphasise the type dependency

@@
 $\tmInr{\tmUnit}$. We consider several interpretations of $\Sigma_{\mathsf{num}}$ in $\Fam(\LatGal)$ and their
-behaviour on the selection-and-sum query from \exref{introduction-example}.
+behaviour on the selection-and-sum query from \exref{introduction-example}. First, we note the type of the
+reverse approximation map in this case.
+The type of the approximation maps depends on the input value.
+Our example input database was $\mathit{db} = [(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$, meaning that the type of the reverse approximation map for this database and any label $l$ is:
+\begin{displaymath}
+  \begin{array}{l}
+    (\partial \mathrm{query})_r(l, \mathit{db}) : \\
+    \quad \partial\sem{\mathsf{num}}(\mathrm{query}\,(l,\mathit{db})) \multimap \mathbb{1} \times (\mathbb{1} \times \partial\sem{\mathsf{num}}(0)) \times (\mathbb{1} \times \partial\sem{\mathsf{num}}(1)) \times (\mathbb{1} \times \partial\sem{\mathsf{num}}(1)) \times \mathbb{1}
+  \end{array}
+\end{displaymath}
+in the category $\JoinSLat$, where $\partial\sem{\mathsf{num}}(x)$ is the lattice of approximations of the
+number $x$ determined by our interpretation. In the codomain, the first four $\mathbb{1}$s correspond to the
+positions of labels in the input, which we are not approximating, and the final $\mathbb{1}$ is the terminator
+of the list. Note how, even if the $\partial\sem{\mathsf{num}}(x)$ does not depend on $x$ the type of the
+output is still dependent on the shape of the input list: type dependency is used in a fundamental way in our
+interpretation.
+
 \begin{enumerate}
@@
   $\sem{\mathrm{add}}_{\PrimOp}$ the embeddings of the usual zero and addition functions via $\Disc$, then the
-  resulting interpretation contains no approximation information. The approximation lattices for all types are
-  always the trivial one-point lattice, and the forward and reverse approximation maps are necessarily the
-  identity.
-\item To get interesting approximation information, we take
-  $\sem{\mathsf{num}}_{\PrimTy} = \Lift(\Disc(\mathbb{R}))$, using the lifting monad from
+  resulting interpretation contains no approximation information. We have
+  $\partial\sem{\mathsf{num}}(x) = \mathbb{1}$ so the type of $(\partial \mathrm{query})_r$ is trivial.
+\item We take $\sem{\mathsf{num}}_{\PrimTy} = \Lift(\Disc(\mathbb{R}))$, using the lifting monad from
   \secref{models-of-total-gps:lifting}, with $\sem{\mathrm{zero}}_{\PrimOp}$ and
-  $\sem{\mathrm{add}}_{\PrimOp}$ defined from the unlifted interpretations above and the monad structure. This
-  interpretation recovers the behaviour given in \exref{introduction-example}: running the reverse
-  approximation map at the input $\mathsf{a}$ with the top approximation maps indeed reveals that only the
-  numbers in the rows tagged with that label in the input are used, and likewise for $\mathsf{b}$.
+  $\sem{\mathrm{add}}_{\PrimOp}$ defined from the unlifted interpretations above and the monad structure. The
+  type of the reverse map now becomes:
+  \begin{displaymath}
+    (\partial \mathrm{query})_r(l, \mathit{db}) : \{\top,\bot\} \multimap \mathbb{1} \times (\mathbb{1} \times \{\top,\bot\}) \times (\mathbb{1} \times \{\top,\bot\}) \times (\mathbb{1} \times \{\top,\bot\}) \times \mathbb{1}
+  \end{displaymath}
+  where every position that corresponds to a number has been tagged with $\top$ for ``present'' and $\bot$ for
+  ``not present''. This interpretation recovers the behaviour given in \exref{introduction-example}: running
+  the reverse approximation map at the input $\mathsf{a}$ with the top approximation maps indeed reveals that
+  only the numbers in the rows tagged with that label in the input are used, and likewise for $\mathsf{b}$.
 \item More quantitative approximation information can be obtained by using the interval approximation
@@
   \begin{displaymath}
-    (\partial\mathrm{query}\,\mathsf{a}\,\mathsf{db})_r([\frac{9}{10},\frac{11}{10}]) = [(\cdot,[-\frac{1}{10},\frac{1}{10}]),(\cdot,\bot),(\cdot,[\frac{9}{10},\frac{11}{10}])]
+    (\partial\mathrm{query})_r(\mathsf{a},\mathsf{db})([\frac{9}{10},\frac{11}{10}]) = \cdot, (\cdot,[-\frac{1}{10},\frac{1}{10}]),(\cdot,\bot),(\cdot,[\frac{9}{10},\frac{11}{10}]), \cdot
   \end{displaymath}
@@
   $\mathsf{a}$ row could be in $[-\frac{1}{10},\frac{1}{10}]$ or the second one could be in
-  $[\frac{9}{10},\frac{11}{10}]$.
+  $[\frac{9}{10},\frac{11}{10}]$, and it does not matter what the number in the second entry was.
 \end{enumerate}
@@
 translation. We use the $T$ monad from \secref{tagging-monad} because it can be defined in terms of the
-language constructed we already have. This requires that we have a signature $\Sigma$ that includes a
+language constructs we already have. This requires that we have a signature $\Sigma$ that includes a
 primitive type to be interpreted as the approximation object $\mathbb{A}$ and primitive operations to be
@@
 
-TODO: describe the syntactic CBN translation; show an example on the query example.
+The monadic CBN translation is standard, and entirely determined by the translation on types, so we only
+define the type translation $\cbn{-}$ here:
+\begin{mathpar}
+  \cbn{\rho} = \rho
+
+  \cbn{\sigma \tySum \tau} = T(\cbn{\sigma}) \tySum T(\cbn{\tau})
+
+  \cbn{\tyUnit} = \tyUnit
+
+  \cbn{\sigma \tyProd \tau} = T(\cbn{\sigma}) \tyProd T(\cbn{\tau})
+
+  \cbn{\sigma \tyFun \tau} = T(\cbn{\sigma}) \tyFun T(\cbn{\tau})
+
+  \cbn{\tyList\;\tau} = \tyList\,(T(\cbn{\tau}))
+\end{mathpar}
+Thus, the CBN translation on types inserts a use of the monad $T(X) = \mathbb{A} \times X$ at the point ``just
+underneath'' every type former. In this case, we are describing the type of data annotated at every
+level. Note that our lists here are still ``strict'', an alternative approach would be to consider ``lazy''
+lists that wrap the tail of every node in a $T$ as well.
+
+We illustrate the effect of the CBN translation on the query example from
+\exref{introduction-example}. Applying $\cbn{-}$ to the type of the database

### 2025-07-11 main/language.tex
[ tex ] more tweaking of the examples

@@
   ``not present''. This interpretation recovers the behaviour given in \exref{introduction-example}: running
-  the reverse approximation map at the input $\mathsf{a}$ with the top approximation maps indeed reveals that
-  only the numbers in the rows tagged with that label in the input are used, and likewise for $\mathsf{b}$.
-\item More quantitative approximation information can be obtained by using the interval approximation
-  interpretation from \exref{intervals-and-maxima-elements} and \secref{interval-approx}. We let
+  the reverse approximation map at the input ``$\mathsf{a}$'' at approximation $\top$ reveals that only the
+  numbers in the rows tagged with ``$\mathsf{a}$'' in the input are used, and likewise for ``$\mathsf{b}$''.
+\item Quantitative approximation information with non-trivial dependency can be obtained by using the interval
+  approximation interpretation from \exref{intervals-and-maxima-elements} and \secref{interval-approx}. We let
   $\sem{\mathsf{num}}_{\PrimTy} = \mathbb{R}_{\mathit{intv}}$ and interpret addition using the morphism given
-  in \secref{interval-approx}.
-
-  Recall that $\mathrm{query}\,\mathsf{a}\,\mathit{db} = 1$, so in the reverse direction we can pick an
-  interval containing $1$ to discover the largest (i.e. least in the order) intervals that will give rise to
-  this output as \emph{independent} changes to the input. For example, if we pick
-  $[\frac{9}{10},\frac{11}{10}]$ as the interval, then:
+  in \secref{interval-approx}. Recall that $\mathrm{query}\,(\mathsf{a},\mathit{db}) = 1$, so in the reverse
+  direction we must choose an interval containing $1$ to discover the largest (i.e. least in the order)
+  intervals that will give rise to this output as \emph{independent} changes to the input. For example, if we
+  pick $[\frac{9}{10},\frac{11}{10}]$ as the interval, then:
   \begin{displaymath}
@@
   \end{displaymath}
-  Indicating that, to achieve an output within $[\frac{9}{10},\frac{11}{10}]$ for this query, either the first
-  $\mathsf{a}$ row could be in $[-\frac{1}{10},\frac{1}{10}]$ or the second one could be in
-  $[\frac{9}{10},\frac{11}{10}]$, and it does not matter what the number in the second entry was.
+  Thus, to achieve an output within $[\frac{9}{10},\frac{11}{10}]$, either the first ``$\mathsf{a}$'' row
+  could be in $[-\frac{1}{10},\frac{1}{10}]$ or the second one could be in $[\frac{9}{10},\frac{11}{10}]$, and
+  the number in the ``$\mathsf{b}$'' row is not relevant.
 \end{enumerate}
 
-We have tested these examples on our fully formalised Agda implementation. See the file \texttt{example.agda}
-in the supplementary material for details.
+We have tested these examples on our Agda implementation. See the file \texttt{example.agda}.
 

### 2025-07-11 main/definability.tex
Delete last sentence.
[ tex ] reword Thm 5.3

@@
   $\Gamma \vdash M : \tau$, where $\Gamma$ and $\tau$ are first-order,
-  the interpretation $\sem{\Gamma \vdash M : \tau}_{\Set}$ in $\Set$
-  and the interpretation
+  the underlying function in the interpretation
   $\sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
-    \JoinSLat^\op)}$ agree on their first components.
+    \JoinSLat^\op)}$ is equal to the interpretation
+  $\sem{\Gamma \vdash M : \tau}_{\Set}$ in $\Set$.
 \end{theorem}
@@
   that is the identity in the first component and projects out the
-  underlying function in the second. We get, for each
-  $\Gamma \vdash M : \tau$, a $g$ such that
+  underlying function in the second. For each
+  $\Gamma \vdash M : \tau$, we obtain a $g$ such that
   $g = \sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
     \JoinSLat^\op)}$ and
-  $\pi_1(g) = \sem{\Gamma \vdash M : \tau}_\Set$. Therefore,
-  $\pi_1(\sem{\Gamma \vdash M : \tau}_{\Fam(\MeetSLat \times
-    \JoinSLat^\op)}) = \sem{\Gamma \vdash M : \tau}_\Set$.
+  $\pi_1(g) = \sem{\Gamma \vdash M : \tau}_\Set$. Substituting $g$
+  yields the result.
 \end{proof}

### 2025-07-11 main/definability.tex
[ tex ] fiddle with the definability text

@@
   \label{thm:language-definability}
-  For all $\Gamma \vdash M : \tau$, where $\Gamma$ and $\tau$ are
-  first-order, there exists a morphism
+  For all $\Gamma \vdash M : \tau$, with $\Gamma, \tau$ first-order,
+  there exists
   $g \in \Fam(\LatGal)(\sem{\Gamma}_{\mathit{fo}},
@@
   $H(g) = (\cong) \circ \sem{\Gamma \vdash M : \tau} \circ (\cong)$,
-  where the isomorphisms arise from
-  \lemref{first-order-agreement-types}.
+  with the isomorphisms from \lemref{first-order-agreement-types}.
 \end{theorem}
@@
 \begin{remark}
-  Note that even if we modified our base interpretation of semantic
-  \GPS along the lines of \remref{further-structure} to give a refined
-  version $\cat{G}$ of $\Fam(\LatGal)$, as long as there is a finite
-  product and coproduct preserving functor
-  $\cat{G} \to \Fam(\MeetSLat \times \JoinSLat^\op)$, then the
-  analogous result to \thmref{language-definability} still holds.
+  If we modified our base interpretation of semantic \GPS as suggested
+  in \remref{further-structure} to give a refined version $\cat{G}$ of
+  $\Fam(\LatGal)$, then if there is a finite bicartesian functor
+  $\cat{G} \to \Fam(\MeetSLat \times \JoinSLat^\op)$, an analogous
+  result to \thmref{language-definability} still holds.
 \end{remark}

### 2025-07-11 main/conclusion.tex
[ tex ] Bit more summary text in conclusion
[ tex ] save a line

@@
 our model, and the use of general theorems, such as Fiore and Simpson's \thmref{glr-definability}, to prove
-properties of the interpretation.
+properties of the interpretation. We have focused on constructions that enable an executable implementation in
+Agda in this work, but have conjectured that there are connections to established notions of categorical
+differentiation in \conref{tangent-stable-fns} and \conref{tangent-fam}.
 
@@
 \cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
-give an account of reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
+model reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
 of \citet{vakar22}.

### 2025-07-11 main/introduction.tex
Delete.

@@
 Previous work on \GPS used a special operational semantics to generate a trace of each execution, and then separately used that trace to compute the Galois connection described above, by re-running forwards or backwards over the trace.
-It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately (but somewhat redundantly) defined ``backwards evaluation'' operation.
+It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately defined ``backwards evaluation'' operation.
 As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vakar22}, which points to a way to develop such an approach.

### 2025-07-11 main/approx-as-tangents.tex
[ tex ] more line saving

@@
   An example that is conditionally multiplicative, but does not admit
-  a backwards map of approximations (from \citet[just
-  before Lemma 12.2.3]{amadio-curien}, originally due to Berry) is given by defining
+  a backwards map of approximations (from \citet[just before Lemma
+  12.2.3]{amadio-curien}, originally due to Berry) is given by
   $\mathrm{unstable} : D \to \{\bot \sqsubseteq \top\}$, where
@@
   and $\mathrm{unstable}(n) = \top$. This is monotone, and preserves
-  meets in every $\downset{x}$. However, there is no ``best'' (i.e.,
-  least) input that gives us any finite output.
+  meets in every $\downset{x}$. But there is no ``best'' (i.e., least)
+  input that gives us any finite output.
 \end{example}

### 2025-07-11 main/approx-as-tangents.tex
the

--- 2025-07-11 main/approx-as-tangents.tex
@@
   the minimal $x_0$ required by stability. This is monotone: if
-  $y \leq y'$, then we know that $y \leq y' \leq f(f^*_x(y'))$ by
+  $y \leq y'$, then we know that $y \leq y' \leq f(f^*_x(y'))$ by the
   definition of $f^*_x$, so $f^*_x(y) \leq f^*_x(y')$ by minimality of

### 2025-07-11 main/approx-as-tangents.tex
Typo.

@@
   lattices in the classical sense: for a type $X$ we can only get
-  suprema and infinma of families in a lower universe level than
+  suprema and infima of families in a lower universe level than
   $X$. This is not necessarily a problem, as \citet{dejong21} show how

### 2025-07-11 main.tex
[ tex ] review mode
Typo.

@@
-\documentclass[acmsmall,screen,anonymous]{acmart}
+\documentclass[acmsmall,screen,anonymous,review]{acmart}
 
@@
   forwards and backwards along a particular execution. In this paper,
-  we explore a analogy between \GPS and differentiable programming,
-  seeing the implementation of \GPS as a kind of automatic
+  we explore an analogy between \GPS and differentiable programming,
+  seeing the implementation of forwards and backwards slicing as a kind of automatic
   differentiation. Using the CHAD approach to automatic differentation
-  due to V{\'a}k{\'a} and collaborators, we reformulate \GPS via a
+  due to V{\'a}k{\'a}r and collaborators, we reformulate \GPS via a
   categorical semantics. In doing so, we are able to explore
   extensions of the \GPS idea to quantitative interval analysis, and
-  to clarify the implicit choices made in existing \GPS
-  instantiations.
+  to clarify the implicit choices made in existing instantiations of this approach.
 \end{abstract}

### 2025-07-11 main/introduction.tex
Start on intro.

@@
 Existing approaches to program slicing are often tied to particular programming languages or implementations.
-In this paper we develop a general categorical approach to program slicing and \todo{complete}.
+In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
+called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
+connection. Our main insight is that this approach to slicing can be seen as a generalised form of
+differentiable programming, with approximations (slices) of elements playing the role of tangents.
 

### 2025-07-11 main/approx-as-tangents.tex
Brackets for consistency with notation elsewhere.

@@
 categorical axiomatisation of the properties of manifolds and smooth
-functions in terms of the presence of tangent bundles $TX$ for every
+functions in terms of the presence of tangent bundles $T(X)$ for every
 object $X$, forward derivatives and additivity of tangents. They
@@
   form a Tangent category where the tangent bundle
-  $TX = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
+  $T(X) = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
   by meets. (2) L-posets and stable functions form a \emph{reverse

### 2025-07-11 main/introduction.tex
Brackets for consistency with notation elsewhere.

@@
 In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
-called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
-connection. Our main insight is that this approach to slicing can be seen as a generalised form of
-differentiable programming, with approximations (slices) of elements playing the role of tangents.
+called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
+backward slicing procedures for a particular execution induce Galois connections between these lattices. Our
+main insight is that this approach to slicing can be seen as a generalised form of differentiable programming,
+with approximations (slices) of elements playing the role of tangents. Our categorical semantics should
+provide a foundation for providing
 

### 2025-07-11 main/language.tex
[ tex ] finish a sketch of the CBN translation

@@
 interpretation.
+\newcommand{\TWO}{\mathbb{2}}
 
@@
   $\sem{\mathrm{add}}_{\PrimOp}$ defined from the unlifted interpretations above and the monad structure. The
-  type of the reverse map now becomes:
-  \begin{displaymath}
-    (\partial \mathrm{query})_r(l, \mathit{db}) : \{\top,\bot\} \multimap \mathbb{1} \times (\mathbb{1} \times \{\top,\bot\}) \times (\mathbb{1} \times \{\top,\bot\}) \times (\mathbb{1} \times \{\top,\bot\}) \times \mathbb{1}
-  \end{displaymath}
+  type of the reverse map now becomes, where $\TWO = \{\top.\bot\}$:
+  \begin{equation}\label{eqn:reverse-query-approx-1}
+    (\partial \mathrm{query})_r(l, \mathit{db}) : \TWO \multimap \mathbb{1} \times (\mathbb{1} \times \TWO) \times (\mathbb{1} \times \TWO) \times (\mathbb{1} \times \TWO) \times \mathbb{1}
+  \end{equation}
   where every position that corresponds to a number has been tagged with $\top$ for ``present'' and $\bot$ for
@@
 
+\newcommand{\Element}{(\TWO \times \TWO)}
+
 We illustrate the effect of the CBN translation on the query example from
-\exref{introduction-example}. Applying $\cbn{-}$ to the type of the database
+\exref{introduction-example}. Applying $\cbn{-}$ to the type of $\mathrm{query}$ yields:
+\begin{displaymath}
+  T(\mathrm{Label}) \times T(\tyList(T(T(\mathrm{Label}) \tyProd T(\mathsf{num})))) \to T(\mathsf{num})
+\end{displaymath}
+Thus, every substructure of the input and output has been annotated with usage information. Under
+interpretation in $\Fam(\MeetSLat \times \JoinSLat^\op)$, the type of the reverse approximation map as a
+morphism in $\JoinSLat$ at the input $\mathit{db}$ is now (suppressing some ``$\times \mathbb{1}$'' for
+readability):
+\begin{displaymath}
+  (\partial \mathrm{query})_r(l, \mathit{db}) : \TWO \multimap \TWO \times (\TWO \times \Element \times \Element \times \Element)
+\end{displaymath}
+Comparing to the type in (\ref{eqn:reverse-query-approx-1}), we now gain much more fine-grained information on
+which parts of the input are used. When $l = \mathsf{a}$, we have
+$(\partial \mathrm{query})_r(\mathsf{a},\mathit{db})(\top) = (\top, (\top, (\top, \top), (\top, \bot), (\top,
+\top)))$, indicating that the execution of this query had examined everything except the number in the second
+entry of the database. With the previous interpretation, we did not have confirmation that the labels in each
+row were actually required.

### 2025-07-11 main/introduction.tex
WIP.

--- 2025-07-11 main/introduction.tex
@@
 In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
-called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
-connection. Our main insight is that this approach to slicing can be seen as a generalised form of
-differentiable programming, with approximations (slices) of elements playing the role of tangents.
+called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
+backward slicing procedures give rise to Galois connections between these lattices. Our main insight is that
+this approach to slicing can be seen as a generalised form of differentiable programming, with approximations
+(slices) of elements playing the role of tangents. Our categorical approach should provide a suitable setting
+for exploring data provenance for a variety of programming languages, and also allows easy experimentation
+with alternative approximation strategies, including quantitative forms of slicing.
 
@@
 In particular, we show that there is a close connection with Stable Domain Theory, a proposed solution to capturing sequentiality by recording more intensional information about programs' sensitivity to approximations.
-\item We also relate \GPS to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
+% \item We also relate \GPS to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
 \end{enumerate}

### 2025-07-11 main/related-work.tex
[ tex ] tweak related work

@@
 conjectured that the categories we have identified in this paper as models of \GPS are Tangent categories. As
-well as providing some substantiation to the claim that \GPS is a form of differentiation, we should be able
-to apply general results about Tangent categories to our setting. % BOB: such as?
+well as providing some substantiation to the claim that \GPS is a form of differentiation, this would also
+clarify the role of higher derivatives in \GPS, which we conjecture are related to \emph{program
+  differencing}.
 

### 2025-07-11 main/related-work.tex
[ tex ] fix non-rectangular paragraph

@@
 compositional semantic structure. These perspectives shed light on the categorical structure of AD and guide
-the design of systems that generalise AD beyond real analysis, including the application to data provenance
+the design of systems that generalise AD% beyond real analysis
+, including the application to data provenance
 and slicing explored in this paper.

### 2025-07-11 main/introduction.tex
Drop indentation in  environment at end of 2.2.

--- 2025-07-11 main/introduction.tex
@@
 In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
-called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
-connection. Our main insight is that this approach to slicing can be seen as a generalised form of
-differentiable programming, with approximations (slices) of elements playing the role of tangents.
+called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
+backward slicing procedures generate Galois connections between these lattices. Our main insight is that this
+approach can be seen as a generalised form of differentiable programming, with slices of elements playing the
+role of tangents. Our categorical approach should provide a suitable setting for exploring data provenance for
+a variety of programming languages, and is easily configured to use alternative approximation strategies,
+including quantitative forms of slicing.
 
@@
 
-\todo{Rewrite this to match the actual contents}
-
 Our primary contribution in this article is to elucidate \GPS by relating it to differentiable programming and automatic differentiation.
 In doing so, we aim to expand the applicability of program slicing, and to potentially transfer efficient automatic differentiation implementation techniques from automatic differentiation to program slicing and provenance tracking.
+
+\todo{Rewrite this to match the actual contents}
 

### 2025-07-11 main/introduction.tex
Some leftmargin parameterisation.
[ tex ] condensed summary in 1.3

@@
 called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
-backward slicing procedures generate Galois connections between these lattices. Our main insight is that this
+backward slicing procedures generate Galois connections between these lattices. Our main contribution is that this
 approach can be seen as a generalised form of differentiable programming, with slices of elements playing the
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=\enummargin]
 \item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
@@
 
-Our primary contribution in this article is to elucidate \GPS by relating it to differentiable programming and automatic differentiation.
+%Our primary contribution in this article is to elucidate \GPS by relating it to differentiable programming and automatic differentiation.
 In doing so, we aim to expand the applicability of program slicing, and to potentially transfer efficient automatic differentiation implementation techniques from automatic differentiation to program slicing and provenance tracking.
 
-\todo{Rewrite this to match the actual contents}
+\GPS, as any program slicing technique, essentially rests on an analysis of how programs intensionally explore their input, in addition to their extensional behaviour.
+Such analysis has been carried out over many years in Domain Theory.
+In \secref{approx-as-tangents}, we use ideas from \citet{berry79}'s \emph{stable domain theory} and develop an analogy between stable functions and smooth functions from mathematical analysis, where stable functions provide a kind of semantic provenance analysis.
+In \secref{models-of-total-gps}, we abstract from stable functions using V{\'a}k{\'a}r \etal{}'s CHAD framework \cite{vakar22,nunes2023} to build models of a higher-order language that automatically compute slices.
+We apply this to a concrete higher-order language in \secref{language} and demonstrate the use of the model on variations of \exref{introduction-example}, highlighting the flexibility of our approach.
+In particular, we show how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.
+We prove two correctness properties in \secref{definability}, relating the higher-order interpretations to first-order ones, proving the crucial Galois connection property.
+\secref{related-work} and \secref{conclusion} discuss additional related and future work.
 
-\begin{enumerate}
-\item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS.
-\item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.
-\item With the benefit of our abstract setting, we can relate \GPS to other parts of denotational semantics.
+% \begin{enumerate}[leftmargin=\enummargin]
+% \item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS.
+% \item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.
+% \item With the benefit of our abstract setting, we can relate \GPS to other parts of denotational semantics.
 In particular, we show that there is a close connection with Stable Domain Theory, a proposed solution to capturing sequentiality by recording more intensional information about programs' sensitivity to approximations.
-% \item We also relate \GPS to categorical treatments of differentiability via Tangent Categories \note{Bob: not sure about this.}
-\end{enumerate}
+% \end{enumerate}
 
-We have \anonymise{\href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}}{formalised this work in Agda}.
+We have formalised our major results in Agda, resulting in an executable implementation built directly from the categorical constructions that we have used to compute the examples in \secref{language:examples}.
+Please consult the file \texttt{everything.agda} in the supplementary material.
+
+% \begin{enumerate}
+% \item We explain how the CHAD framework of Vákár et al. can be adapted to give a general categorical framework for \GPS.
+% \item Using our adapted CHAD framework, we can explain how type structure can be used to control the approximation lattices associated with data points, something that was ``hard coded'' in previous presentations of \GPS.
+% \item With the benefit of our abstract setting, we can relate \GPS to other parts of denotational semantics.
+In particular, we show that there is a close connection with Stable Domain Theory, a proposed solution to capturing sequentiality by recording more intensional information about programs' sensitivity to approximations.
+
+% We have \anonymise{\href{https://github.com/bobatkey/approx-diff}{formalised this work in Agda}}{formalised this work in Agda}.
 Not only does this mean that the constructions and proofs have been checked, but also the construction is executable and can be run on small examples.

### 2025-07-11 main/introduction.tex
[ tex ] squash lattices
Tweak intro.

@@
 called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
-backward slicing procedures generate Galois connections between these lattices. Our main contribution is that this
-approach can be seen as a generalised form of differentiable programming, with slices of elements playing the
-role of tangents. Our categorical approach should provide a suitable setting for exploring data provenance for
-a variety of programming languages, and is easily configured to use alternative approximation strategies,
-including quantitative forms of slicing.
+backward slicing procedures generate Galois connections between these lattices. Our main contribution is that
+this approach can be seen as a generalised form of automatic differentiation, with slices of values playing
+the role of tangents. Our categorical approach should provide a suitable setting for enabling ``automatic''
+data provenance for a variety of programming languages, and is easily configured to use alternative
+approximation strategies, including quantitative forms of slicing.
 
@@
   \begin{center}
-    \begin{tikzpicture}
+    \begin{tikzpicture}[node distance=0.75cm]
       \node (top) at (0,0) {$[(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$};
@@
   \begin{center}
-    \begin{tikzpicture}
+    \begin{tikzpicture}[node distance=0.75cm]
       \node (top) at (0,0) {$1$};

### 2025-07-11 main/approx-as-tangents.tex
Drop indentation in  environment at end of 2.2.
Some leftmargin parameterisation.
[ tex ] further space saving

@@
 
-We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting.
-From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps.
+We motivate our approach by showing how to combine ideas from differential geometry and stable domain theory to reconstruct the ideas of \GPS in a denotational setting.
+% From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps.
 
@@
 Each smooth function induces maps of the (co)tangent spaces:
-\begin{itemize}
+\begin{itemize}[leftmargin=\enummargin]
 \item The \emph{forward derivative} (tangent map, pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$.
@@
   \leavevmode\\[-3ex]
-  \begin{enumerate}
+  \begin{enumerate}[leftmargin=\enummargin]
   \item The function $\mathrm{strictOr}$ is stable. For example, for
@@
 
-\begin{enumerate}
+\begin{enumerate}[leftmargin=*]
 \item Even without completeness, in bounded meet posets and L-posets

### 2025-07-11 main/conclusion.tex
[ tex ] further space saving

@@
 
-\paragraph{Categorical Models of Differentation} We have already identified \conref{tangent-stable-fns} and
-\conref{tangent-fam} as potential connections to categorical models of differentation. It would also be
-interesting to discover any possible links to Differential Linear Logic
-\cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
-model reverse differentiation by \citet{kerjean-pedrot2024} that is similar to the CHAD approach
-of \citet{vakar22}.
+% \paragraph{Categorical Models of Differentation} We have already identified \conref{tangent-stable-fns} and
+% \conref{tangent-fam} as potential connections to categorical models of differentation.
 
@@
 far would likely involve families of bounded lattices indexed by DCPOs, where the order of the DCPO would be
-reflected in embedding-projection relationships between the lattices. There are also other refinements of
-stable domain theory such as \citet{berry79}'s bidomains and \citet{laird07} bistable biorders that separate
-the extensional and stable orders on the same set, in a way that might be similar to
-\exref{intervals-and-maxima-elements}.
+reflected in embedding-projection relationships between the lattices. \citet{berry79}'s bidomains and
+\citet{laird07}'s bistable biorders have separate extensional and stable orders on the same set, in a way that
+might be similar to \exref{intervals-and-maxima-elements}.
 

### 2025-07-11 main/related-work.tex
[ tex ] further space saving

@@
 
-\subsection{Stable Domain Theory}
+\paragraph{Stable Domain Theory}
 
@@
 
-\subsection{Automatic Differentiation}
+\paragraph{Automatic Differentiation}
 
@@
 
-\subsection{\GPS}
+\paragraph{\GPS}
 \label{sec:related-work:galois-slicing}
@@
 
-\subsection{Tangent Categories}
+\paragraph{Tangent Categories and Differential Linear Logic}
 
@@
 the existence of reverse derivatives. In \conref{tangent-stable-fns} and \conref{tangent-fam}, we have
-conjectured that the categories we have identified in this paper as models of \GPS are Tangent categories. As
-well as providing some substantiation to the claim that \GPS is a form of differentiation, this would also
-clarify the role of higher derivatives in \GPS, which we conjecture are related to \emph{program
-  differencing}.
+conjectured that the categories we have identified in this paper as models of \GPS are Tangent
+categories. This would clarify the role of higher derivatives in \GPS, which we conjecture are related to
+\emph{program differencing}. There are likely links to Differential Linear Logic
+\cite{ehrhard_differential_2006}. Differential Linear Logic and the Dialectica translation have been used to
+model reverse differentiation by \citet{kerjean-pedrot2024}.
 
-% \subsection{Differential Linear Logic}
-
-% Differential Linear Logic \cite{dill} is an extension of Linear Logic with a derivative operator that acts
-% like
-
-% \subsection{Lens Categories}
+% \paragraph{Lens Categories}
 

### 2025-07-11 main/introduction.tex
Tweak intro example test to better foreshadow 3.3.

@@
   \label{ex:introduction-example}
-  The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then computing the sum of the numbers:
+  The following program is written in Haskell syntax \cite{haskell}, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then computing the sum of the numbers:
   \begin{displaymath}
@@
   These are not the only choices of approximation lattices that we could have made.
-For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not any of the other data or the overall structure.
+For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not the labels or the structure of the list itself.
 However, other choices are also useful.
 Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information.
-We elaborate on this further in \secref{models-of-total-gps}.
+We elaborate on this further in \secref{semantic-gps}.
 

### 2025-07-11 main/introduction.tex
Tweak.

@@
 
-Previous work on \GPS used a special operational semantics to generate a trace of each execution, and then separately used that trace to compute the Galois connection described above, by re-running forwards or backwards over the trace.
+Previous work on \GPS uses a special operational semantics to generate a trace of each execution, and then uses that trace to compute the Galois connections described above, by re-running forwards or backwards over the trace.
 It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately defined ``backwards evaluation'' operation.

### 2025-07-11 main/introduction.tex
[ tex ] Change title; and tweak intro

@@
 
-Previous work on \GPS uses a special operational semantics to generate a trace of each execution, and then uses that trace to compute the Galois connections described above, by re-running forwards or backwards over the trace.
+Previous work on \GPS used a special operational semantics to generate a trace of each execution, and then uses that trace to compute the Galois connections described above, by re-running forwards or backwards over the trace.
 It would be useful to have a denotational account of \GPS as well, especially if we could provide a semantics where the backwards analysis is baked in, rather than provided by a separately defined ``backwards evaluation'' operation.
-As it happens there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vakar22}, which points to a way to develop such an approach.
+Our thesis, developed in \secref{approx-as-tangents} and \secref{models-of-total-gps} is that there is a close analogy between \GPS and \emph{automatic differentiation} for differentiable programs~\cite{siskind08,elliott18,vakar22}, which points to a way to develop such an approach.
 We have already hinted at this in the description above, but let us now make it explicit.
@@
 
-Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation and use them to model \GPS.
-This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wide range of computational settings.
+Given this close connection between \GPS and differentiable programming, we can take structures intended for modelling automatic differentiation, such as V{\'a}k{\'a}r's CHAD framework and use them to model \GPS.
+This will enable us to generalise and expand the scope of \GPS to act as a foundation for data provenance in a wider range of computational settings.
 

### 2025-07-11 build-suppl.sh
[ build ] tweak supplementary material script

@@ -1,3 +1,3 @@
 #!/bin/bash
 set -xe
-zip -r suppl.zip agda -x "agda/_build/*"
+zip -r suppl.zip agda -x "agda/_build/*" -x "agda/src/unused/*" -x "agda/src/incomplete/*"

### 2025-07-11 main/approx-as-tangents.tex
Tweaks.

@@
 output cannot be assigned a unique minimal input that accounts for
-it. Therefore, continuous functions are incompatible with the central
+it. Therefore, continuous functions are in general incompatible with the central
 idea in \GPS that we should be able to identify a \emph{minimal} part
@@
 \emph{Stability} is a property that can be required of monotone functions, that was invented by \citet{berry79} in an attempt to capture sequentiality.
-This was unsuccessful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to the problem of computing forward and backwards maps of approximations needed in \GPS.
+This was unsuccessful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to the problem of computing the forward and backwards maps of approximations needed in \GPS.
 A textbook description of stable functions in the context of domain theory is given by \citet[Chapter 12]{amadio-curien}.
-We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.
+We start with a property that is weaker than stability, but easier to motivate in connection with derivatives of smooth functions.
 

### 2025-07-11 main/language.tex
[ tex ] condensed summary in 1.3
Tweaks.

@@
 of the context and type, as defined in \figref{semantics:terms}. We have used the notations $\pi_i$ for
-projections, $\prodM{f}{g}$ for pairing, $\coprodM{f}{g}$ for (parameterised) copairing, $!_x$ for morphisms
+projections, $\prodM{f}{g}$ for pairing, $\coprodM{f}{g}$ for (parameterised) copairing, $!_X$ for morphisms
 to the terminal object, and $\lambda$ and $\eval$ for currying and evaluation for exponentials.
@@
 \subsection{Examples}
+\label{sec:language:examples}
 

### 2025-07-11 main/models.tex
Obsolete.
[ tex ] tidy up some hanging paragraphs
Some leftmargin parameterisation.
[ tex ] typo

@@
 
-As similar result is given by \citet{cruttwell2022}, where Euclidean
+A similar result is given by \citet{cruttwell2022}, where Euclidean
 spaces $\RR^n$ and smooth functions are embedded into a category of
@@
 
-\begin{enumerate}
+\begin{enumerate}[leftmargin=\enummargin]
 \item The category $\LatGal$ has bounded lattices as objects and
@@
   \item
-  \begin{enumerate}
+  \begin{enumerate}[leftmargin=\enummargin]
   \item A $\CMon$-category that has biproducts $X \oplus Y$ for all
@@
   hence biproducts:
-  \begin{enumerate}
+  \begin{enumerate}[leftmargin=\enummargin]
   \item In $\FinVect$, morphisms are linear maps and so can be added

### 2025-07-11 main/language.tex
Fix font.
Nicer formatting for CBN translation.

@@
   \begin{displaymath}
-    (\partial\mathrm{query})_r(\mathsf{a},\mathsf{db})([\frac{9}{10},\frac{11}{10}]) = \cdot, (\cdot,[-\frac{1}{10},\frac{1}{10}]),(\cdot,\bot),(\cdot,[\frac{9}{10},\frac{11}{10}]), \cdot
+    (\partial\mathrm{query})_r(\mathsf{a},\mathit{db})([\frac{9}{10},\frac{11}{10}]) = \cdot, (\cdot,[-\frac{1}{10},\frac{1}{10}]),(\cdot,\bot),(\cdot,[\frac{9}{10},\frac{11}{10}]), \cdot
   \end{displaymath}
@@
 define the type translation $\cbn{-}$ here:
-\begin{mathpar}
-  \cbn{\rho} = \rho
-
-  \cbn{\sigma \tySum \tau} = T(\cbn{\sigma}) \tySum T(\cbn{\tau})
-
-  \cbn{\tyUnit} = \tyUnit
-
-  \cbn{\sigma \tyProd \tau} = T(\cbn{\sigma}) \tyProd T(\cbn{\tau})
-
-  \cbn{\sigma \tyFun \tau} = T(\cbn{\sigma}) \tyFun T(\cbn{\tau})
-
-  \cbn{\tyList\;\tau} = \tyList\,(T(\cbn{\tau}))
-\end{mathpar}
+%\begin{mathpar}
+%  \cbn{\rho} = \rho
+%
+%  \cbn{\sigma \tySum \tau} = T(\cbn{\sigma}) \tySum T(\cbn{\tau})
+%
+%  \cbn{\tyUnit} = \tyUnit
+%
+%  \cbn{\sigma \tyProd \tau} = T(\cbn{\sigma}) \tyProd T(\cbn{\tau})
+%
+%  \cbn{\sigma \tyFun \tau} = T(\cbn{\sigma}) \tyFun T(\cbn{\tau})
+%
+%  \cbn{\tyList\;\tau} = \tyList\,(T(\cbn{\tau}))
+%\end{mathpar}
+\[
+\renewcommand{\arraystretch}{1.1}
+\begin{array}{r@{\;=\;}l@{\qquad}r@{\;=\;}l@{\qquad}r@{\;=\;}l}
+\cbn{\rho} & \rho &
+\cbn{\sigma \tySum \tau} & T(\cbn{\sigma}) \tySum T(\cbn{\tau}) &
+\cbn{\sigma \tyProd \tau} & T(\cbn{\sigma}) \tyProd T(\cbn{\tau}) \\
+\cbn{\tyUnit} & \tyUnit &
+\cbn{\sigma \tyFun \tau} & T(\cbn{\sigma}) \tyFun T(\cbn{\tau}) &
+\cbn{\tyList\;\tau} & \tyList\,(T(\cbn{\tau}))
+\end{array}
+\]
 Thus, the CBN translation on types inserts a use of the monad $T(X) = \mathbb{A} \times X$ at the point ``just

### 2025-07-11 main/approx-as-tangents.tex
Lose paragraph break inside enum.

@@
     there is a unique minimal input that achieves it (listed in the
-    first three lines of the definition).
-
-    In terms of \GPS the $\mathrm{gustave}$ function does not present
-    a problem. For any particular run (i.e., input $\mapsto$ output
+    first three lines of the definition). %
+%
+    In terms of \GPS, the $\mathrm{gustave}$ function does not present
+    a problem; for any particular run (i.e., input $\mapsto$ output
     pair), there is an unambiguous minimal input that achieves the

### 2025-07-11 main/models.tex
[ tex ] some final edits

@@
 
-The category $\MeetSLat$ satisfies the hypoetheses of
+The category $\MeetSLat$ satisfies the hypotheses of
 \thmref{fam-closed}, so:
@@
 \end{corollary}
-Unfortunately, neither of the $\LatGal$ or $\FinVect$ satisfy the
+Unfortunately, neither $\LatGal$ nor $\FinVect$ satisfy the
 hypotheses of this theorem, because neither of them have infinite
@@
   However, for the purposes of modelling differentiable programs, this
-  is fatally flawed in that neither of $\LatGal$ nor $\FinVect$ are
+  is fatally flawed in that neither $\LatGal$ nor $\FinVect$ are
   cartesian closed, and there is no way of making them so without
@@
   \item $\MeetSLat$ and $\JoinSLat$ are both $\CMon$-enriched and have
-    finite products for similar reasons to $\LatGal$.
+    finite products similar to $\LatGal$.
   \end{enumerate}
@@
 the vector space operations themselves are finitary. This is the
-solution that \cite{vakar22} uses for the semantics of forward
+solution that \citet{vakar22} use for the semantics of forward
 ($\Fam(\Vect)$) and reverse ($\Fam(\Vect^\op)$) automatic
@@
 behaviour on $\bot$ is determined.)  Scaling by negative numbers is
-also possible with swapping of bounds, as is multiplication.
+also possible with swapping of bounds, as is multiplication. We will
+see an example of the use of these operations in
+\secref{language:examples}.
 
@@
 
-We have seen that the category $\Fam(\MeetSLat \times \JoinSLat^\op)$,
-which is a cartesian closed category with all coproducts, is enough to
-interpret the total higher-order language we define in the next
-section with primitive types and operations defined in
-$\Fam(\LatGal)$. However, we are not guaranteed by construction that
-at first-order type, the interpretations are in fact Galois
-connections. We will rectify this in \secref{definability} using a
-logical relations construction.
+We have seen that the category $\Fam(\LatGal)$ has enough structure to
+express useful approximation maps at first-order and that
+$\Fam(\MeetSLat \times \JoinSLat^\op)$, which is a cartesian closed
+category with all coproducts, is enough to interpret the total
+higher-order language we define in the next section with primitive
+types and operations defined in $\Fam(\LatGal)$. However, we are not
+guaranteed by construction that at first-order type, the
+interpretations are in fact Galois connections. We will rectify this
+in \secref{definability} using a logical relations construction.
 

### 2025-09-11 reviews/popl2026/response.md
Import specific reviewer comments/questions.

We thank our three reviewers for their detailed reviews and suggestions for improving the paper. We address specific comments (C) and questions (Q) asked by each reviewer below, along with some proposed improvements to the paper.

# Responses to Reviewers

## Reviewer A

C1. _Some sections are highly technical and challenging for non-experts to follow._

C2. _Paper lacks empirical validation or comparisons with existing slicing tools_.

C3. _Paper does not address partiality or recursion_.

Q1. _Could the authors provide a high-level "roadmap" linking each theoretical component (e.g., Galois connection, CHAD) to its slicing/AD counterpart? This would improve accessibility._

Q2. _Is there any plan to implement this framework in a practical AD tool and compare slicing accuracy/performance with baseline methods?_

## Reviewer B

C1. _The title and abstract of the paper are misleading, as both refer to automatic differentiation._

C2. _I do not see how the contribution shows that Galois slicing is a "generalised form of automatic differentiation"_.

C3. _The paper does not discuss the existing literature on differential linear logic or probabilistic coherence spaces_.

Q1. _Do we have any counterexamples showing that $\mathrm{Fam}(\mathbf{LatGal})$ is not cartesian closed?_

Q2. _Could a more precise reference be given for Taylor's characterization of stability in terms of Galois connections?_

## Reviewer C

C1. _The evaluation could benefit from examples going beyond toy examples._

Q1. _Could the setting be extended into a source-to-source transformation?_

Q2. _Could the authors provide more end-to-end examples where such richer (quantitative) approximations yield insight unavailable in classical Galois slicing?_

Q3. _What is the computational overhead is of carrying slices alongside values?_

Q4. _How does the framework compare, qualitatively or quantitatively, with existing slicing/provenance techniques ?_

Q5. _Does the work on partiality and CHAD help in your setting?_

# List of Proposed Changes

### 2025-09-11 reviews/popl2026/response.md
draft responses

@@
 
+This is disappointing, and we hope that the addition of more examples in a future version will help to relieve some of the technical burden.
+
 C2. _Paper lacks empirical validation or comparisons with existing slicing tools_.
+
+This is true, and is something that we plan to address in future work.
 
@@
 
+This is true, and is something we plan to address in future work. Nevertheless, total programs do cover a large range of useful applications, such as database queries.
+
 Q1. _Could the authors provide a high-level "roadmap" linking each theoretical component (e.g., Galois connection, CHAD) to its slicing/AD counterpart? This would improve accessibility._
 
+We can certainly make the connections clearer in a future version, but for now:
+
+- Galois connections are the counterpart of forward and reverse tangent maps of real smooth functions.
+The relationship of being a Galois connection replaces the fact that the forward and reverse maps are linear transposes of each other.
+
+- CHAD is an approach to organising the implementation and semantics of AD.
+CHAD interprets a typed language with a real number type into the total category of some indexed category.
+The idea is that the indexing category represents the original functions, and the indexed category represents the tangent maps.
+Two interpretations are considered, one which is "syntactic" and so provides a source-to-source translation, and one which is "semantic" and interprets a term as a formal pairing of its functional meaning and its family of tangent maps.
+In our submission, we only do the semantic one but replace tangents with approximations and tangent maps with Galois connections.
+Because we formulated our results in Agda, the semantic variant is actually executable and was used for our examples, so we didn't prioritise a source-to-source version.
+
 Q2. _Is there any plan to implement this framework in a practical AD tool and compare slicing accuracy/performance with baseline methods?_
+
+We do not know if existing AD tools are appropriate as a basis for implementation, because they are specialised to real number differentiation, but we certainly plan to implement our ideas in a more practical tool to evaluate its performance.
+It may be possible to give a generic CHAD implementation that covers both.
 
@@
 
+We think that "misleading" is too harsh.
+
+We think there are two parts to the reviewer's objection:
+
+1.
+Is it correct that we appropriate "differentiation" and other words from real analysis in our setting?
+This is debatable, and certainly one could argue that absent any kind of definition of derivative as a limit that one should not use the word.
+Our argument is based on the similarity of structures (e.g., existence of tangent maps, meet preservation instead of linearity) outlined in Section 1.2, and on the conjectures about Tangent categories.
+The conjectures about Tangent categories (which in fact we have now proved for the forward case, the reverse case's definition is much more involved) extend the connection to the higher-order derivative case, but do not really say much more than that.
+
+2.
+Whether or not this is a kind of automatic differentiation.
+Our argument here is based on the fact that we are proposing an interpretation of programs that computes forward and backwards maps of additional data at a point alongside the main computation.
+By the CHAD approach, this is exactly what first-order AD does, and so we think that the identification as a generalised form is valid.
+
 C3. _The paper does not discuss the existing literature on differential linear logic or probabilistic coherence spaces_.
+
+We admit that we should have done a more detailed survey of the literature on differential linear logic and how it relates to our setting.
+The connection between the trace and Taylor expansion may have something to say in our setting.
+A difference with DiLL is in the treatment of partiality and non-determinism.
+The Ehrhard survey referenced by the reviewer highlights these as key parts of DiLL, but, as we explained at the start of Section 3, we do not wish to admit either of these in our model for all functions -- they only apply to tangent maps.
+This was our motivation for moving from stable functions to the families construction.
 
@@
 
+We believe it is the case that $\mathrm{Fam}(\mathbf{LatGal})$ is not cartesian closed, but unfortunately we do not have a proof.
+
 Q2. _Could a more precise reference be given for Taylor's characterization of stability in terms of Galois connections?_
+
+FIXME
 
@@
 
+We agree, and plan to do this for future work.
+
 Q1. _Could the setting be extended into a source-to-source transformation?_
+
+Using the CHAD framework, we believe that this is very likely.
 
@@
 
+We plan to do this in future work. The quantitative approximations came quite late in the development and we plan to explore it further.
+
 Q3. _What is the computational overhead is of carrying slices alongside values?_
+
+Probably quite high.
+If the semantics is compiled "as is", the slice transformations will be represented as closures closing over the computation so far.
+A source-to-source transformation, or a more intelligent representation of slice transformations could potentially avoid this.
 
@@
 
+We plan to do a detailed comparison in future work.
+
 Q5. _Does the work on partiality and CHAD help in your setting?_
 
+Most likely, yes.
+
 # List of Proposed Changes

### 2025-09-11 reviews/popl2026/response.md
More response

@@
 The idea is that the indexing category represents the original functions, and the indexed category represents the tangent maps.
-Two interpretations are considered, one which is "syntactic" and so provides a source-to-source translation, and one which is "semantic" and interprets a term as a formal pairing of its functional meaning and its family of tangent maps.
-In our submission, we only do the semantic one but replace tangents with approximations and tangent maps with Galois connections.
+In the CHAD work, two interpretations are considered, one which is "syntactic" and so provides a source-to-source translation, and one which is "semantic" and interprets a term as a formal pairing of its functional meaning and its family of tangent maps.
+The former is interpreted in the latter, and the latter is proved correct with respect to actual derivatives via a special purpose sconing argument similar to our Theorem 5.2.
+In our submission, go directly to the semantic interpretation but replace tangents with approximations and tangent maps with Galois connections.
 Because we formulated our results in Agda, the semantic variant is actually executable and was used for our examples, so we didn't prioritise a source-to-source version.
+Moreover, we replace the specialised sconing / logical relations argument with a use of the Fiore and Simpson result, which is admittedly less elementary but highlights what is actually required for the correctness proof more clearly.
+Note that our approach also applies when $\mathrm{Fam}(\mathbf{LatGal})$ is replaced by $\mathbf{Man}$ and $\mathrm{Fam}(\mathbf{MeetSLat} \times \mathbf{JoinSLat}^{op})$ is replaced by $\mathrm{Fam}(\mathbf{Vect})$, recovering the original CHAD result with less work.
 
@@
 
-We admit that we should have done a more detailed survey of the literature on differential linear logic and how it relates to our setting.
-The connection between the trace and Taylor expansion may have something to say in our setting.
+We admit that we should have done a more detailed survey of the literature on Differential Linear Logic and how it relates to our setting, especially given the common factor of stability.
 A difference with DiLL is in the treatment of partiality and non-determinism.
 The Ehrhard survey referenced by the reviewer highlights these as key parts of DiLL, but, as we explained at the start of Section 3, we do not wish to admit either of these in our model for all functions -- they only apply to tangent maps.
-This was our motivation for moving from stable functions to the families construction.
+This was part of our motivation for moving from stable functions to the families construction.
 
@@
 
-We plan to do this in future work. The quantitative approximations came quite late in the development and we plan to explore it further.
+The quantitative approximations came quite late in the development and we plan to explore it further.
 
@@
 Probably quite high.
-If the semantics is compiled "as is", the slice transformations will be represented as closures closing over the computation so far.
+If the semantics is compiled as a functional program, the slice transformations will be represented as closures closing over the computation so far.
 A source-to-source transformation, or a more intelligent representation of slice transformations could potentially avoid this.
+For example, if one was only interested in forward slicing with the approximation object in Section 3.3.3, one could represent this by a "dual numbers" approach similar to normal forward AD, where the approximation object is interpreted simply as a boolean (formally a pair of a unit (the point) and a boolean (the tangent), and the semantics then tracks this through the input.
 

### 2025-09-12 reviews/popl2026/response.md
more response
add Taylor reference to response

@@
 The relationship of being a Galois connection replaces the fact that the forward and reverse maps are linear transposes of each other.
+
+- Unlike in normal differential geometry, we have no direct analogue of the real numbers.
+The approximation object in Section 3.3.3 is the closest, but does not account for the the interval approximation example in Section 3.3.4.
 
@@
 
-FIXME
+We don't have Taylor's book at hand to find the reference at the moment, but the results are part of Taylor's unpublished work on stable domain theory and its categorical generalisations at this url: https://www.paultaylor.eu/stable/ .
+Specifically, Theorem 1.11 in the manuscript "Quantitative Domains, Groupoids and Linear Logic" ( https://www.paultaylor.eu/stable/quadgl.pdf ).
 

### 2025-09-12 reviews/popl2026/response.md
tweak response

@@
 We address specific comments (C) and questions (Q) asked by each reviewer below, along with some proposed improvements to the paper.
+
+# List of Proposed Changes
+
+1.
+We will add more examples demonstrating how the Galois slicing works on larger programs.
+It is straightforward to add a datatype of arithmetic expressions to the language, where backwards slices of an interpreter for this language yield normal program slicing.
+
+2. We will add more examples demonstrating the interval approximations and their relation to interval analysis.
+
+3.
+We will add material further explaining the connection between our work and the CHAD presentation of Automatic Differentiation.
+See below for a short summary of the connection between CHAD, AD, and our work.
+
+4. We will add a more in-depth discussion of DiLL and how it relates to our work.
 
@@
 
-This is true, and is something we plan to address in future work. Nevertheless, total programs do cover a large range of useful applications, such as database queries.
+This is true, and is something we plan to address in future work.
+Nevertheless, total programs do cover a large range of useful applications such as database queries constructed modularly with higher-order functions.
 
@@
 The relationship of being a Galois connection replaces the fact that the forward and reverse maps are linear transposes of each other.
+
+- In relation to program slicing, the approximations ("tangents") in our approach are the slices of data values, and the Galois connections associated to a given execution of the program are the operations that push slices back and forth.
+The property that the back and forth operations form a Galois connection is the correctness property given by Perera et al.'s prior work.
 
@@
 
-We do not know if existing AD tools are appropriate as a basis for implementation, because they are specialised to real number differentiation, but we certainly plan to implement our ideas in a more practical tool to evaluate its performance.
+We do not know if existing AD tools are appropriate as a basis for implementation, because they are often specialised to real number differentiation, but we certainly plan to implement our ideas in a more practical tool to evaluate its performance.
 It may be possible to give a generic CHAD implementation that covers both.
@@
 Is it correct that we appropriate "differentiation" and other words from real analysis in our setting?
-This is debatable, and certainly one could argue that absent any kind of definition of derivative as a limit that one should not use the word.
+This is debatable, and certainly one could argue that absent any kind of definition of derivative as a limit, or use of linearity, that one should not use the word.
 Our argument is based on the similarity of structures (e.g., existence of tangent maps, meet preservation instead of linearity) outlined in Section 1.2, and on the conjectures about Tangent categories.
-The conjectures about Tangent categories (which in fact we have now proved for the forward case, the reverse case's definition is much more involved) extend the connection to the higher-order derivative case, but do not really say much more than that.
+The conjectures about Tangent categories (which in fact we have now proved for the forward case, the reverse case's definition is much more involved) extend the connection to the higher-order derivative case, but otherwise do not really say much more than that.
 
@@
 Whether or not this is a kind of automatic differentiation.
-Our argument here is based on the fact that we are proposing an interpretation of programs that computes forward and backwards maps of additional data at a point alongside the main computation.
-By the CHAD approach, this is exactly what first-order AD does, and so we think that the identification as a generalised form is valid.
+Our argument here is based on the fact that we are proposing an interpretation of programs that computes forward and backwards maps of associated data along the main computation.
+By the CHAD approach, this is exactly what first-order AD does, and so we think that our identification as a generalised form is valid.
 
@@
 A difference with DiLL is in the treatment of partiality and non-determinism.
-The Ehrhard survey referenced by the reviewer highlights these as key parts of DiLL, but, as we explained at the start of Section 3, we do not wish to admit either of these in our model for all functions -- they only apply to tangent maps.
-This was part of our motivation for moving from stable functions to the families construction.
+The Ehrhard survey referenced by the reviewer highlights these as key parts of DiLL, but, we do not wish to admit either of these in our model for all functions -- they only apply to tangent maps.
+This, as well the fact that approximations for us are not "real data", was part of our motivation for moving from stable functions to the families construction.
 
@@
 
-We agree, and plan to do this for future work.
+We agree, and plan to do this for any future version.
 
@@
 
-Using the CHAD framework, we believe that this is very likely.
+Using the CHAD framework, we believe that this is very likely. The Vákár and Smeding paper already demonstrates this for real-number AD.
 
@@
 
-The quantitative approximations came quite late in the development and we plan to explore it further.
+The quantitative approximations came quite late in the development and we plan to explore the applications of it further.
 
@@
 
-Probably quite high.
-If the semantics is compiled as a functional program, the slice transformations will be represented as closures closing over the computation so far.
+Probably quite high. If the semantics is compiled as a functional program, the slice transformations will be represented as closures closing over the computation so far.
+
 A source-to-source transformation, or a more intelligent representation of slice transformations could potentially avoid this.
+
 For example, if one was only interested in forward slicing with the approximation object in Section 3.3.3, one could represent this by a "dual numbers" approach similar to normal forward AD, where the approximation object is interpreted simply as a boolean (formally a pair of a unit (the point) and a boolean (the tangent), and the semantics then tracks this through the input.
@@
 Most likely, yes.
-
-# List of Proposed Changes

### 2025-09-12 reviews/popl2026/response.md
Minor corrections/tweaks.

@@
-We thank our three reviewers for their detailed reviews and suggestions for improving the paper.
-We address specific comments (C) and questions (Q) asked by each reviewer below, along with some proposed improvements to the paper.
+We thank our three reviewers for their detailed reviews and suggestions for improving the paper. We address specific comments (C) and questions (Q) asked by each reviewer below.
 
@@
 The former is interpreted in the latter, and the latter is proved correct with respect to actual derivatives via a special purpose sconing argument similar to our Theorem 5.2.
-In our submission, go directly to the semantic interpretation but replace tangents with approximations and tangent maps with Galois connections.
+In our submission, we go directly to the semantic interpretation but replace tangents with approximations and tangent maps with Galois connections.
 Because we formulated our results in Agda, the semantic variant is actually executable and was used for our examples, so we didn't prioritise a source-to-source version.

### 2025-09-12 reviews/popl2026/response.md
Minor tweak.
Minor tweaks.

@@
 Our argument here is based on the fact that we are proposing an interpretation of programs that computes forward and backwards maps of associated data along the main computation.
-By the CHAD approach, this is exactly what first-order AD does, and so we think that our identification as a generalised form is valid.
+Under the CHAD approach, this is exactly what first-order AD does, and so we think that the characterisation of our approach as a generalised form of AD valid.
 
@@
 
-We plan to do a detailed comparison in future work.
+We briefly mentioned some qualitative differences with existing slicing/provenance techniques in Section 6. We plan to do a detailed comparison in future work.
 

### 2025-09-12 reviews/popl2026/response.md
final tweaks

@@
 
-This is true, and is something that we plan to address in future work.
+This is true, we will add more examples comparing to existing Galois slicing implementations, and comparison with other tools is something that we plan to address more deeply in future work.
 
@@
 
-We can certainly make the connections clearer in a future version, but for now:
+We will make the connections clearer in a future version, but for now:
 
@@
 Moreover, we replace the specialised sconing / logical relations argument with a use of the Fiore and Simpson result, which is admittedly less elementary but highlights what is actually required for the correctness proof more clearly.
-Note that our approach also applies when $\mathrm{Fam}(\mathbf{LatGal})$ is replaced by $\mathbf{Man}$ and $\mathrm{Fam}(\mathbf{MeetSLat} \times \mathbf{JoinSLat}^{op})$ is replaced by $\mathrm{Fam}(\mathbf{Vect})$, recovering the original CHAD result with less work.
+Note that our approach also applies when $\mathrm{Fam}(\mathbf{LatGal})$ is replaced by $\mathbf{Man}$ and $\mathrm{Fam}(\mathbf{MeetSLat} \times \mathbf{JoinSLat}^{op})$ is replaced by $\mathrm{Fam}(\mathbf{Vect})$, recovering the original CHAD result with less work and at a higher level of abstraction.
 
@@
 
-We do not know if existing AD tools are appropriate as a basis for implementation, because they are often specialised to real number differentiation, but we certainly plan to implement our ideas in a more practical tool to evaluate its performance.
+We do not know if existing AD tools are appropriate as a basis for implementation, because they are often specialised to real number differentiation, but we certainly plan to implement our ideas in a more practical tool than Agda to evaluate its performance.
 It may be possible to give a generic CHAD implementation that covers both.
@@
 
-We admit that we should have done a more detailed survey of the literature on Differential Linear Logic and how it relates to our setting, especially given the common factor of stability.
+We admit that we should have done a more detailed survey of the literature on Differential Linear Logic and how it relates to our setting, especially given the common factor of Stability.
 A difference with DiLL is in the treatment of partiality and non-determinism.
@@
 
-Most likely, yes.
+Most likely, yes, for languages with general recursion.

### 2025-11-12 build-arXiv.sh
Initial version.

@@ -1,0 +1,11 @@
+#!/bin/bash
+set -xe
+
+NAME=arXiv.zip
+ARCHIVE="./$NAME"
+
+make
+
+zip -r - . > $ARCHIVE
+zip -d $ARCHIVE \*.{DS_Store,gitignore}
+zip -d $ARCHIVE *.pdf .git *.sh

### 2025-11-12 build-arXiv.sh
git archive an easier starting point./

@@ -1,11 +1,9 @@
 #!/bin/bash
 set -xe
 
-NAME=arXiv.zip
-ARCHIVE="./$NAME"
+NAME=arXiv
 
 make
 
-zip -r - . > $ARCHIVE
-zip -d $ARCHIVE \*.{DS_Store,gitignore}
-zip -d $ARCHIVE *.pdf .git *.sh
+git archive --format=zip HEAD -o "$NAME.zip"
+zip -d "$NAME.zip" *.pdf .git *.sh agda

### 2025-11-12 build-arXiv.sh
export-ignore rules for .

@@ -6,4 +6,3 @@ NAME=arXiv
 make
 
 git archive --format=zip HEAD -o "$NAME.zip"
-zip -d "$NAME.zip" *.pdf .git *.sh agda
